
\documentclass[11pt,a4paper]{article}
\input{preamble}

\title{Chapter 3: Positive but Not Completely \\ {\small (Wolf Lecture Notes -- Transcription)}}
\date{}

\numberwithin{equation}{section}

\begin{document}
\maketitle
\tableofcontents

\setcounter{section}{2}

\section{Positive, but not completely}

In this chapter we will discuss linear maps which are positive (i.e., they preserve
positive semidefiniteness) but not necessarily completely positive. Although
such operations cannot directly be implemented in a physical device they play
an important role as mathematical tools, in particular in entanglement theory
(see Sec.~3.2).

Recall that a linear map $T:\mcM_d(\CC)\to \mcM_{d'}(\CC)$ is positive if for every
positive semidefinite matrix $A\in \mcM_d(\CC)$ (we write $A\ge 0$) we have $T(A)\ge 0$.
$T$ is positive iff the adjoint map $T^*$ is so, and every positive map is Hermitian
in the sense that $T(X^\dagger)=T(X)^\dagger$ for all $X\in \mcM_d(\CC)$.
By Prop.~2.2 every positive map is a difference of two completely positive maps.
Since this is, however, true for every Hermitian map it does, unfortunately, not provide a
simple parametrization of positive maps.

The paradigm of a map, which is positive but not completely positive, is
the usual matrix transposition $A\mapsto \theta(A):=A^T$ which is defined with respect
to a given orthogonal basis by $\langle i|A^T|j\rangle=\langle j|A|i\rangle$.
The transposition is trace preserving and positive: if $UDU^\dagger$ is the spectral
decomposition of any Hermitian matrix (i.e., $U$ is unitary and $D$ diagonal), then
$(UDU^\dagger)^T=\overline{U}DU^T$ has the same eigenvalues. The fact that transposition is not
completely positive can be easily seen by applying it to half of a maximally entangled state
$\ket{\Omega}=\sum_{i=1}^d \ket{ii}/\sqrt{d}$ (see Example~1.2):
\begin{equation}\label{eq:ch3:3.1}
    (\theta\otimes \id)(\ketbra{\Omega}{\Omega})
    =\frac{1}{d}F,
    \qquad
    F=\sum_{i,j=1}^d \ket{ij}\!\bra{ji}.
\end{equation}

Since the ``flip operator'' $F$ has eigenvalues $\pm 1$ (with multiplicities $d(d\pm 1)/2$
corresponding to symmetric and antisymmetric eigenvectors respectively) $\theta$ fails to be
completely positive. For the partial transpose $(\theta\otimes \id)(X)$ we will occasionally
write $X^{T_A}$ or $X^{T_1}$ referring to a transposition on Alice's/the first tensor factor.

Transposition plays an important role in the discussion of the structure of positive maps and
isometries. We came across one of its appearances already in Wigner's theorem~1.1 and we will see
similar relations in Prop.~3.6 and \dots\ . Among all positive maps transposition gets physically
distinguished by a simple interpretation---as time reversal operation (see Sec.~3.3).

The simplest class of positive maps are those which are build up of a matrix transposition
together with completely positive maps $T_1$ and $T_2$. A positive map which has such a
decomposition in the form\footnote{Here $T_2\theta$ is a concatenation of maps, i.e.,
    $(T_2\theta)(X)=T_2(X^T)$.}
\begin{equation}\label{eq:ch3:3.2}
    T = T_1 + T_2\theta
\end{equation}
is thus called \emph{decomposable}. Note that the order of maps is irrelevant here, since
$\theta T_2\theta$ is completely positive iff $T_2$ is. From $\mcM_d$ to $\mcM_d$ with $d=2$ every
positive map is decomposable, whereas for larger dimensions this is no longer true. Both facts have
far-ranging consequences in entanglement theory (see Sec.~3.2). The existence of indecomposable
positive maps can, with some imagination and the willingness to make a little detour, be traced
back to the history of Hilbert's 17th problem. Since this is a detour of the nicer kind, we will
follow it in Sec.~3.4. We will start, however, by discussing notions of positivity in between
bare positivity and complete positivity:

\subsection{$n$-positivity}

A linear map $T:\mcM_d(\CC)\to \mcM_{d'}(\CC)$ is called $n$-positive (with $n\in\NN$) if $T\otimes \id_n$
is positive. That is, positivity of $T$ is preserved if we add an ``innocent bystander''
of dimension at most $n$. For given dimensions, the set $\mathcal{T}_n$ of $n$-positive maps
forms a closed convex cone, i.e., if $T_i\in \mathcal{T}_n$ then $\sum_i p_i T_i\in \mathcal{T}_n$
for all positive $p_i\in\RR$ and $\mathcal{T}_n$ contains its limit points. Moreover, the dual map
$T^*$ is $n$-positive if and only if $T$ is. Clearly, $n$-positivity implies $m$-positivity for all
$m\le n$. Only if $n=d$, then (by Prop.~1.2) $m$-positivity holds for all $m\in\NN$, i.e., the map is
completely positive. For different $n$ we thus get a chain of inclusions
\begin{equation}\label{eq:ch3:3.3}
    \mathcal{T}_{\mathrm{cp}} := \mathcal{T}_d \subseteq \mathcal{T}_{d-1}\subseteq \cdots \subseteq \mathcal{T}_2 \subseteq \mathcal{T}_1,
\end{equation}
from completely positive maps ($n=d$) to positive maps ($n=1$). As we will see later all these inclusions
are strict. To this end the following characterization is useful:

\begin{proposition}[Choi--Jamiolkowski for $n$-positive maps]
    Let $T:\mcM_d(\CC)\to \mcM_{d'}(\CC)$ be a linear map and
    $\tau:=(T\otimes \id)(\ketbra{\Omega}{\Omega})\in \mcM_{d'}\otimes \mcM_d$ its corresponding
    Choi--Jamiolkowski operator. For $n\le d$ the following are equivalent:
    \begin{enumerate}
        \item $T$ is $n$-positive.
        \item $(\one\otimes P)\tau(\one\otimes P)\ge 0$ for all Hermitian projections $P\in \mcM_d(\CC)$ with $\rank(P)=n$.
        \item $\bra{\psi}\tau\ket{\psi}\ge 0$ for all $\psi\in \CC^{d'}\otimes \CC^d$ with Schmidt-rank $n$.
    \end{enumerate}
\end{proposition}

\begin{proof}
    $T$ is $n$-positive iff $(T\otimes \id_n)(\ketbra{\phi}{\phi})\ge 0$ for all $\phi\in \CC^d\otimes \CC^n$.
    Embedding $\CC^n$ in $\CC^d$ this becomes equivalent to $(T\otimes \id_d)(\ketbra{\phi}{\phi})\ge 0$ for all
    $\phi\in \CC^d\otimes \CC^d$ with Schmidt rank $n$. The latter can be parameterized by
    $\ket{\phi}=(\one\otimes X)\ket{\Omega}$ where $X\in \mcM_d$ has rank $n$. In this way $n$-positivity becomes
    equivalent to
    \begin{equation}\label{eq:ch3:3.4}
        (\one\otimes X)\tau(\one\otimes X)^\dagger \ge 0,
        \qquad \forall X\in \mcM_d(\CC): \rank(X)=n.
    \end{equation}
    From here equivalence with 2.\ follows by expressing $X=\tilde X P$ in terms of the
    projector $P$ onto its support space and an invertible $\tilde X$.
    Equivalence between 3.\ and Eq.~\eqref{eq:ch3:3.4} is seen by taking the expectation value of the l.h.s.\ of
    \eqref{eq:ch3:3.4} with an arbitrary vector $\ket{\tilde \psi}$ and noting that
    $(\one\otimes X)^\dagger\ket{\tilde \psi}=:\ket{\psi}$ has Schmidt rank $n$.
\end{proof}

In particular the last item motivates to investigate extremal overlaps with pure states of a given Schmidt rank:

\begin{lemma}[Overlap with fixed Schmidt rank]
    Let $\phi\in \CC^{d'}\otimes \CC^d$ be a normalized vector with reduced density matrix
    $\tr_2 \ketbra{\phi}{\phi} =:\rho$. Then the maximal overlap with all normalized vectors $\psi$ of Schmidt rank
    $n$ is given by
    \begin{equation}\label{eq:ch3:3.5}
        \sup_{\psi} |\braket{\phi}{\psi}|^2 = \|\rho\|_{(n)},
    \end{equation}
    where $\|\cdot\|_{(n)}$ denotes the Ky-Fan norm, i.e., the sum over the $n$ largest singular values (here,
    eigenvalues).
\end{lemma}

\begin{proof}
    Using the Schmidt decomposition of both vectors we can rewrite the optimization in the form
    \begin{equation}\label{eq:ch3:3.6}
        \sup_{\psi} |\braket{\phi}{\psi}|^2
        = \sup_{U,V,\mu} \left|\tr\!\left[U\sqrt{\mu}\,V\sqrt{\lambda}\right]\right|^2,
    \end{equation}
    where $U,V$ are unitaries, $\lambda$ a diagonal matrix containing the eigenvalues of $\rho$ and $\sqrt{\mu}$ a
    diagonal matrix containing the Schmidt coefficients of $\psi$. The maximum in Eq.~\eqref{eq:ch3:3.6} is
    attained for $U=V=\one$ (which can for instance be seen by exploiting a standard singular value
    inequality\footnote{For two matrices $A,B$ the ordered singular values satisfy $\sum_i s_i(AB)\le \sum_i s_i(A)s_i(B)$;
    see Thm.~IV.2.5 in \dots\ [Bhatia].}). The remaining supremum is then of the
    form $\sup_{\mu}\bigl|\sum_{i=1}^n \sqrt{\mu_i\lambda_i}\bigr|^2$ with
    $\{\lambda_1,\dots,\lambda_n\}$ the $n$ largest eigenvalues of $\rho$. This
    can easily be solved using a Lagrange multiplier for the normalization constraint
    $\sum_{i=1}^n \mu_i=1$ which gives $\mu_i=c\lambda_i$ with $c$ determined by normalization.
    Putting things together we arrive at \eqref{eq:ch3:3.5}.
\end{proof}

Utilizing this Lemma together with item~3.\ in Prop.~3.1 we can now provide
a simple criterion for $n$-positivity in terms of the spectral decomposition
$\tau=\sum_i \nu_i \ketbra{\phi_i}{\phi_i}$ (with $\|\phi_i\|=1$). Denoting by
$\rho_i$ the reduced density operator of $\phi_i$ we obtain:

\begin{proposition}[Criterion for $n$-positivity]
    Consider a Hermitian operator $\tau\in \mcM_{d'}\otimes \mcM_d$ (e.g., the Choi--Jamiolkowski operator of a Hermitian map)
    whose smallest positive non-zero eigenvalue is $\nu_0$. Then with the above notation
    \begin{equation}\label{eq:ch3:3.7}
        \inf_{\psi}\bra{\psi}\tau\ket{\psi} \ge \nu_0 + \sum_{i:\nu_i\le 0} (\nu_i-\nu_0)\|\rho_i\|_{(n)},
    \end{equation}
    if the infimum is taken over all normalized vectors of Schmidt rank $n$. Conversely, if all but one eigenvalue
    $\nu_-$ of $\tau$ are strictly positive, and $\nu_\infty$ is the largest positive eigenvalue, then:
    \begin{equation}\label{eq:ch3:3.8}
        \inf_{\psi}\bra{\psi}\tau\ket{\psi} \le \nu_\infty + (\nu_- - \nu_\infty)\|\rho_-\|_{(n)}.
    \end{equation}
\end{proposition}

\begin{proof}
    By separating positive and negative parts of $\tau$ we obtain
    \begin{align}
        \tau
         &\ge \nu_0 \sum_{i:\nu_i>0} \ketbra{\phi_i}{\phi_i}
        + \sum_{j:\nu_j\le 0} \nu_j \ketbra{\phi_j}{\phi_j},
        \label{eq:ch3:3.9}
        \\
         &= \nu_0 \one + \sum_{i:\nu_i\le 0} (\nu_i-\nu_0)\ketbra{\phi_i}{\phi_i},
        \label{eq:ch3:3.10}
    \end{align}
    from which Eq.~\eqref{eq:ch3:3.7} follows by exploiting Lemma~3.1. The inequality \eqref{eq:ch3:3.9} is
    reversed if we replace $\nu_0$ by $\nu_\infty$, which then leads to Eq.~\eqref{eq:ch3:3.8}.
\end{proof}

Now, let us apply this to the map $T_\eta:\mcM_d\to \mcM_d$ defined by
\begin{equation}\label{eq:ch3:3.11}
    T_\eta(\rho) = \one\,\tr[\rho] - \frac{\rho}{\eta},
    \qquad \eta\in \RR_+.
\end{equation}
For $\eta\ge d$ this map is completely positive. For $\eta<d$ all positive eigenvalues of the corresponding
Choi--Jamiolkowski operator are equal to $1/d$ and there is a single negative eigenvalue
$\nu_-=(1/d-1/\eta)$ with $\|\rho_-\|_{(n)}=n/d$. Hence, the upper and lower bound in Prop.~3.2 actually coincide
and we get that $T_\eta$ is $n$-positive iff $\eta\ge n$.\footnote{For the example $T_\eta$ this can be proven in a
    simpler way (by averaging a Schmidt rank $n$ vector w.r.t.\ the group $\{U\otimes \overline{U}\}$), without invoking
    Prop.~3.2. We nevertheless stated the proposition as it can be applied to situations where $\tau$ does not have such
    a symmetry.} This proves that all the inclusions in the chain \eqref{eq:ch3:3.3} are strict. When $T_\eta$ is
positive (i.e., for $\eta\ge 1$) it is decomposable as $T_\eta\theta$ is completely positive.

\subsection{Positive maps and entanglement theory}

A map, which is positive but not completely positive, like the transposition, does not correspond to a physically
implementable operation. Nevertheless, such maps, and in particular the transposition and the map $T_1$ in
Eq.~\eqref{eq:ch3:3.11}, have become important mathematical tools in the theory of entanglement. Applied as
$(T\otimes \id)$ to a bipartite density matrix they are powerful tools for ``detecting'' entanglement as well as for
recognizing ``useful'' (in the sense of distillable) entanglement.

\paragraph{Detecting entanglement.}
Suppose that $\rho\in \mcM_d\otimes \mcM_{d'}$ is a separable state as defined in Eq.~(1.8), then positivity of $T$
implies that $(T\otimes \id)(\rho)$ remains positive semi-definite, since tensor products and convex combinations of
positive operators are again positive. Hence, for any positive map $T$ a negative eigenvalue of $(T\otimes \id)(\rho)$
proves that $\rho$ is entangled:
\begin{equation}\label{eq:ch3:3.12}
    T\ \text{positive, and}\ \bra{\Psi}(T\otimes \id)(\rho)\ket{\Psi}<0 \ \Longrightarrow\ \rho\ \text{is entangled.}
\end{equation}

We may go one step further and use an $n$-positive $T$. This gives the generalization
\[
    T\ \text{$n$-positive, and}\ \bra{\Psi}(T\otimes \id)(\rho)\ket{\Psi}<0
    \ \Longrightarrow\
    \begin{array}{l}
        \text{Every pure state decomposition}         \\
        \text{of $\rho$ contains at least one vector} \\
        \text{of Schmidt rank larger than $n$.}
    \end{array}
\]
The smallest $n$ such that $\rho$ has a convex decomposition into pure states of Schmidt rank at most $n$ is called
the \emph{Schmidt number} of $\rho$. With this terminology separable states are precisely those with Schmidt number
one.

In order to prove the converse of the above implications it is useful to introduce the concept of witnesses for
(Schmidt number $n+1$) entanglement.

\begin{proposition}[Entanglement witnesses]
    A density operator $\rho\in \mcM_d(\CC)\otimes \mcM_{d'}(\CC)$ has Schmidt number at least $n+1$ iff there exists a
    Hermitian operator $W\in \mcM_d(\CC)\otimes \mcM_{d'}(\CC)$ such that $\tr[W\rho]<0$ but
    $\bra{\psi}W\ket{\psi}\ge 0$ for all vectors $\psi\in \CC^d\otimes \CC^{d'}$ with Schmidt rank $n$.
\end{proposition}

\begin{proof}
    Egg--hyperplane--done. In order to be a bit more precise, let us regard the space of Hermitian operators as a real
    vector space equipped with the Hilbert--Schmidt inner product. Since the set $S_n$ of density operators with Schmidt
    number at most $n$ is compact and convex, every $\rho\notin S_n$ can be separated from it by a hyperplane of Hermitian
    operators $H$ satisfying $\langle H,\widetilde W\rangle=c$ for some constant $c$ and some $\widetilde W=\widetilde
        W^\dagger$. That is, $\langle \rho,\widetilde W\rangle<c$ whereas $\langle \rho_0,\widetilde W\rangle\ge c$ for all
    $\rho_0\in S_n$. Setting $W:=\widetilde W-c\one$ then completes the `only if' part of the proof. The `if' part follows
    from the fact that $S_n$ is the closure of the convex hull of the set of pure states with Schmidt rank $n$.
\end{proof}

An operator $W$ with the properties mentioned in Prop.~3.3 is called \emph{entanglement witness} (for Schmidt number
$n+1$ entanglement). We will follow the literature and use `entanglement witness' referring to the case $n=1$ unless
otherwise specified. Invoking the Choi--Jamiolkowski correspondence (Prop.~2.1) we can assign a positive map to each
witness. This leads to the following characterization:

\begin{proposition}[Positive maps and entanglement]
    A density operator $\rho\in \mcM_d(\CC)\otimes \mcM_{d'}(\CC)$ has Schmidt number at most $n$ iff for all $n$-positive
    maps $T:\mcM_d(\CC)\to \mcM_{d'}(\CC)$ it holds that $(T\otimes \id)(\rho)\ge 0$.
\end{proposition}

\begin{proof}
    We only have to prove the `if' part. For that assume that $\rho\notin S_n$. Then by Prop.~3.3 there is an entanglement
    witness for which $\tr[W\rho]<0$. This defines, by the Choi--Jamiolkowski correspondence, a linear map
    $T^*:\mcM_{d'}(\CC)\to \mcM_d(\CC)$ via
    \begin{equation}\label{eq:ch3:3.13}
        W := (T^*\otimes \id)(\ketbra{\Omega}{\Omega}).
    \end{equation}
    By item~3 in Prop.~3.1 $T^*$ and with it its adjoint $T$ is $n$-positive and by construction
    \begin{equation}\label{eq:ch3:3.14}
        \tr[W\rho] = \bra{\Omega}(T\otimes \id)(\rho)\ket{\Omega} < 0.
    \end{equation}
    Hence, $(T\otimes \id)(\rho)$ cannot be positive.
\end{proof}

We see from Eq.~\eqref{eq:ch3:3.14} that the positive map constructed from a witness is actually more powerful,
in the sense that it detects more entangled states, than the witness we started with: after all $(T\otimes \id)(\rho)$
might have a negative eigenvalue without $\bra{\Omega}(T\otimes \id)(\rho)\ket{\Omega}$ being negative.
A closer look reveals that a positive map $T$ constructed from a witness $W$ `detects' an entangled state $\rho$
iff it is detected by one of the witnesses from the orbit $\{(\one\otimes X)W(\one\otimes X)^\dagger\}$
with $X\in \mcM_{d'}(\CC)$.

Decomposable positive maps correspond to decomposable witnesses which are of the form
\begin{equation}\label{eq:ch3:3.15}
    W = P_1 + P_2^{T_1},
    \qquad P_i\ge 0.
\end{equation}
By duality, the existence of indecomposable witnesses for given dimensions is equivalent to the existence of
entangled states satisfying $\rho^{T_1}\ge 0$:

\begin{proposition}[PPT entangled states]\label{prop:ch3-ppt-entangled}
    \footnote{Admittedly this proposition has mainly and at most pedagogical value. Once examples of entangled PPT states
        are constructed it looses a lot of its charm\ldots}
    There exists an indecomposable positive map $T:\mcM_d(\CC)\to \mcM_{d'}(\CC)$ iff there is an entangled state
    $\rho\in \mcM_d\otimes \mcM_{d'}$ with positive partial transpose $\rho^{T_1}\ge 0$.
\end{proposition}

\begin{proof}
    We exploit the equivalence between indecomposable maps and witnesses. That is, we show that the existence of an
    indecomposable positive map implies that there is an entangled state with $\rho^{T_1}\ge 0$. The converse follows from
    Prop.~3.3.

    Denote by $\mathcal W\subseteq \mcM_d\otimes \mcM_{d'}$ the set of all entanglement witnesses, i.e., Hermitian
    operators with $\bra{\phi\otimes \psi}W\ket{\phi\otimes \psi}\ge 0$ for all $\phi,\psi$, and by
    $\mathcal W_T\subseteq \mathcal W$ all decomposable witnesses. Both $\mathcal W$ and $\mathcal W_T$ are closed convex
    sets. If there is (by assumption) a $W\in \mathcal W\setminus \mathcal W_T$ we can separate it by a hyperplane
    characterized by a Hermitian operator $R$ such that $\tr[WR]<0\le \tr[W_T R]$ for all $W_T\in \mathcal W_T$. Since
    $\mathcal W_T$ contains all positive operators $R$ must be positive so that $\rho:=R/\tr[R]$ is a density matrix. By
    construction $\rho^{T_1}\ge 0$ while $\rho$ is entangled due to $\tr[W\rho]<0$.
\end{proof}

We will see shortly that entangled PPT states exists iff the dimension is larger than $\CC^2\otimes \CC^3$.

\begin{example}[Some positive maps and entanglement witnesses]
    \par\medskip\noindent\textbf{Transposition.}
    Transposition is a positive map. It corresponds to the set of witnesses
    $W=(A\otimes B)F(A\otimes B)^\dagger$ for arbitrary $A,B$.
    The separability criterion $(\theta\otimes \id)(\rho):=\rho^{T_1}\ge 0$
    is called PPT (positive partial transpose) criterion. Although the partial transpose
    depends on the choice of the local basis, the PPT criterion does not: if we take the
    partial transposition with respect to a different (local) basis, labeled by the superscript
    $\widetilde T_1$, we get
    \begin{align}
        \rho^{\widetilde T_1}
         &=(U\otimes \one)\left[((U^\dagger\otimes \one)\rho(U\otimes \one))^{T_1}\right](U^\dagger\otimes \one),
        \label{eq:ch3:3.16}
        \\
         &=((UU^T)\otimes \one) \rho^{T_1}\,((UU^T)^\dagger\otimes \one).
        \label{eq:ch3:3.17}
    \end{align}
    Eqs.~\eqref{eq:ch3:3.16}--\eqref{eq:ch3:3.17} show that the eigenvalues of the partial transpose are basis independent,
    and so is the positivity of $\rho^{T_1}$. Moreover, $\rho^{T_1}\ge 0$ is equivalent to $\rho^{T_2}\ge 0$ since the two
    operators only differ by a global transposition. We will see in \dots\ that the PPT criterion is necessary and
    sufficient for separability in $\CC^2\otimes \CC^2$ and $\CC^2\otimes \CC^3$.

    \par\medskip\noindent\textbf{Reduction criterion.}
    The set of maps $T_n(X):=\tr[X]\one - X/n$, $n\in\NN$ in Eq.~\eqref{eq:ch3:3.11} is $n$-positive (and self-dual, i.e.,
    $T_n^*=T_n$). As a consequence
    \begin{equation}\label{eq:ch3:3.18}
        n\rho_1\otimes \one \ge \rho
        \qquad\text{and}\qquad
        n\one\otimes \rho_2 \ge \rho,
    \end{equation}
    (with $\rho_k$ the $k$th reduced density matrix of $\rho$) is necessary for a bipartite state $\rho$ to have Schmidt
    number $n$. For $n=1$ this is called reduction criterion. As $T_n$ is a decomposable positive map, the reduction
    criterion is generally weaker than the PPT criterion for entanglement detection. For $\CC^2\otimes \CC^2$ the two
    criteria are equivalent since $T_1(X)=(\sigma_y X\sigma_y)^T$ for any $X\in \mcM_2(\CC)$. An entanglement witness
    corresponding to $T_n$ is $W_n=\one/d-\ketbra{\Omega}{\Omega}/n$. That is, a state $\rho$ with Schmidt number $n$ has
    to satisfy $\bra{\Omega}\rho\ket{\Omega}\le n/d$ which is a weak version of Lemma~3.1.

    \par\medskip\noindent\textbf{Breuer--Hall map.}
    $T_{\mathrm{BH}}:\mcM_d\to \mcM_d$ is positive and defined as
    \begin{equation}\label{eq:ch3:3.19}
        T_{\mathrm{BH}}(X)=\tr[X]\one - X - U X^T U^\dagger,
        \quad\text{for any $U$ with $U^T=-U$ and $U^\dagger U\le \one$.}
    \end{equation}
    Positivity is easily seen by applying $T_{\mathrm{BH}}$ to a pure state and noting that due to the anti-symmetry of
    $U$ we have $\bra{\psi}U\ket{\overline{\psi}}=0$ which implies that the subtracted terms are orthogonal. For $d$ even
    there are anti-symmetric unitaries $U^\dagger=U^{-1}=-\overline{U}$ (e.g.\ embeddings of the Pauli matrix $\sigma_y$),
    whereas for $d$ odd these do not exist (which can be seen from the fact that $\det(U)=\det(U^T)=\det(-U)$ would be
    equal to $-\det(U)$ in this case; see Thms.~3.1,3.2). If $U$ is an anti-symmetric unitary $T_{\mathrm{BH}}$ is not
    decomposable.

    \par\medskip\noindent\textbf{Choi-type maps.}
    Let $D\in \mcB(\mcM_d)$ map a matrix $X$ onto a diagonal matrix $D(X)$ with the same diagonal entries, and let $U_{k0}$
    be a cyclic shift as defined in Eq.~(2.24). Then for all $n\in\NN$ with $1\le n\le d-2$ the map
    $T_C\in \mcB(\mcM_d)$ defined by
    \begin{equation}\label{eq:ch3:3.20}
        T_C(X):=(d-n)D(X)-X+\sum_{k=1}^n D\!\left(U_{k0}XU_{k0}^\dagger\right),
    \end{equation}
    is positive and indecomposable. For $n=0$ it is completely positive and for $n=d-1$ it is proportional to the map
    $T_n$ appearing in the context of the reduction criterion.

    \par\medskip\noindent\textbf{Unextendable product bases (UPBs).}
    Unextendable product bases (UPBs) are sets $S=\{\ket{\alpha_j}\otimes \ket{\beta_j}\}$ of normalized and mutually
    orthogonal product vectors in $\CC^d\otimes \CC^{d'}$ ($d'\ge d$) such that there is no product vector, which is
    orthogonal to every element of $S$, and $|S|<dd'$.\footnote{Note that this is only possible if $\{\ket{\alpha_j}\}$
        (resp.\ $\{\ket{\beta_j}\}$) is not a set of mutually orthogonal vectors.} That is, the projection
    $P_S:=\sum_{j=1}^{|S|}\ketbra{\alpha_j}{\alpha_j}\otimes \ketbra{\beta_j}{\beta_j}$ has a kernel which contains no
    product vector. An example of such a UPB in dimension $3\otimes 3$ can be constructed from five real vectors forming
    the apex of a regular pentagonal pyramid, where the height $h$ is chosen such that nonadjacent apex vectors are
    orthogonal. These vectors are
    \begin{equation}\label{eq:ch3:3.21}
        v_j = N\left(\cos\frac{2\pi j}{5}, \sin\frac{2\pi j}{5},\,h\right),
        \qquad j=0,\dots,4,
    \end{equation}
    with $N=2/\sqrt{5+\sqrt{5}}$ and $h=\tfrac12\sqrt{1+\sqrt{5}}$. The UPB is then given by
    $\ket{\alpha_j}\otimes \ket{\beta_j}=\ket{v_j}\otimes \ket{v_{2j \bmod 5}}$. Since any subset of three vectors on
    either side spans the full space, there cannot be a product vector orthogonal to all these states. Based on this or
    any other UPB, one can easily construct an entangled PPT state:
    \begin{equation}\label{eq:ch3:3.22}
        \rho_S \propto (\one - P_S).
    \end{equation}
    Since by construction $\rho_S$ has no product vector in its range it has to be entangled. Moreover, $\rho_S^{T_1}\ge
        0$, since $\{\ket{\overline{\alpha}_j}\otimes \ket{\beta_j}\}$ is again a set of mutually orthogonal and normalized
    vectors. From here we can construct an entanglement witness
    \begin{equation}\label{eq:ch3:3.23}
        W_S = P_S - \epsilon \ketbra{\psi}{\psi},
    \end{equation}
    where $\psi$ is any maximally entangled state of dimension $d$ fulfilling $\bra{\psi}\rho_S\ket{\psi}>0$ and
    \[
        \epsilon := d\inf_{\phi_1,\phi_2}
        \frac{\bra{\phi_1\otimes \phi_2}P_S\ket{\phi_1\otimes \phi_2}}{\|\phi_1\otimes \phi_2\|^2}.
    \]
    So by construction $\tr[W_S\rho_S]<0$ whereas $\tr[W_S\rho]\ge 0$ for all separable states. By the
    Choi--Jamiolkowski correspondence $W_S$ thus defines a positive map which is not decomposable.
\end{example}

\paragraph{Entanglement distillation} \dots\ \textsc{to be written} \dots

\noindent
LOCC \& separable superoperators

\noindent
PPT implies not distillable

\noindent
reduction criterion, distillability (maybe majorization)

\subsection{Transposition and time reversal}

Due to the lack of complete positivity, the transposition (as well as the partial transposition) does not correspond
to a physically implementable operation. However, it can physically be interpreted as ``time reversal'' [?] and
appears as such quite often as a (global) symmetry, in particular in particle physics [?]. Although we will not make
explicit use of this interpretation or symmetry in the following, we will briefly describe this relation.

One way to see that the transposition or complex conjugation (which is the same on Hermitian operators) admits an
interpretation as time reversal, is looking at the Wigner function of a continuous variable system in Schr\"odinger
representation.\footnote{Note that we leave the case $\dim \mathcal H < \infty$ within this section. The underlying
    Hilbert space is now the space $L^2$ of square integrable functions.} If we choose
$\rho=\sum_i \lambda_i \ket{\psi_i}\bra{\psi_i}$ with $\lambda_i\ge 0$ and $\sum_i \lambda_i\|\psi_i\|^2=1$
to be a decomposition of a density operator $\rho$ and define an integral kernel
$D(x;y)=\sum_i \lambda_i \psi_i(x)\overline{\psi_i(y)}$ in the Schr\"odinger representation, then we can write
the Wigner function as
\begin{equation}\label{eq:ch3:3.24}
    W(q,p)=\pi^{-n}\int D(q-x;\,q+x)\,e^{-i2x\cdot p}\,d^n x,
\end{equation}
where $n$ is the number of canonical degrees of freedom, i.e., $x,p,q\in \RR^n$. From Eq.~\eqref{eq:ch3:3.24} we see,
that the complex conjugation or transposition $D(x;y)\mapsto D(y;x)$ is equivalent to a reversal of momenta
$p\mapsto -p$. Due to the basis dependency of the transposition this is, however, no longer true if we choose another
representation. Hence we cannot identify the transposition or complex conjugation with time reversal in general.

Time is a parameter in quantum mechanics. Therefore there cannot be a general transformation in Hilbert space which
directly affects this parameter and acts as $t\mapsto -t$. In this sense the term ``time reversal'' is misleading.
It characterizes a reversal of momenta including angular momenta and spins, rather than a reversal of the arrow of time.
The defining ansatz for time reversal is thus an operator $T$ which acts on Hilbert space in a way that position and
momentum operators transform as
\begin{equation}\label{eq:ch3:3.25}
    TQT^{-1}=Q
    \qquad\text{and}\qquad
    TPT^{-1}=-P,
\end{equation}
and for consistency also $TLT^{-1}=-L$ and $TST^{-1}=-S$ for angular momenta and spins. Left and right multiplication
of the canonical commutation relations $[Q,P]=i\one$ leads with Eq.~\eqref{eq:ch3:3.25} to $TiT^{-1}=-i\one$. Thus $T$
is an ``anti-linear'' operator, and, as it is supposed to be norm preserving, it is ``anti-unitary''.\footnote{An
    anti-linear operator $A$ satisfies $Az=\overline{z}A$ for any complex number $z$. The product of two anti-linear
    operators is thus again a linear operator. If the inverse $A^{-1}$ exists and $\forall\Psi\in \mathcal H:\|A\Psi\|=\|\Psi\|$,
    then we say that $A$ is anti-unitary. We have then $\langle A\Phi|A\Psi\rangle=\langle \Psi|\Phi\rangle$.}

The simplest example of an anti-unitary operator is the complex conjugation $\Gamma$ (which is on Hermitian matrices
equal to the transposition). In fact, in Schr\"odinger representation we may identify the time reversal operator $T$
with complex conjugation, since the latter leaves the position operator unchanged and maps
$-i\frac{d}{dx}\stackrel{\Gamma}{\mapsto} i\frac{d}{dx}$ and therefore changes the sign of momentum and angular
momentum operators. As the time reversal operator should not depend on the representation (resp.\ basis), $T$ cannot be
identified with $\Gamma$ in general. However, every anti-unitary operator can be written as the product of a unitary
operator with complex conjugation.

Let us now turn to spin angular momenta. Consider the standard representation, in which $S_x$ and $S_z$ are real and
$S_y$ is imaginary. In this basis we have
\begin{equation}\label{eq:ch3:3.26}
    \Gamma S_x\Gamma^{-1}=S_x,
    \qquad
    \Gamma S_y\Gamma^{-1}=-S_y,
    \qquad
    \Gamma S_z\Gamma^{-1}=S_z,
\end{equation}
hence we cannot identify $T$ with complex conjugation. However, we may write $T=\Gamma V$, where $V$ has to invert the
signs of $S_x$ and $S_z$, while leaving $S_y$ unchanged. Moreover, $V$ is, as a product of two anti-unitaries, a linear
unitary operator, and it is supposed to act only on the spin degrees of freedom. These requirements are satisfied by
$V=\exp(-i\pi S_y)$, which rotates the spin through the angle $\pi$ about the $y$ axis. Hence, we have with respect to
the considered representation\footnote{Note that time reversal squared is not the identity. The operator $T^2$ in
    Eq.~\eqref{eq:ch3:3.27} has eigenvalues $-1$ for states with total spin $n/2$, with $n$ odd.}
\begin{equation}\label{eq:ch3:3.27}
    T=\Gamma e^{-i\pi S_y},
\end{equation}
and applying the time reversal to a density matrix describing a finite dimensional spin system gives then
$T^{-1}\rho T=V^\dagger \rho^T V$.

Let us now consider (finite dimensional) composite systems, where $\rho\in \mcB(\mathcal H_1\otimes \mathcal H_2)$.
The global time reversal operator is then $T\otimes T$, which is a well defined anti-unitary operator again.
Unfortunately, there exists no operator on Hilbert space, which corresponds to a ``partial time reversal'' of only one
of the two subsystems. In fact, the action of a tensor product of the form $\id\otimes T$ (linear$\otimes$anti-linear)
on a vector $\Psi\in \mathcal H_1\otimes \mathcal H_2$ is not well defined. However, since for product vectors
$(\id\otimes T^{-1})(\ket{\psi}\otimes \ket{\phi})$ gives only rise to a phase ambiguity, the action on the respective
projectors
\begin{equation}\label{eq:ch3:3.28}
    (\id\otimes T^{-1})(\ketbra{\psi}{\psi}\otimes \ketbra{\phi}{\phi})(\id\otimes T)
\end{equation}
becomes well defined without any ambiguity, and Eq.~\eqref{eq:ch3:3.28} describes a linear operator again. Since the
real linear hull of projectors onto product states equals the set of all Hermitian operators, we can in this way define
the ``partial time reversal'' as a map on density matrices. Due to Eq.~\eqref{eq:ch3:3.27} this is equivalent to the
partial transposition up to a local unitary transformation. Note that the latter is irrelevant for the detection of
entanglement since it preserves the eigenvalues.

The time reversal symmetry of a dynamics, i.e., the invariance of a Hamiltonian under time reversal, does not lead to a
conserved quantity (like parity for space inversion). However, it sometimes leads to a degeneracy of the eigenvalues:
Suppose $[H,T]=0$ for some anti-unitary $T$ and Hermitian $H$. Writing $T=\Gamma V$ with $V$ unitary this becomes
equivalent to $HV^\dagger = V^\dagger H^T$. Moreover, $T^2=-1$ is equivalent to $V^T=-V$ being anti-symmetric.

\begin{theorem}[Kramer's theorem]
    If $[H,T]=0$ and $T^2=-1$ for a Hermitian $H$ and anti-unitary $T$, then each eigenvalue of $H$ is at least two-fold
    degenerate.
\end{theorem}

\begin{proof}
    Let $H\ket{\psi}=\lambda\ket{\psi}$. Then $HV^\dagger\ket{\overline{\psi}}=V^\dagger H^T\ket{\overline{\psi}}
        =V^\dagger \overline{H}\ket{\overline{\psi}}=\lambda V^\dagger\ket{\overline{\psi}}$, so $V^\dagger\ket{\overline{\psi}}$
    gives rise to the same eigenvalue. This is degenerate since the two vectors are orthogonal:
    \begin{align}
        \bra{\psi}V^\dagger\ket{\overline{\psi}}
         &= -\bra{\psi}\overline{V}\ket{\overline{\psi}}
        = -\bra{\overline{\psi}}V\ket{\psi},
        \label{eq:ch3:3.29}
        \\
         &= -\bra{\psi}V^\dagger\ket{\overline{\psi}}.
        \label{eq:ch3:3.30}
    \end{align}
\end{proof}

Note that anti-symmetric unitaries only exist in even dimensions.\footnote{In odd dimensions we get that
    $\det(V)=\det(V^T)=-\det(V)$ which contradicts the fact that the determinant of a unitary has to be a phase.} In fact,
the restriction on the required symmetry in the theorem can be relaxed as the same proof shows:

\begin{theorem}[Kramer's theorem II]
    Let $H$ be Hermitian and $HA=AH^T$ for some anti-symmetric $A\neq 0$. Then each eigenvalue of $H$ is at least two-fold
    degenerate.
\end{theorem}

\subsection{From Hilbert's 17th problem to indecomposable maps}

The following pages will somehow be a detour following a historical path. Rather than writing down examples of
indecomposable positive maps (as we did within Example~3.1) we will see how they arose from a problem Hilbert thought
about in 1888. For an overview of Hilbert's 17th problem see [?].

\paragraph{Hilbert's problem:}
A sum of squares (sos) of real polynomials is obviously positive semi-definite. The converse question, whether a real
homogeneous psd polynomial of even degree has a sos decomposition, was posed and answered by David Hilbert in 1888.
Let $H_h(\RR^n)$ be the set of homogeneous polynomials of even degree $h$ in $n$ real variables, and
\begin{align}
    P_h(\RR^n)      &= \{p\in H_h(\RR^n)\mid \forall x\in \RR^n:\ p(x)\ge 0\},
    \label{eq:ch3:3.31}
    \\
    \Sigma_h(\RR^n) &=
    \left\{p\in P_h(\RR^n)\ \middle|\ \exists\{h_k\in H_{h/2}(\RR^n)\}:\ p=\sum_k h_k^2\right\},
    \label{eq:ch3:3.32}
\end{align}
the set of psd polynomials in $H$ and the subset of polynomials having an sos decomposition, respectively.\footnote{Note
    that $P$ and $\Sigma$ are both convex cones. Moreover, they are closed, i.e., if $p_n\to p$ coefficientwise, and each
    $p_n$ is in $P$ resp.\ $\Sigma$, then so is $p$.} It is not difficult to see that $P=\Sigma$ if $n=2$ or $h=2$. In
fact, the latter is nothing but the spectral decomposition of psd matrices. In 1888 Hilbert showed at first, that
every $p\in P_4(\RR^3)$ can be written as the sum of squares of three quadratic forms, i.e., $P_4(\RR^3)=\Sigma_4(\RR^3)$.
Moreover, he proved, albeit in a non-constructive way, that $(h,n)=(2,n),(h,2),(4,3)$ are in fact the only cases where
$P=\Sigma$. In 1900 Hilbert posed a generalization of these questions as his ``17th problem'' at the International
Congress of Mathematics in Paris\footnote{At this congress Hilbert outlined 23 major mathematical problems. Whereas
    some are broad, such as the axiomatization of physics (6th problem), others were very specific and could be solved
    quickly afterwards.}: Does every homogeneous psd form admit a decomposition into a sum of squares of rational
functions? This was answered yes by Artin in 1927 using the Artin--Schreier theory of real closed
fields [?].\footnote{Artin's proof holds for forms with coefficients from a field with unique order, i.e., in
    particular for $\RR$.}

\paragraph{Choi's example:}
Concerning Hilbert's original 1888 work it took nearly 80 years for explicit polynomials $p\in P\setminus \Sigma$ to
appear in the literature. One of these examples was provided by Choi, while investigating one of the numerous ``sum of
squares'' variants of Hilbert's problem. Choi showed in 1975 that there are psd biquadratic forms that cannot be
expressed as the sum of squares of bilinear forms. More precisely, let
\begin{equation}\label{eq:ch3:3.33}
    F(x,y)=\sum_{i,j,k,l=1}^n F_{ijkl}x_ix_jy_ky_l,
    \qquad \forall x,y:\ F(x,y)\ge 0
\end{equation}
be a real psd biquadratic form in $x,y\in \RR^n$. For $n=3$ Choi found a counterexample to the (wrongly
proven\footnote{For the case $x\in \RR^n$, $y\in \RR^m$ with $n=2$ or $m=2$ Calderon [?] proved that there is always a
    sos decomposition of the form in Eq.~(3.34). In [?] Koga claimed that this is also true for arbitrary $n,m$. Finding
    the flaw in Koga's proof, was apparently Choi's motivation for [?].}) conjecture that there is always a decomposition
into a sum of squares of bilinear forms, i.e.
\begin{equation}\label{eq:ch3:3.34}
    F(x,y)
    \stackrel{?}{=}
    \sum_{\alpha}(f^{(\alpha)}(x,y))^2
    =
    \sum_{\alpha}\left(\sum_{i,j=1}^n f_{ij}^{(\alpha)}x_i y_j\right)^2.
\end{equation}
Choi's counterexample, for which he gave very elementary proofs, is
\begin{equation}\label{eq:ch3:3.35}
    F(x,y)=2\sum_{i=1}^3 x_i^2(y_i^2+y_{i+1}^2) - \left(\sum_{j=1}^3 x_jy_j\right)^2,
\end{equation}
with $y_{3+1}\equiv y_1$. Note that $F\in P_4(\RR^6)$ and that every quadratic form appearing in an sos decomposition
of $F$ has to be bilinear. Hence, Eq.~\eqref{eq:ch3:3.35} is an element of $P\setminus \Sigma$ for $(h,n)=(4,6)$.
\footnote{By choosing dependent $x$ and $y$, Choi also specified $F(x,y)$, such that it yielded elements of
    $P\setminus \Sigma$ for $(h,n)=(4,4)$ and $(6,3)$ respectively.}

For the mere beauty of the example let us mention an element of $P\setminus \Sigma$ for $(h,n)=(4,5)$:
\begin{equation}\label{eq:ch3:3.36}
    \sum_{i=1}^5 \prod_{j\neq i}(x_i-x_j).
\end{equation}

\paragraph{Relation to positive maps:}
In our context the main significance of biquadratic forms lies in their relation to linear maps on symmetric
matrices. This relation is in fact the real analog of the correspondence between entanglement witnesses and positive
maps expressed in Eq.~\eqref{eq:ch3:3.13}. Let $S_n$ be the set of all real symmetric $n\times n$ matrices, and let
$\Phi:S_n\to S_n$ be a positive linear map on $S_n$, i.e.\ $\forall x\in\RR^n:\ \Phi(\ketbra{x}{x})\ge 0$. Then $\Phi$
corresponds to a psd biquadratic form $F$ and vice versa via\footnote{$F\ge 0$ implies $\Phi\ge 0$ since any symmetric
    matrix $S\in S_n$ is psd if $\bra{y}S\ket{y}\ge 0$ for every real vector $y\in \RR^n$.}
\begin{equation}\label{eq:ch3:3.37}
    F(x,y)=\bra{y}\Phi(\ketbra{x}{x})\ket{y}.
\end{equation}
Particular instances of such maps are congruence maps of the form $\Phi(A)=K^T AK$, which in turn correspond to psd
biquadratic forms $F(x,y)=f(x,y)^2$, where $f(x,y)=\sum_{ij}K_{ij}x_i y_j$ is a bilinear form. The existence of a psd
biquadratic form, which does not admit an sos decomposition, thus implies that the set of positive maps $\Phi:S_n\to S_n$
properly contains the convex hull of all congruence maps (for $n\ge 3$). The latter is, however, the real analog of
the set of (complex) decomposable positive maps. Let us now see, whether the map
\begin{equation}\label{eq:ch3:3.38}
    \Phi(S)=2\tr[S]\one - \operatorname{diag}(S_{33},S_{11},S_{22}) - S,
\end{equation}
which corresponds to the psd biquadratic form in Eq.~\eqref{eq:ch3:3.35} (see also Example~3.1), represents an
admissible complex non-decomposable positive map as well.

Let $\Lambda$ be a complex extension of $\Phi$ defined as in Eq.~\eqref{eq:ch3:3.38} but on arbitrary complex
$3\times 3$ matrices. Obviously, $\Lambda$ maps Hermitian matrices onto Hermitian matrices. Moreover, if $\Lambda$ is a
decomposable positive map, it acts on symmetric matrices as $\Lambda(S)=\sum_j K_j^*SK_j$ for some complex matrices
$\{K_j\}$. If we decompose the latter with respect to their real and imaginary parts $K_j=R_j+iI_j$ we get
\begin{equation}\label{eq:ch3:3.39}
    \Lambda(S)=\sum_j R_j^T SR_j + I_j^T SI_j \overset{!}{=} \Phi(S),
\end{equation}
contradicting the fact that $\Phi$ in Eq.~\eqref{eq:ch3:3.38} has no such decomposition into congruence maps. Hence,
$\Lambda$ is either not decomposable or not a positive map at all. The latter can, however, be excluded, since we have
for every vector $z\in \CC^3$ with $z_j=x_j\varphi_j$, $x_j=|z_j|$:
\begin{equation}\label{eq:ch3:3.40}
    \Lambda(\ketbra{z}{z})= D^*\,\Phi(\ketbra{x}{x})\,D \ge 0,
    \qquad D=\operatorname{diag}(\varphi_1,\varphi_2,\varphi_3).
\end{equation}
By Prop.~\ref{prop:ch3-ppt-entangled} Choi's map thus proves, albeit in a non-constructive way, the existence of PPT
entangled states in $3\otimes 3$.

The question whether an entanglement witness $W$ is decomposable or not, is equivalent to asking whether the
biquadratic Hermitian psd form
\begin{equation}\label{eq:ch3:3.41}
    W(\Phi,\Psi)=\bra{\Phi\otimes \Psi}W\ket{\Phi\otimes \Psi}
    =\sum_{ijkl} W_{ijkl} \overline{\Phi_i} \overline{\Psi_j} \Phi_k\,\Psi_l
\end{equation}
has an sos decomposition. It is therefore the complex analog of Eq.~\eqref{eq:ch3:3.34}. Apart from examples there is,
however, not much known about forms which do not admit such a decomposition.

Every real unextendable product basis (as the one in Eq.~\eqref{eq:ch3:3.21}) can be used to construct a real-valued
non-decomposable entanglement witness. These witnesses thus correspond to biquadratic psd forms, which do not admit an
sos decomposition and are therefore elements of $P\setminus \Sigma$.

\subsection{Preservation of cones and balls}

\begin{proposition}[Automorphisms and rank preserving maps]
    Consider a linear map $T:\mcM_d(\CC)\to \mcM_d(\CC)$. The following are equivalent:
    \begin{enumerate}
        \item $T$ maps the cone of positive semidefinite matrices onto itself.
        \item $T$ is positive and preserves the rank of Hermitian matrices.
        \item There is an invertible $Y\in \mcM_d(\CC)$ such that $T$ is either of the form
              $T(X)=YXY^\dagger$ or $T(X)=Y X^T Y^\dagger$.
    \end{enumerate}
\end{proposition}

\begin{proof}
    \textit{1 $\Rightarrow$ 2:} Note first that $T$ is bijective and $T^{-1}$ is again positive, linear and
    maps the positive semidefinite cone onto itself. Consider an arbitrary positive semidefinite $P\in \mcM_d(\CC)$ and
    the cone defined by
    \begin{equation}\label{eq:ch3:3.42}
        C(P):=\{A\in \mcM_d(\CC)\mid \exists c>0:\ P\ge cA\ge 0\}.
    \end{equation}
    Note that the subspace spanned by $C(P)$ has dimension $\rank(P)^2$ and that, by the properties of $T$, we have
    $T(C(P))=C(T(P))$. Since $T$ (as a bijection) preserves the dimension of subspaces, this implies that
    $\rank(P)=\rank(T(P))$.

    In order to extend this to Hermitian matrices let us decompose an arbitrary $H=H^\dagger$ into $H=P_+-P_-$ where $P_\pm$
    are positive semidefinite and orthogonal. Then
    $\rank(T(H))\le \rank(T(P_+))+\rank(T(P_-))=\rank(H)$. Applying the same to $T^{-1}(H)$ instead of $H$ gives the converse
    inequality.

    \textit{2 $\Rightarrow$ 3:} Define $\widetilde T(X):=T(\one)^{-1/2}T(X)T(\one)^{-1/2}$. Since $T(\one)$ has full rank,
    $\widetilde T$ inherits the properties of $T$, i.e., it is positive and rank preserving and in addition unital. Now
    recall that $\lambda\in \spec(H)$ iff $H-\lambda\one$ has reduced rank. Using that
    $\widetilde T(H-\lambda\one)=\widetilde T(H)-\lambda\one$ together with the fact that $\widetilde T$ preserves the rank
    for Hermitian $H$ we get that it actually preserves the spectrum. Then Cor.~1.1 completes this part of the proof by
    setting $Y=U(T(\one))^{-1/2}$.

    \textit{3 $\Rightarrow$ 1} should be obvious.
\end{proof}

From the definition of decomposable positive maps in Eq.~\eqref{eq:ch3:3.2} together with the Kraus decomposition in
Thm.~2.1 it is clear that every extreme point within the convex set of decomposable positive maps is of the form
$X\mapsto YXY^\dagger$ or $X\mapsto YX^T Y^\dagger$ (with $Y$ not necessarily invertible). That the converse is true as
well, i.e., that all these maps are extreme point has been proven in \dots.

\begin{proposition}[Confining Lorentz cones]
    Consider a Hermitian matrix $A\in \mcM_d(\CC)$. Then
    \begin{enumerate}
        \item $A\ge 0 \Rightarrow \tr[A]^2 \ge \tr[A^2]$. Conversely,
        \item $A\ge 0 \Leftarrow \tr[A]^2 \ge (d-1)\tr[A^2]$.
    \end{enumerate}
\end{proposition}

\begin{proof}
    1.\ becomes obvious when writing it out in terms of the eigenvalues $\{a_i\in \RR\}$:
    $(\sum_i a_i)^2\ge \sum_i a_i^2$ is certainly true for $a_i\ge 0$ as the l.h.s.\ contains all terms of the r.h.s.\ plus
    some extra positive terms.

    In order to prove the second implication (by contradiction) assume that $A$ has a negative part such that it admits a
    non-trivial decomposition of the form $A=A_+-A_-$ with $A_\pm\ge 0$ and $\tr[A_+A_-]=0$. Denoting by $d_+<d$ the rank
    of $A_+$ we can bound
    \begin{align}
        \tr[A] &\le \tr[A_+] = \|A_+\|_1 \le \sqrt{d_+} \|A_+\|_2,
        \label{eq:ch3:3.43}
        \\
               &\le \sqrt{d-1} \|A_+\|_2 < \sqrt{d-1} \|A\|_2,
        \label{eq:ch3:3.44}
    \end{align}
    where the first inequality is a simple consequence of H\"older's inequality\footnote{For every unitarily invariant
        norm we have $\|CB\|\le \|\,|C|^{p}\|_1^{1/p}\|\,|B|^{q}\|_1^{1/q}$ for all $p>1$ and $1/p+1/q=1$.
        Eq.~\eqref{eq:ch3:3.43} follows from applying this to the trace norm with $p=2$, $C=A_+$ and $B$ the projection onto the
        range of $A_+$.} and the last one uses $\|A\|_2^2=\|A_+\|_2^2+\|A_-\|_2^2$ and $A_-\neq 0$.
\end{proof}

\begin{lemma}[Making positive maps trace preserving]
    Let $T:\mcM_d(\CC)\to \mcM_{d'}(\CC)$ be a positive map for which $T^*(\one)>0$. There exists an invertible
    $X\in \mcM_d(\CC)$ such that $T'(\rho):=T(X\rho X^\dagger)$ is a trace-preserving positive map.
\end{lemma}

\begin{proof}
    Let $\tau:=(T\otimes \id)(\ketbra{\Omega}{\Omega})$ be the Choi--Jamiolkowski operator corresponding to $T$ and $\tau'$
    the one for $T'$. By Prop.~2.1 the map $T'$ is trace preserving if the partial trace $\tau'_B$ equals $\one/d$. Since
    (again by Prop.~2.1)
    \[
        \tau'_B = X^T \tau_B \overline{X}
        = \frac{1}{d}\,X^T (T^*(\one))^T \overline{X},
    \]
    this can be achieved by choosing $X:=(T^*(\one))^{-1/2}$.
\end{proof}

Hildebrand: \dots\ maps preserving the Lorentz cone \dots

Gorini, Sudarshan: \dots\ maps preserving the sphere + Kossakowski and Chru\'sci\'nski \dots

\subsection{Complexity issues}

\subsection{Literature}

\end{document}
