\chapter{Introduction}\label{ch:intro}

The central result is the Fundamental Theorem for translation-invariant
matrix product vectors. The basic object is a family of matrices
$A=\{A^i\}_{i=0}^{d-1}$, which assigns to each chain length~$N$ the vector
$\ket{V^{(N)}(A)}$ with coefficients indexed by
$\sigma=(i_1,\ldots,i_N)$,
\[
    V^{(N)}(A)_\sigma
    = \tr(A^{i_1}\cdots A^{i_N}).
\]
The main argument, for translation-invariant tensors, follows the canonical-form and
basis-of-normal-tensors argument of~\cite{Cirac2016MPDO_arXiv,Cirac2021Matrix},
with the single-block injective case anchored in the earlier representation
theorem of~\cite{PerezGarcia2007Matrix}.

\section*{Notation}

\begin{center}
    \begin{tabular}{ll}
        $\C$, $\N$, $\Z$            & complex numbers, natural numbers, integers       \\
        $\MN{D}$, $\GL_D(\C)$       & complex matrices and invertible complex matrices \\
        $\tr(X)$, $\spn_{\C}(S)$    & trace and complex linear span                    \\
        $A=\{A^i\}_{i=0}^{d-1}$     & a translation-invariant MPS tensor               \\
        $w=(i_1,\ldots,i_L)$, $A^w$ & a word and the product
        $A^{i_1}\cdots A^{i_L}$                                                        \\
        $\ket{V^{(N)}(A)}$          & the length-$N$ matrix product vector             \\
        $V^{(N)}(A)_\sigma$         & the coefficient of $\ket{V^{(N)}(A)}$ at
        $\sigma$                                                                       \\
        $O_{AB}(N)$                 & the bilinear overlap
        $\sum_\sigma V^{(N)}(A)_\sigma\,\overline{V^{(N)}(B)_\sigma}$                  \\
        $\E_A(X)$                   & the transfer map $\sum_i A^i X(A^i)^\dagger$     \\
        $A^{[L]}$                   & the length-$L$ blocked tensor                    \\
    \end{tabular}
\end{center}

In canonical form one writes
\begin{equation}\label{eq:intro_canonical_form}
    A^i = \bigoplus_{j=1}^{g} \mu_j A_j^i,
    \qquad
    B^i = \bigoplus_{k=1}^{h} \nu_k B_k^i,
\end{equation}
where the displayed blocks are normal tensors. If
$\ket{V^{(N)}(A)}=\ket{V^{(N)}(B)}$ for every positive chain length~$N$, then
the bases of normal tensors have the same size. After relabelling the blocks
of $B$, each matched pair satisfies
\begin{equation}\label{eq:intro_matched_pair_gauge}
    B_j^i = \zeta_j X_j A_j^i X_j^{-1},
\end{equation}
for some invertible matrix $X_j$ and nonzero scalar $\zeta_j$, and the equal
case compares the coefficient power sums attached to the relabelled bases. The
block gauges~(\ref{eq:intro_matched_pair_gauge}) and the weight comparison then
combine into a single gauge conjugacy of the two weighted block-diagonal
tensors~(\ref{eq:intro_canonical_form}). For a single injective block this
reduces to the familiar conclusion $B^i = X A^i X^{-1}$ for all physical
indices $i$, the single-block specialization
of~(\ref{eq:intro_matched_pair_gauge}).

Chapters~\ref{ch:mps}--\ref{ch:ft_proof} carry out this argument.
Chapter~\ref{ch:mps} fixes the matrix product vector notation, word
evaluations, transfer maps, gauge equivalence, injectivity, blocking, and
site-periodic tensor families. Chapter~\ref{ch:single} proves the
single-block theorem by trace pairings, linear extension, simplicity of matrix
algebras, and Skolem--Noether. Chapters~\ref{ch:channels}, \ref{ch:schwarz},
\ref{ch:qpf}, and~\ref{ch:spectral} develop the positive-map,
multiplicative-domain, Perron--Frobenius, and overlap-decay tools used to
control normal blocks. Chapters~\ref{ch:wielandt} and~\ref{ch:canonical}
supply the blocking, primitivity, irreducibility, and canonical-form reductions.
Chapter~\ref{ch:canonical} includes the invariant-projection splitting.
Chapter~\ref{ch:bnt} supplies bases of normal tensors and the
power-sum comparison. The
after-blocking reductions in Section~\ref{sec:reduction_to_normal_form} give
the common primitive block decompositions used immediately in
Chapter~\ref{ch:ft_proof}, which records the conditional theorem
statements.
% Maintainer note: the same-index direct-sum lemmas used after a block
% correspondence has already been fixed are collected later in
% the algebraic-FT draft chapter, because they are facts about block-diagonal
% gauges, not part of the canonical-form reduction or of the BNT block-matching
% proof. The remaining BNT matching and fixed-length span inputs are kept as
% explicit hypotheses until the corresponding source steps are supplied.

The chapters after Chapter~\ref{ch:ft_proof} collect applications,
alternative formulations, and later tensor-network material.
Chapter~\ref{ch:symmetry} uses the Fundamental Theorem to obtain virtual
symmetry gauges, projective bond representations, and string-order criteria;
its Kraus-mixing step appeals only to the later channel-representation chapter,
not to any additional input for the Fundamental Theorem.
Parent Hamiltonians, correlations, examples, channel representations, normal
forms, quantum dynamical semigroups, operator convexity, entropy, and matrix
product operators and density operators are then collected.
A later draft treats the direct Fundamental Theorem for periodically repeated
tensors in irreducible form, together with the compatibility theorems for
physical-index rotation (Definition~\ref{def:rotate_physical} in
Chapter~\ref{ch:symmetry}).
% Maintainer note: the rotation definition is internal bookkeeping for the
% periodic MPS material, not an external-paper input or part of the Molnar
% algebraic-FT chapter.
The algebraic and PEPS draft chapters remain in the repository but are omitted
from the generated blueprint while their statements are being stabilized.
These later topics use the Fundamental Theorem, or develop later
finite-dimensional channel theory, rather than enter its proof.
