\chapter{The Single-Block Fundamental Theorem}\label{ch:single}

This chapter proves that if an injective MPS tensor~$A$ generates
the same MPV family as another tensor~$B$ of the same bond dimension,
then $A$ and $B$ are gauge equivalent.
The proof is purely algebraic and does not use the later
overlap or normal-form reductions.
The trace-pairing maps introduced below are auxiliary devices attached
to the chosen generating families $\{A^i\}$ and $\{B^i\}$; they serve
only to construct the linear map that is later shown to be a gauge
transform.

\section{Trace pairing}

\begin{theorem}[Nondegeneracy of the trace pairing]\label{thm:trace_nondeg}
    \lean{Matrix.trace_mul_right_eq_zero_iff}
    \leanok
    For $M \in \MN{D}$,
    $\tr(M N) = 0$ for all $N \in \MN{D}$
    if and only if $M = 0$.
\end{theorem}

\begin{proof}
    \leanok
    Testing against the matrix units $E_{ji}$
    (the matrix with $1$ in position $(j,i)$ and $0$ elsewhere)
    gives $\tr(E_{ji} M) = M_{ij}$.
    Hence if $\tr(MN) = 0$ for all~$N$, then every entry of $M$
    vanishes.
\end{proof}

\begin{lemma}[Same MPV implies trace agreement]\label{lem:same_mpv_trace}
    \lean{MPSTensor.SameMPV.trace_evalWord}
    \leanok
    \uses{def:same_mpv, def:eval_word}
    If $A$ and $B$ generate the same MPV family, then
    $\tr(A^w) = \tr(B^w)$
    for every word~$w$.
\end{lemma}

\begin{proof}
    \leanok
    \uses{def:mpv}
    Viewing the word $w$ as a configuration of length~$|w|$,
    Definition~\ref{def:mpv} gives the corresponding MPV coefficient as
    the trace coefficient $c_w(A)=\tr(A^w)$.
    The same-MPV hypothesis gives equality at every
    system size.
\end{proof}

\begin{definition}[Trace pairing map]\label{def:trace_pairing_map}
    \lean{MPSTensor.traceMulRightPi}
    \leanok
    \uses{def:mps_tensor}
    For a tensor~$A$, attach to the chosen family $\{A^i\}$ the
    auxiliary linear map $\Phi_A : \MN{D} \to \C^d$ defined by
    \[
        \Phi_A(M)_i =   \tr(M \cdot A^i).
    \]
\end{definition}

\begin{theorem}[Injectivity of the trace pairing map]\label{thm:trace_pairing_injective}
    \lean{MPSTensor.traceMulRightPi_ker_eq_bot}
    \leanok
    \uses{def:trace_pairing_map, def:injective}
    If $A$ is injective, then $\ker \Phi_A = \{0\}$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:trace_nondeg}
    If $\Phi_A(M) = 0$, then $\tr(M \cdot A^i) = 0$ for all~$i$.
    Since the $\{A^i\}$ span $\MN{D}$, this gives $\tr(M N) = 0$
    for all~$N$, so $M = 0$ by Theorem~\ref{thm:trace_nondeg}.
\end{proof}

\begin{theorem}[Nonvanishing trace on injective tensors]\label{thm:trace_ne_zero}
    \lean{MPSTensor.trace_ne_zero_of_injective}
    \leanok
    \uses{def:injective, def:same_mpv}
    If $D \ge 1$, $A$ is injective, $A$ and $B$ generate the same
    MPV family, and $B^i = 0$ for all~$i$, then we have a
    contradiction.
\end{theorem}

\begin{proof}
    \leanok
    \uses{lem:same_mpv_trace}
    If $B^i = 0$ for all~$i$, then
    Lemma~\ref{lem:same_mpv_trace} gives $\tr(A^w) = 0$ for every
    word~$w$, in particular for all length-1 words.
    Since the $\{A^i\}$ span $\MN{D}$, the trace functional vanishes
    identically, contradicting $\tr(\Id_D) = D \neq 0$.
\end{proof}

\section{Linear extension}

\begin{theorem}[Existence and uniqueness of the linear extension]\label{thm:linear_ext}
    \lean{MPSTensor.linearExtension_exists_unique}
    \leanok
    \uses{def:injective, def:same_mpv}
    If $A$ is injective and $A$, $B$ generate the same MPV family,
    then there exists a unique linear map
    $T : \MN{D} \to \MN{D}$ satisfying $T(A^i) = B^i$ for all~$i$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:trace_pairing_injective, lem:same_mpv_trace}
    Write $\ell_A : \C^d \to \MN{D}$ for the map that sends
    a coefficient vector $c$ to $\sum_i c_i A^i$
    (linear combination with the $A^i$).
    Applying Lemma~\ref{lem:same_mpv_trace} to all words of length~$2$
    shows $\Phi_A \circ \ell_A = \Phi_B \circ \ell_B$.
    Since $A$ is injective, $\ell_A$ is surjective, giving
    $\mathrm{range}(\Phi_A) \subseteq \mathrm{range}(\Phi_B)$.
    By Theorem~\ref{thm:trace_pairing_injective},
    $\ker \Phi_A = \{0\}$, so rank--nullity gives
    $\dim(\mathrm{range}(\Phi_A)) = \dim(\MN{D})$.
    Hence $\dim(\mathrm{range}(\Phi_B)) \ge \dim(\MN{D})$,
    and rank--nullity again gives $\ker \Phi_B = \{0\}$.
    A left inverse $g$ of $\Phi_B$ gives $T := g \circ \Phi_A$,
    and uniqueness follows because the $\{A^i\}$ span $\MN{D}$.
\end{proof}

\begin{theorem}[Multiplicativity of the linear extension]\label{thm:linear_ext_mul}
    \lean{MPSTensor.linearExtension_mul}
    \leanok
    \uses{def:injective, def:same_mpv}
    If $A$ is injective and $A$, $B$ generate the same MPV family,
    and $T : \MN{D} \to \MN{D}$ is the unique linear map satisfying
    $T(A^i) = B^i$ for all~$i$, then
    $T(MN) = T(M) T(N)$ for all $M, N \in \MN{D}$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:linear_ext, lem:same_mpv_trace}
    \textbf{Step 1} ($\Phi_B \circ T = \Phi_A$).
    Applying Lemma~\ref{lem:same_mpv_trace} to words of length~$2$
    gives $\Phi_B(T(A^i)) = \Phi_B(B^i) = \Phi_A(A^i)$ for all~$i$.
    Since the $\{A^i\}$ span $\MN{D}$, $\Phi_B \circ T = \Phi_A$.

    \textbf{Step 2} ($T(A^i A^j) = B^i B^j$).
    Lemma~\ref{lem:same_mpv_trace} applied to the word $(i,j,k)$ gives
    $\tr(A^i A^j A^k) = \tr(B^i B^j B^k)$ for all $i,j,k$.
    So $\Phi_B(T(A^i A^j))_k = \Phi_A(A^i A^j)_k = \tr(A^i A^j A^k) = \tr(B^i B^j B^k) = \Phi_B(B^i B^j)_k$.
    The same rank--nullity argument used in the proof of
    Theorem~\ref{thm:linear_ext} gives $\ker \Phi_B = \{0\}$.
    Hence $T(A^i A^j) = B^i B^j$.

    \textbf{Step 3} (bilinear extension).
    For fixed~$j$, the maps $M \mapsto T(M \cdot A^j)$ and
    $M \mapsto T(M) \cdot B^j$ agree on the spanning set
    $\{A^i\}$, hence on all of $\MN{D}$.
    Applying this once more with~$j$ replaced by a general
    matrix gives full multiplicativity.
\end{proof}

\section{Simplicity and Skolem--Noether}

\begin{theorem}[Simplicity of $\MN{D}$]\label{thm:simplicity}
    \lean{MPSTensor.linear_mul_endomorphism_bijective}
    \leanok
    A nonzero multiplicative $\C$-linear endomorphism of $\MN{D}$
    is bijective.
\end{theorem}

\begin{proof}
    \leanok
    The kernel of a multiplicative linear map is a two-sided ideal.
    Since $\MN{D}$ is a simple ring
    (it has no nontrivial two-sided ideals),
    the kernel is either $\{0\}$ or $\MN{D}$.
    The latter would force the map to be zero, contradicting
    the hypothesis.
    Hence the map is injective, and surjectivity follows from
    finite-dimensionality.
\end{proof}

\begin{theorem}[Skolem--Noether]\label{thm:skolem_noether}
    \lean{MPSTensor.skolemNoether_matrix}
    \leanok
    Every $\C$-algebra automorphism of $\MN{D}$ is inner:
    for any $f \in \mathrm{Aut}_{\C\text{-alg}}(\MN{D})$,
    there exists $X \in \GL_D(\C)$ such that
    $f(M) = X M X^{-1}$ for all $M$.
\end{theorem}

\begin{proof}
    \leanok
    Via the canonical identification
    $\MN{D} \cong \mathrm{End}_{\C}(\C^D)$,
    the automorphism $f$ induces an automorphism of
    $\mathrm{End}_{\C}(\C^D)$.
    Every algebra automorphism of $\mathrm{End}(V)$ is
    conjugation by an invertible linear map
    (this is a standard result in the theory of central
    simple algebras).
    Translating back to matrices gives the invertible~$X$.
\end{proof}

\begin{lemma}[Promotion to algebra homomorphism]\label{lem:linear_map_to_alg}
    \lean{MPSTensor.linearMapToAlgHom}
    \leanok
    If $T : \MN{D} \to \MN{D}$ is a multiplicative surjective
    $\C$-linear map, then $T(\Id) = \Id$.
    Equivalently, the same underlying function defines a
    $\C$-algebra homomorphism.
\end{lemma}

\begin{proof}
    \leanok
    By surjectivity, there exists $X$ with $T(X) = \Id$.
    Then $T(\Id) = T(X) \cdot T(\Id) = T(X \cdot \Id) = T(X) = \Id$,
    using multiplicativity and $T(X) = \Id$.
    With $T(\Id) = \Id$ established, $T$ preserves the unit
    and hence is a $\C$-algebra homomorphism.
\end{proof}

\section{Proof of the single-block theorem}

\begin{theorem}[Single-block Fundamental Theorem]\label{thm:ft_single}
    \lean{MPSTensor.fundamentalTheorem_singleBlock}
    \leanok
    \uses{def:injective, def:same_mpv, def:gauge_equiv}
    Let $A$ be an injective MPV tensor and let $B$ be an MPV tensor
    of the same bond dimension.
    If $A$ and $B$ generate the same MPV family, then they are gauge
    equivalent: there exists $X \in \GL_D(\C)$ such that
    $B^i = X A^i X^{-1}$ for all~$i$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:linear_ext, thm:linear_ext_mul, thm:simplicity,
          thm:trace_ne_zero, lem:linear_map_to_alg, thm:skolem_noether}
    By Theorem~\ref{thm:linear_ext}, there is a unique linear
    $T$ with $T(A^i) = B^i$.
    By Theorem~\ref{thm:linear_ext_mul}, $T$ is multiplicative.
    $T$ is nonzero: if $T = 0$ then $B^i = 0$ for all~$i$,
    contradicting Theorem~\ref{thm:trace_ne_zero}.
    By Theorem~\ref{thm:simplicity}, $T$ is bijective.
    By Lemma~\ref{lem:linear_map_to_alg}, $T$ is a
    $\C$-algebra automorphism.
    Theorem~\ref{thm:skolem_noether} gives an invertible $X$
    with $T(M) = X M X^{-1}$,
    and evaluating at $M = A^i$ yields $B^i = X A^i X^{-1}$.
\end{proof}

\section{Alternative fixed-length non-TI setup}

The proof above treats a single translation-invariant tensor and equality of
the full MPV family for all system sizes.
The alternative algebraic route of \cite{Molnar2018NormalPEPS} instead keeps a
fixed periodic chain length and allows the local tensors to vary from site to
site.
The first step is therefore a separate bookkeeping layer for finite
non-translation-invariant chains.

\begin{definition}[Periodic non-TI chain data]\label{def:non_ti_chain}
    \lean{MPSChainTensor, mpsChainCoeff, SameChainState, IsInjectiveChain, ChainGaugeEquiv}
    \leanok
    \uses{def:mps_tensor, def:injective}
    Fix a chain length $n$.
    A \emph{non-translation-invariant periodic MPS chain} consists of sitewise
    tensors
    \[
        A_k = \{A_k^i\}_{i=0}^{d-1}, \qquad k = 0, \dots, n-1,
    \]
    where each $A_k^i \in \MN{D}$.
    For a basis configuration $\sigma = (\sigma_0,\dots,\sigma_{n-1})$, its
    coefficient is
    \[
        c_{A,\sigma}
        := \tr\!\bigl(A_0^{\sigma_0} A_1^{\sigma_1} \cdots A_{n-1}^{\sigma_{n-1}}\bigr).
    \]
    Two such chains generate the \emph{same chain state} if these coefficients
    agree for every $\sigma$.
    The chain is \emph{injective} if each local tensor $A_k$ is injective in
    the usual single-site sense.
    Finally, two chains are \emph{cyclically gauge equivalent} if there are
    invertible bond matrices $Z_0,\dots,Z_{n-1}$ such that
    \[
        B_k^i = Z_k A_k^i Z_{k+1}^{-1}
    \]
    for every site $k$ and physical index $i$, with indices understood modulo
    $n$ so that $Z_n = Z_0$.
\end{definition}

\begin{lemma}[Cyclic gauge equivalence is an equivalence relation]
    \label{lem:chain_gauge_equiv_equivalence}
    \lean{MPSChainTensor.GaugeEquiv.refl, MPSChainTensor.GaugeEquiv.symm,
          MPSChainTensor.GaugeEquiv.trans}
    \leanok
    \uses{def:non_ti_chain}
    On fixed-length periodic chains, cyclic gauge equivalence is reflexive,
    symmetric, and transitive.
\end{lemma}

\begin{proof}
    \leanok
    Reflexivity uses the identity gauge on every bond.
    Symmetry inverts each bond gauge, and transitivity multiplies the gauges
    pointwise along the chain.
\end{proof}
