\chapter{Quantum Channels and Positive Maps}\label{ch:channels}

This chapter develops the theory of positive maps and quantum channels
on matrix algebras, following~\cite{Wolf2012Quantum}.

\section{Positive maps and channels}

\begin{definition}[Positive map]\label{def:positive_map}
    \lean{IsPositiveMap}
    \leanok
    A linear map $E : \MN{D} \to \MN{D}$ is \emph{positive}
    if $E(X) \ge 0$ whenever $X \ge 0$, where $X \ge 0$
    means that $X$ is positive semidefinite.
\end{definition}

\begin{remark}\leanok
    We use the Loewner order on matrices: $X \le Y$ means
    $Y - X \ge 0$.
\end{remark}

\begin{definition}[Associated positive linear map]
    \label{def:positive_map_is_positive_linear_map}
    \lean{IsPositiveMap.toPositiveLinearMap}
    \leanok
    \uses{def:positive_map}
    A positive matrix map $E : \MN{D} \to \MN{D}$ determines the same
    linear map regarded as a positive linear map: if $A \le B$, then
    $E(A) \le E(B)$.
\end{definition}

\begin{definition}[Trace-preserving map]\label{def:trace_preserving}
    \lean{IsTracePreservingMap}
    \leanok
    A linear map $E : \MN{D} \to \MN{D}$ is \emph{trace-preserving}
    if $\tr(E(X)) = \tr(X)$ for all $X$.
\end{definition}

\begin{theorem}[Transposition is positive and trace-preserving]
    \label{thm:transpose_positive_trace_preserving}
    \lean{Matrix.transposeLinearMapComplex_isPositiveMap,
        Matrix.transposeLinearMapComplex_isTracePreservingMap}
    \leanok
    \uses{def:positive_map, def:trace_preserving}
    The matrix transposition map $\theta(X)=X^T$ on $\MN{D}$ is positive
    and trace-preserving.
\end{theorem}

\begin{proof}\leanok
    If $X \ge 0$, write $X=C^\dagger C$. Then
    $X^T=C^T\overline C=\overline C^\dagger\overline C$, so $X^T \ge 0$.
    Trace preservation is the identity $\tr(X^T)=\tr(X)$.
\end{proof}

\begin{definition}[Reduction map]
    \label{def:reduction_map}
    \lean{Matrix.reductionMap}
    \leanok
    The reduction map on $\MN{D}$, for $k\in\N$, is
    \[
        T_k(X)=\tr(X)\Id-k^{-1}X.
    \]
\end{definition}

\begin{theorem}[The first reduction map is positive]
    \label{thm:reduction_map_one_positive}
    \lean{Matrix.reductionMap_one_isPositiveMap}
    \leanok
    \uses{def:positive_map, def:reduction_map}
    For $D\ge 1$, the map $T_1(X)=\tr(X)\Id-X$ on $\MN{D}$ is positive.
\end{theorem}

\begin{proof}\leanok
    If $X\ge 0$, diagonalize $X$.  Its eigenvalues
    $\lambda_1,\ldots,\lambda_D$ are nonnegative, and for each $j$,
    \[
        \lambda_j\le \sum_{i=1}^D \lambda_i=\tr(X).
    \]
    Therefore all eigenvalues of $\tr(X)\Id-X$ are nonnegative.
\end{proof}

\begin{theorem}[Self-duality of the reduction map]
    \label{thm:reduction_map_self_dual}
    \lean{Matrix.traceAdjointMap_reductionMap}
    \leanok
    \uses{def:reduction_map}
    The reduction map is self-dual for the trace pairing:
    \[
        \tr(T_k(\rho)X)=\tr(\rho T_k(X)).
    \]
\end{theorem}

\begin{proof}\leanok
    Expand both sides using
    $T_k(Y)=\tr(Y)\Id-k^{-1}Y$ and bilinearity of the trace.  Both sides are
    \[
        \tr(\rho)\tr(X)-k^{-1}\tr(\rho X).
    \]
\end{proof}

\begin{theorem}[Choi matrix of the reduction map]
    \label{thm:choi_matrix_reduction_map}
    \lean{ChoiJamiolkowski.choiMatrix_reductionMap}
    \leanok
    \uses{def:reduction_map}
    For $D\ge 1$, the Choi matrix of
    $T_k(X)=\tr(X)\Id-k^{-1}X$ is
    \[
        \tau(T_k)=D^{-1}\Id-k^{-1}\ket{\Omega}\bra{\Omega}.
    \]
\end{theorem}

\begin{proof}\leanok
    The trace term sends the $(a,b)$ slice of
    $\ket{\Omega}\bra{\Omega}$ to its trace times the identity, giving
    the contribution $D^{-1}\Id$.  The second term is the Choi matrix of
    $k^{-1}$ times the identity map, namely
    $k^{-1}\ket{\Omega}\bra{\Omega}$.
\end{proof}

\begin{theorem}[Conjugation filters preserve positivity]
    \label{thm:conjugation_filters_preserve_positivity}
    \lean{unitaryConjLM_isCPMap, unitaryConjLM_isPositiveMap,
        IsPositiveMap.comp_unitaryConjLM, unitaryConjLM_isTP_of_unitary,
        unitaryConjLM_isChannel_of_unitary}
    \leanok
    \uses{def:cp_map, def:positive_map, def:trace_preserving, def:channel}
    For every $X\in\MN{D}$, the conjugation filter
    \[
        \rho\longmapsto X\rho X^\dagger
    \]
    is completely positive, hence positive.  Therefore, if
    $T:\MN{D}\to\MN{D}$ is positive, then
    $\rho\mapsto T(X\rho X^\dagger)$ is positive.  If in addition
    $X^\dagger X=\Id$, then the conjugation filter is trace-preserving and
    hence is a quantum channel.
\end{theorem}

\begin{proof}\leanok
    Complete positivity follows from the single Kraus operator $X$.
    Positivity follows immediately.  If $X^\dagger X=\Id$, cyclicity of trace
    gives
    \[
        \tr(X\rho X^\dagger)=\tr(X^\dagger X\rho)=\tr(\rho).
    \]
\end{proof}

\begin{theorem}[Filtered trace-normalization criterion]
    \label{thm:filtered_trace_normalization_criterion}
    \lean{IsPositiveMap.comp_unitaryConjLM_positive_tracePreserving}
    \leanok
    \uses{def:positive_map, def:trace_preserving, def:trace_pairing_adjoint,
        thm:conjugation_filters_preserve_positivity}
    Let $T:\MN{D}\to\MN{D}$ be a positive map, and let $X\in\MN{D}$ satisfy
    \[
        X^\dagger T^*(\Id)X=\Id,
    \]
    where $T^*$ is the trace-pairing adjoint.  Then
    \[
        \rho\longmapsto T(X\rho X^\dagger)
    \]
    is positive and trace-preserving.

    This is the algebraic trace-normalization step in the proof of
    \cite[Chapter~3, Lemma ``Making positive maps trace preserving'']
    {Wolf2012Quantum}.  The inverse-square-root choice is recorded in the
    next theorem.
\end{theorem}

\begin{proof}\leanok
    Positivity follows because $X\rho X^\dagger$ is positive whenever
    $\rho$ is positive, and $T$ is positive.  For trace preservation, use the
    trace-pairing adjoint identity and cyclicity of trace:
    \[
        \tr(T(X\rho X^\dagger))
        =\tr(T^*(\Id)X\rho X^\dagger)
        =\tr(X^\dagger T^*(\Id)X\rho)
        =\tr(\rho).
    \]
\end{proof}

\begin{lemma}[Normalizing conjugation by inverse square root]
    \label{lem:conj_inv_sqrt_normalizes}
    \lean{Matrix.conjTranspose_inv_sqrt_mul_self_mul_inv_sqrt_eq_one_of_posDef}
    \leanok
    \uses{}
    Let $A\in\MN{D}$ be positive definite and set $S=A^{1/2}$.  Then
    \[
        (S^{-1})^\dagger A S^{-1}=\Id .
    \]
\end{lemma}

\begin{proof}\leanok
    Since $A$ is positive definite, $S$ is invertible and self-adjoint.
    Moreover $S^2=A$.  Therefore
    \[
        (S^{-1})^\dagger A S^{-1}
        =
        S^{-1}S^2S^{-1}
        =
        \Id .
    \]
\end{proof}

\begin{theorem}[Trace normalization by inverse square root]
    \label{thm:trace_normalization_inverse_square_root}
    \lean{IsPositiveMap.comp_unitaryConjLM_inv_cfc_sqrt_traceAdjointMap_one}
    \leanok
    \uses{def:positive_map, def:trace_preserving, def:trace_pairing_adjoint,
        thm:filtered_trace_normalization_criterion, lem:conj_inv_sqrt_normalizes}
    Let $T:\MN{D}\to\MN{D}$ be a positive map such that $T^*(\Id)$ is positive
    definite.  Put
    \[
        X=(T^*(\Id))^{-1/2}.
    \]
    Then
    \[
        \rho\longmapsto T(X\rho X^\dagger)
    \]
    is positive and trace-preserving.
\end{theorem}

\begin{proof}\leanok
    Let $S=(T^*(\Id))^{1/2}$.  Since $T^*(\Id)$ is positive definite, $S$ is
    invertible and self-adjoint, and $S^2=T^*(\Id)$.  For $X=S^{-1}$,
    \[
        X^\dagger T^*(\Id)X
        =
        S^{-1}S^2S^{-1}
        =
        \Id .
    \]
    The preceding trace-normalization criterion applies.
\end{proof}

\begin{theorem}[Forward Lorentz-cone trace inequality]
    \label{thm:psd_trace_square_le_square_trace}
    \lean{Matrix.PosSemidef.trace_sq_re_le_trace_re_sq}
    \leanok
    If $A \in \MN{D}$ is positive semidefinite, then
    \[
        \operatorname{Re}\tr(A^2) \le \left(\operatorname{Re}\tr(A)\right)^2 .
    \]
\end{theorem}

\begin{proof}\leanok
    Diagonalize $A$ with nonnegative eigenvalues $\lambda_i$. Then
    $\tr(A^2)=\sum_i \lambda_i^2$ and $\tr(A)=\sum_i \lambda_i$, so the
    desired inequality is
    \[
        \sum_i \lambda_i^2 \le \left(\sum_i \lambda_i\right)^2 ,
    \]
    which follows because all mixed products are nonnegative.
\end{proof}

\begin{theorem}[Trace-nonnegative Lorentz-cone converse]
    \label{thm:trace_nonnegative_lorentz_cone_converse}
    \lean{Matrix.IsHermitian.posSemidef_of_trace_re_nonneg_of_card_sub_one_mul_trace_sq_re_le}
    \leanok
    Let $A\in\MN{D}$ be Hermitian and suppose that
    $\operatorname{Re}\tr(A)\ge0$. If
    \[
        (D-1)\operatorname{Re}\tr(A^2)
        \le \left(\operatorname{Re}\tr(A)\right)^2 ,
    \]
    then $A$ is positive semidefinite.

    This is the trace-nonnegative form of the converse in
    \cite[Proposition~3.9]{Wolf2012Quantum}. The unrestricted squared
    implication printed there is false without a trace sign condition; this is
    recorded in
    \path{docs/paper-gaps/wolf_ch3_lorentz_cone_trace_sign.tex}.
\end{theorem}

\begin{proof}\leanok
    Diagonalize $A$ with real eigenvalues $\lambda_i$. Suppose that some
    eigenvalue $\lambda_0$ is negative. Let
    \[
        T=\sum_i\lambda_i,\qquad R=\sum_{i\ne 0}\lambda_i .
    \]
    Since $T\ge0$ and $T=\lambda_0+R$, one has $T<R$, and hence $T^2<R^2$.
    Cauchy's inequality gives
    \[
        R^2\le (D-1)\sum_{i\ne 0}\lambda_i^2
            \le (D-1)\sum_i\lambda_i^2,
    \]
    contradicting the assumed Lorentz inequality. Thus all eigenvalues are
    nonnegative.
\end{proof}

\begin{definition}[Completely positive map]\label{def:cp_map}
    \lean{IsCPMap}
    \leanok
    A linear map $E : \MN{D} \to \MN{D}$ is \emph{completely positive} (CP)
    if it admits a \emph{Kraus representation}: there exist operators
    $\{K_i\}_{i=0}^{r-1}$ with $K_i \in \MN{D}$ such that, for every
    $X \in \MN{D}$,
    \[
        E(X) = \sum_{i=0}^{r-1} K_i X K_i^\dagger.
    \]
    By Choi's theorem, this is equivalent
    to $E \otimes \id_n$ being positive for all~$n$.
\end{definition}

\begin{theorem}[Transposition is not completely positive]
    \label{thm:transpose_not_completely_positive}
    \lean{ChoiJamiolkowski.transposeLinearMapComplex_not_isCPMap}
    \leanok
    \uses{def:cp_map, def:flip_operator, def:fin_two_antisymm_vec,
        thm:choi_transpose_flip, thm:cp_iff_choi_posSemidef}
    If $D \ge 2$, the matrix transposition map $\theta(X)=X^T$ on $\MN{D}$
    is not completely positive.
\end{theorem}

\begin{proof}\leanok
    \uses{def:flip_operator, def:fin_two_antisymm_vec, thm:choi_transpose_flip,
        thm:cp_iff_choi_posSemidef}
    If $\theta$ were completely positive, its Choi matrix would be positive
    semidefinite. By Theorem~\ref{thm:choi_transpose_flip}, this Choi matrix is
    $D^{-1}F$. For the antisymmetric vector
    $v=\ket{01}-\ket{10}$,
    \[
        \bra{v}\,D^{-1}F\,\ket{v} = -\frac{2}{D} < 0,
    \]
    contradicting positive semidefiniteness.
\end{proof}

\begin{theorem}[Entrywise complete positivity from a Kraus representation]
    \label{thm:cp_map_entrywise_complete_positive}
    \lean{IsCPMap.map_cstarMatrix_nonneg}
    \leanok
    \uses{def:cp_map}
    Let $E : \MN{D} \to \MN{D}$ be completely positive in the Kraus sense.
    For every $k$ and every positive block matrix
    $M \in M_k(\MN{D})$, the entrywise image
    \[
        (E(M_{ab}))_{a,b}
    \]
    is again positive in $M_k(\MN{D})$.
\end{theorem}

\begin{proof}\leanok
    Write $E(X)=\sum_i K_i X K_i^\dagger$.  For each $i$, let $d_i$ be
    the block-diagonal matrix with $K_i$ on every diagonal block.  Then
    \[
        (E(M_{ab}))_{a,b}=\sum_i d_i M d_i^\dagger .
    \]
    Each term on the right is positive because conjugation preserves
    positivity, and a finite sum of positive matrices is positive.
\end{proof}

\begin{theorem}[Kraus maps as abstract completely positive maps]
    \label{thm:cp_map_to_abstract_completely_positive}
    \lean{IsCPMap.toCompletelyPositiveMap}
    \leanok
    \uses{def:cp_map, thm:cp_map_entrywise_complete_positive}
    Every Kraus-represented completely positive map
    $E : \MN{D} \to \MN{D}$ determines a completely positive map between
    the matrix $C^*$-algebras in the entrywise positivity sense.
\end{theorem}

\begin{proof}\leanok
    The defining entrywise positivity condition is exactly
    Theorem~\ref{thm:cp_map_entrywise_complete_positive}.
\end{proof}

\begin{theorem}[CP maps are positive]\label{thm:cp_is_positive}
    \lean{IsCPMap.isPositiveMap}
    \leanok
    \uses{def:cp_map, def:positive_map}
    Every completely positive map is positive.
\end{theorem}

\begin{proof}\leanok
    If $X \ge 0$, then each summand $K_i X K_i^\dagger \ge 0$
    (conjugation preserves positive semidefiniteness),
    so the sum is PSD.
\end{proof}

\begin{theorem}[Channels are positive]\label{thm:channel_positive}
    \lean{IsChannel.pos}
    \leanok
    \uses{def:channel, def:positive_map}
    Every quantum channel is positive.
\end{theorem}

\begin{proof}\leanok\uses{thm:cp_is_positive}
    A quantum channel is completely positive by definition, so
    Theorem~\ref{thm:cp_is_positive} applies.
\end{proof}

\begin{theorem}[Positive maps preserve Hermiticity]
    \label{thm:positive_map_preserves_hermitian}
    \lean{IsPositiveMap.map_isHermitian}
    \leanok
    \uses{def:positive_map_is_positive_linear_map}
    If $E$ is positive and $X = X^\dagger$, then $E(X) = E(X)^\dagger$.
\end{theorem}

\begin{proof}\leanok
    Regard $E$ as the associated positive linear map
    (Definition~\ref{def:positive_map_is_positive_linear_map}). For positive
    linear maps between matrix $C^*$-algebras, $E(Y^\ast)=E(Y)^\ast$.
    Since $X=X^\dagger$ means $X=X^\ast$, one obtains
    $E(X)^\dagger=E(X)^\ast=E(X^\ast)=E(X)$.
\end{proof}

\begin{definition}[Quantum channel]\label{def:channel}
    \lean{IsChannel}
    \leanok
    \uses{def:cp_map, def:trace_preserving}
    A linear map $E : \MN{D} \to \MN{D}$ is a \emph{quantum channel}
    if it is both completely positive and trace-preserving (CPTP),
    following~\cite{Wolf2012Quantum}.
\end{definition}

\begin{definition}[Density matrices]\label{def:density_matrices}
    \lean{densityMatrices, mem_densityMatrices}
    \leanok
    The set of \emph{density matrices} in $\MN{D}$ is
    \[
        \mc{D}_D =  \{\rho \in \MN{D} : \rho \ge 0,  \tr(\rho) = 1\}.
    \]
\end{definition}

\begin{theorem}[Density matrices are compact]\label{thm:density_compact}
    \lean{densityMatrices_isCompact}
    \leanok
    \uses{def:density_matrices}
    $\mc{D}_D$ is compact (in the entry-wise topology on $\MN{D}$).
\end{theorem}

\begin{proof}
    \leanok
    The PSD cone is closed
    (preimage of the closed non-negative cone under continuous
    quadratic forms), and if $\rho \ge 0$ with $\tr(\rho) = 1$
    then each entry satisfies $|\rho_{ij}|^2 \le \rho_{ii}\rho_{jj} \le 1$:
    the diagonal entries are non-negative and sum to~$1$, and the
    off-diagonal bound is the PSD Cauchy--Schwarz inequality.
    Hence the matrix entries are uniformly bounded, and Heine--Borel
    gives compactness.
\end{proof}

\begin{theorem}[Density matrices are convex]\label{thm:density_convex}
    \lean{densityMatrices_isConvex}
    \leanok
    \uses{def:density_matrices}
    $\mc{D}_D$ is convex.
\end{theorem}

\begin{proof}\leanok
    Convex combination of PSD matrices is PSD, and trace is linear.
\end{proof}

\begin{theorem}[Density matrices are nonempty]\label{thm:density_nonempty}
    \lean{densityMatrices_nonempty}
    \leanok
    \uses{def:density_matrices}
    For $D \ge 1$, $\mc{D}_D \neq \varnothing$.
\end{theorem}

\begin{proof}\leanok
    $\frac{1}{D} \Id_D$ is a density matrix.
\end{proof}

\begin{theorem}[Channels preserve density matrices]\label{thm:channel_preserves_density}
    \lean{IsChannel.map_densityMatrices}
    \leanok
    \uses{def:channel, def:density_matrices}
    If $E$ is a quantum channel, then $E(\mc{D}_D) \subseteq \mc{D}_D$.
\end{theorem}

\begin{proof}\leanok\uses{thm:cp_is_positive}
    Complete positivity implies positivity
    (Theorem~\ref{thm:cp_is_positive}),
    so $E(\rho) \ge 0$; trace preservation gives
    $\tr(E(\rho)) = \tr(\rho) = 1$.
\end{proof}

\section{Irreducibility}

\begin{definition}[Orthogonal projection]
    \label{def:orthogonal_projection_channel}
    \lean{IsOrthogonalProjection}
    \leanok
    A matrix $P \in \MN{D}$ is an \emph{orthogonal projection} if
    $P = P^\dagger$ and $P^2 = P$.
\end{definition}

\begin{theorem}[Orthogonal projections and involutive-algebra projections]
    \label{thm:orthogonal_projection_star_projection_channel}
    \lean{IsOrthogonalProjection.isStarProjection}
    \lean{IsStarProjection.isOrthogonalProjection}
    \leanok
    \uses{def:orthogonal_projection_channel}
    For a complex square matrix, the two descriptions
    $P = P^\dagger$, $P^2=P$ and $P^* = P$, $P^2=P$ are equivalent.
\end{theorem}

\begin{proof}\leanok
    Since the involution on $\MN{D}$ is conjugate transpose,
    $P^* = P^\dagger$. Therefore
    \[
        P^\dagger = P,\; P^2 = P
        \iff
        P^* = P,\; P^2 = P.
    \]
\end{proof}

\begin{theorem}[Complement of an orthogonal projection]
    \label{thm:orthogonal_projection_complement_channel}
    \lean{IsOrthogonalProjection.one_sub}
    \leanok
    \uses{def:orthogonal_projection_channel,
        thm:orthogonal_projection_star_projection_channel}
    If $P \in \MN{D}$ is an orthogonal projection, then $\Id-P$ is an
    orthogonal projection.
\end{theorem}

\begin{proof}\leanok
    In the involutive algebra $\MN{D}$, $P$ satisfies $P^\dagger=P$ and
    $P^2=P$. Hence
    \[
        (\Id-P)^\dagger=\Id-P,\qquad
        (\Id-P)^2=\Id-2P+P^2=\Id-P .
    \]
    Thus $\Id-P$ is an orthogonal projection.
\end{proof}

\begin{definition}[Irreducible map]\label{def:irreducible_cp}
    \lean{IsIrreducibleMap}
    \leanok
    \uses{def:orthogonal_projection_channel}
    A linear map
    $E : \MN{D} \to \MN{D}$ is \emph{irreducible}
    if whenever $P$ is an orthogonal projection satisfying
    $E(P \MN{D} P) \subseteq P \MN{D} P$,
    then $P = 0$ or $P = \Id$.
    This is \cite[Theorem~6.2(1)]{Wolf2012Quantum}.
    The definition applies to any linear map; complete positivity
    is not required.
\end{definition}

\begin{theorem}[Injectivity implies irreducibility]\label{thm:injective_irreducible}
    \lean{MPSTensor.injective_implies_irreducibleCP}
    \leanok
    \uses{def:injective, def:transfer_map, def:irreducible_cp}
    If $A$ is an injective MPS tensor, then its transfer map $\E_A$
    is irreducible.
\end{theorem}

\begin{proof}
    \leanok
    If $P$ is an invariant projection for $\E_A$,
    then $(\Id - P) A^i P = 0$ for all~$i$.
    Since the $\{A^i\}$ span $\MN{D}$,
    $(\Id - P) M P = 0$ for all $M$,
    so $P = 0$ or $P = \Id$.
\end{proof}

\begin{theorem}[Transfer map is CP]\label{thm:transfer_cp}
    \lean{MPSTensor.transferMap_isCPMap}
    \leanok
    \uses{def:transfer_map, def:cp_map}
    The transfer map $\E_A(X) = \sum_i A^i X (A^i)^\dagger$
    is completely positive.
\end{theorem}

\begin{proof}\leanok
    It is already written in Kraus form, with Kraus operators $\{A^i\}$.
\end{proof}

\begin{theorem}[Transfer map is a channel]\label{thm:transfer_channel}
    \lean{MPSTensor.transferMap_isChannel}
    \leanok
    \uses{def:transfer_map, def:channel}
    If $\sum_i (A^i)^\dagger A^i = \Id$, then the transfer map
    $\E_A(X) = \sum_i A^i X (A^i)^\dagger$
    is a quantum channel (CPTP).
\end{theorem}

\begin{proof}\leanok\uses{thm:transfer_cp}
    Complete positivity is the observation of Theorem~\ref{thm:transfer_cp}.
    Trace preservation:
    $\tr(\E_A(X)) = \sum_i \tr(A^i X (A^i)^\dagger)
        = \tr((\sum_i (A^i)^\dagger A^i) X) = \tr(X)$
    by cyclicity and the normalization hypothesis.
\end{proof}

\section{Ces\`aro fixed-point theorem}

\begin{definition}[Ces\`aro mean]\label{def:cesaro_mean}
    \lean{cesaroMean}
    \leanok
    The \emph{Ces\`aro mean} of a linear map $E$ at order $N$,
    applied to a starting matrix $\rho_0$, is
    \[
        \sigma_N =   \frac{1}{N} \sum_{n=0}^{N-1} E^n(\rho_0).
    \]
\end{definition}

\begin{theorem}[Ces\`aro telescope]\label{thm:cesaro_telescope}
    \lean{cesaroMean_telescope}
    \leanok
    \uses{def:cesaro_mean}
    $E(\sigma_N) - \sigma_N = \frac{1}{N}(E^N(\rho_0) - \rho_0)$.
\end{theorem}

\begin{proof}\leanok\uses{def:cesaro_mean}
    Telescoping: $E(\sigma_N) = \frac{1}{N}\sum_{n=1}^{N} E^n(\rho_0)$
    and subtracting $\sigma_N$ cancels all but the first and last terms.
\end{proof}

\begin{theorem}[Hermitian fixed-point decomposition]\label{thm:hermitian_fixedpoint_parts}
    \lean{IsChannel.posSemidef_parts_of_hermitian_fixedPoint}
    \leanok
    \uses{def:channel}
    Let $E$ be a quantum channel and $X = E(X)$ a Hermitian
    fixed point with decomposition $X = P_+ - P_-$ into orthogonal
    positive and negative parts ($P_\pm \ge 0$, $\tr(P_+ P_-) = 0$).
    Then $E(P_+) = P_+$ and $E(P_-) = P_-$.
    This is \cite[Proposition~6.8]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}
    \leanok
    Let $Q$ be the support projection of $P_+$.
    Trace preservation and positivity give
    $\tr(Q E(P_-)) = 0$
    and $E(P_-) \ge 0$, so $Q E(P_-) = 0$.
    The fixed-point equation
    $E(P_+) - E(P_-) = P_+ - P_-$ then identifies
    $E(P_\pm) = P_\pm$.
\end{proof}

\begin{theorem}[Ces\`aro fixed-point theorem]\label{thm:cesaro_fixedpoint}
    \lean{IsChannel.exists_posSemidef_fixedPoint}
    \leanok
    \uses{def:channel, def:density_matrices, def:cesaro_mean}
    Every quantum channel $E : \MN{D} \to \MN{D}$ (with $D \ge 1$)
    has a nonzero positive semidefinite fixed point:
    there exists $\rho \ge 0$, $\rho \neq 0$, with $E(\rho) = \rho$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:density_compact, thm:density_nonempty,
        thm:channel_preserves_density, thm:cesaro_telescope}
    Start with any $\rho_0 \in \mc{D}_D$
    (Theorem~\ref{thm:density_nonempty}).
    Each Ces\`aro mean $\sigma_N$ lies in $\mc{D}_D$
    by applying channel preservation to every iterate, adding positive
    semidefinite matrices, rescaling by the positive average factor, and using
    trace linearity.
    By compactness (Theorem~\ref{thm:density_compact}),
    extract a convergent subsequence
    $\sigma_{N_k} \to \rho$.
    The telescope identity (Theorem~\ref{thm:cesaro_telescope})
    gives
    $\|E(\sigma_N) - \sigma_N\| = \frac{1}{N}\|E^N(\rho_0) - \rho_0\| \to 0$,
    so $E(\rho) = \rho$ by continuity.
    In~\cite[Theorem~6.11]{Wolf2012Quantum}, existence is proved
    via Brouwer's fixed-point theorem; we use the Ces\`aro
    argument instead.
\end{proof}

\section{Fixed-point projection}

\begin{definition}[Fixed-point projection]\label{def:fixed_point_proj}
    \lean{fixedPointProj}
    \leanok
    Given a linear map $E : \MN{D} \to \MN{D}$ and a fixed point
    $\rho$ with $E(\rho) = \rho$ and $\tr(\rho) \neq 0$,
    the \emph{fixed-point projection} is the rank-one map
    \[
        P(X) =  \frac{\tr(X)}{\tr(\rho)} \rho.
    \]
    This projects onto the span of $\rho$ along the kernel of the
    trace functional. When the fixed point is unique up to scaling,
    this span is the full fixed-point space.
\end{definition}

\begin{theorem}[Power decomposition]\label{thm:pow_decomp}
    \lean{pow_eq_fixedPointProj_add_compl_pow}
    \leanok
    \uses{def:fixed_point_proj, def:trace_preserving}
    If $E$ is trace-preserving and $E(\rho) = \rho$, then
    for $n \ge 1$,
    \[
        E^n =  P + (E - P)^n,
    \]
    where $P$ is the fixed-point projection.
\end{theorem}

\begin{proof}
    \leanok
    \uses{def:fixed_point_proj}
    $P$ is idempotent ($P^2 = P$), the fixed-point relation
    $E(\rho) = \rho$ gives $E \circ P = P$, and
    trace preservation gives $P \circ E = P$.
    These imply $(E-P) \circ P = 0$ and $P \circ (E-P) = 0$,
    so in the binomial expansion of $(P + (E-P))^n$
    all cross terms vanish, leaving $E^n = P^n + (E-P)^n = P + (E-P)^n$.
\end{proof}

\section{Peripheral spectrum and primitivity}

\begin{definition}[Peripheral spectrum]\label{def:peripheral_spectrum}
    \lean{peripheralSpectrum}
    \leanok
    The \emph{peripheral spectrum} of a bounded linear operator $T$
    is the set of spectral values $\mu$ with
    $|\mu|$ equal to the spectral radius of $T$.
\end{definition}

\begin{definition}[Peripheral eigenvalues]\label{def:peripheral_eigenvalues}
    \lean{peripheralEigenvalues}
    \leanok
    The \emph{peripheral eigenvalues} of a linear map $E$ on
    a finite-dimensional space
    are the eigenvalues $\lambda$ on the unit circle:
    $|\lambda| = 1$.
    For a channel, the spectral radius is~$1$
    \cite[Proposition~6.1]{Wolf2012Quantum}, so these coincide with
    the eigenvalues of maximal modulus.

    The condition $|\lambda| = 1$ is appropriate for channels, since the
    spectral radius of a channel is~$1$; for a general linear map with
    spectral radius $r \neq 1$ the condition would generalise to
    $|\lambda| = r$.
\end{definition}

\begin{remark}[Peripheral conventions for channels]%
    \label{rem:peripheral_conventions_channels}
    \leanok
    \uses{def:peripheral_spectrum, def:peripheral_eigenvalues}
    The spectral-radius convention of
    Definition~\ref{def:peripheral_spectrum} is the usual bounded-operator
    notion. For channels the spectral radius is $1$, so the channel arguments
    below use the unit-circle set of Definition~\ref{def:peripheral_eigenvalues}.
\end{remark}

\begin{theorem}[$1$ is a peripheral eigenvalue]\label{thm:one_peripheral}
    \lean{one_mem_peripheralEigenvalues}
    \leanok
    \uses{def:peripheral_eigenvalues}
    If $E$ has a nonzero fixed point $\rho \neq 0$ with $E(\rho) = \rho$,
    then $1$ is a peripheral eigenvalue of $E$.
\end{theorem}

\begin{proof}\leanok
    $\rho$ is an eigenvector with eigenvalue~$1$, and $|1| = 1$.
\end{proof}

\begin{theorem}[Finiteness of peripheral eigenvalues]\label{thm:peripheral_finite}
    \lean{peripheralEigenvalues_finite}
    \leanok
    \uses{def:peripheral_eigenvalues}
    On a finite-dimensional space, the set of peripheral eigenvalues
    is finite.
\end{theorem}

\begin{proof}\leanok
    Peripheral eigenvalues are a subset of all eigenvalues,
    which is a finite set in finite dimensions.
\end{proof}

\begin{theorem}[Power-stable peripheral eigenvalues are roots of unity]\label{thm:peripheral_roots_of_unity}
    \lean{peripheral_isRootOfUnity_of_pow_eigenvalue}
    \leanok
    \uses{def:peripheral_eigenvalues}
    Let $E$ be a linear endomorphism on a finite-dimensional
    space.
    If $\mu$ is a peripheral eigenvalue of $E$
    and $\mu^n$ is an eigenvalue of $E$ for every $n \ge 1$,
    then $\mu$ is a root of unity.
\end{theorem}

\begin{proof}
    \leanok
    Since $\mu^n$ is an eigenvalue for every positive~$n$,
    the pigeonhole principle on the finite eigenvalue set gives
    $\mu^a = \mu^b$ for some $a \neq b$, hence $\mu^{|a-b|} = 1$.
\end{proof}

\begin{theorem}[Peripheral powers imply root of unity]\label{thm:peripheral_powers}
    \lean{peripheral_isRootOfUnity_of_closed_powers}
    \leanok
    \uses{def:peripheral_eigenvalues}
    If the set of peripheral eigenvalues of a linear
    endomorphism~$E$ on a finite-dimensional space is
    \emph{closed under powers}
    ($\mu \in \mathrm{peripheral}(E)$
    and $n \ge 1$ imply
    $\mu^n \in \mathrm{peripheral}(E)$),
    then every peripheral eigenvalue is a root of unity.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:peripheral_roots_of_unity}
    The closure hypothesis provides
    $\mu^n \in \mathrm{peripheral}(E)$ for all positive~$n$,
    which in particular means $\mu^n$ is an eigenvalue.
    Theorem~\ref{thm:peripheral_roots_of_unity} then gives
    $\mu^p = 1$ for some $p > 0$.
\end{proof}

\subsection{Periodicity removal by powering}\label{subsec:periodicity_removal_power}

\begin{remark}\label{rem:periodicity_removal_spectral}\leanok
    The results in this subsection are purely spectral: they use only
    finiteness of a set of complex numbers and the spectral mapping theorem
    for powers. In particular, they do not rely on complete positivity.
    For transfer maps, replacing a tensor by its $p$-blocked tensor replaces
    $\E_A$ by $\E_A^p$, so this powering step is the operator-theoretic
    form of blocking.
\end{remark}

\begin{lemma}[Common exponent for a finite set of roots of unity]\label{lem:common_power_eq_one}
    \lean{exists_common_power_eq_one_of_finite}
    \leanok
    Let $s \subseteq \C$ be a finite set.
    Assume that for every $\mu \in s$ there exists an exponent $p_\mu > 0$
    with $\mu^{p_\mu} = 1$.
    Then there exists $p > 0$ such that $\mu^p = 1$ for all $\mu \in s$.
\end{lemma}

\begin{proof}\leanok
    Take $p$ to be the least common multiple of the exponents $p_\mu$.
\end{proof}

\begin{lemma}[Powering collapses peripheral eigenvalues]\label{lem:peripheral_pow_singleton}
    \lean{peripheralEigenvalues_pow_eq_singleton}
    \leanok
    \uses{def:peripheral_eigenvalues}
    Let $E : \MN{D} \to \MN{D}$ be a linear map, and assume that $E$ has a
    nonzero fixed point $\rho \neq 0$.
    Let $p > 0$.
    If every peripheral eigenvalue $\mu$ of $E$ satisfies $\mu^p = 1$,
    then $\mathrm{peripheral}(E^p) = \{1\}$.
\end{lemma}

\begin{proof}\leanok
    \uses{thm:one_peripheral}
    The fixed point $\rho$ gives $1 \in \mathrm{peripheral}(E^p)$.
    For the reverse inclusion, let $\nu \in \mathrm{peripheral}(E^p)$.
    Spectral mapping gives $\nu = \mu^p$ for some $\mu \in \sigma(E)$.
    From $|\nu| = 1$ and $p > 0$ we obtain $|\mu| = 1$, hence
    $\mu \in \mathrm{peripheral}(E)$.
    The hypothesis $\mu^p = 1$ then forces $\nu = 1$.
\end{proof}

For a normalized MPS tensor, the transfer map is naturally trace-preserving but need
not be unital. In general one should not expect a single gauge choice to make it
both unital and trace-preserving.

\begin{remark}[Bi-canonical special case]\label{rem:bicanonical_special_case}\leanok
    When a Kraus map happens to be both unital and trace-preserving
    (the ``bi-canonical'' case), the Kadison--Schwarz equality at peripheral
    eigenvectors (Theorem~\ref{thm:ks_peripheral_channels}) shows directly that the
    peripheral eigenvalues are closed under powers, and hence are roots of unity
    by Theorem~\ref{thm:peripheral_powers}. It
    requires both normalizations simultaneously. The more general adjoint-fixed-point
    formulation below covers the case where only unitality and a positive definite
    adjoint fixed point are available.
\end{remark}

\subsection{The Kadison--Schwarz inequality and the multiplicative domain}
\label{subsec:ks_input_channels}

\begin{definition}[Kraus maps, adjoints, and normalizations]
    \label{def:kraus_map_channels}
    \lean{KadisonSchwarz.krausMap, KadisonSchwarz.krausAdjointMap,
        KadisonSchwarz.IsUnitalKraus, KadisonSchwarz.IsTPKraus}
    \leanok
    For operators $\{K_i\}_{i=0}^{d-1} \subseteq \MN{D}$, the associated
    \emph{Kraus map} and \emph{adjoint Kraus map} are
    \[
        E(X) = \sum_i K_i X K_i^\dagger,
        \qquad
        E^*(X) = \sum_i K_i^\dagger X K_i.
    \]
    The Kraus family is \emph{unital} when $\sum_i K_i K_i^\dagger = \Id$ and
    \emph{trace-preserving} when $\sum_i K_i^\dagger K_i = \Id$.
\end{definition}

\begin{theorem}[Kadison--Schwarz inequality]
    \label{thm:kadison_schwarz_channels}
    \lean{KadisonSchwarz.kadison_schwarz}
    \leanok
    \uses{def:kraus_map_channels}
    Let $E(X) = \sum_i K_i X K_i^\dagger$ be a unital Kraus map.
    Then for every $X \in \MN{D}$,
    \[
        E(X^\dagger X) \ge E(X)^\dagger E(X).
    \]
    This is \cite[Equation~(5.2)]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}\leanok
    Apply the unital completely positive map entrywise to the positive
    $2 \times 2$ block matrix
    $\bigl(\begin{smallmatrix} X^\dagger X & X^\dagger \\ X & \Id
        \end{smallmatrix}\bigr)$ and take the Schur complement of the lower-right block.
\end{proof}

\begin{theorem}[Hilbert--Schmidt contraction]
    \label{thm:hs_contraction_channels}
    \lean{KadisonSchwarz.hilbertSchmidt_contraction}
    \leanok
    \uses{def:kraus_map_channels, thm:kadison_schwarz_channels}
    If a Kraus map is both unital and trace-preserving, then
    \[
        \tr(E(X)^\dagger E(X)) \le \tr(X^\dagger X)
    \]
    for all $X \in \MN{D}$.
\end{theorem}

\begin{proof}\leanok\uses{thm:kadison_schwarz_channels}
    The Kadison--Schwarz gap is positive semidefinite by
    Theorem~\ref{thm:kadison_schwarz_channels}. Trace preservation then identifies
    the traces of its two terms.
\end{proof}

\begin{theorem}[KS gap decomposition]
    \label{thm:ks_gap_decomp_channels}
    \lean{KadisonSchwarz.ks_gap_eq_sum_squares}
    \leanok
    \uses{def:kraus_map_channels}
    For a unital Kraus map,
    \[
        E(X^\dagger X) - E(X)^\dagger E(X)
        = \sum_i R_i^\dagger R_i,
        \qquad
        R_i := X K_i^\dagger - K_i^\dagger E(X).
    \]
\end{theorem}

\begin{proof}\leanok
    Expand the right-hand side, distribute the products, and use
    $\sum_i K_i K_i^\dagger = \Id$ to recover the Kadison--Schwarz gap.
\end{proof}

\begin{theorem}[KS equality implies Kraus commutation]
    \label{thm:kraus_commute_channels}
    \lean{KadisonSchwarz.kraus_commute_of_ks_equality}
    \leanok
    \uses{thm:ks_gap_decomp_channels}
    Let $E$ be a unital Kraus map.
    If $E(X^\dagger X) = E(X)^\dagger E(X)$, then, for every~$i$,
    \[
        X K_i^\dagger = K_i^\dagger E(X).
    \]
\end{theorem}

\begin{proof}\leanok\uses{thm:ks_gap_decomp_channels}
    The Kadison--Schwarz gap is a sum of positive semidefinite terms
    $\sum_i R_i^\dagger R_i$. If the gap vanishes, each $R_i$ must vanish.
\end{proof}

\begin{theorem}[KS equality for peripheral eigenvectors]
    \label{thm:ks_peripheral_channels}
    \lean{KadisonSchwarz.ks_equality_of_peripheral_eigenvector}
    \leanok
    \uses{def:kraus_map_channels, thm:kadison_schwarz_channels}
    Let $E$ be a Kraus map that is both unital and trace-preserving.
    If $E(X) = \mu X$ with $|\mu| = 1$, then
    \[
        E(X^\dagger X) = E(X)^\dagger E(X).
    \]
\end{theorem}

\begin{proof}\leanok
    The Kadison--Schwarz gap is positive semidefinite by
    Theorem~\ref{thm:kadison_schwarz_channels}. Trace preservation and the relation
    $E(X)=\mu X$ with $|\mu|=1$ show that its trace is zero, hence the gap vanishes.
\end{proof}

\begin{theorem}[Left multiplicative identity from KS equality]
    \label{thm:left_multiplicative_identity_channels}
    \lean{KadisonSchwarz.multiplicative_domain_left}
    \leanok
    \uses{thm:kraus_commute_channels}
    Let $E$ be a unital Kraus map.
    If $E(X^\dagger X) = E(X)^\dagger E(X)$, then for every $Y \in \MN{D}$,
    \[
        E(X^\dagger Y) = E(X)^\dagger E(Y).
    \]
\end{theorem}

\begin{proof}\leanok\uses{thm:kraus_commute_channels}
    The commutation relation of Theorem~\ref{thm:kraus_commute_channels}
    lets one move $X^\dagger$ past each Kraus operator inside the sum.
\end{proof}

\begin{definition}[Multiplicative domain]
    \label{def:multiplicative_domain_channels}
    \lean{KadisonSchwarz.multiplicativeDomain}
    \leanok
    \uses{def:kraus_map_channels}
    The \emph{multiplicative domain} of a Kraus map $E$ is the set of matrices
    that are simultaneously left- and right-multiplicative for $E$.
\end{definition}

\begin{theorem}[Multiplicative domain is a $*$-subalgebra]
    \label{thm:md_star_subalgebra_channels}
    \lean{KadisonSchwarz.multiplicativeDomainStarSubalgebra}
    \leanok
    \uses{def:multiplicative_domain_channels, def:kraus_map_channels}
    For a unital Kraus map, the multiplicative domain is a $*$-subalgebra of
    $\MN{D}$.
\end{theorem}

\begin{proof}\leanok
    The multiplicative identities are stable under addition, multiplication,
    and adjoint, and they contain the identity matrix.
\end{proof}

\subsection{Peripheral closure via adjoint fixed point}%
\label{subsec:peripheral_closure_fixedpoint}

The following results replace the trace-preservation hypothesis by the existence
of a positive definite fixed point for the \emph{adjoint} Kraus map,
matching the weighted-trace argument in
\cite[Appendix~A]{Cirac2016MPDO_arXiv}.

\begin{lemma}[Peripheral closure via adjoint fixed point]%
    \label{lem:peripheral_closed_powers_irred_fp}
    \lean{MPSTensor.peripheralEigenvalues_pow_mem_of_irreducible_unital_of_adjoint_fixedPoint}
    \leanok
    \uses{def:kraus_map_channels, def:irreducible_cp}
    Let $E(X) = \sum_i K_i X K_i^\dagger$ be a Kraus map that is
    unital ($\sum_i K_i K_i^\dagger = \Id$), and assume:
    \begin{enumerate}
        \item the adjoint Kraus map $E^*(X) = \sum_i K_i^\dagger X K_i$
              has a positive definite fixed point $\rho > 0$, and
        \item $E$ is irreducible.
    \end{enumerate}
    Then $\mathrm{peripheral}(E)$ is closed under powers:
    if $\mu \in \mathrm{peripheral}(E)$ then $\mu^n \in \mathrm{peripheral}(E)$
    for every $n \in \N$.
\end{lemma}

\begin{proof}\leanok
    \uses{thm:kadison_schwarz_channels, thm:psd_fp_pd_irred,
        thm:kraus_commute_channels, thm:left_multiplicative_identity_channels}
    Let $E(X) = \mu X$ with $|\mu| = 1$, and set
    $G := E(X^\dagger X) - E(X)^\dagger E(X)$.
    By Kadison--Schwarz, $G \ge 0$.
    Since $E^*(\rho) = \rho$,
    \[
        \tr(\rho G)
        = \tr(\rho E(X^\dagger X)) - \tr(\rho E(X)^\dagger E(X))
        = \tr(E^*(\rho) X^\dagger X) - \tr(\rho E(X)^\dagger E(X))
        = \tr(\rho X^\dagger X) - |\mu|^2 \tr(\rho X^\dagger X)
        = 0.
    \]
    Thus the trace of the Kadison--Schwarz gap against the adjoint
    fixed point vanishes. Because $\rho > 0$ and $G \ge 0$, this forces
    $G = 0$, so $E(X^\dagger X) = E(X)^\dagger E(X)$.
    Hence $X^\dagger X$ is a PSD fixed point, which is
    positive definite by irreducibility
    (Theorem~\ref{thm:psd_fp_pd_irred}).
    So $X$ is invertible.
    By Theorem~\ref{thm:kraus_commute_channels} (KS equality implies Kraus commutation),
    $X$ commutes with each Kraus operator; iterating via the left
    multiplicative identity (Theorem~\ref{thm:left_multiplicative_identity_channels})
    gives $E(X^n) = \mu^n X^n$, so $\mu^n$ is peripheral.
\end{proof}

\begin{theorem}[Roots of unity via adjoint fixed point]%
    \label{thm:peripheral_roots_irred_fp}
    \lean{MPSTensor.peripheral_isRootOfUnity_of_irreducible_unital_of_adjoint_fixedPoint}
    \leanok
    \uses{def:kraus_map_channels, def:irreducible_cp, def:peripheral_eigenvalues}
    Let $E(X) = \sum_i K_i X K_i^\dagger$ be a Kraus map that is
    unital, and assume that the adjoint Kraus map
    $E^*(X) = \sum_i K_i^\dagger X K_i$ has a positive definite fixed
    point and that $E$ is irreducible.
    Then every peripheral eigenvalue of $E$ is a root of unity.
\end{theorem}

\begin{proof}\leanok
    \uses{lem:peripheral_closed_powers_irred_fp, thm:peripheral_powers}
    Combine the closure-under-powers result
    (Lemma~\ref{lem:peripheral_closed_powers_irred_fp})
    with Theorem~\ref{thm:peripheral_powers}.
\end{proof}

\begin{remark}\label{rem:adjoint_fp_more_general}\leanok
    Lemma~\ref{lem:peripheral_closed_powers_irred_fp}
    and Theorem~\ref{thm:peripheral_roots_irred_fp}
    cover the bi-canonical (unital + TP) case as well:
    trace preservation gives
    $E^*(\Id) = \Id$, which is a positive definite fixed
    point of the adjoint.
    The adjoint-fixed-point formulation matches the
    argument in \cite[Appendix~A]{Cirac2016MPDO_arXiv},
    which uses a faithful invariant state rather than
    bi-canonicality.
    Theorem~\ref{thm:peripheral_roots_irred_fp} establishes the
    root-of-unity conclusion in this adjoint-fixed-point formulation.
    In the same adjoint-fixed-point formulation, Chapter~\ref{ch:spectral}
    proves the corresponding cyclic structure: the cyclic group of
    peripheral eigenvalues is Theorem~\ref{thm:peripheral_cyclic_structure},
    and the cyclic projection decomposition is
    Theorem~\ref{thm:cyclic_decomposition_irreducible_schwarz}.
\end{remark}

\begin{definition}[Channel period]\label{def:channel_period}
    \lean{channelPeriod}
    \leanok
    \uses{def:peripheral_eigenvalues}
    The \emph{period} of a channel $E$ is the number of
    peripheral eigenvalues (counted without multiplicity):
    $p(E) := |\mathrm{peripheral}(E)|$.
\end{definition}

\begin{definition}[Primitive map]\label{def:primitive_channel}
    \lean{IsPrimitive}
    \leanok
    \uses{def:peripheral_eigenvalues}
    A linear map is \emph{primitive} if its only peripheral eigenvalue
    is $\lambda = 1$, i.e.\
    $\mathrm{peripheral}(E) = \{1\}$.
    For channels this corresponds to \cite[Theorem~6.7]{Wolf2012Quantum}.
    This definition applies to any linear endomorphism, not only
    to channels.
\end{definition}

\begin{theorem}[Primitivity equals period one]\label{thm:primitive_iff_period_one}
    \lean{isPrimitive_iff_period_one}
    \leanok
    \uses{def:primitive_channel, def:channel_period}
    A linear map with a nonzero fixed point is primitive if and only if
    its period is~$1$.
\end{theorem}

\begin{proof}\leanok\uses{thm:one_peripheral}
    If $\mathrm{peripheral}(E) = \{1\}$, then
    $|\mathrm{peripheral}(E)| = 1$.
    Conversely, if $|\mathrm{peripheral}(E)| = 1$,
    then since $1 \in \mathrm{peripheral}(E)$
    (Theorem~\ref{thm:one_peripheral}), the
    only element is $1$.
\end{proof}

\begin{remark}[Period-one characterization]%
    \label{rem:primitive_period_one_characterization}
    \leanok
    \uses{def:channel_period, thm:primitive_iff_period_one}
    The period $p(E)$ counts peripheral eigenvalues without multiplicity.
    Theorem~\ref{thm:primitive_iff_period_one} is therefore the period-one
    characterization of primitivity. The spectral estimates below are stated
    directly with Definition~\ref{def:primitive_channel}, since they use the
    absence of peripheral eigenvalues other than $1$, not a separate counting
    hypothesis.
\end{remark}

\begin{theorem}[Complementary transfer-map gap for primitive maps]%
    \label{thm:primitive_compl_lt_one}
    \lean{compl_eigenvalue_norm_lt_one_of_primitive}
    \leanok
    \uses{def:primitive_channel, def:fixed_point_proj, def:trace_preserving}
    Let $E : \MN{D} \to \MN{D}$ be trace-preserving, and let
    $\rho \neq 0$ satisfy $E(\rho) = \rho$ and $\tr(\rho) \neq 0$.
    Let
    \[
        P(X) = \frac{\tr(X)}{\tr(\rho)} \rho
    \]
    be the fixed-point projection associated to $\rho$.
    Assume that:
    \begin{enumerate}
        \item $E$ is primitive;
        \item every eigenvalue $\mu$ of $E$ satisfies $|\mu| \le 1$;
        \item whenever $E(X) = X$ and $\tr(X) = 0$, one has $X = 0$.
    \end{enumerate}
    Then every eigenvalue $\nu$ of $E - P$ satisfies $|\nu| < 1$.
\end{theorem}

\begin{proof}
    \leanok
    Let $(E - P)(X) = \nu X$ with $X \neq 0$.
    Then
    \[
        E(X) = \nu X + P(X).
    \]
    If $\tr(X) = 0$, then $P(X) = 0$, so $E(X) = \nu X$ and hence
    $\nu$ is an eigenvalue of $E$ with $|\nu| \le 1$.
    If $|\nu| = 1$, primitivity forces $\nu = 1$, so $X$ is a
    trace-zero fixed point of $E$; by assumption this implies $X = 0$,
    a contradiction. Thus $|\nu| < 1$.

    If instead $\tr(X) \neq 0$, trace preservation and
    $\tr(P(X)) = \tr(X)$ give
    \[
        \tr(X) = \tr(E(X)) = \nu \tr(X) + \tr(P(X))
        = \nu \tr(X) + \tr(X),
    \]
    so $\nu \tr(X) = 0$ and therefore $\nu = 0$.
    In either case, $|\nu| < 1$.
\end{proof}

%% =========================================================
\section{Fixed-point algebra}
\label{sec:fixed_point_algebra}
%% =========================================================

\begin{definition}[Kraus fixed points]\label{def:kraus_fixed_points}
    \lean{Kraus.fixedPoints}\leanok
    \uses{def:kraus_map_channels}
    The \emph{fixed-point set} of a Kraus map
    $E(X) = \sum_i K_i X K_i^\dagger$ is
    $\mathrm{Fix}(E) = \{X \in \MN{D} : E(X) = X\}$.
\end{definition}

\begin{theorem}[Fixed points lie in the multiplicative domain]
    \label{thm:fixedPoints_in_multiplicativeDomain}
    \lean{Kraus.fixedPoints_in_multiplicativeDomain}
    \leanok
    \uses{def:kraus_fixed_points, def:multiplicative_domain_channels,
        def:kraus_map_channels}
    Let $E$ be a trace-preserving Kraus map whose Schr{\"o}dinger-picture map has a
    positive definite fixed point $\rho > 0$.
    Then every adjoint fixed point belongs to the multiplicative domain of the
    adjoint Kraus family.
\end{theorem}

\begin{proof}\leanok
    The Kadison--Schwarz equality for fixed points forces both one-sided
    multiplicative identities, so the fixed point lies in the full
    multiplicative domain.
\end{proof}

\begin{theorem}[Fixed points form a $*$-subalgebra]
    \label{thm:fixedPointsStarSubalgebra}
    \lean{Kraus.fixedPointsStarSubalgebra}\leanok
    \uses{def:kraus_fixed_points, def:kraus_map_channels}
    Let $E$ be a unital Kraus map whose adjoint map $E^*$ has a positive
    definite fixed point $\rho > 0$.
    Then $\mathrm{Fix}(E)$ is a $*$-subalgebra of $\MN{D}$.
    This is \cite[Theorem~6.12]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}\leanok
    \uses{thm:kadison_schwarz_channels, thm:left_multiplicative_identity_channels}
    Closure under addition and scalar multiplication is immediate from
    linearity.
    Closure under $\dagger$: if $E(X) = X$ then $E(X^\dagger) = X^\dagger$
    because fixed points are stable under adjoint.
    Closure under multiplication: if $E(X) = X$ and $E(Y) = Y$, the weighted
    Kadison--Schwarz argument gives
    $E(X^\dagger X) = E(X)^\dagger E(X)$ and, applied to $X^\dagger$,
    also $E(X X^\dagger) = E(X) E(X)^\dagger$. Thus $X$ lies in the
    multiplicative domain, and Theorem~\ref{thm:left_multiplicative_identity_channels}
    yields $E(XY) = E(X) E(Y) = XY$.
\end{proof}

\begin{remark}\leanok
    Theorem~\ref{thm:fixedPoints_in_multiplicativeDomain} is the
    trace-preserving / Heisenberg-picture counterpart of the same
    multiplicative-domain argument.
\end{remark}

\begin{theorem}[Projected elements lie in the multiplicative domain]
    \label{thm:map_mem_multiplicativeDomain_of_idempotent}
    \lean{Kraus.map_mem_multiplicativeDomain_of_idempotent}\leanok
    \uses{def:kraus_fixed_points, def:kraus_map_channels}
    Let $E$ be a unital Kraus map whose adjoint map has a positive definite
    fixed point, and suppose $E^2 = E$. Then $E(X)$ lies in the multiplicative
    domain of $E$ for every $X \in \MN{D}$.
\end{theorem}

\begin{proof}\leanok
    \uses{thm:fixedPointsStarSubalgebra}
    Idempotence gives $E(E(X)) = E(X)$, so $E(X)$ is a fixed point.
    By the argument in Theorem~\ref{thm:fixedPointsStarSubalgebra}, the
    Kadison--Schwarz gap for the fixed point $E(X)$ vanishes, hence
    \[
        E(E(X)^\dagger E(X)) = E(X)^\dagger E(X).
    \]
    Fixed points are closed under $\dagger$, so the same argument applied to
    $E(X)^\dagger$ gives
    \[
        E(E(X) E(X)^\dagger) = E(X) E(X)^\dagger.
    \]
    The vanishing of both Kadison--Schwarz gaps, again by the argument of
    Theorem~\ref{thm:fixedPointsStarSubalgebra}, places $E(X)$ in the
    multiplicative domain.
\end{proof}

\begin{theorem}[Choi--Effros left absorption]
    \label{thm:choi_effros_left}
    \lean{Kraus.choi_effros_left}\leanok
    \uses{def:kraus_fixed_points, def:kraus_map_channels}
    Under the same hypotheses, for all $X, Y \in \MN{D}$,
    \[
        E(E(X) Y) = E(E(X) E(Y)).
    \]
\end{theorem}

\begin{proof}\leanok
    \uses{thm:map_mem_multiplicativeDomain_of_idempotent}
    The previous theorem places $E(X)$ in the multiplicative domain, so
    $E(E(X) Y) = E(E(X)) E(Y) = E(X) E(Y)$.
    Applying the same multiplicativity to $E(X)$ and $E(Y)$ gives
    $E(E(X) E(Y)) = E(E(X)) E(E(Y))$, and idempotence rewrites this as
    $E(X) E(Y)$.
\end{proof}

\begin{theorem}[Choi--Effros right absorption]
    \label{thm:choi_effros_right}
    \lean{Kraus.choi_effros_right}\leanok
    \uses{def:kraus_fixed_points, def:kraus_map_channels}
    Under the same hypotheses, for all $X, Y \in \MN{D}$,
    \[
        E(X E(Y)) = E(E(X) E(Y)).
    \]
\end{theorem}

\begin{proof}\leanok
    \uses{thm:map_mem_multiplicativeDomain_of_idempotent}
    The previous theorem places $E(Y)$ in the multiplicative domain, so
    $E(X E(Y)) = E(X) E(E(Y)) = E(X) E(Y)$.
    Applying the same multiplicativity to $E(X)$ and $E(Y)$ gives
    $E(E(X) E(Y)) = E(E(X)) E(E(Y))$, and idempotence rewrites this as
    $E(X) E(Y)$.
\end{proof}

\begin{theorem}[Choi--Effros symmetric absorption]
    \label{thm:choi_effros_left_eq_right}
    \lean{Kraus.choi_effros_left_eq_right}\leanok
    \uses{def:kraus_fixed_points, def:kraus_map_channels}
    Under the same hypotheses, for all $X, Y \in \MN{D}$,
    \[
        E(E(X) Y) = E(X E(Y)).
    \]
\end{theorem}

\begin{proof}\leanok
    \uses{thm:choi_effros_left, thm:choi_effros_right}
    Both sides equal $E(E(X) E(Y))$ by the two preceding theorems.
\end{proof}

\begin{theorem}[Choi--Effros projected product]
    \label{thm:choi_effros_sandwich}
    \lean{Kraus.choi_effros_sandwich}\leanok
    \uses{def:kraus_fixed_points, def:kraus_map_channels}
    Under the same hypotheses, for all $X, Y \in \MN{D}$,
    \[
        E(E(X) E(Y)) = E(X) E(Y).
    \]
\end{theorem}

\begin{proof}\leanok
    \uses{thm:map_mem_multiplicativeDomain_of_idempotent}
    The previous theorem places $E(X)$ and $E(Y)$ in the multiplicative domain.
    Therefore multiplicativity gives
    \[
        E(E(X) E(Y)) = E(E(X)) E(E(Y)).
    \]
    Since $E$ is idempotent under the standing hypotheses, $E(E(X)) = E(X)$
    and $E(E(Y)) = E(Y)$, so
    \[
        E(E(X) E(Y)) = E(X) E(Y).
    \]
\end{proof}

\begin{definition}[Adjoint fixed points]\label{def:adjoint_fixed_points}
    \lean{Kraus.adjointFixedPoints}\leanok
    \uses{def:kraus_map_channels}
    The \emph{adjoint fixed-point set}
    $\mathrm{Fix}(E^*) = \{X \in \MN{D} : E^*(X) = X\}$,
    where $E^*(X) = \sum_i K_i^\dagger X K_i$.
\end{definition}

\begin{theorem}[Fixed points form a $*$-subalgebra in the Heisenberg picture]
    \label{thm:adjointFixedPointsStarSubalgebra}
    \lean{Kraus.adjointFixedPointsStarSubalgebra, Kraus.fixedPoints_starSubalgebra}\leanok
    \uses{def:adjoint_fixed_points, def:kraus_map_channels}
    If the Kraus map is trace-preserving and has a positive definite
    fixed point $\rho > 0$ with $E(\rho) = \rho$,
    then $\mathrm{Fix}(E^*)$ is a $*$-subalgebra of $\MN{D}$.
\end{theorem}

\begin{proof}\leanok
    \uses{thm:fixedPointsStarSubalgebra}
    Apply Theorem~\ref{thm:fixedPointsStarSubalgebra} to the adjoint
    family $\{K_i^\dagger\}$.
\end{proof}

\begin{definition}[Kraus commutant]\label{def:kraus_commutant}
    \lean{Kraus.krausCommutant, Kraus.krausCommutantStarSubalgebra}\leanok
    \uses{def:kraus_map_channels}
    The \emph{Kraus commutant} is
    $\{K_i, K_i^\dagger\}' = \{X \in \MN{D} :
        [X, K_i] = [X, K_i^\dagger] = 0 \forall i\}$.
    It is a $*$-subalgebra of $\MN{D}$.
\end{definition}

\begin{theorem}[Adjoint fixed points equal the Kraus commutant]
    \label{thm:adjointFixedPoints_eq_krausCommutant}
    \lean{Kraus.adjointFixedPointsStarSubalgebra_eq_krausCommutantStarSubalgebra}\leanok
    \uses{def:adjoint_fixed_points, def:kraus_commutant, def:kraus_map_channels}
    Under the hypotheses of Theorem~\ref{thm:adjointFixedPointsStarSubalgebra},
    the adjoint fixed-point $*$-subalgebra equals the Kraus commutant:
    \[
        \mathrm{Fix}(E^*) = \{K_i, K_i^\dagger\}'.
    \]
    This is \cite[Theorem~6.13]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}\leanok
    \uses{thm:adjointFixedPointsStarSubalgebra}
    ($\supseteq$): if $X$ commutes with all $K_i$ and $K_i^\dagger$,
    then $E^*(X) = \sum_i K_i^\dagger X K_i = X \sum_i K_i^\dagger K_i = X$
    by trace preservation.
    ($\subseteq$): any $*$-subalgebra inside $\mathrm{Fix}(E^*)$
    lies in the Kraus commutant, because for $X \in \mathrm{Fix}(E^*)$
    with $X^\dagger X \in \mathrm{Fix}(E^*)$ the Kadison--Schwarz
    gap vanishes, which forces $X K_i = K_i X$ and
    $X K_i^\dagger = K_i^\dagger X$.
\end{proof}

\begin{theorem}[Kraus commutant inside the adjoint fixed points]
    \label{thm:kraus_commutant_greatest}
    \lean{Kraus.krausCommutantStarSubalgebra_isGreatest_adjointFixedPointStarSubalgebras}
    \leanok
    \uses{thm:adjointFixedPoints_eq_krausCommutant}
    Under the hypotheses above, the Kraus commutant is the largest
    $*$-subalgebra contained in the adjoint fixed-point set.
\end{theorem}

\begin{proof}\leanok
    By Theorem~\ref{thm:adjointFixedPoints_eq_krausCommutant}, every
    $*$-subalgebra inside the adjoint fixed points is automatically contained in
    the Kraus commutant, while the Kraus commutant itself is such a
    $*$-subalgebra.
\end{proof}
