\chapter{Perron--Frobenius Theory for Channels and Transfer Maps}\label{ch:qpf}

This chapter establishes the Perron--Frobenius theory used in subsequent
chapters: existence, positive definiteness, and uniqueness of
fixed points for injective and irreducible transfer maps, canonical
gauge constructions, ergodicity of irreducible channels, the
spectral-radius identification of Perron eigenvalues, and Wolf's
spectral characterization of irreducibility.
The conceptual source is Wolf's treatment of irreducible positive
maps~\cite[Ch.~6, especially Thms.~6.2--6.5 and Cor.~6.3]{Wolf2012Quantum}
and~\cite{Evans1978Spectral}.

\section{Positive definiteness}

\begin{remark}\leanok
	Wolf's general theory treats irreducible positive maps first
	(\cite[Thm.~6.2]{Wolf2012Quantum}), then derives non-degeneracy
	and positive definiteness (\cite[Thm.~6.3]{Wolf2012Quantum}).
	The injective-tensor version (Theorem~\ref{thm:psd_fp_pd}) uses
	a shorter direct kernel argument; the irreducible
	CP-map version (Theorem~\ref{thm:psd_fp_pd_irred}) follows
	via the support-projection argument of~\cite[Thm.~6.2]{Wolf2012Quantum}.
\end{remark}

\begin{theorem}[PSD fixed point is positive definite]\label{thm:psd_fp_pd}
	\lean{MPSTensor.posSemidef_fixedPoint_isPosDef}
	\leanok
	\uses{def:injective, def:transfer_map}
	Let $A$ be an injective MPS tensor and let $\rho \ge 0$,
	$\rho \neq 0$, be a fixed point of the transfer map
	$\E_A(\rho) = \rho$.
	Then $\rho$ is positive definite.
\end{theorem}

\begin{proof}
	\leanok
	Suppose $v^\dagger \rho v = 0$ for some $v \neq 0$.
	Since $\rho \ge 0$, this implies $\rho v = 0$.
	The fixed-point equation gives
	$v^\dagger \rho v = \sum_i ((A^i)^\dagger v)^\dagger
		\rho ((A^i)^\dagger v)$.
	Each term is nonneg; if the sum is zero, then
	$\rho (A^i)^\dagger v = 0$ for all~$i$:
	the kernel of $\rho$ is invariant under all $(A^i)^\dagger$.
	Since the $\{A^i\}$ span $\MN{D}$, we have $\rho M v = 0$
	for every matrix $M$. Given any vector $w$, choose $M$ with
	$M v = w$. Then $\rho w = 0$ for every $w$, so $\rho = 0$,
	contradicting $\rho \neq 0$.
	In~\cite[Thm.~6.3, item~2]{Wolf2012Quantum},
	the same conclusion is reached via the
	growth condition $(\Id + T)^{d-1}(X) > 0$
	of~\cite[Thm.~6.2, condition~(2)]{Wolf2012Quantum};
	here the direct kernel argument is shorter under
	injectivity.
\end{proof}

\begin{theorem}[PSD fixed point is positive definite --- irreducible version]\label{thm:psd_fp_pd_irred}
	\lean{MPSTensor.posSemidef_fixedPoint_isPosDef_of_irreducible}
	\leanok
	\uses{def:irreducible_cp, def:transfer_map}
	If $\E_A$ is irreducible and $\rho \ge 0$, $\rho \neq 0$
	is a fixed point, then $\rho$ is positive definite.
\end{theorem}

\begin{proof}
	\leanok
	\uses{def:irreducible_cp}
	Suppose $\rho$ is not positive definite.
	Then $\rho$ has a nontrivial kernel; let $Q$ be the
	orthogonal projection onto the support of $\rho$
	(the complement of the kernel).
	A direct computation using the fixed-point equation
	$\E_A(\rho) = \rho$ shows that $Q$ is an invariant
	projection for $\E_A$: it satisfies
	$Q \, \E_A(Q X Q) \, Q = \E_A(Q X Q)$ for all~$X$.
	By irreducibility of $\E_A$
	(Definition~\ref{def:irreducible_cp}),
	every invariant projection is $0$ or $\Id$.
	If $Q = 0$ then $\rho = 0$, contradicting $\rho \neq 0$.
	If $Q = \Id$ then $\rho$ is positive definite,
	contradicting our assumption.
	Hence $\rho$ must be positive definite.
	This is the support-projection direction
	of~\cite[Thm.~6.2, condition~(1)]{Wolf2012Quantum}.
\end{proof}

\begin{theorem}[Orthogonal trace condition]\label{thm:orthogonal_trace_cond}
	\lean{orthogonal_trace_pos_of_irreducible_cp}
	\leanok
	\uses{def:irreducible_cp}
	Let $E$ be an irreducible completely positive map on $\MN{D}$,
	and let $A, B \ge 0$ be nonzero positive semidefinite matrices
	with $\tr(B A) = 0$.
	Then there exists $t$ with $1 \le t \le D - 1$ such that
	\[
		\tr\bigl(B \cdot E^t(A)\bigr) > 0.
	\]
\end{theorem}

\begin{proof}
	\leanok
	The growth theorem for irreducible CP maps
	(\cite[Thm.~6.2, item~2]{Wolf2012Quantum}) gives
	$(\Id + E)^{D-1}(A) > 0$ (positive definite).
	Since $B \ge 0$ is nonzero,
	$\tr\bigl(B \cdot (\Id + E)^{D-1}(A)\bigr) > 0$.
	Expanding by the binomial theorem:
	\[
		\tr\Bigl(B \cdot \sum_{k=0}^{D-1} \tbinom{D-1}{k} E^k(A)\Bigr) > 0.
	\]
	Each term $\tr(B \cdot E^k(A)) \ge 0$ since both factors are PSD.
	The $k = 0$ term vanishes by hypothesis
	($\tr(B A) = 0$).
	Hence at least one $k \in \{1, \ldots, D-1\}$ contributes
	a strictly positive term.
	This is~\cite[Thm.~6.2, item~4]{Wolf2012Quantum}:
	item~(2) (growth condition) implies item~(4)
	by a binomial expansion and PSD positivity argument.
\end{proof}

\section{Uniqueness}

\begin{theorem}[Uniqueness of PSD fixed point]\label{thm:psd_fp_unique}
	\lean{MPSTensor.posSemidef_fixedPoint_unique}
	\leanok
	\uses{def:injective, def:transfer_map}
	Let $A$ be an injective MPS tensor.
	If $\rho, \sigma \ge 0$ are both nonzero fixed points of
	$\E_A$, then $\sigma = c \rho$ for some $c \in \C$.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:psd_fp_pd}
	Both $\rho$ and $\sigma$ are positive definite by
	Theorem~\ref{thm:psd_fp_pd}.
	Write $\rho = S S^\dagger$ via a spectral square root and set
	$H = (S^\dagger)^{-1} \sigma S^{-1}$, which is Hermitian
	and positive definite.
	Let $c_0$ be the minimal eigenvalue of~$H$.
	Then $\tau := \sigma - c_0 \rho$ is a PSD fixed point
	which is \emph{not} positive definite
	(it has a zero eigenvalue by construction of~$c_0$).
	By Theorem~\ref{thm:psd_fp_pd}, $\tau$ must be zero,
	giving $\sigma = c_0 \rho$.
	\smallskip\noindent\textbf{Relation to Wolf:}
	In~\cite[Thm.~6.3, item~2]{Wolf2012Quantum}, uniqueness
	is proved via the growth condition
	of~\cite[Thm.~6.2]{Wolf2012Quantum}.
	The same minimal-eigenvalue argument applies here, using the
	injective-tensor positive-definiteness
	(Theorem~\ref{thm:psd_fp_pd}) in place of the
	growth condition.
\end{proof}

\begin{remark}\leanok
	The proof produces a positive real scalar $c_0 > 0$:
	it is the minimal eigenvalue of the positive definite Hermitian
	matrix $H = (S^\dagger)^{-1} \sigma S^{-1}$.
\end{remark}

\begin{theorem}[Uniqueness of positive eigenvalue for irreducible CP maps]%
	\label{thm:eigenvalue_unique_irred}
	\lean{eigenvalue_unique_of_irreducible_cp}
	\leanok
	\uses{def:irreducible_cp}
	Let $D \ge 1$ and let $E$ be an irreducible completely positive map
	on $\MN{D}$.
	Suppose $\rho, \sigma \ge 0$ are nonzero positive semidefinite matrices
	satisfying $E(\rho) = r_1 \rho$ and $E(\sigma) = r_2 \sigma$
	for real scalars $r_1, r_2 > 0$.
	Then $r_1 = r_2$.
\end{theorem}

\begin{proof}
	\leanok
		\uses{thm:pf_eigenvector_existence, thm:psd_fp_pd_irred}
	Extract Kraus operators $K$ for $E$.
	Irreducibility of $E$ passes to the adjoint transfer map
	$E^\dagger(\,\cdot\,) = \sum_i K_i^\dagger\,(\,\cdot\,)\,K_i$,
	which is a positive CP map.
	The Perron--Frobenius existence theorem
	(Theorem~\ref{thm:pf_eigenvector_existence}) together with the
	irreducible positive-definiteness upgrade
	(Theorem~\ref{thm:psd_fp_pd_irred}) gives a positive definite
	eigenvector $\tau > 0$ with eigenvalue $t > 0$:
	$E^\dagger(\tau) = t\,\tau$.
	For any nonzero PSD $X$ with $E(X) = s X$,
	the trace pairing identity yields
	\[
		s \cdot \tr(\tau X)
		  = \tr\bigl(\tau \cdot E(X)\bigr)
		  = \tr\bigl(E^\dagger(\tau) \cdot X\bigr)
		  = t \cdot \tr(\tau X).
	\]
	Since $\tau > 0$ and $X \ge 0$, $X \neq 0$,
	the scalar $\tr(\tau X) > 0$, so $s = t$.
	Applying this to both $(\rho, r_1)$ and $(\sigma, r_2)$
	gives $r_1 = t = r_2$.
	\smallskip\noindent\textbf{Relation to Wolf:}
	This is~\cite[Thm.~6.3, item~3]{Wolf2012Quantum},
	establishing non-degeneracy of the spectral-radius eigenvalue
	for irreducible CP maps. The proof follows Wolf's dual-map
	trace argument.
\end{proof}

\section{Existence and the Perron--Frobenius theorem}

\begin{theorem}[Existence of PSD fixed point]\label{thm:psd_fp_exists}
	\lean{MPSTensor.exists_posSemidef_fixedPoint}
	\leanok
	\uses{def:transfer_map}
	Assume $D \ge 1$. Let $A$ be an MPS tensor with
	$\sum_i (A^i)^\dagger A^i = \Id$ (so that $\E_A$ is
	trace-preserving).
	Then there exists $\rho \ge 0$, $\rho \neq 0$,
	with $\E_A(\rho) = \rho$.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:transfer_channel, thm:cesaro_fixedpoint}
	Under the normalization $\sum_i (A^i)^\dagger A^i = \Id$,
	the transfer map is a quantum channel
	(Theorem~\ref{thm:transfer_channel}).
	The Ces\`aro fixed-point theorem
	(Theorem~\ref{thm:cesaro_fixedpoint}) then provides the
	fixed point; injectivity is not needed for this step.
\end{proof}

\begin{theorem}[Quantum Perron--Frobenius]\label{thm:qpf}
	\lean{MPSTensor.quantum_perron_frobenius}
	\leanok
	\uses{def:injective, def:transfer_map}
	Assume $D \ge 1$. Let $A$ be an injective MPS tensor with
	$\sum_i (A^i)^\dagger A^i = \Id$.
	Then the transfer map $\E_A$ has a unique PSD fixed point
	$\rho$ (up to scaling), and $\rho$ is positive definite.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:psd_fp_exists, thm:psd_fp_pd, thm:psd_fp_unique}
	Combine Theorems~\ref{thm:psd_fp_exists},
	\ref{thm:psd_fp_pd}, and~\ref{thm:psd_fp_unique}.
\end{proof}

\section{Right- and left-canonical gauges}

\begin{theorem}[Right-canonical (unital) gauge]\label{thm:right_canonical_gauge}
	\lean{gauged_unital}
	\leanok
	Let $\E_A(\rho) = \rho$ and let $S$ be an invertible matrix with
	$S S^\dagger = \rho$.
	Define the gauged operators $A'^i = S^{-1} A^i S$.
	Then $\sum_i A'^i (A'^i)^\dagger = \Id$, i.e.\ the gauged
	Kraus map is unital. This is the right-canonical normalization
	used in subsequent chapters.
\end{theorem}

\begin{proof}
	\leanok
	Expand each summand:
	$(S^{-1} A^i S)(S^{-1} A^i S)^\dagger
		= S^{-1} A^i \underbrace{S S^\dagger}_{\rho} (A^i)^\dagger (S^\dagger)^{-1}$.
	Summing over $i$ and using $\sum_i A^i \rho (A^i)^\dagger = \rho$
	gives $S^{-1} \rho (S^\dagger)^{-1}
		= S^{-1} S S^\dagger (S^\dagger)^{-1} = \Id$.
\end{proof}

\begin{theorem}[Left-canonical (trace-preserving) gauge]\label{thm:left_canonical_gauge}
	\lean{gauged_tracePreserving}
	\leanok
	Let $\sigma$ be a fixed point of the adjoint transfer map
	$\sum_i (A^i)^\dagger \sigma A^i = \sigma$,
	and let $S$ be invertible with $S^\dagger S = \sigma$.
	Define $A'^i = S A^i S^{-1}$.
	Then $\sum_i (A'^i)^\dagger A'^i = \Id$, i.e.\ the gauged
	Kraus map is trace-preserving. This is the left-canonical
	normalization used in subsequent chapters.
\end{theorem}

\begin{proof}
	\leanok
	Expand each summand:
	$(S A^i S^{-1})^\dagger (S A^i S^{-1})
		= (S^\dagger)^{-1} (A^i)^\dagger \underbrace{S^\dagger S}_{\sigma} A^i S^{-1}$.
	Summing over $i$ and using $\sum_i (A^i)^\dagger \sigma A^i = \sigma$
	gives $(S^\dagger)^{-1} \sigma S^{-1}
		= (S^\dagger)^{-1} S^\dagger S S^{-1} = \Id$.
\end{proof}

\begin{remark}\leanok
	The right-canonical gauge of Theorem~\ref{thm:right_canonical_gauge}
	and the left-canonical gauge of Theorem~\ref{thm:left_canonical_gauge}
	are generally different similarity transforms: one is built from a fixed point
	of $\E_A$, the other from a fixed point of the adjoint transfer map.
	This section does not claim that a single gauge makes the transfer map
	simultaneously unital and trace-preserving.
\end{remark}

\section{Similarity preserves irreducibility}

\begin{lemma}[Similarity transform preserves irreducibility]%
	\label{lem:similarity_irred}
	\lean{isIrreducibleMap_full_similarity}
	\leanok
	\uses{def:irreducible_cp}
	Let $E$ be an irreducible completely positive map on $\MN{D}$,
	let $C \in \MN{D}$ be invertible ($\det C \neq 0$), and let
	$c > 0$.
	Define the similarity-transformed map
	\[
		E'(X) \;=\; c \cdot C^{-1}\, E\!\bigl(C\,X\,C^\dagger\bigr)\,(C^\dagger)^{-1}.
	\]
	Then $E'$ is also irreducible.
\end{lemma}

\begin{proof}
	\leanok
	\uses{def:irreducible_cp}
	Suppose $Q$ is an invariant projection for $E'$:
	$Q \neq 0$, $Q \neq \Id$, and
	$({\Id} - Q) E'(Q X Q) Q = 0$ for all~$X$.
	Setting $R = C Q C^\dagger$, the support projection~$P$ of $R$
	is an invariant projection for $E$.
	Irreducibility of $E$ forces $P = 0$ or $P = \Id$.
	The invertibility of $C$ then forces $Q = 0$ or $Q = \Id$,
	a contradiction.
	This is the full similarity version of
	\cite[Prop.~6.6]{Wolf2012Quantum};
	the scalar case ($C = \Id$) gives the scaling result stated next.
\end{proof}

\begin{theorem}[Scaling preserves irreducibility]%
	\label{thm:irred_map_smul}
	\lean{isIrreducibleMap_smul}
	\leanok
	\uses{def:irreducible_cp}
	If $E$ is an irreducible map and $c \neq 0$, then $c E$
	is also irreducible.
	This is the scalar special case ($C = \Id$) of
	\cite[Prop.~6.6]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}\leanok
	An invariant projection for $cE$ is an invariant projection
	for $E$ (scale out the nonzero constant from the commutation
	relation).
\end{proof}

\section{Perron--Frobenius eigenvector existence}%
\label{sec:pf_eigenvector}

This section establishes the existence of a positive definite
eigenvector for the adjoint transfer map of an irreducible MPS
tensor, and uses it to construct a TP-gauge normalization.
The PSD-eigenvector step corresponds to
\cite[Thm.~6.5]{Wolf2012Quantum} (general positive maps) and
\cite{Evans1978Spectral}; the upgrade to positive definiteness
uses the irreducible-map theory of
\cite[Thm.~6.3]{Wolf2012Quantum}. The TP-gauge application
follows~\cite[Appendix~A]{Cirac2017MPDO_AnnPhys}.

\begin{theorem}[PSD eigenvector for positive maps]%
	\label{thm:pf_eigenvector_existence}
	\lean{exists_posSemidef_eigenvector}
	\leanok
	\uses{def:positive_map}
	Let $E : \MN{D} \to \MN{D}$ be a positive linear map with
	$D > 0$, and assume that $E(\sigma) \neq 0$ for every nonzero
	PSD matrix $\sigma$.
	Then there exist a nonzero PSD matrix $\rho$ and a positive
	real $r > 0$ such that $E(\rho) = r \rho$.

	This is the finite-dimensional Perron--Frobenius theorem for
	positive maps on matrix algebras
	(\cite[Thm.~6.5]{Wolf2012Quantum}; see also~\cite{Evans1978Spectral}).
	The proof applies Brouwer's fixed-point theorem to the
	normalization map $\rho \mapsto E(\rho) / \tr(E(\rho))$ on the
	compact convex set of density matrices.
\end{theorem}

\begin{theorem}[Adjoint transfer map is nonzero]%
	\label{thm:adjoint_transfer_ne_zero}
	\lean{MPSTensor.adjointTransferMap_ne_zero_of_nonzero}
	\leanok
	\uses{def:transfer_map}
	If some Kraus operator $A^i \neq 0$, then the adjoint transfer
	map $\E_A^\dagger(X) = \sum_i (A^i)^\dagger X A^i$ is nonzero.
\end{theorem}

\begin{proof}\leanok
	If $\E_A^\dagger = 0$ then $\E_A^\dagger(\Id) = 0$,
	so $\sum_i (A^i)^\dagger A^i = 0$.
	Since each summand $(A^i)^\dagger A^i$ is PSD,
	each $(A^i)^\dagger = 0$ and hence each $A^i = 0$.
\end{proof}

\begin{theorem}[Positive-definite adjoint eigenvector for irreducible tensors]%
	\label{thm:posDef_adjoint_eigenvector}
	\lean{MPSTensor.exists_posDef_adjoint_eigenvector}
	\leanok
	\uses{def:irreducible_tensor, def:transfer_map}
	Let $A$ be an irreducible MPS tensor with $D > 0$
	and some $A^i \neq 0$.
	Then there exist a positive definite matrix~$\sigma$
	and a positive real $r > 0$ such that
	$\E_A^\dagger(\sigma) = r \sigma$.

	This combines the general PF existence
	(\cite[Thm.~6.5]{Wolf2012Quantum}) with the
	irreducible upgrade to positive definiteness
	(\cite[Thm.~6.3, item~2]{Wolf2012Quantum});
	the application to the adjoint transfer map follows
	\cite[Appendix~A]{Cirac2017MPDO_AnnPhys}.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:pf_eigenvector_existence, thm:irred_map_smul,
		thm:adjoint_transfer_ne_zero, thm:psd_fp_pd_irred}
	By Theorem~\ref{thm:adjoint_transfer_ne_zero},
	$\E_A^\dagger \neq 0$.
	Since $\E_A^\dagger(X) = \sum_i (A^i)^\dagger X A^i$ is a
	Kraus map, it is positive. Irreducibility of $A$ transfers to the
	adjoint transfer map. More precisely: if $P$ is invariant for
	$\E_A$ (i.e.\ $(\Id - P)K_i P = 0$ for all Kraus operators $K_i$),
	then taking adjoints gives $P K_i^\dagger (\Id - P) = 0$, which
	means $\Id - P$ is invariant for $\E_A^\dagger$. Hence $\E_A$
	irreducible $\Longleftrightarrow$ $\E_A^\dagger$ irreducible, and
	together with $\E_A^\dagger \neq 0$ this implies
	$\E_A^\dagger(\tau) \neq 0$ for every nonzero PSD matrix $\tau$.
	Hence Theorem~\ref{thm:pf_eigenvector_existence} applies to
	$\E_A^\dagger$, giving $\sigma_0 \ge 0$ (PSD, nonzero)
	and $r > 0$ with $\E_A^\dagger(\sigma_0) = r \sigma_0$.
	Rescale: set $T^i = r^{-1/2} (A^i)^\dagger$.
	Then $\E_T(\sigma_0) = \sigma_0$,
	and $\E_T$ is irreducible
	(by Theorem~\ref{thm:irred_map_smul},
	since irreducibility of $\E_A$ transfers to the
	adjoint and is preserved under scaling).
	Since $\E_T$ is irreducible with a PSD fixed point,
	the positive-definiteness upgrade
	(Theorem~\ref{thm:psd_fp_pd_irred})
	gives $\sigma_0$ positive definite.
\end{proof}

\begin{theorem}[TP-gauge existence for irreducible tensors]%
	\label{thm:tp_data_irreducible}
	\lean{MPSTensor.exists_tp_data_of_irreducible}
	\leanok
	\uses{def:irreducible_tensor}
	Let $A$ be an irreducible MPS tensor with $D > 0$
	and some $A^i \neq 0$.
	Then there exist a positive real~$r$, a positive definite
	matrix~$\sigma$, and a gauge-equivalent tensor~$B$ such that
	\[
		B^i = \sigma^{1/2} \cdot (r^{-1/2} A^i) \cdot \sigma^{-1/2}
	\]
	is trace-preserving:
	$\sum_i (B^i)^\dagger B^i = \Id$.

	This is the ``spectral rescaling $+$ TP gauge'' step of
	\cite[Appendix~A]{Cirac2017MPDO_AnnPhys}.
	The similarity is generally non-unitary.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:posDef_adjoint_eigenvector, thm:left_canonical_gauge}
	By Theorem~\ref{thm:posDef_adjoint_eigenvector},
	obtain $\sigma$ positive definite and $r > 0$ with
	$\E_A^\dagger(\sigma) = r \sigma$.
	Set $c = r^{-1/2}$ and $A'^i = c A^i$.
	Then $\E_{A'}^\dagger(\sigma) = \sigma$.
	By the TP gauge construction
	(Theorem~\ref{thm:left_canonical_gauge}),
	$B^i = \sigma^{1/2} A'^i \sigma^{-1/2}$
	satisfies $\sum_i (B^i)^\dagger B^i = \Id$.
\end{proof}

\section{Exponential positivity for irreducible CP maps}

\begin{theorem}[Exponential truncation is positive definite]
	\label{thm:exp_truncation_pd_irred}
	\lean{exp_truncation_posDef_of_irreducible_cp}
	\leanok
	\uses{def:irreducible_cp}
	Let $E$ be an irreducible completely positive map on $\MN{D}$,
	let $A \ge 0$ be nonzero, and let $t > 0$.
	Then the finite exponential truncation
	\[
		\sum_{k=0}^{D-1} \frac{t^k}{k!} E^k(A)
	\]
	is positive definite.
	This is the finite-sum core of \cite[Thm.~6.2, item~(3)]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}
	\leanok
	If a nonzero vector $v$ annihilated this quadratic form, then every term
	$E^k(A)$ with $0 \le k \le D-1$ would annihilate $v$.
	The growth theorem for irreducible CP maps
	(that is, $(\Id + E)^{D-1}(A) > 0$ for irreducible $E$ and nonzero
	PSD $A$, as established in the proof of
	Theorem~\ref{thm:orthogonal_trace_cond}) then forces
	$(\Id + E)^{D-1}(A)$ to annihilate $v$, contradicting its positive
	definiteness.
\end{proof}

\begin{theorem}[Exponential positivity for irreducible CP maps]
	\label{thm:exp_pd_irred}
	\lean{exp_posDef_of_irreducible_cp}
	\leanok
	\uses{def:irreducible_cp, thm:exp_truncation_pd_irred}
	Let $E$ be an irreducible completely positive map on $\MN{D}$,
	let $A \ge 0$ be nonzero, and let $t > 0$.
	Then
	\[
		\exp(tE)(A) = \sum_{k=0}^{\infty} \frac{t^k}{k!} E^k(A)
	\]
	is positive definite.
	This is \cite[Thm.~6.2, item~(3)]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}
	\leanok
	By Theorem~\ref{thm:exp_truncation_pd_irred}, the first $D$ terms already
	form a positive definite matrix.
	Every remaining term is positive semidefinite, so the full exponential
	series is the sum of a positive definite matrix and a positive semidefinite
	tail.
\end{proof}

\section{Ergodicity of irreducible channels}

\begin{theorem}[Unique density-matrix fixed point for an irreducible channel]
	\label{thm:channel_density_fp_irred}
	\lean{IsChannel.exists_unique_density_fixedPoint_of_irreducible}
	\leanok
	\uses{def:channel, def:irreducible_cp}
	Let $E$ be an irreducible quantum channel on $\MN{D}$ with $D > 0$.
	Then there exists a unique density matrix $\sigma$ such that $E(\sigma)=\sigma$.
	Moreover $\sigma$ is positive definite.
	This is the qualitative form of \cite[Cor.~6.3]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:cesaro_fixedpoint}
	Every Ces\`aro mean of a density matrix stays inside the compact convex set
	of density matrices, so a subsequence converges.
	Its limit is a fixed point by the telescoping argument, positive
	definiteness follows from irreducibility, and uniqueness comes from the
	uniqueness of positive semidefinite Perron eigenvectors at eigenvalue~$1$.
\end{proof}

\begin{theorem}[Ces\`aro convergence for irreducible channels]
	\label{thm:channel_cesaro_irred}
	\lean{IsChannel.cesaroMean_tendsto_of_irreducible}
	\leanok
	\uses{def:channel, def:irreducible_cp, def:cesaro_mean,
		  thm:channel_density_fp_irred}
	Let $E$ be an irreducible quantum channel on $\MN{D}$ with $D > 0$,
	and let $\rho$ be any density matrix.
	Then the Ces\`aro means
	\[
		\frac{1}{N} \sum_{n=0}^{N-1} E^n(\rho)
	\]
	converge to the unique positive definite density-matrix fixed point of~$E$.
	This is \cite[Cor.~6.3]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}
	\leanok
	Every subsequential limit of the Ces\`aro means is again a density-matrix
	fixed point.
	By Theorem~\ref{thm:channel_density_fp_irred}, there is only one such
	fixed point, so every convergent subsequence has the same limit.
	Compactness then forces the whole sequence to converge to that limit.
\end{proof}

\section{Spectral radius at the Perron eigenvalue}

\begin{theorem}[Spectral radius is the Perron eigenvalue]
	\label{thm:spectral_radius_pf_irred}
	\lean{spectralRadius_eq_of_posDef_eigenvector_of_irreducible_cp}
	\leanok
	\uses{def:irreducible_cp}
	Let $E$ be an irreducible completely positive map on $\MN{D}$.
	Suppose $\rho > 0$ and $E(\rho) = r\rho$ with $r > 0$.
	Then the spectral radius of $E$ is $r$.
	This is \cite[Thm.~6.3, item~(4)]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}
	\leanok
	Choose a positive definite eigenvector for the adjoint map with the same
	eigenvalue, gauge by its square root, and rescale by $r^{-1}$.
	The resulting map is trace-preserving and has a positive definite fixed
	point, so $1$ is an eigenvalue. The spectral radius of a trace-preserving
	map $F$ satisfies $\rho(F) \le 1$, since $\tr(F^n(X)) = \tr(X)$ for all
	$n$ bounds the growth of iterates. Hence its spectral radius is~$1$.
	Undoing the similarity and scalar rescaling recovers the spectral radius
	of~$E$.
\end{proof}

\begin{theorem}[Real spectral-radius identity]
	\label{thm:spectral_radius_real_pf_irred}
	\lean{spectralRadius_toReal_eq_of_posDef_eigenvector_of_irreducible_cp}
	\leanok
	\uses{thm:spectral_radius_pf_irred}
	Under the hypotheses of Theorem~\ref{thm:spectral_radius_pf_irred}, the
	real-valued spectral radius of $E$ is also equal to~$r$.
\end{theorem}

\begin{proof}
	\leanok
	This is the real-valued reformulation of
	Theorem~\ref{thm:spectral_radius_pf_irred}.
\end{proof}

\section{Spectral characterization of irreducibility}

\begin{definition}[Wolf spectral properties]
	\label{def:has_spectral_properties}
	\lean{HasSpectralProperties}
	\leanok
	\uses{def:cp_map}
	A completely positive map $E$ on $\MN{D}$ has the \emph{spectral properties}
	of \cite[Thm.~6.4]{Wolf2012Quantum} if there exist a positive real number $r$,
	a positive definite right eigenvector $\rho$, and a positive definite left
	eigenvector $\sigma$ for the adjoint map such that
	$E(\rho)=r\rho$, $E^\dagger(\sigma)=r\sigma$, every positive semidefinite
	right eigenvector for eigenvalue $r$ is a scalar multiple of $\rho$, and the
	spectral radius of $E$ is equal to~$r$.
\end{definition}

\begin{theorem}[Irreducibility gives the spectral properties]
	\label{thm:has_spectral_properties_irred}
	\lean{hasSpectralProperties_of_irreducible_cp}
	\leanok
	\uses{def:has_spectral_properties, def:irreducible_cp}
	Every nonzero irreducible completely positive map on $\MN{D}$ has the
	spectral properties of Definition~\ref{def:has_spectral_properties}.
\end{theorem}

\begin{proof}
	\leanok
	Combine Perron--Frobenius existence for the map and its adjoint with the
	uniqueness theorem for positive semidefinite Perron eigenvectors and the
	spectral-radius identity of
	Theorem~\ref{thm:spectral_radius_pf_irred}.
\end{proof}

\begin{theorem}[Spectral properties imply irreducibility]
	\label{thm:irred_from_spectral_properties}
	\lean{isIrreducibleMap_of_hasSpectralProperties}
	\leanok
	\uses{def:has_spectral_properties}
	If a completely positive map on $\MN{D}$ has the spectral properties of
	Definition~\ref{def:has_spectral_properties}, then it is irreducible.
\end{theorem}

\begin{proof}
	\leanok
	Gauge by the positive definite left eigenvector and rescale by the Perron
	eigenvalue to obtain a trace-preserving map with a positive definite fixed
	point.
	A nontrivial invariant projection would then produce, by Ces\`aro averaging in
	its corner algebra, a second positive semidefinite fixed point not proportional
	to the Perron vector, contradicting the uniqueness clause in
	Definition~\ref{def:has_spectral_properties}.
\end{proof}

\begin{theorem}[Irreducibility iff Wolf's spectral properties]
	\label{thm:irred_iff_spectral_properties}
	\lean{isIrreducibleMap_iff_spectral_properties}
	\leanok
	\uses{def:has_spectral_properties, def:irreducible_cp,
		  thm:has_spectral_properties_irred, thm:irred_from_spectral_properties}
	Let $E$ be a nonzero completely positive map on $\MN{D}$.
	Then $E$ is irreducible if and only if it has the spectral properties of
	Definition~\ref{def:has_spectral_properties}.
	This is \cite[Thm.~6.4]{Wolf2012Quantum}.
\end{theorem}

\begin{proof}
	\leanok
	Combine Theorems~\ref{thm:has_spectral_properties_irred}
	and~\ref{thm:irred_from_spectral_properties}.
\end{proof}
