\chapter{Canonical Form Reduction}\label{ch:canonical}

This chapter reduces a general MPS tensor to block-diagonal data via iterated
invariant-subspace decomposition. Two reduction schemes are relevant.
The block-injective reduction aims at injective TP-normalized blocks and the
canonical form conditions of Chapter~\ref{ch:block_perm}. The normal-canonical
reduction uses a weighted family of irreducible TP-normalized blocks with
primitive transfer maps and pairwise distinct nonzero weight moduli, in the
sense of \cite{Cirac2017MPDO_AnnPhys}.
Thus the reductions recorded here are conditional
primitive-block results. They should not be read as the full
translation-invariant canonical-form theorem of
\cite[Theorem~\texttt{Th:TIcanonical}, lines~742--763]{PerezGarcia2007Matrix}.
The normalization convention is also dual to the one used in the cited theorem:
the blocks are written in the unital orientation
$\sum_i A_k^i(A_k^i)^\dagger=\Id$ for
$\E_{A_k}(X)=\sum_i A_k^iX(A_k^i)^\dagger$
\cite[Theorem~\texttt{Th:TIcanonical}, lines~754--758]{PerezGarcia2007Matrix},
whereas this chapter uses
the trace-preserving, or left-canonical, orientation
$\sum_i (A_k^i)^\dagger A_k^i=\Id$.
Passing to the conjugate-transposed Kraus family exchanges the two
orientations by Lemma~\ref{lem:transfer_conjtranspose_adjoint}.
The Perron--Frobenius eigenvector existence and TP-gauge construction
are in Chapter~\ref{ch:qpf}; this chapter uses them to produce
left-canonical normalization data and the normal canonical form data.
The reduction draws on
\cite{PerezGarcia2007Matrix}, \cite{Cirac2021Matrix},
and~\cite[Chapter~6]{Wolf2012Quantum}.

\section{Block-diagonal construction}

\begin{remark}
	See \cref{def:block_diagonal_tensor} for the formalized block-diagonal tensor construction.
\end{remark}

\begin{theorem}[Per-block single-block FT]\label{thm:multi_block_ft}
	\lean{MPSTensor.fundamentalTheorem_multiBlock_blocks}
	\leanok
	\uses{def:block_diagonal_tensor, def:injective}
	If all block tensors $A_k$ are injective and the per-block
	same-MPV condition holds ($A_k$ and $B_k$ generate the same
	MPV family for each~$k$),
	then each pair $(A_k, B_k)$ is gauge equivalent.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:ft_single}
	Apply the single-block fundamental theorem
	(Theorem~\ref{thm:ft_single}) to each block independently.
\end{proof}

\begin{theorem}[Per-block same MPV implies global same MPV]\label{thm:block_to_global_mpv}
	\lean{MPSTensor.sameMPV_toTensorFromBlocks_of_blockSameMPV}
	\leanok
	\uses{def:block_diagonal_tensor, def:same_mpv}
	If $A_k$ and $B_k$ generate the same MPV family for each
	block~$k$, then the block-diagonal tensors
	$\bigoplus_k \mu_k A_k$ and $\bigoplus_k \mu_k B_k$
	also generate the same MPV family.
\end{theorem}

\begin{proof}\leanok\uses{thm:mpv_decomposition}
	By the MPV decomposition (Theorem~\ref{thm:mpv_decomposition}),
	each MPV is $\sum_k \mu_k^N V^{(N)}(A_k)_\sigma$.
	Per-block equality of MPV coefficients gives global equality.
\end{proof}

\begin{theorem}[Global multi-block fundamental theorem]\label{thm:multi_block_global}
	\lean{MPSTensor.fundamentalTheorem_multiBlock_global}
	\leanok
	\uses{def:block_diagonal_tensor, def:injective, def:same_mpv, thm:multi_block_ft}
	If each block $A_k$ is injective and each pair $(A_k, B_k)$ generates
	the same MPV family, then the block-diagonal tensors
	$\bigoplus_k \mu_k A_k$ and $\bigoplus_k \mu_k B_k$ are gauge equivalent.
\end{theorem}

\begin{proof}\leanok\uses{thm:multi_block_ft}
	Apply the per-block single-block fundamental theorem
	(Theorem~\ref{thm:multi_block_ft}) to obtain blockwise gauge
	equivalences from injectivity and equality of MPV families, then
	combine them into a global gauge for the block-diagonal tensors.
\end{proof}

\section{Transfer map normalization}

\begin{theorem}[Scaling preserves injectivity]\label{thm:injective_smul}
	\lean{MPSTensor.isInjective_smul}
	\leanok
	\uses{def:injective}
	If $A$ is injective and $\zeta \neq 0$, then $\zeta A$
	is injective.
\end{theorem}

\begin{proof}\leanok
	Scaling by a nonzero scalar preserves linear independence
	and hence spanning.
\end{proof}

\begin{theorem}[Transfer map under scaling]\label{thm:transfer_smul}
	\lean{MPSTensor.transferMap_smul}
	\leanok
	\uses{def:transfer_map}
	$\E_{\zeta A}(X) = |\zeta|^2 \E_A(X)$.
\end{theorem}

\begin{proof}\leanok\uses{lem:eval_word_smul}
	Each summand: $(\zeta A^i) X (\zeta A^i)^\dagger
		= |\zeta|^2 A^i X (A^i)^\dagger$.
\end{proof}

\begin{theorem}[Phase-scaling preserves normalization]\label{thm:leftCanonical_phase_scaling}
	\lean{MPSTensor.leftCanonical_smul_of_norm_one}
	\leanok
	If $\sum_i (A^i)^\dagger A^i = \Id$ and $|\zeta| = 1$,
	then $\sum_i (\zeta A^i)^\dagger (\zeta A^i) = \Id$.
\end{theorem}

\begin{proof}\leanok
	$(\zeta A^i)^\dagger (\zeta A^i)
		= \bar{\zeta} \zeta (A^i)^\dagger A^i
		= (A^i)^\dagger A^i$,
	since $\bar{\zeta} \zeta = |\zeta|^2 = 1$.
\end{proof}

\begin{theorem}[MPV under block normalization]\label{thm:mpv_normalize}
	\lean{MPSTensor.mpv_toTensorFromBlocks_normalize}
	\leanok
	\uses{def:block_diagonal_tensor, def:mpv}
	For any nonzero scaling factors $\mu_k$, set
	$\eta_k := \mu_k / |\mu_k|$ and $B_k := |\mu_k| A_k$.
	Then $\bigoplus_k \mu_k A_k$ and $\bigoplus_k \eta_k B_k$
	generate the same MPV family.
	Equivalently, the $k$-th block contribution factors as
	\[
		\mu_k^N V^{(N)}(A_k)_\sigma
		= \eta_k^N |\mu_k|^N V^{(N)}(A_k)_\sigma,
	\]
	so the modulus contributes only through $|\mu_k|^N$, while the
	remaining factor $\eta_k$ is a unit-modulus phase.
\end{theorem}

\begin{proof}\leanok\uses{thm:mpv_decomposition, lem:mpv_smul}
	By the MPV decomposition
	(Theorem~\ref{thm:mpv_decomposition}),
	the $k$-th contribution of $\bigoplus_k \mu_k A_k$ is
	$\mu_k^N V^{(N)}(A_k)_\sigma$.
	For $B_k = |\mu_k| A_k$, the MPV-scaling lemma
	(Lemma~\ref{lem:mpv_smul}) gives
	$V^{(N)}(B_k)_\sigma = |\mu_k|^N V^{(N)}(A_k)_\sigma$.
	Hence
	$\eta_k^N V^{(N)}(B_k)_\sigma
	= \eta_k^N |\mu_k|^N V^{(N)}(A_k)_\sigma
	= \mu_k^N V^{(N)}(A_k)_\sigma$
	for every block, so the total MPV families agree.
\end{proof}

\begin{remark}[Why blockwise cancellation cannot be separated without normalization]
	The implication
	\[
		\mc{V}\!\bigl(\textstyle\bigoplus_{k=1}^r \mu_k A_k\bigr)
		=
		\mc{V}\!\bigl(\textstyle\bigoplus_{k=1}^r \mu_k B_k\bigr)
		\quad\Longrightarrow\quad
		\forall k,\ \mc{V}(A_k)=\mc{V}(B_k)
	\]
	is false if one assumes only that the weights $\mu_k$ are pairwise
	distinct and nonzero.
	Already for $r=2$, with $\mu=(1,2)$ and scalar $1\times 1$ blocks,
	\[
		A=(1,3/2), \qquad B=(3,1/2),
	\]
	one has
	\[
		1^N\!\left(1^N-3^N\right)+2^N\!\left((3/2)^N-(1/2)^N\right)=0
	\]
	for every $N \ge 1$, so the two direct sums generate the same MPV
	family, even though the first blocks already disagree at $N=1$.
	The reason is exactly the rescaling freedom described above:
	a nonzero scalar may be moved from the weight $\mu_k$ into the block
	tensor $A_k$, or conversely, without changing the contribution
	$\mu_k^N V^{(N)}(A_k)_\sigma$.
	Thus distinct nonzero weights alone do not identify the individual
	blocks.

	The canonical-form hypotheses remove this ambiguity.
	If each block is first placed in the normalization
	$\sum_i (A_k^i)^\dagger A_k^i = \Id$, then further rescaling is
	restricted to unit-modulus phases, and the strict ordering
	$|\mu_1| > \cdots > |\mu_r|$ used in
	\cite{PerezGarcia2007Matrix,Cirac2021Matrix}
	prevents trading a nontrivial modulus between the weight and the
	block tensor.
\end{remark}

\section{Invariant subspace decomposition}

\begin{theorem}[Unitary conjugation preserves the MPV family]\label{thm:samempv_unitary}
	\lean{MPSTensor.sameMPV_conj_unitary}
	\leanok
	\uses{def:same_mpv}
	If $U$ is a unitary matrix, then $A$ and $U^\dagger A U$
	generate the same MPV family.
\end{theorem}

\begin{proof}\leanok
	By cyclicity of the trace:
	$\tr((U^\dagger A^{i_1} U) \cdots (U^\dagger A^{i_N} U))
		= \tr(A^{i_1} \cdots A^{i_N})$.
\end{proof}

\begin{definition}[Support projection]\label{def:support_proj}
	\lean{MPSTensor.supportProj}
	\leanok
	For a positive semi-definite matrix $\rho \ge 0$,
	the \emph{support projection} $P$ is the orthogonal
	projection onto the range of $\rho$.
	It is constructed via the spectral decomposition:
	if $\rho = U \diag(\lambda_1,\ldots,\lambda_D) U^\dagger$,
	then $P = U \diag(\mathbf{1}_{\lambda_1>0},\ldots,\mathbf{1}_{\lambda_D>0}) U^\dagger$,
	where $\mathbf{1}_{\lambda_j>0}$ is $1$ if $\lambda_j > 0$
	and $0$ otherwise.
\end{definition}

\begin{definition}[Invariant projection predicate]\label{def:has_inv_proj}
	\lean{MPSTensor.HasInvariantProj}
	\leanok
	\uses{def:mps_tensor}
	An MPS tensor $A$ \emph{has an invariant projection}
	if there exists an orthogonal projection $P$ with
	$P \neq 0$, $P \neq \Id$, and
	$(\Id - P) A^i P = 0$ for all~$i$.
	A tensor is \emph{irreducible} if it has no nontrivial invariant projection
	(Definition~\ref{def:irreducible_tensor}).
\end{definition}

\begin{theorem}[PSD fixed point gives invariant projection]\label{thm:fp_inv_proj}
	\lean{MPSTensor.lowerZero_of_posSemidef_fixedPoint}
	\leanok
	\uses{def:transfer_map}
	Let $\rho \ge 0$ be a fixed point of $\E_A$,
	and let $P$ be the support projection of $\rho$.
	Then for all~$i$, $(\Id - P) A^i P = 0$:
	the range of $P$ is invariant under all $A^i$.
\end{theorem}

\begin{proof}
	\leanok
	From $\E_A(\rho) = \rho$ and positivity of each
	summand $A^i \rho (A^i)^\dagger$,
	$(\Id - P) A^i \rho (A^i)^\dagger (\Id - P) = 0$
	for each~$i$.
	Since $\rho$ is supported on the range of $P$,
	this forces $(\Id - P) A^i P = 0$.

	This is the finite-dimensional version of
	\cite[Proposition~6.10]{Wolf2012Quantum}
	(``Stationary subspaces'': the support projection
	of any stationary state is sub-harmonic, i.e.\
	invariant under the channel).
	The equivalence between irreducibility and
	absence of nontrivial invariant projections is
	\cite[Theorem~6.2]{Wolf2012Quantum};
	see also
	\cite[Section~IV.A.1]{Cirac2021Matrix}.
\end{proof}

\begin{theorem}[Two-block decomposition]\label{thm:two_block_decomp}
	\lean{MPSTensor.exists_twoBlock_decomp_of_posSemidef_fixedPoint_strict}
	\leanok
	\uses{def:transfer_map, def:same_mpv2}
	If the transfer map has a PSD fixed point $\rho$ that is
	\emph{not} positive definite (i.e.\ $\rho$ has a nontrivial
	kernel), then there exist MPS tensors $A_1, A_2$ of smaller
	bond dimensions such that the two-block tensor
	$A_1 \oplus A_2$ generates the same MPV family as~$A$:
	$\mc{V}(A) = \mc{V}(A_1 \oplus A_2)$.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:fp_inv_proj, thm:samempv_unitary}
	The support projection $P$ of $\rho$ is neither $0$ nor
	$\Id$ (since $\rho \neq 0$ and $\rho$ is not PD).
	Theorem~\ref{thm:fp_inv_proj} gives
	$(\Id - P) A^i P = 0$ for all~$i$.
	A unitary change of basis that block-diagonalizes $P$ puts
	$A^i$ in upper-triangular block form, with diagonal blocks
	$B_{11}^i$ and $B_{22}^i$.
	Theorem~\ref{thm:samempv_unitary} ensures the MPV is preserved by the
	conjugation.
	The trace of the upper-triangular product equals the sum
	of the diagonal-block traces, so the block-diagonal tensor
	obtained by discarding the off-diagonal blocks has the same MPV
	coefficients as the conjugated tensor, and hence as~$A$.
\end{proof}

\section{Iterated reduction}

\begin{definition}[Irreducible tensor]\label{def:irreducible_tensor}
	\lean{MPSTensor.IsIrreducibleTensor}
	\leanok
	\uses{def:has_inv_proj}
	An MPS tensor~$A$ is \emph{irreducible}
	if it has no nontrivial invariant projection.
\end{definition}

\begin{theorem}[Irreducible block decomposition]\label{thm:irred_decomp}
	\lean{MPSTensor.exists_irreducible_blockDecomp}
	\leanok
	\uses{def:irreducible_tensor}
	Every MPS tensor $A$ admits a block-diagonal tensor
	$\bigoplus_{k=1}^r A_k$
	where each $A_k$ is irreducible and
	$\mc{V}(A) = \mc{V}\!\bigl(\bigoplus_k A_k\bigr)$.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:two_block_decomp, thm:pf_eigenvector_existence}
	By strong induction on the bond dimension~$D$.
	If $A$ is already irreducible, take $r = 1$.
	Otherwise, by Definition~\ref{def:irreducible_tensor},
	$A$ has a nontrivial invariant projection~$P$.
	Diagonalizing~$P$ yields the same two-block upper-triangular
	splitting used in Theorem~\ref{thm:two_block_decomp}; the
	block-diagonal tensor obtained by discarding the off-diagonal part
	has the same MPV family as~$A$, and the two diagonal blocks have
	strictly smaller bond dimensions.
	Iterating this splitting produces a block-diagonal tensor
	$\bigoplus_k A_k$ whose blocks are irreducible.
	Apply the induction hypothesis to each block and concatenate.
	(If one prefers a positive-eigenvector argument without assuming TP
	normalization, the correct replacement for the channel-specific
	Ces\`aro theorem is Theorem~\ref{thm:pf_eigenvector_existence};
	the induction itself needs only the invariant projection.)

	Both \cite[Theorem~4]{PerezGarcia2007Matrix} and
	\cite[Section~IV.A.1]{Cirac2021Matrix} obtain the block decomposition by
	the same invariant-projection splitting used here.
	An alternative derivation, developed
	in~\cite[Theorem~6.14]{Wolf2012Quantum},
	characterises the fixed-point set of a quantum channel
	as a $*$-algebra and then applies the
	Wedderburn--Artin decomposition
	(\cite[Equation~(1.39)]{Wolf2012Quantum})
	to read off the block structure directly.
	The block decomposition is obtained by iterating
	the invariant-projection splitting and using strong
	induction on~$D$ to ensure termination.
\end{proof}

\begin{theorem}[Unitary diagonal positive fixed point in TP gauge]\label{thm:form_II}
	\lean{MPSTensor.exists_unitary_diag_posDef_fixedPoint_of_TP_of_isIrreducibleTensor}
	\leanok
	\uses{def:irreducible_tensor, def:tp_kraus}
	Let $A$ be an irreducible MPS tensor satisfying
	$\sum_i (A^i)^\dagger A^i = \Id$, and assume $D > 0$.
	Then there exist a unitary $U$ and a diagonal positive
	definite matrix $\Lambda$ such that, for $B^i := U^\dagger A^i U$,
	\[
		\sum_i (B^i)^\dagger B^i = \Id
		\qquad\text{and}\qquad
		\E_B(\Lambda) = \Lambda.
	\]
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:psd_fp_pd_irred}
	The TP hypothesis makes $\E_A$ a channel, so it has a nonzero
	PSD fixed point.
	The irreducible tensor condition implies that $\E_A$ is
	irreducible as a CP map, and Theorem~\ref{thm:psd_fp_pd_irred}
	upgrades the fixed point to a positive definite one.
	The spectral theorem diagonalizes this matrix by a unitary
	conjugation, and unitary conjugation preserves the TP
	normalization.
\end{proof}

The Burnside-style bridge from irreducible tensors to full algebra spanning is
developed alongside the Wielandt material in Chapter~\ref{ch:wielandt};
see in particular \cref{thm:irreducible_action_cumulative_span}.

\section{Proportional single-block theorem}

\begin{theorem}[Proportional MPV implies gauge-phase equivalence]\label{thm:proportional_ft}
	\lean{MPSTensor.gaugePhaseEquiv_of_proportionalMPV₂_of_overlap_tendsto_one}
	\leanok
	\uses{def:gauge_phase_equiv, def:injective, def:tp_kraus}
	Let $A$ and $B$ be injective MPS tensors satisfying
	$\sum_i (A^i)^\dagger A^i = \Id$ and
	$\sum_i (B^i)^\dagger B^i = \Id$.
	If $A$ and $B$ generate proportional MPV families
	($V^{(N)}(A)_\sigma = c_N V^{(N)}(B)_\sigma$
	for some $c_N \in \C$) and
	$O_{AA}(N) \to 1$, $O_{BB}(N) \to 1$,
	then $A$ and $B$ are gauge-phase equivalent.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:overlap_decay}
	By contradiction.
	If $A$ and $B$ are not gauge-phase equivalent, then
	$O_{AB}(N) \to 0$ (Theorem~\ref{thm:overlap_decay}).
	From the proportionality hypothesis,
	$O_{AB}(N) = c_N O_{BB}(N)$.
	Since $O_{BB}(N) \to 1$, it follows that $c_N \to 0$.
	But $O_{AA}(N) = |c_N|^2 O_{BB}(N) \to 0$,
	contradicting $O_{AA}(N) \to 1$.

	In~\cite{PerezGarcia2007Matrix}, the proportional case
	is handled as part of the canonical form argument.
	Here it is obtained directly from the overlap-decay theorem.
\end{proof}

\section{Canonical form from primitivity}

The canonical form conditions
(Definition~\ref{def:is_canonical_form})
include an explicit hypothesis
that the self-overlap $O_{A_k A_k}(N) \to 1$
for each block.
The first theorem below derives this from
blockwise primitive fixed-point data.
The second derives the same conclusion from
per-block peripheral-spectrum primitivity.

\begin{theorem}[Canonical form from primitive fixed-point data]%
	\label{thm:canonical_from_primitive_fixedpoint}
	\lean{MPSTensor.isCanonicalForm_of_primitive}
	\leanok
	\uses{def:is_canonical_form}
	Let $(\mu_k, A_k)_{k=1}^r$ be a collection of block tensors and
	scaling factors satisfying:
	\begin{enumerate}
		\item each $A_k$ is injective,
		\item each $A_k$ satisfies the TP normalization
		      $\sum_i (A_k^i)^\dagger A_k^i = \Id$,
		\item $|\mu_1| > \cdots > |\mu_r|$ (strictly decreasing),
		\item all $\mu_k \neq 0$,
		\item each block has positive bond dimension,
		\item each block admits a primitive fixed point.
	\end{enumerate}
	Then $(\mu_k, A_k)$ satisfies the canonical form conditions.
\end{theorem}
% Scope restriction (primitive block hypotheses): not the full source theorem
% (Theorem Th:TIcanonical, lines 742--763), which starts from an arbitrary
% translation-invariant MPS representation. See
% docs/paper-gaps/pgvwc07_ti_canonical_form_scope.tex.

\begin{proof}\leanok
	\uses{thm:primitive_overlap}
	All canonical-form clauses except the overlap condition are already
	assumptions. For each block, the primitive fixed-point hypothesis gives
	$O_{A_k A_k}(N) \to 1$.
\end{proof}

\begin{theorem}[Canonical form from peripheral primitivity]%
	\label{thm:canonical_from_primitive}
	\lean{MPSTensor.isCanonicalForm_of_peripheralPrimitive}
	\leanok
	\uses{def:is_canonical_form, def:primitive_channel}
	Let $(\mu_k, A_k)_{k=1}^r$ be a collection of block
	tensors and scaling factors satisfying:
	\begin{enumerate}
		\item each $A_k$ is injective,
		\item each $A_k$ satisfies the TP normalization
		      $\sum_i (A_k^i)^\dagger A_k^i = \Id$,
		\item $|\mu_1| > \cdots > |\mu_r|$ (strictly decreasing),
		\item all $\mu_k \neq 0$,
		\item each block has positive bond dimension,
		\item each transfer map $\E_{A_k}$ is primitive,
		      i.e. has peripheral spectrum $\{1\}$.
	\end{enumerate}
	Then $(\mu_k, A_k)$ satisfies the canonical form
	conditions (Definition~\ref{def:is_canonical_form}).
\end{theorem}
% Scope restriction (peripheral-primitive block hypotheses): not the full
% source theorem (lines 742--763), which starts from an
% arbitrary translation-invariant MPS representation. See
% docs/paper-gaps/pgvwc07_ti_canonical_form_scope.tex.

\begin{proof}
	\leanok
		\uses{thm:qpf, thm:eigenvalue_bound, thm:psd_fp_unique,
			thm:primitive_compl_lt_one, thm:primitive_overlap}
	Fix a block $k$ and write $E := \E_{A_k}$.
	We check that each primitive block already satisfies the defining
	properties of the canonical form conditions.
	By injectivity and TP normalization, Theorem~\ref{thm:qpf} provides a
	nonzero positive definite fixed point for~$E$.
	To apply the spectral-gap criterion
	(Theorem~\ref{thm:primitive_compl_lt_one}), we check its three hypotheses:
	\begin{enumerate}
		\item $E$ is primitive, since the peripheral spectrum is $\{1\}$ by assumption;
		\item every eigenvalue of $E$ has modulus at most~$1$, because
		      $E = F_{A_k A_k}$ is a normalized mixed transfer operator and
		      Theorem~\ref{thm:eigenvalue_bound} applies;
		\item any trace-zero fixed point of $E$ is zero; this is the
		      trace-zero uniqueness consequence of the PSD fixed-point
		      uniqueness theorem
		      (Theorem~\ref{thm:psd_fp_unique}).
	\end{enumerate}
	Hence $E$ has a spectral gap on the complement of its fixed-point
	projection. Theorem~\ref{thm:primitive_overlap} then gives
	$O_{A_k A_k}(N) \to 1$.
	Since this holds for every block, the overlap clause required in
	Definition~\ref{def:is_canonical_form} follows from the other hypotheses.
\end{proof}

\begin{remark}\label{rem:primitivity_reduces_assumptions}\leanok
	Theorem~\ref{thm:canonical_from_primitive} shows that
	the overlap convergence hypothesis in the canonical form
	conditions is redundant once peripheral-spectrum primitivity is known.
	The argument factors through the spectral-gap formulation of
	Chapter~\ref{ch:spectral}: peripheral-spectrum primitivity is first
	converted to that spectral-gap condition, and the overlap limit is then
	deduced from it.
	In \cite{Cirac2021Matrix}, primitivity is obtained from
	irreducibility plus periodicity removal by blocking to period one.
	Theorem~\ref{thm:blocking_primitivity} is precisely this
	periodicity-removal step under irreducibility and TP normalization.
\end{remark}

\begin{theorem}[Blocking yields a peripheral-spectrum primitive transfer map]\label{thm:blocking_primitivity}
	\lean{MPSTensor.exists_blockTensor_isPrimitive_of_TP_of_isIrreducibleTensor}
	\leanok
	\uses{def:blocked_tensor, def:transfer_map, def:primitive_channel,
		def:tp_kraus, def:irreducible_tensor}
	Let $A$ be an irreducible MPS tensor with $D > 0$.
	Assume that $A$ satisfies the TP normalization
	$\sum_i (A^i)^\dagger A^i = \Id$.
	Then there exists a blocking length $p > 0$ such that the transfer map
	$\E_{A^{[p]}}$ of the blocked tensor $A^{[p]}$ is primitive,
	i.e. its peripheral spectrum is $\{1\}$.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:peripheral_roots_irred_fp, thm:peripheral_finite,
		lem:common_power_eq_one, lem:peripheral_pow_singleton}
	The period-$p$ blocking step is the diagrammatic move
	of grouping the chain into consecutive blocks of length~$p$.
	Pass to the conjugate-transposed Kraus family $K_i := (A^i)^\dagger$.
	The TP normalization of $A$ makes $K$ unital, and irreducibility of $A$
	gives irreducibility of the associated Kraus map.
	The TP fixed-point theorem for $\E_A$ provides a positive definite fixed
	point for the adjoint map of $K$, so Theorem~\ref{thm:peripheral_roots_irred_fp}
	shows that every peripheral eigenvalue is a root of unity.
	Since $\mathrm{peripheral}(\E_A)$ is finite
	(Theorem~\ref{thm:peripheral_finite}),
	Lemma~\ref{lem:common_power_eq_one} gives a common exponent $p$
	such that $\mu^p = 1$ for all $\mu \in \mathrm{peripheral}(\E_A)$.
	Lemma~\ref{lem:peripheral_pow_singleton} then implies
	$\mathrm{peripheral}(\E_A^p) = \{1\}$.
	Finally, blocking corresponds to taking a power of the transfer map,
	so $\E_{A^{[p]}}$ is primitive.
\end{proof}

\begin{remark}\label{rem:blocking_primitivity_appendixA}\leanok
	Theorem~\ref{thm:blocking_primitivity} is the ``periodicity removal by blocking''
	step in \cite[Appendix~A]{Cirac2017MPDO_AnnPhys} for an irreducible block
	already placed in TP gauge. The later existence theorems then reuse
	this step.
\end{remark}

\subsection{Adjoint transport for blocking}

\begin{lemma}[Characteristic polynomial under conjugate transpose]%
	\label{lem:charpoly_conjtranspose}
	\lean{MPSTensor.Matrix.charpoly_conjTranspose}
	\leanok
	For a complex matrix $M$, the characteristic polynomial of $M^\dagger$
	is obtained from the characteristic polynomial of $M$ by coefficientwise
	complex conjugation.
\end{lemma}

\begin{proof}\leanok
	Write $M^\dagger$ as the transpose of the coefficientwise conjugate matrix.
	The characteristic polynomial is invariant under transpose and commutes with
	coefficientwise conjugation.
\end{proof}

\begin{theorem}[Eigenvalues of the adjoint are conjugated]%
	\label{thm:eigenvalue_adjoint_conjugate}
	\lean{MPSTensor.Module.End.hasEigenvalue_adjoint_iff}
	\leanok
	\uses{lem:charpoly_conjtranspose}
	Let $E$ be a linear endomorphism of a finite-dimensional complex inner-product
	space, and let $\mu \in \C$.
	Then $\mu$ is an eigenvalue of $E$ if and only if $\overline{\mu}$ is an
	eigenvalue of the adjoint $E^\dagger$.
\end{theorem}

\begin{proof}\leanok
	\uses{lem:charpoly_conjtranspose}
	Express eigenvalues as roots of the characteristic polynomial, convert the
	adjoint to the conjugate-transposed matrix in an orthonormal basis, and apply
	Lemma~\ref{lem:charpoly_conjtranspose}.
\end{proof}

\begin{theorem}[Primitivity is invariant under adjoint]%
	\label{thm:primitive_adjoint_equiv}
	\lean{MPSTensor.IsPrimitive.adjoint_iff}
	\leanok
	\uses{def:primitive_channel, thm:eigenvalue_adjoint_conjugate}
	A linear endomorphism is primitive if and only if its adjoint is primitive.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:eigenvalue_adjoint_conjugate}
	The peripheral eigenvalues of the adjoint are the complex conjugates of those
	of the original map, so both maps have the same peripheral spectrum on the
	unit circle.
\end{proof}

\begin{lemma}[Transfer map of the conjugate-transposed Kraus family]%
	\label{lem:transfer_conjtranspose_adjoint}
	\lean{MPSTensor.transferMap_conjTranspose_eq_adjoint}
	\leanok
	For an MPS tensor $A$, the transfer map of the conjugate-transposed Kraus
	family $\{(A^i)^\dagger\}$ is the adjoint of $\E_A$ with respect to the
	Frobenius inner product.
\end{lemma}

\begin{proof}\leanok
	Compare both sides against the Frobenius pairing
	$\langle X,Y\rangle = \tr(YX^\dagger)$ and rewrite the Kraus sums.
\end{proof}

\begin{theorem}[Blocking primitivity with a positive-dimension hypothesis]%
	\label{thm:blocking_primitivity_nezero}
	\lean{MPSTensor.exists_blockTensor_isPrimitive}
	\leanok
	\uses{def:blocked_tensor, def:primitive_channel, def:irreducible_tensor}
	Let $A$ be an irreducible MPS tensor with positive bond dimension and
	TP normalization $\sum_i (A^i)^\dagger A^i = \Id$.
	Then there exists a blocking length $p > 0$ such that $\E_{A^{[p]}}$ is
	primitive.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:blocking_primitivity}
	This is Theorem~\ref{thm:blocking_primitivity} with the positive-dimension
	assumption expressed as a typeclass hypothesis.
\end{proof}

\begin{theorem}[Blocking primitivity in left-canonical gauge]%
	\label{thm:blocking_primitivity_leftcanonical}
	\lean{MPSTensor.exists_blockTensor_isPrimitive_of_leftCanonical_of_isIrreducibleTensor}
	\leanok
	\uses{def:blocked_tensor, def:primitive_channel, def:irreducible_tensor}
	Let $A$ be a left-canonical irreducible MPS tensor with positive bond
	dimension.
	Then there exists a blocking length $p > 0$ such that $\E_{A^{[p]}}$ is
	primitive.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:blocking_primitivity}
	This is the left-canonical reformulation of
	Theorem~\ref{thm:blocking_primitivity}.
\end{proof}

\section{Normal canonical form}\label{sec:normal_canonical_form}

\begin{definition}[Normal canonical form conditions]\label{def:is_normal_canonical_form}
		\lean{MPSTensor.IsNormalCanonicalForm}\leanok
		\uses{def:irreducible_tensor, def:primitive_channel, def:tp_kraus}
		A collection of scaling factors $(\mu_k)_{k=1}^r$ and block tensors $(A_k)_{k=1}^r$
		satisfies the \emph{normal canonical form conditions} if:
		\begin{enumerate}
			\item each $A_k$ is irreducible,
			\item each $A_k$ satisfies the TP normalization
			      $\sum_i (A_k^i)^\dagger A_k^i = \Id$,
			\item each transfer map $\E_{A_k}$ is primitive
			      (equivalently, has peripheral spectrum $\{1\}$),
			\item the moduli are non-increasing:
			      $|\mu_1| \ge |\mu_2| \ge \cdots \ge |\mu_r|$,
			\item all $\mu_k \neq 0$,
			\item all bond dimensions are positive.
		\end{enumerate}
		Here ``normal'' is used in the spectral sense of
		\cite{Cirac2017MPDO_AnnPhys}: irreducible blocks whose TP-normalized
		transfer maps are primitive. As noted in the remark after
		Definition~\ref{def:normal}, this is equivalent to the eventual
		block-injectivity formulation of Definition~\ref{def:normal}; see
		\cite[Proposition~3]{Sanz2010Quantum}.
\end{definition}

\begin{theorem}[Self-overlap is derived in normal canonical form]%
		\label{thm:ncf_overlap_tendsto_one}
		\lean{MPSTensor.IsNormalCanonicalForm.overlap_tendsto_one}\leanok
		\uses{def:is_normal_canonical_form, def:mpv_overlap}
		If $(\mu_k, A_k)$ satisfies the normal canonical form conditions, then
		$O_{A_k A_k}(N) \to 1$ for every block~$k$.
\end{theorem}

\begin{proof}\leanok
		\uses{def:is_normal_canonical_form}
		Peripheral-spectrum primitivity is part of the data. Together with
		irreducibility and TP normalization, it yields the spectral-gap
		primitive theorem needed for overlap convergence, so the self-overlap
		condition follows from the other hypotheses rather than appearing as
		a separate assumption.
\end{proof}

\begin{remark}
	See \cref{thm:modulus_one_gauge_irr_tp} for the formalized modulus-one eigenvalue rigidity statement for irreducible trace-preserving blocks.
\end{remark}

\begin{lemma}[Mixed-transfer spectral radius from unit-modulus overlap]%
		\label{lem:overlap_norm_one_implies_spectral_radius_ge_one}
		\lean{MPSTensor.mixedTransferSpectralRadius_ge_one_of_mpvOverlap_norm_tendsto_one}
		\leanok
		\uses{def:mpv_overlap}
		Let $A$ and $B$ be MPS tensors of the same bond dimension.  If
		$\|\langle V^{(N)}(A)\,|\,V^{(N)}(B)\rangle\|\to 1$ as $N\to\infty$,
		then the mixed transfer map has spectral radius at least $1$.
\end{lemma}

\begin{proof}\leanok
		If the mixed transfer
		spectral radius were strictly less than~$1$, the iterated mixed transfer
		map would tend to zero in operator norm, hence so would its trace, and
		via the trace identity
		$\operatorname{tr}(F_{AB}^N) = \langle V^{(N)}(A), V^{(N)}(B)\rangle$
		the overlap itself would tend to~$0$, contradicting the hypothesis.
\end{proof}

\begin{theorem}[Gauge-phase equivalence from unit-modulus overlap]%
		\label{thm:nt_overlap_norm_one_gauge}
		\lean{MPSTensor.gaugePhaseEquiv_of_overlap_norm_tendsto_one_of_irreducible_TP}
		\leanok
		\uses{def:irreducible_tensor, def:tp_kraus, def:gauge_phase_equiv,
			def:mpv_overlap}
		Let $A$ and $B$ be irreducible MPS tensors of the same bond dimension
		satisfying $\sum_i (A^i)^\dagger A^i = \Id$ and
		$\sum_i (B^i)^\dagger B^i = \Id$.  If
		$\|\langle V^{(N)}(A)\,|\,V^{(N)}(B)\rangle\|\to 1$, then $A$ and~$B$
		are gauge-phase equivalent.
		% Same-bond-dimension case of \cite[Lemma~equalMPS,
		% lines~1080--1091]{Cirac2016MPDO_arXiv}; proportionality of MPVs is not assumed.
\end{theorem}

\begin{proof}\leanok
		\uses{lem:overlap_norm_one_implies_spectral_radius_ge_one,
			thm:modulus_one_gauge_irr_tp}
		By Lemma~\ref{lem:overlap_norm_one_implies_spectral_radius_ge_one},
		the mixed transfer spectral radius is at least~$1$.
		Theorem~\ref{thm:modulus_one_gauge_irr_tp} then gives gauge-phase
		equivalence.
\end{proof}

\section{Steps toward canonical form existence}%
\label{sec:canonical_pipeline_1606}

This section combines the components of the canonical-form
construction following \cite[Section~2.3 and Appendix~A]{Cirac2017MPDO_AnnPhys}.
The approach here requires blocking to remove non-trivial periodicity.
A separate reduction for periodic blocks without blocking
is developed in \cite{DeLasCuevas2017Irreducible}; see
Chapter~\ref{ch:periodic_ft_apps} for that direct periodic extension.


\begin{theorem}[Left-canonical normalization for irreducible TP blocks]%
	\label{thm:CFII_data}
	\lean{MPSTensor.exists_CFII_data_of_TP_of_isIrreducibleTensor}
	\leanok
	\uses{def:irreducible_tensor}
	Let $A$ be an irreducible MPS tensor already in TP
	gauge ($\sum_i (A^i)^\dagger A^i = \Id$) with $D > 0$.
	Then there exist a unitary~$U$ and a diagonal positive
	definite matrix~$\Lambda$ such that:
	\begin{enumerate}
		\item $B^i := U^\dagger A^i U$ is TP,
		\item $\E_B(\Lambda) = \Lambda$,
		\item $\Lambda$ is diagonal and positive definite,
		\item $A$ and $B$ generate the same MPV family.
	\end{enumerate}
	This is the left-canonical normalization step (Canonical Form~II)
	of \cite[Appendix~A]{Cirac2017MPDO_AnnPhys}. The unitary
	conjugation occurs only after TP normalization has already been
	achieved; it is a second step inside TP gauge, not a unitary
	normalization of an arbitrary irreducible block.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:form_II, thm:samempv_unitary}
	Apply Theorem~\ref{thm:form_II} to obtain the unitary and
	the diagonal PD fixed point.
	Unitary conjugation preserves trace-preservation and the MPV
	family (Theorem~\ref{thm:samempv_unitary}).
\end{proof}

\begin{theorem}[Irreducible block decomposition with left-canonical normalization]%
	\label{thm:irreducible_block_decomp_with_cfii}
	\lean{MPSTensor.exists_irreducible_blockDecomp_with_CFII}
	\leanok
	\uses{def:irreducible_tensor}
	For any MPS tensor~$A$, there exist irreducible blocks
	$(A_k)_{k=1}^r$ that generate the same MPV family as~$A$.
	Moreover, for each block~$k$,
	if that block has already been placed in TP gauge and $D_k > 0$,
	one obtains a left-canonical normalization (unitary, diagonal PD fixed point,
	MPV preservation).
	This yields the left-canonical normalization; the further reduction to
	normal canonical form is carried out in the subsequent spectral analysis.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:irred_decomp, thm:CFII_data}
	Combine Theorems~\ref{thm:irred_decomp}
	and~\ref{thm:CFII_data}.
\end{proof}

\begin{theorem}[Normal canonical form from a primitive block decomposition]%
		\label{thm:exists_normal_canonical_form}
		\lean{MPSTensor.exists_normalCanonicalForm_of_primitive_blockDecomp}
		\leanok
		\uses{def:is_normal_canonical_form, def:blocked_tensor}
		Suppose $A$ generates the same MPV family as a weighted block-diagonal
		tensor $\bigoplus_k \mu_k A_k$, and assume that:
		\begin{enumerate}
			\item each $A_k$ is irreducible,
			\item each $A_k$ satisfies the TP normalization
			      $\sum_i (A_k^i)^\dagger A_k^i = \Id$,
			\item each transfer map $\E_{A_k}$ is primitive
			      (equivalently, has peripheral spectrum $\{1\}$),
			\item the moduli $|\mu_k|$ are pairwise distinct,
			\item all $\mu_k \neq 0$,
			\item all bond dimensions are positive.
		\end{enumerate}
		Then there exist a blocking length $p > 0$, weights $(\nu_k)$, and blocked
		tensors $(B_k)$ such that $A^{[p]}$ generates the same MPV family as
		$\bigoplus_k \nu_k B_k$, and $(\nu_k, B_k)$ satisfies the normal canonical
		form conditions.
\end{theorem}
% Scope restriction (prepared primitive block decomposition): not the
% full source theorem (lines 742--763), which starts from
% an arbitrary translation-invariant MPS representation. See
% docs/paper-gaps/pgvwc07_ti_canonical_form_scope.tex.

\begin{proof}\leanok
		\uses{def:is_normal_canonical_form}
		No new blocking is needed here: the given blocks are already primitive,
		so one may take $p = 1$.
		The theorem is stated with a blocking length because earlier reduction
		steps may require a common blocking period; under these hypotheses
		that period can be chosen to be $1$.
		Since the moduli $|\mu_k|$ are pairwise distinct, there is a reordering
		of the blocks for which they become strictly decreasing.
		After this reordering, set $\nu_k := \mu_k$ and $B_k := A_k$ in the new
		order.
		Irreducibility, TP normalization, primitivity, nonzero weights, and
		positive bond dimensions are unchanged by the reordering, so the
		reordered family satisfies Definition~\ref{def:is_normal_canonical_form}
		and generates the same MPV family as~$A$.
\end{proof}

\section{Zero-block separation}%
\label{sec:zero_block_separation}

In the irreducible block decomposition, some blocks may have all-zero
Kraus operators. Such blocks contribute to the MPV only at word length
$N = 0$ (where their contribution equals their bond dimension).
The following results separate the zero blocks from the nonzero blocks.

\begin{definition}[Zero MPS tensor]\label{def:zero_mps_tensor}
	\lean{MPSTensor.zeroMPSTensor}
	\leanok
	The \emph{zero MPS tensor} of physical dimension $d$ and bond
	dimension $D$, denoted $\mathbf{0}_{d,D}$, is the tensor whose
	Kraus operators are all zero.
\end{definition}

\begin{theorem}[MPV of zero tensor]\label{thm:mpv_zero_tensor}
	\lean{MPSTensor.mpv_zeroMPSTensor}
	\leanok
	\uses{def:zero_mps_tensor}
	For any spin configuration $\sigma$ of length $N$,
	\[
		V^{(N)}(\mathbf{0}_{d,D})_\sigma
		= \begin{cases} D & \text{if } N = 0, \\ 0 & \text{if } N \ge 1. \end{cases}
	\]
\end{theorem}

\begin{proof}\leanok
	For $N = 0$, $(\mathbf{0}_{d,D})^{\varnothing} = \Id_{D}$,
	so the MPV equals $\tr(\Id_D) = D$.
	For $N \ge 1$, every word product of a zero tensor is zero.
\end{proof}

\begin{theorem}[Zero-block separation]%
	\label{thm:zero_block_separation}
	\lean{MPSTensor.exists_irreducible_blockDecomp_nonzeroBlocks}
	\leanok
	\uses{def:irreducible_tensor, def:zero_mps_tensor}
	Every MPS tensor $A$ admits a block decomposition into:
	\begin{enumerate}
		\item a \emph{trivial block} of bond dimension $D_0$ with all Kraus operators zero
		      (accumulating all all-zero irreducible blocks), and
		\item a finite family of \emph{nontrivial blocks} $(B_k)_{k=1}^r$,
		      each irreducible, each with at least one nonzero Kraus
		      operator, and each with positive bond dimension,
	\end{enumerate}
	such that
	\[
		V^{(N)}(A)_\sigma
		= V^{(N)}(\mathbf{0}_{d,D_0})_\sigma
		+ V^{(N)}\!\bigl(\textstyle\bigoplus_{k=1}^r B_k\bigr)_\sigma
	\]
	for every spin configuration $\sigma$ of length $N$.
	At $N = 0$ this reads $D = D_0 + \sum_{k=1}^r D_k$.
	For $N \ge 1$ the zero-tail term vanishes.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:irred_decomp, thm:mpv_zero_tensor}
	Apply the irreducible block decomposition
	(Theorem~\ref{thm:irred_decomp}) and partition the blocks
	by whether each has a nonzero Kraus operator.
	The all-zero blocks satisfy a dimension bound via
	irreducibility; their total bond dimension is $D_0$.
	The MPV identity follows from the block structure.
\end{proof}

\section{TP-gauge reduction for arbitrary tensors}%
\label{sec:tp_gauge_arbitrary}

\begin{theorem}[TP-gauge reduction for arbitrary tensors with zero-block separation]%
	\label{thm:tp_gauge_arbitrary}
	\lean{MPSTensor.exists_tp_gauge_from_arbitrary_with_zeroTail}
	\leanok
	\uses{def:irreducible_tensor, def:zero_mps_tensor}
	For any MPS tensor $A$, there exist:
	\begin{enumerate}
		\item a zero-tail bond dimension $D_0$,
		\item nontrivial blocks $(B_k)_{k=1}^r$ with nonzero weights
		      $(\mu_k)_{k=1}^r$, each block irreducible and
		      left-canonical ($\sum_i (B_k^i)^\dagger B_k^i = \Id$),
		      with positive bond dimension,
	\end{enumerate}
	such that
	\[
		V^{(N)}(A)_\sigma
		= V^{(N)}(\mathbf{0}_{d,D_0})_\sigma
		+ V^{(N)}\!\bigl(\textstyle\bigoplus_{k=1}^r \mu_k B_k\bigr)_\sigma.
	\]
	This is the furthest unconditional reduction step available for arbitrary tensors
	before periodicity removal.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:zero_block_separation, thm:tp_data_irreducible}
	Apply Theorem~\ref{thm:zero_block_separation} to separate the
	trivial block and nontrivial blocks. Then apply the blockwise TP-gauge theorem
	to the nontrivial blocks. The MPV identity chains through both steps.
\end{proof}

\section{Blocking and primitivity reduction}%
\label{sec:blocking_infrastructure}

\begin{theorem}[Blocking distributes over MPV equivalence]%
	\label{thm:samempv_blocking_distributes}
	\lean{MPSTensor.sameMPV₂_blockTensor_of_sameMPV₂_toTensorFromBlocks}
	\leanok
	\uses{def:blocked_tensor}
	If $A$ generates the same MPV family as $\bigoplus_k \mu_k B_k$,
	then for any $p \ge 0$,
	\[
		A^{[p]} \sim \bigoplus_k \mu_k^p\, B_k^{[p]}.
	\]
\end{theorem}

\begin{proof}\leanok
	The identity $V^{(N)}(A^{[p]})_\sigma = V^{(Np)}(A)_{\sigma_\mathrm{flat}}$
	holds for any tensor by definition of blocking.
	The MPV equivalence of $A$ and $\bigoplus_k \mu_k B_k$ at length $Np$
	then gives the equivalence at length $N$ with weights $\mu_k^p$.
\end{proof}

\begin{theorem}[Blocking preserves equality of assembled weighted blocks]%
	\label{thm:samempv_toTensorFromBlocks_block_power}
	\lean{MPSTensor.sameMPV₂_toTensorFromBlocks_blockPower}
	\leanok
	\uses{thm:block_tensor_to_tensor_from_blocks}
	If two assembled weighted block tensors generate the same MPV family, then
	after any common blocking length $p \ge 0$, the assembled blocked families with
	weights transported to the powers $\mu_k^p$ again generate the same MPV family.
\end{theorem}

\begin{proof}\leanok
	Apply the MPV equality to the flattened blocked word, then rewrite each
	blocked assembled tensor by Theorem~\ref{thm:block_tensor_to_tensor_from_blocks}.
\end{proof}

\begin{definition}[Flattened iterated blocking]%
	\label{def:flattened_iterated_block_tensor}
	\lean{MPSTensor.reindexPhysical,
		MPSTensor.flattenedIteratedBlockTensor}
	\leanok
	\uses{def:blocked_tensor}
	If a tensor $A$ already has physical alphabet $d^p$, its flattened iterated
	blocking by a further length $L$ block is the tensor on the direct alphabet
	$d^{pL}$ obtained
	by first taking the ordinary length-$L$ block over the alphabet $d^p$ and then
	grouping each direct length-$pL$ word into $L$ consecutive length-$p$ words.
\end{definition}

\begin{lemma}[MPVs under physical reindexing and flattened iterated blocking]%
	\label{lem:mpv_flattened_iterated_blocking}
	\lean{MPSTensor.evalWord_reindexPhysical,
		MPSTensor.mpv_reindexPhysical,
		MPSTensor.sameMPV₂_reindexPhysical,
		MPSTensor.mpv_flattenedIteratedBlockTensor}
	\leanok
	\uses{def:flattened_iterated_block_tensor, def:same_mpv2}
	Word evaluation and MPV evaluation after physical reindexing are obtained by
	applying the index map to every physical letter.  Therefore MPV equality between
	two tensors of arbitrary bond dimensions is preserved by the same physical
	reindexing.  For flattened iterated blocking, evaluation is the evaluation of
	the iterated block on the grouped physical configuration.
\end{lemma}

\begin{lemma}[Transport across a physical alphabet equivalence]%
	\label{lem:physical_reindex_equiv_transport}
	\lean{MPSTensor.reindexPhysical_cast_dim,
		MPSTensor.transferMap_reindexPhysical_equiv,
		MPSTensor.leftCanonical_reindexPhysical_equiv,
		MPSTensor.isPrimitive_transferMap_reindexPhysical_equiv,
		MPSTensor.hasInvariantProj_reindexPhysical_equiv,
		MPSTensor.isIrreducibleTensor_reindexPhysical_equiv,
		MPSTensor.isInjective_reindexPhysical_equiv,
		MPSTensor.gaugePhaseEquiv_reindexPhysical_equiv}
	\leanok
	\uses{def:transfer_map, def:primitive_channel}
	A bijection between physical alphabets of sizes $d_1$ and $d_2$ merely
	relabels the Kraus summation.  Therefore physical reindexing commutes with
	bond-dimension casts, leaves the transfer map unchanged, and carries
	left-canonical normalization, transfer-map primitivity, invariant
	projections, irreducibility, injectivity, and gauge-phase equivalence back
	and forth across the bijection.
\end{lemma}

\begin{proof}\leanok
	The transfer-map and left-canonical identities are reindexings of finite sums
	by the chosen bijection.  The primitive-transfer statement follows from equality
	of the transfer maps.
\end{proof}

\begin{theorem}[Flattened iterated blocking agrees with direct blocking]%
	\label{thm:flattened_iterated_blocking}
	\lean{MPSTensor.flattenedIteratedBlockTensor_blockTensor,
		MPSTensor.toTensorFromBlocks_flattenedIteratedBlockTensor,
		MPSTensor.sameMPV₂_toTensorFromBlocks_flattenedIteratedBlockPower}
	\leanok
	\uses{def:flattened_iterated_block_tensor, lem:mpv_flattened_iterated_blocking,
		thm:samempv_toTensorFromBlocks_block_power}
	For a tensor first blocked by $p$, flattened iterated blocking by $L$ agrees
	with direct blocking by $pL$.  The same statement holds for assembled weighted
	block sums:
	flattened iterated blocking is the physical reindexing of the iterated blocked
	assembled family, and full MPV equality of two assembled families is preserved
	after this
	blocking with weights transported to the powers $\mu_k^L$.
\end{theorem}

\begin{proof}\leanok
	The proof is the combinatorial identity identifying a direct length-$pL$ word
	with an $L$-tuple of length-$p$ words.  For assembled block sums, apply the
	ordinary blocking theorem over the already blocked alphabet and then reindex the
	physical alphabet by this grouping map.
\end{proof}

\begin{theorem}[Blocking assembled weighted blocks]%
	\label{thm:block_tensor_to_tensor_from_blocks}
	\lean{MPSTensor.sameMPV₂_blockTensor_toTensorFromBlocks}
	\leanok
	\uses{thm:samempv_blocking_distributes, def:block_diagonal_tensor, def:blocked_tensor}
	For any weighted block tensor $\bigoplus_k \mu_k B_k$ and any blocking
	length $p \ge 0$, its blocked MPV family agrees with the weighted block tensor
	formed from the blocked blocks $B_k^{[p]}$ and the powered weights $\mu_k^p$.
\end{theorem}

\begin{proof}\leanok
	Apply Theorem~\ref{thm:samempv_blocking_distributes} to the identity MPV
	relation for the assembled weighted block tensor.
\end{proof}

\begin{theorem}[Positive-length MPV equality after blocking]%
	\label{thm:samempv2pos_block_tensor}
	\lean{MPSTensor.sameMPV₂Pos_blockTensor}
	\leanok
	\uses{def:same_mpv2_pos, def:blocked_tensor}
	If two tensors agree on all positive system sizes, then their $p$-blocked
	tensors agree on all positive blocked system sizes for every $p \ge 1$.
\end{theorem}

\begin{proof}\leanok
	A positive blocked chain of length $N$ corresponds to an original chain of
	length $Np$, which is positive when both $N$ and $p$ are positive.
\end{proof}

\begin{theorem}[Positive-length equality for powered weighted nonzero blocks]%
	\label{thm:samempv2pos_toTensorFromBlocks_block_power}
	\lean{MPSTensor.sameMPV₂Pos_toTensorFromBlocks_blockPower}
	\leanok
	\uses{thm:block_tensor_to_tensor_from_blocks, thm:samempv2pos_block_tensor,
	      def:block_diagonal_tensor}
	If two weighted nonzero-block tensors agree on all positive system sizes, then
	after any positive common blocking their blocked nonzero-block tensors still agree
	on all positive system sizes, with weights transported from $\mu_k$ to
	$\mu_k^p$.
\end{theorem}

\begin{proof}\leanok
	Use Theorem~\ref{thm:samempv2pos_block_tensor} for the assembled nonzero parts,
	then rewrite each blocked assembled tensor by
	Theorem~\ref{thm:block_tensor_to_tensor_from_blocks}.
\end{proof}

\begin{lemma}[Nonzero weights after blocking powers]%
	\label{lem:block_weights_ne_zero}
	\lean{MPSTensor.blockWeights_ne_zero}
	\leanok
	If every $\mu_k$ is nonzero, then every powered weight $\mu_k^p$ is nonzero.
\end{lemma}

\begin{proof}\leanok
	This is the elementary fact that powers of a nonzero complex number are
	nonzero.
\end{proof}

\begin{lemma}[Nonzero weights after sector replication]%
	\label{lem:replicated_weights_pow_ne_zero}
	\lean{MPSTensor.replicatedWeights_pow_ne_zero}
	\leanok
	Suppose a nonzero-weight block $k$ is replaced by a finite family of cyclic sectors. If
	$\mu_k \ne 0$, then every replicated sector weight $\mu_k^{p_k}$ remains
	nonzero.
\end{lemma}

\begin{proof}\leanok
	Apply Lemma~\ref{lem:block_weights_ne_zero} to the parent block weight and the
	blocking power assigned to that block.
\end{proof}

\begin{lemma}[Nonzero replicated weights with sector phases]%
	\label{lem:replicated_weights_pow_mul_phase_ne_zero}
	\lean{MPSTensor.replicatedWeights_pow_mul_phase_ne_zero}
	\leanok
	If the replicated sector weights are additionally multiplied by nonzero sector
	phases, then the resulting sector weights are still nonzero.
\end{lemma}

\begin{proof}\leanok
	Combine Lemma~\ref{lem:replicated_weights_pow_ne_zero} with nonvanishing of
	the phase factor.
\end{proof}

\begin{theorem}[Primitive channels remain primitive under powers]%
	\label{thm:primitive_pow}
	\lean{MPSTensor.isPrimitive_pow_of_isPrimitive}
	\leanok
	If $\E$ has peripheral spectrum $\{1\}$ (is primitive) and $m \ge 1$,
	then $\E^m$ is also primitive.
\end{theorem}

\begin{proof}\leanok
	The peripheral spectrum of $\E^m$ is the set of $m$-th powers of
	that of $\E$.  Since $\{1\}^m = \{1\}$, the result follows.
\end{proof}

\begin{theorem}[Blocking-period divisibility preserves primitivity]%
	\label{thm:primitive_blocking_divisible}
	\lean{MPSTensor.isPrimitive_transferMap_blockTensor_of_dvd}
	\leanok
	\uses{def:blocked_tensor}
	If the transfer map of $A^{[p]}$ is primitive, $p \mid q$,
	and $q \ge 1$, then the transfer map of $A^{[q]}$ is also primitive.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:primitive_pow}
	Write $q = p \cdot m$.
	Then $\E_{A^{[q]}} = \E_{A^{[p]}}^m$.
	Apply Theorem~\ref{thm:primitive_pow} with $\E = \E_{A^{[p]}}$.
\end{proof}

\begin{theorem}[Iterated blocking multiplies physical dimensions]
	\label{thm:iterated_blocking_phys_dim}
	\lean{MPSTensor.blockPhysDim_blockPhysDim}
	\leanok
	\uses{def:blocked_tensor}
	The physical alphabet obtained by first blocking $m$ sites and then blocking
	$n$ of those sites has the same cardinality as blocking $mn$ original sites:
	\[
		\operatorname{blockPhysDim}(\operatorname{blockPhysDim}(d,m),n)
		= \operatorname{blockPhysDim}(d,mn).
	\]
\end{theorem}

\begin{proof}\leanok
	Both sides are powers of $d$: the left-hand side is $(d^m)^n$, and the
	right-hand side is $d^{mn}$.
\end{proof}

\begin{theorem}[Compatibility of direct and iterated blocking]
	\label{thm:iterated_blocking_samempv_reindex}
	\lean{MPSTensor.sameMPV₂_blockTensor_blockTensor_mul_reindex}
	\leanok
	\uses{def:blocked_tensor, def:same_mpv2}
	If one first blocks $m$ sites and then blocks $n$ of the resulting sites,
	the resulting tensor has the same MPV family as the tensor obtained by
	blocking $mn$ original sites, after the physical labels are identified by
	flattening each iterated block.
\end{theorem}

\begin{proof}\leanok
	For each iterated blocked index, decode the outer block, decode each inner
	block, and concatenate the resulting words. This is exactly the word
	decoded from the corresponding directly blocked index, so the word
	evaluations, and hence all MPV coefficients, agree.
\end{proof}

\begin{theorem}[Block-grouping bijection recovers direct blocking]
	\label{thm:reindex_direct_to_iterated_block_tensor}
	\lean{MPSTensor.reindexPhysical_directToIteratedBlockIndex_blockTensor}
	\leanok
	\uses{def:blocked_tensor}
	Reindexing the physical alphabet of the iterated block tensor by the
	consecutive-grouping bijection returns the tensor obtained by blocking
	$mn$ original sites directly.
\end{theorem}

\begin{proof}
	\leanok
	Apply the tensor-level direct-to-iterated blocking identity and cancel the
	forward and inverse maps of the grouping bijection.
\end{proof}

\begin{corollary}[Block-grouping bijection preserves the MPV family]
	\label{cor:samempv2_reindex_direct_to_iterated_block_tensor}
	\lean{MPSTensor.sameMPV₂_reindexPhysical_directToIteratedBlockIndex_blockTensor}
	\leanok
	\uses{def:same_mpv2, def:blocked_tensor}
	Under the block-grouping bijection, the iterated block tensor and the
	directly blocked tensor have the same matrix-product-vector family.
\end{corollary}

\begin{proof}
	\leanok
	\uses{thm:reindex_direct_to_iterated_block_tensor}
	This is the matrix-product-vector form of the preceding tensor identity.
\end{proof}

\begin{theorem}[Physical-dimension transport and MPV equality]
	\label{thm:samempv_cast_phys_dim}
	\lean{MPSTensor.sameMPV₂_cast_physDim}
	\leanok
	\uses{def:same_mpv2}
	If two physical alphabets have the same cardinality, then the transported
	tensors have the same MPV family if and only if the original tensors do.
\end{theorem}

\begin{proof}\leanok
	After substituting the equality of physical dimensions, the two MPV equality
	statements are the same.
\end{proof}

\begin{theorem}[Physical-dimension transport for block-diagonal tensors]
	\label{thm:tensor_from_blocks_cast_phys_dim}
	\lean{MPSTensor.toTensorFromBlocks_cast_physDim}
	\leanok
	\uses{def:block_diagonal_tensor}
	Identifying equal physical alphabets commutes with forming a weighted
	block-diagonal tensor from a finite block family.
\end{theorem}

\begin{proof}\leanok
	Substitute the equality of physical dimensions; both sides are then the same
	block-diagonal tensor.
\end{proof}

\begin{theorem}[Physical-dimension transport and trace preservation]
	\label{thm:left_canonical_cast_phys_dim}
	\lean{MPSTensor.leftCanonical_cast_physDim}
	\leanok
	\uses{def:tp_kraus}
	After identifying equal physical alphabets, the transported tensor is
	trace-preserving if and only if the original tensor is trace-preserving.
\end{theorem}

\begin{proof}\leanok
	Substitute the equality of physical dimensions; the two normalization sums are
	identical.
\end{proof}

\begin{theorem}[Physical-dimension transport and primitive transfer maps]
	\label{thm:primitive_transfer_cast_phys_dim}
	\lean{MPSTensor.isPrimitive_transferMap_cast_physDim}
	\leanok
	\uses{def:transfer_map, def:primitive_channel}
	After identifying equal physical alphabets, the transported transfer map is
	primitive if and only if the original transfer map is primitive.
\end{theorem}

\begin{proof}\leanok
	Substitute the equality of physical dimensions; the transfer maps coincide.
\end{proof}

\begin{theorem}[Physical-dimension transport and tensor irreducibility]
	\label{thm:irreducible_tensor_cast_phys_dim}
	\lean{MPSTensor.isIrreducibleTensor_cast_physDim}
	\leanok
	\uses{def:irreducible_tensor}
	After identifying equal physical alphabets, the transported tensor is
	irreducible if and only if the original tensor is irreducible.
\end{theorem}

\begin{proof}\leanok
	Substitute the equality of physical dimensions; the irreducibility predicates
	are identical.
\end{proof}

\begin{theorem}[Iterated blocking multiplies periods at the transfer-map level]%
	\label{thm:iterated_blocking_transfer}
	\lean{MPSTensor.transferMap_blockTensor_mul}
	\leanok
	\uses{def:blocked_tensor}
	For any MPS tensor $A$ and integers $m,n \ge 0$,
	\[
		\E_{(A^{[m]})^{[n]}} = \E_{A^{[mn]}}.
	\]
\end{theorem}

\begin{proof}\leanok
	Rewrite the outer blocked transfer map as the $n$-th power of
	$\E_{A^{[m]}}$, then rewrite $\E_{A^{[m]}}$ as $\E_A^m$.
	This gives $(\E_A^m)^n$, which simplifies to $\E_A^{mn}$.
	Finally rewrite $\E_A^{mn}$ back as the transfer map of $A^{[mn]}$.
\end{proof}

\begin{theorem}[Common blocking period for a block family]%
	\label{thm:common_blocking_period}
	\lean{MPSTensor.exists_common_blocking_all_primitive}
	\leanok
	\uses{def:blocked_tensor}
	Let $(B_k)_{k=1}^r$ be a finite family of MPS tensors with
	positive bond dimensions, each admitting a blocking period~$p_k$
	making its transfer map primitive.
	Then $p = \operatorname{lcm}(p_1, \ldots, p_r)$ is a common
	period: $\E_{B_k^{[p]}}$ is primitive for every~$k$.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:primitive_blocking_divisible}
	Each $p_k \mid p$,
	so Theorem~\ref{thm:primitive_blocking_divisible} applies.
\end{proof}

\begin{remark}
	See \cref{def:common_period_blocking} for the formalized common-period blocked family construction.
\end{remark}

\begin{remark}
	See \cref{thm:common_period_blocking_primitive} for the formalized primitivity statement for common-period blocking.
\end{remark}

\begin{theorem}[Formula for common-period blocking]%
	\label{thm:common_period_blocking_apply}
	\lean{MPSTensor.commonPeriodBlocking_apply}
	\leanok
	\uses{def:common_period_blocking}
	Each block of the common-period blocking is exactly the corresponding tensor
	blocked at the common least common multiple.
\end{theorem}

\begin{proof}\leanok
	\uses{def:common_period_blocking}
	This is the definition of the common-period blocking.
\end{proof}

\section{Period removal (cyclic-sector decomposition)}%
\label{sec:cyclic_sectors}

In the non-periodic route of \cite{Cirac2017MPDO_AnnPhys} the only
required step is periodicity removal by blocking: after blocking
each irreducible TP block by the least common multiple of its periods,
the resulting tensor has no nontrivial $p$-periodic vectors
\cite[Section~2.3]{Cirac2017MPDO_AnnPhys}.  The normal/BNT comparison
then proceeds without further cyclic-sector structure.

What follows is the formal derivation of that period-removal step.
The cyclic-sector decomposition itself comes from the peripheral
spectrum of the adjoint transfer map
(\cite[Theorem~6.6]{Wolf2012Quantum}).
The additional lemmas about projection multiplicativity, sector
primitivity, and irreducibility are internal to the period-removal
argument and are not independent paper-level results.

The common-reblocked cyclic-sector family
(Definition~\ref{def:common_blocked_cyclic_sector_family})
is an internal bookkeeping record that reblocks all sectors to a
shared physical alphabet. The reindexing, flattening, and
common-alphabet transport theorems attached to it are internal
plumbing; they are consumed by the normal canonical-form existence
pipeline but do not correspond to independent paper-level
statements.

\begin{theorem}[Compression to a supported projection sector]%
	\label{thm:compressed_sector}
	\lean{MPSTensor.exists_compressedTensor_of_supported_projection}
	\leanok
	Let $A$ be an MPS tensor satisfying
	$\sum_i (A^i)^\dagger A^i = P$ for an orthogonal projection $P$,
	and suppose $P A^i P = A^i$ for every~$i$.
	Then there exist a left-canonical tensor $C$ of bond dimension
	$n = \tr(P)$ and a $\ast$-algebra isomorphism
	$\varphi : \MN{n} \xrightarrow{\sim} P \MN{D} P$
	(preserving both multiplication and adjoint) with
	\[
		V^{(N)}(C)_\sigma = \tr\!\bigl(P \cdot A^\sigma\bigr),
		\qquad
		\varphi\!\bigl(\E_{C^\dagger}(X)\bigr)
			= \E_{A^\dagger}\!\bigl(\varphi(X)\bigr)
	\]
	for every~$\sigma$ and every $X \in \MN{n}$, together with
	$\varphi(XY) = \varphi(X)\varphi(Y)$ and
	$\varphi(X^\dagger) = \varphi(X)^\dagger$.
\end{theorem}

\begin{proof}\leanok
	Diagonalize $P$ via a unitary $U$ (spectral theorem).
	The eigenvalues are in $\{0,1\}$.
	Set $n = \tr P$; let $C^i$ be the top-left $n\times n$ block
	of $U^\dagger A^i U$ (restricted to eigenvalue-$1$ indices).
	TP follows from the projection normalization;
	the MPV identity follows from trace cyclicity.
\end{proof}

\begin{corollary}[Compression preserves positive-length MPV coefficients]%
	\label{cor:compressed_sector_pos_mpv}
	\lean{MPSTensor.exists_compressedTensor_of_supported_projection_pos_mpv}\leanok
	\uses{thm:compressed_sector}
	Under the hypotheses of Theorem~\ref{thm:compressed_sector}, the
	compressed tensor $C$ additionally satisfies
	$V^{(N)}(C)_\sigma = V^{(N)}(A)_\sigma$ for all positive chain
	lengths $N > 0$.
\end{corollary}
\begin{proof}\leanok
    \uses{thm:compressed_sector}
    For $N > 0$ the word $A^{i_1} \cdots A^{i_N}$ is nonempty, so $P$
    acts as the identity on the left by idempotence and the support
    hypothesis.  The MPV identity then follows from trace cyclicity
    via Theorem~\ref{thm:compressed_sector}.
\end{proof}

\begin{theorem}[Block decomposition from commuting projections]%
	\label{thm:blockdecomp_commuting_projs}
	\lean{MPSTensor.exists_blockDecomp_of_commuting_projections}
	\leanok
	\uses{def:irreducible_tensor}
	Let $A$ be a left-canonical MPS tensor, and let
	$(P_k)_{k=1}^m$ be orthogonal projections summing to $\Id$
	that commute with every Kraus operator:
	$P_k A^i = A^i P_k$ for all $i, k$.
	Then there exist left-canonical blocks $(C_k)_{k=1}^m$
	with unit weights such that $A$ generates the same MPV family
	as $\bigoplus_{k=1}^m C_k$.
	Moreover, each sector satisfies the per-sector trace relation
	$V^{(N)}(C_k)_\sigma = \tr(P_k \cdot A^{i_1} \cdots A^{i_N})$
	for every word $\sigma = (i_1, \dots, i_N)$, and each block carries a
	compression linear equivalence
	$\varphi_k : \MN{\dim C_k} \xrightarrow{\sim}
		P_k \MN{D} P_k$
	intertwining the compressed adjoint transfer map with the sector adjoint
	transfer map.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:compressed_sector}
	For each $k$, the tensor $P_k A^i$ is supported on $P_k$
	(using commutativity and $P_k^2 = P_k$).
	The TP condition $\sum_i (P_k A^i)^\dagger (P_k A^i) = P_k$
	follows from $\sum_i (A^i)^\dagger A^i = \Id$.
	Apply Theorem~\ref{thm:compressed_sector} to each sector.
	The MPV identity uses $\sum_k P_k = \Id$ and linearity of trace.
\end{proof}

\begin{theorem}[Block decomposition from adjoint-fixed projections]%
	\label{thm:blockdecomp_adjoint_fixed}
	\lean{MPSTensor.exists_blockDecomp_of_adjoint_fixed_projections}
	\leanok
	\uses{def:irreducible_tensor}
	Let $A$ be a left-canonical MPS tensor, and let
	$(P_k)_{k=1}^m$ be orthogonal projections summing to $\Id$
	that are fixed by the adjoint transfer map:
	$\E_A^\dagger(P_k) = P_k$ for every $k$.
	Then the same block decomposition conclusion holds as in
	Theorem~\ref{thm:blockdecomp_commuting_projs}, including the
	per-sector trace relation.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:blockdecomp_commuting_projs, thm:kraus_commute}
	The Kadison--Schwarz equality condition for a TP map
	(Theorem~\ref{thm:kraus_commute}) implies that a projection fixed by
	$\E_A^\dagger$ commutes with every Kraus operator.
	Apply Theorem~\ref{thm:blockdecomp_commuting_projs}.
\end{proof}

\begin{theorem}[Conjugate-transposed Kraus family for irreducible TP tensors]%
	\label{thm:conjtranspose_kraus_setup}
	\lean{MPSTensor.conjTranspose_kraus_setup}
	\leanok
	\uses{def:irreducible_tensor}
	Let $A$ be a left-canonical irreducible tensor.
	Then the conjugate-transposed family $K_i := (A^i)^\dagger$ is unital
	and irreducible, and there exists a positive-definite matrix $\rho$
	fixed by the adjoint map associated to~$K$.
\end{theorem}

\begin{proof}\leanok
	\uses{def:irreducible_tensor, thm:qpf}
	Left-canonicality gives $\sum_i (A^i)^\dagger A^i = \Id$, so the
	conjugate-transposed family is unital.
	Irreducibility passes to the adjoint map, and the quantum
	Perron--Frobenius theorem provides the required positive-definite fixed
	point.
\end{proof}

\begin{theorem}[One full cyclic turn fixes the sector projections]%
	\label{thm:cyclic_projection_periodic_fixed}
	\lean{MPSTensor.adjointTransferMap_pow_fixes_cyclic_projection}
	\leanok
	If a family of projections $(P_k)_{k=1}^m$ satisfies
	$\E^\dagger(P_{k+1}) = P_k$, then the $m$-th iterate of $\E^\dagger$
	fixes each projection $P_k$.
\end{theorem}

\begin{proof}\leanok
	Iterate the cyclic relation $m$ times.
	Because the indices are taken modulo $m$, the projection returns to its
	starting position.
\end{proof}

\begin{theorem}[Blocked adjoint transfer map as an iterate]%
	\label{thm:blocked_adjoint_transfer_pow}
	\lean{MPSTensor.transferMap_adjoint_blocked_eq_pow}
	\leanok
	\uses{def:blocked_tensor, lem:transfer_conjtranspose_adjoint,
	      thm:iterated_blocking_transfer}
	For every tensor $A$, blocking length $m$, and matrix $X$,
	\[
		\E_{(A^{[m]})^\dagger}(X) = (\E_{A^\dagger})^m(X).
	\]
\end{theorem}

\begin{proof}\leanok
	\uses{lem:transfer_conjtranspose_adjoint, thm:iterated_blocking_transfer}
	Write both transfer maps as Frobenius adjoints of ordinary blocked transfer
	maps. Then use Theorem~\ref{thm:iterated_blocking_transfer}.
\end{proof}

\begin{theorem}[Period removal by blocking (1606.00608, §2.3)]%
	\label{thm:cyclic_sector_decomp_after_blocking}
	\lean{MPSTensor.exists_cyclic_sector_decomp_after_blocking}
	\leanok
	\uses{def:blocked_tensor, thm:blockdecomp_adjoint_fixed}
	Let $A$ be a left-canonical irreducible tensor of period $m$.
	Then the blocked tensor $A^{[m]}$ admits:
	\begin{enumerate}
		\item left-canonical sector tensors $(C_k)_{k=1}^m$,
		\item orthogonal projections $(P_k)_{k=1}^m$ summing to $\Id$ and satisfying
		      $\E_{A^\dagger}(P_{k+1}) = P_k$,
		\item compression linear equivalences
		      $\varphi_k : \MN{\dim C_k} \xrightarrow{\sim} P_k \MN{D} P_k$,
		\item a direct-sum MPV decomposition
		      $A^{[m]} \sim \bigoplus_{k=1}^m C_k$.
	\end{enumerate}
	This is the period-removal step: after blocking by $m$, the
	result admits a unit-weight decomposition with no non-trivial
	$p$-periodic vectors.  The cyclic projections $P_k$ come from
	the peripheral spectrum of the adjoint transfer map
	(\cite[Theorem~6.6]{Wolf2012Quantum}).
\end{theorem}

\begin{proof}\leanok
	\uses{thm:cyclic_projection_periodic_fixed, thm:blocked_adjoint_transfer_pow,
	      thm:blockdecomp_adjoint_fixed}
	The cyclic decomposition of the adjoint transfer map gives projections
	$(P_k)$ with $\E_{A^\dagger}(P_{k+1}) = P_k$.
	Theorem~\ref{thm:cyclic_projection_periodic_fixed} upgrades this to
	$(\E_{A^\dagger})^m(P_k) = P_k$, and
	Theorem~\ref{thm:blocked_adjoint_transfer_pow} identifies
	$(\E_{A^\dagger})^m$ with the adjoint transfer map of the blocked tensor.
	Apply Theorem~\ref{thm:blockdecomp_adjoint_fixed}.
\end{proof}

\begin{theorem}[Period removal for irreducible TP tensors]%
	\label{thm:cyclic_sector_decomp_irr_tp}
	\lean{MPSTensor.exists_cyclic_sector_decomp_of_TP_of_isIrreducibleTensor}
	\leanok
	\uses{def:blocked_tensor, def:irreducible_tensor}
	Let $A$ be a left-canonical irreducible tensor.
	Then there exist $m \ge 1$ and left-canonical blocks
	$(C_k)_{k=1}^m$ such that the $m$-blocked tensor $A^{[m]}$
	admits a unit-weight decomposition into the $C_k$.
	This is the period-removal step of \cite[Section~2.3]{Cirac2017MPDO_AnnPhys}.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:conjtranspose_kraus_setup, thm:cyclic_sector_decomp_after_blocking}
	Apply Theorem~\ref{thm:conjtranspose_kraus_setup} to obtain the
	conjugate-transposed Kraus family, an irreducible adjoint transfer map,
	and a positive-definite fixed point.
	The cyclic peripheral-spectrum theorem then provides a period $m$ and the
	required cyclic spectral data.
	Apply Theorem~\ref{thm:cyclic_sector_decomp_after_blocking}.
\end{proof}

\begin{theorem}[Period removal with primitive irreducible sectors]%
\label{thm:cyclic_sector_decomp_irr_tp_prim_irr}
	\lean{MPSTensor.exists_primitive_irreducible_cyclic_sector_decomp_of_TP_of_isIrreducibleTensor}
	\leanok
	\uses{thm:cyclic_sector_decomp_after_blocking, thm:blocked_sector_unconditional}
	Let $A$ be a left-canonical irreducible tensor. Then there is a period
	$m \ge 1$ and a unit-weight decomposition of $A^{[m]}$ into sector blocks
	$(C_k)_{k=1}^m$ such that every $C_k$ is left-canonical, has primitive
	transfer map, is tensor-irreducible, and has nonzero bond dimension.

	This packages the period-removal step together with the
	internal sector-orbit lift; it is the result consumed by the
	common-blocking construction downstream.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:conjtranspose_kraus_setup, thm:peripheral_cyclic_structure,
		thm:cyclic_sector_decomp_after_blocking, thm:blocked_sector_unconditional}
	The proof repeats the cyclic-sector construction of
	Theorem~\ref{thm:cyclic_sector_decomp_irr_tp}, keeps the projection and
	compression data, and applies
	Theorem~\ref{thm:blocked_sector_unconditional} to the resulting
	sector blocks.
\end{proof}

\begin{definition}[Primitive irreducible cyclic-sector data]%
\label{def:primitive_irreducible_cyclic_sectors}
	\lean{MPSTensor.HasPrimitiveIrreducibleCyclicSectors}
	\leanok
	\uses{def:blocked_tensor, def:block_diagonal_tensor, def:same_mpv2}
	A tensor has primitive irreducible cyclic sectors if some positive
	period-removal length $m$ represents $A^{[m]}$ as a unit-weight block tensor
	whose sector blocks are trace-preserving, have primitive transfer maps, are
	tensor-irreducible, and have positive bond dimensions. This length $m$ is the
	period-removal length, not the later finite blocking length used for common
	refinement or injectivity.
\end{definition}

\begin{theorem}[Primitive cyclic sectors of irreducible TP tensors]%
\label{thm:has_primitive_irreducible_cyclic_sectors_tp_irr}
	\lean{MPSTensor.hasPrimitiveIrreducibleCyclicSectors_of_TP_of_isIrreducibleTensor}
	\leanok
	\uses{def:primitive_irreducible_cyclic_sectors,
		thm:cyclic_sector_decomp_irr_tp_prim_irr}
	Every trace-preserving irreducible tensor satisfies
	Definition~\ref{def:primitive_irreducible_cyclic_sectors}.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:cyclic_sector_decomp_irr_tp_prim_irr}
	This is Theorem~\ref{thm:cyclic_sector_decomp_irr_tp_prim_irr}, restated as
	the predicate of Definition~\ref{def:primitive_irreducible_cyclic_sectors}.
\end{proof}

\begin{definition}[Common cyclic-sector family (internal bookkeeping)]%
\label{def:common_blocked_cyclic_sector_family}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily}
	\leanok
	\uses{def:primitive_irreducible_cyclic_sectors, def:blocked_tensor,
		def:block_diagonal_tensor, def:same_mpv2}
	Internal record bundling common-reblocking data for a finite family of
	nonzero-weight blocks.  It specifies a common physical blocking length $p$,
	per-block period-removal lengths $m_k$, later reblocking lengths $e_k$ with
	$p = m_k e_k$, and the resulting flattened sector family at physical
	dimension $d^p$.  The record retains trace preservation, primitive transfer
	maps, tensor irreducibility, positive bond dimensions, and per-block
	iterated-blocking MPV equivalences.

	This is internal bookkeeping for the normal canonical-form pipeline; it
	does not represent an independent result of \cite{Cirac2017MPDO_AnnPhys}.
\end{definition}

\begin{definition}[Unit weights for a common reblocked cyclic-sector family]%
\label{def:common_blocked_cyclic_sector_flat_weight}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.flatWeight}
	\leanok
	\uses{def:common_blocked_cyclic_sector_family}
	The flattened sectors of a common reblocked cyclic-sector family carry unit
	weights inherited from cyclic period removal.  (The weights are the constant
	value~$1$; the statement that they are nonzero is trivial and not
	exposed as a separate theorem.)
\end{definition}

\begin{definition}[Common-alphabet cyclic sectors (internal)]%
\label{def:common_blocked_cyclic_sector_derived_blocks}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.flatKey,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonSectorBlock,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatDim,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatBlocks,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatBlocksAt,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonSectorTensor,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonReindexedBlock,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatWeight}
	\leanok
	\uses{def:common_blocked_cyclic_sector_family, def:blocked_tensor,
		def:block_diagonal_tensor, thm:iterated_blocking_samempv_reindex}
	Derived sector tensors, bond dimensions, flattened families, and transported
	weights obtained from a \texttt{CommonBlockedCyclicSectorFamily}.  These are
	internal definitions consumed by the normal canonical-form pipeline.
\end{definition}

\begin{theorem}[Common-alphabet sectors retain structural properties (internal)]%
\label{thm:common_blocked_cyclic_sector_derived_properties}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.commonSectorBlock_tp,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonSectorBlock_primitive,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonSectorBlock_irreducible,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonSectorBlock_dim_pos,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatBlocks_tp,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatBlocks_primitive,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatBlocks_irreducible,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatDim_pos,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonBlockWeight_ne_zero,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatWeight_ne_zero,
		MPSTensor.CommonBlockedCyclicSectorFamily.isNormalCanonicalForm_commonFlatBlocks}
	\leanok
	\uses{def:common_blocked_cyclic_sector_derived_blocks,
		thm:tp_primitive_irreducible_extra_blocking,
		thm:left_canonical_cast_phys_dim, thm:primitive_transfer_cast_phys_dim,
		thm:irreducible_tensor_cast_phys_dim, thm:ncf_sorted_tp_primitive_irred}
	Internal lemmas: the common-alphabet sectors are trace-preserving, have
	primitive transfer maps, are tensor-irreducible, and have positive bond
	dimensions.  Original nonzero weights remain nonzero after transport.  If
	transported weights are sorted by strictly decreasing norm, the flattened
	common-sector family is a normal canonical form.

	These are derived properties of the internal
	\texttt{CommonBlockedCyclicSectorFamily} record; they are not independent
	paper-level results.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:tp_primitive_irreducible_extra_blocking,
		thm:left_canonical_cast_phys_dim, thm:primitive_transfer_cast_phys_dim,
		thm:irreducible_tensor_cast_phys_dim, thm:ncf_sorted_tp_primitive_irred}
	Apply the extra-blocking theorem to each period-removal sector, substitute the
	equality between the iterated and common physical alphabets, and use that a
	positive power of a nonzero complex number is nonzero.  The normal canonical
	form statement is Theorem~\ref{thm:ncf_sorted_tp_primitive_irred} applied to
	these derived sector data.
\end{proof}

\begin{theorem}[Common-sector representatives form normal canonical form (internal)]%
\label{thm:common_representative_normal_bnt}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.commonRepresentativeWeight_strictAnti_of_weight_strictAnti,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonRepresentativeBlocksAt_tp,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonRepresentativeBlocksAt_primitive,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonRepresentativeBlocksAt_irreducible,
		MPSTensor.isNormalCanonicalFormBNT_commonRepresentativeBlocksAt}
	\leanok
	\uses{def:common_blocked_cyclic_sector_derived_blocks,
		thm:common_blocked_cyclic_sector_derived_properties,
		def:normal_cf_bnt}
	Internal lemma: for a prescribed common blocking length, choosing one
	representative per original block retains trace-preserving, primitive, and
	tensor-irreducible properties, transports strict weight ordering, and gives a
	normal canonical form with BNT separation when representatives are not
	gauge-phase equivalent.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:common_blocked_cyclic_sector_derived_properties,
		def:normal_cf_bnt}
	Transport the strict ordering of weights and the structural properties to the
	representative blocks at the prescribed length.  Apply the
	normal-canonical-form statement for the representative family.
\end{proof}

\paragraph{Internal flattening and reindexing lemmas.}
The following statements are internal lemmas about the
\texttt{CommonBlockedCyclicSectorFamily} record: they express weight formulas,
blocked-word identifications, and MPV comparisons between direct and iterated
blocking.  They are plumbing consumed by the normal canonical-form pipeline and
do not correspond to independent paper-level results.

\begin{theorem}[Weights of common-alphabet sectors]%
\label{thm:common_flat_weight_apply_of_block}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatWeight_apply_of_block}
	\leanok
	\uses{def:common_blocked_cyclic_sector_derived_blocks}
	Formula for transported weight: a sector $s$ of block $k$ carries weight
	$\mu_k^p$.
\end{theorem}

\begin{proof}\leanok
	Unfold the flattened-sector indexing and the definition of the transported
	weight.
\end{proof}

\begin{theorem}[Equal weights inside one original block]%
\label{thm:common_flat_weight_apply_block_eq}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.commonFlatWeight_apply_block_eq}
	\leanok
	\uses{def:common_blocked_cyclic_sector_derived_blocks}
	All common-alphabet sectors from the same original block carry the same
	transported weight.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:common_flat_weight_apply_of_block}
	Apply the weight formula for each sector of the fixed original block.
\end{proof}

\begin{theorem}[Blocked sectors on a common alphabet (internal flattening)]%
\label{thm:common_blocked_cyclic_sector_reindexed_samempv}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.nestedBlock_sameMPV₂_commonReindexedBlock,
		MPSTensor.CommonBlockedCyclicSectorFamily.commonReindexedBlock_sameMPV₂_commonSectorTensor,
		MPSTensor.CommonBlockedCyclicSectorFamily.sameMPV₂_weightedCommonReindexedBlock_commonFlat}
	\leanok
	\uses{def:common_blocked_cyclic_sector_derived_blocks,
		thm:iterated_blocking_samempv_reindex,
		thm:samempv_cast_phys_dim, thm:samempv_blocking_distributes,
		thm:common_blocked_cyclic_sector_derived_properties}
	Internal flattening lemmas: iterated block $(B_k^{[m_k]})^{[e_k]}$ agrees with
	the directly-blocked tensor after word identification; the weighted sum
	flattens the two-level block/sector sum into one sector family.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:iterated_blocking_samempv_reindex,
		thm:samempv_cast_phys_dim, thm:samempv_blocking_distributes,
		thm:common_blocked_cyclic_sector_derived_properties}
	Use the iterated-blocking comparison theorem, substitute the common
	physical alphabet, compose with the per-block cyclic-sector MPV equivalence,
	and reindex the double sum.
\end{proof}

\begin{definition}[Coordinate grouping for common blocked words (internal)]%
\label{def:common_blocked_cyclic_sector_grouped_block_cast}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.groupedBlockCastAgrees}
	\leanok
	\uses{def:common_blocked_cyclic_sector_derived_blocks,
		thm:iterated_blocking_phys_dim}
	Internal predicate: for one block $k$, the canonical identification of the
	common alphabet with the iterated alphabet agrees with grouping a direct
	word of length $p=m_k e_k$ into $e_k$ consecutive blocks of length $m_k$.
\end{definition}

\begin{theorem}[Flattening direct blocked words (internal digit lemma)]%
\label{thm:common_blocked_cyclic_sector_flatten_word_of_block_cast}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.flattenWordOfBlock_cast_eq}
	\leanok
	Internal digit-level lemma: for $p=m n$, the word obtained by direct
	base-$d$ decoding is the same as decoding as $n$ blocks of length $m$ and
	concatenating.
\end{theorem}

\begin{proof}
	\leanok
	Both sides extract the same base-$d$ digit of the original blocked index.
\end{proof}

\begin{theorem}[Direct and iterated blocking comparison (internal)]%
\label{thm:common_blocked_cyclic_sector_blocked_word_comparison}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.wordOfBlock_eq_iteratedBlockIndex_of_groupedBlockCastAgrees,
		MPSTensor.CommonBlockedCyclicSectorFamily.blockTensor_eq_commonReindexedBlock_of_word_eq,
		MPSTensor.CommonBlockedCyclicSectorFamily.blockTensor_eq_commonReindexedBlock_of_groupedBlockCastAgrees,
		MPSTensor.CommonBlockedCyclicSectorFamily.blockTensor_sameMPV₂_commonReindexedBlock_of_word_eq,
		MPSTensor.CommonBlockedCyclicSectorFamily.blockTensor_sameMPV₂_commonReindexedBlock_of_groupedBlockCastAgrees,
		MPSTensor.CommonBlockedCyclicSectorFamily.sameMPV₂_weightedCanonicalBlock_commonReindexedBlock_of_blockwise,
		MPSTensor.CommonBlockedCyclicSectorFamily.sameMPV₂_weightedCanonicalBlock_commonReindexedBlock_of_word_eq,
		MPSTensor.CommonBlockedCyclicSectorFamily.sameMPV₂_weightedCanonicalBlock_commonReindexedBlock_of_groupedBlockCastAgrees,
		MPSTensor.CommonBlockedCyclicSectorFamily.sameMPV₂_weightedCanonicalBlock_commonFlat_of_blockwise,
		MPSTensor.CommonBlockedCyclicSectorFamily.sameMPV₂_weightedCanonicalBlock_commonFlat_of_word_eq,
		MPSTensor.CommonBlockedCyclicSectorFamily.sameMPV₂_weightedCanonicalBlock_commonFlat_of_groupedBlockCastAgrees}
	\leanok
	\uses{def:common_blocked_cyclic_sector_derived_blocks,
		def:common_blocked_cyclic_sector_grouped_block_cast,
		thm:common_blocked_cyclic_sector_reindexed_samempv}
	Internal reindexing lemmas: direct blocking of each nonzero block has the
	same MPV family as its iterated-blocking reindexed form, and the weighted
	direct sum agrees with the common-sector family.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:mpv_decomposition,
		def:common_blocked_cyclic_sector_grouped_block_cast,
		thm:common_blocked_cyclic_sector_reindexed_samempv}
	The word equality gives equality of block tensors, hence blockwise MPV
	equality.  Expanding both weighted block-diagonal tensors as finite sums and
	composing with the flattening theorem gives the common-sector statement.
\end{proof}

\begin{theorem}[Blocked nonzero part on the common alphabet (internal)]%
\label{thm:common_blocked_cyclic_sector_reindexed_nonzero_part}
	\lean{MPSTensor.CommonBlockedCyclicSectorFamily.sameMPV₂_weightedCanonicalBlock_commonFlat_of_reindexed,
		MPSTensor.CommonBlockedCyclicSectorFamily.sameMPV₂_weightedCanonicalBlock_commonFlatAt_of_reindexed}
	\leanok
	\uses{thm:common_blocked_cyclic_sector_reindexed_samempv,
		def:common_blocked_cyclic_sector_derived_blocks}
	Internal lemma: if the canonical blocked nonzero part agrees with the
	iterated-blocking cyclic-sector tensor as an MPV family, then the
	weighted nonzero part agrees with the common-sector tensor.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:common_blocked_cyclic_sector_reindexed_samempv}
	Compose the assumed MPV equality with the sector decomposition theorem.
\end{proof}

\begin{theorem}[Common reblocking at a prescribed common multiple]%
\label{thm:exists_common_blocked_cyclic_sector_family_common_multiple}
	\lean{MPSTensor.exists_commonBlockedCyclicSectorFamily_of_commonMultiple}
	\leanok
	\uses{def:common_blocked_cyclic_sector_family,
		def:primitive_irreducible_cyclic_sectors,
		thm:tp_primitive_irreducible_extra_blocking,
		thm:samempv_blocking_distributes}
	If every nonzero-weight block has primitive irreducible cyclic sectors and a
	positive integer $p$ is a common multiple of all period-removal lengths, then
	there exists a \texttt{CommonBlockedCyclicSectorFamily} at the prescribed common
	length $p$.  This packages the iterated-blocking construction from period-removal
	sectors to a shared alphabet; the downstream pipeline then selects the least
	common multiple.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:tp_primitive_irreducible_extra_blocking,
		thm:samempv_blocking_distributes,
		thm:iterated_blocking_phys_dim,
		thm:samempv_cast_phys_dim,
		thm:tensor_from_blocks_cast_phys_dim,
		thm:left_canonical_cast_phys_dim,
		thm:primitive_transfer_cast_phys_dim,
		thm:irreducible_tensor_cast_phys_dim}
	For each block with period $m_k$, write the prescribed common multiple as
	$p=m_k e_k$.  Then block every cyclic sector by $e_k$, substitute the equality
	between the iterated and common physical alphabets, and transport the MPV and
	structural properties exactly as in the least-common-multiple construction.
\end{proof}

\begin{theorem}[Common reblocking of per-block cyclic sectors]%
\label{thm:exists_common_blocked_cyclic_sector_family}
	\lean{MPSTensor.exists_commonBlockedCyclicSectorFamily_of_hasPrimitiveIrreducibleCyclicSectors}
	\leanok
	\uses{thm:exists_common_blocked_cyclic_sector_family_common_multiple}
	If every nonzero-weight block in a finite family has primitive irreducible cyclic
	sectors, then the family admits a common reblocked cyclic-sector family.
	The proof chooses the least common multiple of the per-block periods, blocks
	each sector by the corresponding quotient, transports all sectors to the common
	physical alphabet, and preserves TP, primitivity, tensor irreducibility,
	positive bond dimensions, and the unit-weight MPV equivalences.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:exists_common_blocked_cyclic_sector_family_common_multiple}
	Choose the least common multiple $p$ of the per-block periods and apply
	Theorem~\ref{thm:exists_common_blocked_cyclic_sector_family_common_multiple}.
\end{proof}

\subsection{Sector irreducibility lemmas}%
\label{sec:sector_irreducibility_lemmas}

\begin{lemma}[Pairwise orthogonality from projection sum]%
	\label{lem:pairwise_ortho_proj_sum}
	\lean{pairwise_mul_zero_of_orthogonalProjection_sum_one}
	\leanok
	Let $(P_k)_{k=1}^m$ be orthogonal projections summing to the identity:
	$\sum_k P_k = \Id$.
	Then for $i \neq j$ we have $P_i P_j = 0$.
\end{lemma}

\begin{proof}\leanok
	Sandwich the sum between $P_i$ to get
	$\sum_k P_i P_k P_i = P_i$.
	Since the $i$-th summand equals $P_i$ by idempotence,
	all other summands vanish.
	Positivity ($B B^* = 0 \Rightarrow B = 0$) then gives
	$P_i P_k = 0$ for each $k \neq i$.
\end{proof}

\begin{lemma}[Corner preservation from adjoint-fixed projection]%
	\label{lem:corner_preserv_adjoint_fixed}
	\lean{preservesCorner_of_adjoint_fixed_projection}
	\leanok
	\uses{thm:kraus_commute}
	Let $A$ be a left-canonical MPS tensor and let $P$ be an orthogonal
	projection fixed by $\E_A^\dagger$.
	Then $\E_A^\dagger$ preserves the corner algebra
	$P \MN{D} P$.
\end{lemma}

\begin{proof}\leanok
	\uses{thm:kraus_commute}
	The fixed-point condition and the Kadison--Schwarz equality
	imply $[P, A_i] = 0$ for every Kraus operator.
	Thread the idempotent relation $P^2 = P$ through the corner sandwich.
\end{proof}

\begin{lemma}[Orbit-sum fixed point]%
	\label{lem:orbit_sum_fixed}
	\lean{MPSTensor.orbitSumProjection_fixed_of_pow_fix}
	\leanok
	Let $T$ be a linear endomorphism and $Q$ a matrix satisfying $T^m(Q) = Q$.
	Then the orbit sum $\sum_{l=0}^{m-1} T^l(Q)$ is a fixed point of $T$.
\end{lemma}

\begin{proof}\leanok
	Telescoping: $T$ shifts each summand $T^l(Q) \mapsto T^{l+1}(Q)$
	and the periodicity condition $T^m(Q) = Q$ wraps the last term
	$T^{m-1}(Q) \mapsto T^m(Q) = Q$ back to the first.
\end{proof}

\begin{lemma}[Orbit iterate supported on shifted sector]%
	\label{lem:orbit_iterate_shifted_sector}
	\lean{MPSTensor.orbit_iterate_supported_on_shifted_sector}
	\leanok
	Let $(P_k)_{k=1}^m$ be orthogonal projections with cyclic shift
	$T(P_{k+1}) = P_k$, and suppose $T$ is multiplicative on each sector
	corner.
	If $Q$ is supported on sector $P_k$ (i.e.\ $QP_k = Q = P_k Q$), then
	for each $0 \le l < m$ the iterate $T^l(Q)$ is supported on the
	shifted sector $P_{k-l}$.
\end{lemma}

\begin{proof}\leanok
	Induction on $l$.  The base case is the hypothesis $QP_k = Q = P_k Q$.
	For the inductive step, the multiplicative-domain identities and the
	cyclic shift relation $T(P_j) = P_{j-1}$ transfer the support from
	sector $P_j$ to sector $P_{j-1}$.
\end{proof}

\begin{lemma}[Orbit iterate preserves orthogonal projection]%
	\label{lem:orbit_iterate_proj}
	\lean{MPSTensor.orbit_iterate_isOrthogonalProjection}
	\leanok
	\uses{lem:orbit_iterate_shifted_sector}
	Under the same cyclic shift and multiplicative-domain hypotheses as
	Lemma~\ref{lem:orbit_iterate_shifted_sector}, suppose additionally
	that $T$ maps every orthogonal projection supported on a single
	sector to an orthogonal projection.
	If $Q$ is an orthogonal projection supported on $P_k$, then
	$T^l(Q)$ is an orthogonal projection for each $0 \le l < m$.
\end{lemma}

\begin{proof}\leanok
	\uses{lem:orbit_iterate_shifted_sector}
	Induction on $l$.  The base case is the hypothesis on $Q$.
	At each step, Lemma~\ref{lem:orbit_iterate_shifted_sector}
	guarantees that $T^n(Q)$ is supported on a single sector,
	so the one-step preservation hypothesis applies.
\end{proof}

\begin{lemma}[Orbit sum of a full sector projection]%
	\label{lem:orbit_sum_full_sector}
	\lean{MPSTensor.orbitSumProjection_eq_one_of_full_sector}
	\leanok
	\uses{lem:orbit_sum_fixed}
	Let $(P_k)_{k=1}^m$ be orthogonal projections with
	$\sum_k P_k = \Id$ and cyclic shift $T(P_{k+1}) = P_k$.
	Then for every sector index $k$,
	\[
		\sum_{l=0}^{m-1} T^l(P_k) \;=\; \Id.
	\]
\end{lemma}

\begin{proof}\leanok
	By induction, $T^l(P_k) = P_{k-l}$ for each $l$.
	The orbit sum therefore runs over all distinct sectors, and the
	reindexing $l \mapsto k - l$ recovers $\sum_j P_j = \Id$.
\end{proof}

\begin{theorem}[Sector irreducibility from orbit-sum lift]%
	\label{thm:sector_irred_orbit_lift}
	\lean{MPSTensor.isIrreducibleOnCorner_of_cyclic_decomp_mps_of_hLift}
	\leanok
	\uses{thm:cyclic_sector_decomp_irr_tp}
	Let $A$ be a left-canonical MPS tensor with irreducible adjoint transfer
	map, and let $(P_k)_{k=1}^m$ be the cyclic sector projections.
	Suppose the orbit-sum lift hypothesis holds: for every sector $k$ and
	every subprojection $Q \le P_k$ fixed by $(\E_A^\dagger)^m$, there exists
	a global projection $R$ that is invariant under $\E_A^\dagger$ and
	faithfully reflects triviality ($Q = 0 \Leftrightarrow R = 0$,
	$Q = P_k \Leftrightarrow R = \Id$).
	Then $(\E_A^\dagger)^m$ is irreducible on each corner $P_k$.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:cyclic_sector_decomp_irr_tp}
	Direct application of the cyclic-decomposition irreducibility
	theorem from spectral theory.
\end{proof}

\begin{lemma}[Fixed-point upgrade from irreducibility and trace preservation]%
	\label{lem:fix_upgrade_peripheral}
	\lean{MPSTensor.hFixUpgrade_of_peripheral}
	\leanok
	\uses{thm:psd_fp_exists, thm:psd_fp_pd_irred,
	      lem:trace_mul_map_eq_trace_adjointMap_mul,
	      lem:posSemidef_eq_zero_of_posDef_trace_mul_eq_zero}
	Let $A$ be a left-canonical MPS tensor whose transfer map $\E_A$ is
	irreducible, and let $Q$ be an orthogonal projection such that the corner
	$Q \mathcal M_D Q$ is preserved by $((\E_A^\dagger)^m)$. Then
	$((\E_A^\dagger)^m)(Q) = Q$.
\end{lemma}

\begin{proof}\leanok
	\uses{thm:psd_fp_exists, thm:psd_fp_pd_irred,
	      lem:trace_mul_map_eq_trace_adjointMap_mul,
	      lem:posSemidef_eq_zero_of_posDef_trace_mul_eq_zero}
	Let $\rho > 0$ be the distinguished fixed point of $\E_A$. Since
	$Q \mathcal M_D Q$ is preserved, the difference
	$Q - (\E_A^\dagger)^m(Q)$ can be rewritten as
	$Q\,(\E_A^\dagger)^m(\Id - Q)\,Q$, hence it is positive semidefinite.
	The weighted trace pairing with $\rho$ is invariant under $(\E_A^\dagger)^m$,
	so this positive semidefinite matrix has weighted trace zero. Faithfulness of
	the $\rho$-weighted trace therefore forces it to vanish, giving the claim.
\end{proof}

\begin{definition}[Sector fixed-point-algebra rigidity]%
	\label{def:sector_fixed_point_algebra_rigidity}
	\lean{MPSTensor.SectorFixedPointAlgebraRigidity}
	\leanok
	Let $T$ be the adjoint transfer map and $(P_k)_{k=1}^m$ a cyclic family of
	orthogonal sector projections. We say that the sector dynamics satisfies
	fixed-point-algebra rigidity if, for each $k$, the one-step map $T$ is
	multiplicative on the $T^m$-fixed operators supported on the corner
	$P_k \mathcal M_D P_k$: whenever
	\[
		X = X P_k = P_k X,
		\qquad
		Y = Y P_k = P_k Y,
		\qquad
		T^m(X) = X,
		\qquad
		T^m(Y) = Y,
	\]
	one has
	\[
		T(XY) = T(X) T(Y).
	\]
\end{definition}

\begin{remark}\leanok
	This is the abstract multiplicativity property used in the orbit-sum
	construction. Theorem~\ref{thm:sector_fixed_algebra_irred_cyclic} shows
	that it is automatic in the irreducible trace-preserving cyclic setting.
\end{remark}

\begin{lemma}[Projection step from sector fixed-point-algebra rigidity]%
	\label{lem:sector_fixed_algebra_proj_step}
	\lean{MPSTensor.hProjStep_of_sectorFixedPointAlgebraRigidity}
	\leanok
	\uses{def:sector_fixed_point_algebra_rigidity}
	Assume sector fixed-point-algebra rigidity. If $X$ is an orthogonal
	projection supported on $P_k$ and fixed by $T^m$, then $T(X)$ is again an
	orthogonal projection.
\end{lemma}

\begin{proof}\leanok
	\uses{def:sector_fixed_point_algebra_rigidity}
	Kraus maps preserve conjugate transpose, so $T(X)$ is Hermitian. The
	rigidity hypothesis gives
	\[
		T(X)^2 = T(X^2) = T(X),
	\]
	since $X^2 = X$ for an orthogonal projection. Therefore $T(X)$ is again an
	orthogonal projection.
\end{proof}

\begin{theorem}[Sector fixed-point-algebra rigidity from irreducibility]%
	\label{thm:sector_fixed_algebra_irred_cyclic}
	\lean{MPSTensor.sectorFixedPointAlgebraRigidity_of_irreducible_cyclicDecomp}
	\leanok
	\uses{def:sector_fixed_point_algebra_rigidity, lem:orbit_sum_fixed}
	Let $A$ be a left-canonical MPS tensor with irreducible adjoint transfer map,
	and let $(P_k)_{k=1}^m$ be cyclic orthogonal sector projections with
	$\sum_k P_k = \Id$. Then the sector fixed-point-algebra rigidity property
	holds automatically for $\E_A^\dagger$.
\end{theorem}

\begin{proof}\leanok
	\uses{def:sector_fixed_point_algebra_rigidity, lem:orbit_sum_fixed}
	Fix $k$ and let $X,Y$ be operators supported on $P_k$ and fixed by
	$(\E_A^\dagger)^m$. For such an $X$, the orbit sum
	\[
		R_X := \sum_{l=0}^{m-1} (\E_A^\dagger)^l(X)
	\]
	is fixed by $\E_A^\dagger$ by Lemma~\ref{lem:orbit_sum_fixed}. Since the
	trace-preserving hypothesis makes $\E_A^\dagger$ unital, the irreducible
	unital fixed-point theorem implies that $R_X$ is a scalar multiple of
	$\Id$. Compressing by $P_k$ recovers $X$, so $X = c P_k$; the same argument
	gives $Y = c' P_k$. Therefore
	\[
		\E_A^\dagger(XY) = \E_A^\dagger((cc') P_k) = (cc') P_{k-1}
		= \E_A^\dagger(X)\E_A^\dagger(Y),
	\]
	using the cyclic relation $\E_A^\dagger(P_k)=P_{k-1}$.
\end{proof}

\begin{theorem}[Sector fixed-point-algebra rigidity for trace-preserving tensors]%
	\label{thm:sector_fixed_algebra_irred_tp}
	\lean{MPSTensor.sectorFixedPointAlgebraRigidity_of_irreducible_tp}
	\leanok
	\uses{def:sector_fixed_point_algebra_rigidity, thm:sector_fixed_algebra_irred_cyclic}
	Let $A$ be an irreducible trace-preserving MPS tensor, and let
	$(P_k)_{k=1}^m$ be cyclic orthogonal sector projections for
	$\E_A^\dagger$ with $\sum_k P_k = \Id$. Then the sector
	fixed-point-algebra rigidity property holds for $\E_A^\dagger$.
\end{theorem}

\begin{proof}\leanok
	\uses{def:sector_fixed_point_algebra_rigidity, thm:md_characterization,
		thm:sector_fixed_algebra_irred_cyclic}
	The trace-preserving hypothesis makes $\E_A^\dagger$ unital. The cyclic
	shift relation and Kadison--Schwarz equality place each sector projection in
	the multiplicative domain of the adjoint Kraus family, by
	Theorem~\ref{thm:md_characterization}. Hence multiplication by a sector
	projection is preserved by $\E_A^\dagger$.
	Theorem~\ref{thm:sector_fixed_algebra_irred_cyclic} then gives the sector
	fixed-point-algebra rigidity property.
\end{proof}

\begin{theorem}[Sector fixed-point-algebra rigidity from scalar blocked fixed points]%
	\label{thm:sector_fixed_algebra_scalar_blocked}
	\lean{MPSTensor.sectorFixedPointAlgebraRigidity_of_cyclic_decomp_after_blocking_of_scalarBlockedFixedPoints}
	\leanok
	\uses{def:sector_fixed_point_algebra_rigidity}
	Let $A$ be a left-canonical MPS tensor and let $(P_k)_{k=1}^m$ be cyclic
	orthogonal sector projections with $\sum_k P_k = \Id$.
	Suppose that for each $k$ and each element $X \in P_k \MN{D} P_k$ satisfying
	$\E_{C_k}^\dagger(X) = X$ (where $C_k$ is the blocked sector tensor), there
	exists $c \in \mathbb{C}$ with $X = c \cdot \Id$.
	Then the sector fixed-point-algebra rigidity property holds for
	$\E_A^\dagger$.
\end{theorem}

\begin{proof}\leanok
	\uses{def:sector_fixed_point_algebra_rigidity}
	Fix $k$ and let $X, Y$ be operators supported on $P_k$ and fixed by
	$(\E_A^\dagger)^m$. Lift each to the corresponding compressed sector via the
	intertwining equivalence $\varphi_k$; the scalar fixed-point hypothesis yields
	$X = c_X P_k$ and $Y = c_Y P_k$. Then
	\[
		\E_A^\dagger(XY) = \E_A^\dagger((c_X c_Y) P_k) = (c_X c_Y) P_{k-1}
		= \E_A^\dagger(X)\E_A^\dagger(Y),
	\]
	using the cyclic relation $\E_A^\dagger(P_k) = P_{k-1}$ and the linearity of
	$\E_A^\dagger$.
\end{proof}

\begin{lemma}[Orbit-sum construction of the sector lift]%
	\label{lem:orbit_sum_hLift_construction}
	\lean{MPSTensor.hLift_cyclicDecomp_mps_of_fixUpgrade}
	\leanok
	\uses{lem:orbit_sum_fixed, lem:orbit_sum_full_sector,
	      thm:sector_irred_orbit_lift}
	Let $A$ be a left-canonical MPS tensor, $(P_k)_{k=1}^m$ a cyclic family
	of orthogonal sector projections with $\sum_k P_k = \Id$, and set
	$T = \E_A^\dagger$. Assume two auxiliary facts, both discharged from the
	block-diagonal canonical-form analysis of
	\cite{DeLasCuevas2017Irreducible}:
	\begin{enumerate}
		\item \emph{(Projection step.)} For each sector index $k$ and each
			orthogonal projection $X$ supported on $P_k$, the image $T(X)$
			is again an orthogonal projection.
		\item \emph{(Fixed-point upgrade.)} For each sector index $k$ and
			each orthogonal subprojection $Q \le P_k$, if $Q$ is a corner
			fixed point of $T^m$ then $T^m(Q) = Q$.
	\end{enumerate}
	Then for every sector $k$ and every orthogonal subprojection $Q \le P_k$
	which is a corner fixed point of $T^m$, the orbit sum
	\[
		R \;:=\; \sum_{l=0}^{m-1} T^l(Q)
	\]
	is an orthogonal projection, preserved by $T$, with
	$Q = 0 \Leftrightarrow R = 0$ and $Q = P_k \Leftrightarrow R = \Id$.
	In particular $R$ witnesses the orbit-sum lift hypothesis of
	Theorem~\ref{thm:sector_irred_orbit_lift}.
\end{lemma}

\begin{proof}\leanok
	\uses{lem:orbit_sum_fixed, lem:orbit_sum_full_sector,
	      thm:sector_irred_orbit_lift}
	The iterates $T^l(Q)$ are each orthogonal projections (repeatedly using
	the projection-step hypothesis), supported on the shifted sector
	$P_{k-l}$, and are therefore pairwise orthogonal because distinct
	sectors are. This gives both the Hermitian and idempotent conjuncts for
	$R$. Fixedness under $T$ follows from $T^m(Q) = Q$ and
	Lemma~\ref{lem:orbit_sum_fixed}, and corner preservation from the
	adjoint trace-preserving relation. The zero equivalence uses
	$R\cdot Q = Q$ (the diagonal term $l=0$ is $Q$, off-diagonal terms
	vanish by pairwise orthogonality); the full-sector equivalence uses
	$P_k \cdot R \cdot P_k = Q$ and
	Lemma~\ref{lem:orbit_sum_full_sector}.
\end{proof}

\begin{lemma}[Orbit-sum construction, projection-step variant]%
	\label{lem:orbit_sum_hLift_proj_step}
	\lean{MPSTensor.hLift_cyclicDecomp_mps_of_projStep}
	\leanok
	\uses{lem:fix_upgrade_peripheral, lem:orbit_sum_hLift_construction}
	Let $A$ be a left-canonical MPS tensor with irreducible adjoint transfer
	map, and let $(P_k)_{k=1}^m$ be cyclic orthogonal sector projections.
	Assume the projection-step hypothesis of
	Lemma~\ref{lem:orbit_sum_hLift_construction}. Then the orbit-sum lift holds:
	for every orthogonal subprojection $Q \le P_k$ whose corner is preserved by
	$(\E_A^\dagger)^m$, there is a global projection $R$ invariant under
	$\E_A^\dagger$ with the same zero/full-sector tests.
\end{lemma}

\begin{proof}\leanok
	\uses{lem:fix_upgrade_peripheral, lem:orbit_sum_hLift_construction}
	Combine the original orbit-sum construction with
	Lemma~\ref{lem:fix_upgrade_peripheral}, which supplies the fixed-point
	upgrade from irreducibility and trace preservation. The irreducibility
	hypothesis needed there is obtained by transporting irreducibility from the
	adjoint transfer map to the tensor and then back to the primal transfer map.
\end{proof}

\begin{lemma}[Orbit-sum construction, fixed-point-algebra variant]%
	\label{lem:orbit_sum_hLift_fixed_algebra}
	\lean{MPSTensor.hLift_cyclicDecomp_mps_of_sectorFixedPointAlgebraRigidity}
	\leanok
	\uses{def:sector_fixed_point_algebra_rigidity,
	      lem:fix_upgrade_peripheral,
	      lem:sector_fixed_algebra_proj_step,
	      lem:orbit_sum_hLift_construction}
	Let $A$ be a left-canonical MPS tensor with irreducible adjoint transfer map,
	and let $(P_k)_{k=1}^m$ be cyclic orthogonal sector projections. Assume the
	single fixed-point-algebra rigidity hypothesis from
	Definition~\ref{def:sector_fixed_point_algebra_rigidity}. Then the orbit-sum
	lift holds: for every orthogonal subprojection $Q \le P_k$ whose corner is
	preserved by $(\E_A^\dagger)^m$, there is a global projection $R$ invariant
	under $\E_A^\dagger$ with the same zero/full-sector tests.
\end{lemma}

\begin{proof}\leanok
	\uses{def:sector_fixed_point_algebra_rigidity,
	      lem:fix_upgrade_peripheral,
	      lem:sector_fixed_algebra_proj_step,
	      lem:orbit_sum_hLift_construction}
	First apply Lemma~\ref{lem:fix_upgrade_peripheral} to upgrade corner
	preservation of $Q$ under $(\E_A^\dagger)^m$ to an actual fixed-point
	relation. The rigidity hypothesis then yields the one-step projection
	preservation needed along the orbit of $Q$, via
	Lemma~\ref{lem:sector_fixed_algebra_proj_step}. With those two facts in hand,
	the original orbit-sum construction from
	Lemma~\ref{lem:orbit_sum_hLift_construction} applies verbatim. In the
	irreducible trace-preserving cyclic case, the rigidity hypothesis here is
	discharged by Theorem~\ref{thm:sector_fixed_algebra_irred_cyclic}.
\end{proof}

\begin{theorem}[Orbit-sum construction, unconditional variant]%
	\label{thm:orbit_sum_hLift_unconditional}
	\lean{MPSTensor.hLift_cyclicDecomp_mps}
	\leanok
	\uses{lem:orbit_sum_hLift_fixed_algebra, thm:sector_fixed_algebra_irred_cyclic}
	Let $A$ be a left-canonical MPS tensor with irreducible adjoint transfer map,
	and let $(P_k)_{k=1}^m$ be cyclic orthogonal sector projections. Then for
	every orthogonal subprojection $Q \le P_k$ whose corner is preserved by
	$(\E_A^\dagger)^m$, there is a global projection $R$ invariant under
	$\E_A^\dagger$ with the same zero/full-sector tests.
\end{theorem}

\begin{proof}\leanok
	\uses{lem:orbit_sum_hLift_fixed_algebra, thm:sector_fixed_algebra_irred_cyclic}
	Apply Lemma~\ref{lem:orbit_sum_hLift_fixed_algebra} with the rigidity
	hypothesis supplied by
	Theorem~\ref{thm:sector_fixed_algebra_irred_cyclic}.
\end{proof}

\begin{theorem}[Sector irreducibility from the orbit-sum construction]%
	\label{thm:sector_irred_fix_upgrade}
	\lean{MPSTensor.isIrreducibleOnCorner_of_cyclic_decomp_mps_of_fixUpgrade}
	\leanok
	\uses{thm:sector_irred_orbit_lift, lem:orbit_sum_hLift_construction}
	Let $A$ be a left-canonical MPS tensor with irreducible adjoint transfer
	map, and let $(P_k)_{k=1}^m$ be cyclic orthogonal sector projections.
	Assume the projection-step and fixed-point-upgrade hypotheses of
	Lemma~\ref{lem:orbit_sum_hLift_construction}. Then $(\E_A^\dagger)^m$
	is irreducible on each corner $P_k$.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:sector_irred_orbit_lift, lem:orbit_sum_hLift_construction}
	Apply Lemma~\ref{lem:orbit_sum_hLift_construction} to obtain the orbit-sum
	lift, then apply Theorem~\ref{thm:sector_irred_orbit_lift}.
\end{proof}

\begin{theorem}[Sector irreducibility from the projection-step hypothesis]%
	\label{thm:sector_irred_proj_step}
	\lean{MPSTensor.isIrreducibleOnCorner_of_cyclic_decomp_mps_of_projStep}
	\leanok
	\uses{thm:sector_irred_orbit_lift, lem:orbit_sum_hLift_proj_step}
	Let $A$ be a left-canonical MPS tensor with irreducible adjoint transfer
	map, and let $(P_k)_{k=1}^m$ be cyclic orthogonal sector projections.
	Assume only the projection-step hypothesis from
	Lemma~\ref{lem:orbit_sum_hLift_proj_step}. Then $(\E_A^\dagger)^m$
	is irreducible on each corner $P_k$.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:sector_irred_orbit_lift, lem:orbit_sum_hLift_proj_step}
	Apply Lemma~\ref{lem:orbit_sum_hLift_proj_step} and then
	Theorem~\ref{thm:sector_irred_orbit_lift}.
\end{proof}

\begin{theorem}[Sector irreducibility from fixed-point-algebra rigidity]%
	\label{thm:sector_irred_fixed_algebra}
	\lean{MPSTensor.isIrreducibleOnCorner_of_cyclic_decomp_mps_of_sectorFixedPointAlgebraRigidity}
	\leanok
	\uses{thm:sector_irred_orbit_lift, lem:orbit_sum_hLift_fixed_algebra}
	Let $A$ be a left-canonical MPS tensor with irreducible adjoint transfer map,
	and let $(P_k)_{k=1}^m$ be cyclic orthogonal sector projections. Assume only
	the fixed-point-algebra rigidity hypothesis from
	Definition~\ref{def:sector_fixed_point_algebra_rigidity}. Then
	$(\E_A^\dagger)^m$ is irreducible on each corner $P_k$.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:sector_irred_orbit_lift, lem:orbit_sum_hLift_fixed_algebra}
	Apply Lemma~\ref{lem:orbit_sum_hLift_fixed_algebra} and then
	Theorem~\ref{thm:sector_irred_orbit_lift}.
\end{proof}

\begin{theorem}[Unconditional sector irreducibility]%
	\label{thm:sector_irred_unconditional}
	\lean{MPSTensor.isIrreducibleOnCorner_of_cyclic_decomp_mps}
	\leanok
	\uses{thm:sector_irred_fixed_algebra, thm:sector_fixed_algebra_irred_cyclic}
	Let $A$ be a left-canonical MPS tensor with irreducible adjoint transfer map,
	and let $(P_k)_{k=1}^m$ be cyclic orthogonal sector projections. Then
	$(\E_A^\dagger)^m$ is irreducible on each corner $P_k$.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:sector_irred_fixed_algebra, thm:sector_fixed_algebra_irred_cyclic}
	Apply Theorem~\ref{thm:sector_irred_fixed_algebra} with the rigidity
	hypothesis supplied by
	Theorem~\ref{thm:sector_fixed_algebra_irred_cyclic}.
\end{proof}

\section{Reduction to primitive blocks}%
\label{sec:reduction_to_normal_form}

\begin{theorem}[Projection-step hypothesis yields primitive blocked sectors]%
	\label{thm:blocked_sector_proj_step}
	\lean{MPSTensor.primitive_and_irreducible_sectorBlocks_of_cyclic_decomp_after_blocking_of_projStep}
	\leanok
	\uses{thm:cyclic_sector_decomp_after_blocking,
	      thm:sector_irred_proj_step,
	      thm:primitive_adjoint_equiv,
	      lem:orbit_sum_hLift_proj_step}
	Let $A$ be a left-canonical irreducible tensor whose adjoint transfer map has
	peripheral spectrum $\{\gamma^j : 0 \le j < m\}$ for a primitive $m$-th root
	of unity $\gamma$. Let $(C_u)_{u=1}^m$, $(P_u)_{u=1}^m$, and
	$\varphi_u : \MN{\dim C_u} \xrightarrow{\sim} P_u \MN{D} P_u$
	be the blocked cyclic-sector data. If the one-step projection-preservation
	hypothesis of Lemma~\ref{lem:orbit_sum_hLift_proj_step} holds, then every
	blocked sector tensor $C_u$ has primitive transfer map and is irreducible.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:sector_irred_proj_step,
	      thm:primitive_adjoint_equiv}
	Theorem~\ref{thm:sector_irred_proj_step} gives irreducibility of
	$(\E_A^\dagger)^m$ on each corner $P_u$.
	The blocked cyclic-sector data provide compression equivalences
	$\varphi_u$ intertwining the adjoint transfer map of $C_u$ with the
	corresponding corner restriction of $(\E_A^\dagger)^m$.
	The same cyclic peripheral-spectrum argument that produces the sector
	decomposition shows that each corner restriction is primitive.
	Transport primitivity and irreducibility across $\varphi_u$, then use
	Theorem~\ref{thm:primitive_adjoint_equiv} to pass from the adjoint blocked
	map to the primal transfer map of $C_u$.
\end{proof}

\begin{theorem}[Fixed-point-algebra rigidity yields primitive blocked sectors]%
	\label{thm:blocked_sector_fixed_algebra}
	\lean{MPSTensor.primitive_and_irreducible_sectorBlocks_of_cyclic_decomp_after_blocking_of_fixedAlgebraRigidity}
	\leanok
	\uses{def:sector_fixed_point_algebra_rigidity,
	      thm:cyclic_sector_decomp_after_blocking,
	      thm:sector_irred_fixed_algebra,
	      thm:primitive_adjoint_equiv}
	Let $A$ be a left-canonical irreducible tensor whose adjoint transfer map has
	peripheral spectrum $\{\gamma^j : 0 \le j < m\}$ for a primitive $m$-th root
	of unity $\gamma$. Let $(C_u)_{u=1}^m$, $(P_u)_{u=1}^m$, and
	$\varphi_u : \MN{\dim C_u} \xrightarrow{\sim} P_u \MN{D} P_u$
	be the blocked cyclic-sector data. If the fixed-point-algebra rigidity
	hypothesis from Definition~\ref{def:sector_fixed_point_algebra_rigidity}
	holds, then every blocked sector tensor $C_u$ has primitive transfer map and
	is irreducible.
\end{theorem}

\begin{proof}\leanok
	\uses{def:sector_fixed_point_algebra_rigidity,
	      thm:sector_irred_fixed_algebra,
	      thm:primitive_adjoint_equiv}
	Theorem~\ref{thm:sector_irred_fixed_algebra} gives irreducibility of
	$(\E_A^\dagger)^m$ on each corner $P_u$.
	The blocked cyclic-sector data provide compression equivalences
	$\varphi_u$ intertwining the adjoint transfer map of $C_u$ with the
	corresponding corner restriction of $(\E_A^\dagger)^m$.
	The same cyclic peripheral-spectrum argument that produces the sector
	decomposition shows that each corner restriction is primitive.
	Transport primitivity and irreducibility across $\varphi_u$, then use
	Theorem~\ref{thm:primitive_adjoint_equiv} to pass from the adjoint blocked
	map to the primal transfer map of $C_u$.
\end{proof}

\begin{theorem}[Scalar blocked fixed points yield primitive blocked sectors]%
	\label{thm:blocked_sector_scalar_fixed}
	\lean{MPSTensor.primitive_and_irreducible_sectorBlocks_of_cyclic_decomp_after_blocking_of_scalarBlockedFixedPoints}
	\leanok
	\uses{def:sector_fixed_point_algebra_rigidity,
	      thm:sector_fixed_algebra_scalar_blocked,
	      thm:blocked_sector_fixed_algebra,
	      thm:cyclic_sector_decomp_after_blocking,
	      thm:primitive_adjoint_equiv}
	Let $A$ be a left-canonical irreducible tensor whose adjoint transfer map has
	peripheral spectrum $\{\gamma^j : 0 \le j < m\}$ for a primitive $m$-th root
	of unity $\gamma$. Let $(C_u)_{u=1}^m$, $(P_u)_{u=1}^m$, and
	$\varphi_u : \MN{\dim C_u} \xrightarrow{\sim} P_u \MN{D} P_u$
	be the blocked cyclic-sector data. If every $(\E_{C_u}^\dagger)$-fixed element
	in the compressed sector $P_u \MN{D} P_u$ is a scalar multiple of the identity,
	then every blocked sector tensor $C_u$ has primitive transfer map and is
	irreducible.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:sector_fixed_algebra_scalar_blocked,
	      thm:blocked_sector_fixed_algebra}
	Apply Theorem~\ref{thm:sector_fixed_algebra_scalar_blocked} to obtain
	sector fixed-point-algebra rigidity from the scalar fixed-point hypothesis,
	then apply Theorem~\ref{thm:blocked_sector_fixed_algebra}.
\end{proof}

\begin{theorem}[Unconditional primitive blocked sectors]%
	\label{thm:blocked_sector_unconditional}
	\lean{MPSTensor.primitive_and_irreducible_sectorBlocks_of_cyclic_decomp_after_blocking}
	\leanok
	\uses{thm:cyclic_sector_decomp_after_blocking,
	      thm:sector_irred_unconditional,
	      thm:primitive_adjoint_equiv}
	Let $A$ be a left-canonical irreducible tensor whose adjoint transfer map has
	peripheral spectrum $\{\gamma^j : 0 \le j < m\}$ for a primitive $m$-th root
	of unity $\gamma$. Let $(C_u)_{u=1}^m$, $(P_u)_{u=1}^m$, and
	$\varphi_u : \MN{\dim C_u} \xrightarrow{\sim} P_u \MN{D} P_u$
	be the blocked cyclic-sector data. Then every blocked sector tensor $C_u$ has
	primitive transfer map and is irreducible.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:sector_irred_unconditional,
	      thm:primitive_adjoint_equiv}
	Theorem~\ref{thm:sector_irred_unconditional} gives irreducibility of
	$(\E_A^\dagger)^m$ on each corner $P_u$.
	The blocked cyclic-sector data provide compression equivalences
	$\varphi_u$ intertwining the adjoint transfer map of $C_u$ with the
	corresponding corner restriction of $(\E_A^\dagger)^m$.
	The same cyclic peripheral-spectrum argument that produces the sector
	decomposition shows that each corner restriction is primitive.
	Transport primitivity and irreducibility across $\varphi_u$, then use
	Theorem~\ref{thm:primitive_adjoint_equiv} to pass from the adjoint blocked
	map to the primal transfer map of $C_u$.
\end{proof}

\begin{theorem}[Arbitrary tensor to primitive block decomposition]%
	\label{thm:tp_primitive_blockdecomp}
	\lean{MPSTensor.exists_tp_primitive_blockDecomp_after_blocking}
	\leanok
	\uses{def:blocked_tensor, def:zero_mps_tensor}
	For any MPS tensor $A$, there exist a period $p \ge 1$,
	a zero-tail bond dimension $D_0$, nonzero weights $(\mu_k)_{k=1}^r$,
	and left-canonical blocks $(B_k)_{k=1}^r$ with primitive transfer
	maps and positive bond dimensions, such that
	\[
		V^{(N)}\!\bigl(A^{[p]}\bigr)_\sigma
		= V^{(N)}\!\bigl(\mathbf{0}_{d^{[p]},D_0}\bigr)_\sigma
		+ V^{(N)}\!\bigl(\textstyle\bigoplus_k \mu_k B_k\bigr)_\sigma.
	\]
\end{theorem}

\begin{proof}\leanok
	\uses{thm:tp_gauge_arbitrary, thm:common_blocking_period,
	      lem:block_weights_ne_zero, thm:zero_tail_to_tensor_from_blocks_block_power}
	Apply Theorem~\ref{thm:tp_gauge_arbitrary} to get the trivial block
	and left-canonical nontrivial blocks.
	Apply Theorem~\ref{thm:common_blocking_period} to find a common
	blocking period $p$ making all transfer maps primitive. Lemma~\ref{lem:block_weights_ne_zero}
	keeps the blocked weights nonzero, and
	Theorem~\ref{thm:zero_tail_to_tensor_from_blocks_block_power} transports the
	zero-tail/nonzero MPV identity through the blocking step.
\end{proof}

\begin{theorem}[TP-primitive irreducible tensor is normal]%
	\label{thm:normal_tp_primitive_irred}
	\lean{MPSTensor.isNormal_of_tp_primitive_irreducible}
	\leanok
	\uses{def:primitive_mps, def:irreducible_tensor, def:normal}
	Let $A$ be a left-canonical MPS tensor with $D > 0$,
	irreducible transfer map, and peripheral spectrum $\{1\}$.
	Then $A$ is normal.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:normal_from_primitive_posdef, thm:psd_fp_pd_irred}
	Peripheral primitivity and irreducibility yield the spectral-gap
	condition.  The PSD fixed point is positive definite by
	Theorem~\ref{thm:psd_fp_pd_irred}.
	Apply Theorem~\ref{thm:normal_from_primitive_posdef}.
\end{proof}

\begin{theorem}[TP-primitive irreducible tensor is injective after blocking]%
	\label{thm:tp_primitive_irred_block_injective}
	\lean{MPSTensor.exists_blockTensor_isInjective_of_tp_primitive_irreducible,
		MPSTensor.exists_pos_blockTensor_isInjective_of_tp_primitive_irreducible}
	\leanok
	\uses{thm:normal_tp_primitive_irred,
		lem:isNBlkInjective_iff_blockTensor_isInjective}
	Let $A$ be a left-canonical MPS tensor with $D > 0$,
	irreducible transfer map, and peripheral spectrum $\{1\}$.
	Then some blocking $A^{[L]}$ is one-site injective.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:normal_tp_primitive_irred,
		lem:isNBlkInjective_iff_blockTensor_isInjective}
	The previous theorem gives normality, which is eventual block injectivity.
	The blocked-chain equivalence identifies this with one-site injectivity of
	the corresponding blocked tensor.
\end{proof}

\begin{remark}
	See \cref{thm:isNormal_blockTensor} for the formalized statement that blocking preserves normality.
\end{remark}

\subsection{Basis of Normal Tensors}

\begin{theorem}[Sorting by strictly decreasing weight norms]
	\label{thm:bnt_sorted_reindexing}
	\lean{MPSTensor.exists_sorted_reindexing}\leanok
	If nonzero weights $(\mu_k)_{k=1}^r$ have pairwise distinct norms, then there is a permutation of
	$\{1,\dots,r\}$ under which $\|\mu_k\|$ is strictly decreasing.
\end{theorem}

\begin{theorem}[Sorted decomposition preserves the represented MPS family]
	\label{thm:bnt_sorted_blockdecomp}
	\lean{MPSTensor.exists_sorted_blockDecomp_of_distinct_norms}\leanok
	\uses{thm:bnt_sorted_reindexing}
	Under the same distinct-norm hypothesis, the sorted block family is gauge equivalent to the original decomposition and has strictly decreasing norms.
\end{theorem}

\begin{theorem}[Normal canonical form after sorting]
	\label{thm:bnt_sorted_ncf}
	\lean{MPSTensor.exists_sortedNCF_of_distinct_norms}\leanok
	\uses{def:is_normal_canonical_form, thm:bnt_sorted_blockdecomp}
	If each block is irreducible, trace-preserving, primitive, and has positive dimension, then sorting by strictly decreasing weight norms produces a normal canonical form.
\end{theorem}

\begin{definition}[Granular sector decomposition]
	\label{def:trivial_sector_decomp}
	\lean{MPSTensor.trivialSectorDecomp}
	\leanok
	\uses{def:paper_sector_data}
	Given nonzero weights $(\mu_k)_{k=1}^r$ and blocks $(B_k)_{k=1}^r$,
	form a sector decomposition with one basis block for each original block,
	one copy in each sector, and sector weight $\mu_k$ on the $k$th sector.
\end{definition}

\begin{lemma}[Granular sector decomposition preserves the represented family]
	\label{lem:trivial_sector_decomp_same_mpv}
	\lean{MPSTensor.sameMPV₂_trivialSectorDecomp}
	\leanok
	\uses{def:trivial_sector_decomp}
	The granular sector decomposition associated to $(\mu_k,B_k)_k$ represents
	the same MPS family as the original weighted block sum
	$\bigoplus_k \mu_k B_k$.
\end{lemma}

\begin{proof}\leanok
	\uses{def:trivial_sector_decomp}
	Expanding the sector decomposition gives the sum
	$\sum_k \mu_k^N V^{(N)}(B_k)_\sigma$, which is exactly the finite
	block-sum expression for $\bigoplus_k \mu_k B_k$.
\end{proof}

\begin{theorem}[Trivial sector decomposition for the sorted distinct-norm case]
	\label{thm:bnt_trivial_sector}
	\lean{MPSTensor.exists_trivialSectorDecomp_of_sorted_distinct_norms}\leanok
	\uses{def:trivial_sector_decomp, lem:trivial_sector_decomp_same_mpv}
	If all weights are nonzero and strictly decreasing in norm, one obtains a
	sector decomposition with one copy per sector that preserves the represented
	MPS family.
\end{theorem}

\begin{definition}[Norm-class partition]
	\label{def:norm_class_grouping_data}
	\lean{MPSTensor.NormClassGroupingData}\leanok
	The grouping data consists of: norm classes, one representative value per class, class multiplicities, an enumeration of members in each class, and a regrouping identity recovering the full finite sum.
\end{definition}

\begin{definition}[Norm-class enumeration in decreasing order]
	\label{def:norm_class_grouping}
	\lean{MPSTensor.normClassGroupingData}\leanok
	\uses{def:norm_class_grouping_data}
	From any finite weight family, construct norm classes ordered by strictly decreasing norm values.
\end{definition}

\subsection{Equal-Norm Blocks}

\begin{theorem}[Gauge phases have unit modulus for primitive irreducible blocks]
	\label{thm:unit_modulus_phase_tp_prim_irred}
	\lean{MPSTensor.norm_gaugePhase_eq_one_of_irr_TP_primitive}\leanok
	\uses{def:gauge_phase_equiv, def:irreducible_tensor}
	For trace-preserving primitive irreducible blocks, every gauge-phase
	equivalence has a phase factor of modulus~$1$.
\end{theorem}

\begin{proof}\leanok
	\uses{def:gauge_phase_equiv, def:irreducible_tensor}
	Primitivity forces the peripheral spectrum to consist of $\{1\}$ alone,
	so the gauge phase must have modulus one.
\end{proof}

\begin{remark}
	See \cref{lem:restricted_norm_class_collapse_bnt} for the formalized restricted norm-class collapse statement.
\end{remark}


\begin{theorem}[Existential BNT independence from overlap limits]
	\label{thm:bnt_li_from_overlap_limits_exists}
	\lean{MPSTensor.exists_eventually_linearIndependent_of_overlap_tendsto_orthonormal}
	\leanok
	\uses{thm:bnt_overlap_orthonormal_li}
	If the self-overlaps of a finite block family tend to $1$ and the
	cross-overlaps of distinct blocks tend to $0$, then the MPV states are
	linearly independent for all sufficiently large system sizes.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:bnt_overlap_orthonormal_li}
	This is Theorem~\ref{thm:bnt_overlap_orthonormal_li}, restated with an
	explicit threshold $N_0$.
\end{proof}

\begin{theorem}[Granular sector decomposition from eventual BNT independence]
	\label{thm:bnt_sector_decomp_linear_independent}
	\lean{MPSTensor.exists_bnt_sectorDecomp_of_linearIndependent}
	\leanok
	\uses{def:trivial_sector_decomp, lem:trivial_sector_decomp_same_mpv, def:bnt}
	If the MPV states of the blocks are linearly independent for all sufficiently
	large system sizes and all weights are nonzero, then the granular sector
	decomposition represents the same MPS family and satisfies the BNT
	linear-independence condition.
\end{theorem}

\begin{proof}\leanok
	\uses{def:trivial_sector_decomp, lem:trivial_sector_decomp_same_mpv}
	The equality of represented families is
	Lemma~\ref{lem:trivial_sector_decomp_same_mpv}.  The assumed eventual linear
	independence is exactly the BNT condition for the basis blocks of the granular
	sector decomposition.
\end{proof}

\begin{definition}[MPV phase class data]
	\label{def:mpv_phase_class_data}
	\lean{MPSTensor.MPVPhaseEquiv, MPSTensor.MPVPhaseClassData, MPSTensor.mpvPhaseClassData}
	\leanok
	For a finite family of blocks, put $j \sim k$ when there is a nonzero
	scalar factor $\zeta$ such that
	$\ket{V^{(N)}(B_k)}=\zeta^N\ket{V^{(N)}(B_j)}$ for every length~$N$.
	The class data enumerate the finite quotient, choose one representative per class,
	give the scalar relating the representative to each member, and prove that
	distinct representatives are not gauge-phase equivalent.
\end{definition}

\begin{theorem}[Phase-class BNT sector construction]
	\label{thm:bnt_sector_decomp_tp_prim_irr_collapsed}
	\lean{MPSTensor.exists_bnt_sectorDecomp_of_tp_primitive_irr_blocks}
	\leanok
	\uses{def:mpv_phase_class_data, thm:bnt_li_from_overlap_limits_exists}
	For trace-preserving primitive irreducible blocks of positive bond dimension
	with nonzero weights, there is a sector decomposition representing the same
	weighted block sum and satisfying the BNT linear-independence condition.  The
	construction chooses one representative per MPV phase class and absorbs the
	nonzero scalar factors into the sector weights.
\end{theorem}

\begin{proof}\leanok
	\uses{def:mpv_phase_class_data, thm:bnt_li_from_overlap_limits_exists, thm:mpv_decomposition}
	Finite phase classes partition the original block indices.  Expanding the
	sector decomposition gives the original finite sum after replacing each class
	member by its representative and moving the factor $\zeta^N$ into
	$(\zeta\mu_k)^N$.  If two chosen representatives were gauge-phase equivalent,
	they would lie in the same MPV phase class, contradicting their being distinct
	classes.  Therefore the representatives satisfy the separated primitive-normal
	hypothesis, and the overlap-limit criterion
	(Theorem~\ref{thm:bnt_li_from_overlap_limits_exists}) gives the BNT
	linear-independence condition.
\end{proof}


\begin{theorem}[Phase-class BNT sector construction with one-sided overlap data]
	\label{thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_ortho}
	\lean{MPSTensor.exists_bnt_sectorDecomp_of_tp_primitive_irr_blocks_with_overlapOrtho}
	\leanok
	\uses{thm:bnt_sector_decomp_tp_prim_irr_collapsed,
		def:sector_basis_overlap_ortho_hyps}
	For trace-preserving primitive irreducible blocks of positive bond dimension
	with nonzero weights, there is a sector decomposition, obtained by choosing
	one block per MPV phase class, which represents the same weighted block sum,
	satisfies the BNT linear-independence condition, and has a basis satisfying
	the single-family overlap-orthogonality data.  If the original blocks are
	one-site injective, then the chosen basis blocks are one-site injective.
\end{theorem}

\begin{proof}\leanok
	\uses{def:mpv_phase_class_data, thm:bnt_li_from_overlap_limits_exists,
		thm:cross_overlap_zero_split, thm:bnt_sector_decomp_tp_prim_irr_collapsed}
	The proof is the phase-class construction of
	Theorem~\ref{thm:bnt_sector_decomp_tp_prim_irr_collapsed}.  The chosen basis
	blocks keep positive bond dimension, normalization, and any supplied
	one-site injectivity from the original trace-preserving primitive
	irreducible blocks.  Primitivity gives the self-overlap limit, while the
	separation of distinct chosen phase-class blocks gives off-overlap decay.
\end{proof}

\begin{theorem}[Phase-class BNT sector construction with overlap data]
	\label{thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_data}
	\lean{MPSTensor.exists_bnt_sectorDecomp_of_tp_primitive_irr_blocks_with_overlapData}
	\leanok
	\uses{thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_ortho}
	For trace-preserving primitive irreducible blocks of positive bond dimension,
	with nonzero weights, suppose moreover that the blocks are injective. Then
	there is a sector decomposition, obtained by choosing one block per MPV phase
	class, which represents the same weighted block sum, satisfies the BNT
	linear-independence condition, and has chosen basis blocks of positive bond
	dimension which are injective and trace-preserving, have self-overlap tending
	to~$1$, and have mutual overlaps tending to~$0$ for distinct basis blocks.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_ortho}
	Apply Theorem~\ref{thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_ortho}
	and use the supplied injectivity of the original blocks to fill the
	injectivity field required by this theorem's explicit conclusion.
\end{proof}

\begin{theorem}[Chosen phase representatives preserve finite-length spans]
	\label{thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_data_span}
	\lean{MPSTensor.MPVPhaseClassData.representative_mpv_span_eq}
	\lean{MPSTensor.exists_bnt_sectorDecomp_of_tp_primitive_irr_blocks_with_overlapData_and_basisSpan}
	\leanok
	\uses{def:mpv_phase_class_data,
		thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_data}
	For any finite block family, the chosen MPV phase-class representatives span
	the same finite-length MPV subspace as the original blocks at every system
	size. Consequently, for trace-preserving primitive irreducible injective
	blocks with nonzero weights, the one-sided BNT sector construction can be
	chosen with all overlap data from
	Theorem~\ref{thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_data} and
	with this representative-span identity.
\end{theorem}

\begin{proof}\leanok
	\uses{def:mpv_phase_class_data,
		thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_data}
	Each representative is one of the original blocks. Conversely, every original
	block appears in a phase class and differs from that class representative by a
	nonzero scalar power at each length, so its MPV state lies in the
	representative span. The one-sided sector theorem then gives this identity
	for the chosen sector basis.
\end{proof}

\subsection{Cyclic Sectors}

\begin{definition}[Left sector tensor]
	\label{def:left_sector_tensor}
	\lean{MPSTensor.leftSectorTensor}\leanok
	For a projection $P$, the left sector tensor is obtained by left multiplication of each Kraus operator by $P$.
\end{definition}

\begin{lemma}[Commuting letters imply commuting words]
	\label{lem:commutes_words}
	\lean{MPSTensor.commutes_evalWord_of_commutes_letters}\leanok
	If a projection commutes with every Kraus operator, it commutes with every finite word product.
\end{lemma}

\begin{lemma}[Left multiplication identity for sector words]
	\label{lem:left_mul_sector_word}
	\lean{MPSTensor.left_mul_evalWord_leftSectorTensor_of_commutes}\leanok
	\uses{def:left_sector_tensor, lem:commutes_words}
	If a projection commutes with all Kraus operators, then word evaluation for the left sector tensor equals left multiplication by the projection.
\end{lemma}

\begin{lemma}[Sector support]
	\label{lem:left_sector_supported}
	\lean{MPSTensor.leftSectorTensor_supported}\leanok
	\uses{def:left_sector_tensor}
	The left sector tensor is supported on the chosen projection.
\end{lemma}

\begin{theorem}[Adjoint-fixed projections commute with Kraus operators]
	\label{thm:adjoint_fixed_commute_letters}
	\lean{MPSTensor.commutes_letters_of_adjoint_fixed_projection}\leanok
	If a projection is fixed by the adjoint transfer map of a trace-preserving tensor, then it commutes with each Kraus operator.
\end{theorem}

\begin{theorem}[Left-canonical primitive blocking from irreducibility]
	\label{thm:leftcanonical_primitive_blocking}
	\lean{MPSTensor.exists_blockTensor_leftCanonical_isPrimitive}\leanok
	\uses{thm:blocking_primitivity}
	For a trace-preserving irreducible tensor, a positive blocking period yields a blocked tensor that is left-canonical and primitive.
\end{theorem}

\begin{theorem}[Normality from primitivity and a positive-definite fixed point]
	\label{thm:normal_from_primitive}
	\lean{MPSTensor.isNormal_of_isPrimitiveMPS}\leanok
	\uses{def:normal}
	A primitive tensor admitting a positive-definite fixed point of its transfer map is normal.
\end{theorem}

\begin{theorem}[Zero-word evaluation for the zero tensor]
	\label{thm:evalword_zero_allzero}
	\lean{MPSTensor.evalWord_eq_zero_of_all_zero}\leanok
	Every nonempty finite word evaluates to zero when all Kraus operators are zero.
\end{theorem}

\begin{theorem}[Zero MPV for the zero tensor]
	\label{thm:mpv_zero_allzero}
	\lean{MPSTensor.mpv_eq_zero_of_all_zero}\leanok
	\uses{thm:evalword_zero_allzero}
	For a nonempty spin chain, the MPS value of every spin word is zero for the zero tensor.
\end{theorem}

\begin{theorem}[Nonzero Kraus operator in irreducible dimension at least two]
	\label{thm:nonzero_kraus_irred}
	\lean{MPSTensor.exists_nonzero_kraus_of_isIrreducibleTensor}\leanok
	\uses{def:irreducible_tensor}
	An irreducible tensor of bond dimension at least two has at least one nonzero Kraus operator.
\end{theorem}

\begin{theorem}[All-zero tensors are irreducible only in dimension at most one]
	\label{thm:allzero_irred_dim_le_one}
	\lean{MPSTensor.isIrreducibleTensor_allZero_dim_le_one}\leanok
	\uses{thm:nonzero_kraus_irred}
	If every Kraus operator vanishes and the tensor is irreducible, then the bond dimension is at most one.
\end{theorem}

\begin{theorem}[Irreducible block decomposition with trace-preserving gauge]
	\label{thm:irred_blockdecomp_tpgauge}
	\lean{MPSTensor.exists_irreducible_blockDecomp_with_tpGauge}\leanok
	\uses{thm:allzero_irred_dim_le_one}
	After removing a trivial block, one obtains a block decomposition into irreducible blocks with a trace-preserving gauge.
\end{theorem}

The after-blocking canonical-form data consumed by the Fundamental-Theorem
proof is collected separately in Chapter~\ref{ch:canonical_after_blocking};
see in particular
\cref{thm:after_blocking_common_primitive_irreducible_blocks_reindexed}.

\section{Open directions}\label{sec:open_directions}

\begin{remark}[Canonical-form reduction before the equal-case comparison]%
\label{rem:canonical_pipeline_gaps}
	\leanok
	The canonical-form proof in the MPS literature first separates the nilpotent
	part, which contributes only to the length-zero trace, from the nonzero part.
	The nonzero part is decomposed into irreducible blocks, and each block is put in
	trace-preserving gauge.  A block of period $m$ is then blocked over $m$ sites and
	restricted to its cyclic spectral sectors.  These sector tensors are the normal
	blocks used in the comparison theorem: they are trace-preserving, primitive,
	irreducible, and have positive bond dimension.  For a finite nonzero block
	family, one chooses a common multiple of all periods so that every
	sector is expressed over the same blocked physical alphabet.  This is the
	construction of
	\cite[Section~2.3 and Appendix~A]{Cirac2017MPDO_AnnPhys}, and it is the
	canonical-form information used before the basis-of-normal-tensors argument in
	\cite[Proposition~IV.4 and Corollary~IV.5]{Cirac2021Matrix}.

	The bridge statements collected in Chapter~\ref{ch:canonical_after_blocking}
	follow this reduction. There, Theorem~\ref{thm:tp_gauge_arbitrary}
	gives the zero-tail block and the trace-preserving irreducible nonzero block
	family.  Theorem~\ref{thm:ft_after_blocking_per_block_cyclic_live_zero_tail}
	applies the cyclic-sector decomposition to each nonzero block, preserving the
	positive-length MPV equality and the zero-tail identity at $N=0$.
	Theorem~\ref{thm:exists_common_blocked_cyclic_sector_family_common_multiple}
	constructs the common reblocked sector family at any common multiple of the
	period-removal lengths.  Consequently
	Theorem~\ref{thm:ft_after_blocking_common_length_common_sector_theorem} chooses
	one physical blocking length for both tensors, transports the zero-tail
	identities to that length, and supplies cyclic-sector families on a common
	blocked alphabet.  The supporting blocking and transport results, including
	Theorem~\ref{thm:iterated_blocking_samempv_reindex}, identify iterated blocking
	with one blocking by the product length, after identifying the blocked physical
	words.

	The zero-tail and common-sector results in
	Chapter~\ref{ch:canonical_after_blocking} separate the finite-sum comparison
	from the blocked-word assertion.
	Theorem~\ref{thm:common_blocked_cyclic_sector_blocked_word_comparison}
	shows that, once the length-$p$ word read directly from the common blocked
	alphabet equals the word obtained by flattening an iterated $(m_k,e_k)$ block,
	the weighted direct sum of directly blocked nonzero blocks agrees with the
	common-sector family.
	Theorem~\ref{thm:zero_tail_common_flat_of_blocked_word_comparison}
	then rewrites the one-sided zero-tail identity with the weighted common-length
	cyclic sectors under this blockwise word equality.

	The remaining blocked-word equality, after identifying blocked physical
	words, is the equality between the blocked nonzero part and the cyclic-sector
	tensor for the whole weighted family.
	Theorem~\ref{thm:zero_tail_common_flat_of_reindexed} gives the
	one-sided form of this conversion, and
	Theorem~\ref{thm:ft_after_blocking_common_length_common_sector_reindexed}
	states it simultaneously for the two common-length sector families.  The
	auxiliary transport statements,
	Theorems~\ref{thm:zero_tail_common_flat_transport_of_reindexed},
	\ref{thm:zero_tail_common_flat_transport_of_grouped_block_cast}, and
	\ref{thm:samempv_pos_zero_tail_identity_transport}, give the nonzero
	transported weights and the preservation of the positive-length and length-zero
	identities, while
	Theorem~\ref{thm:after_blocking_common_primitive_irreducible_blocks_reindexed}
	collects the resulting primitive irreducible sector decompositions needed for
	comparison.  Theorem~\ref{thm:common_blocked_cyclic_sector_flatten_word_of_block_cast}
	supplies the direct word equality for the chosen coordinates.
	Theorem~\ref{thm:after_blocking_common_primitive_irreducible_blocks_block_span_consequences}
	gives the intermediate span consequence for these common-sector families: a
	common MPV phase cover or block-permutation gauge-phase conclusion supplies
	the finite-length nonzero-block span equality.
	Theorem~\ref{thm:afterBlocking_sectorComparison_reindexed_commonPhaseCover}
	states the exact inputs after this point: equality of the zero-tail
	dimensions, injectivity at the chosen common length, and a common MPV phase cover
	(or the equivalent finite-length span equality) for the two nonzero-sector
	families.  Thus the canonical-form assertions are the blocked-word
	identification for the whole weighted nonzero family, the zero-tail and
	injectivity refinements, and the common-phase or block-permutation gauge-phase
	comparison for the common-length nonzero sectors.

	The sorted normal-canonical-form theorem,
	Theorem~\ref{thm:bnt_sorted_ncf}, assumes pairwise distinct nonzero weight
	norms. This is a useful special form after sorting. Equal-norm families require
	the basis-of-normal-tensors and multiplicity comparison developed later, rather
	than an artificial separation of equal weight norms.
\end{remark}
