\chapter{Block Permutation and Separation}\label{ch:block_perm}

This chapter establishes the algebraic structure of
automorphisms of products of matrix algebras and uses it
to prove that the same-MPV condition on block-diagonal
tensors descends to per-block gauge equivalence.
The results correspond to the canonical form uniqueness
of~\cite{PerezGarcia2007Matrix} and the block separation
results of~\cite{Cirac2021Matrix}. The proof technique
used here---induction on blocks via the spectral
gap---is specific to this formalization.

\section{Block permutation algebra}

\begin{theorem}[Ring isomorphisms permute block ideals]\label{thm:ring_perm}
    \lean{MPSTensor.ringEquiv_pi_simple_permutes_blockIdeals}
    \leanok
    Let $\{R_k\}_{k=1}^r$ be simple rings.
    Every ring isomorphism $T : \prod_k R_k \to \prod_k R_k$
    permutes the block ideals:
    there exists a permutation $\pi$ such that
    $T$ maps the $k$-th block ideal to the $\pi(k)$-th block ideal.
\end{theorem}

\begin{proof}
    \leanok
    Each block ideal $I_k = \{x : x_j = 0 \text{ for } j \neq k\}$
    is an atom in the lattice of two-sided ideals
    (since $R_k$ is simple).
    A ring isomorphism preserves the ideal lattice and hence
    permutes the atoms.
    Injectivity of the induced map on atoms gives the
    permutation~$\pi$.
\end{proof}

\begin{remark}[Ring level versus algebra level]\label{rem:ring_vs_alg}
    Theorem~\ref{thm:ring_perm} is a ring-level statement: it only uses the
    ideal structure of $\prod_k R_k$.
    Theorems~\ref{thm:dim_preserved}
    and~\ref{thm:pi_auto_decomp} require an upgrade to $\C$-algebra
    automorphisms. The reason is threefold: dimension preservation compares
    the blocks as $\C$-vector spaces; Skolem--Noether applies to the resulting
    central simple $\C$-algebras; and the per-block maps extracted from $T$
    must commute with scalar multiplication. Thus the formalization separates
    the ring-level permutation step from the later algebra-level
    decomposition.
\end{remark}

\begin{theorem}[Dimension preservation]\label{thm:dim_preserved}
    \lean{MPSTensor.dim_preserved}
    \leanok
    If $T$ is an algebra automorphism of
    $\prod_{k=1}^r \MN{D_k}$,
    and $\pi$ is the induced permutation, then
    $D_{\pi(k)} = D_k$ for all~$k$.
\end{theorem}

\begin{proof}\leanok\uses{thm:ring_perm}
    The restriction of $T$ to the $k$-th block ideal
    induces a $\C$-algebra isomorphism
    $\MN{D_k} \cong \MN{D_{\pi(k)}}$,
    which forces $D_k = D_{\pi(k)}$
    (by comparing dimensions as $\C$-vector spaces).
\end{proof}

\begin{theorem}[Decomposition of automorphisms of $\prod \MN{D_k}$]\label{thm:pi_auto_decomp}
    \lean{MPSTensor.algEquiv_pi_matrix_decomposition}
    \leanok
    Every $\C$-algebra automorphism $T$ of $\prod_{k=1}^r \MN{D_k}$
    decomposes as a block permutation~$\pi$ followed by
    per-block inner automorphisms:
    there exist $\pi \in S_r$ with $D_{\pi(k)} = D_k$
    and invertible matrices $X_k \in \GL_{D_k}(\C)$ such that,
    for every tuple $M = (M_1,\ldots,M_r)$,
    $T(M)_k = X_k M_{\pi(k)} X_k^{-1}$,
    where $T(M)_k$ denotes the $k$-th block component of $T(M)$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:ring_perm, thm:dim_preserved, thm:skolem_noether}
    Theorem~\ref{thm:ring_perm} gives the permutation $\pi$.
    Theorem~\ref{thm:dim_preserved} gives $D_{\pi(k)} = D_k$.
    The restriction of $T$ to the $k$-th block ideal yields a $\C$-algebra
    isomorphism $\MN{D_k} \cong \MN{D_{\pi(k)}}$; after
    Theorem~\ref{thm:dim_preserved} identifies the two matrix sizes,
    Skolem--Noether (Theorem~\ref{thm:skolem_noether}) makes this
    restriction inner.

    This is a standard result in the theory of
    semisimple algebras;
    see \cite[§IV.A]{Cirac2021Matrix} for the MPS context.
\end{proof}

\section{Product algebra linear extension}

\begin{definition}[Per-block linear extension]\label{def:per_block_linear_ext}
    \lean{MPSTensor.perBlockLinearExtension}
    \leanok
    \uses{def:injective, def:same_mpv}
    Given injective tensors $A_k, B_k$ of bond dimension~$D_k$
    with the same MPV for each block~$k$,
    the \emph{per-block linear extension} $T_k$ is the unique
    linear map $\MN{D_k}\to\MN{D_k}$ satisfying
    $T_k(A_k^i)=B_k^i$ for all~$i$.
    It exists (and is unique) because $\{A_k^i\}$ spans
    $\MN{D_k}$ by injectivity.
\end{definition}

\begin{definition}[Product algebra automorphism]\label{def:pi_alg_equiv}
    \lean{MPSTensor.piAlgEquiv}
    \leanok
    \uses{def:per_block_linear_ext}
    The \emph{product algebra automorphism} is the blockwise assembled map
    $T:\prod_k\MN{D_k}\to\prod_k\MN{D_k}$ defined by
    $T(M)_k := T_k(M_k)$.
    Theorem~\ref{thm:pi_linear_ext} shows that each $T_k$ is a
    $\C$-algebra homomorphism and that this assembled map is indeed a
    $\C$-algebra automorphism.
\end{definition}

\begin{theorem}[Per-block linear extensions assemble]\label{thm:pi_linear_ext}
    \lean{MPSTensor.piAlgEquiv_decomposition}
    \leanok
    \uses{def:injective, def:per_block_linear_ext}
    Let $A_k$ and $B_k$ be injective tensors of bond dimension
    $D_k$ for $k = 1, \ldots, r$, and suppose the per-block
    same-MPV condition holds.
    Then the per-block linear extensions $T_k : A_k^i \mapsto B_k^i$
    assemble into an algebra automorphism of $\prod_k \MN{D_k}$,
    which decomposes as a permutation $\pi$ with
    $D_{\pi(k)} = D_k$ and per-block conjugation matrices $X_k$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:linear_ext, thm:linear_ext_mul, lem:linear_map_to_alg,
          thm:pi_auto_decomp}
    Each $T_k$ exists uniquely and is multiplicative by
    Theorems~\ref{thm:linear_ext} and~\ref{thm:linear_ext_mul}.
    Promotion to algebra homomorphisms
    (Lemma~\ref{lem:linear_map_to_alg}) and composition into
    the product gives an algebra automorphism $T$.
    Theorem~\ref{thm:pi_auto_decomp} decomposes $T$ into
    a permutation plus per-block inner automorphisms.
    Because $T$ is assembled blockwise, it preserves each block ideal,
    so in this theorem the permutation is actually $\pi = \mathrm{id}$.
    A nontrivial block permutation only appears after block separation
    identifies which blocks match globally.
\end{proof}

\begin{theorem}[Per-block same MPV gives global gauge equivalence]\label{thm:per_block_gauge}
    \lean{MPSTensor.fundamentalTheorem_multiBlock_full}
    \leanok
    \uses{def:injective, def:gauge_equiv, def:tensor_from_blocks}
    If all block tensors $A_k$ are injective and per-block
    same-MPV holds, then each pair $(A_k, B_k)$ is gauge
    equivalent, and the block-diagonal tensors
    $\bigoplus_k \mu_k A_k$ and $\bigoplus_k \mu_k B_k$
    are globally gauge equivalent.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:ft_single, thm:multi_block_global}
    Per-block gauge equivalence follows from the single-block
    FT (Theorem~\ref{thm:ft_single}) applied to each block.
    Global gauge equivalence follows by assembling the
    per-block gauge matrices into a block-diagonal matrix
    (Theorem~\ref{thm:multi_block_global}).
\end{proof}

\section{Block separation}

Recall from Definition~\ref{def:mpv_overlap}
and Theorem~\ref{thm:overlap_trace} that the bilinear overlap is
\[
    O_{AB}(N) = \sum_{\sigma \in \{0,\ldots,d{-}1\}^N}
    V^{(N)}(A)_\sigma \overline{V^{(N)}(B)_\sigma} = \Tr(F_{AB}^N).
\]
We use this notation throughout the block-separation argument.

\begin{definition}[Canonical form predicate]\label{def:is_canonical_form}
    \lean{MPSTensor.IsCanonicalForm}
    \leanok
    \uses{def:injective}
    A collection of scaling factors $(\mu_k)_{k=1}^r$ and
    block tensors $(A_k)_{k=1}^r$ satisfies the
    \emph{canonical form predicate} if:
    \begin{enumerate}
        \item each $A_k$ is injective,
        \item each $A_k$ satisfies the TP normalization
              $\sum_i (A_k^i)^\dagger A_k^i = \Id$,
        \item the moduli are strictly decreasing:
              $|\mu_1| > |\mu_2| > \cdots > |\mu_r|$,
        \item all $\mu_k \neq 0$,
        \item the self-overlap converges: $O_{A_k A_k}(N) \to 1$
              for each~$k$.
    \end{enumerate}
    Only this one-sided normalization is part of the predicate;
    no unital condition is assumed here.
    Condition~(5) is a primitivity hypothesis:
    for a primitive channel
    (\cite[Thm.~6.7, item~3]{Wolf2012Quantum}),
    $T^n(\rho)\to\rho_\infty$ for every initial state,
    which forces the self-overlap to converge to~$1$.
    See also \cite[Thm.~6.8]{Wolf2012Quantum}
    for the completely-positive characterisation
    and Chapter~\ref{ch:spectral} for the spectral-gap
    proof.
\end{definition}

\begin{remark}[Normalization convention]\label{rem:tp_vs_unital}
    The convention here differs from
    \cite[Thm.~4]{PerezGarcia2007Matrix}, which uses the unital gauge
    $\sum_i A_k^i (A_k^i)^\dagger = \Id$.
    This chapter instead works in the dual TP gauge
    $\sum_i (A_k^i)^\dagger A_k^i = \Id$.
    A positive-definite fixed point of the adjoint transfer map provides a
    gauge transformation between these two normalizations
    (Theorem~\ref{thm:left_canonical_gauge}).
\end{remark}

\begin{remark}\label{rem:overlap_hypothesis}\leanok
    In the literature~\cite{Cirac2021Matrix},
    condition~(5) is derived from primitivity of the
    transfer map (peripheral spectrum $= \{1\}$).
    Theorem~\ref{thm:canonical_from_primitive} gives this implication:
    per-block primitivity implies overlap convergence,
    so condition~(5) is redundant when primitivity holds.
    The canonical form predicate retains condition~(5)
    explicitly because it can be verified independently of primitivity.
\end{remark}

\begin{remark}[Normal-canonical companion]\label{rem:normal_cf_companion}
    Definition~\ref{def:is_canonical_form} is the block-injective predicate.
    The normal-canonical route instead uses
    Definition~\ref{def:is_normal_canonical_form}: injectivity and the explicit
    self-overlap hypothesis are replaced by irreducibility, TP normalization,
    and primitive transfer maps. Theorem~\ref{thm:ncf_overlap_tendsto_one}
    then supplies the self-overlap limit automatically.
\end{remark}

\begin{theorem}[Vandermonde separation]\label{thm:vandermonde_sep}
    \lean{MPSTensor.vandermonde_separation}
    \leanok
    If $\mu_1,\ldots,\mu_r$ are pairwise distinct complex numbers
    and $\sum_{k=1}^r c_k \mu_k^N = 0$ for
    $N = 0,1,\ldots,r-1$,
    then $c_k = 0$ for all~$k$.
\end{theorem}

\begin{proof}\leanok
    The coefficient vector $c$ satisfies $V c = 0$
    where $V$ is the Vandermonde matrix of the $\mu_k$.
    Injectivity of $\mu$ makes $V$ invertible.
\end{proof}

\begin{remark}
    Theorem~\ref{thm:vandermonde_sep} is not used in the current proof chain.
    It is retained for potential alternative routes; the current path to
    Theorem~12.5 uses BNT linear independence (Chapter~11) rather than
    Newton--Girard identities to match the weight multiset.
\end{remark}

\begin{lemma}[Block separation core]\label{lem:block_sep_core}
    \lean{MPSTensor.block_separation_core}
    \leanok
    \uses{def:is_canonical_form, def:injective}
    Let $((\mu_k), (A_k))$ satisfy the canonical form predicate,
    and let $(B_k)$ be injective tensors satisfying the same TP
    normalization $\sum_i (B_k^i)^\dagger B_k^i = \Id$.
    If $\sum_k \mu_k^N (V^{(N)}(A_k)_\sigma - V^{(N)}(B_k)_\sigma) = 0$
    for all $N,\sigma$,
    then each pair $(A_k, B_k)$ generates the same MPV family.
\end{lemma}

\begin{proof}\leanok
    \uses{thm:overlap_decay, thm:overlap_decay_rect}
    By induction on the number of blocks~$r$.
    In the inductive step, take overlaps with the
    leading block $A_0$; after division by $\mu_0^N$,
    the contributions from $k \ge 1$ decay because
    $|\mu_k / \mu_0| < 1$ (strict weight ordering) and the bilinear overlaps
    $O_{A_0 A_k}(N)$ and $O_{A_0 B_k}(N)$ are bounded
    (Theorems~\ref{thm:overlap_decay_rect}
    and~\ref{thm:overlap_decay} supply the relevant bounds), leaving
    $O_{A_0 A_0}(N) - O_{A_0 B_0}(N) \to 0$.
    Condition~(5) gives $O_{A_0 A_0}(N) \to 1$, hence
    $O_{A_0 B_0}(N) \to 1$.
    Since $O_{A_0 B_0}(N) \to 1 \neq 0$, the bond dimensions must match:
    $D_{A_0} = D_{B_0}$; otherwise Theorem~\ref{thm:overlap_decay_rect}
    would give $O_{A_0 B_0}(N) \to 0$, a contradiction.
    With equal bond dimensions established, the contrapositive of
    Theorem~\ref{thm:overlap_decay} forces $A_0$ and $B_0$
    to be gauge-phase equivalent; the limit $O_{A_0 B_0}(N) \to 1$
    then makes the phase trivial, so $A_0$ and $B_0$
    generate the same MPV family.
    Peeling off the leading block and dividing by
    $\mu_0^N$, the induction hypothesis applies.
\end{proof}

\begin{theorem}[Block separation from canonical form]\label{thm:block_sep}
    \lean{MPSTensor.per_block_sameMPV_of_canonical_form}
    \leanok
    \uses{def:is_canonical_form}
    Let $((\mu_k), (A_k))$ satisfy the canonical form predicate,
    and let $(B_k)$ be injective tensors satisfying the same TP
    normalization $\sum_i (B_k^i)^\dagger B_k^i = \Id$.
    If the block-diagonal tensors
    $\bigoplus_k \mu_k A_k$ and $\bigoplus_k \mu_k B_k$
    generate the same MPV family, then each pair
    $(A_k, B_k)$ generates the same MPV family.
\end{theorem}

\begin{proof}
    \leanok
    \uses{lem:block_sep_core}
    Apply Lemma~\ref{lem:block_sep_core} to the identity obtained from the
    equality of the two block-diagonal MPV families.
    The canonical form predicate supplies injectivity, TP normalization,
    strict nonzero weights, and self-overlap convergence for $(A_k)$, while
    the hypotheses on $(B_k)$ supply the comparison injectivity and
    TP normalization.
\end{proof}

\begin{theorem}[Block separation from normal canonical form]\label{thm:block_sep_normal}
    \lean{MPSTensor.per_block_sameMPV_of_normal_canonical_form}\leanok
    \uses{def:is_normal_canonical_form}
    Let $((\mu_k), (A_k))$ satisfy the normal canonical form predicate,
    and let $(B_k)$ be irreducible tensors of the same bond dimensions, satisfying the same TP
    normalization $\sum_i (B_k^i)^\dagger B_k^i = \Id$.
    If the block-diagonal tensors
    $\bigoplus_k \mu_k A_k$ and $\bigoplus_k \mu_k B_k$
    generate the same MPV family, then each pair
    $(A_k, B_k)$ generates the same MPV family.
\end{theorem}

\begin{proof}\leanok
    \uses{thm:ncf_overlap_tendsto_one, thm:nt_modulus_one_gauge, thm:block_sep}
    The proof follows the same peeling argument as Theorem~\ref{thm:block_sep},
    with Theorem~\ref{thm:nt_modulus_one_gauge} (modulus-one eigenvalue rigidity
    for irreducible tensors) replacing the contrapositive of
    Theorem~\ref{thm:overlap_decay} at the identification step, since the blocks
    are irreducible rather than injective.
    Theorem~\ref{thm:ncf_overlap_tendsto_one}
    replaces the explicit self-overlap hypothesis on the leading block.
    Since both block families already use the same dimension function,
    the conclusion is direct per-block same MPV.
\end{proof}

\begin{theorem}[Canonical form fundamental theorem, same-structure version]\label{thm:ft_canonical}
    \lean{MPSTensor.fundamentalTheorem_canonicalForm}
    \leanok
    \uses{def:is_canonical_form, def:gauge_equiv, def:tensor_from_blocks}
    Under the hypotheses of Theorem~\ref{thm:block_sep}, each pair
    $(A_k, B_k)$ is gauge equivalent. Hence the block-diagonal tensors
    $\bigoplus_k \mu_k A_k$ and $\bigoplus_k \mu_k B_k$
    are gauge equivalent.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:block_sep, thm:per_block_gauge}
    Theorem~\ref{thm:block_sep} gives per-block same MPV.
    Theorem~\ref{thm:per_block_gauge} converts this to
    per-block gauge equivalence and then to global gauge equivalence
    of the block-diagonal tensors.
\end{proof}

\begin{remark}[Relation to the literature]\label{rem:block_perm_literature}
    The results of this chapter correspond to the canonical-form uniqueness
    of~\cite{PerezGarcia2007Matrix} and the block-separation statements
    of~\cite{Cirac2021Matrix}.
    The proof technique used here---an induction on the number of blocks
    driven by the spectral gap of the mixed transfer operators---is specific
    to this formalization.
\end{remark}
