

\chapter{Basis of Normal Tensors}\label{ch:bnt}

This chapter introduces the \emph{basis of normal tensors} (BNT) notion,
the canonical-form variants with BNT separation, the
cross-overlap decay theorem, and the permutation-rigidity
arguments used in the proof of the Fundamental Theorem.
It is the reference for the repeated-copy and equal-modulus coefficient
comparison used later: sector coefficients are compared here, and the
Newton--Girard identities recover the copy weights from those coefficients.
The main references are
\cite{PerezGarcia2007Matrix}, \cite{Cirac2016MPDO_arXiv},
and~\cite{Cirac2021Matrix}.

\section{Basis of normal tensors}

\begin{definition}[CPSV basis of normal tensors]\label{def:bnt_cpsv}
    \lean{MPSTensor.IsCPSVBasisOfNormalTensors}
    \leanok
    A finite family $\{A_j\}_{j=1}^g$ of normal tensors, with bond dimension
    allowed to depend on $j$, is a \emph{basis of normal tensors} for $A$ if:
    \begin{enumerate}
        \item each $A_j$ is normal in the sense of Definition~\ref{def:nt_cpsv};
        \item for every system length $N$ there exist coefficients
              $c_j^{(N)} \in \C$ with
              \[
                  \ket{V^{(N)}(A)}
                  = \sum_{j=1}^{g} c_j^{(N)} \ket{V^{(N)}(A_j)};
              \]
        \item there exists $N_0$ such that for all $N > N_0$ the family
              $\{\ket{V^{(N)}(A_j)}\}_{j=1}^{g}$ is linearly independent.
    \end{enumerate}
    This is the basis-of-normal-tensors definition in
    \cite[\S~II.C, lines~271--274]{Cirac2017MPDO_AnnPhys} and
    \cite[Definition~4.2]{Cirac2021Matrix}.
\end{definition}

\begin{definition}[Basis of Normal Tensors (BNT)]\label{def:bnt}
    \lean{MPSTensor.IsBNT}
    \leanok
    \uses{def:normal, def:mpv}
    A \emph{basis of normal tensors} (BNT) for a
    total tensor $A_{\mathrm{tot}}$ consists of $g$ block tensors
    $(A_1, \ldots, A_g)$ with bond dimensions $(D_1, \ldots, D_g)$
    such that:
    \begin{enumerate}
        \item each $A_j$ is normal
              (Definition~\ref{def:normal}),
        \item at each system size~$N$, the MPV of $A_{\mathrm{tot}}$
              lies in the span of the block MPVs
              $\{\ket{V^{(N)}(A_j)}\}_{j=1}^g$,
        \item for sufficiently large~$N$, the block MPV
              vectors $\{\ket{V^{(N)}(A_j)}\}_{j=1}^g$ are
              linearly independent.
    \end{enumerate}
    This is \cite[Definition~4.2]{Cirac2021Matrix}. In the canonical-form
    characterization of \cite[Section~2.3 and Appendix~A]{Cirac2016MPDO_arXiv},
    a BNT is a minimal family of representatives for the relation
    \[
        \widetilde A_k^i = e^{i\phi_k} X_k A_j^i X_k^{-1}
    \]
    among the normal blocks in the canonical form.
\end{definition}

\begin{lemma}[Finite-index overlap-orthonormality implies eventual linear independence]
    \label{lem:bnt_finite_index_overlap_orthonormal_li}
    \lean{MPSTensor.eventually_linearIndependent_of_finite_overlap_tendsto_orthonormal}
    \leanok
    \uses{def:mpv_overlap}
    Let $I$ be a finite index set and let $(A_i)_{i \in I}$ be a finite
    family of tensors such that $O_{A_iA_i}(N) \to 1$ for each $i$, and
    $O_{A_iA_j}(N) \to 0$ whenever $i \neq j$. Then the MPV states
    $\{\ket{V^{(N)}(A_i)}\}_{i \in I}$ are linearly independent for all
    sufficiently large~$N$.
\end{lemma}

\begin{proof}\leanok
    The Gram matrix of the family $\{\ket{V^{(N)}(A_i)}\}_{i \in I}$
    converges entrywise to the identity matrix. Hence for all sufficiently
    large $N$ it is invertible, so the MPV states are linearly independent.
\end{proof}

\begin{lemma}[Overlap-orthonormality implies eventual linear independence]
    \label{lem:bnt_overlap_orthonormal_li}
    \lean{MPSTensor.eventually_linearIndependent_of_overlap_tendsto_orthonormal}
    \leanok
    \uses{def:mpv_overlap, lem:bnt_finite_index_overlap_orthonormal_li}
    Let $(A_j)_{j=1}^g$ be a finite family of tensors such that
    $O_{A_jA_j}(N) \to 1$ for each $j$ and $O_{A_iA_j}(N) \to 0$
    for $i \neq j$. Then the MPV states
    $\{\ket{V^{(N)}(A_j)}\}_{j=1}^g$ are linearly independent for all
    sufficiently large~$N$.
\end{lemma}

\begin{proof}\leanok
    Apply Lemma~\ref{lem:bnt_finite_index_overlap_orthonormal_li} with
    $I = \{1,\ldots,g\}$.
\end{proof}

\begin{theorem}[Existential BNT independence from overlap limits]%
    \label{thm:bnt_li_from_overlap_limits_exists}
    \lean{MPSTensor.exists_eventually_linearIndependent_of_overlap_tendsto_orthonormal}
    \leanok
    \uses{lem:bnt_overlap_orthonormal_li}
    If the self-overlaps of a finite block family tend to $1$ and the
    cross-overlaps of distinct blocks tend to $0$, then the MPV states are
    linearly independent for all sufficiently large system sizes.
\end{theorem}

\begin{proof}\leanok
    \uses{lem:bnt_overlap_orthonormal_li}
    This is Lemma~\ref{lem:bnt_overlap_orthonormal_li}, restated with an
    explicit threshold $N_0$.
\end{proof}

\begin{lemma}[Eventual coefficient extraction from linear independence]%
    \label{lem:eventual_coeff_eq_of_eventual_li}
    \lean{MPSTensor.coefficient_eventually_eq_of_eventually_linearIndependent}
    \leanok
    If a finite family of vectors is linearly independent for all sufficiently
    large $N$, and two finite linear combinations of that family are equal
    for all sufficiently large $N$, then the corresponding coefficients are
    equal for all sufficiently large $N$.
\end{lemma}

\begin{proof}\leanok
    Move the two linear combinations to one side. Linear independence forces
    each coefficient difference to vanish at every sufficiently large length.
\end{proof}

\begin{lemma}[Overlap-orthonormality for two families]%
    \label{lem:bnt_two_family_overlap_orthonormal_li}
    \lean{MPSTensor.eventually_linearIndependent_of_two_family_overlap_tendsto_orthonormal}
    \leanok
    \uses{def:mpv_overlap, lem:bnt_finite_index_overlap_orthonormal_li}
    Let $(A_j)_{j=1}^{g_A}$ and $(B_k)_{k=1}^{g_B}$ be finite families of
    tensors. Suppose the overlaps inside each family are asymptotically
    orthonormal, and suppose every mixed overlap
    $O_{A_jB_k}(N)$ tends to~$0$. Then the combined MPV family
    \[
        \{\ket{V^{(N)}(A_j)}\}_{j=1}^{g_A}
        \cup
        \{\ket{V^{(N)}(B_k)}\}_{k=1}^{g_B}
    \]
    is linearly independent for all sufficiently large~$N$.
\end{lemma}

\begin{proof}\leanok
    Apply the finite-index Gram-matrix criterion to the disjoint union of the two
    index sets. The within-family hypotheses give the diagonal and
    off-diagonal limits inside each block, and the mixed overlap hypothesis gives
    the off-diagonal limits between the two blocks.
\end{proof}

\begin{definition}[No gauge-phase-equivalent distinct blocks]%
    \label{def:blocks_not_gauge_phase_equiv}
    \lean{MPSTensor.BlocksNotGaugePhaseEquiv}
    \leanok
    \uses{def:gauge_phase_equiv}
    A finite block family has no gauge-phase-equivalent distinct blocks when,
    for any two distinct indices with the same bond dimension, the corresponding
    blocks are not gauge-phase equivalent.
\end{definition}

\begin{definition}[Normal canonical form with BNT separation]\label{def:normal_cf_bnt}
    \lean{MPSTensor.IsNormalCanonicalFormBNT}\leanok
    \uses{def:is_normal_canonical_form, def:blocks_not_gauge_phase_equiv}
    A family is in \emph{normal canonical form with BNT separation}
    if it satisfies Definition~\ref{def:is_normal_canonical_form}
    and, in addition, distinct blocks of the same bond dimension are not
    gauge-phase equivalent, so the blocks form a basis of normal tensors.
    The weight moduli are non-increasing, inherited from
    Definition~\ref{def:is_normal_canonical_form}; they need not be strictly
    decreasing, so equal-modulus blocks are allowed. Grouping into a basis of
    normal tensors is governed by gauge-phase equivalence of the block tensors,
    not by distinctness of the weight moduli.
    Repeated equal-modulus copies of a single basis tensor are not represented
    here as separate blocks; they appear as the individual weights $\mu_{j,q}$
    in the coefficient sum $c_N^{(j)} = \sum_q \mu_{j,q}^N$ used in the
    sector-BNT surface below. This is not the source BNT
    expansion
    \[
        A^i
        = \bigoplus_{j=1}^g \bigoplus_{q=1}^{r_j}
        \mu_{j,q} X_{j,q} A_j^i X_{j,q}^{-1},
    \]
    where repeated copies of the same basis tensor remain visible through
    their coefficients, multiplicities, and gauges.
\end{definition}

\begin{theorem}[Cross-overlap decay from explicit BNT hypotheses]
    \label{thm:cross_overlap_zero_split}
    \lean{MPSTensor.cross_overlap_tendsto_zero_of_separated_bnt_data}
    \leanok
    \uses{def:mpv_overlap, def:gauge_phase_equiv}
    Let $(A_j)_{j=1}^r$ be injective and trace-preserving normalized.
    Assume that distinct equal-dimension blocks are not gauge-phase
    equivalent. Then $O_{A_jA_k}(N) \to 0$ for every pair $j \neq k$.
\end{theorem}

\begin{proof}\leanok\uses{thm:overlap_decay, thm:overlap_decay_rect}
    If the bond dimensions differ, Theorem~\ref{thm:overlap_decay_rect}
    gives the decay. If the bond dimensions agree, the non-equivalence
    hypothesis excludes gauge-phase equivalence, so
    Theorem~\ref{thm:overlap_decay} gives the decay.
\end{proof}

\begin{lemma}[BNT from explicit blockwise hypotheses]
    \label{lem:cf_bnt_is_bnt_split}
    \lean{MPSTensor.isBNT_of_separated_bnt_data}
    \leanok
    \uses{def:bnt, def:mpv_overlap}
    Let $(\mu_k, A_k)_{k=1}^r$ be a weighted block family. Assume that the
    blocks are injective and trace-preserving normalized, that
    $O_{A_kA_k}(N) \to 1$ for each $k$, and that distinct equal-dimension
    blocks are not gauge-phase equivalent. Then the blocks form a basis of
    normal tensors for the block-diagonal tensor $\bigoplus_k \mu_k A_k$.
\end{lemma}

\begin{proof}\leanok
    \uses{thm:mpv_decomposition, lem:bnt_overlap_orthonormal_li, thm:cross_overlap_zero_split}
    Injective blocks are normal. The block-diagonal decomposition theorem gives
    the MPV expansion of $\bigoplus_k \mu_k A_k$ in terms of the block MPVs.
    The self-overlap limits are part of the hypotheses, and
    Theorem~\ref{thm:cross_overlap_zero_split} supplies the off-diagonal
    decay. Therefore Lemma~\ref{lem:bnt_overlap_orthonormal_li} gives the
    eventual linear independence required in Definition~\ref{def:bnt}.
\end{proof}

\begin{remark}[Comparing the two normality conditions]%
    \label{rem:normalcf_bnt_gap}
    Definition~\ref{def:normal_cf_bnt} references the normal canonical form
    conditions (Definition~\ref{def:is_normal_canonical_form}), which encode
    normality via the spectral characterization (primitive transfer map).
    Definition~\ref{def:bnt} requires normality via the algebraic
    characterization (eventual block injectivity,
    Definition~\ref{def:normal}). These are equivalent by Remark~2.20, but
    the equivalence must be invoked explicitly: the Wielandt theory of
    Chapter~\ref{ch:wielandt} provides the implication from primitivity to
    eventual block injectivity needed here.
\end{remark}

\begin{lemma}[Restricted normal-CF-BNT hypotheses give a BNT]
    \label{lem:normal_cf_bnt_is_bnt}
    \lean{MPSTensor.IsNormalCanonicalFormBNT.isBNT}
    \leanok
    \uses{def:normal_cf_bnt, def:bnt}
    If a family is in normal canonical form with BNT separation and each block
    is also known to be normal in the algebraic sense, then the blocks form a
    basis of normal tensors for the associated block-diagonal tensor.
    The blocks here are distinct representatives, so this form does not carry the
    repeated-copy multiplicity data of the full CPSV16 BNT statement; the missing
    multiplicity route is recorded in
    \path{docs/paper-gaps/ft_one_copy_scope_restriction.tex}.
\end{lemma}

\begin{proof}\leanok\uses{rem:normalcf_bnt_gap}
    The normal-canonical hypotheses already give the block-diagonal MPV
    expansion, the self-overlap normalization, and the eventual linear
    independence coming from irreducible trace-preserving overlap decay.
    Supplying algebraic normality for each block adds the remaining clause in
    Definition~\ref{def:bnt}.
\end{proof}

\section{Permutation rigidity for bases of normal tensors}

\begin{lemma}[Invertible change of basis]
    \label{lem:bnt_invertible_change_basis}
    \lean{exists_invertible_changeBasis}
    \leanok
    Let $v_1, \ldots, v_g$ and $w_1, \ldots, w_g$ be two linearly independent
    families in a complex vector space with the same span. Then there exists an
    invertible matrix $U \in \GL_g(\C)$ such that
    \[
        w_j = \sum_{i=1}^g U_{ij} v_i
    \]
    for every $j$.
\end{lemma}

\begin{proof}\leanok
    Because the two families span the same subspace, each $w_j$ has a unique
    expansion in the basis $\{v_i\}_{i=1}^g$; these coefficients form the
    columns of $U$. If $U a = 0$, then the corresponding linear relation among
    the $w_j$ is trivial because the $w_j$ are linearly independent. Thus $U$
    is injective, hence invertible.
\end{proof}

\begin{lemma}[Eventual change of basis from equal spans]
    \label{lem:bnt_eventual_change_basis}
    \lean{MPSTensor.eventually_exists_invertible_changeBasis}
    \leanok
    \uses{lem:bnt_invertible_change_basis, def:mpv_overlap}
    If two finite families of blocks have asymptotically orthonormal overlaps
    within each family and their MPV spans agree for every system size $N$,
    then for all sufficiently large $N$ there exists an invertible matrix
    $U_N$ expressing the MPV states of one family as linear combinations of
    the MPV states of the other.
\end{lemma}

\begin{proof}\leanok\uses{lem:bnt_overlap_orthonormal_li, lem:bnt_invertible_change_basis}
    By Lemma~\ref{lem:bnt_overlap_orthonormal_li}, both families are
    linearly independent for all sufficiently large $N$. At such an $N$, the
    span equality lets us apply Lemma~\ref{lem:bnt_invertible_change_basis}.
\end{proof}

\begin{theorem}[Permutation rigidity for equal block counts]
    \label{thm:bnt_perm_same_card}
    \lean{MPSTensor.exists_perm_dimEq_gaugePhaseEquiv_of_overlapOrtho}
    \leanok
    \uses{def:injective, def:mpv_overlap, def:gauge_phase_equiv}
    Let $(A_j)_{j=1}^g$ and $(B_j)_{j=1}^g$ be injective families of positive
    bond dimension satisfying trace-preserving normalization, with self-overlaps tending to~$1$ and
    off-diagonal overlaps tending to~$0$ within each family. If the spans of
    the MPV states agree for every system size $N$, then there is a permutation
    $\pi$ of $\{1, \ldots, g\}$ such that $A_j$ and $B_{\pi(j)}$ have the
    same bond dimension and are gauge-phase equivalent.
\end{theorem}

\begin{proof}\leanok\uses{lem:bnt_eventual_change_basis, thm:overlap_decay, thm:overlap_decay_rect}
    For large $N$, Lemma~\ref{lem:bnt_eventual_change_basis} gives an
    invertible matrix $U_N$ relating the two MPV families. Fix $j$ and write
    $\ket{V^{(N)}(B_j)}=\sum_i u_{ij}^{(N)}\ket{V^{(N)}(A_i)}$. With
    $G_A(N)=(O_{A_iA_\ell}(N))_{i,\ell}$ and
    $b_j(N)=(O_{A_iB_j}(N))_i$, this gives
    \[
        G_A(N)\overline{u_j(N)}=b_j(N),
        \qquad
        O_{B_jB_j}(N)=u_j(N)^{\top}G_A(N)\overline{u_j(N)}.
    \]
    If all mixed overlaps $O_{A_iB_j}(N)$ tended to $0$, then
    $b_j(N)\to0$; since $G_A(N)\to I$, one obtains
    $\overline{u_j(N)}\to0$ and hence $u_j(N)\to0$. Thus
    $O_{B_jB_j}(N)\to0$, contradicting $O_{B_jB_j}(N)\to1$.
    Thus each $B_j$ has a matched block $A_{f(j)}$ with non-decaying mixed
    overlap. Theorems~\ref{thm:overlap_decay_rect}
    and~\ref{thm:overlap_decay} then force equal bond dimension and
    gauge-phase equivalence for each match. Off-diagonal decay within the
    $B$-family makes $f$ injective, hence bijective, and its inverse is the
    required permutation.
\end{proof}

\begin{lemma}[Permutation rigidity with asymptotically orthonormal overlaps]\label{lem:bnt_perm_simple}
    \lean{MPSTensor.exists_eq_numBlocks_and_equiv_gaugePhase_of_overlapOrtho}
    \leanok
    \uses{def:injective, def:mpv, def:mpv_overlap, def:gauge_phase_equiv}
    Let $(A_j)_{j=1}^{g_A}$ and $(B_k)_{k=1}^{g_B}$ be two
    families of injective tensors of positive bond dimension satisfying the TP normalization
    $\sum_i (A_j^i)^\dagger A_j^i = \Id$ and
    $\sum_i (B_k^i)^\dagger B_k^i = \Id$, whose self-overlaps
    converge to~$1$ and whose cross-overlaps within each
    family converge to~$0$.
    If for every system size~$N$,
    \[
        \spn_{\C}\{\ket{V^{(N)}(A_j)} : 1 \le j \le g_A\}
        =
        \spn_{\C}\{\ket{V^{(N)}(B_k)} : 1 \le k \le g_B\},
    \]
    then $g_A = g_B$, and there is a permutation $\pi$ of the common
    block index set such that $A_j$ and $B_{\pi(j)}$ have the same bond
    dimension and are gauge-phase equivalent for each~$j$.
\end{lemma}

\begin{proof}\leanok\uses{lem:bnt_overlap_orthonormal_li, thm:bnt_perm_same_card}
    By Lemma~\ref{lem:bnt_overlap_orthonormal_li}, both families are
    linearly independent for all sufficiently large $N$. At such an $N$, the
    common span has the same dimension when computed from the $A$-family or
    the $B$-family, so $g_A = g_B$. After identifying the two index sets,
    Theorem~\ref{thm:bnt_perm_same_card} yields the permutation conclusion.
\end{proof}

\begin{theorem}[Non-equivalent irreducible TP blocks are eventually non-proportional]
    \label{thm:not_gauge_phase_eventually_not_proportional}
    \lean{MPSTensor.exists_ge_not_forall_mpv_eq_mul_of_not_gaugePhaseEquiv_of_irreducible_TP}
    \lean{MPSTensor.exists_ge_not_forall_mpv_eq_mul_of_blocksNotGaugePhaseEquiv_of_irreducible_TP}
    \leanok
    \uses{def:gauge_phase_equiv, def:mpv_overlap, def:irreducible_tensor}
    Let $A$ and $B$ be irreducible trace-preserving blocks whose
    self-overlaps converge to $1$. If they are not gauge-phase equivalent,
    then for every lower bound $N_0$ there is $N \ge N_0$ such that
    $\ket{V^{(N)}(A)}$ is not proportional to $\ket{V^{(N)}(B)}$.
    Consequently, in an already separated same-dimensional BNT block family,
    any two distinct blocks have such non-proportional MPV states at
    arbitrarily large lengths.
\end{theorem}

\begin{proof}\leanok
    Suppose that for all sufficiently large $N$ one had
    $\ket{V^{(N)}(A)}=c_N\ket{V^{(N)}(B)}$. Then
    \[
        O_{AA}(N)=|c_N|^2O_{BB}(N),
        \qquad
        O_{AB}(N)=c_NO_{BB}(N).
    \]
    Since $O_{AA}(N)\to1$ and $O_{BB}(N)\to1$, the first identity gives
    $|c_N|\to1$, and hence $|O_{AB}(N)|\to1$. The irreducible
    trace-preserving overlap decay theorem gives $O_{AB}(N)\to0$ when
    $A\not\sim_{\mathrm{gp}}B$, a contradiction.
\end{proof}

\section{Newton--Girard identities}

\begin{theorem}[Newton--Girard trace recursion]\label{thm:newton_girard}
    \lean{Matrix.charpoly_eq_of_forall_trace_pow_eq,
        Matrix.charpoly_eq_of_trace_pow_eq_of_le_card}
    \leanok
    If $A, B \in \MN{n}$ satisfy
    $\tr(A^k) = \tr(B^k)$ for all $k \ge 1$,
    then $A$ and $B$ have the same characteristic polynomial.
    It is enough to assume this equality for $1 \le k \le n$.
\end{theorem}

\begin{proof}
    \leanok
    Let $e_m$ denote the $m$th elementary symmetric polynomial in the
    eigenvalues, with $e_0 = 1$.
    The Newton--Girard recursion
    \[
        m e_m = \sum_{i=1}^m (-1)^{i-1} e_{m-i} \tr(A^i)
    \]
    expresses the coefficients of the characteristic polynomial in terms
    of the power-sum traces $\tr(A^i)$.
    Equal power sums for $A$ and $B$ therefore give equal coefficients,
    hence equal characteristic polynomials.
\end{proof}

\begin{remark}[How the power-sum equality is used]
    The equal-MPV corollary in the source compares scalar power sums only after
    the BNT representatives have already been matched. If the paired basis
    tensors satisfy $B_j=e^{i\phi_j}Y_jA_jY_j^{-1}$, then for every positive
    length $N$ the coefficient $\sum_q \mu_{j,q}^N$ of
    $\ket{V^{(N)}(A_j)}$ in the BNT expansion of $A$ equals the coefficient
    $\sum_q(\nu_{j,q}e^{i\phi_j})^N$ coming from $B$. This scalar equality
    is the input to the Newton--Girard step.

    Thus the power-sum lemma determines the multiset of weights with
    multiplicity, but it is not a replacement for the BNT matching argument. The
    block permutation and the phase gauges must first identify which basis tensor
    is being compared. Once that matching is available, the power-sum lemma
    supplies the multiplicity and weight equality needed to identify the weighted
    block-diagonal gauge.

    The finite-range statement below includes the same-cardinality form, the
    nonzero-entry unequal-cardinality form, and the positive-power conclusion
    after deleting zeros. The nonzero-entry form pads the shorter family by zeros,
    applies the same-cardinality result at length $\max(m,n)$, and then uses the
    nonzero-entry hypothesis to count the padded zeros. The Appendix lemma in the
    source fixes an ordering convention for the two families; here the conclusion
    is stated as multiset equality, so no sorting hypothesis is needed. If
    equality at exponent $0$ is also assumed, the nonzero-entry hypothesis can be
    omitted because $\sum_k \lambda_{a,k}^0=x_a$ and
    $\sum_k \lambda_{b,k}^0=x_b$.
\end{remark}

\begin{theorem}[Finite-range equal power sums imply equal multisets]
    \label{thm:bounded_power_sum_multiset}
    \lean{Matrix.sum_pow_eq_implies_multiset_eq_of_le_card,
        Matrix.sum_pow_eq_implies_nonzero_multiset_eq_of_le_max_length,
        Matrix.sum_pow_eq_implies_multiset_eq_of_le_max_length,
        Matrix.sum_pow_eq_implies_card_eq_and_multiset_eq_of_le_max_card}
    \leanok
    Let $\alpha, \beta : \{0, \ldots, n-1\} \to \C$.
    If $\sum_i \alpha_i^k = \sum_i \beta_i^k$ for $1 \le k \le n$,
    then the multisets $\{\alpha_i\}$ and $\{\beta_i\}$ are equal.
    More generally, if $\alpha : \{0, \ldots, m-1\} \to \C$ and
    $\beta : \{0, \ldots, n-1\} \to \C$ have no zero entries and their power sums
    agree for $1 \le k \le \max(m,n)$, then $m=n$ and the two multisets
    are equal.
    There is also an exponent-zero list form: for lists
    $(\lambda_{a,k})_{k=1}^{x_a}$ and
    $(\lambda_{b,k})_{k=1}^{x_b}$, the hypothesis
    $\sum_{k=1}^{x_a} \lambda_{a,k}^N
        = \sum_{k=1}^{x_b} \lambda_{b,k}^N$ for
    $1 \le N \le \max(x_a,x_b)$ implies
    \[
        \{\lambda_{a,k} \mid \lambda_{a,k}\neq 0\}_{k=1}^{x_a}
        =
        \{\lambda_{b,k} \mid \lambda_{b,k}\neq 0\}_{k=1}^{x_b}
    \]
    as multisets. If the equality is also assumed at $N=0$, then the original
    lists have the same multiset.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:newton_girard}
    Apply the finite-range Newton--Girard theorem to
    $\diag(\alpha)$ and $\diag(\beta)$, then extract the roots of
    $\prod_i (X - \alpha_i)$ and $\prod_i (X - \beta_i)$. For unequal
    cardinalities, pad both lists with zeros to length $\max(m,n)$. The
    equal-root conclusion compares the padded multisets, and the nonzero-entry
    hypothesis makes the number of padded zeros exactly $\max(m,n)-m$ and
    $\max(m,n)-n$, forcing $m=n$.
    Without a nonzero-entry hypothesis, first delete the zero entries; for
    positive $N$, this does not change
    $\sum_k \lambda_k^N$. The nonzero-entry form then compares the remaining
    multisets. In the exponent-zero list form, the equality at $N=0$ gives
    $x_a=x_b$, so the same-cardinality finite-range theorem applies directly.
\end{proof}

\begin{theorem}[Equal power sums imply equal multisets]\label{thm:power_sum_multiset}
    \lean{Matrix.sum_pow_eq_implies_multiset_eq}
    \leanok
    Let $\alpha, \beta : \{0, \ldots, n-1\} \to \C$.
    If $\sum_i \alpha_i^k = \sum_i \beta_i^k$ for all $k \ge 1$,
    then the multisets $\{\alpha_i\}$ and $\{\beta_i\}$ are equal.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:newton_girard}
    Construct diagonal matrices $\diag(\alpha)$ and
    $\diag(\beta)$.
    The power-sum hypothesis gives
    $\tr(\diag(\alpha)^k) = \tr(\diag(\beta)^k)$.
    By Theorem~\ref{thm:newton_girard}, the characteristic
    polynomials agree: $\prod_i (X - \alpha_i) = \prod_i (X - \beta_i)$.
    Extracting roots gives multiset equality.
\end{proof}

\begin{remark}
    Theorems~\ref{thm:newton_girard} and~\ref{thm:power_sum_multiset}
    remain relevant for the equal-MPV source corollary discussed in
    Section~\ref{sec:ft_equal_mpv_comparisons}, where the paper uses a power-sum
    step to recover the multiplicities $r_j$ and the weights $\mu_{j,q}$.
    The BNT multiplicity lemmas handle this recovery directly rather than through
    a separate explicit-coefficient equal-case theorem.
\end{remark}

\section{The BNT canonical form on a sector decomposition}
\label{sec:sector_bnt_api}

This section introduces the BNT canonical-form predicate following
\cite[\S II, lines~287--301]{Cirac2016MPDO_arXiv} and
\cite[Definition~4.2, Proposition~4.3, and lines~1864--1884]{Cirac2021Matrix}.
The predicate is
defined on a sector decomposition with raw sector weights $\mu_{j,q}$ and
sector coefficients $c_N^{(j)} = \sum_q \mu_{j,q}^{\,N}$; no equal-modulus
factorization is assumed.
% Maintainer note (Lean): this is the predicate `IsBNTCanonicalForm` on a
% `SectorDecomposition`. The blueprint prose uses standard mathematical
% notation; the `\lean{...}` tag below carries the link.

\subsection{Sector decompositions and phase matching}
\label{sec:bnt_sector_decompositions}

The sector-decomposition data type, the MPV phase-equivalence predicates, and the
MPV phase-class apparatus are collected here because they are used both by the
BNT canonical-form predicate below and by the equal-MPV comparison of
Chapter~\ref{ch:ft_proof}.

\begin{definition}[Sector weights and sector decomposition]%
    \label{def:sector_weight_data}
    \lean{MPSTensor.SectorWeightData, MPSTensor.SectorDecomposition}
    \leanok
    A sector-weight tuple over $g$ basis blocks consists of multiplicities
    $r_j > 0$, nonzero weights $\mu_{j,q}$ for
    $q \in \{1,\ldots,r_j\}$, and the sector coefficients
    \[
        c_{S,j}^{(N)}
        :=
        \sum_{q=1}^{r_j} \mu_{j,q}^N.
    \]
    A \emph{sector decomposition} consists of the basis blocks
    $(A_j)_{j=1}^g$ together with these multiplicities and weights.
\end{definition}

\begin{lemma}[Sector decomposition formula]%
    \label{lem:paper_decomp_formula}
    \lean{MPSTensor.SectorDecomposition.mpv_toTensor_eq_sum_coeff}
    \leanok
    \uses{def:sector_weight_data}
    For a sector decomposition $S$ with associated tensor $A_{\mathrm{tot}}$,
    the MPV satisfies
    \[
        V^{(N)}(A_{\mathrm{tot}})_\sigma
        =
        \sum_{j=1}^{g} c_{S,j}^{(N)} V^{(N)}(A_j)_\sigma.
    \]
\end{lemma}

\begin{definition}[MPV phase equivalence for nonzero-weight blocks]
    \label{def:mpv_block_phase_equiv}
    \lean{MPSTensor.MPVBlockPhaseEquiv}
    \leanok
    \uses{def:mpv}
    Two blocks $A$ and $B$ are MPV-phase equivalent if there is a nonzero
    scalar $\zeta$ such that, for every system size~$N$ and every physical word,
    \[
        V^{(N)}(B)_\sigma
        =
        \zeta^N V^{(N)}(A)_\sigma.
    \]
    The two bond dimensions are allowed to differ.
\end{definition}

\begin{definition}[Common MPV phase cover]
    \label{def:mpv_common_phase_cover}
    \lean{MPSTensor.MPVCommonPhaseCover}
    \leanok
    \uses{def:mpv_block_phase_equiv}
    For two finite nonzero-weight block families, a common MPV phase cover consists of a third
    finite family of blocks, a map from each nonzero-weight block family onto that common
    family, and MPV-phase equivalences between each block and its common
    representative. Both maps are required to be surjective, so the common family
    contains no unused representative.
\end{definition}

\begin{definition}[MPV phase classes]%
    \label{def:mpv_phase_class_data}
    \lean{MPSTensor.MPVPhaseEquiv, MPSTensor.MPVPhaseClassData, MPSTensor.mpvPhaseClassData}
    \leanok
    For a finite family of blocks, put $j \sim k$ when there is a nonzero
    scalar factor $\zeta$ such that
    \[
        \ket{V^{(N)}(B_k)}
        =
        \zeta^N\ket{V^{(N)}(B_j)}
    \]
    for every length~$N$. The associated tuple enumerates the finite quotient,
    chooses one representative per class, records the scalar relating the
    representative to each member, and proves that distinct representatives are
    not MPV-phase equivalent.
\end{definition}

\begin{definition}[BNT canonical form]%
    \label{def:sector_bnt_cf}
    \lean{MPSTensor.IsBNTCanonicalForm}
    \leanok
    \uses{def:sector_weight_data}
    Given a sector decomposition $P$, the BNT canonical-form predicate
    records the minimal BNT hypotheses of
    \cite{Cirac2016MPDO_arXiv,Cirac2021Matrix} without any equal-modulus
    or strict-ordering condition on the raw sector weights. It asserts that
    each basis block has positive bond dimension, is irreducible,
    and left-canonical; the normalized self-overlap of every basis block
    converges to $1$; the basis MPV family is eventually linearly
    independent; distinct same-dimension basis blocks are not gauge-phase
    equivalent after identifying equal bond dimensions; and the
    \cite[\S II.C, line~246]{Cirac2016MPDO_arXiv} normalization holds,
    \[
        \forall (j,q),\ |\mu_{j,q}| \le 1,
        \qquad
        \exists\, (j_*, q_*) : |\mu_{j_*, q_*}| = 1.
    \]
    The MPV expansion takes the raw two-layer form
    \[
        V^{(N)}(P)_\sigma
        =
        \sum_{j=1}^{g} \left(\sum_{q=1}^{r_j} \mu_{j,q}^N\right)
        V^{(N)}(P_j)_\sigma,
    \]
    matching \cite[\S~II.C, lines~271--301]{Cirac2016MPDO_arXiv}
    and \cite[Definition~4.2, Proposition~4.3, and lines~1864--1884]
    {Cirac2021Matrix}.
    The line-246 normalization admits every example of the source paper
    (in particular $C \oplus D$, $C \oplus (-C)$,
    $C \oplus (1/2) C$, and $C \oplus (-C) \oplus (1/2) C$) and is
    not derivable from the per-block normality hypotheses alone.
\end{definition}

\subsection{Prepared-block construction of BNT canonical forms}%
\label{subsec:paperbnt_supplier}

The CPSV basis-of-normal-tensors predicate in Definition~\ref{def:bnt_cpsv}
is strengthened by the sector-decomposition predicate of
Definition~\ref{def:sector_bnt_cf}, which records multiplicities, span, and
weight-norm conditions. The results below separate the after-blocking
preparation of suitable blocks from the normalized-weight BNT construction.

\begin{lemma}[Common injective blocking for primitive irreducible families]%
    \label{thm:common_injective_blocking_tp_primitive_irreducible_family}
    \lean{MPSTensor.exists_common_injective_blocking_of_tp_primitive_irr_family}
    \leanok
    \uses{thm:tp_primitive_irred_block_injective,
        thm:tp_primitive_irreducible_extra_blocking}
    Let $(B_k)_{k=1}^r$ be a finite family of positive-bond-dimension MPS tensors,
    each trace-preserving, primitive, and tensor-irreducible. Then there is a
    single $L\ge 1$ such that each $B_k^{[L]}$ is one-site injective, and this
    blocking preserves trace preservation, primitivity, and irreducibility for
    every block.
\end{lemma}

\begin{proof}\leanok
    For each block,
    Theorem~\ref{thm:tp_primitive_irred_block_injective} gives a positive length
    at which it becomes one-site injective; the product of these lengths is a
    common length. Injectivity persists under positive further blocking, and
    Theorem~\ref{thm:tp_primitive_irreducible_extra_blocking} preserves the other
    hypotheses.
\end{proof}

\begin{theorem}[Prepared BNT blocks after blocking]%
    \label{thm:prepared_bnt_blocks_after_blocking}
    \lean{MPSTensor.exists_prepared_BNT_blocks_afterBlocking_pos}
    \leanok
    \uses{thm:unconditional_common_primitive_irreducible_blocks,
        thm:common_injective_blocking_tp_primitive_irreducible_family}
    Let $A$ be an MPS tensor. Then there is a blocking length $p\ge 1$, nonzero
    weights $\mu_k$, and blocks $B_k$ over the $p$-blocked alphabet such that:
    \begin{enumerate}
        \item each block has positive bond dimension;
        \item each block is trace-preserving, primitive, tensor-irreducible, and
              one-site injective;
        \item $A^{[p]}$ and $\bigoplus_k \mu_k B_k$ generate the same MPV family
              at every positive length.
    \end{enumerate}
    No weight-modulus normalization is asserted at this stage.
\end{theorem}

\begin{proof}\leanok
    Apply the common primitive irreducible block decomposition to $(A,A)$,
    giving a positive blocking length and a nonzero weighted family of
    trace-preserving primitive irreducible blocks with the positive-length MPV
    identity. Lemma~\ref{thm:common_injective_blocking_tp_primitive_irreducible_family}
    supplies one common further blocking making every block one-site injective
    while preserving the other hypotheses; finally identify the iterated blocked
    alphabet with the direct one.
\end{proof}

\begin{theorem}[BNT from normal blocks with normalized weights]%
    \label{thm:paperbnt_supplier_prepared}
    \lean{MPSTensor.exists_isBNTCanonicalForm_of_tp_primitive_irr_blocks}
    \leanok
    Let $r \in \N$, and for each $k = 1,\ldots,r$ let $D_k \in \N$,
    $\mu_k \in \C$, and let $B_k$ be an MPS tensor of physical dimension $d$ and
    bond dimension $D_k$. Assume:
    \begin{enumerate}
        \item \textbf{positive bond dimensions:} $D_k > 0$ for all $k$,
        \item \textbf{TP normalization:} $\sum_i (B_k^i)^\dagger B_k^i = \Id$,
        \item \textbf{primitivity:} $\E_{B_k}$ is primitive for each $k$,
        \item \textbf{irreducibility:} each $B_k$ is irreducible,
        \item \textbf{nonzero weights:} $\mu_k \neq 0$ for all $k$,
        \item \textbf{bounded weights:} $|\mu_k| \le 1$ for all $k$,
        \item \textbf{global unit witness:} $|\mu_k| = 1$ for some $k$.
    \end{enumerate}
    Then there exists a sector decomposition $P$ such that $P$ and
    $\bigoplus_{k=1}^{r} \mu_k B_k$ generate the same MPV family at every length,
    and $P$ satisfies the BNT conditions of Definition~\ref{def:bnt_cpsv}
    together with the multiplicity, span, and weight-bound conditions of
    Definition~\ref{def:sector_bnt_cf}.
\end{theorem}

\begin{proof}\leanok
    The phase-class quotient groups gauge-phase-equivalent blocks into sectors
    with unit-modulus per-copy scalars. Under the TP, primitive, and irreducible
    hypotheses the phase-class construction applies, and for distinct
    representatives $B_j$, $B_{j'}$ ($j \neq j'$) the cross-overlap decays,
    \[
        \lim_{N\to\infty} O_{B_j B_{j'}}(N) = 0,
        \qquad
        \lim_{N\to\infty} O_{B_j B_j}(N) = 1,
    \]
    which gives the eventual linear-independence condition with no injectivity
    required. The resulting sector decomposition inherits the span and
    multiplicity conditions, and the bounded-weight and global-unit hypotheses
    give the weight-norm conditions.
\end{proof}

\begin{lemma}[Total dimension of a dimension-preserving phase-class quotient]%
    \label{lem:collapsed_bnt_sector_decomp_total_dim}
    \lean{MPSTensor.collapsedBntSectorDecomp_totalDim_eq_sum_dim}
    \leanok
    In the phase-class quotient of the prepared-block BNT supplier, suppose every
    original copy weight is nonzero and every block in a phase class has the bond
    dimension of its representative. Then the total bond dimension of the
    collapsed sector decomposition equals the sum of the bond dimensions of the
    original prepared blocks.
\end{lemma}

\begin{proof}\leanok
    Sum the representative bond dimensions over the copies in each phase class.
    The same-dimension hypothesis replaces each representative dimension by the
    enumerated original block dimension, and the phase-class regrouping identity
    recovers the original direct-sum dimension.
\end{proof}

\begin{lemma}[Prepared phase-equivalent blocks have the same bond dimension]%
    \label{lem:prepared_phase_equiv_dim_eq}
    \lean{MPSTensor.dim_eq_of_MPVBlockPhaseEquiv_of_tp_primitive_irr}
    \leanok
    % Source: Lemma equalMPS (Cirac2016MPDO_arXiv), dimension conclusion.
    Let $X$ and $Y$ be trace-preserving, primitive, irreducible MPS blocks of
    positive bond dimension. If their MPV families differ by a nonzero scalar
    power, then their bond dimensions are equal.
\end{lemma}

\begin{proof}\leanok\uses{thm:overlap_decay_rect_irr_tp}
    Self-overlap normalization forces the scalar relating the two MPV families to
    have unit modulus, so the rectangular overlap has norm tending to $1$. If the
    bond dimensions differed, the rectangular overlap-decay theorem would force
    that overlap to tend to $0$.
\end{proof}

\begin{lemma}[Prepared phase classes preserve bond dimension]%
    \label{lem:prepared_phase_class_dim_eq}
    \lean{MPSTensor.mpvPhaseClassData_dim_eq_of_tp_primitive_irr}
    \leanok
    % Source: prop:char-BNT (Cirac2017MPDO_AnnPhys), with appendix restatement
    % and the same-CF BNT uniqueness note.
    In a prepared family of positive-bond-dimension trace-preserving primitive
    irreducible blocks, every block in an MPV phase class has the bond dimension
    of its representative.
\end{lemma}

\begin{proof}\leanok\uses{lem:prepared_phase_equiv_dim_eq}
    Apply Lemma~\ref{lem:prepared_phase_equiv_dim_eq} to each representative and
    each member of its phase class.
\end{proof}

\begin{lemma}[Prepared collapsed BNT total dimension]%
    \label{lem:prepared_collapsed_bnt_sector_decomp_total_dim}
    \lean{MPSTensor.collapsedBntSectorDecomp_totalDim_eq_sum_dim_of_tp_primitive_irr}
    \leanok
    % Source: canonical-form eq:II_CF1 (Cirac2017MPDO_AnnPhys) and the
    % phase-class quotient prop:char-BNT with its appendix restatement.
    For prepared trace-preserving primitive irreducible blocks of positive bond
    dimension with all copy weights nonzero, the phase-class quotient preserves
    the total bond dimension: the collapsed sector decomposition has total bond
    dimension equal to the sum of the original block dimensions.
\end{lemma}

\begin{proof}\leanok\uses{lem:collapsed_bnt_sector_decomp_total_dim,
        lem:prepared_phase_class_dim_eq}
    Lemma~\ref{lem:prepared_phase_class_dim_eq} supplies the same-dimension
    hypothesis in Lemma~\ref{lem:collapsed_bnt_sector_decomp_total_dim}.
\end{proof}

\begin{theorem}[Conditional BNT form after blocking for an arbitrary tensor]%
    \label{thm:paperbnt_supplier_after_blocking}
    \lean{MPSTensor.exists_isBNTCanonicalForm_afterBlocking_pos}
    \leanok
    \uses{thm:prepared_bnt_blocks_after_blocking,
        thm:paperbnt_supplier_prepared}
    Let $A$ be an MPS tensor. Then there exists a blocking length $p \ge 1$,
    nonzero weights $\mu_k$, and blocks $B_k$ over the blocked alphabet, such
    that:
    \begin{enumerate}
        \item every block has positive bond dimension;
        \item every block is TP-normalized, primitive, irreducible, and
              injective;
        \item $A^{[p]}$ and $\bigoplus_k \mu_k B_k$ generate the same MPV family
              at every positive length.
    \end{enumerate}
    Moreover, under $|\mu_k| \le 1$ for every $k$ and $|\mu_k|=1$ for at least
    one $k$, there exists a sector decomposition $P$ over the blocked alphabet in
    the strengthened BNT canonical form of Definition~\ref{def:sector_bnt_cf}
    such that
    $A^{[p]}$ and $P$ generate the same MPV family at every positive length.
\end{theorem}

\begin{proof}\leanok
    The after-blocking reduction supplies the TP-normalized primitive irreducible
    injective blocks and the positive-length MPV identity. The normalized-weight
    hypotheses are the choice that
    \cite[Section~II.C]{Cirac2017MPDO_AnnPhys} makes;
    Theorem~\ref{thm:paperbnt_supplier_prepared} then produces the sector
    decomposition, following the BNT selection of
    \cite[Appendix~A]{Cirac2017MPDO_AnnPhys}. The missing rescaling step is
    recorded in Remark~\ref{rem:arbitrary_input_reduction_gap}.
\end{proof}

\begin{theorem}[Conditional BNT form after blocking, completed at length zero]%
    \label{thm:paperbnt_supplier_after_blocking_total_dim}
    \lean{MPSTensor.exists_isBNTCanonicalForm_afterBlocking_of_totalDim}
    \leanok
    \uses{thm:paperbnt_supplier_after_blocking,
        def:same_mpv2,
        def:same_mpv2_pos,
        lem:samempv2pos_to_samempv2_of_bonddim_eq}
    In the setting of Theorem~\ref{thm:paperbnt_supplier_after_blocking}, the same
    prepared blocks may be chosen so that the positive-length MPV identity
    upgrades to all lengths whenever the total bond dimension of $P$ equals the
    bond dimension of the blocked input tensor.
\end{theorem}

\begin{proof}\leanok
    The positive-length statement is
    Theorem~\ref{thm:paperbnt_supplier_after_blocking}. At length zero the MPV
    coefficient is the bond dimension, so equality of the two total bond
    dimensions supplies the missing length-zero coefficient, and the
    positive-length identity gives all remaining lengths.
\end{proof}

\begin{remark}[Arbitrary-input reduction]%
    \label{rem:arbitrary_input_reduction_gap}
    Theorem~\ref{thm:paperbnt_supplier_after_blocking} gives the arbitrary-input
    reduction, conditional on the two weight-normalization hypotheses displayed
    there. The preparatory part,
    Theorem~\ref{thm:prepared_bnt_blocks_after_blocking}, produces the
    TP-normalized primitive irreducible injective blocks and the positive-length
    MPV identity but not the weight normalization. Since the all-zero summand
    contributes only at length zero, the nonzero primitive blocks identify the
    MPV family for $N\ge 1$. The proof that the prepared weights can always be
    rescaled to satisfy $|\mu_k|\le 1$ with a unit-modulus witness is the one
    remaining step, recorded in
    \texttt{docs/paper-gaps/cpsv16\_cf\_normalization\_and\_proportional\_comparison.tex};
    the length-zero trace identity and the fully unblocked canonical form remain
    part of the reduction in Section~\ref{sec:reduction_to_normal_form}.
\end{remark}


\begin{lemma}[Raw two-layer sector coefficient identity]%
    \label{lem:sector_bnt_coeff_eq_sum_weight_pow}
    \lean{MPSTensor.SectorDecomposition.coeff_eq_sum_weight_pow}
    \leanok
    \uses{def:sector_weight_data}
    For every sector decomposition $P$, every length $N$, and every basis
    index $j$, the sector coefficient is the raw power sum
    \[
        c_N^{(j)} = \sum_{q=1}^{r_j} \mu_{j,q}^N.
    \]
    This is
    \cite[two-layer BNT and coefficient-expansion displays,
        lines~287--301]{Cirac2016MPDO_arXiv} and
    \cite[lines~1864--1884]{Cirac2021Matrix}.
\end{lemma}

\begin{proof}\leanok
    Definitional unfolding of the sector coefficient.
\end{proof}

\begin{lemma}[Sector coefficient bound by multiplicity]%
    \label{lem:sector_bnt_norm_coeff_le_copies}
    \lean{MPSTensor.SectorDecomposition.norm_coeff_le_copies_of_norm_weight_le_one}
    \leanok
    \uses{lem:sector_bnt_coeff_eq_sum_weight_pow}
    Under the optional global modulus normalization
    \cite[\S II, lines~217--246]{Cirac2016MPDO_arXiv}, every raw weight satisfies
    $|\mu_{j,q}| \le 1$, and the triangle inequality bounds
    the norm of the sector coefficient by the multiplicity:
    \[
        \bigl| c_N^{(j)} \bigr| \le r_j.
    \]
\end{lemma}

\begin{proof}\leanok
    Apply the triangle inequality to the raw power sum from
    Lemma~\ref{lem:sector_bnt_coeff_eq_sum_weight_pow} and bound each term by
    $1$ using $|\mu_{j,q}| \le 1$.
\end{proof}

\begin{lemma}[Cross-overlap decay between distinct basis blocks]%
    \label{lem:sector_bnt_cross_overlap_basis_tendsto_zero}
    \lean{MPSTensor.IsBNTCanonicalForm.cross_overlap_basis_tendsto_zero}
    \leanok
    \uses{def:mpv_overlap, def:sector_bnt_cf}
    If $P$ is a BNT canonical form and $j \neq k$, then
    $O_{P_j P_k}(N) \to 0$ as $N \to \infty$, where $P_j$ and $P_k$ are the
    basis blocks at indices $j$ and $k$. The argument follows the
    normal-tensor overlap dichotomy of
    \cite[lines~1080--1091]{Cirac2016MPDO_arXiv}: differing bond dimensions
    force decay, and matching bond dimensions are ruled out by the
    gauge-phase distinctness clause of
    \cite[lines~264--279]{Cirac2016MPDO_arXiv}.
\end{lemma}

\begin{proof}\leanok
    Case on whether the bond dimensions match. If they differ, apply the
    dimension-mismatch overlap-decay theorem. If they agree, the
    basis-distinctness clause of the BNT canonical form rules out
    gauge-phase equivalence after identifying the common bond dimension, so the
    equal-dimension overlap-decay theorem
    applies.
\end{proof}

\begin{lemma}[Combined-family eventual linear independence]%
    \label{lem:sector_bnt_combined_family_eventually_li}
    \lean{MPSTensor.IsBNTCanonicalForm.combined_family_eventually_li}
    \leanok
    \uses{lem:bnt_two_family_overlap_orthonormal_li,
        lem:sector_bnt_cross_overlap_basis_tendsto_zero}
    Let $P$ and $Q$ each be a BNT canonical form, and suppose every mixed
    overlap $O_{P_j Q_k}(N)$ tends to $0$. Then the union family
    \[
        \bigl\{\ket{V^{(N)}(P_j)}\bigr\}_{j=1}^{g_P}
        \cup
        \bigl\{\ket{V^{(N)}(Q_k)}\bigr\}_{k=1}^{g_Q}
    \]
    is linearly independent for all sufficiently large $N$. This is the
    corollary at \cite[lines~1121--1132]{Cirac2016MPDO_arXiv}, matching
    the BNT linear-independence input recorded at
    \cite[line~1850]{Cirac2021Matrix}.
\end{lemma}

\begin{proof}\leanok
    Feed the per-family normalized self-overlaps (from the canonical-form
    hypotheses), the within-family cross-overlap decay
    (Lemma~\ref{lem:sector_bnt_cross_overlap_basis_tendsto_zero}), and the
    inter-family decay hypothesis into the two-family overlap-orthonormality
    criterion (Lemma~\ref{lem:bnt_two_family_overlap_orthonormal_li}).
\end{proof}

\begin{lemma}[Sector coefficient bound from the line-246 normalization]%
    \label{lem:sector_bnt_norm_coeff_le_copies_struct}
    \lean{MPSTensor.IsBNTCanonicalForm.norm_coeff_le_copies}
    \leanok
    \uses{def:sector_bnt_cf, lem:sector_bnt_norm_coeff_le_copies}
    Under the BNT canonical form, the
    \cite[\S~II.C, line~246]{Cirac2016MPDO_arXiv} normalization yields
    directly
    \[
        \bigl| c_N^{(j)} \bigr| \le r_j
    \]
    for every $N$ and basis index $j$, without an additional explicit
    modulus-bound hypothesis.
\end{lemma}

\begin{proof}\leanok
    Feed the global modulus normalization hypothesis into
    Lemma~\ref{lem:sector_bnt_norm_coeff_le_copies}.
\end{proof}

\begin{lemma}[Unit-modulus weight witness from the line-246 normalization]%
    \label{lem:sector_bnt_weight_unit_exists_of_struct}
    \lean{MPSTensor.IsBNTCanonicalForm.weight_unit_exists_of_struct}
    \leanok
    \uses{def:sector_bnt_cf}
    Under the BNT canonical form, the
    \cite[\S~II.C, line~246]{Cirac2016MPDO_arXiv} unit-modulus existential
    provides indices $(j_*, q_*)$ with
    $|\mu_{j_*, q_*}| = 1$.
\end{lemma}

\begin{proof}\leanok
    Direct projection onto the unit-modulus existential hypothesis of the
    canonical-form predicate.
\end{proof}

\begin{example}[$C \oplus (1/2)C$, unequal-modulus admissibility]%
    \label{ex:sector_bnt_halvedDecomp}
    \lean{MPSTensor.SectorBNT.Examples.halvedDecomp}
    \leanok
    \uses{def:sector_bnt_cf}
    The single-sector two-copy decomposition with raw weights
    $(1, 1/2)$ demonstrates that the BNT canonical form admits
    unequal-modulus copies: the modulus bound holds because
    $|1| \le 1$ and $|1/2| = 1/2 \le 1$, and the
    unit-modulus existential is witnessed by the first copy with
    weight $1$. This is the
    \cite[\S~II.C, line~246]{Cirac2016MPDO_arXiv} normalization in
    positive form, applied to the $C \oplus (1/2)C$ example.
\end{example}

\section{Ces\`aro non-decay and unit-block coefficient}
\label{sec:sector_bnt_cesaro_nondecay}

This section formalizes the analytic input that supplies, for any
sector $j_0$ carrying a unit-modulus weight, the
``sector-$j_0$ coefficient does not asymptotically vanish''
ingredient used in the matching step of
\cite[Appendix MPV proof, line~1182]{Cirac2016MPDO_arXiv}.
The reduction is to a pure-analytic statement: a finite sum of $N$-th
powers of complex numbers in the closed unit disk, with at least one
unit-modulus summand, cannot tend to zero.

The proof is the elementary Ces\`aro-mean identity
$M_T := T^{-1} \sum_{N=0}^{T-1} |c(N)|^2
    = \sum_{p,q} T^{-1} \sum_{N=0}^{T-1} (\mu_p \overline{\mu_q})^N$,
which splits into:
$T^{-1}$ times the partial geometric sum for each pair $(p, q)$; each
pair with $\mu_p \overline{\mu_q} = 1$ contributes the limit $1$, while
each pair with $\mu_p \overline{\mu_q} \neq 1$ contributes the limit
$0$ (since $|(z^T - 1)/(z - 1)| \le 2/|z - 1|$ divided by $T \to 0$).
The diagonal pair $(q^*, q^*)$ at a unit-modulus index $q^*$ contributes,
so $M_T \to \#\{\text{matching pairs}\} \ge 1 > 0$, contradicting
$c(N) \to 0 \Rightarrow |c(N)|^2 \to 0 \Rightarrow M_T \to 0$.

\begin{lemma}[Ces\`aro non-decay of a finite power sum]%
    \label{lem:sector_bnt_cesaro_nondecay}
    \lean{MPSTensor.CesaroNonDecay.sum_pow_not_tendsto_zero_of_unit_modulus}
    \leanok
    Let $\mu : \{0, \ldots, r-1\} \to \C$ satisfy $|\mu_q| \le 1$
    for every $q$ and let some $q^*$ satisfy $|\mu_{q^*}| = 1$. Then the
    sequence $N \mapsto \sum_{q=0}^{r-1} (\mu_q)^N$ does not tend to $0$
    as $N \to \infty$. This is the analytic content of the unit-block
    non-decay step at
    \cite[\S~II.C, line~246]{Cirac2016MPDO_arXiv} and
    \cite[Appendix MPV proof, line~1182]{Cirac2016MPDO_arXiv}.
\end{lemma}

\begin{proof}\leanok
    Ces\`aro-mean argument as described above.
\end{proof}

\begin{lemma}[Unit-block coefficient non-decay]%
    \label{lem:sector_bnt_coeff_not_tendsto_zero_at_unit_block}
    \lean{MPSTensor.IsBNTCanonicalForm.coeff_not_tendsto_zero_at_block}
    \leanok
    \uses{def:sector_bnt_cf, lem:sector_bnt_cesaro_nondecay}
    Let $P$ be a BNT canonical form and let $j_0$ be a sector index such
    that some copy of sector $j_0$ carries a unit-modulus weight, that is,
    there exists $q$ with $|\mu_{j_0, q}| = 1$. Then the
    sector-$j_0$ coefficient $c_N^{(j_0)} = \sum_q \mu_{j_0, q}^N$ does not
    tend to $0$ as $N \to \infty$.

    This reads the \cite[\S~II.C, line~246]{Cirac2016MPDO_arXiv}
    normalization sector by sector, together with the convergence use at
    \cite[Appendix ``Proofs of Section MPS'', line~1244]
    {Cirac2016MPDO_arXiv}. The unit-modulus witness is supplied
    externally (as the hypothesis above) because the source statement is
    global (``at least one of them equals one'' over the two-layer index
    $(j, q)$) and does not pin the witness to a specific sector.
\end{lemma}

\begin{proof}\leanok
    Apply Lemma~\ref{lem:sector_bnt_cesaro_nondecay} to the family
    $q \mapsto \mu_{j_0, q}$, using the modulus bound
    $|\mu_{j_0, q}| \le 1$ from the canonical-form normalization
    together with the per-sector unit-modulus hypothesis.
\end{proof}

\begin{remark}
    Used inside the matching theorem of
    \cite[Appendix MPV proof, line~1182]{Cirac2016MPDO_arXiv} this lemma
    supplies the non-decay step on $c_N^{(j_0)}$. The matching theorem takes
    the unit-modulus sector and witness as explicit parameters, mirroring
    the global character of the source normalization at
    \cite[\S~II.C, line~246]{Cirac2016MPDO_arXiv}.
\end{remark}

\subsection{Matched-sector weight multiset equality}

Given a matched basis-sector pair $(j_0, \tilde k_0)$ and a nonzero
scalar $\zeta \in \C \setminus \{0\}$ such that the matched
coefficient identity
$c_N^{(j_0)}(P) = \zeta^N \cdot c_N^{(\tilde k_0)}(Q)$
holds for every $N$ past some threshold $N_0$, this subsection
records the two algebraic conclusions: the \emph{matched-sector weight
    multiset equality} derived from the coefficient identity by power-sum
recovery (Theorem~\ref{thm:power_sum_multiset}), and the
\emph{matched-sector weight-permutation extractor} upgrading the multiset
equality to an explicit per-copy indexing. Together they realise the
$\mu_{j,q} = \nu_{j,q} \cdot e^{i\varphi_j}$ recovery of
\cite[Appendix MPV proof, lines~1184--1188]{Cirac2016MPDO_arXiv}. At the source level,
the preceding block-selection argument supplies the matched normal sectors
\cite[Appendix MPV proof, line~1182]{Cirac2016MPDO_arXiv}; the sector-coefficient identity
used here is the equal-MPV coefficient comparison of
\cite[Appendix MPV proof, lines~1184--1188]{Cirac2016MPDO_arXiv}, followed by the
global-gauge construction in
\cite[Appendix MPV proof, lines~1189--1192]{Cirac2016MPDO_arXiv}. The coefficient identity
is supplied, in the equal-MPV case, by
Lemma~\ref{lem:sector_bnt_coeff_identity_via_matched_mpv_phase} below. In
the proportional case it is kept as the remaining hypothesis of
Theorem~\ref{thm:sector_bnt_proportional_global_gauge_of_coeff_identity}.

\begin{lemma}[Matched-sector weight multiset equality]%
    \label{lem:sector_bnt_matched_sector_weight_multiset_eq}
    \lean{MPSTensor.matched_sector_weight_multiset_eq}
    \leanok
    \uses{def:sector_weight_data,
        thm:power_sum_multiset}
    Let $P, Q$ be sector decompositions with weights $\mu_{j,q}$ and
    $\nu_{k,q'}$, and fix indices $j_0$, $\tilde k_0$, and a nonzero scalar
    $\zeta \in \C \setminus \{0\}$. Suppose that for every $N > N_0$,
    \[
        c_N^{(j_0)}(P) = \zeta^N \cdot c_N^{(\tilde k_0)}(Q).
    \]
    Then the per-sector multiplicities agree, $r_{j_0}(P) = r_{\tilde k_0}(Q)$,
    and the rescaled per-copy weight multisets coincide.
    This is \cite[Appendix MPV proof, lines~1184--1188]{Cirac2016MPDO_arXiv}
    ($\mu_{j,q} = \nu_{j,q} \cdot e^{i\varphi_j}$ and matching
    multiplicities $r_{a,j} = r_{b,j}$) read on a single matched sector.
\end{lemma}

\begin{proof}\leanok
    \uses{thm:bounded_power_sum_multiset}
    Unfold both sides of the coefficient identity to per-copy power sums
    $\sum_q \mu_{j_0, q}^N$ and $\sum_{q'} \nu_{\tilde k_0, q'}^N$, and
    distribute $\zeta^N$ through the right-hand sum to obtain, for every
    $N > N_0$,
    \[
        \sum_q \mu_{j_0, q}^N = \sum_{q'} (\zeta \cdot \nu_{\tilde k_0, q'})^N.
    \]
    Extend to all positive exponents and apply the unequal-cardinality
    multiset-recovery lemma (Theorem~\ref{thm:bounded_power_sum_multiset}),
    using $\zeta \neq 0$ together with the weight non-vanishing assumption
    to ensure the input weights are nonzero. This yields the multiplicity
    match and the multiset equality.
\end{proof}

\begin{lemma}[Matched-sector weight-permutation extractor]%
    \label{lem:sector_bnt_matched_sector_weight_equiv}
    \lean{MPSTensor.matched_sector_weight_equiv}
    \leanok
    \uses{lem:sector_bnt_matched_sector_weight_multiset_eq}
    Under the hypotheses of
    Lemma~\ref{lem:sector_bnt_matched_sector_weight_multiset_eq}, the
    per-sector multiplicities agree and there is a permutation $\tau$ of
    the per-copy index set identifying the individual weights up to the
    gauge phase: for every $q$,
    \[
        \nu_{\tilde k_0, \tau(q)} = \zeta^{-1} \cdot \mu_{j_0, q}.
    \]
    This is the per-copy realisation of
    \cite[Appendix MPV proof, line~1188]{Cirac2016MPDO_arXiv}
    ($\mu_{j,q} = \nu_{j,q} \cdot e^{i\varphi_j}$): the multiset equality
    from Lemma~\ref{lem:sector_bnt_matched_sector_weight_multiset_eq} is
    upgraded to a concrete indexing between the matched $P$- and
    $Q$-copies.
\end{lemma}

\begin{proof}\leanok
    Invoke Lemma~\ref{lem:sector_bnt_matched_sector_weight_multiset_eq} to
    obtain the cardinality match and the multiset equality.
    From this multiset-map equality extract a permutation matching the
    elements pointwise (proved by induction on the cardinality).
    Compose with the cardinality identification and cancel $\zeta$ via
    $\zeta \neq 0$ to convert the gauge factor on the right of the
    pointwise identity to $\zeta^{-1}$ on the left.
\end{proof}

\section{Strong existential matching by exact linear independence}
\label{sec:sector_bnt_strong_match}

This section lifts the weak existential matching $\exists j, \exists k$ to
the strong existential $\forall k, \exists j$ on the original pair of BNT
canonical forms. The matching is established \emph{exactly}, at fixed
sufficiently large length, with no per-sector unit-modulus hypothesis: the
only normalization used is the global one of
\cite[\S~II.C, line~246]{Cirac2016MPDO_arXiv}, which is part of the BNT
canonical-form predicate.

The mechanism is the one of
\cite[Appendix MPV proof, line~1182]{Cirac2016MPDO_arXiv}, read as an exact
statement rather than an asymptotic projection. Fix a sector $k$ of $Q$ and
suppose, for contradiction, that the cross-overlap of $Q_k$ with every basis
sector $P_j$ tends to zero. Let $T$ be the set of $P$-sectors that decay
against \emph{all} sectors of $Q$. For every sector $j \notin T$ the overlap
dichotomy turns a non-decaying overlap with some $Q_{k(j)}$ into a gauge-phase
equivalence, hence a scalar $\omega_j \in \C^\times$ with
\[
    V^{(N)}(P_j)_\sigma = \omega_j^{N}\, V^{(N)}(Q_{k(j)})_\sigma
    \qquad \text{for every } N \text{ and } \sigma,
\]
so $V^{(N)}(P_j)$ lies in the span of $\{V^{(N)}(Q_k)\}_k$. The family
$\{P_j : j\in T\}\cup\{Q_k\}_k$ then has all pairwise cross-overlaps decaying,
so it is eventually linearly independent
(Lemma~\ref{lem:sector_bnt_combined_family_eventually_li}). Rewriting the
common-MPV identity over this family expresses the relation entirely in an
eventually linearly independent family, so at each fixed large length every
coefficient vanishes \emph{exactly}; in particular the sector coefficient of
each $j\in T$ is eventually zero. This contradicts
Lemma~\ref{lem:sector_bnt_coeff_not_eventually_zero}, which holds for any
nonzero-weight sector with no modulus assumption. Hence the supposed
all-decaying sector cannot occur and the match exists.

Because the contradiction is between exact vanishing at fixed length and exact
non-vanishing of a finite sum of nonzero-weight geometric sequences, it never
passes through a limit of a single sector coefficient; a sub-unit sector, whose
coefficient does decay, is matched by the same argument. The bijective-matching
theorem below applies the strong existential theorem twice (with $(P, Q)$
swapped) to derive the cardinality identity and the per-pair bijection and
gauges on the full basis.

\begin{lemma}[Sector coefficient is not eventually zero]
    \label{lem:sector_bnt_coeff_not_eventually_zero}
    \lean{MPSTensor.IsBNTCanonicalForm.coeff_not_eventually_zero}
    \leanok
    \uses{def:sector_bnt_cf}
    For a BNT canonical form $P$ and any sector $j$, the sector coefficient
    $c_{P,j}^{(N)} = \sum_q \mu_{j,q}^N$ is not eventually zero in $N$. No
    modulus assumption on the copy weights is used; only that they are nonzero.
    The signed-geometric extrapolation shows that if the power sum vanished for
    all large $N$ it would vanish at every exponent, including $N=0$,
    contradicting the positivity of the copy count $r_j = P.\mathrm{copies}(j)$
    (repeated ratios are handled by the same extrapolation, so distinctness of
    the weights is not needed).
\end{lemma}

\begin{proof}\leanok
    If $c_{P,j}^{(N)} = 0$ for all $N \ge M$, the signed geometric-sequence
    vanishing lemma forces $\sum_q \mu_{j,q}^k = 0$ for every exponent $k$, in
    particular $\sum_q 1 = r_j = 0$ at $k=0$, contradicting $r_j > 0$.
\end{proof}

\begin{theorem}[Exact single-sector matching]
    \label{thm:sector_bnt_exact_block_match}
    \lean{MPSTensor.exists_block_match_exact}
    \leanok
    \uses{def:sector_bnt_cf,
        lem:sector_bnt_combined_family_eventually_li,
        lem:sector_bnt_coeff_not_eventually_zero}
    Let $P, Q$ be two BNT canonical forms with the same MPV family (at all
    positive lengths). For every sector $j$ of $P$ there exist a sector $k$ of
    $Q$ with matched bond dimension $\dim_P(j) = \dim_Q(k)$, a gauge-phase
    equivalence after identifying the matched bond dimensions, and a
    non-decaying cross-overlap. No per-sector unit-modulus hypothesis is
    required.
\end{theorem}

\begin{proof}\leanok
    \uses{lem:sector_bnt_combined_family_eventually_li,
        lem:sector_bnt_coeff_not_eventually_zero}
    This is the exact $T$-partition argument described above. Suppose $j$
    decays against every sector of $Q$, and let $T$ be the set of $P$-sectors
    with this property. Sectors outside $T$ contribute MPV vectors lying in the
    span of the $Q$-family, so the common-MPV identity restricts to the
    eventually linearly independent family
    $\{P_{j'} : j'\in T\}\cup\{Q_k\}_k$
    (Lemma~\ref{lem:sector_bnt_combined_family_eventually_li}). Exact
    coefficient extraction at each fixed large length forces the sector
    coefficient of $j$ to be eventually zero, contradicting
    Lemma~\ref{lem:sector_bnt_coeff_not_eventually_zero}. Hence a matched
    sector $k$ exists; the gauge-phase equivalence is recovered from the
    non-decaying overlap by the irreducible trace-preserving overlap
    dichotomy.
\end{proof}

\begin{theorem}[Strong existential matching, full-basis form]
    \label{thm:sector_bnt_forall_unit_k_exists_j_nondecaying_overlap_of_sameMPV}
    \lean{MPSTensor.forall_k_exists_j_nondecaying_overlap_of_sameMPV}
    \leanok
    \uses{def:sector_bnt_cf,
        thm:sector_bnt_exact_block_match}
    Let $P, Q$ be two BNT canonical forms with the same MPV family. Then for
    every sector $k$ of $Q$ there exist a sector $j$ of $P$ with matched bond
    dimension $\dim_P(j) = \dim_Q(k)$, a gauge-phase equivalence after
    identifying the matched bond dimensions, and a non-decaying cross-overlap.
    No per-sector unit-modulus hypothesis is required.
\end{theorem}

\begin{proof}\leanok
    \uses{thm:sector_bnt_exact_block_match}
    Apply Theorem~\ref{thm:sector_bnt_exact_block_match} with the roles of $P$
    and $Q$ swapped and the same-MPV hypothesis flipped by symmetry. The
    resulting conclusion is re-oriented by symmetry of gauge-phase equivalence
    under the bond-dimension identification and symmetry of overlap
    convergence.
\end{proof}

\begin{remark}[Linear independence carries the matching]
    \label{rem:sector_bnt_strong_match_no_partial_li}
    Theorem~\ref{thm:sector_bnt_exact_block_match} routes through the
    linear-independence step
    (Lemma~\ref{lem:sector_bnt_combined_family_eventually_li}), the corollary
    input of \cite[Appendix MPV corollary, lines~1131--1133]
    {Cirac2016MPDO_arXiv}, applied to the family
    $\{P_j : j\in T\}\sqcup\{Q_k\}_k$ on which the residual relation lives. The
    contradiction is exact vanishing of a coefficient at fixed length versus
    Lemma~\ref{lem:sector_bnt_coeff_not_eventually_zero}; no asymptotic
    projection of a single sector coefficient is taken, so no per-sector
    unit-modulus hypothesis enters.
\end{remark}

\subsection{Bijective matching by symmetry}
\label{subsec:sector_bnt_strong_match_bijective_symmetry}

This subsection derives the symmetry argument of
\cite[Appendix MPV proof, line~1182]{Cirac2016MPDO_arXiv}:
$g_a \ge g_b$ and $g_b \ge g_a$ together imply $g_a = g_b$.

The strong existential matching theorem
(Theorem~\ref{thm:sector_bnt_forall_unit_k_exists_j_nondecaying_overlap_of_sameMPV})
gives one direction of the full-basis matching. Applying the same theorem once
more with the roles of $P$ and $Q$ swapped, using the symmetric flip of the
same-MPV hypothesis, gives the other direction. The two matchings, combined
with the basis-distinctness clause of the canonical-form predicate
\cite[\S~II.C, lines~264--279]{Cirac2016MPDO_arXiv}, force injectivity
in both directions.

\paragraph{Scope.}
The matching delivers the literal source conclusion ``$g_a = g_b$ and a
relabelling $B_k = e^{i\phi_k} X_k A_{j_k} X_k^{-1}$''
(\cite[Appendix MPV proof, line~1182]{Cirac2016MPDO_arXiv}) on the full BNT
basis, using only the global normalization of
\cite[\S~II.C, line~246]{Cirac2016MPDO_arXiv} carried by the canonical-form
predicate. No per-sector unit-modulus hypothesis is imposed: the exact
matching (Theorem~\ref{thm:sector_bnt_exact_block_match}) handles sub-unit
sectors on the same footing. The coefficient comparison is not folded into this
matching step; it is recorded separately in the following subsection.

\begin{theorem}[Bijective matching by symmetry]
    \label{thm:sector_bnt_bijective_match_of_sameMPV}
    \lean{MPSTensor.bijective_match_of_sameMPV}
    \leanok
    \uses{thm:sector_bnt_exact_block_match,
        def:sector_bnt_cf}
    Let $P, Q$ be two BNT canonical forms with the same MPV family. Then there
    is a bijection
    \[
        \beta : J(Q) \simeq J(P)
    \]
    such that for every $k \in J(Q)$ the bond dimensions match, the basis
    tensors $P_{\beta(k)}$ and $Q_k$ are gauge-phase equivalent after
    identifying the matched bond dimensions, and the basis cross-overlap does
    not tend to zero.
\end{theorem}

\begin{proof}\leanok
    \uses{thm:sector_bnt_exact_block_match}
    Apply the exact single-sector matching
    (Theorem~\ref{thm:sector_bnt_exact_block_match}) at each sector of $P$, with
    $(P, Q)$ and the original hypothesis, to obtain a raw matched-sector
    function via classical choice; apply it again at each sector of $Q$, with
    $(Q, P)$ and the symmetric flipped hypothesis, to obtain the symmetric raw
    function.

    \emph{Forward injectivity.} Suppose $\varphi_0(k_1) = \varphi_0(k_2) = j$.
    The two matching witnesses give a common centre $P_j$ with two
    gauge-phase equivalences, after identifying the bond dimensions, to
    $Q_{k_1}$ and $Q_{k_2}$. Composing through this centre (transitivity of
    gauge-phase equivalence under these bond-dimension identifications) yields
    a gauge-phase equivalence between $Q_{k_1}$ and $Q_{k_2}$ under the
    dimension equality
    $\dim_Q(k_1) = \dim_Q(k_2)$. The basis-distinctness clause of the
    canonical-form predicate on $Q$
    (\cite[\S~II.C, lines~264--279]{Cirac2016MPDO_arXiv}) then forces
    $k_1 = k_2$. Backward injectivity is symmetric.

    \emph{Bijection.} The two injections give equality of the two finite
    sector counts. Hence the forward injection is a bijection, and the
    matching witnesses are extracted by re-applying the classical-choice
    specification at each index in the domain.
\end{proof}

\subsection{Full-basis coefficient identity and global gauge}
\label{subsec:sector_bnt_full_basis_global_gauge}

The matching theorem above gives the sector phases and gauges. The remaining
coefficient comparison is the step at
\cite[Appendix MPV proof, lines~1187--1188]{Cirac2016MPDO_arXiv}. In the
equal-MPV setting, the full BNT basis and eventual linear independence give the
comparison directly, with no per-sector unit-modulus hypothesis. In the
proportional setting below, the same comparison is recorded as the explicit
hypothesis needed before the copy weights can be assembled into the global
gauge; the sector matching itself needs no per-sector unit-modulus hypothesis
either.

\begin{lemma}[Unit phase from matched normalized BNT blocks]%
    \label{lem:sector_bnt_norm_phase_of_matched_mpv}
    \lean{MPSTensor.IsBNTCanonicalForm.norm_phase_of_matched_mpv}
    \leanok
    \uses{def:sector_bnt_cf}
    Let $P$ and $Q$ be BNT canonical forms. If, for every length $N$ and word
    $\sigma$, a sector $P_j$ and a sector $Q_k$ have MPV families related by
    \[
        V^{(N)}(Q_k)_\sigma=\zeta^N V^{(N)}(P_j)_\sigma
    \]
    then $|\zeta|=1$.
\end{lemma}

\begin{proof}\leanok
    The BNT normalization gives self-overlap limits of modulus $1$ for both
    sectors. The displayed MPV identity scales the $Q_k$ self-overlap by
    $(\zeta\overline{\zeta})^N$ times the $P_j$ self-overlap. Taking limits
    forces $|\zeta|=1$.
\end{proof}

\begin{theorem}[Proportional sector matching with unit phases]%
    \label{thm:sector_bnt_proportional_sector_match_witnesses}
    \lean{MPSTensor.ft_sector_bnt_proportional_sector_match_witnesses}
    \leanok
    \uses{def:sector_bnt_cf,
        def:gauge_phase_equiv,
        thm:gauge_phase_mpv_scaling,
        lem:sector_bnt_norm_phase_of_matched_mpv}
    Let $P$ and $Q$ be BNT canonical forms whose total tensors generate
    eventually nonzero proportional MPV families.
    Then there are a bijection $\beta:J(Q)\simeq J(P)$, bond-dimension
    equalities $\dim P_{\beta(k)}=\dim Q_k$, scalars
    $\zeta_k\in\C$, and block gauges $X_k$ such that $|\zeta_k|=1$ and, for
    every sector $k$ and physical index $i$,
    \[
        Q_k^i=\zeta_k X_k P_{\beta(k)}^i X_k^{-1}
    \]
    Equivalently, for every length $N$ and word $\sigma$,
    \[
        V^{(N)}(Q_k)_\sigma
        =
        \zeta_k^N V^{(N)}(P_{\beta(k)})_\sigma
    \]
    This records the per-block phase conclusion of
    \cite[Appendix MPV theorem statement, lines~1167--1170]
    {Cirac2016MPDO_arXiv}; the proof step at
    \cite[Appendix MPV proof, line~1182]{Cirac2016MPDO_arXiv} is the
    corresponding argument. The following coefficient-comparison step is
    separate.
    % Scope restriction documented in
    % docs/paper-gaps/cpsv16_global_vs_persector_unit_witness.tex.
\end{theorem}

\begin{proof}\leanok\uses{lem:sector_bnt_norm_phase_of_matched_mpv}
    Apply the proportional sector-matching theorem to obtain the bijection and
    dimension-compatible gauge-phase equivalences. Choosing the gauges and
    scalars gives $Q_k^i=\zeta_kX_kP_{\beta(k)}^iX_k^{-1}$, and therefore
    \[
        O_{Q_kQ_k}(N)
        =
        |\zeta_k|^{2N}O_{P_{\beta(k)}P_{\beta(k)}}(N).
    \]
    The normalized self-overlap limits give
    $O_{Q_kQ_k}(N)\to1$ and $O_{P_{\beta(k)}P_{\beta(k)}}(N)\to1$, so
    $|\zeta_k|^{2N}\to1$ and hence $|\zeta_k|=1$. The MPV scalar-power
    identity follows from the same gauge equation:
    $V^{(N)}(Q_k)_\sigma=\zeta_k^NV^{(N)}(P_{\beta(k)})_\sigma$.
\end{proof}

\begin{lemma}[Full-basis coefficient identity from matched phases]%
    \label{lem:sector_bnt_coeff_identity_via_matched_mpv_phase}
    \lean{MPSTensor.coeff_identity_via_matched_mpv_phase}
    \leanok
    \uses{def:sector_bnt_cf}
    Let $P$ be a BNT canonical form, let $Q$ be a sector decomposition, and
    assume $P_{\mathrm{tot}}$ and $Q_{\mathrm{tot}}$ have the same MPV family.
    Suppose that there is a bijection
    \[
        \beta : J(Q) \simeq J(P)
    \]
    and scalars $\zeta_k \in \C$ such that, for every $k \in J(Q)$, every
    length $N$, and every spin configuration $\sigma$ of length $N$,
    \[
        V^{(N)}(Q_k)_\sigma
        =
        \zeta_k^N\,V^{(N)}(P_{\beta(k)})_\sigma .
    \]
    Then for every $k \in J(Q)$ there is $N_0$ such that, for all $N>N_0$,
    \[
        c_N^{(\beta(k))}(P)
        =
        \zeta_k^N\,c_N^{(k)}(Q).
    \]
    This is the coefficient-comparison step of
    \cite[Appendix MPV proof, lines~1187--1188]{Cirac2016MPDO_arXiv} in full-basis
    form.
\end{lemma}

\begin{proof}\leanok
    \uses{lem:eventual_coeff_eq_of_eventual_li}
    For every $N$, expand the equality of total MPVs as
    \[
        \sum_{j\in J(P)} c_N^{(j)}(P)\ket{V^{(N)}(P_j)}
        =
        \sum_{k\in J(Q)} c_N^{(k)}(Q)\ket{V^{(N)}(Q_k)} .
    \]
    For every $N$, the phase relation gives
    $\ket{V^{(N)}(Q_k)}=\zeta_k^N\ket{V^{(N)}(P_{\beta(k)})}$, hence
    \[
        \sum_{j\in J(P)} c_N^{(j)}(P)\ket{V^{(N)}(P_j)}
        =
        \sum_{j\in J(P)}
        \zeta_{\beta^{-1}(j)}^N
        c_N^{(\beta^{-1}(j))}(Q)\ket{V^{(N)}(P_j)} .
    \]
    Lemma~\ref{lem:eventual_coeff_eq_of_eventual_li} applied to the
    $P$-basis gives
    $c_N^{(\beta(k))}(P)=\zeta_k^Nc_N^{(k)}(Q)$ for all sufficiently
    large $N$.
\end{proof}

\begin{lemma}[Full-basis coefficient identity from global gauges]%
    \label{lem:sector_bnt_coeff_identity_via_global_gauge}
    \lean{MPSTensor.coeff_identity_via_global_gauge}
    \leanok
    \uses{def:sector_bnt_cf,
        def:gauge_phase_equiv}
    Let $P$ and $Q$ be BNT canonical forms with the same MPV family. Suppose
    that there is a bijection
    \[
        \beta : J(Q) \simeq J(P)
    \]
    and, for every $k\in J(Q)$, a bond-dimension equality and a gauge-phase
    equivalence between $Q_k$ and $P_{\beta(k)}$. Then for every $k$ there are
    a unit-modulus scalar $\zeta_k$ and a natural number $N_0$ such that, for
    all $N>N_0$,
    \[
        c_N^{(\beta(k))}(P)=\zeta_k^N\,c_N^{(k)}(Q).
    \]
    This is the full-basis equal-MPV coefficient comparison of
    \cite[Appendix MPV proof, lines~1187--1188]{Cirac2016MPDO_arXiv}.
\end{lemma}

\begin{proof}\leanok
    \uses{lem:sector_bnt_norm_phase_of_matched_mpv,
        lem:sector_bnt_coeff_identity_via_matched_mpv_phase}
    For each matched sector, the gauge-phase equivalence gives
    $V^{(N)}(Q_k)_\sigma=\zeta_k^NV^{(N)}(P_{\beta(k)})_\sigma$.
    Lemma~\ref{lem:sector_bnt_norm_phase_of_matched_mpv} gives
    $|\zeta_k|=1$. Applying
    Lemma~\ref{lem:sector_bnt_coeff_identity_via_matched_mpv_phase} to the
    same phases gives
    $c_N^{(\beta(k))}(P)=\zeta_k^Nc_N^{(k)}(Q)$ for all sufficiently
    large $N$.
\end{proof}

\begin{definition}[Copy-weight matching for matched BNT sectors]%
    \label{def:sector_bnt_copy_weight_matching}
    \lean{MPSTensor.SectorBNTCopyWeightMatching}
    \leanok
    Let the sectors of two decompositions $P$ and $Q$ be matched by
    $\beta:J(Q)\simeq J(P)$, and let $\zeta_k$ be the phase attached to
    sector $k$. A copy-weight matching consists of copy permutations
    $\tau_k$ and the following identities: for all sectors $k$ and copies $q$,
    \[
        \nu_{k,q}=\zeta_k^{-1}\,\mu_{\beta(k),\tau_k(q)}.
    \]
    This is the copy-level conclusion supplied by the equal-MPV coefficient comparison in
    \cite[Appendix MPV proof, line~1188]{Cirac2016MPDO_arXiv}. For the
    proportional theorem it remains the separate coefficient-comparison input.
\end{definition}

\begin{lemma}[Copy-weight matching from matched-sector coefficient identities]%
    \label{lem:sector_bnt_copy_weight_matching_of_coeff_identity}
    \lean{MPSTensor.SectorBNTCopyWeightMatching.of_coeff_identity}
    \leanok
    \uses{def:sector_bnt_copy_weight_matching,
        lem:sector_bnt_matched_sector_weight_equiv}
    Let the sectors of two decompositions $P$ and $Q$ be matched by
    $\beta:J(Q)\simeq J(P)$, and let $\zeta_k\neq 0$ be the phase attached to
    sector $k$. Suppose that for every $k$ there is $N_0$ such that, for all
    $N>N_0$,
    \[
        c_N^{(\beta(k))}(P)=\zeta_k^N\,c_N^{(k)}(Q).
    \]
    Then the phases $\zeta_k$ determine a copy-weight matching:
    after a copy permutation inside each matched sector,
    \[
        \nu_{k,q}=\zeta_k^{-1}\,\mu_{\beta(k),\tau_k(q)}.
    \]
    This is, sector by sector, the content of the coefficient comparison in
    \cite[Appendix MPV proof, line~1188]{Cirac2016MPDO_arXiv}.
\end{lemma}

\begin{proof}\leanok
    For all sufficiently large $N$, the coefficient identity in sector $k$ is
    \[
        \sum_r \mu_{\beta(k),r}^{\,N}
        =
        \zeta_k^N\sum_q\nu_{k,q}^{\,N}
        =
        \sum_q(\zeta_k\nu_{k,q})^{N}.
    \]
    Lemma~\ref{lem:sector_bnt_matched_sector_weight_equiv} applies the finite
    power-sum comparison sector by sector, producing a copy permutation
    $\tau_k$ with
    $\mu_{\beta(k),\tau_k(q)}=\zeta_k\nu_{k,q}$. Since $\zeta_k\ne0$, this is
    equivalent to $\nu_{k,q}=\zeta_k^{-1}\mu_{\beta(k),\tau_k(q)}$.
\end{proof}

\begin{theorem}[Global gauge from matched sectors and copy weights]%
    \label{thm:sector_bnt_global_gauge_of_matched_weights}
    \lean{MPSTensor.sector_bnt_global_gauge_of_matched_weights}
    \leanok
    \uses{lem:tensor_from_blocks_globalGauge_conj}
    Suppose that the sectors of two sector decompositions $P$ and $Q$ are matched
    by a bijection $\beta:J(Q)\simeq J(P)$, with bond-dimension equalities,
    copy permutations $\tau_k$, nonzero phases $\zeta_k$, and block gauges
    $X_k$. If, for every sector $k$, copy $q$, and physical index $i$,
    \[
        Q_k^i
        =
        \zeta_k\,X_k P_{\beta(k)}^i X_k^{-1},
        \qquad
        \nu_{k,q}
        =
        \zeta_k^{-1}\,\mu_{\beta(k),\tau_k(q)},
    \]
    then, in the flattened $Q$-coordinates,
    \[
        Q_{\mathrm{flat}}^i
        =
        X\,P_{\mathrm{matched}}^i\,X^{-1},
        \qquad
        X=\bigoplus_k(\mathbf{1}_{r_k}\otimes X_k).
    \]
    This is the direct-sum gauge construction of
    \cite[Appendix MPV proof, lines~1189--1192]{Cirac2016MPDO_arXiv}, separated from
    the coefficient-comparison step that supplies the weight identities.
\end{theorem}

\begin{proof}\leanok
    For each matched copy,
    \[
        \nu_{k,q}Q_k^i
        =
        \bigl(\zeta_k^{-1}\mu_{\beta(k),\tau_k(q)}\bigr)
        \bigl(\zeta_kX_kP_{\beta(k)}^iX_k^{-1}\bigr)
        =
        \mu_{\beta(k),\tau_k(q)}X_kP_{\beta(k)}^iX_k^{-1}.
    \]
    Thus every flattened weighted block is conjugated by its sector gauge
    $X_k$. Lemma~\ref{lem:tensor_from_blocks_globalGauge_conj} assembles these
    block conjugacies into the direct-sum gauge
    $X=\bigoplus_k(\mathbf{1}_{r_k}\otimes X_k)$.
\end{proof}

\begin{theorem}[Global gauge from a copy-weight matching]%
    \label{thm:sector_bnt_global_gauge_of_copy_weight_matching}
    \lean{MPSTensor.sector_bnt_global_gauge_of_copy_weight_matching}
    \leanok
    \uses{def:sector_bnt_copy_weight_matching,
        thm:sector_bnt_global_gauge_of_matched_weights}
    Suppose that the sectors of $P$ and $Q$ are matched, with
    bond-dimension equalities, nonzero phases, and block gauges satisfying
    \[
        Q_k^i=\zeta_k X_kP_{\beta(k)}^iX_k^{-1}.
    \]
    If the same phases also give a copy-weight matching, then the flattened
    $Q$-tensor is obtained from the matched-coordinate $P$-tensor by the
    direct-sum gauge
    \[
        X=\bigoplus_k(\mathbf{1}_{r_k}\otimes X_k).
    \]
    This is only a reformulation of
    Theorem~\ref{thm:sector_bnt_global_gauge_of_matched_weights}.
\end{theorem}

\begin{proof}\leanok
    The copy-weight matching supplies
    $\nu_{k,q}=\zeta_k^{-1}\mu_{\beta(k),\tau_k(q)}$. Together with
    $Q_k^i=\zeta_kX_kP_{\beta(k)}^iX_k^{-1}$, these are exactly the two
    displayed hypotheses of
    Theorem~\ref{thm:sector_bnt_global_gauge_of_matched_weights}.
\end{proof}

\begin{theorem}[Conditional proportional global gauge from a copy-weight matching]%
    \label{thm:sector_bnt_proportional_global_gauge_of_copy_weight_matching}
    \lean{MPSTensor.ft_sector_bnt_proportional_global_gauge_of_copy_weight_matching}
    \leanok
    \uses{def:sector_bnt_copy_weight_matching,
        thm:sector_bnt_proportional_sector_match_witnesses,
        thm:sector_bnt_global_gauge_of_copy_weight_matching}
    Let $P$ and $Q$ be BNT canonical forms whose total tensors generate
    eventually nonzero proportional MPV families.
    Then the proportional sector-matching theorem gives $\beta$, the
    bond-dimension equalities, the unit phases $\zeta_k$, and the block gauges
    $X_k$. If the remaining proportional coefficient comparison supplies a
    copy-weight matching for these same phases, then the flattened
    $Q$-tensor is obtained from the matched-coordinate $P$-tensor by the
    direct-sum gauge
    \[
        X=\bigoplus_k(\mathbf{1}_{r_k}\otimes X_k).
    \]
    % Scope restriction documented in
    % docs/paper-gaps/cpsv16_global_vs_persector_unit_witness.tex.
\end{theorem}

\begin{proof}\leanok
    The proportional sector-matching theorem supplies $\beta$, $X_k$, and
    unit phases $\zeta_k$ satisfying
    $Q_k^i=\zeta_kX_kP_{\beta(k)}^iX_k^{-1}$. Since $|\zeta_k|=1$, each
    $\zeta_k$ is nonzero. A copy-weight matching gives
    $\nu_{k,q}=\zeta_k^{-1}\mu_{\beta(k),\tau_k(q)}$, so
    Theorem~\ref{thm:sector_bnt_global_gauge_of_copy_weight_matching} gives the
    direct-sum gauge.
\end{proof}

\begin{theorem}[Conditional proportional global gauge from coefficient identities]%
    \label{thm:sector_bnt_proportional_global_gauge_of_coeff_identity}
    \lean{MPSTensor.ft_sector_bnt_proportional_global_gauge_of_coeff_identity}
    \leanok
    \uses{lem:sector_bnt_copy_weight_matching_of_coeff_identity,
        thm:sector_bnt_proportional_global_gauge_of_copy_weight_matching}
    Let $P$ and $Q$ be BNT canonical forms whose total tensors generate
    eventually nonzero proportional MPV families.
    Then there are a bijection $\beta:J(Q)\simeq J(P)$, bond-dimension
    equalities, unit-modulus phases $\zeta_k$, and block gauges $X_k$ such that
    \[
        Q_k^i=\zeta_k X_kP_{\beta(k)}^iX_k^{-1}
    \]
    for every sector $k$ and physical index $i$. If these same phases satisfy
    the matched-sector coefficient identities
    \[
        c_N^{(\beta(k))}(P)=\zeta_k^N\,c_N^{(k)}(Q)
    \]
    for all sufficiently large $N$, then they determine a copy-weight matching
    and hence the flattened $Q$-tensor is obtained from the matched-coordinate
    $P$-tensor by the direct-sum gauge
    \[
        X=\bigoplus_k(\mathbf{1}_{r_k}\otimes X_k).
    \]
    % Scope restriction documented in
    % docs/paper-gaps/cpsv16_global_vs_persector_unit_witness.tex.
\end{theorem}

\begin{proof}\leanok
    The sector-matching theorem gives $\beta$, $X_k$, and $|\zeta_k|=1$ with
    $Q_k^i=\zeta_kX_kP_{\beta(k)}^iX_k^{-1}$. Since $\zeta_k\ne0$, the
    coefficient identities
    $c_N^{(\beta(k))}(P)=\zeta_k^Nc_N^{(k)}(Q)$ produce
    $\nu_{k,q}=\zeta_k^{-1}\mu_{\beta(k),\tau_k(q)}$ by
    Lemma~\ref{lem:sector_bnt_copy_weight_matching_of_coeff_identity}.
    Theorem~\ref{thm:sector_bnt_proportional_global_gauge_of_copy_weight_matching}
    then applies to this copy-weight matching.
\end{proof}

\begin{lemma}[Equal-MPV matched copy-weight witnesses]%
    \label{lem:sector_bnt_equal_matched_copy_weight_witnesses}
    \lean{MPSTensor.ft_sector_bnt_equal_matched_copy_weight_witnesses}
    \leanok
    \uses{def:sector_bnt_cf}
    Let $P$ and $Q$ be BNT canonical forms whose total tensors have the same
    MPV family. Then there are a bijection
    $\beta:J(Q)\simeq J(P)$, bond-dimension equalities, unit-modulus phases
    $\zeta_k$, and block gauges $X_k$ such that
    \[
        Q_k^i=\zeta_k X_kP_{\beta(k)}^iX_k^{-1}.
    \]
    For the same phases there is a copy-weight matching: after a permutation
    of the copies inside each matched sector,
    \[
        \nu_{k,q}=\zeta_k^{-1}\mu_{\beta(k),\tau_k(q)}.
    \]
\end{lemma}

\begin{proof}\leanok
    \uses{thm:sector_bnt_bijective_match_of_sameMPV,
        lem:sector_bnt_coeff_identity_via_matched_mpv_phase,
        lem:sector_bnt_copy_weight_matching_of_coeff_identity}
    The full-basis matching theorem gives the bijection and block gauges. The
    self-overlap comparison
    \[
        \langle V_N(Q_k),V_N(Q_k)\rangle
        =
        |\zeta_k|^{2N}
        \langle V_N(P_{\beta(k)}),V_N(P_{\beta(k)})\rangle
    \]
    together with the BNT normalization limits on both sides forces
    $|\zeta_k|=1$. The matched MPV phase identities also give the eventual
    coefficient identities
    \[
        c_N^{(\beta(k))}(P)=\zeta_k^N c_N^{(k)}(Q).
    \]
    Lemma~\ref{lem:sector_bnt_copy_weight_matching_of_coeff_identity} applies
    the finite power-sum comparison in each sector and gives the copy-weight
    identities.
\end{proof}

\begin{theorem}[Full-basis global gauge in matched coordinates]%
    \label{thm:sector_bnt_equal_global_gauge}
    \lean{MPSTensor.ft_sector_bnt_equal_global_gauge}
    \leanok
    \uses{def:sector_bnt_cf}
    Let $P$ and $Q$ be BNT canonical forms whose total tensors have the same
    MPV family. Then there exist:
    \begin{itemize}
        \item a bijection $\beta:J(Q)\simeq J(P)$,
        \item bond-dimension equalities
              $\dim P_{\beta(k)}=\dim Q_k$,
        \item copy-number equalities $r_{\beta(k)}(P)=r_k(Q)$ and
              copy permutations $\tau_k$,
        \item unit-modulus phases $\zeta_k$ with $|\zeta_k|=1$ and block
              gauges $X_k$,
    \end{itemize}
    such that, for every sector $k$, copy $q$, and physical index $i$,
    \[
        Q_k^i
        =
        \zeta_k\,X_k P_{\beta(k)}^i X_k^{-1},
        \qquad
        \nu_{k,q}
        =
        \zeta_k^{-1}\,\mu_{\beta(k),\tau_k(q)}.
    \]
    Moreover, if
    \[
        X=\bigoplus_k (\mathbf{1}_{r_k}\otimes X_k)
    \]
    is the flattened block-diagonal gauge, then in the flattened
    $Q$-coordinates,
    \[
        Q_{\mathrm{flat}}^i
        =
        X\,P_{\mathrm{matched}}^i\,X^{-1}
    \]
    for every physical index $i$. This is the explicit direct-sum gauge
    construction of
    \cite[Appendix MPV proof, lines~1189--1192]{Cirac2016MPDO_arXiv}.
    % Scope restriction documented in
    % docs/paper-gaps/cpsv16_global_vs_persector_unit_witness.tex.
\end{theorem}

\begin{proof}\leanok
    \uses{lem:sector_bnt_equal_matched_copy_weight_witnesses,
        thm:sector_bnt_global_gauge_of_copy_weight_matching}
    Lemma~\ref{lem:sector_bnt_equal_matched_copy_weight_witnesses} gives
    $\beta$, the bond-dimension identifications, the phases $\zeta_k$, the
    block gauges $X_k$, and copy permutations $\tau_k$ satisfying
    \[
        \nu_{k,q}=\zeta_k^{-1}\mu_{\beta(k),\tau_k(q)} .
    \]
    Theorem~\ref{thm:sector_bnt_global_gauge_of_copy_weight_matching} assembles
    the block conjugacies into the flattened direct-sum conjugation by
    $X=\bigoplus_k(\mathbf{1}_{r_k}\otimes X_k)$.
\end{proof}

\begin{theorem}[Literal equal-MPV gauge after sector-coordinate permutation]%
    \label{thm:sector_bnt_equal_mps_gaugeEquiv_literal}
    \lean{MPSTensor.ft_sector_bnt_equal_mps_gaugeEquiv_literal}
    \leanok
    \uses{def:sector_bnt_cf}
    Let $P$ and $Q$ be BNT canonical forms whose total tensors have the same
    MPV family. Then the total bond dimensions agree;
    write their common value as $D$. There is
    \[
        Y\in \GL_D(\C)
    \]
    such that, for every physical index $i$,
    \[
        Q_{\mathrm{tot}}^i
        =
        Y\,P_{\mathrm{tot}}^i\,Y^{-1},
    \]
    where $P_{\mathrm{tot}}^i$ and $Q_{\mathrm{tot}}^i$ are both regarded as
    matrices in $\MN{D}$ using the total bond-dimension equality. This is the
    BNT form of
    \cite[Corollary~II.2 and Appendix MPV proof, lines~1189--1192]
    {Cirac2016MPDO_arXiv}.
\end{theorem}

\begin{proof}\leanok
    \uses{thm:sector_bnt_equal_global_gauge}
    Apply Theorem~\ref{thm:sector_bnt_equal_global_gauge} to obtain the
    flattened gauge after the sectors of $P$ have been reindexed by the
    bijection between basis blocks of $P$ and $Q$. The sector-coordinate
    permutation $\rho$, induced by this bijection, the copy permutations, and
    the bond-dimension equalities, identifies the reindexed $P$-side direct sum
    with the total tensor $P_{\mathrm{tot}}$, transported along the total
    bond-dimension equality. Composing the permutation gauge for $\rho$ with
    the sector-permuted gauge gives the displayed single matrix $Y$.
\end{proof}
