\chapter{BNT and Full Assembly}\label{ch:assembly}

This chapter introduces \emph{Basis Normal Tensors} (BNT),
establishes their permutation rigidity, and assembles the
full Fundamental Theorem of Matrix Product States.
The results combine the canonical form reduction
(Chapter~\ref{ch:canonical}),
block separation (Chapter~\ref{ch:block_perm}),
spectral gap (Chapter~\ref{ch:spectral}),
and Newton--Girard identities into the final statement.
The main references are
\cite{PerezGarcia2007Matrix} and~\cite{Cirac2021Matrix}.

\section{Basis normal tensors}

\begin{definition}[Basis Normal Tensor (BNT)]\label{def:bnt}
    \lean{MPSTensor.IsBNT}
    \leanok
    \uses{def:normal, def:mpv}
    A \emph{basis normal tensor} (BNT) decomposition for a
    total tensor $A_{\mathrm{tot}}$ consists of $g$ block tensors
    $(A_1, \ldots, A_g)$ with bond dimensions $(D_1, \ldots, D_g)$
    such that:
    \begin{enumerate}
        \item each $A_j$ is normal
              (Definition~\ref{def:normal}),
        \item at each system size~$N$, the MPV of $A_{\mathrm{tot}}$
              is a linear combination of the block MPVs:
              $\ket{V^{(N)}(A_{\mathrm{tot}})}
              = \sum_j c_j \ket{V^{(N)}(A_j)}$,
        \item for sufficiently large~$N$, the block MPV
              vectors $\{\ket{V^{(N)}(A_j)}\}_{j=1}^g$ are
              linearly independent.
    \end{enumerate}
    This is \cite[Definition~4.2]{Cirac2021Matrix}.
\end{definition}

\begin{definition}[Canonical form with BNT separation]\label{def:cf_bnt}
    \lean{MPSTensor.IsCanonicalFormBNT}
    \leanok
    \uses{def:is_canonical_form, def:gauge_phase_equiv}
    An \emph{IsCanonicalFormBNT} extends the canonical form
    predicate (Definition~\ref{def:is_canonical_form})
    with the additional requirement that distinct blocks are
    \emph{not} gauge-phase equivalent:
    for $j \neq k$ with $D_j = D_k$,
    $A_j$ and $A_k$ are not gauge-phase equivalent.
\end{definition}

\begin{theorem}[CF-BNT yields BNT]\label{thm:cf_bnt_is_bnt}
    \lean{MPSTensor.IsCanonicalFormBNT.isBNT}
    \leanok
    \uses{def:cf_bnt, def:bnt}
    If $((\mu_k), (A_k))$ satisfies the CF-BNT predicate, then
    the block tensors form a BNT decomposition for the
    block-diagonal tensor $\bigoplus_k \mu_k A_k$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:mpv_decomposition}
    \textbf{Normality}: each block is injective
    (hence $1$-block injective, hence normal).
    \textbf{Spanning}: the MPV decomposition
    (Theorem~\ref{thm:mpv_decomposition}) gives
    $V^{(N)}(A_{\mathrm{tot}})_\sigma = \sum_k \mu_k^N V^{(N)}(A_k)_\sigma$.
    \textbf{Linear independence}: the overlap matrix
    $G_{jk}(N) = O_{A_j A_k}(N)$ converges to the identity
    (diagonal entries $\to 1$ by the self-overlap hypothesis;
    off-diagonal entries $\to 0$ by the spectral gap,
    since distinct CF-BNT blocks are not gauge-phase equivalent).
    A Gram matrix converging to the identity is eventually
    invertible, giving linear independence.
\end{proof}

\section{BNT permutation rigidity}

\begin{theorem}[BNT permutation --- proportional MPV case]\label{thm:bnt_perm_prop}
    \lean{MPSTensor.exists_eq_numBlocks_and_equiv_gaugePhase_of_proportional_decomp}
    \leanok
    \uses{def:bnt, def:gauge_phase_equiv}
    Let $(A_j)_{j=1}^{g_A}$ and $(B_k)_{k=1}^{g_B}$ be two
    BNT decompositions such that the block-diagonal tensors
    generate proportional MPV families with convergent nonzero
    coefficients.
    Then $g_A = g_B$, and there is a permutation
    $\sigma \in S_g$ with $D_{\sigma(k)} = D_k$ and
    $A_{\sigma(k)}$ gauge-phase equivalent to $B_k$ for each~$k$.
\end{theorem}

\begin{proof}
    \leanok
    From the proportional MPV hypothesis and convergent
    coefficients, one extracts (for each pair $(j,k)$) the
    limit $\lim_{N \to \infty} O_{A_j B_k}(N)$.
    By the spectral gap (Chapter~\ref{ch:spectral}),
    cross-block overlaps between non-equivalent blocks
    decay to~$0$.
    The resulting overlap matrix has at most one nonzero entry
    per row and per column, which defines a permutation.
    Nonzero overlap forces gauge-phase equivalence.

    This is \cite[Theorem~4.4]{Cirac2021Matrix},
    proved there using the uniqueness of BNT decompositions.
\end{proof}

\section{Coefficient convergence}

\begin{theorem}[Coefficient ratio decay]\label{thm:coeff_ratio_decay}
    \lean{MPSTensor.IsCanonicalForm.coeff_ratio_tendsto}
    \leanok
    \uses{def:is_canonical_form}
    Under the canonical form predicate, for $k \ge 1$,
    $(\mu_k / \mu_0)^N \to 0$
    as $N \to \infty$.
\end{theorem}

\begin{proof}\leanok
    Since $|\mu_k| < |\mu_0|$ (strict ordering of moduli),
    $|\mu_k / \mu_0| < 1$, so the geometric sequence
    converges to~$0$.
\end{proof}

\section{Newton--Girard identities}

\begin{theorem}[Newton--Girard trace recursion]\label{thm:newton_girard}
    \lean{Matrix.charpoly_eq_of_forall_trace_pow_eq}
    \leanok
    If $A, B \in \MN{n}$ satisfy
    $\tr(A^k) = \tr(B^k)$ for all $k \ge 1$,
    then $A$ and $B$ have the same characteristic polynomial.
\end{theorem}

\begin{proof}
    \leanok
    The Newton--Girard identities express the coefficients
    of the characteristic polynomial recursively in terms
    of the power-sum traces $\tr(A^k)$.
    Equal power sums give equal coefficients, hence equal
    characteristic polynomials.
\end{proof}

\begin{theorem}[Equal power sums imply equal multisets]\label{thm:power_sum_multiset}
    \lean{Matrix.sum_pow_eq_implies_multiset_eq}
    \leanok
    \uses{thm:newton_girard}
    Let $\alpha, \beta : \{1, \ldots, n\} \to \C$.
    If $\sum_i \alpha_i^k = \sum_i \beta_i^k$ for all $k \ge 1$,
    then the multisets $\{\alpha_i\}$ and $\{\beta_i\}$ are equal.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:newton_girard}
    Construct diagonal matrices $\diag(\alpha)$ and
    $\diag(\beta)$.
    The power-sum hypothesis gives
    $\tr(\diag(\alpha)^k) = \tr(\diag(\beta)^k)$.
    By Theorem~\ref{thm:newton_girard}, the characteristic
    polynomials agree: $\prod_i (X - \alpha_i) = \prod_i (X - \beta_i)$.
    Extracting roots gives multiset equality.
\end{proof}

\section{Primitivity bridge}

\begin{definition}[Primitive MPS tensor]\label{def:primitive_mps}
    \lean{MPSTensor.IsPrimitiveMPS}
    \leanok
    \uses{def:mps_tensor}
    A \emph{primitive MPS tensor} is one equipped with a nonzero
    PSD fixed point $\rho$ of the transfer map such that the
    complementary spectral radius
    $\rho(\E_A - P) < 1$
    (where $P$ is the fixed-point projection).
\end{definition}

\begin{theorem}[Primitive overlap convergence]\label{thm:prim_overlap_one}
    \lean{MPSTensor.IsPrimitiveMPS.overlap_tendsto_one}
    \leanok
    \uses{def:primitive_mps, thm:primitive_overlap}
    If $A$ is a primitive MPS tensor, then
    $O_{AA}(N) \to 1$.
\end{theorem}

\begin{proof}\leanok\uses{thm:primitive_overlap}
    Immediate from the primitive overlap convergence
    (Theorem~\ref{thm:primitive_overlap}).
\end{proof}

\section{Full assembly}

\begin{theorem}[Fundamental Theorem --- proportional MPV case]\label{thm:ft_proportional}
    \lean{MPSTensor.fundamentalTheorem_proportionalMPV_CFBNT}
    \leanok
    \uses{def:cf_bnt, def:gauge_phase_equiv}
    Let $((\mu_k^A), (A_k))$ and $((\mu_k^B), (B_k))$ be two
    CF-BNT decompositions.
    If the block-diagonal tensors generate proportional MPV
    families, then:
    \begin{enumerate}
        \item $r_A = r_B$ (same number of blocks),
        \item there exists a permutation $\sigma$ with
              $D_{\sigma(k)}^A = D_k^B$,
        \item each pair $(A_{\sigma(k)}, B_k)$ is gauge-phase
              equivalent.
    \end{enumerate}
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:cf_bnt_is_bnt, thm:bnt_perm_prop,
          thm:coeff_ratio_decay, thm:ft_canonical}
    The CF-BNT predicate produces a BNT decomposition
    (Theorem~\ref{thm:cf_bnt_is_bnt}).
    Coefficient convergence (Theorem~\ref{thm:coeff_ratio_decay})
    provides the convergent nonzero coefficients.
    The BNT permutation theorem
    (Theorem~\ref{thm:bnt_perm_prop}) gives the block count,
    permutation, and per-block gauge-phase equivalence.
\end{proof}

\begin{theorem}[Fundamental Theorem --- equal MPV case]\label{thm:ft_equal}
    \lean{MPSTensor.fundamentalTheorem_equalMPV_CFBNT}
    \leanok
    \uses{def:cf_bnt, def:gauge_equiv}
    Under the same hypotheses as Theorem~\ref{thm:ft_proportional},
    if the MPV families are \emph{equal} (not just proportional),
    then the block-diagonal tensors are gauge equivalent (not just
    gauge-phase equivalent).
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:ft_proportional, thm:power_sum_multiset,
          thm:ft_canonical}
    Theorem~\ref{thm:ft_proportional} gives the permutation
    and per-block gauge-phase equivalence.
    From equal MPVs, the power-sum identity
    $\sum_k (\mu_k^A)^N = \sum_k (\mu_k^B)^N$ holds for all~$N$.
    By Theorem~\ref{thm:power_sum_multiset},
    the multisets $\{\mu_k^A\}$ and $\{\mu_k^B\}$ agree.
    Combining with the permutation and per-block gauge-phase
    equivalence gives global gauge equivalence.

    This is the full Fundamental Theorem of MPS as stated in
    \cite[Theorem~4.1]{Cirac2021Matrix}
    and originally proved in~\cite{PerezGarcia2007Matrix}.
    The key deviation from the literature is the explicit
    use of Newton--Girard identities
    (Theorem~\ref{thm:newton_girard}) to pass from
    equal power sums of scaling factors to
    multiset equality.
\end{proof}