\chapter{Canonical-form data after blocking}\label{ch:canonical_after_blocking}

This chapter collects the canonical-form bridge data consumed by
Chapter~\ref{ch:ft_proof} after the initial blocking and zero-tail reductions.
The statements form an interface between the canonical-form chapter and the
Fundamental-Theorem comparison: they package trace-preserving gauges, primitive
irreducible sector data, common blocked alphabets, and zero-tail bookkeeping.

Conceptually, this material belongs to the blocking part of the canonical-form
construction in \cite[Section~2.3]{Cirac2016MPDO_arXiv} and to the normal/BNT
setup surrounding \cite[Section~IV.A]{Cirac2021Matrix}. The individual blueprint
statements below are implementation bridge lemmas rather than paper-numbered
theorems.

\section{Zero-tail separation (relocated from Chapter~\ref{ch:canonical})}%
\label{sec:zero_block_separation}

In the irreducible block decomposition some blocks may have all-zero Kraus
operators.  Such blocks contribute to the MPV only at length $N=0$, where the
contribution equals their bond dimension.  The three statements below separate
the all-zero blocks from the nonzero blocks; they are consumed throughout this
chapter (zero-tail bookkeeping) and in Chapter~\ref{ch:ft_proof}.

\begin{definition}[Zero MPS tensor]\label{def:zero_mps_tensor}
	\lean{MPSTensor.zeroMPSTensor}
	\leanok
	The \emph{zero MPS tensor} of physical dimension $d$ and bond
	dimension $D$, denoted $\mathbf{0}_{d,D}$, is the tensor whose
	Kraus operators are all zero.
\end{definition}

\begin{theorem}[MPV of zero tensor]\label{thm:mpv_zero_tensor}
	\lean{MPSTensor.mpv_zeroMPSTensor}
	\leanok
	\uses{def:zero_mps_tensor}
	For any spin configuration $\sigma$ of length $N$,
	\[
		V^{(N)}(\mathbf{0}_{d,D})_\sigma
		= \begin{cases} D & \text{if } N = 0, \\ 0 & \text{if } N \ge 1. \end{cases}
	\]
\end{theorem}

\begin{proof}\leanok
	For $N = 0$, $(\mathbf{0}_{d,D})^{\varnothing} = \Id_{D}$,
	so the MPV equals $\tr(\Id_D) = D$.
	For $N \ge 1$, every word product of a zero tensor is zero.
\end{proof}

\begin{theorem}[Zero-block separation]%
	\label{thm:zero_block_separation}
	\lean{MPSTensor.exists_irreducible_blockDecomp_nonzeroBlocks}
	\leanok
	\uses{def:irreducible_tensor, def:zero_mps_tensor}
	Every MPS tensor $A$ admits a block decomposition into:
	\begin{enumerate}
		\item a \emph{trivial block} of bond dimension $D_0$ with all Kraus operators zero
		      (accumulating all all-zero irreducible blocks), and
		\item a finite family of \emph{nontrivial blocks} $(B_k)_{k=1}^r$,
		      each irreducible, each with at least one nonzero Kraus
		      operator, and each with positive bond dimension,
	\end{enumerate}
	such that
	\[
		V^{(N)}(A)_\sigma
		= V^{(N)}(\mathbf{0}_{d,D_0})_\sigma
		+ V^{(N)}\!\bigl(\textstyle\bigoplus_{k=1}^r B_k\bigr)_\sigma
	\]
	for every spin configuration $\sigma$ of length $N$.
	At $N = 0$ this reads $D = D_0 + \sum_{k=1}^r D_k$.
	For $N \ge 1$ the zero-tail term vanishes.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:irred_decomp, thm:mpv_zero_tensor}
	Apply the irreducible block decomposition
	(Theorem~\ref{thm:irred_decomp}) and partition the blocks
	by whether each has a nonzero Kraus operator.
	The all-zero blocks satisfy a dimension bound via
	irreducibility; their total bond dimension is $D_0$.
	The MPV identity follows from the block structure.
\end{proof}

\section{Reduction and structural after-blocking results}

\begin{theorem}[Blockwise TP-gauge normalization of irreducible blocks]%
	\label{thm:blockwise_tp_gauge}
	\lean{MPSTensor.exists_tp_gauge_blockwise}
	\leanok
	\uses{def:irreducible_tensor}
	Suppose $A$ generates the same MPV family as a block-diagonal tensor
	$\bigoplus_{k=1}^r B_k$ with each block $B_k$ irreducible and containing at
	least one nonzero Kraus operator.
	Then, after rescaling the sector weights, one obtains blocks $(C_k)$ such that
	$A \sim \bigoplus_k \mu_k C_k$, each $C_k$ is irreducible, each $C_k$ is
	left-canonical, all $\mu_k \neq 0$, and all bond dimensions are positive.
\end{theorem}

\begin{proof}\leanok
	The Perron--Frobenius TP-gauge construction is applied separately to each
	irreducible block.
	Gauge equivalence preserves irreducibility and the MPV family, and the
	positive scaling factors are absorbed into the new weights.
\end{proof}

\begin{theorem}[Primitive nonzero blocks are normal]%
	\label{thm:normal_live_block_primitive}
	\lean{MPSTensor.isNormal_live_block_of_primitive}
	\leanok
	\uses{def:normal, def:irreducible_tensor}
	A left-canonical irreducible tensor with primitive transfer map and positive
	bond dimension is normal.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:normal_tp_primitive_irred}
	Apply Theorem~\ref{thm:normal_tp_primitive_irred}.
\end{proof}

\begin{theorem}[Blocking preserves irreducibility for primitive irreducible blocks]%
	\label{thm:blocked_irreducible_tp_primitive}
	\lean{MPSTensor.isIrreducibleTensor_blockTensor_of_tp_primitive_irr}
	\leanok
	\uses{def:blocked_tensor, def:irreducible_tensor}
	If $A$ is left-canonical, irreducible, primitive, and has positive bond
	dimension, then every positive blocking length $P$ preserves irreducibility:
	$A^{[P]}$ is again irreducible.
\end{theorem}

\begin{proof}\leanok
	A positive-definite fixed point of $\E_A$ remains fixed by the blocked transfer
	map.
	Primitivity gives uniqueness of positive semidefinite fixed points for the
	blocked map, hence the blocked transfer map is irreducible; this is exactly
	the irreducibility criterion for the blocked tensor.
\end{proof}

\begin{theorem}[Extra blocking of primitive irreducible sector blocks]%
\label{thm:tp_primitive_irreducible_extra_blocking}
	\lean{MPSTensor.tp_primitive_irreducible_extra_blocking}
	\leanok
	\uses{def:blocked_tensor, thm:blocked_irreducible_tp_primitive}
	If a sector block is already trace-preserving, has primitive transfer map, and
	is tensor-irreducible, then every later positive blocking length $k$ preserves
	all three properties. This $k$ is the subsequent common-refinement or
	injectivity length, distinct from the period-removal length that produced the
	sector block.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:blocked_irreducible_tp_primitive}
	Trace preservation follows from blocking a trace-preserving tensor, primitivity
	from taking a positive power of a primitive transfer map, and irreducibility
	from Theorem~\ref{thm:blocked_irreducible_tp_primitive}.
\end{proof}

\begin{theorem}[Sorted TP primitive irreducible data form a normal canonical form]%
	\label{thm:ncf_sorted_tp_primitive_irred}
	\lean{MPSTensor.isNormalCanonicalForm_of_tp_primitive_irr_sorted}
	\leanok
	\uses{def:is_normal_canonical_form}
	Let $(\mu_k, A_k)$ be a finite block family with strictly decreasing norms
	$|\mu_1| > \cdots > |\mu_r|$.
	If every block $A_k$ is left-canonical, primitive, irreducible, has positive
	bond dimension, and every $\mu_k \neq 0$, then $(\mu_k,A_k)$ is a normal
	canonical form.
\end{theorem}

\begin{proof}\leanok
	\uses{def:is_normal_canonical_form}
	This is the defining constructor for normal canonical form with all required
	data supplied by the hypotheses.
\end{proof}

\begin{theorem}[Normal canonical form from an already primitive block decomposition]%
	\label{thm:ncf_from_primitive_input}
	\lean{MPSTensor.exists_normalCanonicalForm_of_primitive_input}
	\leanok
	\uses{def:is_normal_canonical_form}
	If $A$ already admits a block decomposition into irreducible left-canonical
	primitive blocks with pairwise distinct weight norms, nonzero weights, and
	positive bond dimensions, then some blocked tensor $A^{[p]}$ is MPV-equivalent
	to a normal canonical form.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:exists_normal_canonical_form}
	Apply Theorem~\ref{thm:exists_normal_canonical_form} to the given primitive
	block decomposition.
\end{proof}

\begin{theorem}[Common blocking period for two primitive normal tensors]%
	\label{thm:common_period_two_tensors}
	\lean{MPSTensor.bilateral_commonPeriod_blocking_tp_primitive_normal}
	\leanok
	\uses{def:blocked_tensor, def:normal}
	Let $A$ and $B$ be left-canonical normal tensors of positive bond dimensions.
	Suppose $A^{[p_A]}$ and $B^{[p_B]}$ already have primitive transfer maps for
	some positive periods $p_A,p_B$.
	Then there exists a common positive period $p$ such that both
	$A^{[p]}$ and $B^{[p]}$ are left-canonical, primitive, and normal.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:primitive_blocking_divisible, thm:isNormal_blockTensor}
	Take $p$ to be the least common multiple of $p_A$ and $p_B$.
	Primitivity follows from divisibility, and left-canonical normalization and
	normality are both preserved under blocking.
\end{proof}

\begin{theorem}[Structural after-blocking decomposition]%
	\label{thm:ft_after_blocking_structural}
	\lean{MPSTensor.afterBlocking_tpPrimitiveBlockDecompositions_of_sameMPV₂}
	\leanok
	\uses{def:blocked_tensor, def:same_mpv2, def:zero_mps_tensor, def:block_diagonal_tensor}
	If two tensors generate the same MPV family, then each tensor admits some
	blocking after which it decomposes into a zero-tail tensor plus nonzero
	left-canonical blocks with primitive transfer maps and positive bond
	dimensions. Moreover, at every blocked system size and every blocked word,
	the MPV of the blocked tensor is the sum of the zero-tail MPV and the MPV of
	the weighted nonzero-block tensor:
	\[
		V^{(N)}(A^{[p_A]})_\sigma
		= V^{(N)}(Z_{p_A,z_A})_\sigma
		  + V^{(N)}\!\left(\bigoplus_k \mu^A_k A_k\right)_\sigma,
	\]
	and likewise for $B$. The blocked tensors also preserve the original MPV
	equality:
	\[
		A^{[p_A]} \sim B^{[p_A]},
		\qquad
		A^{[p_B]} \sim B^{[p_B]}.
	\]
\end{theorem}

\begin{proof}\leanok
	\uses{thm:tp_primitive_blockdecomp}
	Apply Theorem~\ref{thm:tp_primitive_blockdecomp} separately to the two tensors.
\end{proof}

\begin{remark}[Structural after-blocking decomposition; not the FT block-matching theorem]
  This theorem produces, for each of the two same-MPV tensors independently,
  a TP/primitive/left-canonical block decomposition after blocking.  It does
  \emph{not} derive any block matching, gauge equivalence, or phases between
  the two decompositions; those are produced by the subsequent common-sector
  pipeline in this chapter.  It is structural scaffolding for the FT proof,
  not the multi-block fundamental theorem of
  \cite[Theorem~II.1, lines~349--352]{Cirac2017MPDO_AnnPhys}.
\end{remark}

\begin{theorem}[Zero-tail decomposition after blocking]%
	\label{thm:zero_tail_decomp_block_tensor}
	\lean{MPSTensor.zeroTail_mpv_decomp_blockTensor}
	\leanok
	\uses{def:blocked_tensor, def:zero_mps_tensor}
	Suppose $A$ is expressed as a zero-tail tensor plus a nonzero part $L$ at all
	system sizes. For every positive blocking length $p$, the blocked tensor
	$A^{[p]}$ is expressed as the same zero-tail bond dimension plus $L^{[p]}$.
	The zero-tail term still contributes only at blocked length $0$.
\end{theorem}

\begin{proof}\leanok
	Flatten a blocked word of length $N$ to an original word of length $Np$. Since
	$p>0$, this length is zero exactly when $N=0$, so the zero-tail MPV is
	unchanged by the reblocking convention.
\end{proof}

\begin{theorem}[Zero-tail weighted block decomposition after blocking]%
	\label{thm:zero_tail_to_tensor_from_blocks_block_power}
	\lean{MPSTensor.zeroTail_toTensorFromBlocks_blockPower}
	\leanok
	\uses{thm:zero_tail_decomp_block_tensor, thm:block_tensor_to_tensor_from_blocks,
	      def:block_diagonal_tensor}
	Suppose $A$ is expressed as a zero-tail tensor plus the weighted nonzero-block
	tensor $\bigoplus_k \mu_k B_k$. For every positive blocking length $p$,
	$A^{[p]}$ is expressed as the same zero-tail contribution plus the weighted
	blocked nonzero tensor $\bigoplus_k \mu_k^p B_k^{[p]}$.
\end{theorem}

\begin{proof}\leanok
	Apply Theorem~\ref{thm:zero_tail_decomp_block_tensor} to the assembled
	weighted nonzero tensor and rewrite the blocked assembled tensor using
	Theorem~\ref{thm:block_tensor_to_tensor_from_blocks}.
\end{proof}

\begin{theorem}[Length-zero identity after blocking weighted blocks]%
	\label{thm:zero_tail_identity_to_tensor_from_blocks_block_power}
	\lean{MPSTensor.zeroTail_identity_toTensorFromBlocks_blockPower}
	\leanok
	\uses{thm:block_tensor_to_tensor_from_blocks}
	If the length-zero zero-tail identity holds for two assembled weighted block
	tensors, then after any blocking length $p \ge 0$ the same identity holds for
	the blocked families with weights transported to the powers $\mu_k^p$.
\end{theorem}

\begin{proof}\leanok
	At blocked length zero the flattened configuration is empty.  The blocked
	assembled tensors are rewritten by Theorem~\ref{thm:block_tensor_to_tensor_from_blocks},
	and the original length-zero identity is applied to the empty configuration.
\end{proof}

\begin{theorem}[Zero-tail identity for nonzero block tensors]%
\label{thm:nonzero_block_zero_tail_identity}
	\lean{MPSTensor.nonzeroBlock_positive_sameMPV₂_and_zeroTail_identity_of_sameMPV₂}
	\leanok
	\uses{def:same_mpv2, def:zero_mps_tensor, def:block_diagonal_tensor}
	Suppose two tensors have the same MPV family and each is expressed as a
	zero-tail tensor plus a weighted nonzero-block tensor. Then the two nonzero-block
	tensors agree on every positive system size, and at size $0$ the equality is
	exactly the equality of the two zero-tail contributions plus the two nonzero
	bond-dimension traces.
\end{theorem}

\begin{proof}\leanok
	\uses{def:zero_mps_tensor}
	For $N>0$, the MPV of a zero-tail tensor is zero, so the two decompositions
	can be compared directly through the common MPV family. For $N=0$, the
	zero-tail MPV is its bond dimension, giving the stated length-zero identity.
\end{proof}

\begin{theorem}[Blocked nonzero parts and length zero]%
	\label{thm:nonzero_block_block_power_zero_tail_identity}
	\lean{MPSTensor.nonzeroBlock_blockPower_positive_sameMPV₂_and_zeroTail_identity_of_sameMPV₂}
	\leanok
	\uses{thm:nonzero_block_zero_tail_identity,
	      thm:zero_tail_to_tensor_from_blocks_block_power, def:same_mpv2_pos}
	In the setting of Theorem~\ref{thm:nonzero_block_zero_tail_identity}, after
	any positive common blocking length $p$, the powered weighted nonzero-block tensors
	agree on all positive blocked system sizes. At blocked length $0$, the same
	zero-tail identity holds with weights transported from $\mu_k$ to
	$\mu_k^p$.
\end{theorem}

\begin{proof}\leanok
	Use Theorem~\ref{thm:zero_tail_to_tensor_from_blocks_block_power} on both
	zero-tail decompositions, block the original MPV equality, and apply
	Theorem~\ref{thm:nonzero_block_zero_tail_identity} at the blocked physical
	dimension.
\end{proof}

\begin{theorem}[Equal zero tails give full nonzero-part MPV equality]%
\label{thm:nonzero_block_same_mpv_zero_tail_eq}
	\lean{MPSTensor.nonzeroBlock_sameMPV₂_of_sameMPV₂_of_zeroTail_eq}
	\leanok
	\uses{thm:nonzero_block_zero_tail_identity}
	In the setting of Theorem~\ref{thm:nonzero_block_zero_tail_identity}, if
	the two zero-tail dimensions are equal, then the two nonzero-block tensors have
	the same MPV family, including the length-zero case.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:nonzero_block_zero_tail_identity}
	The positive-length cases are exactly the first conclusion of
	Theorem~\ref{thm:nonzero_block_zero_tail_identity}. At $N=0$, substitute
	the equality of zero-tail dimensions into the length-zero identity and cancel
	the common scalar term.
\end{proof}


\begin{theorem}[Cyclic sectors after the zero-tail split]%
\label{thm:ft_after_blocking_per_block_cyclic_live_zero_tail}
	\lean{MPSTensor.afterBlocking_perBlockCyclicDataWithZeroTail_of_sameMPV₂}
	\leanok
	\uses{thm:tp_gauge_arbitrary, thm:nonzero_block_zero_tail_identity,
		def:primitive_irreducible_cyclic_sectors,
		thm:has_primitive_irreducible_cyclic_sectors_tp_irr}
	If two tensors generate the same MPV family, then after the zero-tail split and
	trace-preserving gauge each side has irreducible nonzero-weight blocks.
	The nonzero-block tensors agree at all positive system sizes, while the
	length-zero case is included as the exact zero-tail identity.
	Moreover each nonzero-weight block has primitive irreducible cyclic sectors. The
	period-removal length for those sectors is kept separate from any later common
	blocking or injectivity length.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:tp_gauge_arbitrary, thm:nonzero_block_zero_tail_identity,
		thm:has_primitive_irreducible_cyclic_sectors_tp_irr}
	Apply the arbitrary-tensor TP-gauge theorem separately to $A$ and $B$. The
	zero-tail identity theorem compares the two resulting nonzero parts at
	positive lengths and gives the $N=0$ identity. Finally apply
	Theorem~\ref{thm:has_primitive_irreducible_cyclic_sectors_tp_irr} to every
	trace-preserving irreducible nonzero-weight block.
\end{proof}

\begin{theorem}[Common blocking for cyclic sectors]%
\label{thm:ft_after_blocking_common_blocked_cyclic_live_zero_tail}
	\lean{MPSTensor.afterBlocking_commonBlockedCyclicDataWithZeroTail_of_sameMPV₂}
	\leanok
	\uses{thm:ft_after_blocking_per_block_cyclic_live_zero_tail,
		thm:exists_common_blocked_cyclic_sector_family,
		thm:common_blocked_cyclic_sector_flat_weight_ne_zero}
	If two tensors generate the same MPV family, then the zero-tail/TP-gauge
	reduction for nonzero-weight blocks can be combined with a common reblocking
	of all per-block cyclic sectors.  Each side obtains a finite flattened sector family
	at one physical blocking length, with trace preservation, primitive transfer
	maps, tensor irreducibility, positive bond dimensions, and nonzero unit
	weights.  The theorem keeps the positive-length equality of the nonzero parts and the exact
	$N=0$ zero-tail identity for the original nonzero-block tensors.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:ft_after_blocking_per_block_cyclic_live_zero_tail,
		thm:exists_common_blocked_cyclic_sector_family,
		thm:common_blocked_cyclic_sector_flat_weight_ne_zero}
	Apply Theorem~\ref{thm:ft_after_blocking_per_block_cyclic_live_zero_tail}.
	Then apply Theorem~\ref{thm:exists_common_blocked_cyclic_sector_family} to the
	per-block cyclic-sector data on each side.  The nonzero unit-weight conclusion
	is Theorem~\ref{thm:common_blocked_cyclic_sector_flat_weight_ne_zero}.
\end{proof}

The next results put the cyclic-sector decompositions on a common blocked
alphabet.  They compare direct blocking with iterated cyclic-sector blocking and
transport the zero-tail identity through this comparison.

\begin{theorem}[Common cyclic sectors after reblocking]%
\label{thm:ft_after_blocking_reindexed_common_sector_live_zero_tail}
	\lean{MPSTensor.afterBlocking_reindexedCommonSectorDataWithZeroTail_of_sameMPV₂}
	\leanok
	\uses{thm:ft_after_blocking_common_blocked_cyclic_live_zero_tail,
		thm:zero_tail_to_tensor_from_blocks_block_power,
		thm:common_blocked_cyclic_sector_reindexed_samempv,
		thm:common_blocked_cyclic_sector_derived_properties}
	If two tensors generate the same MPV family, then after reblocking the nonzero
	blocks one obtains cyclic-sector families over a common alphabet.  These
	families preserve the MPV of the reblocked nonzero part, with nonzero
	transported weights $\mu_k^p$, and retain the zero-tail identity.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:ft_after_blocking_common_blocked_cyclic_live_zero_tail,
		thm:zero_tail_to_tensor_from_blocks_block_power,
		thm:common_blocked_cyclic_sector_reindexed_samempv,
		thm:common_blocked_cyclic_sector_derived_properties}
	Start from the common reblocked cyclic-sector families on the two sides.  Apply
	the zero-tail reblocking theorem to the original nonzero-block decomposition, then
		use the common-alphabet sector decomposition and the derived structural
	properties of the common-sector family.
\end{proof}

\begin{theorem}[Common blocking length for cyclic sectors]%
\label{thm:ft_after_blocking_common_length_common_sector_theorem}
	\lean{MPSTensor.afterBlocking_commonLengthCommonSectorData_of_sameMPV₂}
	\leanok
	\uses{thm:ft_after_blocking_per_block_cyclic_live_zero_tail,
		thm:exists_common_blocked_cyclic_sector_family_common_multiple,
		thm:zero_tail_to_tensor_from_blocks_block_power,
		thm:nonzero_block_block_power_zero_tail_identity,
		thm:common_blocked_cyclic_sector_reindexed_samempv,
		thm:common_blocked_cyclic_sector_derived_properties}
	If two tensors generate the same MPV family, then the cyclic-sector
	decompositions on the two sides can be expressed at one common positive
	blocking length.  At this length the nonzero parts still agree at positive
	system sizes, and the length-zero identity is exactly the zero-tail identity.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:ft_after_blocking_per_block_cyclic_live_zero_tail,
		thm:exists_common_blocked_cyclic_sector_family_common_multiple,
		thm:zero_tail_to_tensor_from_blocks_block_power,
		thm:nonzero_block_block_power_zero_tail_identity,
		thm:common_blocked_cyclic_sector_reindexed_samempv,
		thm:common_blocked_cyclic_sector_derived_properties}
	Start from the per-block cyclic-sector theorem.  Let $p_A$ and $p_B$ be the
	least common multiples of the period-removal lengths on the two sides, and
	use $p=\mathrm{lcm}(p_A,p_B)$.  The prescribed-common-multiple theorem constructs both
	one-sided sector families at this same length.  The zero-tail and positive-
	length equalities are the existing reblocking statements, and the sector
	conclusions are the common-alphabet MPV equalities.
\end{proof}

\begin{theorem}[Length-zero identity under compatible blocking]%
\label{thm:zero_tail_common_flat_of_blocked_word_comparison}
	\lean{MPSTensor.zeroTail_commonFlat_of_blockwise,
		MPSTensor.zeroTail_commonFlat_of_word_eq,
		MPSTensor.zeroTail_commonFlat_of_groupedBlockCastAgrees,
		MPSTensor.zeroTail_commonFlatAt_of_groupedBlockCastAgrees,
		MPSTensor.sameMPV₂Pos_blockTensor_commonFlatAt_of_groupedBlockCastAgrees}
	\leanok
	\uses{thm:zero_tail_to_tensor_from_blocks_block_power,
		thm:common_blocked_cyclic_sector_blocked_word_comparison,
		def:same_mpv2_pos}
	Assume a zero-tail/nonzero decomposition on one side.  If direct blocking of
	each nonzero block agrees with the corresponding iterated cyclic-sector
	blocking, then the zero-tail identity can be rewritten using the common
	cyclic-sector family.  The coordinate-grouping condition supplies this
	agreement at any prescribed common blocking length.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:zero_tail_to_tensor_from_blocks_block_power,
		thm:common_blocked_cyclic_sector_blocked_word_comparison,
		def:same_mpv2_pos}
	Transport the zero-tail/nonzero decomposition to the common blocking length.
	The blockwise comparison then replaces the directly blocked nonzero part by the
	weighted common-sector tensor.
\end{proof}

\begin{theorem}[Length-zero identity under word identification]%
\label{thm:zero_tail_common_flat_of_reindexed}
	\lean{MPSTensor.zeroTail_commonFlat_of_reindexed,
		MPSTensor.zeroTail_commonFlatAt_of_reindexed,
		MPSTensor.sameMPV₂Pos_blockTensor_commonFlatAt_of_reindexed}
	\leanok
	\uses{thm:zero_tail_to_tensor_from_blocks_block_power,
		thm:common_blocked_cyclic_sector_reindexed_nonzero_part}
	Assume a zero-tail/nonzero decomposition on one side.  If the blocked nonzero
	tensor agrees, after identifying blocked words, with the common cyclic-sector
	family, then the same zero-tail identity can be written in that common
	alphabet.  At positive system sizes this says that the blocked tensor and the
	weighted common-sector tensor have the same MPV coefficients.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:zero_tail_to_tensor_from_blocks_block_power,
		thm:common_blocked_cyclic_sector_reindexed_nonzero_part}
	Transport the zero-tail/nonzero decomposition to the common blocking length.
	The word-identification hypothesis then replaces the blocked nonzero tensor by the
	weighted common-sector tensor.  At positive lengths the zero-tail tensor has
	zero MPV coefficients.
\end{proof}

\begin{theorem}[Primitive block decompositions at a common length]%
\label{thm:after_blocking_common_primitive_irreducible_blocks_reindexed}
	\lean{MPSTensor.afterBlocking_commonPrimitiveIrreducibleBlocks_of_reindexedNonzeroParts}
	\leanok
	\uses{thm:ft_after_blocking_common_length_common_sector_theorem,
		thm:common_blocked_cyclic_sector_reindexed_nonzero_part,
		def:same_mpv2_pos}
	Assume that the nonzero part can be identified with the common blocked
	alphabet.  Then two tensors with the same MPV family admit one positive
	blocking length at which both nonzero parts are weighted sums of
	trace-preserving, primitive, tensor-irreducible blocks with positive bond
	dimensions and nonzero weights.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:ft_after_blocking_common_length_common_sector_theorem,
		thm:common_blocked_cyclic_sector_reindexed_nonzero_part}
	Apply Theorem~\ref{thm:ft_after_blocking_common_length_common_sector_theorem}
	to choose the common blocking length and the two common cyclic-sector families.
	The assumed word identification converts the blocked nonzero part into the weighted
	common-sector family on each side.  The zero-tail equations then give the
	displayed decompositions.  The structural properties and nonzero weights come
	from the common-sector families.
\end{proof}

\begin{theorem}[Transport of MPV and length-zero identities]%
\label{thm:samempv_pos_zero_tail_identity_transport}
	\lean{MPSTensor.sameMPV₂Pos_and_zeroTail_identity_of_sameMPV₂}
	\leanok
	\uses{def:same_mpv2, def:same_mpv2_pos}
	Let $L_A,L_B$ be two nonzero parts whose MPV families agree at positive
	lengths and whose zero-tail dimensions satisfy the length-zero identity.  If
	$L_A'$ and $L_B'$ generate the same MPV families as $L_A$ and $L_B$, respectively,
	then $L_A'$ and $L_B'$ satisfy the same positive-length equality and the same
	length-zero zero-tail identity.
\end{theorem}

\begin{proof}\leanok
	\uses{def:same_mpv2, def:same_mpv2_pos}
	Use the two MPV equalities to replace $L_A'$ by $L_A$ and $L_B$ by $L_B'$ in
	the positive-length equality and in the $N=0$ identity.
\end{proof}

\begin{theorem}[Length-zero identity for common sectors]%
\label{thm:zero_tail_common_flat_transport_of_reindexed}
	\lean{MPSTensor.zeroTail_commonFlat_transport_of_reindexed}
	\leanok
	\uses{thm:zero_tail_common_flat_of_reindexed,
		thm:common_blocked_cyclic_sector_reindexed_nonzero_part,
		thm:common_blocked_cyclic_sector_derived_properties}
	If the blocked nonzero part agrees with the common-sector tensor,
	then the zero-tail identity can be written with that common-sector tensor.
	Nonzero original weights remain nonzero after the blocking power.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:zero_tail_common_flat_of_reindexed,
		thm:common_blocked_cyclic_sector_reindexed_nonzero_part,
		thm:common_blocked_cyclic_sector_derived_properties}
	Combine Theorem~\ref{thm:zero_tail_common_flat_of_reindexed} with the
	common-sector MPV equality and the nonzero-weight statement.
\end{proof}

\begin{theorem}[Common sectors from grouped coordinates]%
\label{thm:zero_tail_common_flat_transport_of_grouped_block_cast}
	\lean{MPSTensor.zeroTail_commonFlat_transport_of_groupedBlockCastAgrees}
	\leanok
	\uses{thm:zero_tail_common_flat_of_blocked_word_comparison,
		thm:common_blocked_cyclic_sector_blocked_word_comparison,
		thm:common_blocked_cyclic_sector_derived_properties,
		def:same_mpv2}
	Assume a zero-tail/nonzero decomposition with nonzero weights.  If the common
	blocked alphabet is identified with each iterated alphabet by grouping
	consecutive words, then the zero-tail identity can be written with the
	weighted common-sector family.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:zero_tail_common_flat_of_blocked_word_comparison,
		thm:common_blocked_cyclic_sector_blocked_word_comparison,
		thm:common_blocked_cyclic_sector_derived_properties,
		def:same_mpv2}
	The coordinate-grouping comparison identifies the directly blocked nonzero part
	with the weighted common-sector family.  Combine this with the zero-tail
	identity and the nonzero-weight statement.
\end{proof}

\begin{theorem}[Common sectors at one blocking length]%
\label{thm:ft_after_blocking_common_length_common_sector_reindexed}
	\lean{MPSTensor.afterBlocking_commonLengthCommonSectorData_of_reindexed}
	\leanok
	\uses{thm:ft_after_blocking_common_length_common_sector_theorem,
		thm:common_blocked_cyclic_sector_reindexed_nonzero_part}
	If two tensors generate the same MPV family, choose a common blocking length
	for their cyclic-sector decompositions.  If each blocked nonzero part agrees
	with its common-sector form after identifying words, then all zero-tail
	identities and positive-length MPV equalities can be stated using the common
	sector families.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:ft_after_blocking_common_length_common_sector_theorem,
		thm:common_blocked_cyclic_sector_reindexed_nonzero_part}
	Apply Theorem~\ref{thm:ft_after_blocking_common_length_common_sector_theorem}.
	The two word-identification hypotheses identify the blocked nonzero parts with the
	weighted common-sector tensors.  Substituting these MPV equalities into the
	zero-tail equations gives the rewritten zero-tail identities, the
	positive-length MPV equalities, and the $N=0$ identity.
\end{proof}


\section{Primitive-block reduction (relocated from Chapter~\ref{ch:canonical})}%
\label{sec:ch11b_primitive_block_reduction}

The following two statements were previously presented in
Chapter~\ref{ch:canonical}.  They describe primitive blocked sectors and a
primitive-block decomposition starting from an arbitrary tensor; both are
after-blocking structural results, so they are collected here.

\begin{theorem}[Unconditional primitive blocked sectors]%
	\label{thm:blocked_sector_unconditional}
	\lean{MPSTensor.primitive_and_irreducible_sectorBlocks_of_cyclic_decomp_after_blocking}
	\leanok
	\uses{thm:cyclic_sector_decomp_after_blocking,
	      thm:sector_irred_unconditional,
	      thm:primitive_adjoint_equiv}
	Let $A$ be a left-canonical irreducible tensor whose adjoint transfer map has
	peripheral spectrum $\{\gamma^j : 0 \le j < m\}$ for a primitive $m$-th root
	of unity $\gamma$. Let $(C_u)_{u=1}^m$, $(P_u)_{u=1}^m$, and
	$\varphi_u : \MN{\dim C_u} \xrightarrow{\sim} P_u \MN{D} P_u$
	be the blocked cyclic-sector data. Then every blocked sector tensor $C_u$ has
	primitive transfer map and is irreducible.
\end{theorem}

\begin{proof}\leanok
	\uses{thm:sector_irred_unconditional,
	      thm:primitive_adjoint_equiv}
	Theorem~\ref{thm:sector_irred_unconditional} gives irreducibility of
	$(\E_A^\dagger)^m$ on each corner $P_u$.
	The blocked cyclic-sector data provide compression equivalences
	$\varphi_u$ intertwining the adjoint transfer map of $C_u$ with the
	corresponding corner restriction of $(\E_A^\dagger)^m$.
	The same cyclic peripheral-spectrum argument that produces the sector
	decomposition shows that each corner restriction is primitive.
	Transport primitivity and irreducibility across $\varphi_u$, then use
	Theorem~\ref{thm:primitive_adjoint_equiv} to pass from the adjoint blocked
	map to the primal transfer map of $C_u$.
\end{proof}

\begin{theorem}[Arbitrary tensor to primitive block decomposition]%
	\label{thm:tp_primitive_blockdecomp}
	\lean{MPSTensor.exists_tp_primitive_blockDecomp_after_blocking}
	\leanok
	\uses{def:blocked_tensor, def:zero_mps_tensor}
	For any MPS tensor $A$, there exist a period $p \ge 1$,
	a zero-tail bond dimension $D_0$, nonzero weights $(\mu_k)_{k=1}^r$,
	and left-canonical blocks $(B_k)_{k=1}^r$ with primitive transfer
	maps and positive bond dimensions, such that
	\[
		V^{(N)}\!\bigl(A^{[p]}\bigr)_\sigma
		= V^{(N)}\!\bigl(\mathbf{0}_{d^{[p]},D_0}\bigr)_\sigma
		+ V^{(N)}\!\bigl(\textstyle\bigoplus_k \mu_k B_k\bigr)_\sigma.
	\]
\end{theorem}

\begin{proof}\leanok
	\uses{thm:tp_gauge_arbitrary, thm:common_blocking_period,
	      lem:block_weights_ne_zero, thm:zero_tail_to_tensor_from_blocks_block_power}
	Apply Theorem~\ref{thm:tp_gauge_arbitrary} to get the trivial block
	and left-canonical nontrivial blocks.
	Apply Theorem~\ref{thm:common_blocking_period} to find a common
	blocking period $p$ making all transfer maps primitive. Lemma~\ref{lem:block_weights_ne_zero}
	keeps the blocked weights nonzero, and
	Theorem~\ref{thm:zero_tail_to_tensor_from_blocks_block_power} transports the
	zero-tail/nonzero MPV identity through the blocking step.
\end{proof}
