\chapter{Proof of the Fundamental Theorem}\label{ch:ft_proof}

This chapter records the aperiodic comparison results used in the proof of the
Fundamental Theorem of Matrix Product States.  We first state strict special
cases for canonical-form tensors with already selected BNT
representatives, and then state the sector-decomposition theorem that keeps the
repeated-copy multiplicities appearing in the CPSV BNT comparison.
We then treat the equal-MPV case, including the common-block-structure
specialization that deduces gauge equivalence.
The results combine the canonical form reduction
(Chapter~\ref{ch:canonical}),
block separation (Chapter~\ref{ch:block_perm}),
spectral gap (Chapter~\ref{ch:spectral}),
and the BNT permutation arguments of Chapter~\ref{ch:bnt}.

The theorems in this chapter follow the approach of
\cite{Cirac2017MPDO_AnnPhys}:
they first block the tensor to remove non-trivial periodicity, then
apply the aperiodic fundamental theorem.
Chapter~\ref{ch:periodic_ft_apps} develops the stronger
irreducible-form framework of \cite{DeLasCuevas2017Irreducible},
which gives a fundamental theorem valid for periodic tensors directly,
with an extra $Z$-gauge degree of freedom.

\section{Conditional Block Matching}%
\label{sec:ft_conditional_block_matching}

The following lemma combines the reduction to canonical form
(zero-block separation, TP gauge, blocking, primitivity)
with the restricted normal-CF-BNT proportional comparison.
It is still conditional on the two post-blocking hypotheses left open in
Chapter~\ref{ch:canonical}: irreducibility for the blocked sectors and
pairwise distinct nonzero weight norms.
The direct periodic theorem of Chapter~\ref{ch:periodic_ft_apps} is not
used in this chapter.

\begin{lemma}[Conditional block matching from restricted normal-CF-BNT hypotheses]%
    \label{lem:weak_ft_conditional}
    \lean{MPSTensor.weakFundamentalTheorem_conditional}
    \leanok
    \uses{def:is_normal_canonical_form, def:gauge_phase_equiv}
    Let $(\mu_j^A, A_j)_{j=1}^{g_A}$ and $(\mu_k^B, B_k)_{k=1}^{g_B}$
    be two families in normal canonical form,
    each satisfying the non-degeneracy condition
    that no two distinct blocks in the same family are gauge-phase
    equivalent.
    Let $A_\mathrm{tot}$ and $B_\mathrm{tot}$ be tensors with
    MPV decompositions
    \[
        V^{(N)}(A_\mathrm{tot})_\sigma = \sum_j a_{N,j}\, V^{(N)}(A_j)_\sigma,
        \quad
        V^{(N)}(B_\mathrm{tot})_\sigma = \sum_k b_{N,k}\, V^{(N)}(B_k)_\sigma,
    \]
    with $a_{N,j} \to a_j \neq 0$ and $b_{N,k} \to b_k \neq 0$.
    Suppose moreover that
    $V^{(N)}(A_\mathrm{tot})_\sigma = c_N\, V^{(N)}(B_\mathrm{tot})_\sigma$
    with $c_N \to c \neq 0$.
    Then $g_A = g_B$ and there is a permutation $\pi$ such that
    $D_j^A = D_{\pi(j)}^B$ and $A_j$ is gauge-phase equivalent to
    $B_{\pi(j)}$ for every~$j$.
\end{lemma}

\begin{proof}\leanok
    \uses{lem:ft_proportional_nt}
    This is the restricted normal-CF-BNT proportional comparison
    (Lemma~\ref{lem:ft_proportional_nt}).
    The proof uses split hypotheses.  Normal canonical form supplies
    irreducibility, trace-preserving normalization, primitivity, and the
    self-overlap limit via Theorem~\ref{thm:ncf_overlap_tendsto_one}.  The
    separation of distinct blocks is not derived from normal canonical form; it
    is the explicit no-phase-gauge-equivalence hypothesis in the statement.  The
    nondecaying mixed overlap is then converted to phase-gauge equivalence by
    Theorem~\ref{thm:nt_modulus_one_gauge}.
\end{proof}

\section{Restricted Proportional Comparisons}%
\label{sec:ft_restricted_proportional}

\begin{lemma}[Proportional comparison for restricted CF-BNT hypotheses]\label{lem:ft_proportional}
	\lean{MPSTensor.fundamentalTheorem_proportionalMPV_CFBNT}
	\leanok
	\uses{def:cf_bnt, def:gauge_phase_equiv}
	Let $A_{\mathrm{tot}}$ and $B_{\mathrm{tot}}$ be MPS tensors
	in canonical form with BNT separation, with associated families
	$(\mu_j^A, A_j)_{j=1}^{g_A}$ and $(\mu_k^B, B_k)_{k=1}^{g_B}$.
	Suppose there exist coefficient arrays $a_{N,j}$, $b_{N,k}$ and
	nonzero limits $a_j^\infty$, $b_k^\infty$, $c^\infty$ such that
	\[
		V^{(N)}(A_{\mathrm{tot}})_\sigma
		= \sum_{j=1}^{g_A} a_{N,j} V^{(N)}(A_j)_\sigma,
		\qquad
		V^{(N)}(B_{\mathrm{tot}})_\sigma
		= \sum_{k=1}^{g_B} b_{N,k} V^{(N)}(B_k)_\sigma,
	\]
	with $a_{N,j} \to a_j^\infty \neq 0$ and
	$b_{N,k} \to b_k^\infty \neq 0$, and suppose also that
	\[
		V^{(N)}(A_{\mathrm{tot}})_\sigma
		= c_N V^{(N)}(B_{\mathrm{tot}})_\sigma
	\]
	for a sequence $c_N \to c^\infty \neq 0$.
	Then $g_A = g_B$, and there is a permutation $\pi$ such that
	$D_j^A = D_{\pi(j)}^B$ and $A_j$ is gauge-phase equivalent to
	$B_{\pi(j)}$ for each~$j$.
\end{lemma}

\begin{proof}
	\leanok
	\uses{thm:cross_overlap_zero, thm:bnt_perm_prop}
	Apply Theorem~\ref{thm:bnt_perm_prop} to the two canonical-form
	decompositions with BNT separation.
	This yields the permutation $\pi$ that reindexes the separated sectors
	before comparing them one by one.
	These hypotheses provide injectivity, TP normalization,
	diagonal overlap limits, and off-diagonal decay via
	Theorem~\ref{thm:cross_overlap_zero}.
	With the supplied decomposition identities and convergence
	hypotheses, Theorem~\ref{thm:bnt_perm_prop} yields the equality of
	block counts, the permutation, and the per-block gauge-phase
	equivalence.
\end{proof}

\begin{remark}[Coefficient arrays for canonical-form decompositions]
	\label{rem:cf_coeff_arrays}
	In applications the coefficient arrays in
	Lemma~\ref{lem:ft_proportional} come from the block-diagonal
	decomposition itself. If
	\[
		A_{\mathrm{tot}} = \bigoplus_{j=1}^{g_A} \mu_j^A A_j,
		\qquad
		B_{\mathrm{tot}} = \bigoplus_{k=1}^{g_B} \mu_k^B B_k,
	\]
	then Theorem~\ref{thm:mpv_decomposition} gives, in the one-copy
	strict-weight case,
	\[
		V^{(N)}(A_{\mathrm{tot}})_\sigma
		= \sum_{j=1}^{g_A} (\mu_j^A)^N V^{(N)}(A_j)_\sigma,
		\qquad
		V^{(N)}(B_{\mathrm{tot}})_\sigma
		= \sum_{k=1}^{g_B} (\mu_k^B)^N V^{(N)}(B_k)_\sigma.
	\]
	Thus one takes $a_{N,j} = (\mu_j^A)^N$ and
	$b_{N,k} = (\mu_k^B)^N$ in that specialization. In the
	basis-of-normal-tensors theorem of \cite{Cirac2016MPDO_arXiv}, where one
	allows repeated copies of the same basis tensor, the coefficient attached to
	a basis sector is instead a repeated-copy power sum
	$c_{S,j}^{(N)}=\sum_q \mu_{j,q}^N$, exactly as in the BNT decomposition
	of \cite{Cirac2016MPDO_arXiv}. The sequence
	$c_N$ is the global proportionality factor from the hypothesis
	$V^{(N)}(A_{\mathrm{tot}}) = c_N V^{(N)}(B_{\mathrm{tot}})$.
	Lemma~\ref{lem:ft_proportional} keeps these arrays explicit because
	Theorem~\ref{thm:bnt_perm_prop} is formulated in that abstract form.
\end{remark}

\begin{lemma}[Proportional comparison for restricted normal-CF-BNT hypotheses]\label{lem:ft_proportional_nt}
		\lean{MPSTensor.fundamentalTheorem_proportionalMPV_normalCFBNT}\leanok
		\uses{def:normal_cf_bnt, def:gauge_phase_equiv}
		Let $A_{\mathrm{tot}}$ and $B_{\mathrm{tot}}$ be MPS tensors
		in normal canonical form with BNT separation, with associated families
		$(\mu_j^A, A_j)_{j=1}^{g_A}$ and $(\mu_k^B, B_k)_{k=1}^{g_B}$.
		Suppose there exist coefficient arrays $a_{N,j}$, $b_{N,k}$ and
		nonzero limits $a_j^\infty$, $b_k^\infty$, $c^\infty$ such that
		\[
			V^{(N)}(A_{\mathrm{tot}})_\sigma
			= \sum_{j=1}^{g_A} a_{N,j} V^{(N)}(A_j)_\sigma,
			\qquad
			V^{(N)}(B_{\mathrm{tot}})_\sigma
			= \sum_{k=1}^{g_B} b_{N,k} V^{(N)}(B_k)_\sigma,
		\]
		with $a_{N,j} \to a_j^\infty \neq 0$ and
		$b_{N,k} \to b_k^\infty \neq 0$, and suppose also that
		\[
			V^{(N)}(A_{\mathrm{tot}})_\sigma
			= c_N V^{(N)}(B_{\mathrm{tot}})_\sigma
		\]
		for a sequence $c_N \to c^\infty \neq 0$.
		Then $g_A = g_B$, and there is a permutation $\pi$ such that
		$D_j^A = D_{\pi(j)}^B$ and $A_j$ is gauge-phase equivalent to
		$B_{\pi(j)}$ for each~$j$.
\end{lemma}

\begin{proof}\leanok
		\uses{thm:ncf_overlap_tendsto_one, thm:nt_modulus_one_gauge, rem:normalcf_bnt_gap}
		Normal canonical form with BNT separation provides irreducibility,
		TP normalization, and primitive self-overlap convergence via
		Theorem~\ref{thm:ncf_overlap_tendsto_one}.
		Theorem~\ref{thm:nt_modulus_one_gauge} supplies the leading-block
		identification step: a nondecaying mixed overlap forces the
		corresponding comparison block to be gauge-phase equivalent.
		The same peeling argument as in the block-injective case therefore
		yields the block permutation and gauge-phase equivalence.
		Remark~\ref{rem:normalcf_bnt_gap} explains why the argument stops at
		this separated normal-canonical level.
\end{proof}

\section{Equal-MPV Comparisons}%
\label{sec:ft_equal_mpv_comparisons}

The source corollary for equal MPVs says that if two canonical-form tensors
generate the same MPV family for all system sizes, then their matrix dimensions
coincide and the full tensors are related by one invertible gauge
\cite[Corollary~II.2]{Cirac2017MPDO_AnnPhys}. In the proof, this corollary is
obtained after the proportional Fundamental Theorem by matching the BNT blocks,
then comparing the sector-weight power sums.

We record the comparison statements in the two forms needed below. The
restricted heterogeneous equal-case lemma, stated as
Lemma~\ref{lem:fundamentalTheorem_equalMPV_CFBNT_hetero}, gives the
block-count, bond-dimension, and gauge-phase matching directly from equality of
the block-diagonal MPV families. The common-structure lemma below proves the
global gauge statement when the two sides already share the same block structure.
The general CPSV theorem still passes through sector decompositions, copy-count
recovery, and sector-weight multiset comparison before one obtains a global
block-diagonal gauge.

\begin{lemma}[Common-block equal-MPV comparison]\label{lem:ft_equal_same_structure}
	\lean{MPSTensor.fundamentalTheorem_equalMPV_CFBNT}
	\leanok
	\uses{def:cf_bnt, def:gauge_equiv}
	Fix a common block structure: the same number of blocks~$g$,
	the same bond dimensions $(D_k)_{k=1}^g$, and the same weights
	$(\mu_k)_{k=1}^g$.
	Let $(\mu_k, A_k)_{k=1}^g$ and $(\mu_k, B_k)_{k=1}^g$ be two families in
	canonical form with BNT separation on this common structure.
	If the block-diagonal tensors
	$\bigoplus_k \mu_k A_k$ and $\bigoplus_k \mu_k B_k$
	generate the same MPV family, then each pair $(A_k, B_k)$ is
	gauge equivalent, and hence the corresponding block-diagonal tensors
	are gauge equivalent.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:ft_canonical}
	Since both families share the canonical-form hypotheses, the BNT separation
	condition is not needed.
	The same-structure canonical-form comparison
	(Lemma~\ref{lem:ft_canonical}) then applies directly to this
	common $(g, D_k, \mu_k)$ structure, yielding per-block gauge
	equivalence and hence global gauge equivalence.
\end{proof}

\subsection{Sector multiplicities and weight recovery}%
\label{sec:bnt_power_sum_recovery}

This subsection records the coefficient comparison used when several
canonical-form blocks are phase and gauge transforms of the same basis
normal tensor.  If the $j$th basis block occurs with nonzero weights
$\mu_{j,1},\dotsc,\mu_{j,r_j}$, its length-$N$ contribution is governed by
the power sum
\[
	c_{S,j}^{(N)}=\sum_{q=1}^{r_j}\mu_{j,q}^{N}.
\]
This is the scalar part of the source decomposition
\[
	A^i=\bigoplus_{j=1}^g\bigoplus_{q=1}^{r_j}
		\mu_{j,q}X_{j,q}A_j^iX_{j,q}^{-1}.
\]
The sector tensor used below is the representative form after the gauge
conjugations $X_{j,q}$ have already been absorbed into the choice of basis
block and gauge phases.  Thus BNT linear independence separates the coefficient arrays
$c_{S,j}^{(N)}$, and elementary power-sum algebra recovers both the
multiplicity $r_j$ and the multiset of weights attached to the basis block.
In the later PEPS application this information is not presented as a separate
chain of MPS comparison theorems: Theorem~IV.13 of~\cite{Cirac2017MPDO_AnnPhys} uses
structure coefficients
\[
	c_{\alpha,\beta,\gamma}^{(L)}
	=\operatorname{tr}\!\left(\chi_{\alpha,\beta,\gamma}^L\right),
	\qquad
	m_\gamma=\sum_{\alpha,\beta}c_{\alpha,\beta,\gamma}^{(1)}
		m_\alpha m_\beta,
\]
and the appendix writes the vertical canonical forms as
$\bigoplus_\alpha \mu_\alpha\otimes M_\alpha$ and
$\bigoplus_\beta \nu_\beta\otimes A_\beta$, with
$m_\alpha=\operatorname{tr}(\mu_\alpha)$ and
$n_\beta=\operatorname{tr}(\nu_\beta)$.
Here $\mu_\alpha$ and $\nu_\beta$ are diagonal positive matrices in the
source paper; only their traces are the scalar coefficients compared in this
subsection.
The one-index coefficients $c_{S,j}^{(N)}$ below are the corresponding
power-sum comparison after the basis blocks have already been matched.  The
geometric extrapolation and Newton--Girard steps are external finite-sequence
facts used to recover the lists of weights from those power sums.

\begin{definition}[Sector weights and sector decomposition]%
\label{def:paper_sector_data}
	\lean{MPSTensor.SectorWeightData, MPSTensor.SectorDecomposition}
	\leanok
	A sector-weight tuple over $g$ basis blocks consists of
	multiplicities $r_j > 0$, weights
	$\mu_{j,q} \neq 0$ for $q \in \{1,\dotsc,r_j\}$, and the
	\emph{sector coefficients}
	\[
		c_{S,j}^{(N)} := \sum_{q=1}^{r_j} \mu_{j,q}^N
	\]
	for the tuple $S$.
	A \emph{sector decomposition} consists of the basis blocks
	$(A_j)_{j=1}^g$ together with these multiplicities and weights.
\end{definition}

\begin{lemma}[Sector decomposition formula]%
\label{lem:paper_decomp_formula}
	\lean{MPSTensor.SectorDecomposition.mpv_toTensor_eq_sum_coeff}
	\leanok
	\uses{def:paper_sector_data}
	For a sector decomposition $S$ with associated tensor $A_{\mathrm{tot}}$,
	the MPV satisfies
	\[
		V^{(N)}(A_{\mathrm{tot}})_\sigma
		= \sum_{j=1}^{g} c_{S,j}^{(N)} V^{(N)}(A_j)_\sigma.
	\]
\end{lemma}

\begin{lemma}[Eventual coefficient comparison from BNT linear independence]%
\label{lem:coeff_eventually_eq}
	\lean{MPSTensor.SectorWeightData.coeff_eventually_eq_of_sameMPV}
	\leanok
	\uses{def:paper_sector_data, def:bnt}
	If two sector-weight tuples $S, T$ over the same basis, with coefficient arrays
	$c_{S,j}^{(N)}$ and $c_{T,j}^{(N)}$, produce equal total MPVs
	and the basis is eventually linearly independent (the BNT property),
	then there is $N_0$ such that
	\[
		\forall N>N_0,\ \forall j,\qquad
		c_{S,j}^{(N)}=c_{T,j}^{(N)}.
	\]
\end{lemma}

\begin{lemma}[Eventual coefficient equality recovers copy counts]%
\label{lem:copies_eq_of_eventually_coeff_eq}
	\lean{MPSTensor.SectorWeightData.copies_eq_of_eventually_coeff_eq}
	\leanok
	\uses{def:paper_sector_data}
	If two sector-weight tuples $S,T$ over the same basis have
	\[
		c_{S,j}^{(N)}=c_{T,j}^{(N)}
	\]
	for every basis block $j$ and all sufficiently large $N$, then
	$r_j^S=r_j^T$ for every~$j$.
\end{lemma}

\begin{proof}
	\leanok
	For a fixed basis sector \(j\), the coefficient equality is equality of
	the power sums of the two nonzero weight families for all sufficiently large
	exponents.  Moving one side to the other gives a finite signed sum of
	nonzero geometric sequences which vanishes eventually.  The resulting
	linear recurrence extends the identity to exponent \(0\), where the two
	power sums are \(r_j^S\) and \(r_j^T\).
\end{proof}

\begin{lemma}[BNT recovery of copy counts and sector weights]%
\label{lem:bnt_power_sum_recover_copies_weights}
	\lean{MPSTensor.SectorWeightData.exists_copies_eq_and_weight_multiset_eq_of_sameMPV_bnt}
	\leanok
	\uses{def:paper_sector_data, def:bnt}
	Let $S$ and $T$ be two sector-weight tuples over the same basis, without assuming
	their multiplicities agree. If the basis is eventually linearly
	independent and the total MPVs of $S$ and $T$ agree, then there are
	copy-count equalities $r_j^S = r_j^T$ for every basis block~$j$, and,
	after transporting along these equalities, the sector weight multisets
	are equal for every~$j$:
	\[
		\{\mu_{j,q}^{S}:1\le q\le r_j^S\}
		=
		\{\mu_{j,q}^{T}:1\le q\le r_j^T\}.
	\]
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:coeff_eventually_eq, lem:copies_eq_of_eventually_coeff_eq,
		thm:power_sum_multiset}
	BNT linear independence and equality of total MPVs give eventual
	coefficient equality by Lemma~\ref{lem:coeff_eventually_eq}.  Thus, for each
	basis sector \(j\), the two finite nonzero weight families satisfy
	\[
		\sum_{q=1}^{r_j^S}(\mu_{j,q}^S)^N
		=
		\sum_{q=1}^{r_j^T}(\mu_{j,q}^T)^N
	\]
	for all sufficiently large \(N\).  Lemma~\ref{lem:copies_eq_of_eventually_coeff_eq}
	applies the signed-geometric-sequence argument at exponent \(0\), so
	\(r_j^S=r_j^T\).  After transporting the right-hand index set along this
	copy-count equality, the identities for positive powers become equality of
	the power sums of two same-cardinality weight lists.  Newton--Girard
	(Theorem~\ref{thm:power_sum_multiset}) then recovers equality of the sector
	weight multisets.
\end{proof}

\begin{lemma}[Phase matching recovers copy counts and sector weights]%
\label{lem:phase_match_recover_copies_weights}
	\lean{MPSTensor.fundamentalTheorem_equalMPV_sectorDecomposition_hetero_of_phaseMatch_exists_copies}
	\leanok
	\uses{def:paper_sector_data}
	Let $P$ and $Q$ be sector decompositions. Assume there is a permutation
	$\pi$ matching the basis blocks and, for each $j$, a nonzero complex
	number $\zeta_j$ such that
	\[
		V^{(N)}(Q_{\pi(j)})_\sigma = \zeta_j^N V^{(N)}(P_j)_\sigma
	\]
	for every $N$ and $\sigma$. If the total MPV families of $P$ and $Q$
	agree and the basis blocks of $P$ are eventually linearly independent,
	then the copy counts satisfy $r_j^P=r_{\pi(j)}^Q$, and after multiplying
	the weights of $Q_{\pi(j)}$ by $\zeta_j$ the sector weight multisets agree.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:paper_decomp_formula, lem:bnt_power_sum_recover_copies_weights}
	Absorb the factors $\zeta_j$ into the sector weights of $Q$ and keep the
	basis blocks of $P$. The phase-matching hypothesis and the sector
	decomposition formula show that the resulting decomposition has the same
	total MPV family as~$Q$, hence also as~$P$. One is then in the
	shared-basis situation of Lemma~\ref{lem:bnt_power_sum_recover_copies_weights},
	which recovers both the copy-count equalities and the claimed equality of
	weight multisets.
\end{proof}

\begin{lemma}[Two-basis sector comparison after phase matching]%
\label{lem:phase_match_sector_weights}
	\lean{MPSTensor.fundamentalTheorem_equalMPV_sectorDecomposition_hetero_of_phaseMatch}
	\leanok
	\uses{def:paper_sector_data}
	Let $P$ and $Q$ be sector decompositions. Assume there is a permutation
	$\pi$ matching the basis blocks, the multiplicities satisfy
	$r_j^P = r_{\pi(j)}^Q$, and for each $j$ there is a nonzero complex number
	$\zeta_j$ such that
	\[
		V^{(N)}(Q_{\pi(j)})_\sigma = \zeta_j^N V^{(N)}(P_j)_\sigma
	\]
	for every $N$ and $\sigma$. If the total MPV families of $P$ and $Q$
	agree and the basis blocks of $P$ are eventually linearly independent,
	then for each $j$ the sector weight multiset of $P_j$ equals the
	multiset obtained from that of $Q_{\pi(j)}$ by multiplication by~$\zeta_j$.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:phase_match_recover_copies_weights}
	Lemma~\ref{lem:phase_match_recover_copies_weights} gives copy-count
	equalities and the corresponding weight-multiset equality. The recovered
	copy-count equalities coincide with the assumed ones, so the weight-multiset
	equality gives the conclusion.
\end{proof}

\begin{lemma}[Gauge-phase paired bases imply phase-matched sector comparison]%
\label{lem:gauge_phase_matched_basis_sector_weights}
	\lean{MPSTensor.fundamentalTheorem_equalMPV_sectorDecomposition_hetero_of_matched_basis}
	\leanok
	\uses{def:paper_sector_data, def:gauge_phase_equiv}
	Let $P$ and $Q$ be sector decompositions. Assume there is a permutation
	$\pi$ matching the basis blocks, the multiplicities satisfy
	$r_j^P = r_{\pi(j)}^Q$, and each matched pair of basis blocks is
	gauge-phase equivalent. If the total MPV families of $P$ and $Q$ agree
	and the basis blocks of $P$ are eventually linearly independent, then
	for each $j$ the sector weight multiset of $P_j$ equals the multiset
	obtained from that of $Q_{\pi(j)}$ after multiplication by the
	corresponding gauge phase.
\end{lemma}

\begin{proof}
	\leanok
	\uses{thm:gauge_phase_mpv_scaling, lem:phase_match_sector_weights}
	Gauge-phase equivalence gives the required phase-scaling identity for the
	MPVs by Theorem~\ref{thm:gauge_phase_mpv_scaling}. The conclusion then
	follows immediately from
	Lemma~\ref{lem:phase_match_sector_weights}.
\end{proof}

\begin{definition}[Sector basis matching witness]%
\label{def:sector_basis_matching}
	\lean{MPSTensor.SectorBasisMatching}
	\leanok
	\uses{def:paper_sector_data, def:gauge_phase_equiv}
	A \emph{sector basis matching} between two sector decompositions $P$ and $Q$
	consists of
	\begin{enumerate}
		\item a basis permutation (bijection)
		      $\pi : \{1,\dots,g_P\} \simeq \{1,\dots,g_Q\}$,
		\item per-block multiplicity equalities $r_j^P = r_{\pi(j)}^Q$,
		\item per-block bond-dimension equalities $D_j^P = D_{\pi(j)}^Q$, and
		\item per-block gauge-phase equivalences $P_j \simeq_{\mathrm{gp}} Q_{\pi(j)}$
		      after transporting $P_j$ across the bond-dimension equality.
	\end{enumerate}
	The hypotheses consumed by
	Lemma~\ref{lem:gauge_phase_matched_basis_sector_weights} are exactly the fields of
	this matching tuple.
\end{definition}

\begin{definition}[Sector basis pre-matching]%
\label{def:sector_basis_pre_matching}
	\lean{MPSTensor.SectorBasisPreMatching}
	\leanok
	\uses{def:paper_sector_data, def:gauge_phase_equiv}
	A \emph{sector basis pre-matching} between sector decompositions $P$
	and $Q$ consists of a permutation of their basis blocks, equality of the
	matched bond dimensions, and gauge-phase equivalence of each matched pair.
	It does not include the copy-count equality; that equality is recovered
	from total MPV equality and BNT linear independence in
	Lemma~\ref{lem:sector_basis_pre_matching_to_matching}.
\end{definition}

\begin{definition}[Single-family overlap-orthogonality hypotheses for sector bases]%
\label{def:sector_basis_overlap_ortho_hyps}
	\lean{MPSTensor.SectorBasisOverlapOrthoHypotheses}
	\leanok
	\uses{def:paper_sector_data, def:mpv_overlap}
	For one sector decomposition $P$, this condition states positive basis
	dimensions, trace-preserving normalization of all basis blocks,
	self-overlap limits equal to~$1$, and off-overlap limits equal to~$0$.
\end{definition}

\begin{definition}[Primitive overlap-span hypotheses for sector bases]%
\label{def:sector_basis_overlap_span_hyps}
	\lean{MPSTensor.SectorBasisOverlapSpanHypotheses}
	\leanok
	\uses{def:paper_sector_data, def:mpv_overlap, def:mpv_state,
		def:sector_basis_overlap_ortho_hyps}
	For sector decompositions $P$ and $Q$, this condition states the
	primitive overlap-rigidity hypotheses for their basis blocks: nonzero
	bond dimensions, injectivity, trace-preserving normalization,
	self-overlap limits equal to~$1$, off-overlap limits equal to~$0$, and
	equality of the finite-length spans of the basis MPV states for every
	system size.
\end{definition}

\begin{lemma}[Single-family overlap hypotheses plus span give overlap-span hypotheses]%
\label{lem:sector_basis_overlap_ortho_to_span}
	\lean{MPSTensor.SectorBasisOverlapOrthoHypotheses.to_overlapSpan}
	\leanok
	\uses{def:sector_basis_overlap_ortho_hyps, def:sector_basis_overlap_span_hyps}
	If two sector bases each satisfy the single-family overlap-orthogonality hypotheses,
	and one also supplies one-site injectivity of all their basis blocks together
	with equality of the finite-length MPV spans, then the two bases satisfy the
	primitive overlap-span hypotheses.
\end{lemma}

\begin{proof}
	\leanok
	\uses{def:sector_basis_overlap_ortho_hyps, def:sector_basis_overlap_span_hyps}
	This is just collecting the one-family fields on the left and right, and adding
	the remaining injectivity and span-comparison hypotheses.
\end{proof}

\begin{lemma}[A pre-matching gives a sector basis matching]%
\label{lem:sector_basis_pre_matching_to_matching}
	\lean{MPSTensor.SectorBasisPreMatching.exists_sectorBasisMatching_of_sameMPV}
	\leanok
	\uses{def:sector_basis_pre_matching, def:sector_basis_matching}
	Let $M$ be a sector basis pre-matching between $P$ and $Q$. If the total
	MPV families agree and the basis MPVs of $P$ are eventually linearly
	independent, then the copy-count equalities recovered by the phase-matched
	BNT coefficient comparison turn $M$ into a full sector basis matching.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:phase_match_recover_copies_weights}
	Gauge-phase equivalence gives the required phase-scaling identities for
	the corresponding basis MPVs:
	\[
		V^{(N)}(Q_{\pi(j)})_\sigma
		=
		\zeta_j^N V^{(N)}(P_j)_\sigma.
	\]
	Lemma~\ref{lem:phase_match_recover_copies_weights} then recovers the
	missing copy-count equalities, which are precisely the extra fields needed
	for Definition~\ref{def:sector_basis_matching}.
\end{proof}

\begin{theorem}[Overlap rigidity gives a sector basis matching]%
\label{thm:sector_basis_matching_from_overlap_ortho_span}
	\lean{MPSTensor.exists_sectorBasisMatching_of_overlapOrtho_span_sameMPV}
	\leanok
	\uses{def:paper_sector_data, def:sector_basis_matching, def:mpv_overlap}
	Assume the basis blocks of two sector decompositions have nonzero bond
	dimensions, are trace-preserving and injective, have asymptotically
	orthonormal overlaps within each family, and span the same MPV subspace at
	every system size. If the total MPV families also agree, then there exists a
	sector basis matching between the two decompositions.
\end{theorem}

\begin{proof}
	\leanok
	\uses{lem:bnt_perm_simple, lem:sector_basis_pre_matching_to_matching}
	The permutation-rigidity lemma for bases of normal tensors gives a
	permutation, the matched bond dimensions, and the gauge-phase equivalences
	of the basis blocks, hence a sector basis pre-matching. Asymptotic
	orthonormality gives the BNT linear-independence condition, and
	Lemma~\ref{lem:sector_basis_pre_matching_to_matching} recovers the
	copy-count equalities from equality of the total MPV families.
\end{proof}

\begin{lemma}[Primitive overlap-span hypotheses give a sector basis matching]%
\label{lem:sector_basis_overlap_span_to_matching}
	\lean{MPSTensor.SectorBasisOverlapSpanHypotheses.exists_sectorBasisMatching}
	\leanok
	\uses{def:sector_basis_overlap_span_hyps, def:sector_basis_matching}
	If two sector decompositions satisfy the primitive overlap-span hypotheses
	and their total MPV families agree, then there exists a sector basis matching
	between them.
\end{lemma}

\begin{proof}
	\leanok
	\uses{thm:sector_basis_matching_from_overlap_ortho_span}
	The collected hypotheses are exactly the nonzero-dimension, injectivity,
	normalization, overlap, and span assumptions of
	Theorem~\ref{thm:sector_basis_matching_from_overlap_ortho_span}. Applying
	that theorem gives the matching witness.
\end{proof}


\begin{theorem}[Span equality for representative sector bases]%
\label{thm:bnt_sector_decomp_pair_overlap_span_of_block_span}
	\lean{MPSTensor.exists_bnt_sectorDecomp_pair_with_overlapSpan_of_block_span_eq}
	\leanok
	\uses{def:sector_basis_overlap_span_hyps,
		thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_data_span}
	Let two nonzero-weight block families be trace-preserving, primitive, irreducible,
	injective, and carry nonzero weights. If their finite-length MPV spans agree
	for every system size, then there exist sector decompositions $P$ and $Q$
	whose total tensors have the same MPVs as the two weighted nonzero-block tensors,
	whose bases satisfy BNT linear independence, and whose sector bases satisfy
	the primitive overlap-span hypotheses.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_data_span}
	Apply the one-sided construction with the representative-span identity on both
	sides. The one-sided positive-dimension, injectivity, normalization, and
	overlap fields fill all fields of
	Definition~\ref{def:sector_basis_overlap_span_hyps} except the two-family
	span equality. That span equality follows by transporting the assumed
	nonzero-block span equality through the two representative-span identities.
\end{proof}

\begin{definition}[MPV phase equivalence for nonzero-weight blocks]
\label{def:mpv_block_phase_equiv}
	\lean{MPSTensor.MPVBlockPhaseEquiv}
	\leanok
	\uses{def:mpv}
	Two blocks $A$ and $B$ are MPV-phase equivalent if there is a nonzero
	scalar $\zeta$ such that, for every system size $N$ and every physical word,
	the MPV coefficient of $B$ equals $\zeta^N$ times the MPV coefficient of $A$.
	The two bond dimensions are allowed to differ.
\end{definition}

\begin{definition}[Common MPV phase cover]
\label{def:mpv_common_phase_cover}
	\lean{MPSTensor.MPVCommonPhaseCover}
	\leanok
	\uses{def:mpv_block_phase_equiv}
	For two finite nonzero-weight block families, a common MPV phase cover consists of a third
	finite family of blocks, a map from each nonzero-weight block family onto that common
	family, and MPV-phase equivalences between each block and its common
	representative. Both maps are required to be surjective, so the common family
	contains no unused representative.
\end{definition}

\begin{lemma}[Common MPV phase cover gives nonzero-block span equality]%
\label{lem:nonzero_block_span_eq_of_common_phase_cover}
	\lean{MPSTensor.mpv_span_eq_of_common_phase_cover}
	\leanok
	\uses{def:mpv_block_phase_equiv, def:mpv_state}
	Let two finite nonzero-weight block families map onto a third finite family of common
	blocks. Suppose both maps are surjective, and each block has finite-length
	MPV vectors equal to a nonzero scalar power times the MPV vectors of its common
	block. Then, for every system size, the two nonzero-weight block families span the same
	MPV subspace.
\end{lemma}

\begin{proof}
	\leanok
	Each nonzero-block MPV vector lies in the span of the common family because it is a
	scalar multiple of its common block's MPV vector. Surjectivity gives the reverse
	inclusion, so each nonzero-block span equals the common span; transitivity gives
	the equality of the two nonzero-block spans.
\end{proof}

\begin{lemma}[Common phase cover gives nonzero-block span equality]%
\label{lem:mpv_common_phase_cover_span_eq}
	\lean{MPSTensor.MPVCommonPhaseCover.span_eq}
	\leanok
	\uses{def:mpv_common_phase_cover}
	A common MPV phase cover for two nonzero-weight block families gives equality of their
	finite-length MPV spans at every system size.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:nonzero_block_span_eq_of_common_phase_cover}
	The common cover specifies the common family, the two surjective class maps,
	and the MPV-phase equivalences. The span lemma then applies directly.
\end{proof}

\begin{definition}[Common phase cover from a sector basis pre-matching]%
\label{def:sector_basis_pre_matching_common_phase_cover}
	\lean{MPSTensor.SectorBasisPreMatching.commonPhaseCover}
	\leanok
	\uses{def:sector_basis_pre_matching, def:mpv_common_phase_cover}
	A sector basis pre-matching between sector decompositions $P$ and $Q$
	canonically gives a common MPV phase cover of their basis block families.  The
	common family is the basis of $P$, the class map from $P$ is the identity, and
	the class map from $Q$ is the inverse of the pre-matching permutation.
\end{definition}

\begin{lemma}[Sector basis pre-matching gives basis-span equality]%
\label{lem:sector_basis_pre_matching_span_eq}
	\lean{MPSTensor.SectorBasisPreMatching.span_eq}
	\leanok
	\uses{def:sector_basis_pre_matching_common_phase_cover,
		lem:mpv_common_phase_cover_span_eq}
	A sector basis pre-matching between $P$ and $Q$ implies equality, for every
	system size, of the finite-length MPV spans of the two sector bases.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:mpv_common_phase_cover_span_eq}
	Apply the common phase-cover span lemma to the common phase cover constructed from
	the pre-matching.
\end{proof}

\begin{lemma}[Pre-matching supplies primitive overlap-span hypotheses]%
\label{lem:sector_basis_pre_matching_to_overlap_span}
	\lean{MPSTensor.SectorBasisPreMatching.to_overlapSpan}
	\leanok
	\uses{def:sector_basis_pre_matching, def:sector_basis_overlap_ortho_hyps,
		def:sector_basis_overlap_span_hyps}
	If two sector bases satisfy the one-family overlap-orthogonality hypotheses and their
	basis blocks are one-site injective, then a sector basis pre-matching between
	them establishes the finite-length span equality and hence the full primitive
	overlap-span hypotheses.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:sector_basis_pre_matching_span_eq,
		lem:sector_basis_overlap_ortho_to_span}
	The pre-matching gives the basis-span equality. Combining that equality with
	the one-family overlap hypotheses and injectivity is exactly the construction of
	Lemma~\ref{lem:sector_basis_overlap_ortho_to_span}.
\end{proof}

\begin{lemma}[Common phase cover from a BNT proportional-decomposition conclusion]%
\label{lem:common_phase_cover_of_proportional_decomposition}
	\lean{MPSTensor.nonempty_mpvCommonPhaseCover_of_proportionalDecompositionConclusion}
	\leanok
	\uses{def:mpv_common_phase_cover}
	If a BNT proportional-decomposition comparison has produced a permutation of
	the two basis families together with matched bond dimensions and gauge-phase
	equivalences, then the two basis families have a common MPV phase cover.
\end{lemma}

\begin{proof}
	\leanok
	\uses{def:mpv_block_phase_equiv}
	Take the left basis family as the common family. The comparison permutation and
	its inverse give the two class maps, while each gauge-phase equivalence gives
	the required MPV-phase equivalence.
\end{proof}

\begin{lemma}[Common phase cover from a bijective phase matching]%
\label{lem:common_phase_cover_of_equiv_phase}
	\lean{MPSTensor.nonempty_mpvCommonPhaseCover_of_equiv_phase}
	\leanok
	\uses{def:mpv_common_phase_cover, def:mpv_block_phase_equiv}
	If two block families are matched by a bijection and corresponding blocks have
	the same matrix product vectors up to a nonzero scalar power depending on the
	length, then the two families have a common MPV phase cover.
\end{lemma}

\begin{proof}
	\leanok
	Take the first block family as the common family. The bijection and its inverse
	give the two class maps, and the assumed phase match gives the second
	family's phase factors.
\end{proof}

\begin{lemma}[Span equality from a proportional-decomposition comparison]%
\label{lem:nonzero_block_span_eq_of_proportional_decomposition}
	\lean{MPSTensor.mpv_span_eq_of_proportionalDecompositionConclusion}
	\leanok
	\uses{lem:common_phase_cover_of_proportional_decomposition,
		lem:mpv_common_phase_cover_span_eq}
	A BNT proportional-decomposition comparison gives equality, for every system
	size, of the finite-length MPV spans of the two nonzero-weight block families.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:common_phase_cover_of_proportional_decomposition,
		lem:mpv_common_phase_cover_span_eq}
	First form the common MPV phase cover from the proportional-decomposition
	comparison. The common phase-cover span lemma then gives the stated span equality.
\end{proof}

\begin{lemma}[Span equality from a bijective phase matching]%
\label{lem:nonzero_block_span_eq_of_equiv_phase}
	\lean{MPSTensor.mpv_span_eq_of_equiv_phase}
	\leanok
	\uses{lem:common_phase_cover_of_equiv_phase,
		lem:mpv_common_phase_cover_span_eq}
	A bijective phase matching gives equality, for every system size, of the
	finite-length MPV spans of the two block families.
\end{lemma}

\begin{proof}
	\leanok
	First form the common MPV phase cover from the bijective phase matching. The
	common phase-cover span lemma then gives the stated span equality.
\end{proof}

\begin{theorem}[Overlap-span hypotheses from a common phase cover]%
\label{thm:bnt_sectorDecomp_pair_overlapSpan_of_commonPhaseCover}
	\lean{MPSTensor.exists_bnt_sectorDecomp_pair_with_overlapSpan_of_commonPhaseCover}
	\leanok
	\uses{def:mpv_common_phase_cover, def:sector_basis_overlap_span_hyps}
	Let two block families be trace-preserving, primitive, irreducible, injective,
	and carry nonzero weights. If they have a common MPV phase cover, then there
	exist BNT sector decompositions $P$ and $Q$ for
	the two weighted nonzero parts, and the sector bases satisfy the primitive
	overlap-span hypotheses.
\end{theorem}

\begin{proof}
	\leanok
	\uses{lem:mpv_common_phase_cover_span_eq,
		thm:bnt_sector_decomp_pair_overlap_span_of_block_span}
	The common cover gives finite-length span equality for the two nonzero-weight
	block families. The phase-class representative construction then transports
	this span equality to the BNT sector bases and supplies the remaining
	one-family overlap hypotheses.
\end{proof}

\begin{theorem}[Overlap-span hypotheses from proportional decomposition]%
\label{thm:bnt_sectorDecomp_pair_overlapSpan_of_proportionalDecompositionConclusion}
	\lean{MPSTensor.exists_bnt_sectorDecomp_pair_with_overlapSpan_of_proportionalDecompositionConclusion}
	\leanok
	\uses{def:sector_basis_overlap_span_hyps}
	Let two block families be trace-preserving, primitive, irreducible, injective,
	and carry nonzero weights. If a BNT proportional-decomposition comparison
	supplies a permutation, equal bond
	dimensions, and gauge-phase equivalences between the two families, then the
	phase-class representative BNT sector decompositions satisfy the primitive
	overlap-span hypotheses.
\end{theorem}

\begin{proof}
	\leanok
	\uses{lem:common_phase_cover_of_proportional_decomposition,
		thm:bnt_sectorDecomp_pair_overlapSpan_of_commonPhaseCover}
	First turn the proportional-decomposition comparison into a common MPV phase cover.
	The preceding theorem then converts that cover into the overlap-span hypotheses
	for the constructed sector bases.
\end{proof}

\begin{lemma}[Separated normal canonical-form hypotheses produce a common phase cover]%
\label{lem:common_phase_cover_of_separated_normal_bnt}
	\lean{MPSTensor.nonempty_mpvCommonPhaseCover_of_separated_normalCFBNT_data}
	\leanok
	\uses{def:mpv_common_phase_cover, def:is_normal_canonical_form,
		def:blocks_not_gauge_phase_equiv, thm:bnt_perm_prop_nt}
	For two normal canonical-form block families with no gauge-phase-equivalent
	distinct equal-dimension blocks, proportional decompositions whose coefficient
	limits are nonzero give a common MPV phase cover of the two basis families.
	This is the BNT construction of the common cover; it does not assume equality of
	finite-length spans.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:common_phase_cover_of_proportional_decomposition,
		thm:bnt_perm_prop_nt}
	The irreducible trace-preserving BNT permutation theorem gives the permutation
	matching the basis blocks, bond-dimension equalities, and gauge-phase
	equivalences. Applying the previous lemma to that conclusion yields the
	common phase cover.
\end{proof}

\begin{lemma}[Separated normal canonical-form hypotheses give nonzero-block span equality]%
\label{lem:separated_normal_bnt_span_eq}
	\lean{MPSTensor.mpv_span_eq_of_separated_normalCFBNT_data}
	\leanok
	\uses{lem:common_phase_cover_of_separated_normal_bnt,
		lem:mpv_common_phase_cover_span_eq}
	For two separated normal canonical-form block families as in
	Lemma~\ref{lem:common_phase_cover_of_separated_normal_bnt}, the BNT
	comparison implies equality, at every system size, of the finite-length
	MPV spans of the two basis families.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:common_phase_cover_of_separated_normal_bnt,
		lem:mpv_common_phase_cover_span_eq}
	The separated normal-form comparison first gives a common MPV phase cover.
	Applying the common phase-cover span lemma to that cover gives the span equality.
\end{proof}

\begin{theorem}[Sector comparison via a sector-basis matching]%
\label{thm:sector_matching_weight_multisets}
	\lean{MPSTensor.fundamentalTheorem_equalMPV_sectorDecomposition_hetero_of_sectorMatching}
	\leanok
	\uses{def:sector_basis_matching, lem:gauge_phase_matched_basis_sector_weights}
	Given two sector decompositions $P$ and $Q$, a sector basis matching
	(Definition~\ref{def:sector_basis_matching}) between them, eventual
	linear independence of the basis MPV states of $P$, and equality of the
	total MPV families of $P$ and $Q$, one obtains nonzero phases $(\zeta_j)_j$
	such that for each $j$ the sector weight multiset of $P_j$ equals the
	multiset obtained from that of $Q_{\pi(j)}$ by multiplication by~$\zeta_j$.
\end{theorem}

\begin{proof}
	\leanok
	\uses{lem:gauge_phase_matched_basis_sector_weights}
	The sector-basis matching contains exactly the permutation, copy, and
	gauge-phase hypotheses of Lemma~\ref{lem:gauge_phase_matched_basis_sector_weights},
	so the conclusion follows directly from that lemma.
\end{proof}

\begin{lemma}[Cancellation of leftover zero blocks]\label{lem:sameMPV2_live_of_zeroTail_eq}
	\lean{MPSTensor.sameMPV₂_live_of_sameMPV₂_with_zeroTail_eq}
	\leanok
	\uses{def:same_mpv2, def:mpv}
	Let $A$ and $B$ have the same MPV family, each decomposed as an all-zero
	leftover block plus a nonzero part. If the leftover zero-block dimensions
	agree, then the nonzero parts themselves have the same MPV family, including
	at length zero.
\end{lemma}

\begin{proof}
	\leanok
	\uses{thm:mpv_zero_tensor}
	At positive lengths the all-zero leftover block vanishes, so the nonzero
	parts agree. Theorem~\ref{thm:mpv_zero_tensor} gives the exact value
	\[
		V^{(0)}(\mathbf{0}_{d,D_0})_\varnothing=D_0,
	\]
	so equality of the two leftover zero-block dimensions gives cancellation
	at length zero as well.
\end{proof}

\begin{theorem}[Sector comparison from leftover zero blocks and span hypotheses]%
\label{thm:afterBlocking_sectorComparison_zeroTail_of_blockSpan}
	\lean{MPSTensor.afterBlocking_sectorComparison_zeroTail_of_blockSpan}
	\leanok
	\uses{lem:sameMPV2_live_of_zeroTail_eq,
		thm:bnt_sector_decomp_pair_overlap_span_of_block_span,
		lem:sector_basis_overlap_span_to_matching,
		thm:sector_matching_weight_multisets}
	Let $A$ and $B$ have the same MPV family. Fix $p>0$ and suppose $A^{[p]}$
	and $B^{[p]}$ each decompose as an all-zero leftover block plus a weighted
	nonzero part, with all blocks in the nonzero part trace-preserving,
	primitive, irreducible, injective, and with nonzero weights. If the two
	leftover zero-block dimensions agree and the two nonzero-weight block
	families have equal
	finite-length MPV spans at every system size, then there exist a period
	$p' > 0$ and sector decompositions $P,Q$ such that $A^{[p']}$ and
	$B^{[p']}$ agree with the corresponding sector tensors at every positive
	length, $P$ and $Q$ have the same full MPV family, both have BNT bases, and
	their sector weights satisfy the matched sector-weight conclusion.
\end{theorem}

\begin{proof}
	\leanok
	\uses{lem:sameMPV2_live_of_zeroTail_eq,
		thm:bnt_sector_decomp_pair_overlap_span_of_block_span,
		lem:sector_basis_overlap_span_to_matching,
		thm:sector_matching_weight_multisets}
	The finite-length span equality gives the primitive overlap-span hypotheses
	for the BNT sector bases of the two nonzero parts. Cancellation of leftover
	zero blocks identifies the two nonzero parts as full MPV families, while the
	leftover zero-block summands vanish at positive lengths, giving the
	positive-length comparison of $A^{[p]}$ and $B^{[p]}$ with the two sector
	decompositions. The sector-basis matching and two-basis comparison theorems
	then give the sector-weight conclusion.
\end{proof}

The following common-primitive hypotheses are not separate statements in
arXiv:1606.00608.  They name the extra formal assumptions used here before the
zero-block sector comparison: span equality, phase-cover hypotheses, and
BNT comparison hypotheses for the common primitive sectors.

\begin{definition}[Span hypotheses for common primitive sectors]%
\label{def:common_primitive_span_hypotheses}
	\lean{MPSTensor.CommonPrimitiveSpanHypotheses}
	\leanok
	\uses{def:mpv_state, def:injective}
	For two common primitive nonzero-sector families at one blocking length, this
	condition consists of the hypotheses needed before applying the
	zero-block sector comparison: equality of the two leftover zero-block
	dimensions, one-site injectivity of all blocks on both sides, and equality
	of the finite-length MPV spans of the two block families for every system
	size.
\end{definition}

\begin{theorem}[Phase covers supply span hypotheses]%
\label{thm:common_primitive_span_hypotheses_of_common_phase_cover}
	\lean{MPSTensor.CommonPrimitiveSpanHypotheses.of_commonPhaseCover}
	\leanok
	\uses{def:common_primitive_span_hypotheses, def:mpv_common_phase_cover,
		lem:mpv_common_phase_cover_span_eq}
	If the two common primitive nonzero-sector families have equal leftover
	zero-block dimensions, are one-site injective, and are equipped with a common MPV phase cover, then
	they satisfy the span hypotheses of
	Definition~\ref{def:common_primitive_span_hypotheses}.
\end{theorem}

\begin{proof}
	\leanok
	\uses{lem:mpv_common_phase_cover_span_eq}
	The zero-block dimension equality and injectivity fields are exactly the supplied hypotheses.  The
	common phase cover gives the finite-length span equality by
	Lemma~\ref{lem:mpv_common_phase_cover_span_eq}.
\end{proof}

\begin{definition}[Phase-cover hypotheses for common primitive sectors]%
\label{def:common_primitive_phase_cover_hypotheses}
	\lean{MPSTensor.CommonPrimitivePhaseCoverHypotheses}
	\leanok
	\uses{def:mpv_common_phase_cover, def:injective}
	For two common primitive nonzero-sector families at one blocking length, this
	condition is the phase-cover analogue of
	Definition~\ref{def:common_primitive_span_hypotheses}: equality of the two
	leftover zero-block dimensions, one-site injectivity of all blocks on both sides, and a
	common MPV phase cover for the two block families.
\end{definition}

\begin{theorem}[Phase-cover hypotheses imply span hypotheses]%
\label{thm:common_primitive_phase_cover_hypotheses_to_span_hypotheses}
	\lean{MPSTensor.CommonPrimitivePhaseCoverHypotheses.toSpanHypotheses}
	\leanok
	\uses{def:common_primitive_phase_cover_hypotheses,
		def:common_primitive_span_hypotheses}
	If two common primitive nonzero-sector families satisfy the phase-cover
	hypotheses of
	Definition~\ref{def:common_primitive_phase_cover_hypotheses}, then they
	satisfy the span hypotheses of
	Definition~\ref{def:common_primitive_span_hypotheses}.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:common_primitive_span_hypotheses_of_common_phase_cover}
	The zero-block dimension equality and the two injectivity assertions are already part of
	the phase-cover hypotheses.  Apply
	Theorem~\ref{thm:common_primitive_span_hypotheses_of_common_phase_cover} to a
	common MPV phase cover for the two block families.
\end{proof}

\begin{definition}[Proportional-decomposition hypotheses for common primitive sectors]%
\label{def:common_primitive_proportional_hypotheses}
	\lean{MPSTensor.CommonPrimitiveProportionalHypotheses}
	\leanok
	\uses{def:common_primitive_phase_cover_hypotheses}
	For two common primitive nonzero-sector families at one blocking length, this
	condition is the proportional-decomposition analogue of
	Definition~\ref{def:common_primitive_phase_cover_hypotheses}: equality of the two
	leftover zero-block dimensions, one-site injectivity of all blocks on both sides, and a
	BNT proportional-decomposition conclusion for the two block families
	(a permutation, equal bond dimensions, and gauge-phase equivalences).
\end{definition}

\begin{theorem}[Proportional-decomposition hypotheses imply phase-cover hypotheses]%
\label{thm:common_primitive_proportional_hypotheses_to_phase_cover_hypotheses}
	\lean{MPSTensor.CommonPrimitiveProportionalHypotheses.toPhaseCoverHypotheses}
	\leanok
	\uses{def:common_primitive_proportional_hypotheses,
		def:common_primitive_phase_cover_hypotheses}
	If the proportional-decomposition hypotheses of
	Definition~\ref{def:common_primitive_proportional_hypotheses} hold, then the
	phase-cover hypotheses of
	Definition~\ref{def:common_primitive_phase_cover_hypotheses} hold.
\end{theorem}

\begin{proof}
	\leanok
	\uses{lem:common_phase_cover_of_proportional_decomposition}
	The zero-block dimension equality and injectivity assumptions transfer directly.
	The common MPV phase cover is provided by the general proportional-decomposition
	lemma applied to the two block families.
\end{proof}

\begin{theorem}[Proportional-decomposition hypotheses imply span hypotheses]%
\label{thm:common_primitive_proportional_hypotheses_to_span_hypotheses}
	\lean{MPSTensor.CommonPrimitiveProportionalHypotheses.toSpanHypotheses}
	\leanok
	\uses{thm:common_primitive_proportional_hypotheses_to_phase_cover_hypotheses,
		def:common_primitive_span_hypotheses}
	If the proportional-decomposition hypotheses hold, then the span hypotheses
	hold by passing through the phase-cover hypotheses.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:common_primitive_proportional_hypotheses_to_phase_cover_hypotheses,
		thm:common_primitive_phase_cover_hypotheses_to_span_hypotheses}
	The proportional hypotheses give the phase-cover hypotheses via the preceding
	theorem, and the phase-cover hypotheses imply the span hypotheses.
\end{proof}

\begin{definition}[BNT matching hypotheses for common primitive sectors]%
\label{def:common_primitive_bnt_cover_hypotheses}
	\lean{MPSTensor.CommonPrimitiveBNTCoverHypotheses,
		MPSTensor.CommonPrimitiveBNTCoverHypotheses.ofNormalCanonicalFormBNT,
		MPSTensor.CommonPrimitiveBNTCoverHypotheses.ofNormalCanonicalFormBNT_zeroTailIdentity,
		MPSTensor.CommonPrimitiveBNTCoverHypotheses.ofCommonRepresentatives_zeroTailIdentity,
		MPSTensor.CommonPrimitiveBNTCoverHypotheses.ofCommonRepresentativeBNTCoverHypotheses,
		MPSTensor.zeroTail_eq_of_proportionalDecompositionConclusion}
	\leanok
	\uses{def:is_normal_canonical_form, def:blocks_not_gauge_phase_equiv,
		def:injective, thm:ncf_sorted_tp_primitive_irred}
	For two common primitive nonzero-sector families at one blocking length, this
	condition states the hypotheses needed to apply the BNT comparison theorem:
	normal canonical form on both sides, no phase-gauge-equivalent distinct
	blocks in either family, equality \(z_A=z_B\) of the leftover all-zero block
	dimensions, one-site injectivity, and proportionality of the two block MPV
	families.  Its output is the source-form matching
	\(B_{\pi(j)}^i=e^{i\phi_j}X_jA_j^iX_j^{-1}\), hence
	\(\ket{V^{(N)}(B_{\pi(j)})}=e^{iN\phi_j}\ket{V^{(N)}(A_j)}\).

	% Maintainer note: the structure name contains "cover" because the next
	% step constructs this matching as a common MPV phase cover.  This wording is
	% not part of the source terminology and is intentionally not displayed.
\end{definition}

\begin{definition}[BNT-cover inputs for representative common-sector families]%
\label{def:common_representative_bnt_cover_hypotheses}
	\lean{MPSTensor.CommonRepresentativeBNTCoverHypotheses}
	\leanok
	\uses{def:common_primitive_bnt_cover_hypotheses,
		def:common_sector_relabeling_hypothesis}
	For representative common-sector families after choosing one representative
	per original nonzero block, this structure collects the remaining hypotheses
	needed for the BNT comparison: nonzero weights, strict weight ordering by norm,
	BNT separation (no gauge-phase-equivalent distinct representatives), the
	length-zero identity of the representative nonzero parts, one-site
	injectivity of each representative block, and proportional-decomposition
	data linking the two representative families.
\end{definition}

\begin{theorem}[BNT matching hypotheses produce a common phase cover]%
\label{thm:common_primitive_bnt_cover_hypotheses_to_common_phase_cover}
	\lean{MPSTensor.CommonPrimitiveBNTCoverHypotheses.toMPVCommonPhaseCover}
	\leanok
	\uses{def:common_primitive_bnt_cover_hypotheses,
		def:mpv_common_phase_cover}
	If two common primitive nonzero-sector families satisfy the BNT matching
	hypotheses, then they have a common MPV phase cover.
\end{theorem}

\begin{proof}
	\leanok
	\uses{lem:common_phase_cover_of_separated_normal_bnt}
	The normal canonical-form hypotheses, gauge-phase separation, and proportional
	decomposition are exactly the hypotheses of
	Lemma~\ref{lem:common_phase_cover_of_separated_normal_bnt}.
\end{proof}

\begin{theorem}[BNT matching hypotheses imply phase-cover hypotheses]%
\label{thm:common_primitive_bnt_cover_hypotheses_to_phase_cover_hypotheses}
	\lean{MPSTensor.CommonPrimitiveBNTCoverHypotheses.toCommonPrimitivePhaseCoverHypotheses}
	\leanok
	\uses{def:common_primitive_bnt_cover_hypotheses,
		def:common_primitive_phase_cover_hypotheses}
	If two common primitive nonzero-sector families satisfy the BNT matching
	hypotheses, then they satisfy the phase-cover hypotheses of
	Definition~\ref{def:common_primitive_phase_cover_hypotheses}.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:common_primitive_bnt_cover_hypotheses_to_common_phase_cover}
	The zero-block dimension equality and injectivity assumptions are part of the
	BNT matching hypotheses. The preceding theorem supplies the common MPV phase cover.
\end{proof}

\begin{theorem}[Common primitive sectors with span hypotheses]%
\label{thm:afterBlocking_sectorComparison_zeroTail_of_reindexedNonzeroParts_spanHypotheses}
	\lean{MPSTensor.afterBlocking_sectorComparison_zeroTail_of_reindexedNonzeroParts_spanHypotheses}
	\leanok
	\uses{thm:after_blocking_common_primitive_irreducible_blocks_reindexed,
		def:common_primitive_span_hypotheses,
		thm:afterBlocking_sectorComparison_zeroTail_of_blockSpan}
	Assume $A$ and $B$ have the same MPV family and that the canonically blocked
	nonzero parts agree with the same families written after identifying
	blocked physical words.  The common primitive theorem then produces one
	positive blocking length with
	\(A^{[p]}\sim\mathbf{0}_{d^p,z_A}\oplus A_{\mathrm{nz}}\) and
	\(B^{[p]}\sim\mathbf{0}_{d^p,z_B}\oplus B_{\mathrm{nz}}\), and
	trace-preserving primitive irreducible nonzero-sector families with positive
	bond dimensions and nonzero weights.  If these produced families also satisfy
	Definition~\ref{def:common_primitive_span_hypotheses}, then the leftover
	zero-block sector comparison conclusion holds: after a positive blocking,
	there are sector decompositions $P,Q$ agreeing with $A$ and $B$ at all
	positive lengths, with $P$ and $Q$ generating the same full MPV family, with
	BNT bases, and with their sector weights matched up to a permutation and
	nonzero phases.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:after_blocking_common_primitive_irreducible_blocks_reindexed,
		thm:afterBlocking_sectorComparison_zeroTail_of_blockSpan}
	Apply the common primitive theorem to obtain the nonzero-sector families.
	The span hypotheses provide \(z_A=z_B\), one-site injectivity, and
	finite-length nonzero-sector span equality.  These are exactly the additional
	hypotheses required by the leftover-zero-block comparison theorem.
\end{proof}

\begin{theorem}[Common primitive irreducible sectors with conditional block-span consequences]%
\label{thm:afterBlocking_commonSector_blockSpan_of_reindexedNonzeroParts}
	\lean{MPSTensor.afterBlocking_commonSector_blockSpan_of_reindexedNonzeroParts}
	\leanok
	\uses{thm:after_blocking_common_primitive_irreducible_blocks_reindexed,
		lem:mpv_common_phase_cover_span_eq}
	Let $A$ and $B$ have the same MPV family and assume the blocked-word
	identification for the common-sector families. The common primitive
	irreducible-block theorem supplies one positive blocking length and
	trace-preserving, primitive, irreducible nonzero-sector families on both sides.
	This theorem records that, for exactly those families, either a common MPV
	phase cover or a BNT proportional-decomposition conclusion supplies the
	finite-length block-span equality required by
	Theorem~\ref{thm:afterBlocking_sectorComparison_zeroTail_of_blockSpan}.
	Thus the remaining mathematical inputs -- the blocked-word identification
	and the later common-phase or BNT matching comparison -- are kept explicit.
\end{theorem}

\begin{definition}[Blocked-word identification for common sectors]%
\label{def:common_sector_relabeling_hypothesis}
	\lean{MPSTensor.CommonSectorRelabelingHypothesis}
	\leanok
	This is the one-sided assertion that the canonically blocked weighted nonzero
	part agrees, as an MPV family, with the same common cyclic-sector family after
	identifying blocked physical words. It isolates the coordinate equality between
	direct and iterated blocking. The
	coordinate-grouping theorem below supplies this condition in the theorems that
	take it as a hypothesis, so it is not a new paper-level mathematical hypothesis.
\end{definition}

\begin{definition}[Global coordinate-grouping hypothesis for common sectors]%
\label{def:common_grouped_block_cast_hypothesis}
	\lean{MPSTensor.CommonGroupedBlockCastHypothesis}
	\leanok
	\uses{def:common_blocked_cyclic_sector_grouped_block_cast}
	This is the corresponding global assertion at the level of blocked words: every
	common blocked cyclic-sector family satisfies the coordinate-grouping condition
	for each original nonzero block.  Equivalently, the canonical identification with
	each iterated blocked alphabet agrees with grouping the direct blocked word into
	consecutive words of the period-removal length.  The next lemma proves this
	assertion from the concrete word-cast construction.
\end{definition}

	\begin{lemma}[Canonical coordinate grouping]%
	\label{lem:common_grouped_block_cast_of_flatten_word}
		\lean{MPSTensor.CommonGroupedBlockCastHypothesis.of_flattenWordOfBlock_cast_eq}
		\leanok
		\uses{def:common_grouped_block_cast_hypothesis,
			lem:common_blocked_cyclic_sector_flatten_word_of_block_cast}
		The canonical coordinate grouping holds for every common blocked
		cyclic-sector family.  The proof identifies each grouped block with the
		corresponding direct blocked word.
	\end{lemma}

	\begin{proof}\leanok
		\uses{lem:common_blocked_cyclic_sector_flatten_word_of_block_cast}
		Apply the grouped-block cast comparison to each original nonzero block.
	\end{proof}

	\begin{lemma}[Coordinate grouping gives the word identification]%
\label{lem:common_grouped_block_cast_to_relabeling}
		\lean{MPSTensor.CommonGroupedBlockCastHypothesis.toRelabelingHypothesis}
	\leanok
	\uses{def:common_grouped_block_cast_hypothesis,
		def:common_sector_relabeling_hypothesis,
		lem:common_blocked_cyclic_sector_blocked_word_comparison}
		If the global coordinate-grouping hypothesis holds, then the blocked-word
		identification used by the common-sector structural theorem holds.  The
	proof is exactly the weighted finite-sum comparison for direct and iterated
	blocking applied to each common cyclic-sector family.
\end{lemma}

\begin{proof}
	\leanok
	\uses{lem:common_blocked_cyclic_sector_blocked_word_comparison}
	Apply the coordinate-grouping comparison to each one-sided family and sum over
	the original nonzero blocks with the transported powers of the weights.
\end{proof}

\begin{theorem}[Unconditional common primitive block decompositions]%
\label{thm:unconditional_common_primitive_irreducible_blocks}
	\lean{MPSTensor.unconditional_commonPrimitiveIrreducibleBlocks}
	\leanok
	\uses{thm:after_blocking_common_primitive_irreducible_blocks_reindexed}
	Let $A$ and $B$ have the same MPV family.  Then there is a positive blocking
	length at which both blocked tensors decompose into a leftover all-zero
	block and a weighted direct sum of trace-preserving, primitive,
	tensor-irreducible blocks with positive bond dimensions and nonzero weights.
	The two nonzero-sector families have the same positive-length MPV family,
	and the leftover zero blocks satisfy the length-zero identity.
\end{theorem}

\begin{proof}
	\leanok
	\uses{lem:common_blocked_cyclic_sector_flatten_word_of_block_cast,
		lem:common_grouped_block_cast_to_relabeling,
		thm:after_blocking_common_primitive_irreducible_blocks_reindexed}
	The direct digit computation gives the global coordinate-grouping hypothesis.
	This implies the blocked-word identification for common sectors.  Applying the
	common primitive decomposition theorem with this identification gives the
	unconditional decomposition.
\end{proof}

\begin{theorem}[Common sectors from phase covers]%
\label{thm:afterBlocking_sectorComparison_reindexed_commonPhaseCover}
	\lean{MPSTensor.afterBlocking_sectorComparison_zeroTail_of_reindexedNonzeroParts_commonPhaseCover}
	\leanok
	\uses{def:common_primitive_phase_cover_hypotheses,
		def:common_sector_relabeling_hypothesis}
	Let $A$ and $B$ have the same MPV family.  Assume the blocked-word
	identification for the common-sector families.  The common-sector structural
	theorem then produces a common blocking length with
	\(A^{[p]}\sim\mathbf{0}_{d^p,z_A}\oplus A_{\mathrm{nz}}\) and
	\(B^{[p]}\sim\mathbf{0}_{d^p,z_B}\oplus B_{\mathrm{nz}}\), and
	trace-preserving, primitive, irreducible nonzero-sector families on both sides. If those
	families satisfy the phase-cover hypotheses of
	Definition~\ref{def:common_primitive_phase_cover_hypotheses}, then the sector
	decompositions obtained from the two sides have BNT bases, agree as full MPV
	families, and their sector weights satisfy the matched sector-weight conclusion.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:afterBlocking_sectorComparison_zeroTail_of_reindexedNonzeroParts_spanHypotheses,
		thm:common_primitive_phase_cover_hypotheses_to_span_hypotheses}
	Use Theorem~\ref{thm:common_primitive_phase_cover_hypotheses_to_span_hypotheses}
		to form the span hypotheses of
	Definition~\ref{def:common_primitive_span_hypotheses}.  Then apply the
	common-sector comparison theorem with those span hypotheses.
\end{proof}

\begin{theorem}[Sector comparison from relabeled common sectors and proportional-decomposition data]%
\label{thm:afterBlocking_sectorComparison_zeroTail_of_reindexedNonzeroParts_proportional}
	\lean{MPSTensor.afterBlocking_sectorComparison_zeroTail_of_reindexedNonzeroParts_proportional}
	\leanok
	\uses{def:common_primitive_proportional_hypotheses,
		def:common_sector_relabeling_hypothesis,
		thm:afterBlocking_sectorComparison_reindexed_commonPhaseCover}
	Assume the blocked-word identification for the common-sector families.
	The structural theorem supplies one common positive blocking length with
	trace-preserving, primitive, irreducible nonzero-sector families on both
	sides. If the remaining comparison assertions for exactly those families
	are available -- equality of the two zero-tail dimensions, injectivity at
	that blocking level, and a BNT proportional-decomposition comparison --
	then the common-sector families satisfy the proportional-decomposition
	hypotheses, which convert to common phase-cover hypotheses and give the
	sector-weight comparison conclusion.
\end{theorem}

\begin{theorem}[Common sectors from BNT matching hypotheses]%
\label{thm:afterBlocking_sectorComparison_reindexed_bntCover}
	\lean{MPSTensor.afterBlocking_sectorComparison_zeroTail_of_reindexedNonzeroParts_bntCover}
	\leanok
	\uses{def:common_primitive_bnt_cover_hypotheses,
		def:common_sector_relabeling_hypothesis}
	Let $A$ and $B$ have the same MPV family. Assume the blocked-word
	identification for the common-sector families.  The common-sector structural
	theorem gives common primitive nonzero-sector families. If those families
	satisfy the BNT matching hypotheses, then the same sector-weight comparison
	conclusion holds.
\end{theorem}

\begin{proof}
	\leanok
	\uses{thm:common_primitive_bnt_cover_hypotheses_to_phase_cover_hypotheses,
		thm:afterBlocking_sectorComparison_reindexed_commonPhaseCover}
	Convert the BNT matching hypotheses for the common primitive nonzero-sector
	families to common phase-cover hypotheses, and then apply the common-sector
	comparison theorem with a common phase cover.
\end{proof}

\begin{theorem}[Sector comparison from normal-CF-BNT common primitive data]%
\label{thm:afterBlocking_sectorComparison_zeroTail_of_commonPrimitiveNormalBNTData_zeroTailIdentity}
	\lean{MPSTensor.afterBlocking_sectorComparison_zeroTail_of_commonPrimitiveNormalBNTData_zeroTailIdentity}
	\leanok
	\uses{def:common_primitive_bnt_cover_hypotheses,
		def:common_sector_relabeling_hypothesis,
		thm:afterBlocking_sectorComparison_reindexed_bntCover}
	Assume the blocked-word identification for the common-sector families.
	The structural theorem supplies common primitive nonzero-sector families;
	if those families are in normal canonical form with BNT separation and
	also carry proportional-decomposition data and one-site injectivity, then
	the zero-tail sector comparison follows. The zero-tail equality is derived
	from the length-zero identity and the proportional BNT comparison.
\end{theorem}

\begin{definition}[After-blocking BNT matching hypotheses]%
	\label{def:after_blocking_sector_comparison_hypotheses}
	\lean{MPSTensor.BlockedNormalFormHypotheses}
	\leanok
	\uses{def:common_primitive_bnt_cover_hypotheses}
	For two tensors with the same MPV family, the structural reduction supplies
	a common blocking with
	\(A^{[p]}\sim\mathbf{0}_{d^p,z_A}\oplus A_{\mathrm{nz}}\),
	\(B^{[p]}\sim\mathbf{0}_{d^p,z_B}\oplus B_{\mathrm{nz}}\), and
	\(z_A=z_B\), together with common primitive nonzero-sector families. These
	hypotheses add the BNT matching hypotheses for those families. The
	blocked-word identification for common cyclic sectors is
	derived from the grouped-block cast theorem.

	The hypotheses include the sectors'
	trace-preserving normalization, primitivity, irreducibility, and positive
	bond-dimension evidence before returning BNT matching hypotheses for
	those blocks. Thus the assumptions apply to the common primitive sectors
	obtained after blocking, not to an arbitrary chosen decomposition.
\end{definition}

\begin{definition}[After-blocking sector-comparison conclusion]%
\label{def:after_blocking_sector_comparison_conclusion}
	\lean{MPSTensor.BlockedNormalFormSectorMatching}
	\leanok
	\uses{def:paper_sector_data}
	The conclusion consists of a positive blocking length and sector
	decompositions $P,Q$ at that length. The blocked tensors agree with
	$P$ and $Q$ at every positive length, while $P$ and $Q$ generate the same
	full MPV family. Both sector decompositions have BNT bases. Their basis
	sectors are matched by a permutation, the corresponding multiplicities agree,
	and the sector-weight multisets agree after multiplying the weights on one
	side by nonzero sector phases.
\end{definition}

\begin{lemma}[After-blocking comparison from supplied BNT matching hypotheses]%
\label{lem:after_blocking_sector_comparison_of_supplied_bnt_cover}
	\lean{MPSTensor.fundamentalTheorem_afterBlocking_from_blockedNormalFormHypotheses}
	\leanok
	\uses{def:after_blocking_sector_comparison_hypotheses,
			def:after_blocking_sector_comparison_conclusion}
	Let $A$ and $B$ generate the same MPV family. Assume the after-blocking
	BNT matching hypotheses of
	Definition~\ref{def:after_blocking_sector_comparison_hypotheses}. Then the
	source-paper sector-comparison conclusion of
	Definition~\ref{def:after_blocking_sector_comparison_conclusion} is
	nonempty.

	The conclusion matches the sector-decomposition part of the
	Cirac--P\'erez-Garc\'ia--Schuch--Verstraete Fundamental Theorem: after
	blocking, the tensor families are represented by BNT sector decompositions
	whose basis sectors and weights match up to permutation and nonzero phases,
	as in \cite{Cirac2016MPDO_arXiv} and \cite{Cirac2021Matrix}.

	This lemma assumes the BNT matching hypotheses. It does not prove
	them from equality of MPV families. Under those hypotheses, together with
	\(A^{[p]}\sim\mathbf{0}_{d^p,z_A}\oplus A_{\mathrm{nz}}\),
	\(B^{[p]}\sim\mathbf{0}_{d^p,z_B}\oplus B_{\mathrm{nz}}\), \(z_A=z_B\),
	trace-preserving normalization, primitivity, irreducibility, and positive
	bond dimensions, the sector-decomposition conclusion is nonempty.
\end{lemma}

	\begin{proof}
		\leanok
		\uses{lem:common_grouped_block_cast_of_flatten_word,
			lem:common_grouped_block_cast_to_relabeling,
			thm:afterBlocking_sectorComparison_reindexed_bntCover}
	Apply Lemma~\ref{thm:afterBlocking_sectorComparison_reindexed_bntCover} with
	the word-identification derived from
	Lemma~\ref{lem:common_grouped_block_cast_to_relabeling} and the BNT matching
	hypotheses from
	Definition~\ref{def:after_blocking_sector_comparison_hypotheses}. The proof
	then records the resulting blocking length, sector decompositions, BNT bases,
	permutation, multiplicity equalities, phases, and sector-weight
	multiset equalities as the conclusion structure.
\end{proof}

\begin{remark}[Equal-case comparison after canonical reduction]\notready
	The equal-case proof in the literature begins with both tensors in canonical
	form.  One chooses a basis of normal tensors for each tensor, so that every
	canonical-form block is a phase times a gauge transform of exactly one basis
	tensor; this is the basis characterization of
	\cite{Cirac2021Matrix} and \cite{Cirac2016MPDO_arXiv}.
	Equality, or proportionality, of the MPV families then pairs the two bases by
	phase and gauge, as in the Fundamental Theorem of Matrix Product Vectors.
	In the equal case, one expands each tensor in its basis of normal tensors and
	compares, for every basis tensor and every length $N$, the power sums
	\[
		\sum_q \mu_{j,q}^N = \sum_q (\nu_{j,q}e^{i\phi_j})^N.
	\]
	Newton identities, equivalently the elementary power-sum lemma used in the
	CPSV argument, recover the multiplicities and the weights.  The per-sector
	gauge matrices are then placed on the diagonal to obtain the global conjugating
	matrix; this is the equal-case corollary of the Fundamental Theorem in
	\cite{Cirac2016MPDO_arXiv}, restated in \cite{Cirac2021Matrix}.

	% Maintainer note: the comparison statements in this chapter cover this
	% latter part of the argument.  Lemma~\ref{lem:ft_equal_same_structure}
	% proves the common block-structure special case, and
	% Lemma~\ref{lem:bnt_power_sum_recover_copies_weights} recovers the copy
	% counts and sector-weight multisets from the multiplicities and weights.
	% The basis-of-normal-tensors construction is available for
	% trace-preserving primitive irreducible nonzero blocks with positive bond
	% dimensions and nonzero weights in
	% Theorem~\ref{thm:bnt_sector_decomp_tp_prim_irr_collapsed}.  When these
	% blocks are injective,
	% Theorems~\ref{thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_data}
	% and~\ref{thm:bnt_sector_decomp_tp_prim_irr_collapsed_overlap_data_span}
	% give the asymptotic orthogonality and finite-length span preservation
	% needed to separate coefficients.
	% Theorem~\ref{thm:bnt_sector_decomp_pair_overlap_span_of_block_span} then
	% turns equality of nonzero-sector spans into the hypotheses used by the
	% sector comparison theorem.

	Common phases are used here only to obtain the required span hypothesis.  A common
	MPV phase cover gives equality of finite-length nonzero-block spans by
	Lemma~\ref{lem:mpv_common_phase_cover_span_eq}.  A sector basis pre-matching
	gives the same span equality and the primitive overlap-span hypotheses by
	Lemma~\ref{lem:sector_basis_pre_matching_span_eq} and
	Lemma~\ref{lem:sector_basis_pre_matching_to_overlap_span}.  The BNT comparison
	lemmas produce such common phases from a proportional-decomposition conclusion in
	Lemmas~\ref{lem:common_phase_cover_of_proportional_decomposition}
	and~\ref{lem:common_phase_cover_of_separated_normal_bnt}.  These common phase-cover
	facts are not a separate form of the source theorem; they are used only to
	obtain the finite-length span equality required by the sector comparison.

	Chapter~\ref{ch:canonical} supplies the canonical-form reduction needed before
	these comparison theorems can be applied: a common physical blocking length
	with \(A^{[p]}\sim\mathbf{0}_{d^p,z_A}\oplus A_{\mathrm{nz}}\),
	\(B^{[p]}\sim\mathbf{0}_{d^p,z_B}\oplus B_{\mathrm{nz}}\), and
	\(z_A=z_B\), plus cyclic-sector families for both tensors on a common
	blocked alphabet.  The blocked-word relabelling and the
	common primitive irreducible nonzero-sector decompositions are now part of the
	reduction, culminating in
	Theorem~\ref{thm:after_blocking_common_primitive_irreducible_blocks_reindexed}.
	Lemma~\ref{thm:afterBlocking_sectorComparison_zeroTail_of_reindexedNonzeroParts_spanHypotheses}
		then states the hypotheses: \(z_A=z_B\), injectivity of the common
	nonzero-sector blocks, and
	finite-length MPV span equality.
	Lemma~\ref{thm:afterBlocking_sectorComparison_reindexed_commonPhaseCover}
	adds the zero-block dimension equality and injectivity hypotheses to the common phase-cover theorem and
	therefore gives the same sector-weight conclusion.  Establishing these
	hypotheses is the mathematical passage from equality of total MPV vectors to the
	componentwise power-sum identities used in
	\cite[Corollary~IV.5]{Cirac2021Matrix}.
\end{remark}
