
%% =========================================================
%% Chapter: The Algebraic Fundamental Theorem
%% Covers the fixed-system-size approach of arXiv:1804.04964
%% (Moln\'ar, Garre-Rubio, P\'erez-Garc\'ia, Schuch, Cirac, NJP 2018)
%% =========================================================

\chapter{The Algebraic Fundamental Theorem}%
\label{ch:algebraic_ft}

This chapter develops the one-dimensional MPS results
of~\cite{Molnar2018NormalPEPS}: the Fundamental Theorem for injective
and normal MPS on a finite periodic chain of fixed length.
The argument is purely algebraic, working directly with $\MN{D}$.
Two preparatory lemmas---the virtual bond gauge
(Lemma~\ref{lem:virtual_bond_gauge}) and the proportionality lemma
(Lemma~\ref{lem:tensor_proportional})---carry the entire argument,
yielding the main theorem and its corollaries for translation-invariant
chains, non-translation-invariant descriptions of translation-invariant states,
and blocked chains.

Throughout, $D$ is the bond dimension and $d$ is the physical dimension
with $d \ge D^2$.
The extensions to injective and normal PEPS on arbitrary graphs and the
improved system-size bound ($2L+1$ in place of $3L$) are not pursued in
this chapter.
The non-translation-invariant chain data, including same chain state,
injectivity, and cyclic gauge equivalence, is set up below.

%% ---------------------------------------------------------
\section{Alternative fixed-length non-TI setup}
\label{sec:non_ti_setup}
%% ---------------------------------------------------------

The algebraic proof of \cite{Molnar2018NormalPEPS} keeps a
fixed periodic chain length and allows the local tensors to vary from site to
site.
The first step is therefore a separate setup for finite
non-translation-invariant chains.

\begin{definition}[Periodic non-TI chain data]\label{def:non_ti_chain}
    \lean{MPSChainTensor, MPSChainTensor.coeff, MPSChainTensor.SameState, MPSChainTensor.IsInjective, MPSChainTensor.GaugeEquiv}
    \leanok
    \uses{def:mps_tensor, def:injective}
    Fix a chain length $n$.
    A \emph{non-translation-invariant periodic MPS chain} consists of sitewise
    tensors
    \[
        A_k = \{A_k^i\}_{i=0}^{d-1}, \qquad k = 0, \dots, n-1,
    \]
    where each $A_k^i \in \MN{D}$.
    For a basis configuration $\sigma = (\sigma_0,\dots,\sigma_{n-1})$, its
    coefficient is
    \[
        c_{A,\sigma}
        := \tr\!\bigl(A_0^{\sigma_0} A_1^{\sigma_1} \cdots A_{n-1}^{\sigma_{n-1}}\bigr).
    \]
    Two such chains generate the \emph{same chain state} if these coefficients
    agree for every $\sigma$.
    The chain is \emph{injective} if each local tensor $A_k$ is injective in
    the usual single-site sense.
    Finally, two chains are \emph{cyclically gauge equivalent} if there are
    invertible bond matrices $Z_0,\dots,Z_{n-1}$ such that
    \[
        B_k^i = Z_k A_k^i Z_{k+1}^{-1}
    \]
    for every site $k$ and physical index $i$, with indices understood modulo
    $n$ so that $Z_n = Z_0$. Locally, this gauge relation is
    represented by
    \begin{center}
        \TNGaugeConjugation{Z_k}{i}{Z_{k+1}^{-1}}
    \end{center}
    where the black node denotes the site tensor named in the surrounding
    formula and the two red side nodes denote the bond gauges on the
    neighboring virtual legs.
\end{definition}

\begin{lemma}[Cyclic gauge equivalence is an equivalence relation]
    \label{lem:chain_gauge_equiv_equivalence}
    \lean{MPSChainTensor.GaugeEquiv.refl}
    \lean{MPSChainTensor.GaugeEquiv.symm}
    \lean{MPSChainTensor.GaugeEquiv.trans}
    \leanok
    \uses{def:non_ti_chain}
    On fixed-length periodic chains, cyclic gauge equivalence is reflexive,
    symmetric, and transitive.
\end{lemma}

\begin{proof}
    \leanok
    Reflexivity uses the identity gauge on every bond.
    Symmetry inverts each bond gauge, and transitivity multiplies the gauges
    pointwise along the chain.
\end{proof}

%% ---------------------------------------------------------
\section{One-sided inverse and decomposition}%
\label{sec:decomposition_map}
%% ---------------------------------------------------------

If $A$ is injective, the linear combination map
$\Phi_A(c) := \sum_{i} c_i A^i : \C^d \to \MN{D}$
is surjective and admits a right inverse.

\begin{lemma}[Existence of a right inverse]\label{lem:exists_rightInverse}
    \lean{MPSTensor.IsInjective.exists_rightInverse}
    \leanok
    \uses{def:injective}
    If $A$ is injective, then there exists a $\C$-linear map
    $\Psi : \MN{D} \to \C^d$ satisfying
    $\Phi_A \circ \Psi = \mathrm{id}_{\MN{D}}$.
\end{lemma}

\begin{proof}
    \leanok
    Surjectivity of $\Phi_A$ yields a linear section by choice.
\end{proof}

\begin{definition}[Decomposition map]\label{def:decomposition_map}
    \lean{MPSTensor.decompositionMap}
    \leanok
    \uses{def:injective, lem:exists_rightInverse}
    Let $A$ be an injective MPS tensor.
    The \emph{decomposition map} $\Psi_A : \MN{D} \to \C^d$ is a
    fixed choice of right inverse of $\Phi_A$.
\end{definition}

\begin{lemma}[Decomposition property]\label{lem:exists_decomposition}
    \lean{MPSTensor.IsInjective.exists_decomposition}
    \leanok
    \uses{def:decomposition_map}
    If $A$ is injective, then for every $X \in \MN{D}$ there exist
    coefficients $c \in \C^d$ such that
    \[
        X = \sum_{i=0}^{d-1} c_i\, A^i.
    \]
\end{lemma}

\begin{proof}
    \leanok
    Take $c = \Psi_A(X)$.
    Then $\Phi_A(c) = \Phi_A(\Psi_A(X)) = X$.
\end{proof}

\begin{remark}[Non-uniqueness of the decomposition]
    \label{rem:decomposition_nonunique}
    When $d > D^2$, the map $\Phi_A$ has a nontrivial kernel, so
    the section $\Psi_A$ is not unique.
    The statements that follow depend only on
    $\Phi_A \circ \Psi_A = \mathrm{id}$, so the choice does not
    affect the conclusions.
\end{remark}

%% ---------------------------------------------------------
\section{Physical realisation of virtual insertions}%
\label{sec:phys_realize}
%% ---------------------------------------------------------

For an injective tensor $A$, the physical realisation map encodes the
effect of inserting a virtual matrix $X$ on the bond as a $d \times d$
operation on the physical indices.

\begin{definition}[Physical realisation map]\label{def:phys_realize}
    \lean{MPSTensor.physRealize}
    \leanok
    \uses{def:decomposition_map}
    Let $A$ be an injective tensor.
    For $X \in \MN{D}$, define the $d \times d$ matrix
    $O_A(X) \in \MN{d}$ by
    \[
        O_A(X)_{ij} := (\Psi_A(A^i X))_j,
    \]
    so that row~$i$ of $O_A(X)$ contains the decomposition coefficients
    of $A^i X$ in the spanning set $\{A^j\}$.
    Diagrammatically, the inserted virtual matrix is re-expressed as a
    physical operation on the same local tensor:
    \begin{center}
        \TNPhysicalRealization{X}{O_A(X)}
    \end{center}
    The map $X \mapsto O_A(X)$ is $\C$-linear.
\end{definition}

\begin{lemma}[Defining property of the physical realisation map]%
    \label{lem:physRealize_spec}
    \lean{MPSTensor.physRealize_spec}
    \leanok
    \uses{def:phys_realize, def:decomposition_map}
    For every $X \in \MN{D}$ and every physical index $i$,
    \[
        A^i\, X = \sum_{j=0}^{d-1} O_A(X)_{ij}\, A^j.
    \]
\end{lemma}

\begin{proof}
    \leanok
    By definition, $O_A(X)_{ij} = (\Psi_A(A^i X))_j$.
    Applying $\Phi_A$ gives
    $\sum_j O_A(X)_{ij} A^j = \Phi_A(\Psi_A(A^i X)) = A^i X$,
    where the last step uses $\Phi_A \circ \Psi_A = \mathrm{id}$.
\end{proof}

\begin{theorem}[Physical realisation is multiplicative]%
    \label{thm:physRealize_mul}
    \lean{MPSTensor.physRealize_mul}
    \leanok
    \uses{def:phys_realize, lem:physRealize_spec, def:injective}
    Let $A$ be an injective MPS tensor.
    For all $X, Y \in \MN{D}$,
    \[
        O_A(XY) = O_A(X)\, O_A(Y).
    \]
\end{theorem}

\begin{proof}
    \leanok
    Applying Lemma~\ref{lem:physRealize_spec} to $X$ yields
    \[
        A^i X = \sum_{j=0}^{d-1} O_A(X)_{ij}\, A^j.
    \]
    Multiply by $Y$ on the right and apply the linear map $\Psi_A$:
    \[
        \Psi_A(A^i XY)
        = \sum_{j=0}^{d-1} O_A(X)_{ij}\, \Psi_A(A^j Y).
    \]
    Taking the $k$th component of both sides gives
    \[
        O_A(XY)_{ik}
        = \sum_{j=0}^{d-1} O_A(X)_{ij}\, O_A(Y)_{jk}
        = (O_A(X)\, O_A(Y))_{ik}.
    \]
    Since this holds for all $i$ and $k$, the matrices are equal.
\end{proof}

\begin{remark}[General injective case]
    \label{rem:physRealize_mul_general}
    \uses{thm:physRealize_mul, def:injective, def:decomposition_map}
    When $d > D^2$, the decomposition map $\Psi_A$ need not be unique.
    Nevertheless, Theorem~\ref{thm:physRealize_mul} is already the
    general injective statement: for the fixed choice of $\Psi_A$ in
    Definition~\ref{def:decomposition_map}, the associated physical
    realisation map $O_A$ is multiplicative.
    This supersedes the earlier basis-only formulation.
\end{remark}

%% ---------------------------------------------------------
\section{Intermediate lemmas}%
\label{sec:intermediate}
%% ---------------------------------------------------------

\begin{lemma}[Blocking preserves injectivity]\label{lem:blocking_preserves_injectivity}
    \lean{MPSTensor.chainCombinedTensor_isInjective}
    \leanok
    \uses{def:injective, def:blocked_tensor}
    Let $A_0, \ldots, A_{m-1}$ be injective MPS tensors with compatible
    bond dimensions.
    Then the blocked tensor
    $(A_0 \otimes \cdots \otimes A_{m-1})^{(i_0,\ldots,i_{m-1})}
        := A_0^{i_0} A_1^{i_1} \cdots A_{m-1}^{i_{m-1}}$
    is again injective.
\end{lemma}

\begin{proof}
    \leanok
    The span of the blocked matrices contains all products
    $A_0^{i_0} \cdots A_{m-1}^{i_{m-1}}$.
    Since each $\{A_k^i\}_i$ spans $\MN{D}$, these products span
    all of $\MN{D}$.
\end{proof}

\begin{lemma}[Centre of $\MN{D}$]\label{lem:center_scalar}
    \lean{Matrix.isScalar_of_commute_span_eq_top}
    \leanok
    If $Z \in \MN{D}$ commutes with every element of $\MN{D}$, then
    $Z = \lambda \Id_D$ for some $\lambda \in \C$.
\end{lemma}

\begin{proof}
    \leanok
    Standard: $Z$ commutes with every matrix unit $E_{ij}$, which
    forces all off-diagonal entries to vanish and all diagonal entries
    to agree.
\end{proof}

%% ---------------------------------------------------------
\section{Same state forces a virtual bond gauge}%
\label{sec:virtual_bond_gauge}
%% ---------------------------------------------------------

This section presents the 3-site virtual-bond statement used later in the
proof of the PEPS fundamental theorem. Here the hypothesis is already the
all-length MPV equality of the combined tensors, and the conclusion is the
conjugacy of the middle-bond insertion data.

\begin{lemma}[Virtual bond gauge -- 3-site case]%
    \label{lem:virtual_bond_gauge}
    \lean{MPSTensor.virtual_bond_gauge}
    \leanok
    \uses{thm:physRealize_mul, thm:simplicity, thm:skolem_noether}
    Let $A = (A_0, A_1, A_2)$ and $B = (B_0, B_1, B_2)$ be two
    3-site injective periodic chains with common bond dimension
    $D \ge 1$ and $\mc{V}(A) = \mc{V}(B)$ (same MPV family).
    Then at the middle bond there exists an invertible matrix
    $Z \in \GL_D(\C)$ such that the virtual-insertion trace
    coefficients agree after conjugation:
    \[
        \tr(A_0^{i_0} X\, A_1^{i_1} A_2^{i_2})
        = \tr(B_0^{i_0} Z^{-1} X Z\, B_1^{i_1} B_2^{i_2})
    \]
    for all $X \in \MN{D}$ and all physical indices $i_0, i_1, i_2$.

    In \cite[Lemma~1]{Molnar2018NormalPEPS}, the statement is given for
    arbitrary chain length $n \ge 3$ and any bond position, with
    uniqueness of $Z$ up to scalar.
    The chain-level generalization to arbitrary $n \ge 3$
    and any bond position is not pursued here.
\end{lemma}

\begin{proof}
    \leanok
    \uses{thm:physRealize_mul, thm:simplicity, thm:skolem_noether,
        lem:blocking_preserves_injectivity}
    The $\mc{V}(A) = \mc{V}(B)$ hypothesis gives, for every $X \in \MN{D}$,
    equality of the virtual-insertion trace coefficients:
    $\tr(A_0^{i_0} X\, A_1^{i_1} A_2^{i_2})
     = \tr(B_0^{i_0} Y\, B_1^{i_1} B_2^{i_2})$
    for some $Y$ depending on $X$.
    Since all three site tensors on each side are injective,
    the physical realisation maps $O_A, O_B$ are algebra homomorphisms
    by Theorem~\ref{thm:physRealize_mul}.
    Setting $T(X) = Y$ defines a $\C$-linear multiplicative map,
    hence a nonzero $\C$-algebra endomorphism of $\MN{D}$, which is
    bijective by Theorem~\ref{thm:simplicity}.
    Skolem--Noether (Theorem~\ref{thm:skolem_noether}) gives
    $Z \in \GL_D(\C)$ with $T(X) = Z^{-1} X Z$.
\end{proof}


%% ---------------------------------------------------------
\section{Aligned gauges force proportionality}%
\label{sec:tensor_proportional}
%% ---------------------------------------------------------

After applying Lemma~\ref{lem:virtual_bond_gauge} to each bond and
absorbing the gauge matrices, one obtains two chains whose
virtual-insertion operations agree on every bond.
Lemma~2 of~\cite{Molnar2018NormalPEPS} shows that this agreement
forces the local tensors to be proportional.

\begin{lemma}[Aligned gauges force proportionality --
    Lemma~2 of \protect\cite{Molnar2018NormalPEPS}]%
    \label{lem:tensor_proportional}
    \lean{MPSTensor.tensor_proportional}
    \leanok
    \uses{def:injective, def:decomposition_map, lem:center_scalar}
    Let $A_1, A_2$ and $B_1, B_2$ be injective MPS tensors with
    the same bond dimension $D$.
    Suppose that for all $X \in \MN{D}$ (on the internal bond),
    all $Y \in \MN{D}$ (on the external bond), and all physical indices $i,j$,
    \begin{align}
        \tr(A_1^{i}\, X\, A_2^{j})
         &= \tr(B_1^{i}\, X\, B_2^{j})
        \label{eq:internal_bond}\\
        \tr(A_2^{j}\, Y\, A_1^{i})
         &= \tr(B_2^{j}\, Y\, B_1^{i})
        \label{eq:external_bond}
    \end{align}
    Diagrammatically, inserting $X$ on the bond between the two sites
    (Eq.~\ref{eq:internal_bond}):
    \begin{center}
        \TNInternalTraceInsertion{i}{j}{X}
    \end{center}
    while the insertion $Y$ on the outer trace bond
    (Eq.~\ref{eq:external_bond}) is:
    \begin{center}
        \TNExternalTraceInsertion{i}{j}{Y}
    \end{center}
    so the two hypotheses distinguish the internal and external virtual
    loops of the two-site periodic chain.

    Then there exists a nonzero scalar $\lambda \in \C^{\times}$ such
    that
    \[
        A_1^{i} = \lambda\, B_1^{i}
        \qquad\text{and}\qquad
        A_2^{j} = \lambda^{-1}\, B_2^{j}
    \]
    for all physical indices $i, j$.
\end{lemma}

\begin{proof}
    \leanok
    \uses{def:decomposition_map, lem:center_scalar}
    By cyclicity and non-degeneracy of the trace pairing,
    the conditions in Eqs.~(\ref{eq:internal_bond}) and~(\ref{eq:external_bond})
    give, for all physical indices $i,j$,
    \[
        \begin{aligned}
        A_2^{j} A_1^{i} &= B_2^{j} B_1^{i},\\
        A_1^{i} A_2^{j} &= B_1^{i} B_2^{j}.
        \end{aligned}
    \]
    Write $\Id_D = \sum_j c_j A_2^j$ using the decomposition map
    $\Psi_{A_2}$, and define $Z := \sum_j c_j B_2^j \in \MN{D}$.
    This is the point where the two-site comparison collapses to a single
    boundary matrix:
    \begin{center}
        \TNBoundaryRegrow{Z}{i}{j}{2}
        \qquad
        \TNLocalEqualityStep{Z}{\lambda \Id}{i}
    \end{center}
    The first product identity gives $A_1^i = Z B_1^i$, and the second
    gives $A_1^i = B_1^i Z$.
    Hence $Z$ commutes with every $B_1^i$; since $\{B_1^i\}$ spans
    $\MN{D}$, $Z$ is central.
    By Lemma~\ref{lem:center_scalar}, $Z = \lambda \Id_D$ for some
    $\lambda \in \C^{\times}$ (nonzero since the tensors are nonzero).
    Thus $A_1^i = \lambda B_1^i$, and substituting into the first
    product identity forces $A_2^j = \lambda^{-1} B_2^j$.
\end{proof}

%% ---------------------------------------------------------
\section{The Fundamental Theorem for injective chains}%
\label{sec:ft_chain}
%% ---------------------------------------------------------

The chain result below combines the site and physical indices into a single
physical index and applies the single-block Fundamental Theorem to the resulting
tensor. The fixed-length same-state statement requires a separate same-state
reduction hypothesis.

\begin{theorem}[Fundamental Theorem for injective chains]%
    \label{thm:fundamentalTheorem_injective_chain}
    \lean{MPSChainTensor.fundamentalTheorem_injective_chain}
    \leanok
    \uses{def:non_ti_chain, def:same_mpv}
    Let $\mathbf{A} = (A_0, \ldots, A_{n-1})$ and
    $\mathbf{B} = (B_0, \ldots, B_{n-1})$ be two $n$-site chains with
    common bond dimension $D$, and assume that $\mathbf{A}$ is injective.
    Suppose that, after combining the site index and physical index into a
    single MPS tensor, the resulting two tensors generate the same MPV
    family. Then $\mathbf{A}$ and $\mathbf{B}$ are cyclically gauge
    equivalent.
\end{theorem}

\begin{proof}
    \leanok
    \uses{lem:blocking_preserves_injectivity, thm:ft_single}
    Choose one site of $\mathbf{A}$; its injectivity implies injectivity of
    the combined tensor by Lemma~\ref{lem:blocking_preserves_injectivity}.
    Apply the single-block Fundamental Theorem
    (Theorem~\ref{thm:ft_single}) to the two combined tensors.
    This yields one invertible matrix $X$ such that every combined matrix of
    $\mathbf{B}$ equals the corresponding combined matrix of
    $\mathbf{A}$ conjugated by $X$.
    Reading this identity back at each site gives
    $B_k^i = X A_k^i X^{-1}$ for all $k$ and $i$, so the constant choice
    $Z_k = X$ is a cyclic gauge.
\end{proof}


\begin{lemma}[Uniform site rescaling rescales the combined tensor]%
    \label{lem:chainCombinedTensor_smul_chain}
    \lean{MPSTensor.chainCombinedTensor_smul_chain}
    \leanok
    \uses{def:non_ti_chain}
    Let $\mathbf{A} = (A_0, \ldots, A_{n-1})$ be an $n$-site chain and let
    $\zeta \in \C$. If $\zeta \mathbf{A}$ denotes the chain obtained by
    replacing each site tensor $A_k$ with $\zeta A_k$, and if
    $\widetilde{\mathbf{A}}$ denotes the combined tensor of $\mathbf{A}$,
    then
    \[
        \widetilde{\zeta \mathbf{A}} = \zeta\, \widetilde{\mathbf{A}}.
    \]
\end{lemma}

\begin{proof}
    \leanok
    \uses{def:non_ti_chain}
    Combining the site and physical indices does not change which local matrix
    is read at the component $(k,i)$, so both sides evaluate to
    $\zeta A_k^i$ there.
\end{proof}

\begin{theorem}[Injective-chain Fundamental Theorem up to a global phase]%
    \label{thm:fundamentalTheorem_injective_chain_gaugePhase}
    \lean{MPSChainTensor.fundamentalTheorem_injective_chain_gaugePhase}
    \leanok
    \uses{thm:fundamentalTheorem_injective_chain, def:gauge_phase_equiv,
        def:non_ti_chain}
    Let $\mathbf{A} = (A_0, \ldots, A_{n-1})$ and
    $\mathbf{B} = (B_0, \ldots, B_{n-1})$ be two $n$-site chains with common
    bond dimension $D$, and assume that $\mathbf{A}$ is injective.
    Suppose that the combined tensors $\widetilde{\mathbf{A}}$ and
    $\widetilde{\mathbf{B}}$ are gauge-phase equivalent.
    Then there exist invertible matrices $Z_0, \ldots, Z_{n-1} \in \GL_D(\C)$
    and a nonzero scalar $\zeta \in \C^{\times}$ such that
    \[
        B_k^i = \zeta\, Z_k A_k^i Z_{k+1}^{-1}
    \]
    for all sites $k$ and physical indices $i$, where $k+1$ is interpreted
    cyclically modulo $n$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{lem:chainCombinedTensor_smul_chain,
        thm:gauge_invariance, thm:fundamentalTheorem_injective_chain}
    If the combined tensor of $\mathbf{B}$ differs from that of $\mathbf{A}$
    by a gauge and a nonzero scalar $\zeta$, rescale each site of
    $\mathbf{B}$ by $\zeta^{-1}$. Lemma~\ref{lem:chainCombinedTensor_smul_chain}
    removes this common scalar from the combined tensor, so
    Theorem~\ref{thm:fundamentalTheorem_injective_chain} applies to the
    rescaled chain. Multiplying the resulting sitewise relations by $\zeta$
    gives the displayed conclusion.
\end{proof}

\subsection{From fixed-length chain states to combined-tensor MPV agreement}

\begin{definition}[Same-state reduction hypothesis]
    \label{def:same_state_reduction_hyp}
    \lean{MPSChainTensor.SameStateBridgeHyp}
    \leanok
    \uses{def:non_ti_chain, def:same_mpv}
    A \emph{same-state reduction hypothesis} for dimensions $(d,D)$ is the
    assertion that whenever $\mathbf{A}$ is an injective $n$-site chain with
    $n \ge 3$ and $\mathbf{A},\mathbf{B}$ generate the same chain state,
    then the combined tensors of $\mathbf{A}$ and $\mathbf{B}$ generate the
    same MPV family.
\end{definition}

\begin{lemma}[Same state gives combined-tensor MPV agreement]
    \label{lem:sameMPV_chainCombined_of_sameState}
    \lean{MPSChainTensor.sameMPV_chainCombined_of_sameState}
    \leanok
    \uses{def:same_state_reduction_hyp, def:non_ti_chain, def:same_mpv}
    Assume a same-state reduction hypothesis for $(d,D)$.
    Let $\mathbf{A}$ and $\mathbf{B}$ be $n$-site chains with common bond
    dimension $D$, with $n \ge 3$, and assume that $\mathbf{A}$ is
    injective.
    If $\mathbf{A}$ and $\mathbf{B}$ generate the same chain state, then the
    combined tensors of $\mathbf{A}$ and $\mathbf{B}$ generate the same MPV
    family.
\end{lemma}

\begin{proof}
    \leanok
    \uses{def:same_state_reduction_hyp}
    This is exactly the defining property of the same-state reduction
    hypothesis.
\end{proof}

\begin{theorem}[Fundamental Theorem for injective chains from same state]
    \label{thm:fundamentalTheorem_injective_chain_of_sameState}
    \lean{MPSChainTensor.fundamentalTheorem_injective_chain_of_sameState}
    \leanok
    \uses{def:same_state_reduction_hyp, def:non_ti_chain}
    Assume a same-state reduction hypothesis for $(d,D)$.
    Let $\mathbf{A}$ and $\mathbf{B}$ be $n$-site chains with common bond
    dimension $D$, with $n \ge 3$, and assume that $\mathbf{A}$ is
    injective.
    If $\mathbf{A}$ and $\mathbf{B}$ generate the same chain state, then
    $\mathbf{A}$ and $\mathbf{B}$ are cyclically gauge equivalent.
\end{theorem}

\begin{proof}
    \leanok
    \uses{lem:sameMPV_chainCombined_of_sameState,
        thm:fundamentalTheorem_injective_chain}
    Apply Lemma~\ref{lem:sameMPV_chainCombined_of_sameState} to convert the
    fixed-length chain-state hypothesis into equality of the combined-tensor
    MPV families, then invoke
    Theorem~\ref{thm:fundamentalTheorem_injective_chain}.
\end{proof}

\begin{remark}[Relationship to the translation-invariant all-sizes theorem]
    \label{rem:ft_vs_singleBlock}
    \uses{thm:fundamentalTheorem_injective_chain, def:injective}
    Theorem~\ref{thm:fundamentalTheorem_injective_chain} treats a fixed chain
    length and assumes equality of the combined-tensor MPV families.
    The single-block Fundamental Theorem
    (Theorem~\ref{thm:ft_single}) instead studies one tensor at all system
    sizes and concludes exact gauge equivalence
    $B^i = X A^i X^{-1}$.
    The chain theorem is recovered by converting the chain into one combined
    tensor; the price is that the hypothesis is all-length MPV equality for
    that combined tensor rather than equality of the chain state at one fixed
    length.
\end{remark}

\subsection{Finite-length MPV agreement}

\begin{definition}[Finite-length MPV agreement]\label{def:same_mpv_from}
    \lean{MPSTensor.SameMPVFrom}
    \leanok
    \uses{def:same_mpv}
    Two tensors $A$ and $B$ of the same bond dimension satisfy
    \emph{finite-length MPV agreement from $N_0$ onward} if they generate
    the same MPV family for all system sizes $N \ge N_0$, meaning that for
    every $N \ge N_0$ and $\sigma \in [d]^N$,
    \[
        V^{(N)}(A)_\sigma = V^{(N)}(B)_\sigma
    \]
\end{definition}

\begin{theorem}[Finite-length agreement implies full agreement]%
    \label{thm:sameMPV_of_sameMPVFrom}
    \lean{MPSTensor.sameMPV_of_sameMPVFrom_of_injective}
    \leanok
    \uses{def:same_mpv_from, def:injective, def:same_mpv}
    Let $A$ be an injective tensor of bond dimension $D \ge 1$, and let
    $B$ be a tensor of the same bond dimension. If $A$ and $B$ generate
    the same MPV family for every system size $N \ge N_0$, then they
    generate the same MPV family for every system size.
\end{theorem}

\begin{proof}
    \leanok
    \uses{def:same_mpv_from, def:injective}
    Write $1 = \sum_i c_i A^i$ (by injectivity) and set
    $S = \sum_i c_i B^i$.
    \begin{enumerate}
        \item Show that the length-$N_0$ word span of $B$ is all of
              $\MN{D}$ via the linear relation between the
              length-$N_0$ word-evaluation maps of $A$ and $B$
              induced by trace agreement at length $N_0 + 1$,
              together with a finrank transfer argument.
        \item Derive $S = 1$ by trace nondegeneracy: for every
              $M \in \MN{D}$, $\tr(M(S - 1)) = 0$.
        \item Downward induction on $k = N_0 - |w|$: decompose
              $\tr(w_A)$ and $\tr(w_B)$ using $1 = \sum c_i A^i$
              and $S = 1$ respectively, then apply the inductive
              hypothesis at length $|w| + 1$.
    \end{enumerate}
\end{proof}

\begin{theorem}[Strengthened single-block FT (finite-length)]%
    \label{thm:ft_singleBlock_finiteLength}
    \lean{MPSTensor.fundamentalTheorem_singleBlock_finiteLength}
    \leanok
    \uses{thm:sameMPV_of_sameMPVFrom, thm:ft_single, def:gauge_equiv}
    Let $A$ be an injective tensor of bond dimension $D \ge 1$, and let
    $B$ be a tensor of the same bond dimension. If $A$ and $B$ agree on
    all MPV coefficients for system sizes $N \ge N_0$ (for any threshold
    $N_0$), then $A$ and $B$ are gauge equivalent.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:sameMPV_of_sameMPVFrom, thm:ft_single}
    Apply Theorem~\ref{thm:sameMPV_of_sameMPVFrom} to obtain
    equality of the MPV families for all system sizes, then invoke the
    single-block Fundamental Theorem (Theorem~\ref{thm:ft_single}).
\end{proof}

%% ---------------------------------------------------------
\section{Corollaries}%
\label{sec:ft_corollaries}
%% ---------------------------------------------------------

\subsection{Translation-invariant chains}

\begin{theorem}[Single gauge for translation-invariant chains]
    \label{thm:ti_tensors_single_gauge}
    \lean{MPSChainTensor.ti_tensors_single_gauge}
    \leanok
    \uses{def:injective, def:same_mpv}
    Let $A$ and $B$ be MPS tensors of bond dimension $D$, and fix a chain
    length $n \ge 1$.
    Assume that $A$ is injective and that the combined tensors of the two
    constant chains $(A, \ldots, A)$ and $(B, \ldots, B)$ generate the same
    MPV family.
    Then there exists an invertible matrix $X \in \GL_D(\C)$ such that
    \[
        B^i = X\, A^i\, X^{-1}
    \]
    for every physical index~$i$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{lem:blocking_preserves_injectivity, thm:ft_single}
    Bundle the constant chains into one tensor.
    Injectivity of $A$ implies injectivity of the combined tensor by
    Lemma~\ref{lem:blocking_preserves_injectivity}, and the combined-tensor MPV
    equality is exactly the hypothesis needed for the single-block
    Fundamental Theorem.
    The resulting gauge matrix is independent of the site because every site
    carries the same local tensor.
\end{proof}

\begin{corollary}[Translation-invariant collapse to one gauge]\label{cor:ft_ti_chain}
    \lean{MPSChainTensor.ti_tensors_collapse_to_single_gauge}
    \leanok
    \uses{def:injective, def:same_mpv}
    Let $A$ and $B$ be MPS tensors of bond dimension $D$, and fix a chain
    length $n \ge 1$.
    Assume that $A$ is injective and that the combined tensors of the two
    constant chains $(A, \ldots, A)$ and $(B, \ldots, B)$ generate the same
    MPV family.
    Then there exist an invertible matrix $Z \in \GL_D(\C)$ and a scalar
    $\lambda \in \C^{\times}$ with $\lambda^n = 1$ such that
    \[
        B^i = \lambda\, Z^{-1}\, A^i\, Z
    \]
    for every physical index~$i$.
\end{corollary}

\begin{proof}
    \leanok
    \uses{thm:ti_tensors_single_gauge}
    Apply Theorem~\ref{thm:ti_tensors_single_gauge} to obtain
    $B^i = X A^i X^{-1}$.
    Set $Z = X^{-1}$ and $\lambda = 1$.
    Then $\lambda^n = 1$ and the displayed identity is exactly the same
    gauge relation rewritten in the form used by the corollary.
\end{proof}

\subsection{Non-translation-invariant descriptions of translation-invariant states}

\begin{corollary}[Translation-invariant states from nonuniform chains]%
    \label{cor:non_ti_to_ti}
    \notready
    \uses{def:same_state_reduction_hyp, def:non_ti_chain}
    Assume a same-state reduction hypothesis for $(d,D)$.
    Let $\mathbf{A} = (A_0, \ldots, A_{n-1})$ be an injective
    non-translation-invariant chain of length $n \ge 3$.
    Suppose the state generated by $\mathbf{A}$ is translation-invariant,
    i.e.\ it equals the state generated by the cyclically shifted chain
    $(A_1, \ldots, A_{n-1}, A_0)$.
    Then $\mathbf{A}$ admits a translation-invariant description with the
    same bond dimension: there exists an injective tensor $B$ such that
    the constant chain $(B, \ldots, B)$ generates the same state as
    $\mathbf{A}$.
    Explicitly, $B^i = A_0^i\, Z_0^{-1}$ (right multiplication,
    \emph{not} conjugation), where $Z_0$ is extracted from the cyclic shift.
\end{corollary}

\begin{proof}
    \uses{thm:fundamentalTheorem_injective_chain_of_sameState}
    Apply Theorem~\ref{thm:fundamentalTheorem_injective_chain_of_sameState}
    to $\mathbf{A}$ and its cyclic shift.
    This gives invertible gauges $Z_0, \ldots, Z_{n-1}$ with
    $A_{k+1}^i = Z_k^{-1} A_k^i Z_{k+1}$.
    Iterating these identities yields
    $A_k^i = L_k^{-1} A_0^i R_k$, where the products $L_k, R_k$ satisfy
    $R_k L_{k+1}^{-1} = Z_0^{-1}$.
    Substituting into the chain trace shows that the same state is generated
    by the constant tensor $B^i := A_0^i Z_0^{-1}$.
    Since $A_0$ is injective and $Z_0$ is invertible, $B$ is injective.
\end{proof}

\noindent Symmetry-theoretic consequences of the Fundamental Theorem,
including virtual representations and string order, are developed in
Chapter~\ref{ch:symmetry}.
\subsection{Blocked chains}

\begin{lemma}[Block injectivity and blocked-tensor injectivity]
    \label{lem:isNBlkInjective_iff_blockTensor_isInjective}
    \lean{MPSTensor.isNBlkInjective_iff_blockTensor_isInjective}
    \leanok
    \uses{def:blocked_tensor, def:injective}
    A tensor $A$ is $L$-block injective if and only if its $L$-blocked
    tensor $A^{[L]}$ is injective.
\end{lemma}

\begin{proof}
    \leanok
    The blocked physical alphabet is just a reindexing of words of length $L$.
    Decoding blocked indices therefore identifies the two spanning families,
    so the two injectivity conditions are equivalent.
\end{proof}

\begin{definition}[Constant blocked chain]
    \label{def:blocked_chain}
    \lean{MPSChainTensor.blockedChain}
    \leanok
    \uses{def:blocked_tensor}
    For an MPS tensor $A$, a blocking length $L$, and a chain length $n$, the
    \emph{blocked chain} is the constant $n$-site chain whose local tensor is
    the blocked tensor $A^{[L]}$.
\end{definition}

\begin{lemma}[Blocked chains are injective]
    \label{lem:blocked_chain_isInjective}
    \lean{MPSChainTensor.blockedChain_isInjective}
    \leanok
    \uses{def:blocked_chain, lem:isNBlkInjective_iff_blockTensor_isInjective}
    If $A$ is $L$-block injective, then the constant blocked chain built from
    $A^{[L]}$ is injective at every site.
\end{lemma}

\begin{proof}
    \leanok
    \uses{lem:isNBlkInjective_iff_blockTensor_isInjective}
    By Lemma~\ref{lem:isNBlkInjective_iff_blockTensor_isInjective}, the local
    blocked tensor $A^{[L]}$ is injective, and therefore every site of the
    constant blocked chain is injective.
\end{proof}

\begin{theorem}[Fundamental Theorem for blocked chains]
    \label{thm:fundamentalTheorem_blockedChain}
    \lean{MPSChainTensor.fundamentalTheorem_blockedChain}
    \leanok
    \uses{def:blocked_chain, def:same_mpv}
    Let $A$ and $B$ be MPS tensors of bond dimension $D$, let $L$ be a
    blocking length, and let $n$ be a chain length.
    Assume that $A$ is $L$-block injective and that the combined tensors of the
    two constant blocked chains built from $A^{[L]}$ and $B^{[L]}$ generate the
    same MPV family.
    Then the two blocked chains are gauge equivalent.
\end{theorem}

\begin{proof}
    \leanok
    \uses{lem:blocked_chain_isInjective,
        thm:fundamentalTheorem_injective_chain}
    By Lemma~\ref{lem:blocked_chain_isInjective}, the blocked chain built from
    $A^{[L]}$ is injective.
    Apply Theorem~\ref{thm:fundamentalTheorem_injective_chain} to the two
    constant blocked chains.
\end{proof}

\begin{remark}
    The theorem above is the blocked-chain result established here.
    Recovering sitewise gauges for the original unblocked tensors gives the
    normal-MPS corollary of \cite{Molnar2018NormalPEPS}; that step is outside
    this blocked-chain statement.
\end{remark}

%% ---------------------------------------------------------
\section{Product algebra construction and per-block gauge}%
\label{sec:pi_algebra}
%% ---------------------------------------------------------

The chain Fundamental Theorem (Theorem~\ref{thm:fundamentalTheorem_injective_chain})
handles non-translation-invariant periodic chains.
A parallel line of argument addresses block-diagonal tensors:
given a family of injective block tensors $\{A_k^i\}_{k=0}^{r-1}$
with block dimensions $D_k$ and a second family $\{B_k^i\}$ with
per-block $\mc{V}(A_k) = \mc{V}(B_k)$, one seeks per-block gauge
equivalence $B_k^i = X_k\, A_k^i\, X_k^{-1}$ for each~$k$
independently, and a global block-permutation plus inner-automorphism
decomposition.

\subsection{Per-block linear extension}

\begin{definition}[Per-block linear extension]%
    \label{def:perBlockLinearExtension}
    \lean{MPSTensor.perBlockLinearExtension}
    \leanok
    \uses{def:injective, def:same_mpv}
    Let $A_k, B_k$ be MPS tensors of bond dimension $D_k$ with
    $A_k$ injective and $\mc{V}(A_k) = \mc{V}(B_k)$ for each block
    $k \in \{0, \ldots, r{-}1\}$.
    The \emph{per-block linear extension} $T_k : \MN{D_k} \to \MN{D_k}$
    is the unique $\C$-linear map satisfying $T_k(A_k^i) = B_k^i$
    for every physical index~$i$.
\end{definition}

\begin{lemma}[Per-block multiplicativity]%
    \label{lem:perBlockLinearExtension_mul}
    \lean{MPSTensor.perBlockLinearExtension_mul}
    \leanok
    \uses{def:perBlockLinearExtension}
    For each block~$k$, the per-block linear extension $T_k$ is
    multiplicative:
    $T_k(MN) = T_k(M)\, T_k(N)$ for all $M, N \in \MN{D_k}$.
\end{lemma}

\begin{proof}\leanok\uses{def:perBlockLinearExtension}
    Since $\{A_k^i\}$ spans $\MN{D_k}$, it suffices to check on spanning
    elements, where the result follows from the MPV identity.
\end{proof}

\begin{lemma}[Per-block bijectivity]%
    \label{lem:perBlockLinearExtension_bijective}
    \lean{MPSTensor.perBlockLinearExtension_bijective}
    \leanok
    \uses{lem:perBlockLinearExtension_mul}
    Each $T_k$ is bijective.
\end{lemma}

\begin{proof}\leanok\uses{lem:perBlockLinearExtension_mul}
    A nonzero multiplicative linear endomorphism of $\MN{D_k}$ is
    automatically bijective: injectivity follows from the simplicity
    of $\MN{D_k}$ (the kernel is a proper ideal, hence zero), and
    surjectivity from finite dimension.
    Non-vanishing of $T_k$ is established via the trace pairing: if
    $T_k = 0$, then all $B_k^i = 0$, contradicting the non-vanishing
    of the trace pairing for injective tensors.
\end{proof}

\subsection{Product algebra automorphism}

\begin{definition}[Product algebra automorphism]%
    \label{def:piAlgEquiv}
    \lean{MPSTensor.piAlgEquiv}
    \leanok
    \uses{def:perBlockLinearExtension, lem:perBlockLinearExtension_mul,
        lem:perBlockLinearExtension_bijective}
    The \emph{product algebra automorphism} is the $\C$-algebra
    equivalence
    \[
        T : \prod_{k=0}^{r-1} \MN{D_k}
        \xrightarrow{ \sim }
        \prod_{k=0}^{r-1} \MN{D_k}
    \]
    defined by $T(M)_k = T_k(M_k)$, where each $T_k$ is the per-block
    linear extension.
    It is an algebra automorphism since each $T_k$ is a multiplicative,
    unital, bijective linear map.
    Diagrammatically, this is the componentwise map on the product algebra
    obtained by applying the corresponding extension to each summand.
\end{definition}

\begin{theorem}[Block-permutation decomposition]%
    \label{thm:piAlgEquiv_decomposition}
    \lean{MPSTensor.piAlgEquiv_decomposition}
    \leanok
    \uses{def:piAlgEquiv, thm:skolem_noether}
    The product algebra automorphism $T$ decomposes as a block
    permutation composed with per-block inner automorphisms: there exist
    a permutation $\sigma \in S_r$ (with $D_{\sigma(k)} = D_k$ for
    all~$k$) and invertible matrices $X_k \in \GL_{D_k}(\C)$ such that
    \[
        (T(M))_{\sigma(k)} = X_k\, M\, X_k^{-1}
    \]
    for all $M \in \MN{D_k}$ (viewed as supported in block~$k$) and
    all~$k$.
    Equivalently, the product-algebra automorphism first permutes the
    simple matrix blocks and then acts inside each block by an inner
    conjugation.
\end{theorem}

\begin{proof}\leanok\uses{def:piAlgEquiv, thm:skolem_noether}
    The automorphism $T$ permutes the block ideals of $\prod_k \MN{D_k}$
    (each block ideal is simple, and an automorphism sends simple ideals
    to simple ideals of the same dimension).
    This gives the permutation $\sigma$.
    On each block, the restricted map is a $\C$-algebra automorphism of
    $\MN{D_k}$; by Skolem--Noether (Theorem~\ref{thm:skolem_noether}),
    it is inner. Thus $T$ splits algebraically into the permutation of
    block ideals followed by inner conjugation on each surviving block.
\end{proof}

\subsection{Nondegeneracy of the product trace pairing}

\begin{lemma}[Nondegeneracy of the product trace pairing]%
    \label{lem:piTrace_mul_right_eq_zero}
    \lean{MPSTensor.piTrace_mul_right_eq_zero}
    \leanok
    Let $M = (M_0, \ldots, M_{r-1}) \in \prod_{k} \MN{D_k}$.
    If $\sum_{k} \tr(M_k N_k) = 0$ for all
    $N = (N_0, \ldots, N_{r-1})$, then $M = 0$.
\end{lemma}

\begin{proof}\leanok
    Choose $N$ supported on a single block~$k$ (all other components
    zero).
    The hypothesis reduces to $\tr(M_k N_k) = 0$ for all $N_k$, so
    $M_k = 0$ by nondegeneracy of the matrix trace pairing on
    $\MN{D_k}$.
    Since $k$ was arbitrary, $M = 0$.
\end{proof}

\begin{lemma}[Injectivity of the product Gram map]%
    \label{lem:piTraceMulRightPi_ker_eq_bot}
    \lean{MPSTensor.piTraceMulRightPi_ker_eq_bot}
    \leanok
    \uses{def:injective}
    Let $\{A_k\}$ be a family of injective block tensors.
    The product Gram map
    $M \mapsto (k, i \mapsto \tr(M_k \cdot A_k^i))$
    is injective.
\end{lemma}

\begin{proof}\leanok\uses{def:injective}
    Reduces to per-block injectivity of the Gram map
    $M_k \mapsto (i \mapsto \tr(M_k \cdot A_k^i))$, which holds by
    injectivity of $A_k$ (the matrices $\{A_k^i\}$ span $\MN{D_k}$,
    so the trace pairing against them separates points).
\end{proof}

%% ---------------------------------------------------------
\section{Gauge equivalence for direct sums}%
\label{sec:multi_block_ft}
%% ---------------------------------------------------------

\begin{definition}[Flattened block-diagonal gauge]%
    \label{def:global_gauge_of_blocks}
    \lean{MPSTensor.blockDiagonalGL, MPSTensor.globalGaugeOfBlocks}
    \leanok
    \uses{def:tensor_from_blocks}
    Given per-block gauges $X_k \in \GL_{D_k}(\C)$, first form the
    dependent block-diagonal unit $\bigoplus_k X_k$ and then reindex the
    dependent sum of block coordinates to the flattened bond index
    $\Fin(\sum_k D_k)$.  The resulting element of
    $\GL_{\sum_k D_k}(\C)$ is the global gauge used by the assembled tensor
    $\bigoplus_k \mu_k A_k$.
\end{definition}

\begin{lemma}[Explicit direct-sum conjugation by the flattened gauge]%
    \label{lem:tensor_from_blocks_globalGauge_conj}
    \lean{MPSTensor.toTensorFromBlocks_eq_globalGaugeOfBlocks_conj,
        MPSTensor.gaugeEquiv_toTensorFromBlocks_of_blockConj}
    \leanok
    \uses{def:global_gauge_of_blocks, def:tensor_from_blocks, def:gauge_equiv}
    If each block satisfies $B_k^i=X_k A_k^i X_k^{-1}$, then the weighted
    direct sums satisfy, for every physical index~$i$,
    \[
        \bigoplus_k \mu_k B_k^i
        = X\left(\bigoplus_k \mu_k A_k^i\right)X^{-1},
    \]
    where $X$ is the flattened block-diagonal gauge of
    Definition~\ref{def:global_gauge_of_blocks}.  Hence the two assembled
    tensors are gauge equivalent with this explicit witness.
\end{lemma}

\begin{proof}\leanok
    The block-diagonal matrix product is computed blockwise.  The scalar weight
    $\mu_k$ commutes with the per-block conjugation, so each diagonal block of
    the right-hand side is $\mu_k B_k^i$.  Reindexing from the dependent block
    coordinates to $\Fin(\sum_k D_k)$ preserves multiplication and inverses.
\end{proof}

\begin{lemma}[Gauge equivalence of direct sums from per-block MPV equality]%
    \label{lem:fundamentalTheorem_multiBlock_full}
    \lean{MPSTensor.fundamentalTheorem_multiBlock_full}
    \leanok
    \uses{def:injective, def:same_mpv, def:gauge_equiv,
        def:tensor_from_blocks, lem:tensor_from_blocks_globalGauge_conj}
    Let $\{A_k\}_{k=0}^{r-1}$ and $\{B_k\}_{k=0}^{r-1}$ be two families
    of MPS tensors with block dimensions $D_k$, each $A_k$ injective, and
    $\mc{V}(A_k) = \mc{V}(B_k)$ for all~$k$.
    Then:
    \begin{enumerate}
        \item Per-block gauge equivalence:
              $B_k^i = X_k\, A_k^i\, X_k^{-1}$ for some
              $X_k \in \GL_{D_k}(\C)$, for every $k$ and $i$.
        \item Global gauge equivalence: the block-diagonal tensors
              $\bigoplus_k \mu_k A_k$ and $\bigoplus_k \mu_k B_k$ are
              gauge equivalent.
    \end{enumerate}
    Explicitly, for $X=\bigoplus_k X_k$,
    \[
        \bigoplus_k \mu_k B_k^i
        = X \left(\bigoplus_k \mu_k A_k^i\right) X^{-1}.
    \]
\end{lemma}

\begin{proof}\leanok\uses{thm:multi_block_ft, thm:multi_block_global}
    Per-block gauge equivalence follows from applying the single-block
    Fundamental Theorem to each block, using injectivity of
    $A_k$ and $\mc{V}(A_k) = \mc{V}(B_k)$.
    With $X=\bigoplus_k X_k$, the displayed identity follows by multiplying
    block-diagonal matrices block by block.
\end{proof}

\begin{lemma}[Explicit per-block gauges]%
    \label{lem:fundamentalTheorem_multiBlock_explicit}
    \lean{MPSTensor.fundamentalTheorem_multiBlock_explicit}
    \leanok
    \uses{lem:fundamentalTheorem_multiBlock_full}
    Under the hypotheses of
    Lemma~\ref{lem:fundamentalTheorem_multiBlock_full}, there exist
    invertible matrices $X_k \in \GL_{D_k}(\C)$ such that
    \[
        B_k^i = X_k\, A_k^i\, X_k^{-1}
    \]
    for every block $k$ and every physical index $i$.
\end{lemma}

\begin{proof}
    \leanok
    \uses{lem:fundamentalTheorem_multiBlock_full}
    Extract the gauge matrix from the per-block part of
    Lemma~\ref{lem:fundamentalTheorem_multiBlock_full} for each block.
\end{proof}

\begin{lemma}[Product-algebra decomposition of per-block MPV equality]%
    \label{lem:fundamentalTheorem_multiBlock_decomposition}
    \lean{MPSTensor.fundamentalTheorem_multiBlock_decomposition}
    \leanok
    \uses{lem:fundamentalTheorem_multiBlock_full}
    Assume that every block dimension is nonzero.
    Under the hypotheses of
    Lemma~\ref{lem:fundamentalTheorem_multiBlock_full}, the product-algebra
    automorphism determined by the per-block linear extensions decomposes into
    a block permutation together with inner conjugation on each block: there
    exist a permutation $\sigma$ of the blocks, compatible block-dimension
    equalities, and invertible matrices $X_k$ such that, after reindexing the
    target block $\sigma(k)$ back to block $k$, the action on the $k$th matrix
    block is $M \mapsto X_k M X_k^{-1}$.
\end{lemma}

\begin{proof}
    \leanok
    \uses{thm:piAlgEquiv_decomposition}
    Apply Theorem~\ref{thm:piAlgEquiv_decomposition} to the product-algebra
    automorphism built from the per-block linear extensions.
\end{proof}

\begin{lemma}[Per-block equality of MPV families and per-block gauge equivalence]%
    \label{lem:perBlock_sameMPV_iff_gaugeEquiv}
    \lean{MPSTensor.perBlock_sameMPV_iff_gaugeEquiv}
    \leanok
    \uses{def:injective, def:same_mpv, def:gauge_equiv}
    Let $\{A_k\}$ be a family of injective block tensors.
    Then $\mc{V}(A_k) = \mc{V}(B_k)$ for all~$k$ if and only if
    $A_k$ and $B_k$ are gauge equivalent for all~$k$.
\end{lemma}

\begin{proof}\leanok\uses{def:gauge_equiv}
    The forward direction applies the single-block Fundamental Theorem to
    each block.
    The reverse direction holds because gauge equivalence preserves all
    traces of word evaluations.
\end{proof}

%% ---------------------------------------------------------
\section{Block separation from ordered canonical-form data}%
\label{sec:block_separation}
%% ---------------------------------------------------------

Lemma~\ref{lem:fundamentalTheorem_multiBlock_full} assumes
$\mc{V}(A_k)=\mc{V}(B_k)$ for every block~$k$.
In practice, one starts from the weaker global hypothesis that the
block-diagonal tensors satisfy the same MPV family:
$\mc{V}\!(\bigoplus_k \mu_k A_k) =
    \mc{V}\!(\bigoplus_k \mu_k B_k)$.
Block separation extracts the per-block identities from the global one
under canonical-form normalization plus a strict separation of the
weight moduli.
The first step is the weighted block-sum identity
\[
    \sum_k \mu_k^N V^{(N)}(A_k)_\sigma = \sum_k \mu_k^N V^{(N)}(B_k)_\sigma,
\]
which is then combined with overlap estimates and the leading-block peel.

\begin{definition}[Canonical-form data with ordered weights]\label{def:isCanonicalForm}
    \lean{MPSTensor.IsCanonicalForm}
    \leanok
    \uses{def:injective}
    A family of block tensors $\{A_k\}_{k=0}^{r-1}$ with scaling
    weights $\mu_0, \ldots, \mu_{r-1} \in \C^{\times}$ has
    \emph{canonical-form data with ordered weights} if:
    \begin{enumerate}
        \item each $A_k$ is injective,
        \item each $A_k$ is left-canonical:
              $\sum_i (A_k^i)^{\dagger}\, A_k^i = \Id_{D_k}$,
        \item the weights are non-increasing by norm and nonzero:
              $|\mu_0| \ge |\mu_1| \ge \cdots \ge |\mu_{r-1}| > 0$,
        \item the self-overlaps are normalised:
              $O_{A_k A_k}(N) \to 1$ as $N \to \infty$.
    \end{enumerate}
    This records the normalized ordered data used before the final
    block-separation argument.  Equal-modulus blocks are still allowed here,
    as in the canonical decompositions in arXiv:1606.00608.  The strict
    inequality needed for the leading-block peeling argument is a separate
    hypothesis in the block-separation lemmas below, and at the BNT level it
    is obtained only after grouping repeated representatives.
\end{definition}

\begin{definition}[Normal canonical-form data with ordered weights]\label{def:normal_canonical_form}
    \lean{MPSTensor.IsNormalCanonicalForm}
    \leanok
    \uses{def:isCanonicalForm}
    A family has \emph{normal canonical-form data with ordered weights} if it
    satisfies Definition~\ref{def:isCanonicalForm} with each block additionally
    \emph{irreducible} (no nontrivial invariant projection) and
    \emph{primitive} (the transfer map has spectral gap).
\end{definition}

\begin{lemma}[Block separation from weighted word identities]%
    \label{lem:block_separation_all_words}
    \lean{MPSTensor.block_separation_all_words}
    \leanok
    \uses{def:same_mpv}
    Let $\{A_k\}$ and $\{B_k\}$ be block families with strictly ordered,
    nonzero weights $\mu_k$.
    Assume that every $A_k$ and $B_k$ is injective and left-canonical, that
    each self-overlap $O_{A_kA_k}(N)$ tends to $1$, and that for every system
    size $N$ and every word $\sigma$ one has the weighted identity
    \[
        \sum_k \mu_k^N \bigl(V^{(N)}(A_k)_\sigma - V^{(N)}(B_k)_\sigma\bigr)=0.
    \]
    Then $\mc{V}(A_k)=\mc{V}(B_k)$ for every block $k$.
\end{lemma}

\begin{proof}
    \leanok
    The proof is the peeling argument described below for canonical-form data
    equipped with a separate strict ordering of the weight moduli.
    The strict weight ordering isolates the leading block, the overlap bounds
    show that the tail decays geometrically after dividing by the dominant
    weight, and one then iterates on the remaining blocks.
\end{proof}

\begin{lemma}[Block separation from canonical-form data and strict weights]%
    \label{lem:per_block_sameMPV_of_canonical_form}
    \lean{MPSTensor.per_block_sameMPV_of_canonical_form}
    \leanok
    \uses{def:isCanonicalForm, def:same_mpv}
    Let $\{A_k\}$ have canonical-form data with ordered weights
    $\{\mu_k\}$, and assume in addition that the weight norms are strictly
    decreasing.  Let $\{B_k\}$ be another family with each $B_k$ injective
    and left-canonical.
    If $\mc{V}\!(\bigoplus_k \mu_k A_k)
        = \mc{V}\!(\bigoplus_k \mu_k B_k)$,
    then $\mc{V}(A_k) = \mc{V}(B_k)$ for each block~$k$ individually.
    The proof isolates the leading block and treats the remaining blocks
    as an exponentially decaying tail with
    $\rho = |\mu_1|/|\mu_0| < 1$.
\end{lemma}

\begin{proof}
    \leanok
    \uses{def:isCanonicalForm}
    The proof proceeds by strong induction on the number of blocks~$r$.
    For $r = 1$, the weighted sum degenerates and the result is
    immediate after cancelling $\mu_0^N \ne 0$.

    For $r \ge 2$, the global identity
    $\sum_k \mu_k^N\, \bigl(V^{(N)}(A_k)_\sigma
        - V^{(N)}(B_k)_\sigma\bigr) = 0$
    holds for all $N$ and $\sigma$.
    \[
        \mu_0^N (V^{(N)}(A_0)_\sigma - V^{(N)}(B_0)_\sigma)
        =
        - \sum_{k \ge 1} \mu_k^N
        (V^{(N)}(A_k)_\sigma - V^{(N)}(B_k)_\sigma).
    \]
    Dividing by $\mu_0^N$ shows that the right-hand side is
    $O(\rho^N)$ with $\rho = |\mu_1|/|\mu_0| < 1$.
    Multiplying by the MPV components of $A_0$ and summing over $\sigma$
    converts this to an overlap identity.
    Dividing by $|\mu_0|^{2N}$ and using the strict norm ordering
    $|\mu_1| / |\mu_0| < 1$, the tail terms
    $(k \ge 1)$ decay geometrically with ratio
    $\rho = |\mu_1| / |\mu_0|$.
    The leading term ($k = 0$) therefore converges:
    $O_{A_0 B_0}(N) \to 1$ as $N \to \infty$.
    Combined with the normalised self-overlap
    $O_{A_0 A_0}(N) \to 1$, this gives
    $\mc{V}(A_0) = \mc{V}(B_0)$ via the gauge-phase equivalence
    criterion.
    Subtracting the leading block and applying the inductive hypothesis
    yields $\mc{V}(A_k) = \mc{V}(B_k)$ for all remaining $k$.
\end{proof}

\begin{lemma}[Block separation from separated canonical data]%
    \label{lem:per_block_sameMPV_of_separated_canonical_data}
    \lean{MPSTensor.per_block_sameMPV_of_separated_canonical_data}
    \leanok
    \uses{def:same_mpv}
    Suppose the weights are nonzero with strictly decreasing norms, the
    $A$-blocks are injective and left-canonical with normalized self-overlaps,
    the $B$-blocks are injective and left-canonical, and the block-diagonal
    tensors generate the same MPV family.
    Then $\mc{V}(A_k)=\mc{V}(B_k)$ for every block $k$.
\end{lemma}

\begin{proof}
    \leanok
    \uses{lem:block_separation_all_words}
    Equality of the block-diagonal MPV families gives the weighted word identity
    required by Lemma~\ref{lem:block_separation_all_words}.
    The separated hypotheses supply exactly the injectivity,
    left-canonicality, strict-weight, and self-overlap assumptions needed
    there.
\end{proof}

\begin{lemma}[Gauge comparison from separated canonical data]%
    \label{lem:fundamentalTheorem_of_separated_canonical_data}
    \lean{MPSTensor.fundamentalTheorem_of_separated_canonical_data}
    \leanok
    \uses{lem:per_block_sameMPV_of_separated_canonical_data}
    Under the hypotheses of
    Lemma~\ref{lem:per_block_sameMPV_of_separated_canonical_data}, every
    block pair $A_k,B_k$ is gauge equivalent, and the block-diagonal
    tensors are globally gauge equivalent.
\end{lemma}

\begin{proof}
    \leanok
    \uses{lem:per_block_sameMPV_of_separated_canonical_data,
        lem:fundamentalTheorem_multiBlock_full}
    First apply
    Lemma~\ref{lem:per_block_sameMPV_of_separated_canonical_data} to get
    $\mc{V}(A_k)=\mc{V}(B_k)$ for every block~$k$.
    Then apply
    Lemma~\ref{lem:fundamentalTheorem_multiBlock_full}.
\end{proof}

\begin{lemma}[Explicit gauge from separated canonical data]%
    \label{lem:fundamentalTheorem_of_separated_canonical_data_explicit}
    \lean{MPSTensor.fundamentalTheorem_of_separated_canonical_data_explicit}
    \leanok
    \uses{lem:fundamentalTheorem_of_separated_canonical_data}
    Under the same hypotheses, there exist invertible matrices
    $X_k \in \GL_{D_k}(\C)$ such that
    \[
        B_k^i = X_k\, A_k^i\, X_k^{-1}
    \]
    for every block $k$ and physical index $i$.
\end{lemma}

\begin{proof}
    \leanok
    \uses{lem:per_block_sameMPV_of_separated_canonical_data,
        lem:fundamentalTheorem_multiBlock_explicit}
    Combine Lemma~\ref{lem:per_block_sameMPV_of_separated_canonical_data} with
    the matrices $X_k$ from Lemma~\ref{lem:fundamentalTheorem_multiBlock_explicit}.
\end{proof}

\begin{lemma}[Gauge comparison from canonical-form data and strict weights]%
    \label{lem:fundamentalTheorem_canonicalForm}
    \lean{MPSTensor.fundamentalTheorem_canonicalForm}
    \leanok
    \uses{lem:per_block_sameMPV_of_canonical_form,
        lem:fundamentalTheorem_multiBlock_full}
    Under the hypotheses of
    Lemma~\ref{lem:per_block_sameMPV_of_canonical_form},
    every block pair $A_k, B_k$ is gauge equivalent, and the
    block-diagonal tensors are globally gauge equivalent.
\end{lemma}

\begin{proof}\leanok
    \uses{lem:per_block_sameMPV_of_canonical_form,
        lem:fundamentalTheorem_multiBlock_full}
    Lemma~\ref{lem:per_block_sameMPV_of_canonical_form} gives per-block
    $\mc{V}(A_k) = \mc{V}(B_k)$.
    Lemma~\ref{lem:fundamentalTheorem_multiBlock_full} converts this
    to per-block and global gauge equivalence.
\end{proof}

\begin{lemma}[Explicit gauge from canonical-form data and strict weights]%
    \label{lem:fundamentalTheorem_canonicalForm_explicit}
    \lean{MPSTensor.fundamentalTheorem_canonicalForm_explicit}
    \leanok
    \uses{lem:fundamentalTheorem_of_separated_canonical_data_explicit}
    Under the same hypotheses, there exist invertible matrices
    $X_k \in \GL_{D_k}(\C)$ such that
    $B_k^i = X_k\, A_k^i\, X_k^{-1}$
    for every block $k$ and physical index $i$.
\end{lemma}

\begin{proof}\leanok
    \uses{lem:fundamentalTheorem_of_separated_canonical_data_explicit}
    Apply Lemma~\ref{lem:fundamentalTheorem_of_separated_canonical_data_explicit}
    with the strict-ordering data from the separated hypotheses.
\end{proof}

%% ---------------------------------------------------------
\section{Heterogeneous equal-case block matching}%
\label{sec:hetero_equal_case_ft}
%% ---------------------------------------------------------

\begin{lemma}[Heterogeneous equal-case block matching]%
    \label{lem:fundamentalTheorem_equalMPV_CFBNT_hetero}
    \lean{MPSTensor.fundamentalTheorem_equalMPV_CFBNT_hetero}
    \leanok
    \uses{def:same_mpv, def:cf_bnt, def:gauge_phase_equiv}
    Let $\{\mu_j A_j\}_{j=0}^{r_A-1}$ and $\{\nu_k B_k\}_{k=0}^{r_B-1}$ be
    two canonical-form BNT families, possibly with different block counts and
    different bond dimensions in corresponding blocks.
    If the block-diagonal tensors generate the same MPV family, then
    the two families have the same number of blocks, the blocks can be matched
    by a permutation preserving the bond dimensions, and each matched pair is
    gauge-phase equivalent.
    This is the block-matching conclusion, not yet the final global weighted
    gauge-equivalence statement for the block-diagonal tensors.
\end{lemma}

\begin{proof}
    \leanok
    The proof is the block-matching argument carried out in the supporting
    modules for the heterogeneous equal case: one first produces a
    non-decaying cross-overlap for each block, applies the overlap dichotomy to
    show that the matched blocks must have equal dimension and differ only by a
    phase gauge, and finally upgrades the resulting matching to a permutation of
    all blocks.
\end{proof}

%% ---------------------------------------------------------
\section{Translation-invariant reduction}%
\label{sec:ti_reduction}
%% ---------------------------------------------------------

\begin{theorem}[Translation-invariant reduction corollary]%
    \label{thm:ti_reduction_corollary}
    \lean{MPSChainTensor.ti_reduction_corollary}
    \leanok
    \uses{def:injective, def:same_mpv}
    Let $A$ and $B$ be MPS tensors of bond dimension $D$ with $A$
    injective.
    Consider the two constant translation-invariant chains
    $(A, \ldots, A)$ and $(B, \ldots, B)$ of length $n \ge 1$.
    If the combined tensors of the two chains generate the same MPV
    family, then:
    \begin{enumerate}
        \item $B$ is gauge equivalent to $A$: there exists
              $X \in \GL_D(\C)$ with $B^i = X\, A^i\, X^{-1}$ for
              all~$i$.
        \item $B$ is injective.
    \end{enumerate}
\end{theorem}

\begin{proof}\leanok\uses{thm:ti_tensors_single_gauge}
    Theorem~\ref{thm:ti_tensors_single_gauge} gives a single gauge matrix
    $X$ with $B^i = X A^i X^{-1}$ for all $i$.
    Since gauge equivalence preserves injectivity, $B$ is injective as well.
\end{proof}

\begin{theorem}[Translation-invariant reduction from same state]%
    \label{thm:ti_reduction_of_sameState}
    \lean{MPSChainTensor.ti_reduction_of_sameState}
    \leanok
    \uses{def:same_state_reduction_hyp, thm:ti_reduction_corollary}
    Assume a same-state reduction hypothesis for $(d,D)$.
    Let $A$ and $B$ be MPS tensors of bond dimension $D$, and let $n \ge 3$.
    If the two constant chains $(A,\ldots,A)$ and $(B,\ldots,B)$ generate the
    same chain state, with $A$ injective, then the same conclusion as in
    Theorem~\ref{thm:ti_reduction_corollary} holds: $B$ is gauge equivalent to
    $A$, and $B$ is injective.
\end{theorem}

\begin{proof}\leanok
    \uses{lem:sameMPV_chainCombined_of_sameState,
        thm:ti_reduction_corollary}
    The same-state reduction hypothesis upgrades equality of the two constant
    chain states to equality of their combined-tensor MPV families.
    Then Theorem~\ref{thm:ti_reduction_corollary} applies.
\end{proof}

\subsection{Global symmetry via gauge composition}

The virtual representation theorem
(Theorem~\ref{thm:virtual_rep_injective}) relies on the fact that
composing two gauge transformations yields another gauge up to a
scalar factor.

\begin{lemma}[Gauge ratio commutes with tensor matrices]%
    \label{lem:gauge_ratio_commutes}
    \lean{MPSTensor.gauge_ratio_commutes}
    \leanok
    \uses{def:gauge_equiv}
    Diagrammatically, the hypothesis is that the two local gauges
    \begin{center}
        \TNGaugeConjugation{X}{i}{X^{-1}}
        \qquad and\qquad
        \TNGaugeConjugation{Y}{i}{Y^{-1}}
    \end{center}
    produce the same target tensor $B^i$ from the same source tensor
    $A^i$.
    If $X, Y \in \GL_D(\C)$ both conjugate an MPS tensor $A$ to the same
    tensor $B$ (i.e.\ $B^i = X A^i X^{-1} = Y A^i Y^{-1}$ for all $i$),
    then $Y^{-1} X$ commutes with every $A^i$.
\end{lemma}

\begin{proof}\leanok
    Direct matrix calculation: conjugating both gauge relations and
    composing gives $(Y^{-1} X) A^i = A^i (Y^{-1} X)$.
\end{proof}

\begin{lemma}[Gauge ratio is scalar for injective tensors]%
    \label{lem:gauge_ratio_isScalar}
    \lean{MPSTensor.gauge_ratio_isScalar}
    \leanok
    \uses{lem:gauge_ratio_commutes, def:injective, lem:center_scalar}
    Under the same hypotheses, if $A$ is additionally injective, then
    $Y^{-1} X = \lambda \Id_D$ for some scalar $\lambda \in \C$.
\end{lemma}

\begin{proof}\leanok\uses{lem:gauge_ratio_commutes, lem:center_scalar}
    The same two-gauge picture shows that the ratio $Y^{-1}X$ acts
    invisibly on the local tensor.
    By Lemma~\ref{lem:gauge_ratio_commutes}, $Y^{-1}X$ commutes with
    all $A^i$.
    Since $\{A^i\}$ spans $\MN{D}$, the matrix $Y^{-1}X$ commutes
    with all of $\MN{D}$.
    By Lemma~\ref{lem:center_scalar}, it is a scalar matrix.
\end{proof}

\begin{theorem}[Projective multiplication law for gauge matrices]%
    \label{thm:gaugeMatrix_projective_mul}
    \lean{MPSTensor.gaugeMatrix_projective_mul}
    \leanok
    \uses{lem:gauge_ratio_isScalar, def:injective,
        def:on_site_symmetric, lem:twisted_tensor_mul}
    Let $A$ be an injective MPS tensor with on-site symmetry under a
    group representation $U : G \to \GL_d(\C)$.
    For each $g \in G$, let $X(g) \in \GL_D(\C)$ satisfy
    $\widetilde{A}_g^i = X(g)\, A^i\, X(g)^{-1}$ for all~$i$.
    Then for all $g, h \in G$, there exists $c(g,h) \in \C$
    such that
    \[
        X(h)\, X(g) = c(g,h)\, X(g \cdot h).
    \]
    In fact $c(g,h) \ne 0$ since both sides are invertible, so
    $c(g,h) \in \C^{\times}$. Diagrammatically, the two candidate gauges
    are competing local replacements for the same twisted tensor:
    \begin{center}
        \TNLocalEqualityStep{X(h)\,X(g)}{X(gh)}{i}
    \end{center}
    and both sides conjugate $A^i$ to the same twisted tensor
    $\widetilde{A}_{gh}^i$.
\end{theorem}

\begin{proof}\leanok
    \uses{lem:gauge_ratio_isScalar, lem:twisted_tensor_mul}
    By the composition law $\widetilde{A}_{gh} = \widetilde{(\widetilde{A}_h)}_g$
    (Lemma~\ref{lem:twisted_tensor_mul}), both $X(h)\,X(g)$ and $X(gh)$
    conjugate $A$ to $\widetilde{A}_{gh}$.
    Lemma~\ref{lem:gauge_ratio_isScalar} then gives the scalar factor.
\end{proof}

%% ---------------------------------------------------------
\section{Closing remarks}%
\label{sec:closing}
%% ---------------------------------------------------------

\begin{remark}
    Appendix~A of~\cite{Molnar2018NormalPEPS} strengthens the
    normal-MPS bound from $n \ge 3L$ to $n \ge 2L + 1$ by a more
    refined blocking argument.
    The injective PEPS theorem of Chapter~\ref{ch:peps_ft} and the normal
    PEPS theorem on arbitrary graphs both rest on the same two core lemmas
    (Lemma~\ref{lem:virtual_bond_gauge} and
    Lemma~\ref{lem:tensor_proportional}); the graph geometry enters only
    through the blocking step that reduces to a three-site chain.
\end{remark}
