\chapter{Exponential Decay of Correlations}\label{ch:correlations}

This chapter records the transfer-matrix expression for thermodynamic-limit
correlators of a normal MPS and the resulting exponential decay estimate.
The key mechanism is spectral: after removing the leading eigenvalue
contribution, the connected part is governed by subleading eigenvalues of the
transfer map.

When the MPS tensor generates a parent Hamiltonian
(Chapter~\ref{ch:parent_hamiltonian}), the correlator quantifies spatial
correlations in the ground state.  The exponential decay rate is set by the
spectral properties of the transfer map, i.e.\ by the modulus of its
subleading eigenvalues.

The spectral analysis of the transfer map and the mixed transfer operator,
including the eigenvalue bound $\rho(F_{AB}) \le 1$, the strict spectral
gap $\rho(F_{AB}) < 1$ for non-equivalent blocks, and the resulting
overlap decay, is developed in Chapter~\ref{ch:spectral}.  This chapter
focuses on the \emph{correlator} side: one-point and two-point expectations,
their spectral decomposition, and quantitative bounds on decay rates.

\section{Connected correlators from the transfer map}

\begin{definition}[One-point expectation]\label{def:one_point_expectation}
    \lean{MPSTensor.onePointExpectation}
    \leanok
    \uses{def:transfer_map, thm:qpf}
    Fix an MPS tensor and a right fixed point of its transfer map.
    Write the tensor as $A$ and the fixed point as $\rho^R$.
    Under the hypotheses of Theorem~\ref{thm:qpf}, that fixed point exists
    and is unique up to scaling.
    For a local observable $X \in \MN{D}$ define
    \[
        \langle X_0 \rangle := \tr(X\rho^R).
    \]
\end{definition}

\begin{definition}[Two-point expectation at distance $n$]\label{def:two_point_expectation}
    \lean{MPSTensor.twoPointExpectation}
    \leanok
    \uses{def:transfer_map, def:one_point_expectation}
    For observables $X,Y \in \MN{D}$ and $n\ge 0$,
    \[
        \langle X_0 Y_n \rangle
        := \tr\!\left(Y\,\E_A^n(X\rho^R)\right).
    \]
    Equivalently: insert $X$ at site $0$, propagate by $\E_A^n$, then insert
    $Y$ and take the trace.
\end{definition}

\begin{definition}[Connected correlator]\label{def:connected_correlator}
    \lean{MPSTensor.connectedCorrelator}
    \leanok
    \uses{def:two_point_expectation, def:one_point_expectation}
    The connected two-point function is
    \[
        C(X,Y;n) := \langle X_0Y_n\rangle - \langle X_0\rangle\langle Y_0\rangle.
    \]
\end{definition}

\section{Spectral expansion and exponential decay}

\begin{theorem}[Sum of exponentials]\label{thm:correlator_sum_exponentials}
    \lean{MPSTensor.connectedCorrelator_eq_sum}
    \leanok
    \uses{def:connected_correlator, thm:spectral_radius_le_one, thm:qpf}
    For a normal MPS, the connected correlator admits a spectral expansion of
    the form
    \[
        C(X,Y;n) = \sum_{j=1}^{D^2-1} c_j\,\lambda_j^n,
    \]
    where $\lambda_j$ are subleading transfer-matrix eigenvalues.
\end{theorem}

\begin{proof}
    The leading eigenvalue contribution ($\lambda_1=1$) equals
    $\langle X_0\rangle\langle Y_0\rangle$ and cancels in the connected part.
    Apply spectral decomposition to $\E_A^n$.
\end{proof}

\begin{theorem}[Exponential decay bound]\label{thm:correlator_exponential_bound}
    \lean{MPSTensor.connectedCorrelator_bound}
    \leanok
    \uses{thm:correlator_sum_exponentials}
    If $|\lambda_j| \le |\lambda_2| < 1$ for all subleading eigenvalues, then
    there exists $C_{XY}\ge 0$ such that
    \[
        |C(X,Y;n)| \le C_{XY}\,|\lambda_2|^n.
    \]
\end{theorem}

\begin{proof}
    Combine the sum-of-exponentials expansion with the triangle inequality.
\end{proof}

\begin{remark}[Spectral hypothesis and the spectral gap]%
    \label{rem:spectral_hypothesis}
    \uses{thm:correlator_exponential_bound, thm:primitive_compl_lt_one}
    The hypothesis ``$|\lambda_2| < 1$'' in
    Theorem~\ref{thm:correlator_exponential_bound} is precisely the
    \emph{spectral gap} condition for the transfer map $\E_A$.
    For primitive channels, the complementary spectral gap
    $\rho(\E_A - P) < 1$ is established in
    Theorem~\ref{thm:primitive_compl_lt_one} (Chapter~\ref{ch:channels});
    for irreducible trace-preserving tensors, the analogous result is
    Theorem~\ref{thm:spectral_gap_irr_tp}.
    Thus the exponential decay of connected correlations follows
    directly from the block separation theory once the spectral
    decomposition is in hand.
\end{remark}

\begin{definition}[Correlation length]\label{def:correlation_length}
    \lean{MPSTensor.correlationLength}
    \leanok
    \uses{thm:correlator_exponential_bound}
    The correlation length associated with $\lambda_2$ is
    \[
        \xi := -\frac{1}{\log|\lambda_2|}.
    \]
\end{definition}

\begin{lemma}[Correlation length is positive]\label{lem:correlation_length_pos}
    \lean{MPSTensor.correlationLength_pos}
    \leanok
    \uses{def:correlation_length}
    If $|\lambda_2| < 1$ and $\lambda_2 \ne 0$, then $\xi > 0$.
\end{lemma}

\begin{proof}
    \leanok
    Since $0 < |\lambda_2| < 1$, we have $\log|\lambda_2| < 0$, so
    $\xi = -1/\log|\lambda_2| > 0$.
\end{proof}

\section{Quantitative spectral gap bounds}

The spectral gap theorems in Chapter~\ref{ch:channels} establish
$\rho(\E_A - P) < 1$ for primitive channels qualitatively.  This section
records the \emph{quantitative} refinements that give explicit constants
and rates for the single-tensor transfer map $\E_A$.
All three quantitative results below are now fully proved.

\begin{theorem}[Exponential convergence of injective primitive channels]%
    \label{thm:exponential_convergence_primitive}
    \lean{MPSTensor.exponential_convergence_of_primitive}
    \leanok
    \uses{thm:primitive_compl_lt_one, def:fixed_point_proj,
          thm:primitive_overlap}
    Fix an injective, trace-preserving MPS tensor $A$ with positive
    definite fixed point $\rho$.
    Let
    \[
        P(X) = \frac{\tr(X)}{\tr(\rho)}\rho
    \]
    be the corresponding fixed-point projection.
    Then there exist $C > 0$ and $0 < \delta \le 1$ such that for all
    $n \ge 0$ and $X \in \MN{D}$,
    \[
        \|\E_A^n(X) - P(X)\|
        \;\le\; C\,(1-\delta)^n\,\|X\|.
    \]
    The spectral gap $\delta$ exists by primitivity
    (Theorem~\ref{thm:primitive_compl_lt_one}, the complementary spectral gap).
\end{theorem}

\begin{proof}
    \leanok
    Injectivity implies primitivity of the transfer map.
    The complementary spectral radius $\rho(\E_A - P) < 1$ then follows from
    the primitive spectral gap theorem.  The Gelfand formula yields a
    geometric bound $\|\,(\E_A - P)^n\| \le C\,(1-\delta)^n$ for suitable
    constants, giving the claimed estimate.
\end{proof}

\begin{theorem}[Correlation length bound]\label{thm:correlation_length_bound}
    \lean{MPSTensor.correlation_length_bound}
    \leanok
    \uses{def:correlation_length, thm:exponential_convergence_primitive,
          def:injective}
    For an injective trace-preserving MPS tensor, there exist
    $C > 0$ and $\xi > 0$ such that for all $n \ge 0$ and all
    traceless $X \in \MN{D}$ (i.e.\ $\tr(X) = 0$),
    \[
        \|\E_A^n(X)\|
        \;\le\; C\,e^{-n/\xi}\,\|X\|.
    \]
    Traceless matrices lie in $\ker P$, where $P$ is the fixed-point
    projection.  Since $\E_A - P$ has spectral radius strictly less
    than~$1$, the Gelfand formula gives exponential decay at rate
    $\xi = -1/\log\rho(\E_A - P)$.
\end{theorem}

\begin{proof}
    \leanok
    For traceless $X$ we have $P(X) = 0$, so $\E_A^n(X) = (\E_A - P)^n(X)$.
    The exponential convergence theorem gives a geometric bound
    $C\,(1-\delta)^n\,\|X\|$.  Rewriting $(1-\delta)^n = e^{-n/\xi}$ with
    $\xi = -1/\log(1-\delta)$ yields the claimed estimate.
\end{proof}

\begin{theorem}[Spectral gap from injectivity]%
    \label{thm:spectral_gap_wielandt}
    \lean{MPSTensor.spectral_gap_of_injective}
    \leanok
    \uses{thm:spectral_gap, thm:wielandt_bound}
    For an injective trace-preserving MPS tensor, there exists
    $\delta > 0$ such that every eigenvalue $\mu \neq 1$ of the
    transfer map satisfies $|\mu| \le 1 - \delta$.
\end{theorem}
\begin{proof}
    \leanok
    Injectivity implies eventual full Kraus rank. This yields primitivity,
    hence a spectral gap for the complementary map. Since the transfer map has
    only finitely many eigenvalues, one may choose a uniform
    $\delta > 0$ separating $1$ from the remaining spectrum.
\end{proof}

\begin{remark}[Status of quantitative bounds]\label{rem:quantitative_status}
    \uses{thm:exponential_convergence_primitive,
          thm:correlation_length_bound}
    All three quantitative bounds above are now fully proved:
    exponential convergence of primitive irreducible channels,
    the correlation length bound for injective tensors, and
    the spectral gap from injectivity.
\end{remark}

\section{Relation to parent Hamiltonians}

\begin{remark}\label{rem:correlations_parent_hamiltonian}
    \uses{def:parent_hamiltonian, lem:parent_hamiltonian_annihilates,
          def:correlation_length, thm:correlator_exponential_bound}
    When the MPS tensor $A$ generates a parent Hamiltonian $H_N(A,L)$
    (Definition~\ref{def:parent_hamiltonian}) whose ground state is the
    matrix product vector (Lemma~\ref{lem:parent_hamiltonian_annihilates}),
    the exponential decay bound
    (Theorem~\ref{thm:correlator_exponential_bound}) shows that ground-state
    correlations decay with a finite correlation length $\xi$.
\end{remark}

\section{Renormalization fixed points and zero correlation length}

\begin{definition}[Renormalization fixed point]\label{def:is_rfp}
    \lean{MPSTensor.IsRFP}
    \leanok
    \uses{def:transfer_map}
    An MPS tensor $A$ is a \emph{renormalization fixed point} (RFP) when its
    transfer map is idempotent, $\E_A \circ \E_A = \E_A$.
    See \cite[Definition~3.2]{Cirac2021Matrix}.
\end{definition}

\begin{definition}[Local orthogonality]\label{def:is_locally_orthogonal}
    \lean{MPSTensor.IsLocallyOrthogonal}
    \leanok
    \uses{def:is_rfp}
    A single BNT block is \emph{locally orthogonal} when its self-transfer map
    is idempotent. For a single tensor this coincides with the RFP condition
    (Definition~\ref{def:is_rfp}); in the full multi-block BNT setting, local
    orthogonality additionally requires that mixed transfer operators vanish.
    See \cite[Definition~3.5]{Cirac2021Matrix}.
\end{definition}

\begin{theorem}[RG flow convergence for canonical-form tensors]%
    \label{thm:rg_flow_converges_of_cf}
    \lean{MPSTensor.rg_flow_converges_of_cf}
    \leanok
    \uses{def:is_canonical_form, def:transfer_map, def:fixed_point_proj,
          thm:exponential_convergence_primitive}
    For a tensor in canonical form with blocks $(A_k)$ and weights $(\mu_k)$,
    the iterated blocking $\E_{A_k}^{2^n}$ converges pointwise to an
    idempotent linear map $\E_\infty$ as $n\to\infty$.
    That is, there exists $\E_\infty$ with $\E_\infty^2 = \E_\infty$ such that
    $\E_{A_k}^{2^n}(\rho) \to \E_\infty(\rho)$ for all density matrices $\rho$.
    This is \cite[Appendix~B]{Cirac2021Matrix}.
\end{theorem}
\begin{proof}\leanok
    Each block is injective and left-canonical, so its transfer map is a
    primitive channel with a unique positive-definite fixed point $\rho_0$.
    The exponential convergence bound
    $\lVert \E^n(X) - P(X)\rVert \le C(1-\delta)^n \lVert X\rVert$
    (Theorem~\ref{thm:exponential_convergence_primitive}) then gives
    pointwise convergence $\E^n(X)\to P(X)$, and composing with the
    subsequence $2^n\to\infty$ yields the result.
    The witness $\E_\infty = P$ is the rank-one fixed-point projection,
    which is idempotent.
\end{proof}

\begin{remark}\label{rem:zero_correlation_length}
    \uses{thm:spectral_radius_le_one, def:correlation_length, def:is_rfp}
    The identity $\E^2=\E$ is equivalent to vanishing subleading spectrum;
    equivalently, $\xi=0$ and connected correlations vanish at nonzero
    separation.
\end{remark}

\begin{definition}[Correlations independent of distance]\label{def:is_cid}
    \lean{MPSTensor.IsCID}
    \leanok
    \uses{def:connected_correlator}
    An MPS tensor has \emph{correlations independent of distance} (CID) when
    its connected two-point function $C(X,Y;n)$ is constant in the
    separation $n$ for all local observables $X$ and $Y$.
\end{definition}

\begin{theorem}[CID implies RFP]\label{thm:isCID_implies_isRFP}
    \lean{MPSTensor.isCID_implies_isRFP}
    \leanok
    \uses{def:is_cid, def:is_rfp, def:transfer_map}
    If an MPS tensor admits a positive definite right fixed point of its
    transfer map and has correlations independent of distance, then
    its transfer map is idempotent.
    This is the reverse direction of \cite[Theorem~3.8]{Cirac2021Matrix}.
\begin{proof}\leanok
    Trace nondegeneracy and invertibility of the fixed point reduce
    idempotence to equality of the connected correlator at separations
    $n=2$ and $n=1$.
\end{proof}
\end{theorem}

\begin{theorem}[Zero correlation length iff idempotent transfer]%
    \label{thm:zcl_iff_idempotent_transfer}
    \lean{MPSTensor.zcl_iff_idempotent_transfer}
    \leanok
    \uses{def:is_rfp, def:is_cid, def:is_locally_orthogonal}
    An MPS tensor has zero correlation length (i.e.\ is both locally
    orthogonal and CID) if and only if its transfer map is idempotent.
    This is \cite[Theorem~3.8]{Cirac2021Matrix}.
\begin{proof}\leanok
    The forward direction extracts local orthogonality (which equals
    idempotence by definition). The reverse uses
    $\E^n = \E$ for $n \ge 1$ to show the connected correlator is
    separation-independent.
\end{proof}
\end{theorem}
