\section{Simple local structure from SAL and ZCL}

For the simple MPDO case in Appendix~C.2 of \cite{Cirac2016MPDO_arXiv},
the strong-area-law hypothesis is used locally through the equality case of
strong subadditivity for a normalized three-site reduced state.

\begin{definition}[Local $\eta$-structure for a three-site reduced state]
    \label{def:mpdo_eta_structure}
    \lean{MPOTensor.EtaStructure}
    \leanok
    \uses{def:entropy_quantum_markov_decomposition}
    For a three-site density operator $\rho_{ABC}$, the local $\eta$-structure
    is the quantum Markov decomposition on the middle subsystem $B$ produced by
    equality in strong subadditivity. Concretely, after a unitary change of
    basis on $B$, one has a finite direct-sum decomposition
    \[
        B = \bigoplus_k (b_1^{(k)} \otimes b_2^{(k)})
    \]
    such that
    \[
        \rho_{ABC} = \bigoplus_k \rho_{A b_1}^{(k)} \otimes \rho_{b_2 C}^{(k)}.
    \]
\end{definition}

\begin{theorem}[SAL implies local $\eta$-structure]
    \label{thm:sal_implies_eta_structure}
    \lean{MPOTensor.sal_implies_eta_structure}
    \leanok
    \uses{def:ssa_equality, def:mpdo_eta_structure}
    If a normalized three-site reduced state satisfies the equality case of
    strong subadditivity --- the local entropy form of the simple-MPDO strong
    area law --- then it admits the local $\eta$-structure of
    Definition~\ref{def:mpdo_eta_structure}.
\end{theorem}

\begin{proof}
    \leanok
    \uses{thm:entropy_ssa_equality_quantum_markov}
    This is exactly the forward implication of the Hayashi equality
    characterization: equality in strong subadditivity yields a quantum Markov
    decomposition on the middle subsystem.
\end{proof}

\begin{definition}[Injective inverse map for a simple MPO tensor]
    \label{def:mpdo_injective_inverse_layer}
    \lean{MPOTensor.IsInjective}
    \lean{MPOTensor.inverseTensor}
    \lean{MPOTensor.physRealize}
    \lean{MPOTensor.physRealizeLeft}
    \leanok
    \uses{def:mpo_tensor, def:mpo_to_mps, def:injective, def:decomposition_map, def:phys_realize}
    A simple MPO tensor $K$ is called injective if its doubled-index MPS tensor
    is injective. For such a tensor, the chosen decomposition map of the
    doubled-index MPS furnishes a concrete inverse tensor $K^{-1}$ together
    with right- and left-physical realization maps that turn virtual bond
    insertions back into local physical operators.
\end{definition}

\begin{theorem}[Inverse-map identities for an injective simple MPO tensor]
    \label{thm:mpdo_injective_inverse_layer_spec}
    \lean{MPOTensor.inverseTensor_spec}
    \lean{MPOTensor.physRealize_spec}
    \lean{MPOTensor.physRealize_mul}
    \lean{MPOTensor.physRealizeLeft_spec}
    \leanok
    \uses{def:mpdo_injective_inverse_layer}
    For an injective simple MPO tensor, contracting $K^{-1}$ with $K$ recovers
    the matrix units on the virtual bond space. Likewise, every right virtual
    insertion admits a physical realization, this realization is multiplicative,
    and the analogous left-insertion statement also holds.
\end{theorem}

\begin{definition}[Explicit neighboring $\eta$-operator data]
    \label{def:mpdo_explicit_eta_operators}
    \lean{MPOTensor.etaOperators}
    \lean{MPOTensor.ExplicitEtaOperators}
    \leanok
    \uses{def:mpdo_eta_structure}
    Fix a Hayashi decomposition witness $h_\eta$. An explicit neighboring
    $\eta$-family is a dependent family $(\eta_{k,h})$ indexed by sector pairs,
    where $\eta_{k,h}$ acts on the neighboring bond space
    $B_k^{R} \otimes B_h^{L}$. Requiring positivity for each pair gives the
    corresponding positive $\eta$-data.
\end{definition}

\begin{definition}[Trace matrices of explicit $\eta$-data]
    \label{def:mpdo_explicit_eta_trace_matrix}
    \lean{MPOTensor.ExplicitEtaOperators.traceMatrix}
    \lean{MPOTensor.ExplicitEtaOperators.traceMatrix_apply}
    \lean{MPOTensor.ExplicitEtaOperators.traceMatrixRe}
    \lean{MPOTensor.ExplicitEtaOperators.traceMatrixRe_apply}
    \leanok
    \uses{def:mpdo_explicit_eta_operators}
    Every explicit neighboring $\eta$-family determines a sector-by-sector trace
    matrix. We give both its complex-valued form and the real-part version,
    which is the direct data for the real Perron--Frobenius matrix $T$ used
    in the rank-one step.
\end{definition}


\begin{theorem}[Positivity of bipartite partial traces]
    \label{thm:matrix_possemidef_partial_traces}
    \lean{Matrix.PosSemidef.traceLeft}
    \lean{Matrix.PosSemidef.traceRight}
    \leanok
    If $\rho$ is positive semidefinite on a bipartite space $A \otimes B$, then
    both reduced operators $\tr_A \rho$ and $\tr_B \rho$ are positive
    semidefinite.
\end{theorem}

\begin{proof}
    \leanok
    Write each partial trace as the finite sum, over the traced index, of a
    principal submatrix of $\rho$. Principal submatrices of a positive
    semidefinite matrix are positive semidefinite, and finite sums of positive
    semidefinite matrices remain positive semidefinite.
\end{proof}

\begin{definition}[Hayashi--Markov extraction of explicit $\eta$-operators]
    \label{def:mpdo_eta_of_hayashi_markov}
    \lean{MPOTensor.ExplicitEtaOperators.ofHayashiMarkov}
    \leanok
    \uses{def:mpdo_eta_structure, def:mpdo_explicit_eta_operators,
        thm:matrix_possemidef_partial_traces}
    Every Hayashi decomposition witness $h_\eta$ canonically produces an
    explicit neighboring $\eta$-family by Kronecker-multiplying the
    sector-reduced states on the neighboring bond spaces:
    \[
        \eta_{k,h}
        := \tr_C(\rho_{b_2 C}^{(k)})
        \otimes \tr_A(\rho_{A b_1}^{(h)}).
    \]
    Positive semidefiniteness of each $\eta_{k,h}$ follows from the fact
    that partial traces preserve positivity and that positive
    semidefiniteness is closed under Kronecker products. This is the
    sector-reduced extraction; the remaining connection to
    \cite[Appendix~C.2]{Cirac2016MPDO_arXiv} is the identification of this
    family with the inverse-map
    construction in the original simple-MPDO tensor coordinates.
\end{definition}

\begin{theorem}[Trace matrix of the Hayashi--Markov $\eta$-operators]
    \label{thm:mpdo_eta_of_hayashi_markov_trace_matrix}
    \lean{MPOTensor.ExplicitEtaOperators.traceMatrix_ofHayashiMarkov}
    \lean{MPOTensor.ExplicitEtaOperators.traceMatrixRe_ofHayashiMarkov}
    \leanok
    \uses{def:mpdo_eta_of_hayashi_markov,
        def:mpdo_explicit_eta_trace_matrix}
    The trace matrix induced by the extracted family is entrywise
    $T_{k,h} = 1$ in both its complex-valued form and its real-valued form,
    matching the normalization $T_{k,h} = a_k b_h$ used in
    $a_k = b_h = 1$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{def:mpdo_eta_of_hayashi_markov,
        def:mpdo_explicit_eta_trace_matrix}
    Each $\rho_{b_2 C}^{(k)}$ is a density matrix, hence has unit trace; the
    partial trace $\tr_C$ preserves the trace. The analogous statement holds
    for $\tr_A \rho_{A b_1}^{(h)}$. Multiplicativity of the trace under
    Kronecker products gives $T_{k,h} = 1 \cdot 1 = 1$, and taking real parts
    gives the real-valued statement.
\end{proof}

\begin{theorem}[Entrywise nonnegativity of the real trace matrix]
    \label{thm:mpdo_explicit_eta_trace_matrix_nonneg}
    \lean{MPOTensor.ExplicitEtaOperators.traceMatrixRe_nonneg}
    \leanok
    \uses{def:mpdo_explicit_eta_trace_matrix}
    Positivity of the neighboring operators gives entrywise nonnegativity of
    the real trace matrix. Matrix-level positive semidefiniteness of $T$ is a
    separate structural condition.
\end{theorem}

\begin{proof}
    \leanok
    The trace of a positive semidefinite neighboring operator is non-negative in
    the complex order. Taking real parts gives the stated entrywise
    nonnegativity.
\end{proof}

\begin{definition}[Auxiliary matrix predicates for the rank-one step]
    \label{def:matrix_trace_rankone_predicates}
    \lean{Matrix.TracePowersConstant}
    \lean{Matrix.HasRankOneFactorization}
    \lean{Matrix.PrimitiveTracePowersConstantImpliesRankOne}
    \leanok
    The auxiliary predicates state three matrix-theoretic ingredients used in
    the rank-one step of \cite[Appendix~C.2, Lemma~C.5]{Cirac2016MPDO_arXiv}:
    constant traces of all
    positive powers, existence of a rank-one factorization $T = a b^\top$, and
    the conditional Perron--Frobenius implication from primitivity together with
    constant trace powers to such a factorization.

    This conditional implication is not a globally provable theorem. As a
    universal statement over primitive non-negative matrices it is false; a
    concrete $3 \times 3$ counterexample appears in the archive, and the
    deviation from the source is documented in
    \path{docs/paper-gaps/cpgsv17_pf_rank_one.tex}. The
    corrected criterion proved here adds positive semidefiniteness and trace
    normalization, which provide the diagonalizability needed to turn the
    trace-power condition into a rank-one factorization.
\end{definition}

\begin{theorem}[Trace square of a Hermitian matrix]
    \label{thm:matrix_hermitian_trace_square_eigenvalues}
    \lean{Matrix.IsHermitian.trace_sq_eq_sum_eigenvalues_sq}
    \leanok
    Let $T$ be a finite Hermitian matrix over a real or complex scalar field,
    with Hermitian eigenvalues $\lambda_i$. Then
    \[
        \tr(T^2) = \sum_i \lambda_i^2.
    \]
\end{theorem}

\begin{proof}
    \leanok
    The spectral theorem writes $T$ as a unitary conjugate of the diagonal
    matrix of its Hermitian eigenvalues. Squaring commutes with this conjugation,
    and cyclicity of trace reduces the identity to the trace of the squared
    diagonal matrix.
\end{proof}

\begin{theorem}[PSD trace-power rank-one criterion]
    \label{thm:matrix_psd_trace_power_rank_one}
    \lean{Matrix.PosSemidef.trace_powers_constant_implies_rank_one}
    \lean{Matrix.primitive_trace_powers_constant_implies_rank_one_of_posSemidef}
    \leanok
    \uses{def:matrix_trace_rankone_predicates}
    Let $T$ be a finite real square matrix. If $T$ is positive semidefinite,
    normalized by $\tr(T)=1$, and has constant traces on all positive powers,
    then $T$ has a rank-one factorization. Consequently, under the same
    positive-semidefinite and trace-normalization hypotheses, the conditional
    rank-one criterion used in \cite[Lemma~C.5]{Cirac2016MPDO_arXiv} follows.
\end{theorem}

\begin{proof}
    \leanok
    \uses{def:matrix_trace_rankone_predicates, thm:matrix_hermitian_trace_square_eigenvalues}
    The spectral theorem diagonalizes the positive semidefinite matrix with
    non-negative real eigenvalues. The trace gives that their sum is $1$, while
    the second trace moment gives that the sum of their squares is also $1$.
    Hence exactly one eigenvalue is equal to $1$ and all others vanish, so the
    diagonal form is rank one. Unitary conjugation preserves the existence of an
    outer-product factorization.
\end{proof}

% Unfaithful: thm:sal_zcl_implies_rank_one_t mirrors the Lean theorem
% MPOTensor.sal_zcl_implies_rank_one_T, whose conditional Perron--Frobenius
% hypothesis is false as a universal statement (counterexample in
% TNLean/Archive/PerronFrobeniusRankOneCounterexample.lean); the source
% Lemma~C.5 derives the factorization from Perron--Frobenius fixed-point
% theory. Documented in docs/paper-gaps/cpgsv17_pf_rank_one.tex; eliminated by
% the PSD-corrected thm:psd_trace_powers_rank_one_t.
\begin{theorem}[SAL and ZCL imply rank-one $T$ (conditional form)]
    \label{thm:sal_zcl_implies_rank_one_t}
    \lean{MPOTensor.sal_zcl_implies_rank_one_T}
    \leanok
    \uses{def:matrix_trace_rankone_predicates}
    Let $T$ be a finite real square matrix. Assume that $T$ is primitive,
    that $\tr(T) = 1$, and that every positive power of $T$ has the same trace
    as $T$. If one further supplies the Perron--Frobenius hypothesis that
    this particular primitive non-negative matrix with constant trace powers is
    rank one, then
    \[
        T = a b^\top
    \]
    for some vectors $a$ and $b$, with normalization
    \[
        a \cdot b = 1.
    \]
\end{theorem}

\begin{proof}
    \leanok
    \uses{def:matrix_trace_rankone_predicates}
    The Perron--Frobenius hypothesis gives the factorization $T = a b^\top$.
    Taking traces and using $\tr(a b^\top) = a \cdot b$ converts the
    normalization $\tr(T) = 1$ into $a \cdot b = 1$.
\end{proof}

\begin{theorem}[PSD-corrected rank-one $T$ criterion]
    \label{thm:psd_trace_powers_rank_one_t}
    \lean{MPOTensor.sal_zcl_implies_rank_one_T_of_posSemidef}
    \leanok
    \uses{def:matrix_trace_rankone_predicates}
    Let $T$ be a finite real square matrix. Assume that $T$ is primitive,
    positive semidefinite, normalized by $\tr(T)=1$, and has constant traces on
    all positive powers. Then
    \[
        T = a b^\top
    \]
    for some vectors $a$ and $b$, with $a \cdot b = 1$.
\end{theorem}

\begin{proof}
    \leanok
    \uses{def:matrix_trace_rankone_predicates, thm:sal_zcl_implies_rank_one_t,
        thm:matrix_psd_trace_power_rank_one}
    The positive-semidefinite matrix criterion diagonalizes $T$, uses the first
    two trace moments to show that exactly one eigenvalue is nonzero, and hence
    supplies the auxiliary rank-one criterion. The conditional statement of
    \cite[Lemma~C.5]{Cirac2016MPDO_arXiv} then gives the factorization and
    trace normalization.
\end{proof}
