\section{Normal PEPS Fundamental Theorem}\label{sec:peps_normal_capstone}

The pieces now combine. The blocking hypotheses of a normal PEPS, that the
tensors are injective after blocking around every edge and that each site
separates two injective regions whose complements are also injective, are what
the injective theorem needs once union closure is in hand. The Fundamental
Theorem for normal PEPS, Theorem~\ref{thm:fundamentalTheorem_normalPEPS_connected},
follows, with the gauges unique up to a scalar, the formalization of
\cite[Theorem~3]{Molnar2018NormalPEPS}. The corollaries specialize it to closed
chains, translation-invariant matrix product states, and the square lattice.

\begin{definition}[Normal PEPS blocking hypotheses]%
    \label{def:peps_normalPEPSBlockingHypotheses}
    \lean{TNLean.PEPS.NormalPEPSBlockingHypotheses}
    \lean{TNLean.PEPS.NormalPEPSBlockingHypotheses.injective_chain_at_edge}
    \lean{TNLean.PEPS.%
        NormalPEPSBlockingHypotheses.endpoint_disjoint_cover_at_edge}
    \lean{TNLean.PEPS.%
        NormalPEPSBlockingHypotheses.endpoint_injective_disjoint_cover_at_edge}
    \lean{TNLean.PEPS.NormalPEPSBlockingHypotheses.mem_withSite_iff}
    \leanok
    \uses{def:peps_normalEdgeBlockingHypotheses,
        def:peps_normalOneSiteSeparationHypotheses}
    The general normal theorem assumes both the edge-centred blockings into
    three injective regions and the one-site comparison regions with injective
    complements. These two assumptions are collected as the normal PEPS
    blocking hypotheses; the same hypotheses give the corresponding edge and
    one-site consequences without restating the separate edge and one-site
    assumptions.
    \begin{center}
        \TNPEPSNormalBlockingHypotheses
    \end{center}
\end{definition}

\begin{theorem}[Square-lattice normal blocking hypotheses from margin covers]%
    \label{thm:peps_normalSquarePEPSBlockingHypotheses_of_marginCovers}
    \lean{TNLean.PEPS.normalSquarePEPSBlockingHypotheses_of_marginCovers}
    \lean{TNLean.PEPS.normalSquarePEPSBlockingHypotheses_edgeBlocking_of_marginCovers}
    \lean{TNLean.PEPS.normalSquarePEPSBlockingHypotheses_blockingData_of_marginCovers}
    \lean{TNLean.PEPS.%
        normalSquarePEPSBlockingHypotheses_oneSiteSeparation_of_marginCovers}
    \lean{TNLean.PEPS.%
        normalSquarePEPSBlockingHypotheses_injective_chain_of_marginCovers}
    \lean{TNLean.PEPS.%
        normalSquarePEPSBlockingHypotheses_endpoint_disjoint_cover_of_marginCovers}
    \lean{TNLean.PEPS.%
        normalSquarePEPSBlockingHypotheses_endpoint_injective_disjoint_cover_of_marginCovers}
    \lean{TNLean.PEPS.%
        normalSquarePEPSBlockingHypotheses_mem_withSite_iff_of_marginCovers}
    \leanok
    \uses{thm:peps_normalSquareTranslatedEdgeBlockingHypotheses_of_windows,
        def:peps_normalPEPSBlockingHypotheses,
        def:peps_normalOneSiteSeparationHypotheses}
    Given rectangular injectivity hypotheses, finite union closure, oriented
    margin-and-cover hypotheses at every edge, and one-site separation, the finite
    square lattice satisfies the general normal blocking hypotheses. The
    edge-blocking hypotheses are exactly the margin-cover construction, the
    one-edge datum at an edge is the datum supplied there, and the one-site
    separation is exactly the supplied one-site separation. Hence these hypotheses
    give the three injective regions, the endpoint-disjoint-cover facts at
    each edge, and the one-site membership formula as consequences of the same
    normal blocking hypotheses.
\end{theorem}
\begin{proof}\leanok
    Combine the edge-blocking hypotheses constructed from the margin-cover
    hypotheses with the given one-site separation hypotheses.
\end{proof}

\begin{theorem}[Fundamental Theorem for normal PEPS]%
    \label{thm:fundamentalTheorem_normalPEPS}
    \notready
    \uses{thm:peps_normalEdgeBlockingHypotheses_injectiveChain,
        def:peps_normalPEPSBlockingHypotheses,
        thm:peps_oneVertexComplementComparison,
        def:GaugeEquivPEPS}
    Suppose two normal PEPS generate the same state, can be blocked into
    three-site injective chains around every edge, and for every site admit
    two injective regions with injective complements that differ exactly at
    that site. Then their defining tensors are related by a local gauge, and
    the gauges are unique up to the scalar freedom allowed by the graph. This
    is the general normal-PEPS theorem stated after
    \cite[Theorem~3]{Molnar2018NormalPEPS}.
    A connected-graph form, with the blocking hypothesis stated as the
    source's proof uses it (one blocking shared by the two states, each
    distinguished edge the entire bond of its chain), is
    Theorem~\ref{thm:fundamentalTheorem_normalPEPS_connected}; its per-edge
    uniqueness clause is
    Theorem~\ref{thm:fundamentalTheorem_normalPEPS_gauge_unique}.
    The source's remark following the theorem, that for a translationally
    invariant system the gauges are also translationally invariant when the
    proportionality constants are not absorbed into them, is realized on the
    torus by
    Theorem~\ref{thm:fundamentalTheorem_normalTorusPEPS_unconditional}: its
    gauge family is carried onto every edge by the translations from one
    reference matrix per orientation class, and its scalar $\lambda$ is kept
    separate rather than absorbed.
\end{theorem}

\begin{definition}[Doubly injective regions]%
    \label{def:peps_regionInjectivityDataPair}
    \lean{TNLean.PEPS.regionInjectivityDataPair}
    \leanok
    \uses{def:peps_regionInjectivityData}
    For two region-injectivity assignments on the same vertex set, a region is
    doubly injective when it is injective for each of the two.  The general
    normal theorem blocks two states by one shared geometry, and its final
    comparison contracts both tensors over the same regions, so its blocking
    hypotheses are instantiated at the doubly injective assignment of the two
    tensors.
\end{definition}

\begin{theorem}[Equal bond dimensions from the normal blocking hypotheses]%
    \label{thm:peps_bondDim_eq_of_normalBlocking}
    \lean{TNLean.PEPS.bondDim_eq_of_normalBlocking}
    \leanok
    \uses{def:peps_normalPEPSBlockingHypotheses,
        def:peps_regionInjectivityDataPair}
    Let $A$ and $B$ be PEPS tensors on a finite simple graph with the same
    positive physical dimension and positive bond dimensions, generating the
    same state and satisfying the normal blocking hypotheses at the doubly
    injective assignment of the two tensors, with each distinguished edge the
    only edge of the graph joining its two adjacent blocks.  Then the bond
    dimensions of $A$ and $B$ agree on every edge.
    In the source the equality is part of the isomorphism lemma of
    \cite[Section~3]{Molnar2018NormalPEPS}: the insertion correspondence
    $X\mapsto Y$ on the chosen bond is an algebra isomorphism between the two
    full bond matrix algebras, so the bond dimensions on the two sides agree;
    the normal case inherits the argument through the blocked three-site
    chains.  The single-crossing condition is the source's blocking around an
    edge made explicit: the chosen edge is the entire bond between the first
    two parties of the blocked chain.
    \begin{proof}
        \leanok
        \uses{thm:peps_edgeBlockedThreeSiteInjective,
            thm:peps_edgeBlockedInsertionAlgebraIsomorphism}
        Fix an edge.  Blocking both networks into the shared three regions
        around it produces two three-site chains on the three-vertex complete
        graph whose super-sites carry the blocked-region weight families,
        injective by hypothesis.  Each coarse state coefficient is a positive
        merge-collapse constant, a product of bond dimensions over the edges
        not incident to each region, times the corresponding original state
        coefficient, so the two coarse states are proportional with ratio the
        two constants.  Multiplying one super-site of each chain by the other
        chain's constant yields chains with equal states, unchanged bond
        dimensions, and the super-site injectivities preserved.  The matched
        matrix insertions on the distinguished coarse bond of the two scaled
        chains form an algebra isomorphism between the two full bond matrix
        algebras
        (Theorem~\ref{thm:peps_edgeBlockedInsertionAlgebraIsomorphism}), and
        an isomorphism of full matrix algebras forces equal matrix sizes.
        Under the single-crossing hypothesis the coarse bond carries the
        original bond dimension of the distinguished edge for each tensor, so
        the two original bond dimensions agree.
    \end{proof}
\end{theorem}

\begin{theorem}[Fundamental Theorem for normal PEPS on a connected graph]%
    \label{thm:fundamentalTheorem_normalPEPS_connected}
    \lean{TNLean.PEPS.fundamentalTheorem_normalPEPS}
    \leanok
    % Reading note: this is the general normal theorem stated after Theorem 3
    % of \cite{Molnar2018NormalPEPS} on a connected graph.  The shared
    % blockings and the single-crossing hypothesis are the source's blocking
    % hypothesis made explicit, not restrictions; both readings are recorded
    % in \path{docs/paper-gaps/peps_normal_ft_section3_route.tex}.
    \uses{def:peps_normalPEPSBlockingHypotheses,
        def:peps_regionInjectivityDataPair, def:GaugeEquivPEPS}
    Let $A$ and $B$ be PEPS tensors on a connected finite simple graph with
    the same positive physical dimension and positive bond dimensions,
    generating the same state.  Suppose the normal blocking
    hypotheses hold at the doubly injective assignment of the two tensors:
    around every edge the network blocks into three doubly injective regions
    with the edge endpoints in the first two, and every site admits two doubly
    injective regions, with doubly injective complements, differing exactly at
    that site.  Suppose moreover that at every edge the distinguished edge is
    the only edge of the graph joining its two adjacent blocks.  Then the
    defining tensors are related by a local gauge: there are invertible
    matrices, one on each edge, whose endpoint actions transform $A_v$ into
    $B_v$ at every vertex, with no leftover scalar.
    This is the general normal-PEPS theorem stated after
    \cite[Theorem~3]{Molnar2018NormalPEPS}
    (Theorem~\ref{thm:fundamentalTheorem_normalPEPS}), with the source's
    blocking hypothesis stated explicitly as its proof uses it, in two ways,
    neither of which restricts the source statement.  First, the blockings
    are shared between the two states, each region injective for both
    tensors: the source blocks both states by one blocking (``\emph{they}
    can be blocked''), its isomorphism lemma takes the two blocked chains
    over the same tripartition with all six blocks injective, and its final
    comparison contracts both tensors over the same one-site-different
    regions, whose injectivity for both tensors its comparison lemma
    requires.  Second, the single-crossing hypothesis is the explicit
    statement of the source's blocking \emph{around} an edge: the chosen edge
    is the entire bond between the first two parties of the three-site chain.
    In the source, blocking around an edge groups all vertices except the two
    endpoints of the edge, so the edge is the bond between the first two
    parties; the regions in the proof of
    \cite[Theorem~3]{Molnar2018NormalPEPS} enlarge those parties to injective
    blocks meeting only along the distinguished edge, and the gauge of the
    isomorphism lemma lives on that edge of the original network. The
    equality of the bond dimensions is not assumed; it is derived from the
    blocking hypotheses
    (Theorem~\ref{thm:peps_bondDim_eq_of_normalBlocking}), as in the source's
    isomorphism lemma. The
    positive bond dimensions and the connectivity of the graph are the
    counterexample-backed corrections carried over from the injective
    Fundamental Theorem (Theorem~\ref{thm:fundamentalTheorem_PEPS}); the
    absorption of the per-vertex constants into the gauges is exactly the step
    that requires connectivity.
    \begin{proof}
        \leanok
        \uses{thm:peps_bondDim_eq_of_normalBlocking,
            lem:peps_injectiveRegion_union_pos,
            thm:peps_region_edge_gauge_of_coeff_transfer,
            thm:gaugeConsistency}
        The bond dimensions of the two tensors agree on every edge by
        Theorem~\ref{thm:peps_bondDim_eq_of_normalBlocking}.
        The union closure of injective regions for either tensor follows from
        the positive bond dimensions and
        Lemma~\ref{lem:peps_injectiveRegion_union_pos}.  At every edge the
        shared blocking datum restricts to a datum for each tensor over the
        same three regions, so the per-edge gauge of the region insertion
        transfer (Theorem~\ref{thm:peps_region_edge_gauge_of_coeff_transfer})
        and its orientation-adapted absorption give a family of invertible
        matrices realizing the absorbed inserted-coefficient equality at every
        edge.  A gauge preserves blocked-region linear independence, so at
        every site the one-site comparison applies to the gauge-absorbed
        tensor: comparing the smaller region with its one-site completion
        yields two proportionality scalars whose ratio is a nonzero scalar
        $\lambda_v$ with $A_v$ equal to $\lambda_v$ times the gauge action on
        $B$ at $v$.  When the smaller region is empty its block is the count
        of global virtual configurations, and when the completion is the full
        vertex set its block is the closed state coefficient; in these
        degenerate cases the proportionality holds with scalar one, and every
        other comparison region has a boundary edge by connectivity.  The
        per-vertex relations against the nonzero closed state force
        $\prod_v \lambda_v = 1$, so on the connected graph a spanning-tree
        construction realizes the reciprocals of the scalars as oriented
        incidence products of edge scalars, which fold into the gauges as in
        the absorption step of the injective Fundamental Theorem
        (Theorem~\ref{thm:gaugeConsistency}).
    \end{proof}
\end{theorem}

\begin{theorem}[Uniqueness of the normal-PEPS gauges up to a constant]%
    \label{thm:fundamentalTheorem_normalPEPS_gauge_unique}
    \lean{TNLean.PEPS.fundamentalTheorem_normalPEPS_gauge_unique}
    \leanok
    \uses{def:peps_normalPEPSBlockingHypotheses,
        def:peps_regionInjectivityDataPair, def:gaugeVertex}
    Let $A$ and $B$ be PEPS tensors on a finite simple graph with equal bond
    dimensions, the bond dimensions of the first tensor positive, satisfying
    the normal blocking hypotheses at the doubly injective assignment of the
    two tensors.  Suppose two families of invertible matrices, one on each
    edge, each realize the scalar-free gauge relation of
    Theorem~\ref{thm:fundamentalTheorem_normalPEPS_connected}: at every vertex
    the endpoint actions of the family transform $A_v$ into $B_v$.  Then the
    two families are proportional at every edge: $X'_e = c\cdot X_e$ for a
    nonzero scalar $c$ depending on the edge.  This is the uniqueness clause
    of the general normal-PEPS theorem stated after
    \cite[Theorem~3]{Molnar2018NormalPEPS}, read against the scalar-free
    relation; neither equality of the two states nor connectivity is needed.
    \begin{proof}
        \leanok
        \uses{thm:peps_gaugeConj_eq_of_coeffIdentities,
            lem:peps_injectiveRegion_union_pos}
        The inserted-edge coefficient is multilinear in the site tensors, one
        factor per site, so each scalar-free per-vertex relation makes every
        inserted coefficient of $B$ equal the corresponding inserted
        coefficient of the gauge-absorbed $A$: each family realizes the
        absorbed inserted-coefficient equality at every edge, with the roles
        of the two tensors exchanged.  On the first block of the edge's frame,
        whose blocked map and complement blocked map are injective for $A$ by
        the blocking hypotheses and union closure
        (Lemma~\ref{lem:peps_injectiveRegion_union_pos}), the region insertion
        transfer is determined by the absorbed equality
        (Theorem~\ref{thm:peps_gaugeConj_eq_of_coeffIdentities}), so the two
        orientation-adapted absorbing gauges induce the same conjugation and
        differ by a nonzero scalar; unwinding the orientation adaptation gives
        the proportionality of the gauges themselves.
    \end{proof}
\end{theorem}

\begin{theorem}[Cycle blocking hypotheses from arc injectivity]%
    \label{thm:peps_cycleNormalPEPSBlockingHypotheses}
    \lean{TNLean.PEPS.cycleArcFrom}
    \lean{TNLean.PEPS.NormalCycleArcInjectivityHypotheses}
    \lean{TNLean.PEPS.cycleEdgeBlockingData}
    \lean{TNLean.PEPS.cycleOneSiteSeparation}
    \lean{TNLean.PEPS.cycleNormalPEPSBlockingHypotheses}
    \lean{TNLean.PEPS.isCrossingEdge_cycleEdge_iff}
    \leanok
    \uses{def:peps_normalPEPSBlockingHypotheses,
        def:peps_regionInjectivityData}
    A closed chain of $n$ sites is the cycle graph on $n$ vertices, and a
    block of consecutive sites is an arc: the vertices reached clockwise
    from a starting site.  Suppose $0 < L$, $3L \le n$, every arc
    of $L$ consecutive sites is injective for a region-injectivity
    assignment, and injectivity is closed under unions.  Then the chain
    satisfies the normal blocking hypotheses: around every edge it blocks
    into the $L$ sites on one side, the $L$ sites on the other, and the
    remaining $n - 2L \ge L$ sites, all injective and covering the chain;
    and at every site the $L + 1$ sites starting there and the $L$ sites
    starting at its successor are injective regions with injective
    complements differing exactly at that site.  Moreover the only
    nearest-neighbour pair joining the two blocks around an edge is the edge
    itself, since the far ends of the two blocks are separated by the
    remaining $n - 2L \ge L$ sites.
\end{theorem}
\begin{proof}\leanok
    Membership in an arc is a cyclic-distance inequality, so the
    disjointness, covering, and unique-crossing claims are modular
    arithmetic on the distances.  Arcs of length between $L$ and $n$ are
    unions of arcs of length $L$, hence injective by union closure; this
    covers the two complements at a site and the third block around an
    edge.
\end{proof}

\begin{corollary}[Fundamental Theorem for normal MPS on a closed chain]%
    \label{cor:fundamentalTheorem_normalMPS_cycle}
    \lean{TNLean.PEPS.fundamentalTheorem_normalMPS_cycle}
    \leanok
    \uses{def:peps_regionInjectivityDataOf, def:GaugeEquivPEPS}
    Let $A$ and $B$ be tensors on the closed chain of $n$ sites (the cycle
    graph on $n$ vertices, one tensor per site with two virtual legs), with
    the same positive physical dimension and positive bond dimensions, and
    let $0 < L$ with $n \ge 3L$.  Suppose blocking any $L$ consecutive sites
    gives an injective tensor, for $A$ and for $B$, and the two MPS generate
    the same state.  Then the defining tensors are related by a local gauge:
    there are invertible matrices, one on each bond, whose endpoint actions
    transform $A_i$ into $B_i$ at every site with no leftover scalar.  This
    is the source's family $Z_i$ ($i = 1, \dots, n$, $n + 1 \equiv 1$) with
    $B_i = Z_i^{-1} A_i Z_{i+1}$, the first corollary following the general
    normal-PEPS theorem of \cite[Section~3]{Molnar2018NormalPEPS}, the
    statement for MPS without translation invariance at a fixed system size.
    The positive physical and bond dimensions are the counterexample-backed
    corrections carried over from the injective Fundamental Theorem
    (Theorem~\ref{thm:fundamentalTheorem_PEPS}), and $0 < L$ is implicit in
    the source's blocking of $L$ consecutive sites; the equality of the bond
    dimensions is not assumed but derived, and the connectivity of the cycle
    is proved, not assumed.
\end{corollary}
\begin{proof}\leanok
    \uses{thm:peps_cycleNormalPEPSBlockingHypotheses,
        thm:fundamentalTheorem_normalPEPS_connected,
        lem:peps_injectiveRegion_union_pos}
    Positive bond dimensions give union closure of injective regions for
    each tensor (Lemma~\ref{lem:peps_injectiveRegion_union_pos}), hence for
    the doubly injective assignment.  The arc-injectivity hypothesis then
    yields the normal blocking hypotheses on the cycle with the unique
    crossing property
    (Theorem~\ref{thm:peps_cycleNormalPEPSBlockingHypotheses}), the cycle is
    connected, and the Fundamental Theorem for normal PEPS on a connected
    graph (Theorem~\ref{thm:fundamentalTheorem_normalPEPS_connected})
    supplies the scalar-free local gauge.
\end{proof}

\begin{corollary}[Uniqueness of the closed-chain gauges up to a constant]%
    \label{cor:fundamentalTheorem_normalMPS_cycle_gauge_unique}
    \lean{TNLean.PEPS.fundamentalTheorem_normalMPS_cycle_gauge_unique}
    \leanok
    \uses{def:peps_regionInjectivityDataOf, def:gaugeVertex}
    In the setting of
    Corollary~\ref{cor:fundamentalTheorem_normalMPS_cycle}, with the bond
    dimensions of the two tensors equal and positive and the arcs of $L$
    consecutive sites injective for each tensor, any two families of
    invertible bond matrices realizing the scalar-free gauge relation are
    proportional at every bond: $Z'_i = c \cdot Z_i$ for a nonzero scalar
    $c$ depending on the bond.  This is the uniqueness clause of the first
    corollary following the general normal-PEPS theorem of
    \cite[Section~3]{Molnar2018NormalPEPS}: the gauges $Z_i$ are unique up
    to a multiplicative constant.  Neither equality of the two states nor
    connectivity is needed, as in the general uniqueness clause.
\end{corollary}
\begin{proof}\leanok
    \uses{thm:peps_cycleNormalPEPSBlockingHypotheses,
        thm:fundamentalTheorem_normalPEPS_gauge_unique,
        lem:peps_injectiveRegion_union_pos}
    The arc-injectivity hypothesis and union closure from the positive bond
    dimensions (Lemma~\ref{lem:peps_injectiveRegion_union_pos}) assemble the
    normal blocking hypotheses on the cycle
    (Theorem~\ref{thm:peps_cycleNormalPEPSBlockingHypotheses}), and the
    per-edge uniqueness of the general theorem
    (Theorem~\ref{thm:fundamentalTheorem_normalPEPS_gauge_unique}) gives the
    proportionality at every bond.
\end{proof}

\begin{definition}[Cycle tensor of a matrix tensor]%
    \label{def:cycleTensorOfMPS}
    \lean{TNLean.PEPS.cycleSuccEdge}
    \lean{TNLean.PEPS.cycleLeftIncident}
    \lean{TNLean.PEPS.cycleRightIncident}
    \lean{TNLean.PEPS.cycleTensorOfMPS}
    \leanok
    \uses{def:peps_tensor, def:mps_tensor}
    Every edge of the cycle graph on $n \ge 3$ vertices joins a site to its
    cyclic successor, so sites and bonds are in bijection, and every site has
    exactly two incident bonds: the bond shared with its cyclic predecessor
    and the bond shared with its cyclic successor.  The \emph{cycle tensor}
    of a tensor $A = (A^i)_{i=0}^{d-1}$ of $D \times D$ matrices places one
    copy of $A$ at every site of the closed chain of $n$ sites, every bond of
    dimension $D$, with the row index of $A^i$ read from the bond shared with
    the predecessor and the column index from the bond shared with the
    successor.
\end{definition}

\begin{lemma}[State coefficients of the cycle tensor]%
    \label{lem:stateCoeff_cycleTensorOfMPS}
    \lean{MPSTensor.evalWord_ofFn_apply}
    \lean{MPSTensor.trace_evalWord_eq_sum_cyclic}
    \lean{TNLean.PEPS.stateCoeff_cycleTensorOfMPS}
    \leanok
    \uses{def:cycleTensorOfMPS, def:peps_stateCoeff, def:mpv}
    For $n \ge 3$ and every configuration
    $\sigma = (i_1, \ldots, i_n)$, the state coefficient of the cycle
    tensor of $A$ at $\sigma$ is the matrix product vector component
    $V^{(n)}(A)_\sigma = \tr\!(A^{i_1} \cdots A^{i_n})$.
\end{lemma}
\begin{proof}\leanok
    An entry of a word product is a sum over paths of bond indices pinned at
    the two ends of products of matrix entries read along the path, by
    induction on the word; closing the two pinned ends into a loop, the trace
    of a closed word product is the sum over all cyclic bond configurations
    of the products of entries around the loop.  Contracting the cycle tensor
    sums the per-site entries over all assignments of bond indices, which is
    that cyclic sum after matching each bond with the site on its
    predecessor side.
\end{proof}

\begin{lemma}[Arc injectivity of the cycle tensor from block injectivity]%
    \label{lem:regionBlockedTensorInjective_cycleTensorOfMPS}
    \lean{TNLean.PEPS.isRegionBoundaryEdge_cycleArcFrom_iff}
    \lean{TNLean.PEPS.regionBlockedWeight_cycleTensorOfMPS}
    \lean{TNLean.PEPS.matrix_eq_zero_of_isNBlkInjective}
    \lean{TNLean.PEPS.regionBlockedTensorInjective_cycleTensorOfMPS}
    \leanok
    \uses{def:cycleTensorOfMPS, def:l_blk_injective,
        def:peps_regionInjectivityDataOf}
    Let $0 < L < n$ with $n \ge 3$, let the bond dimension be positive, and
    suppose $A$ is $L$-block injective.  Then for every arc of $L$
    consecutive sites of the closed chain, the blocked tensor of the arc of
    the cycle tensor of $A$ is injective.
\end{lemma}
\begin{proof}\leanok
    An arc of $L < n$ consecutive sites has exactly two boundary-crossing
    bonds, the bond entering its first site and the bond leaving its last
    site, by cyclic-distance arithmetic.  The blocked weight of the arc at a
    boundary assignment $(a, b)$ and a physical word $x$ is
    $D^{\,n - (L+1)}\,(A^{x_1} \cdots A^{x_L})_{a b}$: the bonds inside
    the arc contract to the matrix product, the two boundary bonds stay
    pinned, and each bond away from the arc contributes a free factor $D$.
    A linear relation among the blocked components over the boundary pairs
    is a matrix $C$ with $\tr\!(C^{\mathsf T} A^{x_1} \cdots A^{x_L}) = 0$
    for every word $x$; the word products span the full matrix algebra by
    $L$-block injectivity, and the trace pairing is nondegenerate, so
    $C = 0$.
\end{proof}

\begin{corollary}[Fundamental Theorem for translation-invariant normal MPS
        on a closed chain, matrix form]%
    \label{cor:fundamentalTheorem_normalMPS}
    \lean{TNLean.PEPS.fundamentalTheorem_normalMPS}
    \leanok
    \uses{def:l_blk_injective, def:mpv}
    Let $n$ be positive, let $0 < L$ with $3L \le n$, and let $A$ and $B$ be
    tensors of $D \times D$ matrices over physical dimension $d$, each
    $L$-block injective, whose matrix product vectors agree at system size
    $n$: $V^{(n)}(A)_\sigma = V^{(n)}(B)_\sigma$ for every configuration
    $\sigma$.  Then there are invertible matrices $Z_v$, one per bond of the
    closed chain ($v = 1, \ldots, n$, $n + 1 \equiv 1$), with
    \[
        B^i = Z_v^{-1}\, A^i\, Z_{v+1}
        \qquad\text{for every site } v \text{ and every } i .
    \]
    This is the first corollary following the general normal-PEPS theorem of
    \cite[Section~3]{Molnar2018NormalPEPS} specialized to one
    site-independent tensor: the source states the corollary for
    site-dependent families, and its conclusion is the per-bond gauge family
    delivered here.  The source's translation-invariant corollary (a single
    gauge $Z$ and a constant $\lambda$ with $\lambda^n = 1$) refines this
    conclusion and is not part of this statement.  Positivity of the
    physical and bond dimensions is not assumed: a vanishing bond dimension
    makes the conclusion trivial, and a vanishing physical dimension
    contradicts block injectivity.
\end{corollary}
\begin{proof}\leanok
    \uses{cor:fundamentalTheorem_normalMPS_cycle,
        lem:stateCoeff_cycleTensorOfMPS,
        lem:regionBlockedTensorInjective_cycleTensorOfMPS,
        def:cycleTensorOfMPS, def:gaugeVertex}
    The cycle tensors of $A$ and $B$ generate the same state by
    Lemma~\ref{lem:stateCoeff_cycleTensorOfMPS}, and every arc of
    $L$ consecutive sites is injective for each by
    Lemma~\ref{lem:regionBlockedTensorInjective_cycleTensorOfMPS}, so the
    closed-chain corollary
    (Corollary~\ref{cor:fundamentalTheorem_normalMPS_cycle}) produces one
    invertible matrix per edge relating the two cycle tensors.  At each site
    the gauge action contracts the two incident matrices against the site
    matrix, with the matrix of the entering bond acting by its
    transposed inverse when the site is the stored second endpoint; reading
    the per-site identity through the two incident bonds turns the per-edge
    family into per-bond gauges, the transpose of the edge matrix away from
    the seam and the inverse on the seam edge, whose conjugation relation is
    exactly $B^i = Z_v^{-1} A^i Z_{v+1}$.
\end{proof}

\begin{corollary}[Uniqueness of the per-bond gauges up to one constant,
        matrix form]%
    \label{cor:fundamentalTheorem_normalMPS_gauge_unique}
    \lean{TNLean.PEPS.fundamentalTheorem_normalMPS_gauge_unique}
    \leanok
    \uses{def:l_blk_injective}
    Let $n$ be positive, let $0 < L$ with $3L \le n$, and let $A$ and $B$ be
    tensors of $D \times D$ matrices, each $L$-block injective.  If two
    families $Z$ and $Z'$ of invertible matrices on the bonds of the closed
    chain both satisfy $B^i = Z_v^{-1} A^i Z_{v+1}$ at every site, then they
    are proportional by a single constant: there is a nonzero scalar $c$
    with $Z'_v = c\,Z_v$ at every bond.  This is the uniqueness clause of
    the first corollary following the general normal-PEPS theorem of
    \cite[Section~3]{Molnar2018NormalPEPS}, for one site-independent
    tensor.  Equality of the two states is not needed.
\end{corollary}
\begin{proof}\leanok
    \uses{cor:fundamentalTheorem_normalMPS_cycle_gauge_unique,
        lem:regionBlockedTensorInjective_cycleTensorOfMPS,
        def:cycleTensorOfMPS, def:gaugeVertex}
    Each per-bond family converts back into a per-edge family — the
    transpose of the bond gauge away from the seam, the inverse on the seam
    edge — realizing the per-site gauge identity between the cycle tensors,
    so the closed-chain uniqueness clause
    (Corollary~\ref{cor:fundamentalTheorem_normalMPS_cycle_gauge_unique})
    gives a constant on every edge; undoing the conversion gives a constant
    per bond.  The relation $B^i = Z_v^{-1} A^i Z_{v+1}$ for both families
    forces the scalar $\mu_v^{-1}\mu_{v+1}$ to fix the nonzero tensor $B$,
    so consecutive constants agree, and going around the cycle merges them
    into one.
\end{proof}

\begin{lemma}[Consecutive per-bond gauges are proportional]%
    \label{lem:gaugeFamily_succ_proportional}
    \lean{TNLean.PEPS.gaugeFamily_succ_proportional}
    \leanok
    \uses{def:l_blk_injective, def:eval_word}
    Let $n$ be positive, let $A$ and $B$ be tensors of $D \times D$ matrices
    over physical dimension $d$ with $A$ being $L$-block injective, and let
    $Z_v$ ($v = 1, \ldots, n$, $n + 1 \equiv 1$) be invertible matrices with
    $B^i = Z_v^{-1} A^i Z_{v+1}$ at every site $v$ and every $i$.  Then for
    every site $v$ there is a nonzero scalar $c$ with $Z_{v+1} = c\,Z_v$.
\end{lemma}
\begin{proof}\leanok
    \uses{def:eval_word}
    Iterating the per-bond relation along a word $x$ of length $L$ from the
    starting site $v$ gives
    $B^{x_1} \cdots B^{x_L} = Z_v^{-1} A^{x_1} \cdots A^{x_L} Z_{v+L}$.
    Comparing the iterations from $v$ and from $v + 1$ over the word
    products $A^{x_1} \cdots A^{x_L}$, which span the full matrix algebra by
    $L$-block injectivity, the comparison extends linearly to every matrix
    $W$: $Z_v^{-1} W Z_{v+L} = Z_{v+1}^{-1} W Z_{v+1+L}$.  The empty word
    ($W$ the identity) pins the two bond transports
    $Z_v^{-1} Z_{v+L} = Z_{v+1}^{-1} Z_{v+1+L}$ to the same invertible
    matrix; cancelling it leaves $Z_v^{-1} W Z_v = Z_{v+1}^{-1} W Z_{v+1}$
    for every $W$, so $Z_{v+1} Z_v^{-1}$ commutes with the full matrix
    algebra, hence is a scalar, nonzero by invertibility.
\end{proof}

\begin{corollary}[Fundamental Theorem for translation-invariant normal MPS
        on a closed chain, single-gauge form]%
    \label{cor:fundamentalTheorem_normalMPS_translationInvariant}
    \lean{TNLean.PEPS.fundamentalTheorem_normalMPS_translationInvariant}
    \leanok
    \uses{def:l_blk_injective, def:mpv}
    Let $n$ be positive, let $0 < L$ with $3L \le n$, and let $A$ and $B$ be
    tensors of $D \times D$ matrices over physical dimension $d$, each
    $L$-block injective, whose matrix product vectors agree at system size
    $n$: $V^{(n)}(A)_\sigma = V^{(n)}(B)_\sigma$ for every configuration
    $\sigma$.  Then there are an invertible matrix $Z$ and a constant
    $\lambda$ with $\lambda^n = 1$ such that
    \[
        B^i = \lambda\, Z^{-1} A^i Z \qquad\text{for every } i .
    \]
    This is the translation-invariant corollary of
    \cite[Section~3]{Molnar2018NormalPEPS}, the statement for TI MPS
    following the first corollary after the general normal-PEPS theorem.
    Positivity of the bond dimension is not assumed: a vanishing bond
    dimension makes the conclusion trivial.
\end{corollary}
\begin{proof}\leanok
    \uses{cor:fundamentalTheorem_normalMPS, lem:gaugeFamily_succ_proportional}
    The matrix-form corollary
    (Corollary~\ref{cor:fundamentalTheorem_normalMPS}) gives per-bond gauges
    $Z_v$ with $B^i = Z_v^{-1} A^i Z_{v+1}$, and consecutive gauges are
    proportional (Lemma~\ref{lem:gaugeFamily_succ_proportional}):
    $Z_{v+1} = c_v Z_v$ with $c_v$ nonzero.  Multiplying the relation by
    $Z_v$ on the left gives $Z_v B^i = A^i Z_{v+1}$ at every site; comparing
    this at sites $v$ and $v+1$ through the proportionalities yields
    $c_v\,(Z_v B^i) = c_{v+1}\,(Z_v B^i)$, and $Z_v B^{i_0} \neq 0$ for some
    $i_0$ because $B$ is $L$-block injective with positive bond dimension,
    so all the scalars equal a single constant $\lambda := c_0$.  Iterating
    $Z_{v+1} = \lambda Z_v$ from the starting bond gives
    $Z_k = \lambda^k Z_0$, and after $n$ steps the closed chain returns to
    the starting bond: $Z_0 = \lambda^n Z_0$, so $\lambda^n = 1$ since
    $Z_0 \neq 0$.  The relation at the starting site then reads
    $B^i = \lambda\, Z_0^{-1} A^i Z_0$.
\end{proof}

\begin{corollary}[Uniqueness of the translation-invariant gauge up to a
        constant]%
    \label{cor:fundamentalTheorem_normalMPS_translationInvariant_gauge_unique}
    \lean{TNLean.PEPS.fundamentalTheorem_normalMPS_translationInvariant_gauge_unique}
    \leanok
    \uses{def:l_blk_injective}
    Let $0 < L$ and let $A$ and $B$ be tensors of $D \times D$ matrices,
    each $L$-block injective.  If invertible matrices $Z$, $Z'$ and
    constants $\lambda$, $\lambda'$ satisfy
    $B^i = \lambda\, Z^{-1} A^i Z$ and $B^i = \lambda'\, Z'^{-1} A^i Z'$ for
    every $i$, then there is a nonzero scalar $c$ with $Z' = c\,Z$.  This is
    the uniqueness clause of the translation-invariant corollary of
    \cite[Section~3]{Molnar2018NormalPEPS} for TI MPS.  No system size
    enters the statement: equality of the two states is not needed, nor are
    the root-of-unity conditions on $\lambda$ and $\lambda'$.
\end{corollary}
\begin{proof}\leanok
    \uses{def:eval_word}
    The scalar $\lambda$ is nonzero because $B$ is $L$-block injective with
    positive bond dimension (a vanishing bond dimension makes the
    conclusion trivial), so some $B^{i_0} \neq 0$.  Iterating each relation
    along a word $x$ of length $L$ gives
    $B^{x_1} \cdots B^{x_L} = \lambda^L\, Z^{-1} A^{x_1} \cdots A^{x_L} Z$
    and the primed analogue; the word products span the full matrix algebra
    by $L$-block injectivity, so
    $\lambda^L\, Z^{-1} W Z = \lambda'^L\, Z'^{-1} W Z'$ for every matrix
    $W$.  The identity pins $\lambda^L = \lambda'^L$; cancelling the
    nonzero factor leaves $Z^{-1} W Z = Z'^{-1} W Z'$ for every $W$, so
    $Z' Z^{-1}$ commutes with the full matrix algebra, hence is a nonzero
    scalar.
\end{proof}

\begin{lemma}[Block injectivity extends to longer blocks]%
    \label{lem:isNBlkInjective_of_le}
    \lean{MPSTensor.isNBlkInjective_of_le}
    \leanok
    \uses{def:l_blk_injective}
    Let $A$ be a tensor of $D \times D$ matrices over physical dimension
    $d$, $L$-block injective with $0 < L$.  Then $A$ is $m$-block injective
    for every $m \ge L$.  This is the chain form of the statement that any
    region of at least the blocking size is injective, derived in the
    closing section of \cite{Molnar2018NormalPEPS} (``Normal MPS:
    alternative proof'') from its union lemma.
\end{lemma}
\begin{proof}\leanok
    \uses{def:eval_word}
    It suffices to pass from length $m$ to $m + 1$.  Write the identity as
    a combination of length-$L$ word products and peel the leading letter
    of each: any matrix $P$ becomes a combination of terms
    $A^{i}\,(W P)$ with $W$ a word product of length $L - 1$.  Each factor
    $W P$ is a combination of length-$m$ word products by the induction
    hypothesis, so each term is a combination of length-$(m+1)$ word
    products.
\end{proof}

\begin{lemma}[Word transport between same-state networks]%
    \label{lem:exists_mpvTransport}
    \lean{MPSTensor.exists_mpvTransport}
    \lean{MPSTensor.exists_evalWordTransport}
    \leanok
    \uses{def:l_blk_injective, def:mpv, def:eval_word}
    Let $A$ and $B$ be tensors of $D \times D$ matrices over physical
    dimension $d$, let $n = k + q$, and suppose the length-$k$ and
    length-$q$ word products of $A$ each span the full matrix algebra.  If
    the matrix product vectors of $A$ and $B$ agree at system size $n$,
    then there is a linear map $\Lambda$ of the matrix algebra with
    $\Lambda(A^{x_1} \cdots A^{x_k}) = B^{x_1} \cdots B^{x_k}$ for every
    word $x$ of length $k$.  The same holds when, instead of the traces,
    the word products of the two tensors are themselves equal at length
    $n$.  This is the operator-transport mechanism underlying Lemma~5 of
    \cite{Molnar2018NormalPEPS}: an operator expanded over the window
    products of one network is re-read in the other.
\end{lemma}
\begin{proof}\leanok
    \uses{def:mpv, def:eval_word}
    Pair the matrix algebra against the complementary word products: let
    $\Psi_A(M)$ be the family of traces
    $\tr(M\,A^{y_1} \cdots A^{y_q})$ over the length-$q$ words $y$, and
    $\Psi_B$ its $B$-analogue.  $\Psi_A$ is injective because the length-$q$
    products span and the trace pairing is nondegenerate.  On the spanning
    family of length-$k$ products the two pairings match:
    $\Psi_A(A^{x}) = \Psi_B(B^{x})$, since both compute closed-chain
    coefficients of total length $n$, equal by hypothesis.  Hence the range
    of $\Psi_A$ lies in that of $\Psi_B$, forcing $\Psi_B$ to be injective
    by dimension count, and a left inverse of $\Psi_B$ composed with
    $\Psi_A$ is the transport.  For the matrix-product version replace the
    traces by the products themselves.
\end{proof}

\begin{lemma}[Bond-operator extraction from overlapping windows]%
    \label{lem:overlapWindow_exists_bondOperator}
    \lean{MPSTensor.overlapWindow_exists_bondOperator}
    \lean{MPSTensor.bondOperator_unique}
    \leanok
    \uses{def:l_blk_injective, def:eval_word}
    Let $0 < L$ with $2L + 1 \le n$, let $B$ be a tensor of $D \times D$
    matrices over physical dimension $d$, $L$-block injective, and let
    $C_0, \ldots, C_L$ be deformed window tensors: each $C_j$ assigns a
    matrix to every word of length $L$, read as replacing the blocked
    tensor of the $L$-site window starting $j$ sites left of a fixed bond
    of the closed chain on $n$ sites.  Suppose the deformed states agree
    for consecutive window positions: for every $j < L$, every prefix $p$
    of $j$ sites, every assignment $u$ on the $L + 1$ sites covered by the
    two windows, and every suffix $s$ on the remaining sites,
    \[
        \tr\bigl(B^{p}\, C_j(u_1 \cdots u_L)\, B^{u_{L+1}}\, B^{s}\bigr)
        = \tr\bigl(B^{p}\, B^{u_1}\, C_{j+1}(u_2 \cdots u_{L+1})\,
        B^{s}\bigr),
    \]
    where $B^{p}$ and $B^{s}$ denote the word products over the prefix and
    suffix.  Then the deformation is a virtual operation on the bond: there
    is a matrix $X$ with
    \[
        C_0(a) = B^{a_1} \cdots B^{a_L}\, X
        \quad\text{and}\quad
        C_L(b) = X\, B^{b_1} \cdots B^{b_L}
    \]
    for all length-$L$ words $a$, $b$, and $X$ is unique.  This is Lemma~5
    of \cite{Molnar2018NormalPEPS} (its closing section, ``Normal MPS:
    alternative proof''), specialized to one site-independent tensor; the
    source states it for site-dependent tensors on five sites with $L = 2$,
    with the deformed window tensors arising as physical operators
    contracted with the blocked windows they act on; the site-dependent
    form, in the setting in which the source states the lemma, is
    Lemma~\ref{lem:chain_overlapWindow_exists_bondOperator}, and this
    statement is its constant specialization.  The
    algebra-homomorphism clause of the source lemma is established with the
    insertion correspondence in
    Lemma~\ref{lem:exists_conjugation_of_mpv_eq}.
\end{lemma}
\begin{proof}\leanok
    \uses{lem:chain_overlapWindow_exists_bondOperator, def:eval_word}
    Place the tensor at every site of the site-dependent form
    (Lemma~\ref{lem:chain_overlapWindow_exists_bondOperator}): the arc
    products of the constant chain are the word products of the repeated
    tensor, and $L$-block injectivity of the tensor is window injectivity
    of the constant chain.
\end{proof}

\begin{lemma}[Conjugate word products at one length from Lemma~5]%
    \label{lem:exists_conjugation_of_mpv_eq}
    \lean{MPSTensor.exists_conjugation_of_mpv_eq}
    \leanok
    \uses{def:l_blk_injective, def:mpv, def:eval_word}
    Let $0 < L$ with $2L + 1 \le n$, and let $A$ and $B$ be tensors of
    $D \times D$ matrices over physical dimension $d$, each $L$-block
    injective, whose matrix product vectors agree at system size $n$.
    Then there is an invertible matrix $Z$ with
    \[
        B^{w_1} \cdots B^{w_n} = Z\, A^{w_1} \cdots A^{w_n}\, Z^{-1}
    \]
    for every word $w$ of length $n$.  This captures the insertion
    correspondence of Lemma~5 of \cite{Molnar2018NormalPEPS} and its
    algebra-homomorphism clause for one site-independent tensor: the map
    sending an insertion on one network to the corresponding insertion on
    the other is an automorphism of the matrix algebra, hence inner.
\end{lemma}
\begin{proof}\leanok
    \uses{lem:exists_mpvTransport, lem:overlapWindow_exists_bondOperator,
        lem:isNBlkInjective_of_le, thm:skolem_noether}
    A virtual insertion $X$ on a bond of the $A$-network is realized on
    each of the $L + 1$ overlapping windows around the bond by deformed
    window tensors placing $X$ after the first $L - j$ letters.
    Transporting them to the $B$-network by the word transport
    (Lemma~\ref{lem:exists_mpvTransport}) preserves all closed-chain
    coefficients, so the transported deformations generate one common
    state, and the bond-operator extraction
    (Lemma~\ref{lem:overlapWindow_exists_bondOperator}) produces
    $Y = \Phi(X)$ with $\Lambda(A^{a} X) = B^{a}\,\Phi(X)$ and
    $\Lambda(X A^{b}) = \Phi(X)\,B^{b}$ on all window words.  Uniqueness
    of the bond operator makes $\Phi$ linear, unital and multiplicative,
    hence an automorphism of the full matrix algebra, and by
    Skolem--Noether (Theorem~\ref{thm:skolem_noether})
    $\Phi(X) = Z X Z^{-1}$ for an invertible $Z$.  Pairing the insertions
    against all closed-chain words of length $n$ and using nondegeneracy of
    the trace pairing transfers the conjugation to the word products.
\end{proof}

\begin{lemma}[Rigidity of word products at one length]%
    \label{lem:proportional_of_evalWord_eq}
    \lean{MPSTensor.proportional_of_evalWord_eq}
    \leanok
    \uses{def:l_blk_injective, def:eval_word}
    Let $0 < L$ with $2L + 1 \le n$ and $0 < D$, and let $A$ and $C$ be
    tensors of $D \times D$ matrices over physical dimension $d$, each
    $L$-block injective, with equal word products at length $n$:
    $C^{w_1} \cdots C^{w_n} = A^{w_1} \cdots A^{w_n}$ for every word $w$ of
    length $n$.  Then there is a constant $\lambda$ with $\lambda^n = 1$
    and $C^i = \lambda A^i$ for every $i$.
\end{lemma}
\begin{proof}\leanok
    \uses{lem:exists_mpvTransport, lem:isNBlkInjective_of_le,
        def:eval_word}
    Transport word products of $A$ to those of $C$ at the window length
    $L$, at $L + 1$, at the complementary length $n - 1 - L$, and at
    $n - L$ (Lemma~\ref{lem:exists_mpvTransport}, matrix-product version;
    the lengths lie between $L$ and $n - L$ by $2L + 1 \le n$).  For a
    transport pair at lengths $k$ and $k + 1$, the transported identity
    $\Lambda_k(I)$ commutes with every letter of $C$ — its two one-letter
    extensions agree, $C^i\,\Lambda_k(I) = \Lambda_{k+1}(A^i) =
        \Lambda_k(I)\,C^i$ — hence with the full matrix algebra by block
    injectivity, so it is a scalar $c_k I$, nonzero because the transport
    is surjective onto the spanning word products, hence bijective.
    Peeling one letter off the full-length equality through the spanning
    windows of lengths $L$ and $n - 1 - L$ then leaves
    $C^i = (c_L\,c_{n-1-L})^{-1} A^i =: \lambda A^i$; evaluating any
    nonzero length-$n$ word product gives $\lambda^n = 1$.
\end{proof}

\begin{corollary}[Fundamental Theorem for translation-invariant normal MPS
        at $n \ge 2L + 1$, single-gauge form]%
    \label{cor:fundamentalTheorem_normalMPS_translationInvariant_of_overlap}
    \lean{TNLean.PEPS.fundamentalTheorem_normalMPS_translationInvariant_of_overlap}
    \leanok
    \uses{def:l_blk_injective, def:mpv}
    Let $0 < L$ with $2L + 1 \le n$, and let $A$ and $B$ be tensors of
    $D \times D$ matrices over physical dimension $d$, each $L$-block
    injective, whose matrix product vectors agree at system size $n$:
    $V^{(n)}(A)_\sigma = V^{(n)}(B)_\sigma$ for every configuration
    $\sigma$.  Then there are an invertible matrix $Z$ and a constant
    $\lambda$ with $\lambda^n = 1$ such that
    \[
        B^i = \lambda\, Z^{-1} A^i Z \qquad\text{for every } i .
    \]
    This is the strengthened closed-chain corollary of
    \cite{Molnar2018NormalPEPS} (the corollary after Lemma~5 in its closing
    section ``Normal MPS: alternative proof''), improving the system-size
    bound of
    Corollary~\ref{cor:fundamentalTheorem_normalMPS_translationInvariant}
    from $n \ge 3L$ to $n \ge 2L + 1$.  The uniqueness clause of the source
    corollary — the gauge $Z$ is unique up to a multiplicative constant —
    needs no system size and is
    Corollary~\ref{cor:fundamentalTheorem_normalMPS_translationInvariant_gauge_unique}.
    Positivity of the bond dimension is not assumed: a vanishing bond
    dimension makes the conclusion trivial.
\end{corollary}
\begin{proof}\leanok
    \uses{lem:exists_conjugation_of_mpv_eq, lem:proportional_of_evalWord_eq}
    By Lemma~\ref{lem:exists_conjugation_of_mpv_eq} the word products of
    $B$ at length $n$ are those of $A$ conjugated by an invertible matrix.
    Absorbing the conjugation into $B$ — which preserves $L$-block
    injectivity — leaves two tensors with equal word products at length
    $n$, and Lemma~\ref{lem:proportional_of_evalWord_eq} makes them
    proportional letterwise by an $n$-th root of unity.  Undoing the
    absorption gives $B^i = \lambda\, Z^{-1} A^i Z$.
\end{proof}

\begin{corollary}[Fundamental Theorem for translation-invariant normal MPS
        at $n \ge 2L + 1$, per-bond form]%
    \label{cor:fundamentalTheorem_normalMPS_of_overlap}
    \lean{TNLean.PEPS.fundamentalTheorem_normalMPS_of_overlap}
    \leanok
    \uses{def:l_blk_injective, def:mpv}
    Let $n$ be positive, let $0 < L$ with $2L + 1 \le n$, and let $A$ and
    $B$ be tensors of $D \times D$ matrices over physical dimension $d$,
    each $L$-block injective, whose matrix product vectors agree at system
    size $n$.  Then there are invertible matrices $Z_v$, one per bond of
    the closed chain ($v = 1, \ldots, n$, $n + 1 \equiv 1$), with
    \[
        B^i = Z_v^{-1}\, A^i\, Z_{v+1}
        \qquad\text{for every site } v \text{ and every } i .
    \]
    This is the per-bond gauge family of
    Corollary~\ref{cor:fundamentalTheorem_normalMPS}, delivered at the
    source's strengthened bound $n \ge 2L + 1$ in place of $n \ge 3L$
    (\cite{Molnar2018NormalPEPS}, the corollary after Lemma~5 in its
    closing section ``Normal MPS: alternative proof'').
\end{corollary}
\begin{proof}\leanok
    \uses{cor:fundamentalTheorem_normalMPS_translationInvariant_of_overlap}
    The single-gauge form
    (Corollary~\ref{cor:fundamentalTheorem_normalMPS_translationInvariant_of_overlap})
    gives $Z$ and $\lambda$ with $\lambda^n = 1$ and
    $B^i = \lambda\,Z^{-1} A^i Z$.  The family $Z_v := \lambda^{v} Z$
    dresses the gauge with powers of the root of unity: away from the seam
    the consecutive powers contribute the factor $\lambda$, and across the
    seam the wraparound contributes $\lambda^{1-n} = \lambda$ because
    $\lambda^n = 1$, so $B^i = Z_v^{-1} A^i Z_{v+1}$ at every site.
\end{proof}


The source states its Lemma~5 for site-dependent tensors: five sites
carrying tensors $A_1, \ldots, A_5$ whose two-site windows are injective.
The entries above specialize to one repeated tensor; the remaining entries
of this section formalize the site-dependent setting itself.  A closed
chain carries one tensor per site, windows are arcs of consecutive sites,
and the closed-chain corollary delivers one gauge per bond.  Throughout,
all sites share one physical dimension $d$ and all bonds one bond
dimension $D$; the source poses no restriction on the local dimensions of
the site-dependent tensors, and the uniform-dimension restriction is
recorded in the route note
(\path{docs/paper-gaps/peps_normal_ft_section3_route.tex}).

\begin{definition}[Arc products and window injectivity of a site-dependent
        chain]%
    \label{def:chain_arc_window_injective}
    \lean{MPSChainTensor.arcEval}
    \lean{MPSChainTensor.IsWindowInjective}
    \leanok
    \uses{def:non_ti_chain}
    Let a closed chain on $n \ge 1$ sites carry one tensor of $D \times D$
    matrices over physical dimension $d$ per site, $A_0, \dots, A_{n-1}$.
    The \emph{arc product} of a word $x = x_1 \cdots x_\ell$ read from
    site $s$ is
    \[
        A_{(s)}^{x} = A_{s}^{x_1}\, A_{s+1}^{x_2} \cdots
        A_{s+\ell-1}^{x_\ell},
    \]
    site indices cyclic.  The chain is \emph{window injective at length
        $L$} when, for every site $s$, the arc products of the length-$L$
    words read from $s$ span the full matrix algebra.  This is the
    site-dependent setting of the closing section of
    \cite{Molnar2018NormalPEPS} (``Normal MPS: alternative proof''): an
    MPS on five sites in which the blocking of any two consecutive tensors
    is injective, for a general window length $L$.
\end{definition}

\begin{lemma}[Arc spanning from window injectivity]%
    \label{lem:isWindowInjective_arc_span}
    \lean{MPSChainTensor.IsWindowInjective.arc_span}
    \leanok
    \uses{def:chain_arc_window_injective}
    Let a site-dependent chain on $n \ge 1$ sites be window injective at
    length $L$ with $0 < L$.  Then for every $m \ge L$ and every site $s$,
    the arc products of the length-$m$ words read from $s$ span the full
    matrix algebra.  This is the site-dependent form of the remark that
    any region of at least the blocking size is injective
    (\cite{Molnar2018NormalPEPS}, closing section,
    cf.\ Lemma~\ref{lem:isNBlkInjective_of_le}).
\end{lemma}
\begin{proof}\leanok
    \uses{def:chain_arc_window_injective}
    Write the identity as a combination of window products at the starting
    site and peel the leading letter of each: any matrix becomes a
    combination of terms $A_s^{i}\,(W P)$ with $W$ an arc product of
    length $L - 1$ read from $s + 1$.  Each factor $W P$ is a combination
    of length-$m$ arc products read from $s + 1$ by induction, so each
    term is a combination of length-$(m+1)$ arc products read from $s$.
\end{proof}

\begin{lemma}[Bond-operator extraction from overlapping windows,
        site-dependent form]%
    \label{lem:chain_overlapWindow_exists_bondOperator}
    \lean{MPSChainTensor.overlapWindow_exists_bondOperator}
    \lean{MPSChainTensor.eq_of_span_mul_left}
    \leanok
    \uses{def:chain_arc_window_injective}
    Let $0 < L$ with $2L + 1 \le n$, let the closed chain on $n$ sites
    carry one tensor of $D \times D$ matrices per site, window injective
    at length $L$, and fix a site $r$.  Let $C_0, \ldots, C_L$ be deformed
    window tensors around the bond $(r+L-1,\, r+L)$: each $C_j$ assigns a
    matrix to every word of length $L$, read as replacing the blocked
    tensor of the $L$-site window starting at site $r + j$.  Suppose the
    deformed states agree for consecutive window positions: for every
    $j < L$, every prefix $p$ of $j$ sites read from $r$, every assignment
    $u$ on the $L + 1$ sites covered by the two windows, and every suffix
    $s$ on the remaining sites,
    \[
        \tr\bigl(A_{(r)}^{p}\, C_j(u_1 \cdots u_L)\,
        A_{r+j+L}^{u_{L+1}}\, A_{(r+j+L+1)}^{s}\bigr)
        = \tr\bigl(A_{(r)}^{p}\, A_{r+j}^{u_1}\,
        C_{j+1}(u_2 \cdots u_{L+1})\, A_{(r+j+L+1)}^{s}\bigr).
    \]
    Then the deformation is a virtual operation on the bond: there is a
    matrix $X$ with
    \[
        C_0(a) = A_{(r)}^{a}\, X
        \quad\text{and}\quad
        C_L(b) = X\, A_{(r+L)}^{b}
    \]
    for all length-$L$ words $a$, $b$, and $X$ is unique.  This is
    Lemma~5 of \cite{Molnar2018NormalPEPS} in the site-dependent setting
    in which the source states it, with the deformed window tensors
    arising as physical operators contracted with the blocked windows
    they act on; Lemma~\ref{lem:overlapWindow_exists_bondOperator} is its
    constant specialization.
\end{lemma}
\begin{proof}\leanok
    \uses{lem:isWindowInjective_arc_span, def:chain_arc_window_injective}
    Comparing the windows at $j$ and $j + 1$ through the complementary
    arc of $n - L - 1 \ge L$ sites, whose arc products span by
    Lemma~\ref{lem:isWindowInjective_arc_span}, strips the state equality
    to the letter level:
    $C_j(u_1 \cdots u_L)\,A_{r+j+L}^{u_{L+1}} =
        A_{r+j}^{u_1}\,C_{j+1}(u_2 \cdots u_{L+1})$ on every window of
    $L + 1$ sites.  Chaining the $L$ relations along the $2L$ sites
    flanking the bond turns the leftmost window into the rightmost one:
    $C_0(a)\,A_{(r+L)}^{b} = A_{(r)}^{a}\,C_L(b)$ for all length-$L$
    words.  Writing the identity as a combination
    $\sum_b \alpha_b A_{(r+L)}^{b} = I$, the matrix
    $X := \sum_b \alpha_b\,C_L(b)$ satisfies $A_{(r)}^{a} X = C_0(a)$;
    the mirrored combination of the $C_0$-family collapses onto the same
    matrix, giving $X\,A_{(r+L)}^{b} = C_L(b)$, and stripping the
    spanning factors makes $X$ unique.
\end{proof}

\begin{lemma}[Arc transport between same-state chains]%
    \label{lem:exists_arcTransport}
    \lean{MPSChainTensor.exists_arcTransport}
    \leanok
    \uses{def:chain_arc_window_injective, def:non_ti_chain}
    Let two site-dependent chains $A$, $B$ on $n$ sites generate the same
    closed-chain state, let $k + q = n$, and suppose the $A$-arc products
    of length $k$ read from a site $s$ and of length $q$ read from
    $s + k$ each span the full matrix algebra.  Then there is a linear
    map $\Lambda$ of the matrix algebra with
    $\Lambda(A_{(s)}^{x}) = B_{(s)}^{x}$ for every word $x$ of length
    $k$.
\end{lemma}
\begin{proof}\leanok
    \uses{lem:exists_mpvTransport, def:chain_arc_window_injective}
    As for one repeated tensor (Lemma~\ref{lem:exists_mpvTransport}):
    pair the matrix algebra against the complementary arc products.  The
    pairing is injective on the $A$-side by spanning and trace
    nondegeneracy, and the two pairings match on the spanning arc family
    because both compute closed-chain coefficients of total length $n$,
    equal after rotating the trace of the closed chain to the starting
    site $s$.
\end{proof}

\begin{lemma}[The insertion correspondence is an algebra homomorphism,
        site-dependent form]%
    \label{lem:chain_exists_insertionHom}
    \lean{MPSChainTensor.exists_insertionHom}
    \lean{MPSChainTensor.insertionHom_unique}
    \leanok
    \uses{def:chain_arc_window_injective, def:non_ti_chain}
    Let $0 < L$ with $2L + 1 \le n$, let $A$ and $B$ be site-dependent
    chains on $n$ sites, each window injective at length $L$, generating
    the same closed-chain state, and fix a site $r$.  Then there is a
    linear map $\Phi$ of the matrix algebra, unital and multiplicative,
    such that for every matrix $X$ and every word $w$ of length $n$,
    \[
        \tr\bigl(X\, A_{(r+L)}^{w}\bigr)
        = \tr\bigl(\Phi(X)\, B_{(r+L)}^{w}\bigr),
    \]
    and $\Phi$ is the unique linear map with this pairing property.  This
    is the closing clause of Lemma~5 of \cite{Molnar2018NormalPEPS} — the
    maps $O_1 \mapsto X$ and $O_3^T \mapsto X$ are uniquely defined and
    are algebra homomorphisms — in the site-dependent setting of the
    source, with the algebra structure carried by the insertions: $\Phi$
    sends an insertion on the bond $(r+L-1,\, r+L)$ of the first chain to
    the bond operator extracted on the second.
\end{lemma}
\begin{proof}\leanok
    \uses{lem:chain_overlapWindow_exists_bondOperator,
        lem:exists_arcTransport, lem:isWindowInjective_arc_span}
    A virtual insertion $X$ on the bond is realized on each of the
    $L + 1$ overlapping windows around it by deformed window tensors
    placing $X$ after the first $L - j$ letters of the window starting at
    site $r + j$.  Each deformation is transported to the second chain by
    the arc transport at its own starting site
    (Lemma~\ref{lem:exists_arcTransport}); the same-state hypothesis
    makes the transported deformations generate one common state, and the
    bond-operator extraction
    (Lemma~\ref{lem:chain_overlapWindow_exists_bondOperator}) produces
    $\Phi(X)$.  Uniqueness of the bond operator makes $\Phi$ linear,
    unital and multiplicative; the pairing against all closed-chain words
    follows by expanding the insertion over the window products, and pins
    $\Phi$ through the spanning arcs of length $n$
    (Lemma~\ref{lem:isWindowInjective_arc_span}) and nondegeneracy of the
    trace pairing.
\end{proof}

\begin{lemma}[Conjugate arc products at every bond, site-dependent form]%
    \label{lem:chain_exists_conjugation_of_sameState}
    \lean{MPSChainTensor.exists_conjugation_of_sameState}
    \leanok
    \uses{def:chain_arc_window_injective, def:non_ti_chain}
    Let $0 < L$ with $2L + 1 \le n$, and let $A$ and $B$ be
    site-dependent chains on $n$ sites, each window injective at length
    $L$, generating the same closed-chain state.  Then for every site $p$
    there is an invertible matrix $Z$ with
    \[
        B_{(p)}^{w} = Z\, A_{(p)}^{w}\, Z^{-1}
    \]
    for every word $w$ of length $n$.  The statement for one repeated
    tensor is Lemma~\ref{lem:exists_conjugation_of_mpv_eq}.
\end{lemma}
\begin{proof}\leanok
    \uses{lem:chain_exists_insertionHom, thm:skolem_noether}
    The insertion correspondence at the bond
    (Lemma~\ref{lem:chain_exists_insertionHom}) is unital and nonzero,
    hence an automorphism of the full matrix algebra, and by
    Skolem--Noether (Theorem~\ref{thm:skolem_noether}) it is conjugation
    by an invertible $Z$; pairing insertions against all closed-chain
    words read from $p$ and using nondegeneracy of the trace pairing
    transfers the conjugation to the arc products.
\end{proof}

\begin{corollary}[Fundamental Theorem for site-dependent normal closed
        chains at $n \ge 2L + 1$]%
    \label{cor:fundamentalTheorem_normalMPSChain_of_overlap}
    \lean{TNLean.PEPS.fundamentalTheorem_normalMPSChain_of_overlap}
    \leanok
    \uses{def:chain_arc_window_injective, def:non_ti_chain}
    Let $0 < L$ with $2L + 1 \le n$, and let $A$ and $B$ be
    site-dependent chains on $n$ sites, each window injective at length
    $L$, generating the same closed-chain state.  Then the chains are
    gauge equivalent: there are invertible matrices $Z_v$, one per bond
    of the closed chain, with
    \[
        B_v^i = Z_v\, A_v^i\, Z_{v+1}^{-1}
    \]
    for every site $v$ and every $i$, indices cyclic.  This is the
    closed-chain corollary of the closing section of
    \cite{Molnar2018NormalPEPS} assembled from its site-dependent
    Lemma~5 at every bond; the source displays the corollary for
    translation-invariant tensors
    (Corollary~\ref{cor:fundamentalTheorem_normalMPS_translationInvariant_of_overlap}),
    with the site-dependent Lemma~5 as its engine.
\end{corollary}
\begin{proof}\leanok
    \uses{lem:chain_exists_conjugation_of_sameState,
        lem:chain_overlapWindow_exists_bondOperator,
        lem:isWindowInjective_arc_span}
    For a vanishing bond dimension the conclusion is trivial.  Otherwise
    the per-bond conjugations
    (Lemma~\ref{lem:chain_exists_conjugation_of_sameState}) give window
    covariance with nonzero scalars: on every arc of $m$ sites with
    $L \le m$ and $m + L \le n$, the conjugation at the near end
    intertwines the arc products of the two chains across the far bond,
    the bond-operator extraction
    (Lemma~\ref{lem:chain_overlapWindow_exists_bondOperator}) produces
    the connecting matrix, and the conjugation at the far end pins it to
    a nonzero scalar multiple of the gauge there through the centralizer
    of the full matrix algebra, so
    $B_{(p)}^{u} = c_{p,m} \cdot Z_p\, A_{(p)}^{u}\, Z_{p+m}^{-1}$ on all
    length-$m$ words $u$.  Comparing the covariance on windows of lengths
    $L + 1$ and $L$ through the spanning window products leaves the
    letterwise relation $B_v^i = \mu_v\, Z_v\, A_v^i\, Z_{v+1}^{-1}$ with
    nonzero scalars $\mu_v$; iterating it once around the closed chain
    against a nonzero arc product forces $\prod_v \mu_v = 1$, and
    dressing the gauges with the partial products of the scalars absorbs
    them, including across the seam.
\end{proof}


\begin{theorem}[Fundamental Theorem for normal PEPS, vertex-injective case]%
    \label{thm:fundamentalTheorem_normalPEPS_of_vertexInjective}
    \lean{TNLean.PEPS.fundamentalTheorem_normalPEPS_of_vertexInjective}
    \leanok
    \uses{thm:fundamentalTheorem_PEPS, def:peps_normalPEPSBlockingHypotheses,
        def:GaugeEquivPEPS, def:IsVertexInjective}
    Suppose PEPS tensors $A$ and $B$ are vertex injective with positive bond
    dimensions on a connected graph, generate the same state, and $A$ satisfies
    the normal blocking hypotheses. Then their defining tensors are related by
    a local gauge.
    This is the vertex-injective specialization of the general normal-PEPS
    theorem: when each single tensor is already injective the gauge follows
    from the injective Fundamental Theorem, so the region blocking hypotheses
    are not used. The general region-injective theorem, in which only blocked
    regions are injective, remains open; the missing step is the
    region-level insertion-algebra isomorphism that produces the per-edge gauge
    after blocking.
\end{theorem}
\begin{proof}\leanok
    Vertex injectivity, positive bond dimensions, same state, and connectivity
    are exactly the hypotheses of the injective Fundamental Theorem, which
    supplies the local gauge.
\end{proof}

\begin{definition}[Oriented endpoint scalar]%
    \label{def:edgeScalarAt}
    \lean{TNLean.PEPS.edgeScalarAt}
    \leanok
    \uses{def:edgeGaugeAt}
    An edge-scalar family $c=(c_e)_{e\in E}$ determines the endpoint scalar
    $c_{v,e}$ by $c_{v,e}=c_e$ at the lower endpoint of $e$ and
    $c_{v,e}=c_e^{-1}$ at the upper endpoint.
\end{definition}

\begin{definition}[Vertex-balanced edge scalars]%
    \label{def:IsVertexBalanced}
    \lean{TNLean.PEPS.IsVertexBalanced}
    \leanok
    \uses{def:edgeScalarAt}
    An edge-scalar family is vertex-balanced when
    $\prod_{e\ni v} c_{v,e}=1$ for every vertex $v$.
\end{definition}
