\section{Square-Lattice Normal PEPS}\label{sec:peps_normal_square}

On the square lattice the injective regions are built from $2\times3$ and
$3\times2$ coordinate rectangles. This section fixes the geometry: the nested
regions $R\subseteq S$ with $S\setminus R$ a single site, which realize the
one-site-different injective regions the normal theorem requires at each vertex,
together with the local window $T$ used around an edge. The union closure of
Section~\ref{sec:peps_normal_union} builds these regions from injective
rectangles, so a rectangular injectivity hypothesis yields their injectivity.

\begin{definition}[Complement of a finite region]%
    \label{def:peps_regionComplement}
    \lean{TNLean.PEPS.regionComplement}
    \lean{TNLean.PEPS.mem_regionComplement}
    \leanok
    For a finite vertex region $R\subseteq V$, its complement is
    $V\setminus R$. Equivalently, a vertex lies in the complement exactly
    when it does not lie in $R$.
\end{definition}

\begin{definition}[Square-lattice coordinate regions]%
    \label{def:peps_squareLatticeCoordinateRegions}
    \lean{TNLean.PEPS.SquareLatticeVertex}
    \lean{TNLean.PEPS.squareLatticeHorizontalNeighbor}
    \lean{TNLean.PEPS.squareLatticeVerticalNeighbor}
    \lean{TNLean.PEPS.squareLatticeGraph}
    \lean{TNLean.PEPS.IsHorizontalSquareLatticeEdge}
    \lean{TNLean.PEPS.IsVerticalSquareLatticeEdge}
    \lean{TNLean.PEPS.squareLatticeRectangle}
    \lean{TNLean.PEPS.squareLatticeFull}
    \lean{TNLean.PEPS.squareLatticeCoordinateInterval}
    \lean{TNLean.PEPS.squareLatticeContiguousRectangle}
    \lean{TNLean.PEPS.IsTwoByThreeSquareLatticeProduct}
    \lean{TNLean.PEPS.IsThreeByTwoSquareLatticeProduct}
    \lean{TNLean.PEPS.IsTwoByThreeContiguousSquareLatticeRectangle}
    \lean{TNLean.PEPS.IsThreeByTwoContiguousSquareLatticeRectangle}
    \lean{TNLean.PEPS.normalSquareRegionR}
    \lean{TNLean.PEPS.normalSquareRegionS}
    \lean{TNLean.PEPS.normalSquareRegionTVerticalBlock}
    \lean{TNLean.PEPS.normalSquareRegionTHorizontalBlock}
    \lean{TNLean.PEPS.normalSquareRegionTHole}
    \lean{TNLean.PEPS.normalSquareRegionT}
    \lean{TNLean.PEPS.normalSquareEdgeComplementRegion}
    \lean{TNLean.PEPS.squareLatticeGraph_adj}
    \lean{TNLean.PEPS.squareLatticeGraph_adj_right}
    \lean{TNLean.PEPS.squareLatticeGraph_adj_up}
    \lean{TNLean.PEPS.squareLatticeRightEdge}
    \lean{TNLean.PEPS.squareLatticeUpEdge}
    \lean{TNLean.PEPS.squareLatticeRightEdge_isHorizontal}
    \lean{TNLean.PEPS.squareLatticeUpEdge_isVertical}
    \lean{TNLean.PEPS.squareLatticeEdge_horizontal_or_vertical}
    \lean{TNLean.PEPS.squareLatticeEdge_not_horizontal_and_vertical}
    \lean{TNLean.PEPS.horizontalSquareLatticeEdge_coords}
    \lean{TNLean.PEPS.horizontalSquareLatticeEdge_eq_rightEdge}
    \lean{TNLean.PEPS.verticalSquareLatticeEdge_coords}
    \lean{TNLean.PEPS.verticalSquareLatticeEdge_eq_upEdge}
    \leanok
    The finite $n\times m$ square lattice is represented by pairs of
    coordinates. The corresponding square-lattice graph has horizontal and
    vertical nearest-neighbor edges; the two elementary adjacency lemmas record
    the right-neighbor and upper-neighbor cases used for translated edge
    blockings, and the corresponding coordinate edges name those two
    nearest-neighbor edges as elements of the edge set. These named right and
    upward edges have the expected horizontal and vertical orientations.
    Every edge of this
    graph is horizontal or vertical, but not both. Since edges are represented
    by ordered endpoint pairs, a horizontal edge is oriented from left to right
    and is exactly the coordinate right edge from its left endpoint; similarly,
    a vertical edge is oriented from bottom to top and is exactly the
    coordinate upward edge from its lower endpoint.
    A coordinate rectangle is a product of a horizontal coordinate set and a
    vertical coordinate set. The $2\times3$ and
    $3\times2$ coordinate-product predicates assert only that these two
    coordinate sets have cardinalities $2,3$ and $3,2$, respectively.
    Contiguous coordinate rectangles are products of contiguous coordinate
    intervals, and the contiguous $2\times3$ and $3\times2$ predicates name the
    rectangular blocks used in the normal square-lattice theorem. The
    displayed region $R$ is the union of a $2\times3$ contiguous rectangle
    with the $3\times2$ contiguous rectangle occupying its upper two rows and
    extending one column to the right. The displayed region $S$ is the union
    of two $2\times3$ contiguous rectangles shifted by one horizontal
    coordinate, hence the full $3\times3$ square spanned by them. The
    displayed local-window model for $T$ is the complement, inside a local
    $5\times6$ coordinate window, of the vertical $2\times3$ edge block
    and the horizontal $3\times2$ edge block shown in the source picture.
    The finite-lattice complementary block is the whole finite square lattice
    with those two edge blocks removed; this is the block denoted $A_3$ in
    the proof of \cite[Theorem~3]{Molnar2018NormalPEPS}.
    The source shapes are:
    \begin{center}
        \TNPEPSNormalRegionsRS

        \vspace{0.5em}

        \TNPEPSNormalRegionT
    \end{center}
    % Maintainer note: injectivity of this finite-lattice complementary block
    % is not yet proved; this definition only names the region.
\end{definition}

\begin{lemma}[Basic identities for square-lattice coordinate regions]%
    \label{lem:peps_squareLatticeCoordinateRegion_identities}
    \lean{TNLean.PEPS.mem_squareLatticeRectangle}
    \lean{TNLean.PEPS.mem_squareLatticeCoordinateInterval}
    \lean{TNLean.PEPS.mem_squareLatticeContiguousRectangle}
    \lean{TNLean.PEPS.mem_normalSquareRegionR}
    \lean{TNLean.PEPS.mem_normalSquareRegionS}
    \lean{TNLean.PEPS.mem_normalSquareRegionTVerticalBlock}
    \lean{TNLean.PEPS.mem_normalSquareRegionTHorizontalBlock}
    \lean{TNLean.PEPS.mem_normalSquareRegionTHole}
    \lean{TNLean.PEPS.mem_normalSquareRegionT}
    \lean{TNLean.PEPS.mem_normalSquareEdgeComplementRegion}
    \lean{TNLean.PEPS.isTwoByThreeContiguousSquareLatticeRectangle_of_bounds}
    \lean{TNLean.PEPS.isThreeByTwoContiguousSquareLatticeRectangle_of_bounds}
    \lean{TNLean.PEPS.card_squareLatticeRectangle}
    \lean{TNLean.PEPS.squareLatticeFull_eq_univ}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions}
    Membership in a coordinate rectangle is coordinatewise membership, the
    contiguous coordinate interval has the expected coordinate inequalities,
    membership in a contiguous coordinate rectangle is the conjunction of the
    four bounding inequalities, and membership in $R$ is the disjunction of
    membership in the $2\times3$ lower-left piece or in the shifted
    $3\times2$ upper piece. Membership in $S$ is the disjunction of
    membership in the left $2\times3$ piece or in the horizontally shifted
    $2\times3$ piece. Membership in the local-window $T$ region is
    membership in the local $5\times6$ window together with nonmembership in
    the union of the two displayed edge blocks. Bounded contiguous coordinate
    rectangles of sizes $2\times3$ and $3\times2$ satisfy, respectively,
    the predicates for
    contiguous $2\times3$ and contiguous $3\times2$ square-lattice
    rectangles. The cardinality of a coordinate rectangle is the
    product of the two coordinate cardinalities, and the full coordinate
    rectangle is the whole finite vertex set.
\end{lemma}
\begin{proof}\leanok
    The coordinate-rectangle statements follow from membership and cardinality
    identities for Cartesian products of finite coordinate sets. The
    contiguous-rectangle statement then unfolds the two coordinate intervals.
    The membership statements for $R$, $S$, and the two edge blocks removed
    from $T$ additionally unfold membership in a union of two such
    rectangles, while membership in $T$ also unfolds the set difference from
    the local $5\times6$ window. The shape predicates follow by using the
    displayed lower-left corners as witnesses.
\end{proof}

\begin{lemma}[The normal regions $R$ and $S$ are unions of contiguous
        rectangles]%
    \label{lem:peps_normalSquareRegionRS_rectangular_decomposition}
    \lean{TNLean.PEPS.normalSquareRegionR_rectangular_decomposition}
    \lean{TNLean.PEPS.normalSquareRegionS_rectangular_decomposition}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions,
        lem:peps_squareLatticeCoordinateRegion_identities}
    If the displayed $3\times3$ window containing $R$ and $S$ lies in
    the finite square lattice, then $R$ is the union of a contiguous
    $2\times3$ rectangle and a contiguous $3\times2$ rectangle, while
    $S$ is the union of two contiguous $2\times3$ rectangles.
\end{lemma}
\begin{proof}\leanok
    This is the defining description of $R$ and $S$, together with the
    coordinate bounds which make the displayed rectangles contiguous
    $2\times3$ and $3\times2$ rectangles.
\end{proof}

\begin{lemma}[The edge blocks removed from $T$ are contiguous rectangles]%
    \label{lem:peps_normalSquareRegionTHole_rectangular_decomposition}
    \lean{TNLean.PEPS.normalSquareRegionTVerticalBlock_rectangular}
    \lean{TNLean.PEPS.normalSquareRegionTHorizontalBlock_rectangular}
    \lean{TNLean.PEPS.normalSquareRegionTHole_rectangular_decomposition}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions,
        lem:peps_squareLatticeCoordinateRegion_identities}
    If the two edge blocks cut out of the displayed $T$-region lie in the
    finite square lattice, then they are respectively a contiguous
    $2\times3$ rectangle and a contiguous $3\times2$ rectangle.
\end{lemma}
\begin{proof}\leanok
    This is the defining description of the two removed edge blocks, together
    with the coordinate bounds which make them valid contiguous rectangles.
\end{proof}

\begin{lemma}[The local $T$-region and its removed edge blocks fill the window]%
    \label{lem:peps_normalSquareRegionT_window_complement}
    \lean{TNLean.PEPS.normalSquareRegionTHole_subset_window}
    \lean{TNLean.PEPS.normalSquareRegionTVerticalBlock_disjoint_horizontalBlock}
    \lean{TNLean.PEPS.normalSquareRegionT_disjoint_THole}
    \lean{TNLean.PEPS.normalSquareRegionT_disjoint_verticalBlock}
    \lean{TNLean.PEPS.normalSquareRegionT_disjoint_horizontalBlock}
    \lean{TNLean.PEPS.normalSquareRegionT_union_THole}
    \lean{TNLean.PEPS.normalSquareRegionT_union_verticalBlock_union_horizontalBlock}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions,
        lem:peps_squareLatticeCoordinateRegion_identities}
    The two edge blocks removed from the displayed local $T$-region lie inside
    the local $5\times6$ coordinate window. They are disjoint from each other
    and from this local $T$ region, and their union with it is exactly that
    local window.
\end{lemma}
\begin{proof}\leanok
    This follows by unfolding the definition of $T$ as the set-theoretic
    complement of the two displayed edge blocks inside the local window.
\end{proof}

\begin{lemma}[The edge-complementary block fills the finite lattice]%
    \label{lem:peps_normalSquareEdgeComplementRegion}
    \lean{TNLean.PEPS.normalSquareRegionT_subset_edgeComplementRegion}
    \lean{TNLean.PEPS.normalSquareEdgeComplementRegion_disjoint_THole}
    \lean{TNLean.PEPS.normalSquareEdgeComplementRegion_union_THole}
    \lean{TNLean.PEPS.normalSquareEdgeComplementRegion_union_verticalBlock_union_horizontalBlock}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions,
        lem:peps_squareLatticeCoordinateRegion_identities}
    The finite-lattice complementary block around the edge is disjoint from
    the two displayed edge blocks, and its union with them is the full finite
    square lattice. The local-window $T$ region is contained in this
    finite-lattice complementary block.
\end{lemma}
\begin{proof}\leanok
    This is the set-theoretic complement of the two displayed edge blocks in
    the full finite square lattice. The containment of the local-window
    $T$ region follows because its points also avoid the two removed edge
    blocks.
\end{proof}

\begin{definition}[Rectangular covers of square-lattice regions]%
    \label{def:peps_normalSquareRectangleCover}
    \lean{TNLean.PEPS.SquareLatticeRectangleCover}
    \lean{TNLean.PEPS.NormalSquareRegionTRectangleCover}
    \lean{TNLean.PEPS.NormalSquareEdgeComplementRectangleCover}
    \lean{TNLean.PEPS.normalSquareVerticalEdgeComplementRegion}
    \lean{TNLean.PEPS.NormalSquareVerticalEdgeComplementRectangleCover}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions,
        lem:peps_normalSquareEdgeComplementRegion}
    A rectangular cover of a square-lattice region is a nonempty finite family
    of regions whose union is exactly the given region, and each member is a
    contiguous $2\times3$ or $3\times2$ rectangle. The two cases used in the
    proof are the cover of the local displayed $T$-region and the cover of the
    finite-lattice complementary block $A_3$.
    \begin{center}
        \TNPEPSNormalRectangleCover
    \end{center}
\end{definition}
% Maintainer note: this definition records sufficient coordinate criteria.
% The concrete local T cover remains to be constructed.

\begin{lemma}[The present local $T$-window and $A_3$ cover obstructions]%
    \label{lem:peps_normalSquareRegionT_no_five_by_six_cover}
    \lean{TNLean.PEPS.not_squareLatticeRectangleCover_of_forced_points}
    \lean{TNLean.PEPS.not_normalSquareRegionT_rectangleCover_at_origin}
    \lean{TNLean.PEPS.not_normalSquareRegionT_rectangleCover_five_by_six}
    \lean{TNLean.PEPS.not_normalSquareVerticalRegionT_rectangleCover_at_origin}
    \lean{TNLean.PEPS.not_normalSquareVerticalRegionT_rectangleCover_six_by_five}
    \lean{TNLean.PEPS.not_normalSquareEdgeComplementRectangleCover_at_origin}
    \lean{TNLean.PEPS.not_normalSquareEdgeComplementRectangleCover_five_by_seven}
    \lean{TNLean.PEPS.not_normalSquareVerticalEdgeComplementRectangleCover_at_origin}
    \lean{TNLean.PEPS.not_normalSquareVerticalEdgeComplementRectangleCover_seven_by_five}
    \leanok
    \uses{def:peps_normalSquareRectangleCover,
        lem:peps_normalSquareRegionT_window_complement}
    The present origin-based local-window model for the displayed $T$-region
    admits no nonempty finite cover by contained contiguous $2\times3$ and
    $3\times2$ rectangles whenever the ambient lattice contains the
    $5\times6$ window. The origin-based rotated local-window model used for
    vertical edges has the analogous obstruction whenever the ambient lattice
    contains the $6\times5$ window. The normalized $5\times6$ and
    $6\times5$ cases are recorded as corollaries. The present origin-based
    horizontal edge-complement model has the analogous obstruction whenever
    the ambient lattice contains the $5\times7$ frame; the normalized
    $5\times7$ case is again recorded as a corollary. The present
    origin-based vertical edge-complement model has the corresponding
    obstruction whenever the ambient lattice contains the $7\times5$ frame;
    the normalized $7\times5$ case is recorded as a corollary.
\end{lemma}
% Maintainer note: this is a local-model obstruction, not a statement of the
% source theorem. It records that the current coordinate model is not yet the
% source cover needed for the $T$-injectivity sentence.
\begin{proof}\leanok
    The lower-left point of the local window belongs to the present $T$
    model. Any contiguous $2\times3$ rectangle containing it also contains
    a point of the removed vertical edge block, while any contiguous
    $3\times2$ rectangle containing it also contains a point of the removed
    horizontal edge block. The rotated statement is the same coordinate
    argument with the two shapes interchanged. The horizontal
    edge-complement statement uses the same lower-left point, and so does the
    normalized vertical edge-complement statement. These cases are all
    instances of the same point-forcing obstruction to an exact rectangular
    cover.
\end{proof}

\begin{lemma}[Injectivity from a rectangular cover]%
    \label{lem:peps_normalSquareRectangleCover_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.injective_of_rectangleCover}
    \leanok
    \uses{def:peps_normalSquareRectangleCover,
        def:peps_normalRectangleInjectivityHypotheses,
        lem:peps_regionInjectivityUnionClosure_finite}
    If a square-lattice region has a nonempty finite cover by contiguous
    $2\times3$ and $3\times2$ rectangles, then rectangular injectivity and
    finite union closure imply that the region is injective.
\end{lemma}
\begin{proof}\leanok
    Apply rectangular injectivity to every rectangle in the cover, then apply
    finite union closure to the nonempty family.
\end{proof}

\begin{lemma}[The normal regions $R$ and $S$ differ in one site]%
    \label{lem:peps_normalSquareRegionR_regionS_difference}
    \lean{TNLean.PEPS.normalSquareRegionR_subset_regionS}
    \lean{TNLean.PEPS.mem_normalSquareRegionS_sdiff_regionR}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions,
        lem:peps_squareLatticeCoordinateRegion_identities}
    For the displayed regions $R$ and $S$ with the same lower-left corner,
    $R$ is contained in $S$. The difference $S\setminus R$ is exactly
    the lower-right site of the $3\times3$ square: in coordinates, the site
    one obtains by shifting two steps horizontally and zero steps vertically
    from the lower-left corner.
\end{lemma}
\begin{proof}\leanok
    Expand the two displayed unions. The only point of the second
    $2\times3$ piece of $S$ which is not already in the union defining
    $R$ is its lower-right point.
\end{proof}

\begin{definition}[Normal rectangular injectivity hypotheses]%
    \label{def:peps_normalRectangleInjectivityHypotheses}
    \lean{TNLean.PEPS.NormalRectangleInjectivityHypotheses}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses}
    \leanok
    \uses{def:peps_regionInjectivityData,
        def:peps_squareLatticeCoordinateRegions}
    In the square-lattice normal theorem, every $2\times 3$ and
    $3\times 2$ rectangular region is assumed injective. The corresponding
    abstract hypotheses specify which finite regions have these two shapes and
    assert injectivity for all of them. In the coordinate square-lattice
    specialization, these two families are the contiguous coordinate
    $2\times 3$ and $3\times 2$ rectangles.
\end{definition}

\begin{lemma}[Coordinate rectangular injectivity gives abstract rectangular
        injectivity]%
    \label{lem:peps_squareLatticeRectangleInjectivity_toRectangular}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.toRectangular}
    \leanok
    \uses{def:peps_normalRectangleInjectivityHypotheses}
    The coordinate square-lattice rectangular injectivity hypotheses determine
    abstract rectangular injectivity hypotheses, with the two abstract
    rectangular families taken to be the contiguous coordinate $2\times 3$
    and $3\times 2$ rectangles.
\end{lemma}
\begin{proof}\leanok
    This is the direct specialization of the abstract rectangular families to
    the two contiguous coordinate-rectangle predicates.
\end{proof}

\begin{lemma}[Coordinate rectangular injectivity gives injective rectangles]%
    \label{lem:peps_squareLatticeRectangleInjectivity_rectangles}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.rect23_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.rect32_injective}
    \leanok
    \uses{def:peps_normalRectangleInjectivityHypotheses,
        lem:peps_squareLatticeCoordinateRegion_identities}
    Under the coordinate square-lattice rectangular injectivity hypotheses,
    every bounded contiguous $2\times3$ rectangle and every bounded
    contiguous $3\times2$ rectangle is injective.
\end{lemma}
\begin{proof}\leanok
    This is the defining rectangular injectivity hypothesis applied to the
    coordinate witnesses for the two rectangle shapes.
\end{proof}

\begin{lemma}[Rectangular injectivity gives injective $T$-edge blocks]%
    \label{lem:peps_normalSquareRegionT_edgeBlocks_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.tVerticalBlock_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.tHorizontalBlock_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.tHole_injective_of_union}
    \leanok
    \uses{def:peps_regionInjectivityData,
        def:peps_normalRectangleInjectivityHypotheses,
        lem:peps_normalSquareRegionTHole_rectangular_decomposition}
    For the displayed local $T$-region inside its $5\times6$ coordinate window,
    if the smaller edge-block bounds
    $x_0 + 5 \leq n$ and $y_0 + 5 \leq m$ hold in the finite $n\times m$
    square lattice, then the coordinate square-lattice rectangular injectivity
    hypotheses imply that the vertical $2\times3$ block and horizontal
    $3\times2$ block removed from the window are injective. If, in addition,
    region injectivity is closed under unions, then the union of these two
    removed edge blocks is injective.
\end{lemma}
% Maintainer note: this does not prove injectivity of the finite-lattice
% complementary block used as the third region in
% \cite[Theorem~3]{Molnar2018NormalPEPS}.
\begin{proof}\leanok
    Apply the rectangular injectivity hypotheses to the already identified
    contiguous rectangle shapes of the two edge blocks. The final assertion is
    one application of union closure to these two injective blocks.
\end{proof}

\begin{lemma}[Injectivity from a rectangular cover of the local $T$-region]%
    \label{lem:peps_normalSquareRegionT_injective_of_cover}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.regionT_injective_of_cover}
    \lean{TNLean.PEPS.NormalSquareVerticalRegionTRectangleCover}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.verticalT_inj_of_cover}
    \leanok
    \uses{def:peps_normalSquareRectangleCover,
        def:peps_normalRectangleInjectivityHypotheses,
        lem:peps_normalSquareRectangleCover_injective,
        lem:peps_regionInjectivityUnionClosure_finite}
    If the displayed local $T$-region has a nonempty finite cover by
    contiguous $2\times3$ and $3\times2$ rectangles, then rectangular
    injectivity and finite union closure imply that $T$ is injective. The
    rotated vertical local $T$-region has the same conditional injectivity
    criterion.
\end{lemma}
% Current source-comparison status: docs/paper-gaps/peps_normal_ft_section3_route.tex.
\begin{proof}\leanok
    Apply rectangular injectivity to every rectangle in the cover, then apply
    finite union closure to the nonempty family.
\end{proof}

\begin{lemma}[Injectivity from a rectangular cover of the edge complement]%
    \label{lem:peps_normalSquareEdgeComplementRegion_injective_of_cover}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.edgeComplement_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.%
        verticalEdgeComplement_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.%
        edgeComplement_injective_of_T_cover}
    \leanok
    \uses{def:peps_normalSquareRectangleCover,
        def:peps_normalRectangleInjectivityHypotheses,
        lem:peps_normalSquareRectangleCover_injective,
        lem:peps_regionInjectivityUnionClosure_finite,
        lem:peps_normalSquareEdgeComplementRegion_coverOfT}
    If the finite-lattice complementary block $A_3$ has a nonempty finite
    cover by contiguous $2\times3$ and $3\times2$ rectangles, then
    rectangular injectivity and finite union closure imply that $A_3$ is
    injective. The same criterion applies to every vertical edge-complementary
    block: a nonempty finite cover of that block by contiguous $2\times3$ and
    $3\times2$ rectangles implies injectivity of the vertical
    edge-complementary block. In the normalized horizontal-edge $5\times7$
    frame, it is enough to give such a cover of the local $T$-region, since
    adjoining the two top-collar $3\times2$ rectangles gives a cover of $A_3$.
\end{lemma}
% Current source-comparison status: docs/paper-gaps/peps_normal_ft_section3_route.tex.
\begin{proof}\leanok
    Apply rectangular injectivity to every rectangle in the cover, then apply
    finite union closure to the nonempty family. For the normalized
    $5\times7$ edge complement, first extend the local $T$-cover by the
    two top-collar rectangles.
\end{proof}

\begin{lemma}[A cover of $T$ extends to a cover of the edge complement]%
    \label{lem:peps_normalSquareEdgeComplementRegion_coverOfT}
    \lean{TNLean.PEPS.normalSquareEdgeComplementRectangleCoverOfT}
    \lean{TNLean.PEPS.normalSquareVerticalEdgeComplementRectangleCoverOfVerticalT}
    \leanok
    \uses{def:peps_normalSquareRectangleCover,
        lem:peps_normalSquareEdgeComplementRegion_topCollar}
    In the normalized horizontal-edge $5\times7$ frame with lower-left corner
    $(0,0)$, any rectangular cover of the local $T$-region extends to a
    rectangular cover of the edge-complementary block $A_3$ by adjoining the
    two top-collar contiguous $3\times2$ rectangles. In the normalized
    vertical-edge $7\times5$ frame with lower-left corner $(0,0)$, any
    rectangular cover of the rotated local $T$-region extends to a rectangular
    cover of the vertical edge-complementary block by adjoining the two
    right-collar contiguous $2\times3$ rectangles.
\end{lemma}
\begin{proof}\leanok
    Index the new cover by the old cover indices together with two extra
    indices for the collar rectangles. For the horizontal block, the
    decomposition is
    \[
        A_3 =
        T \cup (\{0,1,2\}\times\{5,6\})
        \cup (\{2,3,4\}\times\{5,6\}).
    \]
    For the vertical block, the decomposition is
    \[
        A_3^{\mathrm v} =
        T^{\mathrm v} \cup (\{5,6\}\times\{0,1,2\})
        \cup (\{5,6\}\times\{2,3,4\}).
    \]
\end{proof}

\begin{lemma}[The normalized edge complements have collars]%
    \label{lem:peps_normalSquareEdgeComplementRegion_topCollar}
    \lean{TNLean.PEPS.normalSquareEdgeComplementRegion_eq_T_union_topCollar}
    \lean{TNLean.PEPS.normalSquareEdgeComplementRegion_eq_T_union_rightCollar}
    \lean{TNLean.PEPS.normalSquareVerticalRegionTHole}
    \lean{TNLean.PEPS.normalSquareVerticalRegionT}
    \lean{TNLean.PEPS.mem_normalSquareVerticalRegionTHole}
    \lean{TNLean.PEPS.mem_normalSquareVerticalRegionT}
    \lean{TNLean.PEPS.normalSquareVerticalEdgeComplement_eq_verticalT_union_rightCollar}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.topCollar_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.rightCollar_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.verticalComp_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.verticalComp_inj_of_cover}
    \leanok
    \uses{lem:peps_normalSquareEdgeComplementRegion,
        lem:peps_normalSquareRegionT_window_complement,
        def:peps_regionInjectivityData,
        lem:peps_squareLatticeRectangleInjectivity_rectangles}
    In the normalized horizontal-edge $5\times7$ frame, the finite-lattice
    complementary block $A_3$ is the local-window $T$-region together with
    two top-collar contiguous $3\times2$ rectangles. Consequently, if the
    local $T$-region is injective, then rectangular injectivity and union
    closure imply that $A_3$ is injective. There is also a transposed
    $7\times5$ horizontal-frame collar identity with two contiguous
    $2\times3$ right-collar rectangles. For the rotated vertical
    edge-blocking regions, the corresponding complement is the rotated local
    $T$-region together with the same two right-collar rectangles; hence rotated
    $T$-injectivity, and hence a rectangular cover of rotated $T$, implies
    injectivity of the vertical complementary block.
    \begin{center}
        \TNPEPSNormalEdgeComplementTopCollar
    \end{center}
\end{lemma}
% Current source-comparison status: docs/paper-gaps/peps_normal_ft_section3_route.tex.
\begin{proof}\leanok
    The two set identities follow by expanding the coordinate inequalities.
    The injectivity statements apply union closure first to the local
    $T$-region and the first collar rectangle, and then to the second collar
    rectangle.
\end{proof}

\begin{lemma}[A large rectangle is injective]%
    \label{lem:peps_normalSquareWideRectangle_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.wideRectangle_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.shortRectangle_injective}
    \leanok
    \uses{lem:peps_injectiveRegion_union,
        lem:peps_squareLatticeRectangleInjectivity_rectangles}
    A contiguous rectangle whose width is at least two and whose height is at
    least three is the union of the source $2\times3$ rectangles sliding over
    it, so rectangular injectivity and union closure prove it injective. The
    transposed statement holds for a contiguous rectangle of width at least
    three and height at least two.
\end{lemma}
% Current source-comparison status: docs/paper-gaps/peps_normal_ft_section3_route.tex.
\begin{proof}\leanok
    Each cell of the rectangle lies in one of the contained source windows, so
    the rectangle equals the finite union of those windows; union closure then
    gives injectivity.
\end{proof}

\begin{lemma}[Interior edge-complement injectivity]%
    \label{lem:peps_normalSquareInteriorEdgeComplement_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.interiorEdgeComplement_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.interiorVerticalEdgeComplement_injective}
    \leanok
    \uses{lem:peps_normalSquareEdgeComplementRegion,
        lem:peps_normalSquareWideRectangle_injective,
        lem:peps_squareLatticeRectangleInjectivity_rectangles}
    Let the two edge blocks be removed at an interior position of a sufficiently
    large lattice, with one row and column of room below and to the left and
    further room above and to the right. Then the complementary block $A_3$ is
    the union of four full-length bands around the removed blocks --- below,
    above, left, and right --- together with two filler rectangles inside the
    bounding box of the removed blocks. Each band is a large contiguous
    rectangle and each filler is a source rectangle, all avoiding the removed
    blocks, so rectangular injectivity and union closure prove $A_3$ injective
    with no rectangular-cover hypothesis. The transposed statement holds for the
    rotated vertical complementary block.

    \emph{Scope restriction (interior edge).} The removed blocks must sit in the
    interior. The fixed origin window has its corner forced and so has no
    rectangular cover; the source object is the full-lattice complement $A_3$,
    which has room only at an interior edge. The source comparison is recorded
    in \path{docs/paper-gaps/peps_normal_ft_section3_route.tex}.
\end{lemma}
% Current source-comparison status: docs/paper-gaps/peps_normal_ft_section3_route.tex.
\begin{proof}\leanok
    The six pieces cover the complement of the removed blocks, by expanding the
    coordinate inequalities. Each band is injective by the large-rectangle
    lemma and each filler by rectangular injectivity, so union closure gives
    injectivity of the union.
\end{proof}

\begin{lemma}[Cover-free interior every-edge blocking]%
    \label{lem:peps_normalSquareInteriorEdgeBlocking}
    \lean{TNLean.PEPS.horizontalSquareLatticeEdge_blockingDatum_interior}
    \lean{TNLean.PEPS.verticalSquareLatticeEdge_blockingDatum_interior}
    \lean{TNLean.PEPS.NormalSquareInteriorEdgeDatum}
    \lean{TNLean.PEPS.normalSquareEdgeBlockingHypotheses_of_interiorData}
    \leanok
    \uses{lem:peps_normalSquareInteriorEdgeComplement_injective,
        lem:peps_normalSquareHorizontalTranslatedEdge_blockingData,
        def:peps_normalEdgeBlockingHypotheses}
    For an interior horizontal or vertical edge with the corresponding interior
    margins, the translated edge-blocking datum is built with no rectangular
    cover: its complementary-block injectivity is the interior $A_3$
    injectivity. A family of these cover-free interior data over all edges
    assembles the normal edge-blocking hypotheses, with the injective chain,
    endpoint membership, pairwise disjointness, and coverage at every edge.

    \emph{Scope restriction (interior edge).} The construction supplies data
    only at edges with interior margins. Edges near the lattice boundary of the
    open rectangular model lack such margins; on a translation-invariant lattice
    every edge is interior. The source comparison is recorded in
    \path{docs/paper-gaps/peps_normal_ft_section3_route.tex}.
\end{lemma}
% Current source-comparison status: docs/paper-gaps/peps_normal_ft_section3_route.tex.
\begin{proof}\leanok
    Each interior edge is horizontal or vertical; the corresponding interior
    blocking datum discharges its complementary-block injectivity through the
    interior $A_3$ injectivity. Assembling these per-edge data over all edges
    gives the edge-blocking hypotheses.
\end{proof}

\begin{lemma}[Normalized edge-blocking data]%
    \label{lem:peps_normalSquareHorizontalEdge_blockingData}
    \lean{TNLean.PEPS.normalSquareHorizontalEdge}
    \lean{TNLean.PEPS.normalSquareHorizontalEdge_isHorizontal}
    \lean{TNLean.PEPS.normalSquareHorizontalEdgeRed}
    \lean{TNLean.PEPS.normalSquareHorizontalEdgeBlue}
    \lean{TNLean.PEPS.normalSquareHorizontalEdgeComplement}
    \lean{TNLean.PEPS.normalSquareHorizontalEdge_blockingDatum}
    \lean{TNLean.PEPS.normalSquareHorizontalEdge_blockingData}
    \lean{TNLean.PEPS.normalSquareHorizontalEdge_blockingData_of_TCover}
    \lean{TNLean.PEPS.normalSquareHorizontalEdge_injective_chain}
    \lean{TNLean.PEPS.normalSquareHorizontalEdge_blockingDatum_of_TCover}
    \lean{TNLean.PEPS.normalSquareHorizontalEdge_injective_chain_of_TCover}
    \lean{TNLean.PEPS.normalSquareVerticalEdge}
    \lean{TNLean.PEPS.normalSquareVerticalEdge_isVertical}
    \lean{TNLean.PEPS.normalSquareVerticalEdgeRed}
    \lean{TNLean.PEPS.normalSquareVerticalEdgeBlue}
    \lean{TNLean.PEPS.normalSquareVerticalEdgeComplement}
    \lean{TNLean.PEPS.normalSquareVerticalEdge_blockingDatum}
    \lean{TNLean.PEPS.normalSquareVerticalEdge_blockingData}
    \lean{TNLean.PEPS.normalSquareVerticalEdge_blockingData_of_verticalT}
    \lean{TNLean.PEPS.normalSquareVerticalEdge_blockingData_of_verticalTCover}
    \lean{TNLean.PEPS.normalSquareVerticalEdge_blockingDatum_of_verticalT}
    \lean{TNLean.PEPS.normalSquareVerticalEdge_blockingDatum_of_verticalTCover}
    \lean{TNLean.PEPS.normalSquareVerticalEdge_injective_chain}
    \lean{TNLean.PEPS.normalSquareVerticalEdge_injective_chain_of_verticalT}
    \lean{TNLean.PEPS.normalSquareVerticalEdge_injective_chain_of_verticalTCover}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions,
        lem:peps_normalSquareEdgeComplementRegion,
        lem:peps_normalSquareRegionT_edgeBlocks_injective,
        lem:peps_normalSquareEdgeComplementRegion_topCollar}
    In the normalized $5\times7$ horizontal-edge frame, the red block is the
    vertical $2\times3$ block, the blue block is the horizontal $3\times2$
    block, and the complementary block is the finite-lattice edge complement
    $A_3$. The distinguished edge is a horizontal edge of the square-lattice
    graph. The left endpoint of the distinguished edge lies in the red block,
    the right endpoint lies in the blue block, the three regions are pairwise
    disjoint, and their union is the whole finite square lattice.
    If the local $T$-region is injective, rectangular injectivity and union
    closure give injectivity of all three blocks; in particular a rectangular
    cover of $T$ suffices. The normalized $7\times5$ vertical-edge frame
    has the analogous red and blue rectangular blocks, a distinguished vertical
    edge, and the same set-theoretic blocking data; at this stage its
    complementary-block injectivity is either kept as an explicit hypothesis or
    obtained from
    rotated $T$-injectivity, and therefore from a rectangular cover of
    rotated $T$, by the vertical collar lemma. In both cases the named
    blocking datum supplies the corresponding three injective regions for the
    edge-blocked chain.
\end{lemma}
% Maintainer note: horizontal and vertical complement injectivity are reduced
% to local and rotated T-covers, respectively. The translated every-edge
% construction remains a separate step.
\begin{proof}\leanok
    The endpoint memberships are coordinate checks. Rectangular injectivity
    gives the rectangular red and blue blocks. For the horizontal edge, the
    collar lemma gives the complement, and the rectangular cover criterion
    removes the direct $T$-injectivity input when such a cover is supplied.
    For the vertical edge, complement injectivity is an explicit input or is
    obtained from the rotated $T$-cover. The disjointness and cover
    statements are finite-set complement identities.
\end{proof}

\begin{lemma}[Translated horizontal edge-blocking data]%
    \label{lem:peps_normalSquareHorizontalTranslatedEdge_blockingData}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdge}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdgeRed}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdgeBlue}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdgeComplement}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdge_isHorizontal}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdge_eq_rightEdge}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdge_blockingDatum}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdge_injective_chain}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdge_blockingDatum_of_complementCover}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdge_injective_chain_of_complementCover}
    \lean{TNLean.PEPS.horizontalTranslatedEdge_zero}
    \lean{TNLean.PEPS.horizontalTranslatedRed_zero}
    \lean{TNLean.PEPS.horizontalTranslatedBlue_zero}
    \lean{TNLean.PEPS.horizontalTranslatedComp_zero}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions,
        lem:peps_normalSquareEdgeComplementRegion,
        lem:peps_normalSquareEdgeComplementRegion_injective_of_cover}
    For any coordinate origin whose horizontal blocking window lies in the
    finite square lattice, there is a distinguished horizontal edge with the
    same relative position as in the normalized picture. The translated red
    region is the $2\times3$ block, the translated blue region is the
    $3\times2$ block, and the complementary region is the finite-lattice
    complement of their union. If this complementary region is injective,
    these three regions form the one-edge blocking datum; if it has a
    rectangular cover, rectangular injectivity and union closure supply that
    complementary injectivity. In both forms the datum supplies the three
    injective regions used in the edge-blocked chain. The distinguished edge
    is the corresponding coordinate right edge. At coordinate origin $0,0$,
    the translated edge and the three translated regions specialize to the
    normalized horizontal picture.
\end{lemma}
\begin{proof}\leanok
    The endpoint memberships and horizontal orientation are coordinate
    checks. Rectangular injectivity gives the red and blue blocks. The
    complementary block is supplied either as an input or from a rectangular
    cover, and the disjointness and covering identities follow from the
    finite-set complement formulas.
\end{proof}

\begin{lemma}[Translated vertical edge-blocking data]%
    \label{lem:peps_normalSquareVerticalTranslatedEdge_blockingData}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdge}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdgeRed}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdgeBlue}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdgeComplement}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdge_isVertical}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdge_eq_upEdge}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdge_blockingDatum}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdge_injective_chain}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdge_blockingDatum_of_complementCover}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdge_injective_chain_of_complementCover}
    \lean{TNLean.PEPS.verticalTranslatedEdge_zero}
    \lean{TNLean.PEPS.verticalTranslatedRed_zero}
    \lean{TNLean.PEPS.verticalTranslatedBlue_zero}
    \lean{TNLean.PEPS.verticalTranslatedComp_zero}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions,
        lem:peps_normalSquareEdgeComplementRegion_injective_of_cover}
    For any coordinate origin whose vertical blocking window lies in the
    finite square lattice, there is a distinguished vertical edge with the same
    relative position as in the normalized vertical picture. The translated
    red region is the $3\times2$ block, the translated blue region is the
    $2\times3$ block, and the complementary region is the finite-lattice
    complement of their union. If this complementary region is injective,
    these three regions form the one-edge blocking datum; if it has a
    rectangular cover, rectangular injectivity and union closure supply that
    complementary injectivity. In both forms the datum supplies the three
    injective regions used in the edge-blocked chain. The distinguished edge
    is the corresponding coordinate upward edge. At coordinate origin $0,0$,
    the translated edge and the three translated regions specialize to the
    normalized vertical picture.
\end{lemma}
\begin{proof}\leanok
    The endpoint memberships and vertical orientation are coordinate checks.
    Rectangular injectivity gives the red and blue blocks. The complementary
    block is supplied either as an input or from a rectangular cover, and the
    disjointness and covering identities follow from the finite-set complement
    formulas.
\end{proof}

\begin{definition}[Translated edge windows]%
    \label{def:peps_normalSquareTranslatedEdgeWindow}
    \lean{TNLean.PEPS.NormalSquareTranslatedEdgeWindow}
    \lean{TNLean.PEPS.NormalSquareTranslatedEdgeWindow.blockingDatum}
    \lean{TNLean.PEPS.NormalSquareTranslatedEdgeWindow.injective_chain}
    \lean{TNLean.PEPS.NormalSquareTranslatedEdgeWindow.endpoint_disjoint_cover}
    \lean{TNLean.PEPS.normalSquareTranslatedEdgeWindow_rightEdge_margins}
    \lean{TNLean.PEPS.normalSquareTranslatedEdgeWindow_upEdge_margins}
    \lean{TNLean.PEPS.not_leftBoundaryRightEdge_window}
    \lean{TNLean.PEPS.not_leftBoundaryRightEdge_window_seven}
    \lean{TNLean.PEPS.not_rightMarginRightEdge_window}
    \lean{TNLean.PEPS.not_rightMarginRightEdge_window_seven}
    \lean{TNLean.PEPS.not_bottomBoundaryUpEdge_window}
    \lean{TNLean.PEPS.not_bottomBoundaryUpEdge_window_seven}
    \lean{TNLean.PEPS.not_topMarginUpEdge_window}
    \lean{TNLean.PEPS.not_topMarginUpEdge_window_seven}
    \lean{TNLean.PEPS.not_fourSidedBoundaryWindow_seven}
    \lean{TNLean.PEPS.not_forall_normalSquareTranslatedEdgeWindow_seven}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdgeWindow}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdgeWindow}
    \lean{TNLean.PEPS.normalSquareHorizontalTranslatedEdge_sub_eq_rightEdge}
    \lean{TNLean.PEPS.normalSquareVerticalTranslatedEdge_sub_eq_upEdge}
    \lean{TNLean.PEPS.squareLatticeRightEdgeWindow}
    \lean{TNLean.PEPS.squareLatticeUpEdgeWindow}
    \lean{TNLean.PEPS.squareLatticeRightEdgeWindow_of_margins}
    \lean{TNLean.PEPS.squareLatticeUpEdgeWindow_of_margins}
    \lean{TNLean.PEPS.horizontalSquareLatticeEdgeWindow}
    \lean{TNLean.PEPS.horizontalSquareLatticeEdgeWindow_of_margins}
    \lean{TNLean.PEPS.verticalSquareLatticeEdgeWindow}
    \lean{TNLean.PEPS.verticalSquareLatticeEdgeWindow_of_margins}
    \lean{TNLean.PEPS.IsNormalSquareHorizontalEdgeMargins}
    \lean{TNLean.PEPS.IsNormalSquareVerticalEdgeMargins}
    \lean{TNLean.PEPS.NormalSquareEdgeMarginCover}
    \lean{TNLean.PEPS.NormalSquareEdgeMarginCover.window}
    \lean{TNLean.PEPS.NormalSquareEdgeMarginCover.blockingDatum}
    \lean{TNLean.PEPS.NormalSquareEdgeMarginCover.injective_chain}
    \lean{TNLean.PEPS.NormalSquareEdgeMarginCover.endpoint_disjoint_cover}
    \lean{TNLean.PEPS.normalSquareEdgeMarginCover_rightEdge_bounds}
    \lean{TNLean.PEPS.normalSquareEdgeMarginCover_upEdge_bounds}
    \lean{TNLean.PEPS.not_leftBoundaryRightEdge_marginCover}
    \lean{TNLean.PEPS.not_leftBoundaryRightEdge_marginCover_seven}
    \lean{TNLean.PEPS.not_rightMarginRightEdge_marginCover}
    \lean{TNLean.PEPS.not_rightMarginRightEdge_marginCover_seven}
    \lean{TNLean.PEPS.not_bottomBoundaryUpEdge_marginCover}
    \lean{TNLean.PEPS.not_bottomBoundaryUpEdge_marginCover_seven}
    \lean{TNLean.PEPS.not_topMarginUpEdge_marginCover}
    \lean{TNLean.PEPS.not_topMarginUpEdge_marginCover_seven}
    \lean{TNLean.PEPS.not_fourSidedBoundaryMarginCover_seven}
    \leanok
    \uses{lem:peps_normalSquareHorizontalTranslatedEdge_blockingData,
        lem:peps_normalSquareVerticalTranslatedEdge_blockingData}
    A translated edge window records that a chosen edge is one of the
    translated horizontal or vertical edge pictures, together with a
    rectangular cover of the complementary block. Such a window gives the
    one-edge blocking datum for that edge, hence the three injective regions
    used in the edge-blocked chain. The coordinate right and upward edges
    obtain such windows directly from the corresponding translated picture.
    Equivalently, an actual coordinate right or upward edge obtains such a
    window when it has the necessary room around it for the translated
    $5\times7$ or $7\times5$ frame and the complementary cover is supplied.
    The same criterion applies to an arbitrary horizontal or vertical
    square-lattice edge after identifying it with its coordinate right or
    upward representative. The oriented margin-and-cover datum is the same
    criterion written as per-edge data in the rectangular coordinate model;
    it directly supplies the one-edge blocking datum, its three injective
    regions, and the identities
    $u_e\in R_e$, $v_e\in B_e$,
    $R_e\cap B_e=R_e\cap C_e=B_e\cap C_e=\varnothing$, and
    $R_e\cup B_e\cup C_e=V$. The boundary obstruction lemmas show that these
    translated-window and margin-cover criteria are interior criteria, not the
    final every-edge construction in the open square-lattice model.
    % Detailed source-comparison status:
    % docs/paper-gaps/peps_normal_ft_section3_route.tex.
\end{definition}

\begin{theorem}[Construction from translated edge windows]%
    \label{thm:peps_normalSquareTranslatedEdgeBlockingHypotheses_of_windows}
    \lean{TNLean.PEPS.normalSquareTranslatedEdgeBlockingHypotheses_of_windows}
    \lean{TNLean.PEPS.normalSquareTranslatedEdgeBlockingHypotheses_blockingData_of_windows}
    \lean{TNLean.PEPS.normalSquareTranslatedEdgeBlockingHypotheses_injective_chain_of_windows}
    \lean{TNLean.PEPS.%
        normalSquareTranslatedEdgeBlockingHypotheses_endpoint_disjoint_cover_of_windows}
    \lean{TNLean.PEPS.normalSquareEdgeBlockingHypotheses_of_marginCovers}
    \lean{TNLean.PEPS.normalSquareEdgeBlockingHypotheses_blockingData_of_marginCovers}
    \lean{TNLean.PEPS.normalSquareEdgeBlockingHypotheses_injective_chain_of_marginCovers}
    \lean{TNLean.PEPS.normalSquareEdgeBlockingHypotheses_endpoint_disjoint_cover_of_marginCovers}
    \lean{TNLean.PEPS.not_forall_normalSquareEdgeMarginCover_seven}
    \leanok
    \uses{def:peps_normalSquareTranslatedEdgeWindow,
        def:peps_normalEdgeBlockingHypotheses}
    Given rectangular injectivity hypotheses, finite union closure, and a
    translated edge window at every edge of the finite square lattice, the
    corresponding one-edge blockings determine the normal edge-blocking
    hypotheses. This is the conditional construction;
    constructing such windows for all edges of the $7\times7$ lattice remains
    separate; the current open rectangular translated-window criterion does not
    by itself supply such a family. The one-edge datum recovered from the
    resulting hypotheses is the datum supplied by the corresponding translated
    window, and the resulting hypotheses give the three injective regions at
    every edge, together with
    $u_e\in R_e$, $v_e\in B_e$,
    $R_e\cap B_e=R_e\cap C_e=B_e\cap C_e=\varnothing$, and
    $R_e\cup B_e\cup C_e=V$. Equivalently, it is enough to supply the oriented
    margin-and-cover data at every edge; the resulting hypotheses recover the
    corresponding datum, give the same three injectivity conclusions, and give
    the same endpoint, disjointness, and covering conclusions. The obstruction
    theorem records that this interior margin-cover criterion does not by
    itself supply data at every edge of the open $7\times7$ lattice.
    % Detailed source-comparison status:
    % docs/paper-gaps/peps_normal_ft_section3_route.tex.
\end{theorem}
\begin{proof}\leanok
    Apply the general construction for one-edge blocking data to the datum
    supplied by the chosen translated window at each edge.
\end{proof}

\begin{lemma}[Conditional injectivity of the normal $R$ and $S$ regions]%
    \label{lem:peps_normalSquareRegionRS_injective_of_unionClosure}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.regionR_injective_of_union}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.regionS_injective_of_union}
    \leanok
    \uses{def:peps_regionInjectivityData,
        def:peps_normalRectangleInjectivityHypotheses,
        lem:peps_normalSquareRegionRS_rectangular_decomposition}
    If region injectivity is closed under unions, then the coordinate
    square-lattice rectangular injectivity hypotheses imply that the displayed
    regions $R$ and $S$ are injective.
\end{lemma}
\begin{proof}\leanok
    Region $R$ is the union of one injective $2\times3$ rectangle and one
    injective $3\times2$ rectangle. Region $S$ is the union of two
    injective $2\times3$ rectangles. Apply union closure in each case.
\end{proof}

\begin{lemma}[Injectivity of the normal $R$, $S$, and $T$ regions]%
    \label{lem:peps_normalSquareRegionRST_injective}
    \lean{TNLean.PEPS.regionBlockedTensorInjective_normalSquareRegionR}
    \lean{TNLean.PEPS.regionBlockedTensorInjective_normalSquareRegionS}
    \lean{TNLean.PEPS.regionBlockedTensorInjective_normalSquareRegionT}
    \leanok
    \uses{thm:peps_regionInjectivityUnionClosure_of_overlap,
        lem:peps_normalSquareRegionRS_injective_of_unionClosure,
        lem:peps_normalSquareRegionT_injective_of_cover}
    Fix a square-lattice PEPS tensor with every bond dimension positive whose
    blocked tensor is injective on every contiguous $2\times3$ and $3\times2$
    rectangle. Then the displayed regions $R$ and $S$ are injective. If
    moreover the displayed region $T$ has a cover by such rectangles, then $T$
    is injective too. The union closure used here is supplied by the
    overlapping union lemma, so no single-vertex injectivity is assumed.
\end{lemma}
\begin{proof}\leanok
    Apply the conditional injectivity of $R$ and $S$
    (Lemma~\ref{lem:peps_normalSquareRegionRS_injective_of_unionClosure}) and the
    conditional cover criterion for $T$
    (Lemma~\ref{lem:peps_normalSquareRegionT_injective_of_cover}), with the union
    closure discharged by
    Theorem~\ref{thm:peps_regionInjectivityUnionClosure_of_overlap}.
\end{proof}

% The comparison square and comparison region at a parametric offset, used to
% place the one-site comparison pair $R\subseteq S=R\cup\{v\}$ afresh at every
% vertex of the open square lattice.

\begin{definition}[Comparison square, comparison region, and interior window]%
    \label{def:peps_normalSquareComparisonRegions}
    \lean{TNLean.PEPS.normalSquareComparisonSquare}
    \lean{TNLean.PEPS.normalSquareComparisonRegion}
    \lean{TNLean.PEPS.IsNormalSquareComparisonWindow}
    \leanok
    \uses{def:peps_squareLatticeCoordinateRegions}
    At a coordinate offset $(a,b)$ the comparison square is the $3\times3$
    contiguous coordinate rectangle with lower-left corner $(a,b)$. The
    comparison region is this square with its lower-left corner removed,
    presented as the union of the $2\times3$ contiguous rectangle at
    $(a+1,b)$ and the $3\times2$ contiguous rectangle at $(a,b+1)$. These
    are the displayed regions $S$ and $R$ of the final one-site comparison,
    placed at an arbitrary offset rather than at a fixed origin. The interior
    window at $(a,b)$ for a finite $n\times m$ lattice is the placement margin
    $4\le a$, $a+9\le n$, $4\le b$, $b+9\le m$, under which every boundary edge
    of the comparison square and of the comparison region carries the interior
    margins of the open edge-blocking geometry.
\end{definition}

\begin{lemma}[Membership in the comparison region]%
    \label{lem:peps_normalSquareComparisonRegion_membership}
    \lean{TNLean.PEPS.mem_normalSquareComparisonRegion}
    \lean{TNLean.PEPS.notMem_normalSquareComparisonRegion}
    \lean{TNLean.PEPS.insert_normalSquareComparisonRegion}
    \leanok
    \uses{def:peps_normalSquareComparisonRegions,
        lem:peps_squareLatticeCoordinateRegion_identities}
    For any offset $(a,b)$ a vertex with coordinates $(x,y)$ lies in the
    comparison region at $(a,b)$ if and only if
    \[
        (a+1 \le x < a+3 \;\text{and}\; b \le y < b+3)
        \;\lor\;
        (a \le x < a+3 \;\text{and}\; b+1 \le y < b+3).
    \]
    Let $v$ be a lattice vertex and place the comparison region at $v$'s own
    coordinates. Then $v$ lies outside that region, and inserting $v$ back in
    recovers the full $3\times3$ comparison square placed at $v$'s coordinates.
\end{lemma}
\begin{proof}\leanok
    Membership unfolds the union of the two contiguous rectangles into the
    displayed disjunction of coordinate bounds. The corner offset of $v$ fails
    both clauses of the disjunction, so $v$ lies outside; and a vertex lies in
    the insertion exactly when it equals $v$ or already lies in the region,
    which is the membership condition of the $3\times3$ square, the corner
    being the one cell of the square that both pieces omit.
\end{proof}

\begin{lemma}[Band decomposition of the comparison square complement]%
    \label{lem:peps_normalSquareComparisonSquare_complementBands}
    \lean{TNLean.PEPS.compl_normalSquareComparisonSquare_eq_bands}
    \leanok
    \uses{def:peps_normalSquareComparisonRegions}
    On a finite $n\times m$ lattice, if $a+3\le n$ and $b+3\le m$ then the
    complement of the comparison square at $(a,b)$ is the union of the four
    coordinate bands
    \begin{align*}
        &\{(x,y) : 0 \le x < a,\; 0 \le y < m\}
        \;\cup\;
        \{(x,y) : a+3 \le x < n,\; 0 \le y < m\} \\
        \;\cup\;
        &\{(x,y) : a \le x < a+3,\; 0 \le y < b\}
        \;\cup\;
        \{(x,y) : a \le x < a+3,\; b+3 \le y < m\}.
    \end{align*}
\end{lemma}
\begin{proof}\leanok
    Expand the set-difference membership into coordinate inequalities and verify
    that a vertex lies outside the square exactly when it lies in one of the four
    displayed bands.
\end{proof}

\begin{lemma}[Band decomposition of the comparison region complement]%
    \label{lem:peps_normalSquareComparisonRegion_complementBands}
    \lean{TNLean.PEPS.compl_normalSquareComparisonRegion_eq_bands}
    \leanok
    \uses{def:peps_normalSquareComparisonRegions}
    On a finite $n\times m$ lattice, if $2\le a$, $1\le b$, $a+3\le n$, and
    $b+3\le m$ then the complement of the comparison region at $(a,b)$ is the
    union of the same four bands of
    Lemma~\ref{lem:peps_normalSquareComparisonSquare_complementBands} together
    with the $3\times2$ corner rectangle
    \[
        \{(x,y) : a-2 \le x < a+1,\; b-1 \le y < b+1\},
    \]
    which contains the omitted corner $(a,b)$.
\end{lemma}
\begin{proof}\leanok
    Expand the set-difference membership into coordinate inequalities and verify
    that a vertex lies outside the region exactly when it lies in one of the four
    bands or in the corner rectangle. The hypotheses $2\le a$ and $1\le b$ keep
    the corner rectangle's lower-left coordinates $(a-2,b-1)$ within the lattice.
\end{proof}

\begin{lemma}[Injectivity of the comparison region and the comparison square]%
    \label{lem:peps_normalSquareComparison_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.comparisonRegion_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.comparisonSquare_injective}
    \leanok
    \uses{def:peps_normalRectangleInjectivityHypotheses,
        def:peps_regionInjectivityData,
        lem:peps_squareLatticeRectangleInjectivity_rectangles,
        lem:peps_normalSquareWideRectangle_injective}
    Assume rectangular injectivity and union closure. On a finite $n\times m$
    lattice, if $a+3\le n$ and $b+3\le m$ then the comparison region at $(a,b)$
    is injective, being the union of a $2\times3$ and a $3\times2$ rectangle,
    and the comparison square at $(a,b)$ is injective, being a $3\times3$
    rectangle.
\end{lemma}
\begin{proof}\leanok
    Apply union closure to the rectangular injectivity of the contained
    $2\times3$ and $3\times2$ rectangles for the region, the square being a
    single large rectangle.
\end{proof}

\begin{lemma}[Injectivity of the comparison complements]%
    \label{lem:peps_normalSquareComparison_complement_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.compl_comparisonSquare_injective}
    \lean{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.compl_comparisonRegion_injective}
    \leanok
    \uses{def:peps_normalRectangleInjectivityHypotheses,
        def:peps_regionInjectivityData,
        lem:peps_normalSquareWideRectangle_injective,
        lem:peps_normalSquareComparisonSquare_complementBands,
        lem:peps_normalSquareComparisonRegion_complementBands}
    Assume rectangular injectivity and union closure. On a finite $n\times m$
    lattice, if $2\le a$, $2\le b$, $a+5\le n$, and $b+5\le m$ then the
    complement of the comparison square at $(a,b)$ is injective, being the union
    of its four surrounding bands, and the complement of the comparison region
    at $(a,b)$ is injective, being that union together with one further corner
    rectangle.
\end{lemma}
\begin{proof}\leanok
    Rewrite each complement by its band decomposition and apply union closure to
    the bands, each a large contiguous rectangle, and to the corner rectangle.
\end{proof}

\begin{definition}[Normal edge-blocking hypotheses]%
    \label{def:peps_normalEdgeBlockingHypotheses}
    \lean{TNLean.PEPS.NormalEdgeBlockingData}
    \lean{TNLean.PEPS.NormalEdgeBlockingHypotheses}
    \leanok
    \uses{def:peps_edge, def:peps_regionInjectivityData}
    Around one edge, the normal proof uses a blocking into three injective
    regions: a red region containing one endpoint, a blue region containing the
    other endpoint, and an injective complementary region. The three regions
    are pairwise disjoint and cover the whole vertex set. The edge-blocking
    hypotheses are precisely this one-edge datum at every edge; the red,
    blue, and complementary regions are read off from it. The local blocking
    picture used in the proof of
    \cite[Theorem~3]{Molnar2018NormalPEPS}:
    \begin{center}
        \TNPEPSNormalEdgeBlockingHypotheses
    \end{center}
\end{definition}

\begin{theorem}[Injectivity supplied by normal edge blocking]%
    \label{thm:peps_normalEdgeBlockingHypotheses_injectiveChain}
    \lean{TNLean.PEPS.NormalEdgeBlockingData.injective_chain}
    \lean{TNLean.PEPS.NormalEdgeBlockingData.endpoint_disjoint_cover}
    \lean{TNLean.PEPS.NormalEdgeBlockingData.endpoint_injective_disjoint_cover}
    \lean{TNLean.PEPS.NormalEdgeBlockingHypotheses.ofBlockingData}
    \lean{TNLean.PEPS.NormalEdgeBlockingHypotheses.ofBlockingData_blockingData}
    \lean{TNLean.PEPS.NormalEdgeBlockingHypotheses.injective_chain_at_edge}
    \lean{TNLean.PEPS.NormalEdgeBlockingHypotheses.endpoint_disjoint_cover_at_edge}
    \lean{TNLean.PEPS.%
        NormalEdgeBlockingHypotheses.endpoint_injective_disjoint_cover_at_edge}
    \leanok
    \uses{def:peps_normalEdgeBlockingHypotheses}
    A family of one-edge blocking data is precisely the normal edge-blocking
    hypotheses. If $H$ is assembled from the family $d_e$, then the one-edge
    datum that $H$ associates with $e$ is $d_e$. Write the endpoints of $e$
    as $u_e<v_e$ in the canonical vertex ordering. If the three regions are
    $R_e$, $B_e$, and $C_e$, then
    \[
        u_e \in R_e,\qquad v_e \in B_e,\qquad
        R_e\cap B_e=R_e\cap C_e=B_e\cap C_e=\varnothing,\qquad
        R_e\cup B_e\cup C_e=V,
    \]
    and the underlying PEPS tensor is injective on $R_e$, $B_e$, and $C_e$.
\end{theorem}
\begin{proof}\leanok
    By construction, the one-edge datum at each edge $e$ is $d_e$.
    The displayed endpoint, disjointness, coverage, and injectivity statements
    are exactly the corresponding assumptions of that one-edge datum.
\end{proof}

\begin{definition}[Normal one-site separation hypotheses]%
    \label{def:peps_normalOneSiteSeparationHypotheses}
    \lean{TNLean.PEPS.NormalOneSiteSeparationHypotheses}
    \leanok
    \uses{def:peps_regionInjectivityData, def:peps_regionComplement}
    For each site, the normal theorem assumes two injective regions with
    injective complements that differ exactly at that site. This is the
    one-site comparison used after the edge gauges have been obtained.
    Diagrammatically, in the square-lattice proof this is the comparison
    between the source regions $R$ and $S=R\cup\{v\}$:
    \begin{center}
        \TNPEPSNormalOneSiteSeparation
    \end{center}
\end{definition}

\begin{theorem}[Membership in a one-site separated region]%
    \label{thm:peps_normalOneSiteSeparation_mem_withSite_iff}
    \lean{TNLean.PEPS.NormalOneSiteSeparationHypotheses.mem_withSite_iff}
    \leanok
    \uses{def:peps_normalOneSiteSeparationHypotheses}
    For the two regions associated to a site $v$, a vertex $w$ belongs to
    the region containing $v$ if and only if either $w=v$, or $w$
    belongs to the comparison region omitting $v$.
\end{theorem}
\begin{proof}\leanok
    If $w=v$, use the distinguished-site membership and nonmembership
    assumptions. If $w\neq v$, use the equality of the two regions away
    from the distinguished site.
\end{proof}

\begin{definition}[Normal square-lattice blocking regions]%
    \label{def:peps_normalSquareBlockingRegions}
    \lean{TNLean.PEPS.NormalSquareBlockingRegions}
    \lean{TNLean.PEPS.normalSquareBlockingRegions_of_TCover}
    \leanok
    \uses{def:peps_normalRectangleInjectivityHypotheses,
        def:peps_regionInjectivityData,
        lem:peps_normalSquareRegionRS_injective_of_unionClosure,
        lem:peps_normalSquareRegionT_injective_of_cover}
    In the translationally invariant square-lattice proof, the source paper
    assumes that every $2\times 3$ and $3\times 2$ region is injective.
    The formal statement here is the corresponding abstract, scope-restricted
    hypothesis bundle: it records the rectangular injectivity assumptions, the
    size bound $n,m\geq 7$ with $|V|=nm$, and three finite regions
    $R$, $S$, and $T$ whose injectivity is assumed directly. It does
    not by itself assert that these regions are the displayed square-lattice
    unions, nor that their translated variants give the edge-centred blocking
    around every edge. Conversely, for the coordinate square lattice with
    $n,m\geq7$, the rectangular hypotheses, union closure, and a rectangular
    cover of the displayed $T$-region yield this abstract $R,S,T$
    structure, with $R$ and $S$ obtained from their rectangular
    decompositions and $T$ obtained from the supplied cover.
    % Current source-comparison status: docs/paper-gaps/peps_normal_ft_section3_route.tex.
    The source regions are:
    \begin{center}
        \TNPEPSNormalRegionsRS

        \vspace{0.5em}

        \TNPEPSNormalRegionT
    \end{center}
\end{definition}

\begin{theorem}[Normal edge blocking to a three-site injective chain]%
    \label{thm:peps_normalEdgeBlockingToInjectiveChain}
    \notready
    \uses{lem:peps_injectiveRegion_union,
        def:peps_normalSquareBlockingRegions,
        def:peps_normalEdgeBlockingHypotheses,
        thm:peps_normalEdgeBlockingHypotheses_injectiveChain,
        thm:peps_edgeBlockedInsertionAlgebraIsomorphism}
    Under the square-lattice normality hypotheses of
    \cite[Theorem~3]{Molnar2018NormalPEPS}, the regions $R$, $S$, and
    $T$, together with their translated variants, give an edge-centred
    blocking into red, blue, and complementary injective regions. These are
    the regions on which the three-site injective-chain insertion argument is
    applied:
    \begin{center}
        \TNPEPSNormalEdgeBlockingReduction
    \end{center}
\end{theorem}

\begin{theorem}[Translationally invariant normal gauge absorption]%
    \label{thm:peps_tiNormalGaugeAbsorption}
    \notready
    \uses{thm:peps_normalEdgeBlockingToInjectiveChain,
        thm:peps_edge_gauge_absorption}
    In the translationally invariant normal square-lattice case, the edge
    gauges obtained after blocking are represented by one horizontal matrix
    $X$ and one vertical matrix $Y$, unique up to scalar. Absorbing these
    matrices into $B$ gives a tensor $\widetilde B$, equivalently the final
    local gauge formula of \cite[Theorem~3]{Molnar2018NormalPEPS}:
    \begin{center}
        \TNPEPSTINormalGaugeAbsorption
    \end{center}
    The torus form of this absorption, with the gauge family carried onto
    every edge by the translations from one matrix per orientation class, is
    part of Theorem~\ref{thm:fundamentalTheorem_normalTorusPEPS_unconditional}.
    The present open-lattice entry awaits a translation-invariance predicate
    on the open lattice, which would make the per-edge gauges of the edge
    blocking one horizontal and one vertical matrix.
\end{theorem}

\begin{theorem}[Fundamental Theorem for normal square-lattice PEPS]%
    \label{thm:fundamentalTheorem_normalSquarePEPS}
    \notready
    \uses{thm:peps_tiNormalGaugeAbsorption,
        thm:peps_oneVertexComplementComparison,
        def:GaugeEquivPEPS}
    Let $A$ and $B$ be translationally invariant normal square-lattice PEPS
    tensors such that every $2\times 3$ and $3\times 2$ region is
    injective. If they generate the same state on an $n\times m$ region with
    $n,m\geq 7$, then they are related by horizontal and vertical gauges
    $X,Y$, up to a scalar $\lambda$ satisfying
    $\lambda^{nm}=1$. This is
    \cite[Theorem~3]{Molnar2018NormalPEPS}.
    Theorem~\ref{thm:fundamentalTheorem_normalSquarePEPS_unconditional} proves
    an interior-window form of this statement from the source hypotheses alone,
    without translation invariance and with per-vertex scalars in place of the
    single $\lambda$; the gauge-equivalence conclusion on a general connected
    graph, from region injectivity alone, is
    Theorem~\ref{thm:fundamentalTheorem_normalPEPS_connected}. The present
    source-exact entry awaits a translation-invariance predicate on the open
    lattice, which would make the per-edge gauges one horizontal and one
    vertical matrix and the per-vertex scalars a single $\lambda$ with
    $\lambda^{nm}=1$.
\end{theorem}

\begin{theorem}[Fundamental Theorem for normal square-lattice PEPS,
        interior-window form]%
    \label{thm:fundamentalTheorem_normalSquarePEPS_unconditional}
    \lean{TNLean.PEPS.fundamentalTheorem_normalSquarePEPS_unconditional}
    \leanok
    \uses{thm:fundamentalTheorem_normalSquarePEPS,
        def:peps_normalRectangleInjectivityHypotheses,
        def:applyGauge,
        def:peps_edgeInsertedCoeff}
    Let $A$ and $B$ be PEPS tensors on the open $n\times m$ square lattice
    with the same positive physical dimension, equal and positive bond
    dimensions, generating the same state, and such that every contiguous
    $2\times 3$ and $3\times 2$ coordinate rectangle of either tensor is
    injective.  Then there is a family of invertible matrices, one on each
    edge, such that:
    \begin{enumerate}
        \item at every edge with interior margins (for a horizontal edge with
              left endpoint $(x,y)$: $3\le x$, $x+7\le n$, $4\le y$,
              $y+6\le m$; for a vertical edge the mirrored margins), inserting
              any matrix on the edge of $A$ gives the same coefficient as
              inserting it on the corresponding edge of the gauge-absorbed $B$,
              for every physical configuration; and
        \item every vertex $v=(x,y)$ of the interior comparison window
              $4\le x\le n-9$, $4\le y\le m-9$ carries a nonzero scalar
              $\lambda_v$ such that $A_v$ equals $\lambda_v$ times the gauge
              action of the family on $B$ at $v$, on every virtual configuration
              of the incident edges and every physical index.
    \end{enumerate}
    This is the open-lattice form of \cite[Theorem~3]{Molnar2018NormalPEPS},
    restricted in two ways. First, the conclusion reaches only the interior:
    the absorbed equality holds at the edges with interior margins and the
    per-vertex relation at the window vertices, while vertices near the
    lattice boundary are not reached, the blocking frames of the open lattice
    requiring interior margins. Second, the scalars of distinct vertices are
    not asserted equal, and no condition $\lambda^{nm}=1$ is asserted: the
    single $\lambda$ of the source comes from translation invariance, which is
    not assumed here, and without it a single scalar is false --- rescaling
    the components of $A$ at two interior vertices by $\mu$ and $\mu^{-1}$
    preserves the state and the rectangular injectivity but separates the two
    scalars. No lower bound on $n$ or $m$ is assumed; both conclusions are
    empty when the lattice is too small to contain interior edges or window
    vertices.
    \begin{proof}
        \leanok
        \uses{lem:peps_injectiveRegion_union_pos}
        The union closure of injective regions follows from the positive bond
        dimensions and Lemma~\ref{lem:peps_injectiveRegion_union_pos}. The
        translated blocking frames at the edges with interior margins produce
        per-edge gauges, whose absorption into $B$ gives the absorbed
        inserted-coefficient equality at every such edge. A gauge preserves
        blocked-region linear independence, so at every window vertex the
        region comparison applies to the gauge-absorbed tensor: comparing the
        $3\times 3$ square at the vertex minus its corner with its one-site
        completion yields two proportionality scalars $c_R$ and $c_S$, every
        boundary edge of either region carrying interior margins, and the
        inserted-site extraction gives the per-vertex relation with the ratio
        $\lambda_v = c_S/c_R$.
    \end{proof}
\end{theorem}
