\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Proof-Route Alignment for the Periodic Overlap Dichotomy
  (de las Cuevas--Cirac--Schuch--P\'erez-Garc\'ia)}
\author{}
\date{2026-06-04}

\begin{document}
\maketitle

\section{Scope}

This note records, in mathematical terms, where the formal development of the
periodic overlap dichotomy chooses a proof route that differs from the appendix
argument of \cite[Appendix~A]{DeLasCuevas2017Irreducible}, and where its
statements restrict the source.\footnote{Formal files:
\leanid{TNLean/MPS/Periodic/Overlap/}. Local source:
\leanid{Papers/1708.00029/main.tex}. Paper-level umbrella:
\href{https://github.com/LionSR/TNLean/issues/619}{issue \#619}; proposition-level
subtracking: \href{https://github.com/LionSR/TNLean/issues/81}{issue \#81}
(self-overlap limit and non-matching decay,
\href{https://github.com/LionSR/TNLean/issues/448}{issue \#448}; sector-match
propagation and gauge assembly,
\href{https://github.com/LionSR/TNLean/issues/450}{issue \#450}).}

The source proposition is \emph{equal-or-orthogonal-generalized}.\footnote{Stated
in \cite{DeLasCuevas2017Irreducible}, lines 589--609, and proved in Appendix~A,
lines 903--1118.} None of the deviations below is unfaithful in the
smuggled-hypothesis sense: each formal statement is a faithful (or explicitly
restricted) version of the source, and the route substitutions are
mathematically equivalent.

\section{Different-period decay via the peripheral spectrum}

\textbf{Source.}\footnote{\cite{DeLasCuevas2017Irreducible}, Appendix~A,
lines 917--950.} For periods $m_a\neq m_b$, the paper blocks at
$p=\operatorname{lcm}(m_a,m_b)$ and uses Lemma~\emph{bdcf}: the blocks
$P_uA^{(p)}$ and $Q_vB^{(p)}$ are non-repeated normal tensors. Applying the
one-site translation operator to $Q_vB^{(p)}$ versus $Q_{v+1}B^{(p)}$ yields a
contradiction unless $\lim_N|\langle V_N(A)|V_N(B)\rangle|=0$.

\textbf{\Lean{} route.}\footnote{Formal declaration:
\leanid{periodicOverlap_tendsto_zero_of_ne_period} in \leanid{Case1.lean}.}
The same obstruction is realized spectrally. For a tensor $A$, write
$\mathcal E_A$ and $\mathcal E_A^*$ for the transfer map and its adjoint,
defined by
\begin{align}
    \mathcal E_A(Y) &= \sum_i A^iYA^{i\dagger},&
    \mathcal E_A^*(Y) &= \sum_i A^{i\dagger}YA^i .
\end{align}
If $D_a=D_b$ and $B^i=\zeta\,XA^iX^{-1}$, the comparison is
\begin{align}
    \mathcal E_B(Y)
        &= \zeta\bar\zeta\,
           X\mathcal E_A(X^{-1}YX^{-\dagger})X^\dagger .
    \tag{Case 1}
\end{align}
Eigenvalue uniqueness for an irreducible completely positive map
(\cite{Wolf2012Quantum}, Theorem~6.3) forces $|\zeta|=1$, so $\mathcal E_A$ and
$\mathcal E_B$ have equal peripheral spectra. An $m$-periodic tensor has
peripheral spectrum $\{e^{2\pi i r/m}:0\le r<m\}$, so equality forces $m_a=m_b$.
Thus different periods exclude gauge-phase equivalence, and
\emph{equal-or-orthogonal} (the irreducible trace-preserving overlap dichotomy)
gives the decay. This substitutes the paper's lcm-blocking + translation
argument; it is complete and proved.

\section{Sector non-repetition via the blocked fixed-point structure}

\textbf{Source.}\footnote{Lemma~\emph{bdcf} in
\cite{DeLasCuevas2017Irreducible}, lines 404--423.} For a periodic block,
$\mathcal E_A$ is irreducible with peripheral spectrum $\{\omega^r\}$,
$\omega=e^{2\pi i/m}$. The blocked map $\mathcal E_A^m$ then has $1$ as its only
modulus-one eigenvalue (multiplicity $m$), with fixed-point set exactly
$\{P_u\Lambda_AP_u\}_u$, while the adjoint $\mathcal E_A^{*m}$ has fixed points
$\{P_u\}_u$. If, for $u\neq v$, a gauge-phase equivalence
$C_u^{\mathbf i}=e^{i\xi}UC_v^{\mathbf i}U^\dagger$ held with $U=P_uUP_v$
($UU^\dagger=P_u$, $U^\dagger U=P_v$), then
\begin{align}
    \mathcal E_A^m(U)
        &= \sum_{\mathbf i} C_u^{\mathbf i}UC_v^{\mathbf i\dagger} \\
        &= e^{i\xi}U
           \sum_{\mathbf i} C_v^{\mathbf i}C_v^{\mathbf i\dagger} \\
        &= e^{i\xi}U .
        \tag{$\mathcal E_{C_v}(P_v)=P_v$}
\end{align}
so $U$ is a modulus-one eigenvector of $\mathcal E_A^m$. As the only such
eigenvalue is $1$ with the diagonal fixed points $\{P_w\Lambda_AP_w\}$ and
$U=P_uUP_v$ is off-diagonal for $u\neq v$, this is a contradiction.

\textbf{\Lean{} status.}\footnote{Formal declaration:
\leanid{not_gaugePhaseEquiv_of_orthogonal_cyclicSector_traces} in
\leanid{SelfOverlap.lean}; the proof is still open.} The lemma statement is the
faithful non-repetition fact:
distinct cyclic sectors (orthogonal projectors $P_uP_v=0$) are not gauge-phase
equivalent. An earlier draft proposed to discharge it by a ``BDCF converse''
(orthogonal sectors are eventually linearly independent, so a gauge-phase
equivalence would force two sector matrix-product-vector families to be
proportional). That is \emph{not} the source argument. The faithful route is the
spectral one above, which reuses the peripheral-spectrum / fixed-point machinery
already invoked in the different-period proof.\footnote{Formal declaration:
\leanid{period_eq_of_gaugePhaseEquiv_of_isPeriodic} in \leanid{Case1.lean}.}
The docstrings have been realigned to the source argument; the remaining work is
to port it.

\textbf{Signature correction (periodicity is load-bearing).} An adversarial
proof-attempt found that the lemma was \emph{underspecified}: with only the bare
cyclic-projection data (orthogonal $P_k$ summing to $1$, the adjoint shift
$\mathcal E_A^{*}(P_{k+1})=P_k$, blocked-letter commutation, the trace formula,
and orthogonality), the conclusion is \emph{false} --- a reducible $A=B\oplus B$
(for an irreducible period-$m$ block $B$) admits the same cyclic-projection data
yet has gauge-phase-equivalent distinct sectors. The spectral argument genuinely
requires $A$ to be a periodic (irreducible) block: that is what makes $\mathcal
E_A$ irreducible with the stated peripheral spectrum and the sectors primitive.
The hypothesis \leanid{hP : IsPeriodic m A} was therefore added to the lemma
signature (it is supplied at the unique call site
\leanid{sectorBlocks_not_gaugePhaseEquiv_of_ne}). The \leanid{sorry} now sits on a
\emph{true} statement; the remaining work is to port the spectral proof.

\section{Sector-match case: propagation and the phase assembly}

\textbf{Source.}\footnote{\cite{DeLasCuevas2017Irreducible}, Appendix~A,
lines 961--1117; the initial blocked equation \emph{eq:Nm} appears at
lines 978--984.} Given one matched pair, applying the
translation operator $T^l$ to the blocked equation \emph{eq:Nm}
and invoking Theorem~2.10 of \cite{CiracPerezGarcia2017Matrix} produces, at each
offset, a phase and a unitary $U_{\tilde v+l}=P_{\tilde u+l}U_{\tilde
v+l}Q_{\tilde v+l}$, so the offset
$v-u=q$ is constant. Then each blocked product
$F_u$ becomes injective after $N_0$ repetitions, with a right
inverse $\Omega_u$; contracting the concatenated products
with the $\Omega_u$ recovers per-site proportionality $A_u^i=\kappa_v\,
e^{i\eta/m}\,B_v^i$. The phase
bookkeeping is load-bearing: $\prod_v\kappa_v=1$ (\emph{eq:prodkappaprop}) and
$|\kappa_v|=1$ from $\|\sum_iA_u^{i\dagger}A_u^i\|=1$, so $\kappa_v=e^{i\theta_v}$
with $\sum_v\theta_v=0$; choosing $\phi_v$ with $\theta_v=\phi_v-\phi_{v+1}$
telescopes the per-sector phases into a single global phase $\xi=\eta/m$ and a
single global unitary $U=\sum_u e^{i\phi_{u+q}}P_uU_{u+q}Q_{u+q}$, giving
$A^i=e^{i\xi}UB^iU^\dagger$.

\textbf{\Lean{} status.}\footnote{Formal declarations:
\leanid{sectorMatch_propagation},
\leanid{sectorTensor_proportional_of_blockedMatch},
\leanid{repeatedBlocks_of_blockedSectorGaugePhase}, and
\leanid{periodicOverlap_gaugeEquiv_of_sector_match} in \leanid{Case3.lean}.}
The statements track this two-stage structure: the $T^l$ + Theorem~2.10
staircase, followed by the $\Omega_u$ contraction and the
$\kappa/\theta/\phi$ phase assembly. The top-level theorem now invokes these
assembly lemmas and is proved modulo the two remaining leaf obligations. The
docstrings have been realigned to cite the source equation labels and line
ranges (replacing earlier published-version labels such as ``A.8'' and
``A.14--A.18'' that do not occur in the local source), and the
$\kappa/\theta/\phi$ phase assembly is now stated explicitly as load-bearing.

\textbf{Formalized single-site inputs.} The single-site off-diagonal
decomposition is now formalized from the cyclic adjoint-shift
relation.\footnote{The formal declarations are
\leanid{offDiag_shift_of_adjoint_cyclic_shift},
\leanid{eq_sum_offDiag_of_adjoint_cyclic_shift},
\leanid{IsCyclicSectorDecomp.offDiag_shift}, and
\leanid{IsCyclicSectorDecomp.eq_sum_offDiag}.} It gives
$P_{u+1}A^i=A^iP_u$ and $A^i=\sum_u P_{u+1}A^iP_u$ for a chosen cyclic
sector decomposition. This is the inverse-indexed form of
\emph{eq:Aoffdiag}, lines~335--339; the later \emph{eq:Auprop}, line~1014,
defines the sector letters $A_u^i=P_uA^iP_{u+1}$.

The construction of the cyclic sectors also now records the letter-level
corner identity.\footnote{Formal declaration:
\leanid{exists_cyclic_sector_decomp_with_letter_after_blocking_of_isPeriodic}.} If
$\varphi_u$ is the compression isomorphism from the compressed sector algebra
onto $P_uM_DP_u$, then
\[
    \varphi_u(C_u^{\mathbf i})=
        P_u (A^{(m)})^{\mathbf i} P_u .
\]
This is the exposed formal version of the blocked corner-letter relation in
Lemma~\emph{bdcf} and \emph{Cu}. It supplies the missing correspondence between
a compressed sector gauge and the ambient corner in the Step~2 spectral route;
it does not yet construct the resulting off-diagonal modulus-one eigenvector.

Let $O_{C,D}(N):=\langle V_N(C)\mid V_N(D)\rangle$. The one-site
sector-overlap translation is also formalized.\footnote{Formal declaration:
\leanid{sectorOverlap_succ_eq_of_cyclicSectorDecomp}.} It proves, for every
$N>0$,
\[
    O_{A_{u+1},B_{v+1}}(N)=O_{A_u,B_v}(N).
\]
It is the overlap-sum form of the translation step in lines~985--1002, using
the single-site relations inside the paired trace sum and the repository's
inverse cyclic indexing convention.

\textbf{Remaining Case-3 input.} After these single-site inputs, the open
Case-3 leaf is the contraction and phase-assembly step:
\begin{itemize}
  \item \emph{Per-sector corner unitaries}: upgrading the compression LinearEquiv
    $\varphi_u$ to a multiplicative $\ast$-isomorphism with a unitary intertwiner
    $U_v=P_uU_vQ_v$ on $\mathrm{Fin}\,D$ (\emph{eq:Cvprop}); and the $m$-factor
    cyclic $\Omega_u$-contraction with the $\eta$ bookkeeping (the two-site
    \leanid{tensor_proportional} is its $m=2$ shadow) plus the $\kappa/\theta/\phi$
    telescoping into the global unitary $U=\sum_u e^{i\phi_{u+q}}P_uU_{u+q}Q_{u+q}$.
\end{itemize}
This is the remaining obligation for the repeated-blocks assembly
theorem.\footnote{Formal declaration:
\leanid{repeatedBlocks_of_blockedSectorGaugePhase}.} The one-step
transport theorem is closed by the single-site overlap theorem above together
with the overlap-to-gauge theorem and the proved sector primitivity.\footnote{
The formal declarations are
\leanid{sectorGaugePhaseEquiv_succ_of_cyclicTransport} and
\leanid{gaugePhaseEquiv_of_overlap_norm_tendsto_one_of_irreducible_TP}.}

\section{Scope restriction: independence without spanning}

\textbf{Source.}\footnote{\cite{DeLasCuevas2017Irreducible}, lines 604--608
and Lemma~\emph{Lem1t}, lines 511--519. The phrase about the name
``basis of periodic vectors'' appears at line 611.} As a consequence of the
proposition and Lemma~\emph{Lem1t}, for $N\ge N_0$ the non-zero members of
$\{|V_N(A_j)\rangle\}_{j\in J}$ are linearly independent \emph{and} span
$|V_N(A)\rangle$; the spanning half is what ``justifies the name basis of
periodic vectors''.

\textbf{\Lean{} statement.}\footnote{Formal declaration:
\leanid{periodicBasis_eventuallyLinearlyIndependent} in \leanid{Dichotomy.lean};
the matching blueprint entry is in
\path{blueprint/src/chapter/ch22_periodic_ft.tex}.} Only the independence
half is stated; the spanning clause is dropped. The basis condition is encoded
directly as the pairwise non-repetition hypothesis \leanid{hNonrep} together with
a common-period divisibility \leanid{hDiv}, rather than derived from a basis of
periodic tensors, and the \emph{Lem1t} almost-orthonormal-implies-independent
lemma is not used. Restoring the spanning clause and routing through
\emph{Lem1t} is tracked under
\href{https://github.com/LionSR/TNLean/issues/81}{issue \#81}.

\begin{thebibliography}{9}
\bibitem{DeLasCuevas2017Irreducible}
G.~de las Cuevas, J.~I.~Cirac, N.~Schuch, and D.~P\'erez-Garc\'ia,
\emph{Irreducible forms of matrix product states: Theory and applications},
J.~Math.~Phys.\ \textbf{58}, 121901 (2017), arXiv:1708.00029.

\bibitem{CiracPerezGarcia2017Matrix}
J.~I.~Cirac, D.~P\'erez-Garc\'ia, N.~Schuch, and F.~Verstraete,
\emph{Matrix product states and projected entangled pair states: Concepts,
symmetries, theorems}, Ann.\ Phys.\ \textbf{378}, 100 (2017)
(``Ci15'' in arXiv:1708.00029).

\bibitem{Wolf2012Quantum}
M.~M.~Wolf, \emph{Quantum Channels and Operations: Guided Tour}, unpublished
lecture notes (2012).
\end{thebibliography}

\end{document}
