\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{docs/paper-gaps/command}

\title{Theorem Surface Audit for Canonical Form, BNT, and the Non-periodic MPS Fundamental Theorem}
\author{}
\date{2026-05-05}

\begin{document}
\maketitle

\section{Purpose}

This note compares the theorem surface in the formal development with the
statements in the source papers.  It is not an issue index.  The goal is to
separate four kinds of results:
\begin{enumerate}
  \item named statements appearing directly in the papers;
  \item cited mathematical results used by the papers;
  \item auxiliary formal statements needed to make an implicit proof step exact;
  \item stronger or alternative statements that should not be presented as part
        of the paper-level proof unless they become necessary.
\end{enumerate}

\section{Named source statements}

The 2016 source has a small number of named statements in the canonical-form
and BNT part:
\begin{itemize}
  \item normal tensor and canonical form definitions
    \cite[lines~233--255]{Cirac2016MPDO_arXiv};
  \item basis of normal tensors definition and characterization
    \cite[lines~271--280]{Cirac2016MPDO_arXiv};
  \item injective and block-injective canonical form definitions, together with
    the quantum-Wielandt and block-injective-after-blocking assertions
    \cite[lines~317--345]{Cirac2016MPDO_arXiv};
  \item the proportional Fundamental Theorem and the equal-MPV corollary
    \cite[lines~347--356]{Cirac2016MPDO_arXiv}.
\end{itemize}
The 2021 review gives the same normal tensor, BNT, proportional theorem, and
equal-MPV corollary in Section~IV
\cite[lines~1828--1899]{Cirac2021Matrix}.

These are the paper-level targets.  They should remain the prominent theorem
statements in the blueprint.

\section{Why the formalization has more declarations}

The paper proof suppresses finite-dimensional and finite-index operations that
must be named in a proof assistant.  In this part of the development the extra
formal statements usually fall into one of the following categories.

\paragraph{Reduction statements.}
The canonical-form proof splits the paper's ``put the tensor in canonical
form'' step into invariant-subspace extraction, block-diagonal decomposition,
trace-preserving normalization, period-removing blocking, and primitive-block
normality.  These are not new mathematical claims beyond the source reduction,
but they need separate hypotheses and conclusions so later theorems can apply
them.

\paragraph{Common blocking statements.}
The paper says that periodicity is removed by blocking a common multiple of the
periods.  The formalization must record positivity of the chosen block length,
persistence of injectivity under further blocking, and the simultaneous
application to a finite family of representative sectors.  These statements are
auxiliary, but they are mathematically faithful to the period-removal step.

\paragraph{Wielandt proof-chain statements.}
The source proof of the quantum-Wielandt bound uses Lemma~1, eigenvalue
extraction, vector spreading, and the fixed-length rank-one step as an argument,
not as separate MPS theorem statements
\cite[Theorem~1 proof]{Sanz2010Quantum}.  Formal conjunctions that bundle
these inputs, such as the \emph{Wielandt analysis} and \emph{Wielandt chain}
declarations, should therefore be lemmas.  The visible theorem is the
cumulative or fixed-length Wielandt conclusion, with the stronger fixed-length
claim distinguished from the cumulative span bound.

\paragraph{BNT comparison statements.}
The source BNT characterization selects one representative from each
phase-gauge equivalence class and compares the power sums of weights.  The
formalization separates this into overlap decay, eventual linear independence,
permutation rigidity, sector-weight comparison, and phase absorption.  The
heterogeneous block-matching theorem is a genuine formal conclusion for the
block-matching part, while the final global weighted gauge equivalence remains
the paper-level equal-MPV target.
The overlap-to-linear-independence and change-of-basis declarations are
linear-algebra support lemmas for the permutation-rigidity theorem, not
additional BNT theorems from the source.
The repeated-copy part should be written around the displayed coefficients
\[
  c_{S,j}^{(N)}=\sum_{q=1}^{r_j^S}(\mu^S_{j,q})^N
\]
and the resulting multiset equality
\[
  \{\!\{\mu^S_{j,q}:1\le q\le r_j^S\}\!\}
  =
  \{\!\{\mu^T_{j,q}:1\le q\le r_j^T\}\!\}.
\]
Here the double braces denote multisets, so repeated weights are counted with
their multiplicities.
In this notation, BNT linear independence separates the coefficient arrays, and
Newton--Girard recovers the weights.  The formal declarations that only package
these two algebra steps in sector-data notation are support lemmas, even when
their names mention the Fundamental Theorem.

\section{External lemmas need explanations}

When a result is needed but the MPS source only cites it, the formal
development should give a pedagogical explanation at the point where the result
enters the proof.  A citation alone is not enough in this situation.  The
explanation should include:
\begin{enumerate}
  \item the exact mathematical statement used;
  \item the hypotheses in the MPS notation, after blocking or reindexing if
        necessary;
  \item a short proof idea, or a precise indication that the proof is imported
        from the cited source;
  \item the formal declaration that supplies the result, if it has already been
        formalized.
\end{enumerate}

This applies in particular to the following families of external input.
\begin{itemize}
  \item The quantum Wielandt theorem supplies finite-length matrix-algebra span
    from normality or primitivity \cite{Sanz2010Quantum}.  The paper cites it in
    the canonical-form reduction; the formal text should say whether it is using
    cumulative span, exact-length span, injectivity of the map $\Gamma_L$, or
    block-injectivity of a direct sum.
  \item The block-injective canonical-form bound uses the older MPS
    injectivity and block-injectivity argument of
    \cite{PerezGarcia2007MPSReps}.  The formal statement should identify the
    common blocked physical alphabet and the direct-sum matrix algebra that is
    being spanned.  The direct-sum proof input in the older source is recorded
    in \path{docs/paper-gaps/david2006_direct_sum_input.tex}.
  \item Vandermonde inversion and Newton--Girard identities are elementary, but
    they are not proved in the MPS source.  When they are used to recover
    coefficients, multiplicities, or weight multisets, the blueprint should
    explain the finite linear system or the power-sum recursion.
    Declarations that only restate these finite-sequence facts in BNT notation
    should be lemmas, even when they are applied inside a paper-level theorem.
  \item Perron--Frobenius and peripheral-spectrum facts for quantum channels are
    standard inputs from finite-dimensional channel theory, but they are not
    interchangeable.  A formal theorem should specify whether it uses
    irreducibility, primitivity, trace preservation, a faithful fixed point, or
    a rank-one peripheral projection.
\end{itemize}

This rule also gives a retirement test.  If an external-looking lemma has no
source citation, no proof explanation, and no current proof path needing it, it
should not remain as a headline theorem in the paper-level blueprint.

\section{Dependency-chain classification}

The non-periodic Fundamental Theorem route should be read with the following
separation of proof roles.

\subsection{Final target trace-back}

For the non-periodic route, the source target is the Fundamental Theorem for
canonical-form MPS and its equal-MPV corollary, after the tensor has been put in
normal form and the periodicity has been removed by blocking.  In the notation
of \cite[lines~347--356]{Cirac2016MPDO_arXiv}, the proportional statement
compares two canonical-form tensors with the same family of MPVs up to a scalar
sequence and concludes a block permutation, phases, and blockwise gauges.  The
equal-MPV corollary is the special case in which the scalar sequence is
identically \(1\), but the proof still passes through the BNT coefficient
comparison.

The formal route currently has three distinct layers that should not be
identified with one another.
\begin{enumerate}
  \item The single-block injective theorem
    \path{MPSTensor.fundamentalTheorem_singleBlock} is the injective-block
    gauge theorem.  It is a source-level theorem only for one injective block,
    not for the multi-block BNT comparison.
  \item The strict canonical-form BNT declarations
    \path{MPSTensor.fundamentalTheorem_proportionalMPV_CFBNT} and
    \path{MPSTensor.fundamentalTheorem_equalMPV_CFBNT} are strict
    representative forms.  Their exact restrictions are recorded in the
    paragraph on paper-level targets below.
  \item The sector-decomposition theorem
    \path{MPSTensor.fundamentalTheorem_equalMPV_sectorDecomposition_hetero_of_sectorMatching}
    is the current formal statement closest to the BNT part of the source
    proof: from a supplied sector-basis matching, linear independence, and
    equality of total MPV families, it proves the corresponding equality of
    sector-weight multisets, up to the phases in the matching.  The source-level
    equal-MPV corollary is obtained only after the sector-basis matching itself
    is derived from the source hypotheses.
\end{enumerate}

Thus the proof gap is not that the paper has many hidden Fundamental Theorems.
Rather, the paper suppresses the formal separation between deriving the BNT
comparison data from the source hypotheses and applying the existing comparison
theorems to that data.  The remaining theorem-level work is to obtain the
sector matching, copy-count equalities, and fixed-length direct-sum span
hypotheses from the source assumptions.  Once those inputs are available, the
existing strict and sector-decomposition statements combine to give the
paper-level theorem; they are not additional paper theorems.

\paragraph{Paper-level targets.}
The proportional canonical-form theorem and the equal-MPV corollary are the
source targets.  Their source proof is formulated after choosing BNT
representatives and retaining the repeated-copy coefficients
\(\mu_{j,q}\) in the sums
\[
  \sum_q \mu_{j,q}^N.
\]
The current strict-representative declarations
\path{fundamentalTheorem_proportionalMPV_CFBNT} and
\path{fundamentalTheorem_equalMPV_CFBNT} are therefore special-case statements.
The proportional comparison assumes explicit coefficient data; the equal-MPV
common-block comparison assumes the repeated-copy structure has already been
matched into common weights and dimensions.  Neither statement should be cited
as the general arXiv:1606.00608 theorem unless the missing BNT multiplicity
comparison has already been discharged.  Heterogeneous variants that state the
permutation, phases, gauges, and weight matching for possibly different sector
indices are legitimate theorem statements only when their hypotheses are
obtained from the same BNT comparison.

\paragraph{Conditional after-blocking statements.}
The after-blocking comparison theorems are useful only after the source
canonical-form, BNT selection, block-injective span, and BNT matching inputs have
been supplied.  The common-phase-cover and BNT-cover formulations are
conditional statements using those inputs.  They should stay visibly conditional
until the hypotheses are produced from the paper's assumptions.  In particular,
the finite-length span or phase-cover assumptions are not extra hypotheses in
the source theorem; they are the formal names for proof inputs that still have
to be obtained from the source argument.

\paragraph{Data structures and transport lemmas.}
Objects such as \path{SectorBasisOverlapSpanHypotheses} and
\path{CommonPrimitiveBNTCoverHypotheses}
are formal comparison data for the proof.  They should not be cited as paper
theorems.  Lemmas converting one data structure into another are needed in the
formal proof chain, but they are auxiliary unless they state one of the source
mathematical claims directly.

\paragraph{External finite-dimensional inputs.}
The formal route also contains named statements for linear algebra that the
paper uses without proof: Vandermonde separation, Newton--Girard recovery,
finite-dimensional Perron--Frobenius structure, Burnside--Jacobson algebra
generation, and quantum Wielandt bounds.  These may be important mathematical
theorems in their own right, but in this proof chain they enter as cited inputs.
Their statements should be explained, and their formal names should normally be
lemma-level when they are merely restated in MPS notation.

\section{Current demotions}

The following declarations are auxiliary and should not be read as independent
paper-level theorems.
\begin{itemize}
  \item The explicit-coefficient equal-MPV result is named
    \[
      \texttt{fundamentalTheorem\_equalMPV\_of\_explicit\_coefficients}.
    \]
    It is now lemma-level.  It is not the full equal-MPV corollary from the
    papers, since it assumes the coefficient arrays and their nonzero limits.
  \item The proportional result
    \[
      \texttt{fundamentalTheorem\_proportionalMPV\_CFBNT}
    \]
    is stated with the same explicit coefficient arrays, their limits, and
    non-vanishing assumptions.  The source theorem derives the relevant BNT
    coefficient comparison from the proportional MPV hypothesis; that extraction
    is not part of this formal statement.  This is a genuine restriction of the
    formal route, not merely a theorem-versus-lemma presentation choice.
  \item The unused fixed-size BNT Vandermonde MPV declarations have been
    removed.  The scalar Vandermonde separation lemma remains because it is a
    small algebraic input elsewhere, but it is not presented as part of the
    main BNT comparison route.
  \item The equal-span permutation rigidity statement with asymptotically
    orthonormal overlaps is a lemma.  It is a useful span route, but the
    proportional BNT permutation theorem is the theorem used by the main
    proportional Fundamental Theorem in Chapter~11.
  \item The older theorem from proportional-decomposition data to sector
    comparison without leftover-zero-block bookkeeping has been removed.  The
    active after-blocking route keeps this bookkeeping visible because the
    formal equality relation includes the empty word.  This is not terminology
    from \cite{Cirac2016MPDO_arXiv}; the source only says that canonical forms
    may have zero blocks through the inequality
    \(\sum_k D_k \le D\) \cite[lines~217--219]{Cirac2016MPDO_arXiv}.
  \item The power-sum multiset recovery statement
    \[
      \texttt{power\_sum\_eq\_implies\_multiset\_eq}
    \]
    is a lemma-level restatement of the elementary Newton--Girard input used in
    the equal-MPV proof.  Likewise,
    \[
      \texttt{perBlock\_sameMPV\_of\_equalMPV\_CFBNT}
    \]
    is only the extraction of the blockwise equality conclusion from the
    common-block equal-MPV lemma.  Neither should be counted as an additional source-level
    Fundamental Theorem.
  \item The implication
    \[
      \texttt{sameMPV}_{2}\texttt{\_implies\_proportionalMPV}_{2}
    \]
    from equal MPV families to proportional MPV families with proportionality
    constant \(1\) is also lemma-level.  The displayed subscript \(2\) denotes
    the Unicode subscript \(2\) used in the Lean declaration name.  It is a
    logical conversion between hypotheses, not a separate source theorem.
  \item The BNT sector-weight recovery declarations
    \path{weight_multiset_eq_of_sameMPV_bnt} and
    \path{exists_copies_eq_and_weight_multiset_eq_of_sameMPV_bnt}
    are lemma-level.  They formalize the coefficient comparison
    \(c_{S,j}^{(N)}=c_{T,j}^{(N)}\) and the elementary recovery of copy counts
    and weight multisets; they are not additional BNT theorems beyond the
    source's basis-normal-tensor comparison.
  \item The phase-matched sector-comparison adapters
    \path{fundamentalTheorem_equalMPV_sectorDecomposition_hetero_of_phaseMatch}
    and
    \path{fundamentalTheorem_equalMPV_sectorDecomposition_hetero_of_matched_basis}
    are also lemma-level.  They transport the same coefficient and multiset
    formulas across a supplied phase or matched-basis identification.  The
    theorem-level surface begins only when the proof supplies the actual
    matching data from the BNT hypotheses.
\end{itemize}

\section{Statements to keep visibly conditional}

Several declarations should stay in the development, but their statements
should remain visibly conditional until the missing paper input is produced.
\begin{itemize}
  \item The conditional after-blocking Fundamental Theorem assumes the
    post-blocking sector hypotheses.  It should not be cited as the unconditional
    canonical-form theorem.
  \item The explicit-coefficient equal case assumes coefficient arrays rather
    than deriving the BNT power-sum data from bare equality of MPVs.
\end{itemize}

\section{Retirement criterion}

An auxiliary declaration should be removed or moved out of the main blueprint
route when all three conditions hold:
\begin{enumerate}
  \item it is not a statement in the cited paper and is not a cited outside
        result;
  \item no current proof of a paper-level target uses it;
  \item it states a stronger finite-length, span, or synchronization hypothesis
        than the source proof requires.
\end{enumerate}
If a declaration is still used but only as a formal intermediate result, it
should be a lemma or a clearly conditional theorem, not a headline replacement
for the paper theorem.

\bibliographystyle{alpha}
\bibliography{docs/paper-gaps/references}

\end{document}
