\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{The SAL to Eta-Local Structure Step in
Cirac--Perez-Garcia--Schuch--Verstraete 2017 Appendix C.2}
\author{}
\date{2026-05-27}

\begin{document}
\maketitle

\section{Source statement}

Appendix C.2 of Cirac--P\'erez-Garcia--Schuch--Verstraete proves that a simple
MPDO satisfying the strong area law has the commuting nearest-neighbor product
form
\[
  \sigma^{(N)}({\cal K})\propto \prod_{n=1}^N B_{n,n+1},
\]
where $B_{n,n+1}\geq 0$ and the translated two-site operators commute
\cite[Appendix~C.2, Proposition~C.6]{Cirac2016MPDO_arXiv}. In the local
source file this is Proposition \texttt{3to4},
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:1571--1593}.

The proof uses the sector decomposition obtained earlier in Appendix C.2:
\[
  \tilde\sigma
  =
  \bigoplus_{k_1,\ldots,k_N}
  \bigotimes_{n=1}^N \eta_{k_n,k_{n+1}},
\]
with $\eta_{k_n,k_{n+1}}\geq 0$ acting between the right subspace of site $n$
and the left subspace of site $n+1$. It then defines
\[
  B_{n,n+1}
  =
  \sum_{k_n,k_{n+1}}
  \operatorname{Id}_{H_{L,n}^{(k_n)}}\otimes
  \eta_{k_n,k_{n+1}}\otimes
  \operatorname{Id}_{H_{R,n+1}^{(k_{n+1})}},
\]
regarded as the identity away from sites $n,n+1$. With this definition,
$\tilde\sigma=\prod_n B_{n,n+1}$ and the commutation of the translated
$B$-operators follows from their sector-local action.

\section{Current formalization}

The formal development currently records this as an eta-local structure for
$K$.\footnote{The formal structure is
\leanid{MPOTensor.EtaLocalStructureData}.} It contains the assembled positive
translation-invariant two-site bond and its realization of the finite-chain
MPO. From this structure, the development proves the commuting-form property
and the GSNNCH--ZCL consequences.

The two scope-restricted declarations are assembly
statements.\footnote{These declarations are
\leanid{MPOTensor.SimpleMPDOBlockedRFPData.ofSALZCLAndEtaLocalStructure} and
\leanid{MPOTensor.simple_mpdo_rfp_chain_of_sal_zcl_and_etaLocalStructure}.}
They assume the eta-local structure instead of deriving it from the SAL local
coordinates. They also carry the ZCL hypothesis needed for the subsequent RFP
consequences. Thus they formalize the final assembly step after the
$B_{n,n+1}$ have been constructed, not the source theorem that constructs those
bonds from SAL.

\section{Missing step}

The missing source-faithful theorem must construct the formal eta-local
structure \leanid{MPOTensor.EtaLocalStructureData} from the Appendix C.2 local
data attached to a simple MPDO satisfying SAL. It must not add ZCL as a
hypothesis for this construction. Concretely, it must identify the local
$\eta_{k,h}$ family with the original tensor coordinates, define the
translation-invariant bond
$B_{n,n+1}$ above, prove $B_{n,n+1}\geq 0$, prove the commutation relations,
and prove the finite-chain realization
\[
  \sigma^{(N)}(K)=c_N\prod_{n=1}^N B_{n,n+1}
\]
for positive scalars $c_N$.

ZCL enters only after this commuting product form has been obtained, in the
later step that derives the GSNNCH--ZCL and RFP conclusions.

This is tracked by
\href{https://github.com/LionSR/TNLean/issues/823}{issue \#823}. Until that
construction exists, the scope-restricted assembly statements must not be
presented as the full formalization of Proposition~C.6.

\section{Classification}

This note records an open gap for the SAL-to-bond construction of
\leanid{MPOTensor.EtaLocalStructureData}. The two assembly declarations above
are scope restrictions: they formalize only the post-assembly step after the
nearest-neighbor bonds have already been constructed.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
