\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{./command}

\title{Block-Diagonal Boundary Conditions in the Degenerate Parent-Hamiltonian Ground Space}
\author{}
\date{2026-06-11}

\begin{document}
\maketitle

\section{Notation}

For a tensor $A=(A_i)_{i=1}^d$ with bond dimension $D$, write
\begin{align}
  \Gamma_N^A(X)(i_1,\ldots,i_N)
    =\tr(A_{i_1}\cdots A_{i_N}X)
    \qquad (X\in M_D(\mathbb C)).
\end{align}
Set
\begin{align}
  \mathcal G_N^A
  =
  \{\Gamma_N^A(X):X\in M_D(\mathbb C)\},
  \qquad
  V^{(N)}(A)=\Gamma_N^A(I).
\end{align}
For the ring with periodic boundary conditions, with $\operatorname{res}_{s,L}$
denoting restriction to the length-$L$ interval beginning at
$s\in \mathbb Z/N\mathbb Z$, the local ground-space condition is
\begin{align}
  \mathcal K_{N,L}(A)
  =
  \bigcap_{s\in \mathbb Z/N\mathbb Z}
  \{\psi:\operatorname{res}_{s,L}(\psi)\in\mathcal G_L^A\}.
\end{align}
When weights are present, write
\begin{align}
  \widehat A_i=\bigoplus_{j=1}^b\mu_jA_i^j,
  \qquad
  \mu_j\ne0,
\end{align}
for the corresponding block-diagonal tensor.  The letters $g$ and $b$ both
denote the number of blocks below: $g$ follows the review's block-injective
canonical form notation, while $b$ follows
\cite[Theorem~12]{PerezGarcia2007MPSReps}.

\section{The source assertion}

Cirac et al. describe the degenerate parent-Hamiltonian ground space for a
matrix product state with more than one block in canonical form in Section~IV.C
of \cite{Cirac2021Matrix}.\footnote{For
	traceability, the periodic-boundary block-component comparison was tracked by
	\href{https://github.com/LionSR/TNLean/issues/2651}{issue \#2651}, which is
	now closed. The remaining derivation of the source comparison currently
	assumed by the conditional theorems is tracked by
	\href{https://github.com/LionSR/TNLean/issues/2971}{issue \#2971}. Pull
	request~\#2724 records the conditional reduction on the main branch, reducing
	the unconditional target to the boundary-crossing comparison identities.
	Earlier reductions were tracked by
	\href{https://github.com/LionSR/TNLean/issues/2641}{issue \#2641},
	\href{https://github.com/LionSR/TNLean/issues/905}{issue \#905}, and
	\href{https://github.com/LionSR/TNLean/issues/870}{issue \#870}. The relevant
	source lines are \path{Papers/2011.12127/TN-Review-main.tex}:2112--2128.}
The source's block-injective canonical form is a block-diagonal
physical--virtual correspondence. If
\begin{align}
  \widetilde A^i=\bigoplus_{j=1}^g A_j^i,
\end{align}
then
\begin{align}
  \forall X=(X_j)_j\in\bigoplus_{j=1}^g M_{D_j}(\mathbb C),
  \qquad
  \exists (c_i(X))_i,\qquad
  X=\sum_i c_i(X)\widetilde A^i.
\end{align}
After blocking, the corresponding word-span statement is
\begin{align}
  \exists m_0\quad \forall m\ge m_0,\qquad
  \operatorname{span}\{(A_1^\omega,\ldots,A_g^\omega):|\omega|=m\}
  =
  \prod_{j=1}^g M_{D_j}(\mathbb C).
\end{align}

The current Lean development separates the product-span input from
the finite blocking hypothesis.  First, it proves the one-step intersection
identity from \cite[Theorem~12]{PerezGarcia2007MPSReps} under the displayed
product-span equation.  Then the
normalized basis-of-normal-tensors (BNT) theorems derive that product span in
the finite Condition C1 range from the injectivity of
\begin{align}
  \Gamma_{L_0}^{A_j}:M_{D_j}(\mathbb C)
    \to(\mathbb C^d)^{\otimes L_0},
\end{align}
together with the separated normalized block hypotheses.  Thus the current
Lean statements no longer replace Condition C1 by a length-one span condition
in the boundary-decomposition path.  The remaining mathematical boundary lies
later in the argument: after the cyclic constraints have been propagated into
the trace-form sum, the boundary conditions themselves must be compared.

The resulting parent-Hamiltonian theorem in \cite[Section~IV.C]{Cirac2021Matrix}
is
\begin{align}
  N\ge L_0+1
  \quad\Longrightarrow\quad
  \ker(H_N)=\operatorname{span}\{V^{(N)}(A_j):j=1,\ldots,g\}.
\end{align}
The sentence preceding the theorem restricts the proof to block-diagonal
boundary conditions.  After opening the periodic boundary at the chosen cut, this
means boundary matrices of the form
\begin{align}
  X=\bigoplus_{j=1}^g X_j,
  \qquad
  Y=\bigoplus_{j=1}^g Y_j
  \qquad
  (X_j,Y_j\in M_{D_j}(\mathbb C)).
\end{align}

\section{The block-diagonal intersection identity}

The older representation paper \cite[Theorem \emph{Degeneracy of the ground
space v2}]{PerezGarcia2007MPSReps} gives the finite-chain identity used by the
source argument.\footnote{Source
location: \path{Papers/quant-ph_0608197/MPSarchive.tex}:1424--1456.} It fixes
the block count and local length by
\begin{align}
  b=g,
  \qquad
  L\ge 3(b-1)(L_0+1)+1.
\end{align}
Thus
\begin{align}
  m\ge L
\end{align}
is the range used in the intersection step:
\begin{align}
  \mathbb C^d\otimes\left(\bigoplus_j \mathcal G_m^{A_j}\right)
  \cap
  \left(\bigoplus_j \mathcal G_m^{A_j}\right)\otimes\mathbb C^d
  =
  \bigoplus_j \mathcal G_{m+1}^{A_j}.
\end{align}
For a vector $\psi$ in the left-hand side,
\begin{align}
  \psi
  &=
  \sum_j
  \sum_{i_1,\ldots,i_{m+1}}
  \tr\!\left(A^j_{i_{m+1}} C^j_{i_1} A^j_{i_2}\cdots A^j_{i_m}\right)
  |i_1\cdots i_{m+1}\rangle \notag\\
  &=
  \sum_j
  \sum_{i_1,\ldots,i_{m+1}}
  \tr\!\left(D^j_{i_{m+1}} A^j_{i_1} A^j_{i_2}\cdots A^j_{i_m}\right)
  |i_1\cdots i_{m+1}\rangle.
\end{align}
The direct-sum separation lemma and the one-block injectivity condition identify,
for every $j,i_1,i_{m+1}$,
\begin{align}
  A^j_{i_{m+1}} C^j_{i_1}
  =
  D^j_{i_{m+1}} A^j_{i_1}.
\end{align}
The source line writes \(E^j=\sum_k C^j_kA^j_k\), without an adjoint on the
second factor. The normalization used in the next line is
\(\sum_kA^j_kA^{j\dagger}_k=I\), so the calculation requires the adjointed
matrix
\begin{align}
  E^j=\sum_{k} C^j_{k}A^{j\dagger}_{k}.
\end{align}
Together with the normalization
\begin{align}
  \sum_{i_1} A^j_{i_1}A^{j\dagger}_{i_1}=I,
\end{align}
the displayed matrix identity gives
\begin{align}
  A^j_{i_{m+1}} E^j A^j_{i_1}
  &=
  \sum_k
  A^j_{i_{m+1}} C^j_k A^{j\dagger}_k A^j_{i_1} \notag\\
  &=
  D^j_{i_{m+1}}
  \left(\sum_k A^j_k A^{j\dagger}_k\right)
  A^j_{i_1} \notag\\
  &=
  D^j_{i_{m+1}} A^j_{i_1}
  =
  A^j_{i_{m+1}} C^j_{i_1},
\end{align}
which puts the vector in $\bigoplus_j \mathcal G_{m+1}^{A_j}$.

\section{The remaining boundary condition}

The present Lean statements prove the one-step identity under the product-span
hypothesis
\begin{align}
  \operatorname{span}
  \{(A^1_w,\ldots,A^b_w): |w|=m\}
  =
  \prod_{j=1}^b M_{D_j}(\mathbb C),
\end{align}
and, in the normalized BNT case, give the internal direct sums
\begin{align}
  \bigvee_j \mathcal G_m^{A_j}
  =
  \bigoplus_j \mathcal G_m^{A_j},
  \qquad
  \bigvee_j \mathcal G_{m+1}^{A_j}
  =
  \bigoplus_j \mathcal G_{m+1}^{A_j}.
\end{align}
The proved one-step conclusion is therefore
\begin{align}
  \mathbb C^d\otimes\left(\bigoplus_j \mathcal G_m^{A_j}\right)
  \cap
  \left(\bigoplus_j \mathcal G_m^{A_j}\right)\otimes\mathbb C^d
  =
  \bigoplus_j \mathcal G_{m+1}^{A_j}
\end{align}
for all sufficiently large $m$ satisfying the product-span hypotheses.\footnote{
Lean declarations:
\leanid{MPSTensor.pgvwc07_directSum_restriction_intersection_of_wordTupleSpanTop},
\leanid{MPSTensor.groundSpace_iSupIndep_of_wordTupleSpanTop}, and
\leanid{MPSTensor.pgvwc07_directSum_restriction_intersection_of_ge_of_bnt_directSum_unital}.}
The corresponding finite-Condition-C1 declarations then propagate the cyclic
local constraints to
\begin{align}
  \mathcal K_{N,L}(\widehat A)\subseteq
  S_N:=\bigvee_{j=1}^b\mathcal G_N^{A_j},
\end{align}
and make the sum defining $S_N$ internal.\footnote{Lean declarations:
\leanid{MPSTensor.chainGroundSpace_toTensorFromBlocks_le_iSup_groundSpace_of_ge_of_bnt_directSum_unital_c1},
\leanid{MPSTensor.chainGroundSpace_toTensorFromBlocks_le_iSup_and_iSupIndep_of_bnt_unital_c1},
\leanid{MPSTensor.exists_unique_sum_groundSpace_of_chainGroundSpace_toTensorFromBlocks_of_bnt_unital_c1},
and
\leanid{MPSTensor.chainGroundSpace_toTensorFromBlocks_two_inclusions_and_iSupIndep_of_bnt_unital_c1}.}
What remains is the source boundary-condition comparison: starting with a
vector in the periodic-boundary space $\mathcal K_{N,L}(\widehat A)$, produce
block-diagonal boundary matrices whose block components again belong to the
blockwise periodic-boundary spaces.

The present boundary isolates the remaining source comparison in the coordinates
obtained by opening the periodic boundary for boundary-crossing windows.  Suppose
\begin{align}
  \psi
  =
  \Gamma_N^{\widehat A}\!\left(\bigoplus_{j=1}^b X_j\right)
  \in \mathcal K_{N,L}(\widehat A).
\end{align}
For a cyclic interval beginning at \(i\) and crossing the boundary cut, put
\(a=i+L-N\).  The word \(\beta\) records the part of the interval before the
chosen cut, while \(\rho\) records the outside word of length \(N-L\).  These
are the blocked words obtained by opening the source boundary-index comparison.
The conditional theorem assumes that, for every block \(j\) and every such
outside word \(\rho\), there is a matrix \(E_{j,i,\rho}\) such that for every
word \(\beta\) of length \(a\),
\begin{align}
  \mu_j^N X_j A^j_\beta A^j_\rho
  =
  A^j_\beta E_{j,i,\rho}.
\end{align}
Under the usual block-injectivity and normalization hypotheses this assumption
implies
\begin{align}
  \Gamma_N^{A_j}(\mu_j^N X_j)\in\mathcal K_{N,L}(A_j)
  \qquad (1\le j\le b).
\end{align}
Thus the trace representation of \(\psi\) is not the obstruction. The
periodic component theorem proves the displayed conclusion from these
boundary-crossing coordinate identities. The remaining source-level work in this note is
to present the \(C^j,D^j,E^j\) comparison from
\cite[Theorem~12]{PerezGarcia2007MPSReps}
as an unconditional hypothesis-free equality theorem in the source range.\footnote{Formal declarations:
\leanid{MPSTensor.blockDiagonal_boundary_component_chainGroundSpace_of_complementary_word_identities},
\leanid{MPSTensor.blockDiagonal_boundary_component_chainGroundSpace_of_complementary_word_identities_of_injective},
\leanid{MPSTensor.blockDiagonal_boundary_component_chainGroundSpace_of_pgvwc_comparison_of_injective},
\leanid{MPSTensor.exists_blockDiagonal_boundary_chainGroundSpace_of_complementary_identities_bnt_c1}, and
\leanid{MPSTensor.chainGroundSpace_toTensorFromBlocks_eq_iSup_and_iSupIndep_of_complementary_identities}.}

The live blueprint records that larger source statement with explicit citations
and separates these opened-boundary coordinates from the paper's boundary indices,
without claiming that the source statement is proved.\footnote{Blueprint locations:
\path{def:parent_hamiltonian_ground_space_bnt} for the block-diagonal
periodic-boundary ground space,
\path{lem:isup_chain_ground_space_block_le_assembled} for the forward
inclusion, \path{thm:block_diagonal_ground_space_intersection} for the
block-diagonal intersection identity from
\cite[Theorem~12]{PerezGarcia2007MPSReps},
\path{thm:parent_ground_space_le_bnt_span} for the parent-ground-space
containment, and \path{thm:gs_eq_bnt_span} for the final span statement, all in
\path{blueprint/src/chapter/ch13_parent_hamiltonian_bnt_consequences.tex}.} With this notation,
\begin{align}
  \sum_{j=1}^b\mathcal K_{N,L}(A_j)
  =
  \left\{
    \sum_{j=1}^b \psi_j:
    \psi_j\in\mathcal K_{N,L}(A_j)
  \right\}.
\end{align}
The source-level boundary-condition statement not yet stated without assuming
the boundary-crossing coordinate identity is
\begin{align}
  \left[
    b\ge2
    \ \wedge\
    L_0\ge1
    \ \wedge\
    (\forall j,\ \mu_j\ne 0)
    \ \wedge\
    L_0<L
    \ \wedge\
    L\ge 3(b-1)(L_0+1)+1
    \ \wedge\
    N\ge L+L_0
  \right]
  \Longrightarrow
  \notag\\
  \mathcal K_{N,L}(\widehat A)
  =
  \sum_{j=1}^b \mathcal K_{N,L}(A_j).
\end{align}

The following implication is not available.  The trace-form space
$\mathcal G_N^A$ is not contained in the periodic-boundary ground space
$\mathcal K_{N,L}(A)$ for an arbitrary boundary matrix $X$.  If the
cyclic window crosses the chosen virtual boundary, then a vector
$\Gamma_N^A(X)$ has coefficients of the form
\begin{align}
  \tr\!\left(A^{\omega_{\mathrm{tail}}}
    X A^{\omega_{\mathrm{head}}} A^\rho\right),
\end{align}
where $\omega_{\mathrm{tail}}\omega_{\mathrm{head}}$ is the word seen in the
cyclic window and $\rho$ is the outside word.  This is not, for general $X$,
of the form
\begin{align}
  \tr\!\left(A^{\omega_{\mathrm{tail}}}
    A^{\omega_{\mathrm{head}}}Y_\rho\right)
\end{align}
with one matrix $Y_\rho$ independent of the window word.  The missing step is
therefore a boundary-condition comparison, not the inclusion
$\mathcal G_N^A\subseteq\mathcal K_{N,L}(A)$.
The source theorem uses the multi-block hypothesis, the positive common
injectivity length, the nonzero-weight hypothesis, and the length range of
\cite[Theorem~12]{PerezGarcia2007MPSReps}:
\begin{align}
  b\ge2,
  \qquad
  L_0\ge1,
  \qquad
  \forall j,\ \mu_j\ne0,
  \qquad
  L_0<L,
  \qquad
  L\ge 3(b-1)(L_0+1)+1,
  \qquad
  N\ge L+L_0.
\end{align}
For $b\ge2$ and $L_0\ge1$, the source length bound implies
$L\ge 3L_0+4>L_0$; the inequality $L_0<L$ is displayed separately because the
one-block uniqueness input is used in that range.
The one-block case is the injective parent-Hamiltonian theorem rather than the
degenerate block-injective boundary theorem.
The finite-Condition-C1 propagation theorem currently proved in Lean is stated
in the more conservative range
\begin{align}
  L\ge (L_0+1)+3(b-1)(L_0+1)+1.
\end{align}
The periodic-boundary comparison reductions now separate into three forms: a
decomposition into blockwise vectors with periodic boundary conditions, a
block-diagonal boundary representation whose components already belong to the
corresponding blockwise periodic-boundary spaces, and the boundary-crossing
coordinate formulation displayed above.\footnote{Lean declarations:
\leanid{MPSTensor.chainGroundSpace_toTensorFromBlocks_le_iSup_of_boundary_decomposition},
\leanid{MPSTensor.chainGroundSpace_toTensorFromBlocks_le_iSup_of_blockDiagonal_boundary_groundSpaceMap},
\leanid{MPSTensor.chainGroundSpace_toTensorFromBlocks_eq_iSup_and_iSupIndep_of_bnt_c1_boundary_decomposition},
\leanid{MPSTensor.chainGroundSpace_toTensorFromBlocks_eq_iSup_and_iSupIndep_of_bnt_c1_blockBoundary},
and
\leanid{MPSTensor.chainGroundSpace_toTensorFromBlocks_eq_iSup_and_iSupIndep_of_complementary_identities}.}
These reductions do not yet state the source theorem without assuming the
boundary-crossing comparison.
For
\begin{align}
  X=\bigoplus_j X_j,
\end{align}
the boundary-matrix formula is
\begin{align}
  \Gamma_N^{\widehat A}(X)(i_1,\ldots,i_N)
  &=
  \sum_j \mu_j^N
  \tr(A^j_{i_1}\cdots A^j_{i_N}X_j) \notag\\
  &=
  \sum_j \mu_j^N\Gamma_N^{A_j}(X_j)(i_1,\ldots,i_N).
\end{align}
For the next display, abbreviate
\begin{align}
  \Gamma_j(X)=\Gamma_N^{A_j}(X),
  \qquad
  V_j=V^{(N)}(A_j).
\end{align}
Thus the desired proof with periodic boundary conditions is the composition of
the two implications
\begin{align}
  \psi\in\mathcal K_{N,L}(\widehat A)
  &\Longrightarrow
  \exists (X_j)_{j=1}^b,\quad
  \psi=\sum_{j=1}^b\Gamma_j(X_j),
  \notag\\
  &\hphantom{\Longrightarrow}
  \forall\,1\le j\le b,\quad
  \Gamma_j(X_j)\in\mathcal K_{N,L}(A_j)
  ,
\end{align}
and, for every $X_1,\ldots,X_b$,
\begin{align}
  \forall\,1\le j\le b,\quad
  \Gamma_j(X_j)\in\mathcal K_{N,L}(A_j)
  &\Longrightarrow
  \exists (c_j)_{j=1}^b,
  \notag\\
  &\hphantom{\Longrightarrow}
  \forall\,1\le j\le b,\quad
  \Gamma_j(X_j)=c_j V_j
  .
\end{align}
The composition gives
\begin{align}
  \psi
  &=
  \sum_{j=1}^b\Gamma_j(X_j)
  =
  \sum_{j=1}^b c_j V_j,
  \notag\\
  \psi
  &\in
  \operatorname{span}\{V_j:j=1,\ldots,b\}.
\end{align}

\section{Conclusion}

This is not a source-error claim. The formal development has proved the
surrounding block-diagonal reductions and a conditional propagation step. In the
finite-Condition-C1 Lean range
\begin{align}
  L\ge (L_0+1)+3(b-1)(L_0+1)+1,
\end{align}
assume the boundary-crossing coordinate identities displayed above. Then, for
every block $j$,
\begin{align}
  \psi\in\mathcal K_{N,L}(\widehat A)
  \quad\Longrightarrow\quad
  \Gamma_N^{A_j}(\mu_j^NX_j)\in\mathcal K_{N,L}(A_j).
\end{align}
This is the conditional reduction recorded by pull request~\#2724.  There are
now two proved conditional routes to the boundary-condition comparison. The
boundary-trace route records the following implication. Assume the common
word-span property for the middle-word length \(m\). If, for every
boundary-crossing interval, word
$\beta$ before the cut, outside word $\rho$, and middle word $w$ of length \(m\),
the two boundary trace expressions from \cite[Theorem~12]{PerezGarcia2007MPSReps}
satisfy
\begin{align}
  \sum_j\tr\!\left(A^j_\beta C^j_\rho A^j_w\right)
  =
  \sum_j\tr\!\left((\mu_j^NX_jA^j_\beta)A^j_\rho A^j_w\right),
\end{align}
then
\begin{align}
  \Gamma_N^{A_j}(\mu_j^NX_j)\in\mathcal K_{N,L}(A_j)
\end{align}
for every block $j$ in the block-injective range.\footnote{Lean declaration:
\leanid{MPSTensor.blockDiagonal_boundary_component_chainGroundSpace_of_trace_decomposition_of_injective}.}
There is also an explicit-boundary version: once
\(\psi=\Gamma_N^{\widehat A}(\bigoplus_jX_j)\) is fixed, the same equality of
boundary trace expressions gives
\begin{align}
  \Gamma_N^{A_j}(\mu_j^NX_j)\in\mathcal K_{N,L}(A_j)
\end{align}
for every block \(j\), without using any span assumption at the short
boundary-crossing tail length \(N-i\).\footnote{Lean declaration:
\leanid{MPSTensor.exists_blockDiagonal_boundary_chainGroundSpace_of_trace_decomposition_of_boundary}.}
At the fixed wrapped-tail length supplied by a single boundary-crossing cyclic
window, the boundary trace expressions are now derived from the local
constraint once the block-diagonal boundary representation is fixed: for
\(N<i+L\), outside word \(\rho\) of length \(N-L\), and wrapped tail word
\(\alpha\) of length \(N-i\), there are matrices \(C^j_{i,\rho}\) such that
\begin{align}
  \sum_j\tr\!\left(A^j_\beta C^j_{i,\rho}A^j_\alpha\right)
  =
  \sum_j\tr\!\left((\mu_j^NX_jA^j_\beta)A^j_\rho A^j_\alpha\right).
\end{align}
This fixed-tail statement is recorded as
\leanid{MPSTensor.blockDiagonal_boundary_crossing_trace_decomposition_of_boundary}.
It should not be read as the common-middle-word trace hypothesis above: the
fixed cyclic window gives the wrapped tail length \(N-i\), whereas the
separating trace-decomposition implication is stated at a common product-span
length \(m\).

In the normalized BNT finite range, this common middle-word span is no longer an
external assumption. If
\begin{align}
  m=(L_0+1)+(b-1)\bigl((L_0+1)+((L_0+1)+(L_0+1))\bigr),
  \qquad
  L\ge m+1,
\end{align}
then the normalized block-separation theorem gives
\begin{align}
  \operatorname{span}\{(A^1_w,\ldots,A^b_w):|w|=m\}
  =
  \prod_j M_{D_j}(\mathbb C).
\end{align}
Consequently the boundary-trace implication in the finite range only keeps
the displayed trace equality and the block-diagonal boundary representation
as external inputs.\footnote{Lean declarations:
\leanid{MPSTensor.exists_blockDiagonal_boundary_chainGroundSpace_of_trace_decomposition_bnt_c1_span}
and
\leanid{MPSTensor.chainGroundSpace_toTensorFromBlocks_eq_iSup_and_iSupIndep_of_trace_decomposition_bnt_c1_span}.}

The \(C^j,D^j\) opened-boundary route is closer to the proof of
\cite[Theorem~12]{PerezGarcia2007MPSReps}.
The words \(\beta\) and \(\rho\) are local word coordinates introduced after
opening and blocking the cyclic boundary comparison; they reindex the source
end-site comparison \(i_1,i_{m+1}\) by blocked words, and are not additional
hypotheses or terminology from the paper.
For a fixed boundary-crossing interval starting at site \(i\), the local
block-sum constraint and the word-tuple span of the wrapped tail word of
length \(N-i\) now give matrices
\(C^j_{i,\rho}\) satisfying
\begin{align}
  A^j_\beta C^j_{i,\rho}
  =
  \bigl((\mu_j^NX_j)A^j_\beta\bigr)A^j_\rho .
\end{align}
This is formalized as
\leanid{MPSTensor.blockDiagonal_boundary_crossing_pgvwc_comparison_of_sum_mem_iSup}
and, once a block-diagonal boundary representation of \(\psi\) is fixed, into
\leanid{MPSTensor.exists_blockDiagonal_boundary_chainGroundSpace_of_crossing_pgvwc_comparison_of_boundary}.
The current BNT specialization then obtains that boundary representation through
\leanid{MPSTensor.exists_blockDiagonal_boundary_chainGroundSpace_of_crossing_pgvwc_comparison_bnt_c1}.
Thus the crossing comparison is not an independent trace-comparison assumption:
the present proof boundary is the crossing-tail word-span hypothesis and the
still-used block-diagonal boundary representation of a vector
\(\psi\in\mathcal K_{N,L}(\widehat A)\).
This crossing-tail span hypothesis should not be confused with the
large-length BNT product-span bound: if the boundary-crossing interval begins
at \(i=N-1\), the required tail length is \(N-i=1\), so it is not supplied by
the lower bound on \(L\) alone.

After \href{https://github.com/LionSR/TNLean/issues/2651}{issue \#2651} was
closed, the remaining source-comparison target is tracked by
\href{https://github.com/LionSR/TNLean/issues/2971}{issue \#2971}: remove the
remaining external boundary-representation input, and replace the
span-dependent crossing-tail route by the \(C^j,D^j,E^j\) comparison from
\cite[Theorem~12]{PerezGarcia2007MPSReps} in the source range
\begin{align}
  b\ge2,
  \qquad
  L_0\ge1,
  \qquad
  L\ge 3(b-1)(L_0+1)+1,
  \qquad
  N\ge L+L_0.
\end{align}
After this boundary-condition comparison is proved, the conditional equality
formalized in the current finite Condition C1 range gives
\begin{align}
  \mathcal K_{N,L}(\widehat A)
  =
  \sum_{j=1}^b\mathcal K_{N,L}(A_j).
\end{align}
The final degenerate
parent-Hamiltonian ground-space equality follows by applying the one-block
uniqueness theorem to each block.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
