\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{CPGSV17 Blocked-Basis Chi Families and Uniform BNT Labels}
\author{}
\date{2026-05-24}

\begin{document}
\maketitle

\section{Source statement}

Cirac--P\'erez-Garc\'ia--Schuch--Verstraete 2017 state in
\cite[Theorem~IV.13(ii)]{Cirac2017MPDO_AnnPhys} that there is a set of
diagonal matrices
\[
    \chi_{\alpha,\beta,\gamma}
\]
with positive entries such that, for every length $L$, the operators
$O_L(M_\alpha)$ linearly span an algebra whose structure coefficients satisfy
\[
    c^{(L)}_{\alpha,\beta,\gamma}
      = \tr(\chi_{\alpha,\beta,\gamma}^{L}).
\]
The important point for the present note is that the labels
$\alpha,\beta,\gamma$ are BNT labels, not basis labels depending on $L$.
Thus the same diagonal matrix $\chi_{\alpha,\beta,\gamma}$ is used for all
positive lengths; only the exponent changes.

Here $O_L(M_\alpha)$ denotes the length-$L$ operator associated to the
normal tensor labelled by $\alpha$ in the basis of normal tensors. The phrase
``BNT labels'' refers to these labels $\alpha,\beta,\gamma$. The support
algebra $\mathcal A_n$ below is the formal support algebra chosen for the
blocked length $n$ in the Lean development.

\section{Formalized blocked-basis analogue}

The current blocked coefficient construction in
\path{TNLean/MPS/MPDO/AlgebraStructure.lean} chooses bases of the support
algebras $\mathcal A_n$. Its multiplication coefficients have indices
\[
    i,j \in \mathrm{Basis}(\mathcal A_n),
    \qquad
    k \in \mathrm{Basis}(\mathcal A_{2n}),
\]
because multiplication is represented as
\[
    \mathcal A_n \times \mathcal A_n \longrightarrow \mathcal A_{2n}.
\]
The declarations
\leanid{MPOTensor.AlgebraStructureData.BlockedStructureChiFamily} and
\leanid{MPOTensor.AlgebraStructureData.HasBlockedStructureChiTracePowerForm}
therefore express a positive-length, length-dependent blocked-basis analogue:
\[
    c^{(n)}_{i,j,k}
      = \sum_r \chi_{n,i,j,k,r}^{\,n}
      = \tr(\chi_{n,i,j,k}^{\,n}).
\]
Here $n$ is the source blocking length of the two factors in
$\mathcal A_n$, while the coefficient $k$ is taken in the basis of
$\mathcal A_{2n}$. The predicate is intentionally restricted to
$0 < n$, avoiding the degenerate length-zero constraint
$c^{(0)}_{i,j,k} = \dim(\chi_{0,i,j,k})$, which is not part of the intended
positive-chain statement.

\section{Gap}

Verdict: scope restriction.

This blocked-basis analogue is not the uniform BNT-label statement of
Theorem~IV.13(ii). It allows the diagonal matrix to depend on $n$, because
the available indices themselves depend on the blocked bases
$\mathcal A_n$ and $\mathcal A_{2n}$. The source theorem instead provides
one matrix for each triple of BNT labels and uses that same matrix for every
length $L$.

There is a second scope restriction in the operation being compared. In the
paper, the coefficients $c^{(L)}_{\alpha,\beta,\gamma}$ belong to the
same-length operator algebra:
\[
    O_L(M_\alpha) O_L(M_\beta)
      = \sum_\gamma c^{(L)}_{\alpha,\beta,\gamma} O_L(M_\gamma).
\]
The current Lean coefficient family comes from the blocked support-algebra
product
\[
    \mathcal A_n \times \mathcal A_n \longrightarrow \mathcal A_{2n},
\]
so the two input factors have source length $n$, while the result is expanded
in a basis at length $2n$. The exponent $n$ in
\[
    c^{(n)}_{i,j,k}=\tr(\chi_{n,i,j,k}^{\,n})
\]
is therefore the source length of the two factors, not the codomain length.
This convention is only a blocked-basis analogue; it is not a replacement for
the same-length multiplication statement in Theorem~IV.13(ii).

This note isolates the chi-uniformity and same-length coefficient comparison
part of Theorem~IV.13(ii).  The idempotent condition
\[
    m_\gamma =
      \sum_{\alpha,\beta} c^{(1)}_{\alpha,\beta,\gamma} m_\alpha m_\beta
\]
from paper lines 981--985 is a separate condition in the same theorem; it is
not represented by the blocked-basis trace-power predicate.

Consequently,
\leanid{MPOTensor.AlgebraStructureData.HasBlockedStructureChiTracePowerForm}
should not be cited as a formalization of Theorem~IV.13(ii). It is a local
blocked-basis coefficient statement, useful only after the
uniform BNT-label statement has been separated from the chosen blocked bases.

\section{Relation to the algebra-to-fusion converse}

Let $E=E_1$ be the one-site transfer map.  The present algebra predicate gives
the fixed-point-space identity
\[
    \operatorname{Fix}\bigl((E^n)^\dagger\bigr)
      = \operatorname{Fix}(E^\dagger),
    \qquad n>0.
\]
This rules out root-of-unity peripheral eigenvalues of $E^\dagger$, but it does
not imply the fusion-side idempotence equation $E^2=E$.

The converse direction $(\mathrm{ii})\Rightarrow(\mathrm{i})$ in the proof of
Theorem~IV.13 uses a stronger coefficient argument.  In Appendix~C.4,
lines~2046--2085, the authors compare the two BNT expansions coming from the
one-site and two-site decompositions.  The source identity
\texttt{eq:algebra-appendix} is stated at lines~1933--1936 and used in this
converse at lines~2048--2050:
\[
    \sum_\gamma \tr(\nu_\gamma^L) O_L(A_\gamma)
      =
    \sum_{\alpha,\beta,\gamma}
      \tr\bigl((\mu_\alpha\otimes\mu_\beta
        \otimes\chi_{\alpha,\beta,\gamma})^L\bigr) O_L(M_\gamma).
\]
Since both $A_\gamma$ and $M_\gamma$ are BNT, the comparison first gives,
up to relabelling (line~2053),
\[
    A_\gamma
      = e^{i\theta_\gamma} Z_\gamma M_\gamma Z_\gamma^{-1}.
\]
Proposition~IV.12 then removes the phase and identifies $Z_\gamma$ with a
unitary $U_\gamma$, as in line~2057, so
\[
    A_\gamma = U_\gamma M_\gamma U_\gamma^{-1}
\]
with $U_\gamma$ unitary.  Since, for all sufficiently large $L$, the operators
$O_L(A_\gamma)=O_L(M_\gamma)$ are linearly independent, and all coefficients are
positive, the simple-matrix lemma (lines~1155--1163, used at line~2058) gives
the multiset identity
\[
    \{\nu_{\gamma,k}\}_k
      =
    \{\mu_{\alpha,k_1}\mu_{\beta,k_2}
      \chi_{\alpha,\beta,\gamma,k_3}\}_{\alpha,\beta,k_1,k_2,k_3}.
\]
Summing this identity gives
\[
    \tr(\nu_\gamma)
      = \sum_{\alpha,\beta}m_\alpha m_\beta
          c^{(1)}_{\alpha,\beta,\gamma}
      = m_\gamma .
\]
Only after this trace equality is obtained do the normalizing maps in
\cite[lines~2065--2085]{Cirac2017MPDO_AnnPhys} have the form
\[
    \widetilde R_2\Bigl(\bigoplus_\gamma X_\gamma\Bigr)
      = \bigoplus_\gamma \frac{\nu_\gamma}{m_\gamma}\otimes X_\gamma,
    \qquad
    \widetilde R_1\Bigl(\bigoplus_\gamma X_\gamma\Bigr)
      = \bigoplus_\gamma \frac{\mu_\gamma}{m_\gamma}\otimes X_\gamma .
\]
These are the maps used in
$T=\widetilde R_2\widetilde T R_1$ and
$S=\widetilde R_1\widetilde S R_2$.  The first denominator is normalized by
$\tr(\nu_\gamma)=m_\gamma$, proved above; the second by the defining equality
$m_\gamma=\tr(\mu_\gamma)$ for the original block coefficient.

The present algebra predicate
\leanid{MPOTensor.IsRFP_MPDO_via_algebra} does not include these BNT-label
coefficients, the positive $\chi_{\alpha,\beta,\gamma}$, or the length-one
idempotent trace identity.  It proves the fixed-point-space equation above, not
the displayed BNT coefficient comparison.  Thus the source-faithful route to the
fusion formulation must construct the BNT-label theorem witness from the MPDO
tensor and then use the one-site retract criterion
\leanid{MPOTensor.fusionIsometryData_one_iff_isRFP}.

\section{BNT-label coefficient objects and remaining elimination plan}

The same-length coefficient family $c^{(L)}_{\alpha,\beta,\gamma}$, indexed
by fixed BNT labels, is represented separately from the blocked-basis
coefficients.\footnote{\leanid{MPOTensor.BNTLabelCoefficientFamily}.}  When the
Appendix C.4 construction has already produced a diagonal
$\chi_{\alpha,\beta,\gamma}$-family, there is also a canonical coefficient
family
\[
    c^{(L)}_{\alpha,\beta,\gamma}
      = \sum_r \chi_{\alpha,\beta,\gamma,r}^{\,L}.
\]
This is recorded separately so that the $\chi$-family can serve as the single
source of those coefficients in the source-side special case.\footnote{%
\leanid{MPOTensor.BNTLabelCoefficientFamily.ofChi},
\leanid{MPOTensor.BNTLabelCoefficientFamily.ofChi_coeff},
\leanid{MPOTensor.BNTLabelCoefficientFamily.%
ofChi_coeff_eq_trace_matrix_pow},
\leanid{MPOTensor.BNTLabelCoefficientFamily.%
ofChi_hasPositiveLengthChiTracePowerForm}.}  A
same-length BNT product predicate records the algebra identity
\[
    O_L(M_\alpha)O_L(M_\beta)
      = \sum_\gamma c^{(L)}_{\alpha,\beta,\gamma}O_L(M_\gamma)
\]
for an abstract family of length-$L$ BNT operators.\footnote{%
\leanid{MPOTensor.BNTLabelOperatorFamily.HasSameLengthProductForm}.}  The
idempotent condition
\[
    m_\gamma =
      \sum_{\alpha,\beta} c^{(1)}_{\alpha,\beta,\gamma}m_\alpha m_\beta
\]
is also represented as a predicate on the trace scalars
$m_\alpha=\tr(\mu_\alpha)$.\footnote{%
\leanid{MPOTensor.BNTLabelTraceScalarFamily.HasIdempotentCoefficientForm}.}  A
positive BNT-label $\chi$-witness consists of a length-independent diagonal
family $\chi_{\alpha,\beta,\gamma}$, positivity of every diagonal entry, and
the positive-length trace-power identity.\footnote{%
\leanid{MPOTensor.PositiveBNTLabelChiTracePowerForm}.}
If the coefficients are chosen canonically from the same $\chi$-family, then
positivity of the diagonal entries alone gives such a witness.\footnote{%
\leanid{MPOTensor.PositiveBNTLabelChiTracePowerForm.ofChi}.}
A blocked-basis comparison predicate records that the chosen blocked
multiplication coefficients are the pullback of the source-length BNT-label
coefficients along a source label map from $\mathcal A_n$ and a target label
map from $\mathcal A_{2n}$.  The label-assignment datum is separated from
the coefficient equality, so the Appendix C.3 construction target can be stated
without also assuming the comparison formula.\footnote{%
\leanid{MPOTensor.BNTBlockedBasisLabelAssignment},
\leanid{MPOTensor.BNTBlockedBasisLabelAssignment.blockedLabel},
\leanid{MPOTensor.BNTBlockedBasisCoefficientComparison}.}  This predicate
inherits the second scope restriction above: its left hand side is expanded in a
basis of $\mathcal A_{2n}$, while the source theorem's BNT product law is
same-length. It therefore records the formal blocked-basis comparison
predicate; constructing the source label maps from an MPDO tensor is still part
of the Appendix C.3--C.4 witness construction. The predicate is not itself the
same-length product statement of Theorem~IV.13(ii).
These declarations specify the BNT-label coefficient objects and the comparison
statement, but they do not yet construct the operators, trace scalars,
coefficients, comparison maps, or $\chi$-matrices from an MPDO tensor.
Once both the same-length BNT product form and a positive BNT-label
$\chi$-witness are available, the same-length product expansion with
coefficients written as $\tr(\chi_{\alpha,\beta,\gamma}^L)$
is formalized as a direct corollary.\footnote{%
\leanid{MPOTensor.BNTLabelOperatorFamily.HasSameLengthProductForm.%
eq_sum_chi_trace_pow},
\leanid{MPOTensor.BNTLabelOperatorFamily.HasSameLengthProductForm.%
eq_sum_ofChi_trace_pow}.}
Once both the comparison predicate and a positive BNT-label $\chi$-witness are
available, the coefficient-level blocked trace-power formula is formalized as a
direct corollary.\footnote{%
\leanid{MPOTensor.BNTBlockedBasisCoefficientComparison.blocked_coeff_eq_trace_pow},
\leanid{MPOTensor.BNTBlockedBasisCoefficientComparison.%
blocked_coeff_eq_ofChi_trace_pow},
\leanid{MPOTensor.BNTBlockedBasisCoefficientComparison.%
blocked_coeff_eq_pulledBlockedChi_trace_pow}.}
The same hypotheses also give a positive blocked-basis $\chi$-witness by
pulling back the uniform BNT-label $\chi$-family along the
comparison maps.\footnote{%
\leanid{MPOTensor.BNTBlockedBasisCoefficientComparison.%
pulledBlockedChiFamily_toDiagonal_of_pos},
\leanid{MPOTensor.BNTBlockedBasisCoefficientComparison.%
pulledBlockedChi_tracePowerCoeff_of_pos},
\leanid{MPOTensor.BNTBlockedBasisCoefficientComparison.%
pulledBlockedChi_dim_of_pos},
\leanid{MPOTensor.BNTBlockedBasisCoefficientComparison.%
pulledBlockedChi_trace_matrix_pow_of_pos},
\leanid{MPOTensor.BNTBlockedBasisCoefficientComparison.%
toPositiveBlockedStructureChiTracePowerForm}.}
Likewise, once the idempotent coefficient condition and the positive
BNT-label $\chi$-witness are available, the length-one idempotent identity is
formalized with the coefficients rewritten as traces of the corresponding
$\chi$-matrices.\footnote{%
\leanid{MPOTensor.BNTLabelTraceScalarFamily.HasIdempotentCoefficientForm.%
eq_sum_chi_trace},
\leanid{MPOTensor.BNTLabelTraceScalarFamily.HasIdempotentCoefficientForm.%
eq_sum_ofChi_trace}.}

The BNT-label objects above are now also gathered into one theorem-data record:
fixed coefficients, same-length operators and product law, trace scalars and
idempotent condition, positive BNT-label $\chi$-witness, and blocked-basis
comparison.\footnote{\leanid{MPOTensor.BNTLabelTheoremData}.}  This record is
not an additional mathematical assumption in the source theorem; it is the
single formal object collecting the source data supplied by the Appendix
C.3--C.4 construction from an MPDO tensor.  There is also a construction for the
source-side special case in which the coefficient family is the one canonically
determined by the same $\chi$-family; it takes the product law, idempotent
law, and blocked-basis comparison for those canonical coefficients.\footnote{%
\leanid{MPOTensor.BNTLabelTheoremData.ofChi}.}  From such
data, the source
predicates themselves and the corresponding equation-level consequences are
available: the same-length product law, the idempotent scalar law, the BNT-label
$\chi$-positivity and trace-power predicates, the coefficient trace-power
formula, positivity of the diagonal $\chi$-entries, the same-length chi-trace
product law, the chi-trace idempotent law, the blocked-basis coefficient
trace-power formula, the blocked-basis coefficient comparison, the positive
blocked $\chi$-witness, the pulled-back blocked trace-power predicate, and the
pulled-back blocked-basis $\chi$-family with positive entries and trace-power
formula are immediate consequences.\footnote{%
\leanid{MPOTensor.BNTLabelTheoremData.same_length_product_form},
\leanid{MPOTensor.BNTLabelTheoremData.idempotent_coefficient_form},
\leanid{MPOTensor.BNTLabelTheoremData.positive_chi_pos_entries},
\leanid{MPOTensor.BNTLabelTheoremData.positive_chi_trace_power},
\leanid{MPOTensor.BNTLabelTheoremData.labelAssignment},
\leanid{MPOTensor.BNTLabelTheoremData.sourceLabel},
\leanid{MPOTensor.BNTLabelTheoremData.targetLabel},
\leanid{MPOTensor.BNTLabelTheoremData.same_length_product_eq_sum},
\leanid{MPOTensor.BNTLabelTheoremData.idempotent_eq_sum},
\leanid{MPOTensor.BNTLabelTheoremData.coeff_eq_trace_pow},
\leanid{MPOTensor.BNTLabelTheoremData.coeff_eq_ofChi_coeff},
\leanid{MPOTensor.BNTLabelTheoremData.same_length_product_form_ofChi},
\leanid{MPOTensor.BNTLabelTheoremData.idempotent_coefficient_form_ofChi},
\leanid{MPOTensor.BNTLabelTheoremData.same_length_product_eq_sum_ofChi},
\leanid{MPOTensor.BNTLabelTheoremData.idempotent_eq_sum_ofChi},
\leanid{MPOTensor.BNTLabelTheoremData.chi_entry_pos},
\leanid{MPOTensor.BNTLabelTheoremData.blocked_coeff_eq},
\leanid{MPOTensor.BNTLabelTheoremData.blockedComparison_ofChi},
\leanid{MPOTensor.BNTLabelTheoremData.blocked_coeff_eq_ofChi},
\leanid{MPOTensor.BNTLabelTheoremData.same_length_product_eq_sum_chi_trace_pow},
\leanid{MPOTensor.BNTLabelTheoremData.idempotent_eq_sum_chi_trace},
\leanid{MPOTensor.BNTLabelTheoremData.blocked_coeff_eq_trace_pow},
\leanid{MPOTensor.BNTLabelTheoremData.toPositiveBlockedStructureChiTracePowerForm},
\leanid{MPOTensor.BNTLabelTheoremData.positiveBlockedChi},
\leanid{MPOTensor.BNTLabelTheoremData.positiveBlockedChi_toDiagonal_of_pos},
\leanid{MPOTensor.BNTLabelTheoremData.%
positiveBlockedChi_tracePowerCoeff_of_pos},
\leanid{MPOTensor.BNTLabelTheoremData.positiveBlockedChi_dim_of_pos},
\leanid{MPOTensor.BNTLabelTheoremData.%
positiveBlockedChi_trace_matrix_pow_of_pos},
\leanid{MPOTensor.BNTLabelTheoremData.positiveBlockedChi_tracePower},
\leanid{MPOTensor.BNTLabelTheoremData.positiveBlockedChi_posEntries},
\leanid{MPOTensor.BNTLabelTheoremData.positiveBlockedChi_entry_pos},
\leanid{MPOTensor.BNTLabelTheoremData.%
blocked_coeff_eq_positiveBlockedChi_trace_pow}.}

The same data can be stated in the existential form closest to the paper: one
object records the BNT-label type, the same-length operator spaces, their
algebraic structure, and the corresponding theorem data.\footnote{%
\leanid{MPOTensor.BNTLabelTheoremWitness},
\leanid{MPOTensor.HasBNTLabelTheoremWitness},
\leanid{MPOTensor.BNTLabelTheoremWitness.toTheoremData},
\leanid{MPOTensor.BNTLabelTheoremWitness.coeffs},
\leanid{MPOTensor.BNTLabelTheoremWitness.operators},
\leanid{MPOTensor.BNTLabelTheoremWitness.traceScalars},
\leanid{MPOTensor.BNTLabelTheoremWitness.positiveChi},
\leanid{MPOTensor.BNTLabelTheoremWitness.blockedComparison},
\leanid{MPOTensor.BNTLabelTheoremWitness.same_length_product_form},
\leanid{MPOTensor.BNTLabelTheoremWitness.idempotent_coefficient_form},
\leanid{MPOTensor.BNTLabelTheoremWitness.positive_chi_pos_entries},
\leanid{MPOTensor.BNTLabelTheoremWitness.positive_chi_trace_power},
\leanid{MPOTensor.BNTLabelTheoremWitness.labelAssignment},
\leanid{MPOTensor.BNTLabelTheoremWitness.sourceLabel},
\leanid{MPOTensor.BNTLabelTheoremWitness.targetLabel},
\leanid{MPOTensor.BNTLabelTheoremWitness.positiveBlockedChi}.}  The existential
nonempty statement can also be unpacked directly into a concrete witness
carrying the source predicates, the corresponding same-length and idempotent
equations written with $c^{(L)}_{\alpha,\beta,\gamma}$, and the same two
equations with coefficients written as $\chi$-traces; the blocked-basis
$\chi$-family is also identified as the pullback of the BNT-label
$\chi$-family along the comparison maps.\footnote{%
\leanid{MPOTensor.HasBNTLabelTheoremWitness.exists_source_predicates},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.%
exists_positive_length_coeff_eq_ofChi},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.exists_source_ofChi_predicates},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.exists_source_ofChi_equations},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.exists_blocked_chi_pullback},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.%
exists_blocked_coefficient_comparison},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.%
exists_blocked_coefficient_comparison_ofChi},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.%
exists_source_coefficient_equations},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.%
exists_source_chi_trace_equations}.}
The existential
statement has the analogous canonical construction from a positive
$\chi$-family, again leaving the product, idempotent, and blocked-comparison
statements as inputs rather than reconstructing them from an MPDO tensor.\footnote{%
\leanid{MPOTensor.BNTLabelTheoremWitness.ofChi},
\leanid{MPOTensor.BNTLabelTheoremWitness.hasBNTLabelTheoremWitness},
\leanid{MPOTensor.BNTLabelTheoremWitness.%
ofChi_hasBNTLabelTheoremWitness}.}  Such a witness
immediately gives the source predicates and their consequences: the coefficient
trace-power formula, the same-length product law, the idempotent scalar law,
positivity of the diagonal $\chi$-entries, the blocked-basis coefficient
comparison, the same-length chi-trace product law, the chi-trace idempotent law,
the blocked-basis coefficient trace-power formula, the positive blocked-basis
$\chi$-witness, and the pulled-back blocked trace-power predicate and
$\chi$-family for the chosen algebra-structure data.\footnote{%
\leanid{MPOTensor.BNTLabelTheoremWitness.same_length_product_form},
\leanid{MPOTensor.BNTLabelTheoremWitness.idempotent_coefficient_form},
\leanid{MPOTensor.BNTLabelTheoremWitness.positive_chi_pos_entries},
\leanid{MPOTensor.BNTLabelTheoremWitness.positive_chi_trace_power},
\leanid{MPOTensor.BNTLabelTheoremWitness.sourceLabel},
\leanid{MPOTensor.BNTLabelTheoremWitness.targetLabel},
\leanid{MPOTensor.BNTLabelTheoremWitness.coeff_eq_trace_pow},
\leanid{MPOTensor.BNTLabelTheoremWitness.coeff_eq_ofChi_coeff},
\leanid{MPOTensor.BNTLabelTheoremWitness.same_length_product_form_ofChi},
\leanid{MPOTensor.BNTLabelTheoremWitness.idempotent_coefficient_form_ofChi},
\leanid{MPOTensor.BNTLabelTheoremWitness.same_length_product_eq_sum_ofChi},
\leanid{MPOTensor.BNTLabelTheoremWitness.idempotent_eq_sum_ofChi},
\leanid{MPOTensor.BNTLabelTheoremWitness.same_length_product_eq_sum},
\leanid{MPOTensor.BNTLabelTheoremWitness.idempotent_eq_sum},
\leanid{MPOTensor.BNTLabelTheoremWitness.chi_entry_pos},
\leanid{MPOTensor.BNTLabelTheoremWitness.blocked_coeff_eq},
\leanid{MPOTensor.BNTLabelTheoremWitness.blockedComparison_ofChi},
\leanid{MPOTensor.BNTLabelTheoremWitness.blocked_coeff_eq_ofChi},
\leanid{MPOTensor.BNTLabelTheoremWitness.same_length_product_eq_sum_chi_trace_pow},
\leanid{MPOTensor.BNTLabelTheoremWitness.idempotent_eq_sum_chi_trace},
\leanid{MPOTensor.BNTLabelTheoremWitness.blocked_coeff_eq_trace_pow},
\leanid{MPOTensor.BNTLabelTheoremWitness.%
toPositiveBlockedStructureChiTracePowerForm},
\leanid{MPOTensor.BNTLabelTheoremWitness.%
positiveBlockedChi_toDiagonal_of_pos},
\leanid{MPOTensor.BNTLabelTheoremWitness.%
positiveBlockedChi_tracePowerCoeff_of_pos},
\leanid{MPOTensor.BNTLabelTheoremWitness.positiveBlockedChi_dim_of_pos},
\leanid{MPOTensor.BNTLabelTheoremWitness.%
positiveBlockedChi_trace_matrix_pow_of_pos},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.%
positive_blocked_chi_witness},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.%
exists_blocked_chi_trace_power_form},
\leanid{MPOTensor.HasBNTLabelTheoremWitness.%
exists_blocked_coeff_eq_trace_pow},
\leanid{MPOTensor.BNTLabelTheoremWitness.positiveBlockedChi_tracePower},
\leanid{MPOTensor.BNTLabelTheoremWitness.positiveBlockedChi_posEntries},
\leanid{MPOTensor.BNTLabelTheoremWitness.positiveBlockedChi_entry_pos},
\leanid{MPOTensor.BNTLabelTheoremWitness.%
blocked_coeff_eq_positiveBlockedChi_trace_pow}.}  The remaining step is still the
construction of this existential witness from the MPDO tensor.

To eliminate the remaining gap, the formalization should introduce:
\begin{enumerate}
    \item construction of the comparison maps relating the BNT-label
        coefficients to the chosen blocked bases of $\mathcal A_n$;
    \item construction of the same-length BNT operator family and product law
        from the MPDO tensor, together with the comparison to the blocked
        support-algebra product
        $\mathcal A_n \times \mathcal A_n \to \mathcal A_{2n}$;
    \item construction of the positive BNT-label
        $\chi_{\alpha,\beta,\gamma}$-witness from the paper's Appendix C.3
        and C.4 hypotheses;
    \item construction of the idempotent condition
        $m_\gamma =
          \sum_{\alpha,\beta} c^{(1)}_{\alpha,\beta,\gamma}
          m_\alpha m_\beta$ from the MPDO tensor;
    \item construction of the single
        \leanid{MPOTensor.BNTLabelTheoremWitness} from these source
        objects, and hence of the underlying
        \leanid{MPOTensor.BNTLabelTheoremData} record.  The blocked-basis
        trace-power statement, with its
        source-length exponent convention, is then a derived corollary of the
        uniform BNT-label theorem, rather than a separate replacement for the
        theorem itself.
\end{enumerate}

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
