\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}

\input{command}

\title{Finite-Length Block Separation in CPGSV17 Proposition IV.3}
\author{}
\date{2026-04-30}

\begin{document}
\maketitle

\section{The assertion}

This note concerns the finite-length block-separation step in Proposition IV.3
of Cirac, P\'erez-Garc\'ia, Schuch, and Verstraete
\cite{Cirac2016MPDO_arXiv,Cirac2017MPDO_AnnPhys}.\footnote{For traceability,
this note corresponds to \href{https://github.com/LionSR/TNLean/issues/1043}{issue \#1043}.
Related formal work is recorded in
\href{https://github.com/LionSR/TNLean/issues/587}{issue \#587} and
\href{https://github.com/LionSR/TNLean/issues/822}{issue \#822}; earlier proof attempts are
\href{https://github.com/LionSR/TNLean/pull/682}{PR \#682} and
\href{https://github.com/LionSR/TNLean/pull/813}{PR \#813}.}

Let $A$ be a tensor in canonical form, and let
$A_1,\ldots,A_g$ be a basis of normal tensors for $A$. The paper characterizes
such a basis by minimality: no two chosen normal tensors may be related by an
invertible similarity together with a phase. It then defines block-injective
canonical form as the statement that, after passing to these blocks, simultaneous
word evaluations span the whole product algebra
\[
  \bigoplus_{j=1}^g M_{D_j}(\mathbb C).
\]
Equivalently, for every block tuple $X=(X_j)_j$ there should be a linear
combination of common physical words whose evaluation on block $j$ gives $X_j$
for every $j$.

The source first invokes the quantum Wielandt theorem: after at most $D^4$
blockings, each normal tensor becomes injective. It then cites David2006 for the
cross-block step: if $L$ is a blocking length after which every normal block is
injective, then after $3(L+1)(D-1)$ blockings the tensor is block
injective. The proposition records the resulting bound as follows:
\[
  \text{after blocking at most } 3D^5 \text{ spins, every tensor in canonical
  form is in block-injective canonical form.}
  \tag{CPGSV17, Prop. IV.3}
\]
Here $D$ is the bond-dimension parameter used in the paper.
The local source is
\path{Papers/1606.00608/MPDO-22-12-17-2.tex}, lines 271--345. The cited
reference is the paper's item \texttt{David2006}, namely D. P\'erez-Garc\'ia,
F. Verstraete, M. M. Wolf, and J. I. Cirac, ``Matrix Product State
Representations'', \emph{Quantum Information and Computation} \textbf{7}
(2007), 401--430.

\section{The formal replacement}

The formal treatment does not yet derive this proposition from the
canonical-form and basis-of-normal-tensors hypotheses. Instead, it makes the
finite-length conclusion explicit. The horizontal canonical-form data assume
the following trace-separation property: there is a length $N$ such
that, if matrices $\Delta_j\in M_{D_j}(\mathbb C)$ satisfy
\[
  \sum_{j=1}^g \operatorname{tr}(\Delta_j A_j^w)=0
  \qquad\text{for every word } w \text{ of length } N,
\]
then $\Delta_j=0$ for every $j$. By nondegeneracy of the trace pairing on the
product algebra, this is the dual form of the finite-length spanning property
above.

Several exact sufficient hypotheses have been proved. The first is the
finite-length product-algebra span itself: the tuples
\[
  w \longmapsto (A_1^w,\ldots,A_g^w)
\]
span $\bigoplus_j M_{D_j}(\mathbb C)$ for one common length. The second is an
entrywise linear-independence criterion: the scalar functions
$w\mapsto (A_j^w)_{\alpha\beta}$, indexed by the block and by a matrix entry in
that block, are linearly independent at one common length. This implies the
product-algebra span, hence the trace-separation property. The third is a
selector-word formulation of Proposition IV.3: after a common blocking length
each block is injective, and after a further finite length there are word
polynomials which are the identity on one chosen block and vanish on all the
others. Concatenating an injective prefix with such a selector suffix gives the
full product-algebra span.

In the formal files these three criteria are named
\texttt{WordTupleSpanTop}, \texttt{wordEntryFamily}, and
\texttt{PropBlockInjective}. The field corresponding to the trace-separation
conclusion is \texttt{HorizontalCFData.biCF}, and the constructors from the three
finite-length hypotheses to horizontal canonical-form data are in
\path{TNLean/MPS/MPDO/BiCFDerivation.lean}.  In particular, the relevant
formal implications include \texttt{hasBiCF\_of\_wordTupleSpanTop},
\texttt{hasBiCF\_of\_wordEntryFamily\_linearIndependent},
\texttt{hasBiCF\_of\_propBlockInjective}, and the corresponding
\texttt{HorizontalCFData} constructors such as
\texttt{horizontalCFData\_of\_propBlockInjective}.  The blueprint records the
same state in \path{blueprint/src/chapter/ch02b_mpdo.tex}, lines 708--787.
Thus Proposition IV.3 has not been asserted without proof. The formal statements
prove that several finite-length witnesses are sufficient, while the derivation
of such a witness from the full canonical-form and BNT hypotheses remains open.

\section{The counterexample boundary}

A tempting shortcut is false. One cannot recover the finite-length biCF witness
from blockwise injectivity, left-canonicality, and nonzero weights alone. Take
two one-dimensional scalar blocks over a one-letter alphabet, with both block
tensors equal to $1$ and with weights $1$ and $2$. Each block is injective and
left-canonical, and the weights are nonzero and distinct. However every word
evaluation is the same on the two blocks. The tuple span is the diagonal line in
$\mathbb C\oplus\mathbb C$, not the whole product algebra. Equivalently,
choosing $\Delta_1=1$ and $\Delta_2=-1$ makes the trace pairing vanish on every
word while the tuple $(\Delta_1,\Delta_2)$ is nonzero.

This example is formalized in
\path{TNLean/MPS/MPDO/BiCFDerivation.lean}, lines 1363--1463, and is recorded in
\path{docs/counterexamples.md}, lines 29--35. It shows that the extra
finite-length witness included in the formal hypotheses is genuinely
stronger than the other horizontal canonical-form fields. It does not refute
CPGSV17 Proposition IV.3. The duplicate scalar blocks are related by similarity
and phase, and therefore do not form a minimal basis of normal tensors in the
sense of the source characterization.

\section{Conclusion}

The present status is provisional and should not be read as a source erratum.
The formal statements replace the paper's finite-length block-separation theorem
by explicit finite-length span, linear-independence, or selector hypotheses. A
small counterexample shows that those hypotheses cannot be derived from the
weaker fields alone. The full source theorem remains plausible under the actual
canonical-form and minimal BNT hypotheses.

What remains to be formalized is the David2006/CPGSV17 finite-length
block-separation argument: from minimality of the BNT blocks, together with
single-block injectivity after Wielandt blocking, one must produce a common
finite selector or entrywise linear-independence witness, with the paper's bound
$3(L+1)(D-1)$ and hence the stated $3D^5$ corollary. Until that argument is
available, formal uses of biCF should keep the finite-length witness as an
explicit hypothesis or field.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}