\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Normal Parent-Hamiltonian Range Reduction in Cirac--Perez-Garcia--Schuch--Verstraete 2021}
\author{}
\date{2026-04-30}

\begin{document}
\maketitle

\section{The assertion}

The range-reduction step for normal matrix product states appears in
Section~IV.C of Cirac--P\'erez-Garc\'ia--Schuch--Verstraete
\cite{Cirac2021Matrix}.\footnote{For traceability, this note corresponds to
\href{https://github.com/LionSR/TNLean/issues/1042}{issue \#1042}. The principal
range-reduction comparison is recorded in
\href{https://github.com/LionSR/TNLean/issues/588}{issue \#588}; the
periodic-boundary comparison is recorded in
\href{https://github.com/LionSR/TNLean/issues/2368}{issue \#2368}; its current
coordinate form is recorded in
\href{https://github.com/LionSR/TNLean/issues/2405}{issue \#2405};
and the parent-Hamiltonian consequences are recorded in
\href{https://github.com/LionSR/TNLean/issues/460}{issue \#460}.}

Let $A$ be a normal MPS tensor, and let $L_0>0$ be its injectivity length: after
blocking $L_0$ consecutive sites, the resulting tensor is injective. The source
first proves an intersection step for two overlapping local Hamiltonian terms of
range $L_0+1$. These terms act on sites $1,\ldots,L_0+1$ and
$2,\ldots,L_0+2$. The source denotes the three regions by $A$, $B$, and $C$:
the first site, the middle region consisting of sites $2,\ldots,L_0+1$, and the
last site. The tensor associated to the middle region $B$ is injective. For a
state in the intersection
\[
  \mathcal G_{1,\ldots,L_0+1}\cap \mathcal G_{2,\ldots,L_0+2},
\]
the argument inverts the tensor associated with the first-site region $A$,
restores the middle region $B$, and then uses injectivity of the enlarged
region to invert the joint region. It concludes that every vector in the
intersection already lies in
$\mathcal G_{1,\ldots,L_0+2}$; the reverse inclusion is immediate.
\footnote{Source location:
\path{Papers/2011.12127/TN-Review-main.tex}:2049--2077, especially the
middle-region argument around the displayed figures following
\path{eq:4:initiate-intersection-proof}.}

The source then says that this argument can be iterated. Hence the ground space
obtained from all length-$L_0+1$ terms agrees with the ground space obtained from
all longer length-$L_0+k$ terms. When $k=L_0$, the proof either appeals to the
previous $2L_0$-range uniqueness theorem or applies the same kind of argument
at the periodic boundary. The resulting theorem states that a normal
MPS which becomes injective after blocking $L_0$ sites has a unique ground state
for the parent Hamiltonian whose interaction range is $L_0+1$.
\footnote{Source location:
\path{Papers/2011.12127/TN-Review-main.tex}:2078--2079 for the iteration and
periodic-boundary closure-property sentence, and
\path{Papers/2011.12127/TN-Review-main.tex}:2087--2090 for the theorem labeled
\path{thm:4:unique-gs-L0_plus_1}.}

\section{The formal statement}

The corresponding formal statement keeps the intended hypotheses, with one
formalization-added range restriction, $L_0+1<N$, recorded as a scope
restriction in Section~\ref{sec:scope-minimal-ring}. In
mathematical terms, it assumes that $A$ is normal, that $A$ is injective
after blocking a positive number $L_0$ of sites, and that the ring has length
$N\ge 2$ with an interaction range $L$ satisfying $L_0<L\le N$, together with the
strict bound $L_0+1<N$. The desired
conclusion is that the periodic-boundary ground space is the line spanned by
the periodic MPS vector
$\Omega_N(A)$:
\[
  \mathcal G_{N,L}(A)=\mathbb C\,\Omega_N(A).
\]
In the formal statement this is the equality between the periodic-boundary ground
space and the subspace \(\mathbb C\,\Omega_N(A)\). The hard inclusion into
\(\mathbb C\,\Omega_N(A)\) is the remaining range-reduction lemma.
\footnote{Formal locations:
\path{TNLean/MPS/ParentHamiltonian/UniqueGroundState.lean}:902 for
\path{MPSTensor.chainGroundSpace_le_mpvSubmodule_of_normal_range_reduction},
and
\path{TNLean/MPS/ParentHamiltonian/UniqueGroundState.lean}:936 for
\path{MPSTensor.chainGroundSpace_eq_mpvSubmodule_normal}. The formal hypotheses
are named \path{IsNormal A} and \path{IsNBlkInjective A L0}, together with
\path{0 < L0}, \path{2 <= N}, \path{L0 + 1 < N}, and
\path{L0 < L <= N}.}

The blueprint records the corresponding closure-property step in
``Closure-property step toward uniqueness of the ground state'' in
\path{blueprint/src/chapter/ch13_parent_hamiltonian_normal_closing.tex}:1403--1417, labelled
\path{thm:normal_range_reduction_containment}. Its proof sketch separates the
open-chain intersection iteration from the comparison at the periodic boundary,
matching the source distinction between the intersection property and the
closure property.


The first mismatch is that the available intersection theorem is not a literal
expression of the proof through the region $B$. That theorem was written for a
tensor that is already injective at one site. Its proof decomposes the identity as a linear combination of the one-site matrices
$A_j$ and then iterates this single-letter intersection property through
contiguous windows.
\footnote{Formal locations:
\path{TNLean/MPS/ParentHamiltonian/IntersectionProperty.lean}:281--284 for
\path{MPSTensor.groundSpace_intersection},
\path{TNLean/MPS/ParentHamiltonian/IntersectionProperty.lean}:341--364 for the
use of \path{decompositionMap hA 1}, and
\path{TNLean/MPS/ParentHamiltonian/CyclicWindow.lean}:153--163 for the contiguous
iteration using that theorem.}
For a normal tensor with injectivity length $L_0$, the available inverse is the
inverse for words of length $L_0$, not a one-site inverse. This is not a
contradiction in the source argument; it is a mismatch between the diagrammatic
region-by-region proof and the available formal statements for intersection
properties.

The formal argument therefore separates the proof into two parts instead of
reproducing the compressed source proof verbatim. The open-chain component has been
proved by a block-injective range-reduction lemma: if every contiguous
window of length $L_0+1$ lies in the corresponding local MPS ground space, then
the full open chain lies in the full MPS ground space. Combining this with cyclic
window monotonicity gives an open-chain consequence from the cyclic length-$L$
constraints. This supplies the formal analog of the iterated open-boundary
part of the source proof, without expressing the inverse for the region $B$
through the older single-site intersection lemma.
\footnote{Formal locations:
\path{TNLean/MPS/ParentHamiltonian/UniqueGroundState.lean}:266 for
\path{MPSTensor.contiguous_mem_groundSpace_of_isNBlkInjective}, and
\path{TNLean/MPS/ParentHamiltonian/UniqueGroundState.lean}:328 for
\path{MPSTensor.chainGroundSpace_le_groundSpace_of_isNBlkInjective}.}

\section{The remaining commutation identity}

After the open-boundary range reduction, a periodic-boundary ground state can
be written in trace form with a matrix $X$. The point left to prove is that the
cyclic length-$L_0+1$ constraints force $X$ to commute with the one-site
matrices of $A$; block injectivity then forces $X$ to be scalar, and hence the
state lies in \(\mathbb C\,\Omega_N(A)\).

This is why the open-chain containment cannot simply be reversed. Let
\(\Gamma_N(X)\) denote the length-\(N\) trace-form vector
\(\Gamma_N(X)_\omega=\tr(A^\omega X)\). A vector
\(\Gamma_N(X)\in G_N(A)\) need not lie in the periodic-boundary ground space.
For a boundary-crossing window, the restriction has trace coefficients with \(X\)
between the two parts of the window word,
\begin{align}
  \tr\!\left(A^{\omega_{\mathrm{tail}}}X
    A^{\omega_{\mathrm{head}}}A^\rho\right),
\end{align}
whereas membership in \(G_L(A)\) would require the same coefficients to be
written as
\begin{align}
  \tr\!\left(A^{\omega_{\mathrm{tail}}}
    A^{\omega_{\mathrm{head}}}Y_\rho\right)
\end{align}
for a matrix \(Y_\rho\) depending only on the labels outside the window.
The closure-property comparison supplies the missing movement of \(X\) through
the boundary-crossing word.

In the coordinate form used here, \(\mu\) denotes the nonempty word placed on
the sites outside the two cyclic windows and \(\eta\) is the letter used on the
remaining outside sites. This is a coordinate form of the periodic-boundary
closure-property sentence in Section~IV.C, lines 2078--2079. The reduction
isolates the movement of \(X\) as an identity between boundary-restriction
matrices in the closure-property comparison, which implies the one-site
commutation identity.
After choosing boundary-restriction matrices for the two boundary-crossing
restrictions, the verified one-sided boundary-product lemma gives, for every
physical letter \(j\),
\[
  W_\eta A^j=A^\mu A^jX,
  \qquad
  XA^jA^\mu=A^jV_\eta .
\]
Here \(W_\eta\) is the boundary-restriction matrix for the boundary-crossing
restriction beginning at the last site, and \(V_\eta\) is the
boundary-restriction matrix for the second boundary-crossing restriction.
\footnote{Formal locations:
\path{TNLean/MPS/ParentHamiltonian/BoundaryClosingStripping.lean}:338 for
\path{MPSTensor.closure_property_boundary_restrictions_eq_of_groundSpaceMap},
and
\path{TNLean/MPS/ParentHamiltonian/BoundaryClosingStripping.lean}:371--372 for the use of
\path{MPSTensor.closure_property_boundary_one_sided_products_of_groundSpaceMap}.}
This local matrix comparison gives the following commutation identity through
the block-injective commutation lemma: for every physical letter \(k\),
\[
  XA^k=A^kX.
\]
This is the one-site commutation step in the periodic-boundary closure-property
argument.\footnote{
In the formal proof of
\leanid{MPSTensor.closure_property_boundary_restrictions_eq_of_groundSpaceMap},
the commutation identity is obtained at
\path{TNLean/MPS/ParentHamiltonian/BoundaryClosingStripping.lean}:376--378.}

For an \(N\)-site vector \(\psi\),
\(\operatorname{Res}^{\tau}_{i,L}(\psi)\) denotes the \(L\)-site vector whose
\(\sigma\)-coordinate is the coordinate of \(\psi\) obtained by placing
\(\sigma\) in the cyclic window of length \(L\) beginning at \(i\), and by
using \(\tau\) on the remaining sites.

Once this identity is available, the boundary-restriction equality lemma proves
that the two boundary-restriction matrices have the same first-letter products:
\[
  W_\eta A^j=V_\eta A^j,
\]
for every \(\eta\) and \(j\). The first-letter restriction lemma then identifies
the two cyclic restrictions
\[
  \operatorname{Res}^{\tau^+_\eta(\mu)}_{M,L_0+1}(\psi)
  =
  \operatorname{Res}^{\tau^-_\eta(\mu)}_{M+1-L_0,L_0+1}(\psi).
\]
Thus the equality of boundary restrictions is no longer the primitive open
comparison: it is a checked consequence of the one-sided products and the
commutation identity.\footnote{Formal locations:
\path{TNLean/MPS/ParentHamiltonian/BoundaryMatrixBlock.lean}:134 for
\leanid{MPSTensor.boundary_restrictions_eq_of_commutes_and_one_sided}, and
\path{TNLean/MPS/ParentHamiltonian/BoundaryClosingStripping.lean}:379 for its
use in
\leanid{MPSTensor.closure_property_boundary_restrictions_eq_of_groundSpaceMap}.}

The block-injective commutation lemma reduces the commutation identity itself
to this local matrix equation. It states that, if \(A\) is
\(L_0\)-block-injective and there are matrices \(Y_\nu\) such that, for every
length-\(L_0\) word \(\alpha\) and every complementary word \(\nu\),
\[
  XA^\alpha A^\nu=A^\alpha Y_\nu,
\]
then \(XA^k=A^kX\) for every physical letter \(k\).\footnote{Formal location:
\path{TNLean/MPS/ParentHamiltonian/BoundaryMatrixBlock.lean}:46 for
\leanid{MPSTensor.boundary_matrix_commutes_of_isNBlkInjective_of_block_matEq}.}
The trace rotation at the last cyclic window has now been isolated as the
identity
\[
  \operatorname{tr}\left(A^j X A^\alpha A^\nu\right)
  =
  \operatorname{tr}\left(A^j A^\alpha Y_\nu\right),
\]
formalized as
\leanid{MPSTensor.closure_property_boundary_block_window_trace_eq_of_groundSpaceMap}.
These identities probe the matrix difference only by the single letters
\(A^j\), and therefore do not separate matrices for a general
\(L_0\)-block-injective tensor. The remaining mathematical task is to obtain
the corresponding trace identities against all length-\(L_0\) word products
\(A^\beta\), or an equivalent coordinate comparison; those products span the
full matrix algebra by block injectivity. After that equation is proved, the
block-injective commutation lemma supplies the one-site commutation identity,
the boundary-restriction equality lemma supplies the equality of the two
boundary restrictions, and the periodic-boundary comparison follows.\footnote{The
present source-labelled statement for this step is
\leanid{MPSTensor.closure_property_boundary_block_window_equation_of_groundSpaceMap}.}

Equivalently, the same boundary phenomenon can be recognized in coordinate
terms. For an \(N\)-site vector \(\psi\), the two restrictions are obtained from
the two boundary-crossing words
\[
  \tau^{+}_{\eta}(\mu)_i =
  \begin{cases}
    \mu_{i-L_0}, & L_0\le i<N-1,\\
    \eta,        & \text{otherwise},
  \end{cases}
  \qquad
  \tau^{-}_{\eta}(\mu)_i =
  \begin{cases}
    \mu_{i-1}, & 1\le i<N-L_0,\\
    \eta,      & \text{otherwise}.
  \end{cases}
\]
The direct comparison of these two restrictions, and the left-multiplied
coordinate comparison obtained after taking first-letter restrictions, are
consequences of the one-site commutation identity \(XA^k=A^kX\).

\section{Scope restriction: the minimal ring}
\label{sec:scope-minimal-ring}

The source theorem \path{thm:4:unique-gs-L0_plus_1}
(\path{Papers/2011.12127/TN-Review-main.tex}:2087--2090) asserts a unique ground
state for the range-$(L_0+1)$ parent Hamiltonian on a normal MPS, placing no
lower bound on the ring length $N$ beyond the $N\ge L_0+1$ needed to define a
range-$(L_0+1)$ term. The minimal case $N=L_0+1$, in which the single
length-$(L_0+1)$ window already wraps the whole ring, is therefore covered by the
source statement.

The formal statements add the strict hypothesis $L_0+1<N$, so they cover only
$N\ge L_0+2$ and exclude that minimal ring. The restriction is forced by the
periodic-boundary closure-property argument: the comparison ranges over a
complement word $\mu\colon \mathrm{Fin}\,(N-(L_0+1))\to\mathrm{Fin}\,d$, which is
the empty function exactly when $N=L_0+1$, so the two boundary-crossing windows
coincide and the comparison carries no information. This is a genuine narrowing
of the source theorem rather than a typo: the conclusion at $N=L_0+1$ is true
(it is the source's minimal case), but the present proof does not reach it.

Elimination plan: prove the empty-complement case of the periodic-boundary
argument. When $N=L_0+1$, the complement word has length zero and the two
boundary-crossing words coincide; one must still use the cyclic local constraints
to identify the periodic ground space with the MPS line. Thus the missing step is
not a single-window kernel computation, but a separate cyclic-boundary argument
in the degenerate complement case. Alternatively, extend the witness-comparison
lemma to include the empty complement word. Either route removes $L_0+1<N$ from
\path{MPSTensor.wrapped_mirror_witness_agree_of_chainGroundSpace},
\path{MPSTensor.chainGroundSpace_le_mpvSubmodule_of_normal_range_reduction},
\path{MPSTensor.chainGroundSpace_eq_mpvSubmodule_normal},
\path{MPSTensor.parentHamiltonian_unique_gs_normal}, and, through this
range-reduction theorem, the range-$2L_0$ corollary
\path{MPSTensor.parentHamiltonian_unique_gs_injective}. In the latter corollary,
the bound $L_0+1<N$ is automatic from $2L_0\le N$ for $1<L_0$ and only excludes
the two-site case $L_0=1$, $N=2$. The narrowing is recorded in these declarations
with a \texttt{Scope restriction} docstring marker.

\section{Conclusion}

The classification is an open gap. It identifies a difference between the
compressed source proof and the available formal lemmas; it is not a
source-error claim. The cited Cirac--Perez-Garcia--Schuch--Verstraete 2021
argument is not refuted. The source compresses two mathematical steps which the
formalization must keep separate: the $B$-region inverting and growing-back
open-chain range reduction, and the movement of the trace-form matrix \(X\)
through the boundary-crossing word in the periodic-boundary comparison.

The open-chain part has been proved under the intended block-injective
hypotheses, though not by reusing the older single-site intersection theorem.
The last mathematical point is the commutation identity \(XA^k=A^kX\) for every
physical letter \(k\). The present formalization proves that the full
boundary-restriction equality follows from this identity and the one-sided
boundary products. If this local matrix equation proves the commutation identity,
the remaining difference is only expository. If it reveals an additional
mathematical obstruction, the conclusion should be revised accordingly.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
