\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{The Martingale Principal-Angle Estimate for MPS Parent Hamiltonians}
\author{}
\date{2026-06-04}

\begin{document}
\maketitle

\section{The source assertion}

The martingale-method paragraph in Cirac--Perez-Garcia--Schuch--Verstraete 2021
can be compared with the quantitative estimate isolated in the formal
parent-Hamiltonian proof.\footnote{For
    traceability, the finite-row reduction is recorded in
    \href{https://github.com/LionSR/TNLean/issues/1044}{issue \#1044}. The remaining
    quantitative estimate is recorded in
    \href{https://github.com/LionSR/TNLean/issues/952}{issue \#952}, and the broader
    parent-Hamiltonian theorem in
    \href{https://github.com/LionSR/TNLean/issues/460}{issue \#460}.}
The relevant source passage is the discussion of gaps for MPS parent Hamiltonians
in \cite[Section~IV.C]{Cirac2021Matrix}, where the local paragraph
\path{sec:4:hams-gaps} develops the inequalities labelled
\path{eq:4:martingale-1} and \path{eq:4:martingale-2}.

Let
\[
    H=\sum_i h_i
\]
be a frustration-free Hamiltonian whose local terms are projections. The
source recalls that a gap lower bound $\gamma>0$ is equivalent, at the
quadratic-form level, to
\[
    H^2\ge \gamma H.
\]
Since $h_i^2=h_i$, the expansion of $H^2$ separates diagonal terms,
overlapping products, and non-overlapping products. The non-overlapping
products are treated as non-negative. Thus it is enough to control each
overlapping pair by an anticommutator estimate
\begin{equation}
    h_i h_j+h_j h_i
    \ge -c_{ij}(1-\gamma)(h_i+h_j),
    \tag{\path{eq:4:martingale-2}}
\end{equation}
where the coefficients $c_{ij}$ are chosen so that the contributions in each
row add up to one. The formal row-sum algebra uses the slightly more flexible
condition that, for each fixed $i$, the sum of the coefficients attached to
terms overlapping $h_i$ is at most one.

The same paragraph interprets this estimate as a lower bound on the minimum
non-zero angle between the ground spaces of overlapping regions. It then cites
the martingale method and the MPS parent-Hamiltonian results of
Fannes--Nachtergaele--Werner and Nachtergaele, together with the later
Kastoryano--Lucia formulation, for the conclusion that MPS parent Hamiltonians
have a uniform positive spectral gap after a suitable blocking
\cite{Fannes1992Finitely,Nachtergaele1996Spectral,Kastoryano2018Martingale}.
In the convention of this note, the Kastoryano--Lucia formulation cited there
is the estimate for
\begin{align}
    \delta(\ell):=\|q_i q_j-P\|,
\end{align}
where $q_i$ and $q_j$ are the projectors onto the ground spaces
$\ker h_i$ and $\ker h_j$, $P$ is the projector onto
$\ker h_i\cap\ker h_j=\ker(h_i+h_j)$, and $\ell$ is the size of the
overlap region. The review writes the displayed product with the symbols
$h_i,h_j$; in the convention fixed above, those symbols must be read as the
ground-space projectors $q_i,q_j$, not as the excitation projections $h_i$ used
in \path{eq:4:martingale-2}. If one inserted the excitation projections into
$\|h_ih_j-P\|$, the norm would be at least one on the common ground space. With
the ground-space projectors, the cited result states that $\delta(\ell)$ decays
exponentially with universal constants whenever the frustration-free
Hamiltonian is gapped, and that this implies the martingale condition because
both quantities depend only on the principal angles between
$\ker h_i$ and $\ker h_j$.
The review text does not spell out the numerical constants or the precise
finite cyclic blocking convention needed for the formal statement below.
In particular, the intersection identity
\[
    \ker h_i\cap\ker h_j=\ker(h_i+h_j)
\]
identifies the common local ground space, but it is not a substitute for the
quantitative principal-angle estimate.

The source route that has to be matched is more precise than the review
paragraph. Kastoryano--Lucia define, for finite regions $A$ and $B$,
\[
    \delta(A,B)
      =\|(P_A-P_{A\cup B})(P_B-P_{A\cup B})\|
      =\|P_A P_B-P_{A\cup B}\|
\]
under frustration-freeness, and identify this with the cosine of the first
principal angle between the two reduced ground-space projectors
\cite[Definition~2 and Remark~2]{Kastoryano2018Martingale}. Nachtergaele's
martingale hypotheses C3 and C3$'$ instead use the projection estimates
\[
    \|G_{\Lambda_{n+1}\setminus\Lambda_{n-l}}E_n\|\le\varepsilon_l,
    \qquad
    \|G_{\Lambda_{n+p}\setminus\Lambda_{n-l}}E_n^{(p)}\|\le\eta_l,
\]
with thresholds $\varepsilon_l<1/\sqrt{l+1}$ and
$\eta_l<1/\sqrt{2}$, and Theorem~3 gives the corresponding gap lower bounds
\cite[Equations~(2.4)--(2.5) and Theorem~3]{Nachtergaele1996Spectral}. In
Section~6, Nachtergaele proves the GVBS/MPS gap by verifying C3$'$ using the
projection estimates inherited from Fannes--Nachtergaele--Werner
\cite[Section~6]{Nachtergaele1996Spectral}. The remaining formal task is to
convert one of these source estimates to the cyclic-window inequality used
below, with constants in the same blocking convention.

\section{Finite-row martingale reduction}

The current development separates the finite-row martingale algebra from the
remaining MPS-specific principal-angle estimate. Fix an interaction range
$L>1$ and an $N$-site periodic chain with $N\ge 2L$. Let $p_i$ denote the
Hilbert-space representative of the length-$L$ local parent-Hamiltonian
projection at the cyclic window starting at $i$. Write $i\sim_L j$ when the two
length-$L$ cyclic windows intersect and $i\ne j$.
For each fixed $i$, the number of indices $j$ with $i\sim_L j$ is at most
\[
    m=2(L-1).
\]
The row-sum reduction specializes the source coefficients to
\[
    c_{ij}=\frac{1}{2(L-1)}
\]
for overlapping cyclic windows and to $0$ otherwise. The row-cardinality bound
then gives $\sum_j c_{ij}\le 1$.

With this choice, the constant in the local statement is
\[
    \gamma_L=\frac{1}{4L}.
\]
The ordered overlapping estimate sufficient for the gap bound is
\begin{equation}
    \operatorname{Re}\langle p_i v,p_j v\rangle
    \ge
    -\left(1-\frac{1}{4L}\right)\frac{1}{2(L-1)}
    \operatorname{Re}\langle p_i v,v\rangle
    \tag{F}
\end{equation}
for every $N\ge 2L$, every off-diagonal overlapping pair $i\sim_L j$, and every
vector $v$ in the finite chain Hilbert space. Given this estimate, the already
formalized row-sum argument proves $H_{N,L}^2\ge \gamma_L H_{N,L}$, and hence a
norm lower bound with constant $\gamma_L$ on the orthogonal complement of the
ground space.

The proof also isolates a sufficient norm-compression form of the same
input:
\begin{equation}
    \|p_i p_j v\|
    \le
    \left(1-\frac{1}{4L}\right)\frac{1}{2(L-1)}
    \|p_i v\|.
    \tag{N}
\end{equation}
For a symmetric projection $p_i$ one has
\[
    \operatorname{Re}\langle p_i v,p_j v\rangle
    =\operatorname{Re}\langle p_i v,p_i p_j v\rangle
    \ge -\|p_i v\| \|p_i p_j v\|.
\]
Thus (N), together with $\operatorname{Re}\langle p_i v,v\rangle=\|p_i v\|^2$,
implies (F). This is a sufficient projection-geometry comparison used in the
proof: a norm-compression estimate for each overlapping pair is converted
by Cauchy--Schwarz to the ordered cross-term estimate, and the finite row-sum
argument converts those ordered estimates into the global gap bound. The Lean
development also records the source anticommutator row-sum reduction directly,
without passing through this stronger one-sided estimate.\footnote{The
    current formal anchors include the anticommutator row-sum reductions
    \leanid{crossTerm_sum_bound_of_anticommutator_rowCol},
    \leanid{quadraticForm_sum_projections_of_anticommutator_rowCol}, and
    \leanid{parentHamiltonianES_gap_bound_of_anticommutator_local_term_bounds}.
    The finite-overlap anticommutator reductions are
    \leanid{quadraticForm_sum_projections_of_finite_overlap_anticommutator},
    \leanid{parentHamiltonianES_gap_bound_of_finite_overlap_anticommutator}, and
    \leanid{parentHamiltonianES_gap_bound_of_cyclic_window_overlap_anticommutator}.
    The corresponding final gap statements are
    \leanid{parentHamiltonianES_gap_bound_of_anticommutator} and
    \leanid{parentHamiltonian_gapped_of_anticommutator} in
    \path{TNLean/MPS/ParentHamiltonian/Martingale/Gap.lean}.
    The cyclic-window compression reductions are
    \leanid{parentHamiltonianES_gap_bound_of_cyclic_window_overlap_ordered_cross_term},
    \leanid{parentHamiltonianES_gap_bound_of_cyclic_window_overlap_norm_bound},
    and
    \leanid{parentHamiltonianES_gap_bound_of_cyclic_window_overlap_norm_bound_of_le}
    in \path{TNLean/MPS/ParentHamiltonian/Martingale/Reduction.lean}. The
    corresponding constant-$\eta$ reductions are
    \leanid{parentHamiltonianES_gap_bound_of_cyclic_window_overlap_norm_bound_of_lt}
    in the same file and
    \leanid{parentHamiltonianES_gap_bound_of_overlap_norm_constant},
    \leanid{parentHamiltonian_gapped_of_overlap_norm_constant} in
    \path{TNLean/MPS/ParentHamiltonian/Martingale/Gap.lean}. The
    projection-geometry reductions used by these statements are in
    \path{TNLean/Analysis/ProjectionGeometry.lean}. The declaration names
    identify the intermediate ordered cross-term estimate and the sufficient
    norm-compression estimates; this note uses ``anticommutator estimate'' for
    the source condition and ``norm compression'' for the local sufficient
    strengthening.}

The anticommutator row-sum formulation, including its finite-overlap
cyclic-window specialization, is now the formal condition matching the source.
The flexible constant formulation records a sufficient norm-compression variant:
instead of requiring the special coefficient in (N), it assumes that for some
$\eta\ge 0$,
\begin{equation}
    \eta\,2(L-1)<1
    \tag{E}
\end{equation}
and that every overlapping cyclic-window pair satisfies
\begin{equation}
    \|p_i p_j v\|\le \eta\|p_i v\|.
    \tag{N$_\eta$}
\end{equation}
The finite-row argument then gives a positive gap with lower bound
$1-\eta\,2(L-1)$. The special displayed estimate (N) is one sufficient choice
of $\eta$, but the source comparison need only supply a compression constant
strictly smaller than $1/(2(L-1))$ for the blocking convention used here.

\section{What remains to compare}

The source discussion and the formal theorem are aligned at the qualitative
level: both use the martingale expansion, non-negativity of non-overlapping
terms, row-sum coefficients for overlapping terms, and principal-angle estimates
for local ground spaces. The point still requiring a mathematical check is
quantitative and conventional.

First, the source-matching formal theorem asks for the cyclic-window
anticommutator estimate
\[
    p_i p_j+p_jp_i
    \ge
    -\left(1-\frac{1}{4L}\right)\frac{1}{2(L-1)}(p_i+p_j)
\]
for overlapping off-diagonal pairs. The review text states the same martingale
condition with general coefficients $c_{ij}$ whose rows add to one, and it
refers to blocking and principal angles to obtain a positive gap. It does not by
itself identify the Fannes--Nachtergaele--Werner, Nachtergaele, or
Kastoryano--Lucia estimates with this coefficient and this fixed cyclic-window
convention. A proof may instead derive a row-sum anticommutator estimate with
different positive constants and then specialize the final gap constant
accordingly.

Second, the blocking parameter in the cited martingale argument must be matched
with the formal convention. The formal statement is uniform for all $N\ge 2L$
and for the length-$L$ cyclic windows used in the formal finite-chain
Hilbert-space realization. A source proof that produces some positive gap after
increasing or changing the blocking may still be mathematically sufficient for the
existential MPS gap theorem, but it would not immediately be the displayed
all-$L$ estimate with constant $1/(4L)$ unless the constants are converted to
the same blocking convention.

Consequently the remaining comparison is not whether the row-sum algebra is
valid; that part has been isolated and formalized. The remaining comparison is
whether the cited FNW--Nachtergaele--Kastoryano principal-angle estimates supply
the required cyclic-window anticommutator estimate, with constants independent
of the chain length, after the same blocking convention has been fixed.
The qualitative fact that two finite-dimensional subspaces have a well-defined
principal-angle decomposition is not enough: the martingale reduction needs a
uniform numerical bound after the same blocking convention has been fixed.

The norm-compression route remains available as a stronger sufficient target:
any estimate (N$_\eta$) with $\eta\,2(L-1)<1$ gives a positive gap through the
already formalized reduction. There is, however, a directional point. If
(N$_\eta$) is imposed for all vectors
$v$, then $p_i v=0$ implies $p_i p_j v=0$. Thus $p_j$ maps $\ker p_i$ into
$\ker p_i$. This is stronger than merely knowing that the
smallest non-zero principal angle between the two local ground spaces is
positive. A proof using compression must therefore identify a principal-angle
estimate that gives the directed compressed-product bound (N$_\eta$). A direct
proof can instead target the anticommutator form
\[
    h_i h_j+h_jh_i\ge -c_{ij}(1-\gamma)(h_i+h_j)
\]
used in \cite[Section~IV.C]{Cirac2021Matrix}.

\section{Relation to the blueprint}

The blueprint item labelled \path{lem:martingale_mps} in
\path{blueprint/src/chapter/ch13_parent_hamiltonian_spectral_gap.tex}
separates the finite-row martingale reduction from the still-open quantitative
estimate.
It does not assert that the displayed all-$L>1$ cyclic-window bound has already
been obtained from the cited principal-angle estimates. Instead, it records the
source anticommutator estimate as the hypothesis of
Theorem \path{thm:anticommutator_gap_bound}, and records norm-compression
estimates as sufficient stronger hypotheses.

The source-matching theorem entry
\path{thm:anticommutator_gap_bound} points to
\leanid{parentHamiltonianES_gap_bound_of_anticommutator} in
\path{TNLean/MPS/ParentHamiltonian/Martingale/Gap.lean}. Its hypothesis is the
cyclic-window anticommutator lower bound with coefficient
\[
    \left(1-\frac{1}{4L}\right)\frac{1}{2(L-1)}.
\]
The theorem entries
\path{thm:overlap_norm_gap_bound} and
\path{thm:overlap_norm_gap_bound_constant} point respectively to the special
coefficient theorem and to the constant-$\eta$ theorem in
\path{TNLean/MPS/ParentHamiltonian/Martingale/Gap.lean}. The special coefficient
theorem is the specialization of the constant-$\eta$ theorem with
\[
    \eta=\left(1-\frac{1}{4L}\right)\frac{1}{2(L-1)},
    \qquad
    1-\eta\,2(L-1)=\frac{1}{4L}.
\]
The two norm-compression theorem entries are therefore conditional on the same
stronger input, specialized either to this displayed coefficient or to an
arbitrary coefficient satisfying (E). Thus all finite-row, cyclic-window,
non-overlap, anticommutator, norm-compression, and spectral-theorem reductions
are proved, while the MPS-specific anticommutator estimate remains to be proved
or matched precisely to the cited martingale estimates.

Once the quantitative comparison with the cited martingale estimates is settled,
the blueprint should say that the cyclic-window anticommutator estimate follows
from those estimates under the convention used here. If the proof instead goes
through norm compression, the blueprint should say explicitly that the
compression bound (N$_\eta$) with $\eta\,2(L-1)<1$ is a stronger sufficient
estimate implying the anticommutator route.

\section{Conclusion}

The conclusion is a difference between the review paragraph and the formal
quantitative statement, with one mathematical comparison still to be made. The
source states the martingale condition in terms of general row-sum coefficients
and invokes principal-angle estimates to obtain a positive MPS parent-Hamiltonian
gap. The formal proof contains the direct finite-overlap anticommutator
reduction for cyclic windows, specializes the row coefficients to
$c_{ij}=1/(2(L-1))$ on overlapping pairs, and leaves the remaining MPS-specific
work as the source anticommutator estimate. It also records the stronger
norm-compression route
\[
    \|p_i p_j v\|\le \eta\|p_i v\|,
    \qquad
    \eta\,2(L-1)<1.
\]
The earlier coefficient
\[
    \left(1-\frac{1}{4L}\right)\frac{1}{2(L-1)}
\]
is a sufficient special case for this stronger route, not the only possible way
to use the formal martingale reduction.

The Cirac--Perez-Garcia--Schuch--Verstraete 2021 statement is not being
challenged. The comparison records that the available formal argument is more
explicit than the review paragraph: the exact constants and the cyclic-window
blocking convention still have to be compared with the FNW, Nachtergaele, and
Kastoryano--Lucia estimates.

\textbf{Verdict:} open gap.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
