\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{The Perron--Frobenius Rank-One Step in CPGSV17 Appendix C.2}
\author{}
\date{2026-04-30}

\begin{document}
\maketitle

\section{The assertion}

Appendix C.2 of Cirac--P\'erez-Garc\'ia--Schuch--Verstraete relates the strong
area law and zero correlation length for a simple matrix product density operator
to an outer-product factorization of a finite trace matrix
\cite[Appendix~C.2, Lemmas~C.3--C.4]{Cirac2016MPDO_arXiv}. The relevant source
and formal locations are listed for comparison.
\footnote{For traceability, this note corresponds to
\href{https://github.com/LionSR/TNLean/issues/1041}{issue \#1041}. The
underlying question was isolated in
\href{https://github.com/LionSR/TNLean/issues/832}{issue \#832}; the finite
counterexample was added in
\href{https://github.com/LionSR/TNLean/pull/849}{pull request \#849}; and the
positive-semidefinite replacement theorem was added in
\href{https://github.com/LionSR/TNLean/pull/917}{pull request \#917}. The
source passage is
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:1394--1499}.}

The paper constructs positive neighboring operators $\eta_{k,h}$ and defines
\[
  T_{k,h}=\operatorname{tr}(\eta_{k,h}).
\]
Thus $T$ is entrywise nonnegative. Under the strong area law the source argues
that $T$ is primitive, in the sense that some power of $T$ has all entries
strictly positive. Under zero correlation length it obtains
\[
  \operatorname{tr}(T^N)=\operatorname{tr}(T)
\]
for every positive integer $N$, with $\operatorname{tr}(T)=1$ after
normalization. The proof then invokes fixed-point theory for primitive
nonnegative matrices and concludes that $T$ has one eigenvalue equal to $1$
and all other eigenvalues equal to $0$. From this it concludes the desired
factorization
\[
  T_{k,h}=a_k b_h.
\]

The last implication is the point at issue. Primitivity and the equalities of
positive-power traces control the nonzero eigenvalues of $T$, but they do not
exclude nilpotent Jordan structure on the generalized eigenspace for the
eigenvalue $0$. Such nilpotent structure is invisible to traces of positive
powers and can prevent $T$ from having rank one.

\section{The counterexample}

Consider the real $3\times 3$ matrix
\[
  T=
  \begin{pmatrix}
    \tfrac12 & \tfrac12 & 0\\
    \tfrac16 & \tfrac16 & \tfrac23\\
    \tfrac13 & \tfrac13 & \tfrac13
  \end{pmatrix}.
\]
All entries are nonnegative, and
\[
  T^2=
  P:=\frac13
  \begin{pmatrix}
    1&1&1\\
    1&1&1\\
    1&1&1
  \end{pmatrix}.
\]
Hence $T$ is primitive, since $T^2$ has all entries strictly positive. Also
$P^2=P$ and $PT=TP=P$, so $T^N=P$ for every $N\geq 2$. Since
$\operatorname{tr}(T)=\operatorname{tr}(P)=1$, it follows that
\[
  \operatorname{tr}(T^N)=\operatorname{tr}(T)=1
\]
for every positive integer $N$.

Nevertheless $T$ is not an outer product. If $T=a b^{\top}$, then the
entry $T_{1,1}=1/2$ implies $a_1\neq 0$. The entry $T_{1,3}=0$ then
forces $b_3=0$, but this contradicts $T_{2,3}=2/3$. Thus the universal
matrix implication
\[
  \text{primitive} + \text{constant positive-power traces}
  \Longrightarrow \text{rank one}
\]
is false for finite entrywise nonnegative real matrices.
\footnote{The formal counterexample is
\path{TNLean/Archive/PerronFrobeniusRankOneCounterexample.lean}, especially the
matrix definition at lines 63--65 and the theorem \texttt{counterexample} at
lines 150--154. It is also summarized in
\path{docs/counterexamples.md:19--27}.}

This example has the form $T=P+\frac16 xy^{\top}$, where
$x=(1,-1,0)^{\top}$ and $y=(1,1,-2)^{\top}$. The perturbation satisfies
$(xy^{\top})^2=0$ and is annihilated by $P$ on both sides. It therefore
leaves all traces of positive powers unchanged from the Perron projector while
raising the rank of $T$.

\section{The corrected statement}

The formal development proves a corrected sufficient matrix theorem. Let $T$
be a finite real square matrix. If $T$ is positive semidefinite,
$\operatorname{tr}(T)=1$, and
$\operatorname{tr}(T^N)=\operatorname{tr}(T)$ for every positive integer
$N$, then $T$ has an outer-product factorization $T=a b^{\top}$.
\footnote{The formal statements are
\texttt{Matrix.PosSemidef.trace\_powers\_constant\_implies\_rank\_one} and
\texttt{Matrix.primitive\_trace\_powers\_constant\_implies\_rank\_one\_of\_posSemidef}
in \path{TNLean/Algebra/PerronFrobenius/RankOne.lean:181--240}.}
Primitivity is not needed for this corrected matrix theorem once positive
semidefiniteness and trace normalization are assumed.

The proof is spectral. Positive semidefiniteness makes $T$ Hermitian and
diagonalizable with nonnegative real eigenvalues $\lambda_i$. The trace
normalization gives $\sum_i \lambda_i=1$. The trace-power hypothesis at
$N=2$ gives $\sum_i \lambda_i^2=1$. For nonnegative numbers with sum one,
this forces exactly one $\lambda_i$ to be $1$ and all the others to be
$0$. Therefore the diagonal form has rank one, and unitary conjugation gives
an outer-product factorization of $T$.

This corrected theorem should not be read as a refutation of the intended MPDO
result. It identifies the extra matrix structure that is sufficient to justify
the rank-one step. It remains to show that the trace matrix arising from the neighboring operators
has this positive-semidefinite or Hermitian structure
$\eta_{k,h}$. Positivity of each $\eta_{k,h}$ gives
$T_{k,h}=\operatorname{tr}(\eta_{k,h})\geq 0$ entry by entry, but entrywise
nonnegativity is not matrix-level positive semidefiniteness or Hermitian
symmetry of $T$.
\footnote{The entrywise nonnegativity result is
\texttt{MPOTensor.ExplicitEtaOperators.traceMatrixRe\_nonneg} in
\path{TNLean/MPS/MPDO/SimpleLocalStructure.lean:306--318}. The
positive-semidefinite reformulation for the rank-one conclusion is
\texttt{MPOTensor.sal\_zcl\_implies\_rank\_one\_T\_of\_posSemidef} in
\path{TNLean/MPS/MPDO/SimpleLocalStructure.lean:350--366}. The corresponding
blueprint discussion is
\path{blueprint/src/chapter/ch02b_mpdo.tex:568--680}.}

\section{Conclusion}

The proof in Appendix C.2 contains an unjustified matrix step as written: a
primitive entrywise nonnegative matrix with constant traces of positive powers
need not have rank one. The obstruction is a nilpotent Jordan component at the
zero eigenvalue, which is not detected by the trace identities.

The formal development proves a corrected sufficient theorem: positive
semidefiniteness, trace normalization, and the same trace-power hypothesis imply
the required rank-one factorization. What remains for the unconditional MPDO
argument is not another Perron--Frobenius statement for arbitrary nonnegative
matrices, but an MPDO-specific proof that the trace matrix obtained from the
$\eta$-operators has the Hermitian or positive-semidefinite structure needed by
the corrected theorem. Thus the source proof has a genuine unsupported
intermediate assertion, while the intended MPDO theorem itself has not been
disproved.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}