\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Projective Reading of Proportional MPVs in CPSV16}
\author{}
\date{2026-05-11}

\begin{document}
\maketitle

\section{Source phrase}

Cirac--P\'erez-Garc\'ia--Schuch--Verstraete state Theorem~\texttt{thm1}
in \cite{Cirac2016MPDO_arXiv}, lines~1170--1192, for tensors whose
matrix-product vectors are ``proportional to each other''. The displayed
source statement does not spell out whether the proportionality scalar at a
fixed length is required to be nonzero. The two formal readings are
\begin{align}
    \exists c_N,  V^{(N)}(A)=c_N V^{(N)}(B),\\
    \exists c_N\ne 0,  V^{(N)}(A)=c_N V^{(N)}(B).
\end{align}
\begin{align}
    V^{(N)}(A)=0\cdot V^{(N)}(B).
\end{align}
The scalar multiple with coefficient $0$ gives no projective point.

\section{Formal reading}

The formal predicate \leanid{NonzeroProportionalMPV₂} requires, at each
length $N$, a scalar $c_N \neq 0$ such that
\begin{align}
    V^{(N)}(A) = c_N V^{(N)}(B).
\end{align}
This is the projective reading of proportionality: the two vectors determine
the same point in projective space. It is the reading used by the proof of
Theorem~\texttt{thm1}. In the block-selection step, the paper assumes that all
overlaps of a fixed block on one side with the blocks on the other side tend
to zero, expands the two BNT decompositions, and invokes Corollary~\texttt{Lem1}
to contradict proportionality. If $c_N$ were allowed to be zero at a
length where the vector on the left also vanished, the identity
$V^{(N)}(A)=0\cdot V^{(N)}(B)$ would not express equality of projective
MPV families and would not supply the nonzero scalar used in the contradiction.

To make this last use explicit, write
$a_j(N)=V^{(N)}(A_j)$ and $b_\ell(N)=V^{(N)}(B_\ell)$ for the BNT-sector
vectors at length $N$. Let $g_A$ and $g_B$ be the numbers of sectors in the two
decompositions, and let $\mu_j,\nu_\ell\in\mathbb C$ be the corresponding
sector weights in the one-representative form used in the block-selection
argument. The proportionality relation has the form
\begin{align}
    \sum_{j=1}^{g_A}\mu_j^N a_j(N)
    = c_N\sum_{\ell=1}^{g_B}\nu_\ell^N b_\ell(N).
\end{align}
Projecting this equality to a selected $B$-sector $k$ gives
\begin{align}
    c_N\nu_k^N\langle b_k(N),b_k(N)\rangle=0.
\end{align}
The contradiction is therefore the incompatibility of this equality with
\begin{align}
    c_N\ne 0,\qquad
    \nu_k\ne 0,\qquad
    \langle b_k(N),b_k(N)\rangle\ne 0.
\end{align}
Under the canonical-form hypotheses used in the source argument, the selected
BNT block states have nonzero norm on the tail of lengths used in the
contradiction; in that range the weak scalar-multiple predicate already forces
the proportionality scalar to be nonzero. Thus the formal predicate records
the projective form needed at the use site, rather than adding a separate
asymptotic assumption to the block-selection argument.

\section{Effect on the theorem}

The nonzero condition does not change the intended conclusion of the source
theorem. It records the conventional mathematical meaning of proportional
nonzero vectors in the source argument. The weaker scalar-multiple predicate
\leanid{ProportionalMPV₂}, where $c_N$ may vanish, remains available for
older auxiliary lemmas, but it is not used as the paper-faithful hypothesis
for the proportional Fundamental Theorem.

\section{Elimination plan}

No extra mathematical theorem is needed for the present projective reading of
CPSV16 lines~1170--1192. If a literal weak-scalar reading is later required,
it should be formalized as a separate bridge theorem: prove from the source
canonical-form hypotheses that the relevant MPV vectors do not vanish at the
lengths used in the argument, then derive
\leanid{NonzeroProportionalMPV₂} from the weak predicate before applying the
block-selection proof.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
