\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Canonical-Form Weight Normalization and the Proportional MPV Comparison}
\author{}
\date{2026-05-10}

\begin{document}
\maketitle

\section{The source argument}

Cirac--P\'erez-Garc\'ia--Schuch--Verstraete 2016 builds the multi-block route to
the Fundamental Theorem on top of the canonical form
\cite[Section~II]{Cirac2016MPDO_arXiv}. A tensor $A=(A^i)_{i=1}^d$ is in
canonical form when, after a basis change of the bond, it decomposes as
\begin{equation}
  A^i=\bigoplus_{k=1}^r \mu_k\,A_k^i,
  \tag{\path{eq:II_CF1}}
\end{equation}
where each block tensor $A_k$ is normal and the scalars $\mu_k$ are nonzero.
The text immediately fixes a normalization on the weights:
\begin{quote}
  ``we can always choose $|\mu_k|\le 1$ and at least one of them equals one,
  something which we will assume from now on.''
  \cite[paragraph after \texttt{eq:II\_CF1}]{Cirac2016MPDO_arXiv}
\end{quote}
This is a definitional convention applied throughout the rest of the paper; it
is not derived later. Combined with the strict modulus ordering used by the BNT
arrangement, it implies that the dominant block has $|\mu_1|=1$ and every
sub-dominant block has $|\mu_k|<1$.

The same paper proves the proportional Fundamental Theorem
\cite[Theorem~\texttt{thm1} and Corollary~\texttt{II\_cor2}]{Cirac2016MPDO_arXiv}.
Suppose two tensors $A$ and $B$ are presented in their basis-of-normal-tensor
expansions, with representative families $(A_j)_{j=1}^g$ and $(B_j)_{j=1}^g$
and BNT coefficients $\mu_{j,q_a}$ and $\nu_{j,q_b}$ for
$q_a=1,\ldots,r_{a,j}$ and $q_b=1,\ldots,r_{b,j}$. After Theorem~\texttt{thm1}
identifies the representatives, $B_j=e^{i\phi_j}Y_j A_j Y_j^{-1}$, the
trace-pair argument in the proof of Corollary~\texttt{II\_cor2} reduces to the
scalar identity
\begin{equation}
  \sum_{q=1}^{r_{a,j}}\mu_{j,q}^N
    =\sum_{q=1}^{r_{b,j}}\bigl(\nu_{j,q}e^{i\phi_j}\bigr)^N,
  \qquad\text{for all }j\text{ and all }N.
  \tag{\path{eq:II_cor2_power_sum}}
\end{equation}
The cited Lemma~\texttt{Lem:app\_simple} then concludes that
$r_{a,j}=r_{b,j}=:r_j$ and that the multisets of weights coincide,
$\mu_{j,q}=\nu_{j,q}e^{i\phi_j}$. The matching dimensions and the global gauge
formula follow.

The mathematical structure of this proof is that the coefficient information
is read off at the level of power sums, using all $N$ simultaneously as input
to a separation lemma for sums of $N$th powers. No convergence of any
$N$-indexed quantity is invoked.

\section{The formal canonical form}

The formal counterpart of the canonical form for the BNT route is the structure
\texttt{IsNormalCanonicalFormBNT}.\footnote{%
\path{TNLean/MPS/BNT/Construction.lean}, lines 181--189.}
Its mathematical content is: a strict-modulus ordering on the weight family
$\mu : \mathrm{Fin}\,r\to\mathbb C$, the nonvanishing $\mu_k\ne 0$ for every
$k$, and the per-block normality data inherited from the underlying canonical
form. The dominant weight is unconstrained in modulus.

Comparing this with the source definition, the field
$|\mu_1|=1$ (in source enumeration; index $0$ in the formal one) is missing.
The blueprint BNT chapter records the strict-modulus ordering in the
BNT-separated canonical-form definitions. Separately, the canonical-form
chapter records the decomposition $\mu_k=\eta_k|\mu_k|$ with $\eta_k$ on the
unit circle. Neither place imposes $|\mu_1|=1$.\footnote{%
The BNT-separated definitions are
\texttt{def:cf\_bnt} and \texttt{def:normal\_cf\_bnt} in
\path{blueprint/src/chapter/ch10_bnt.tex}, lines 70--100 and 119--141,
respectively. The phase--modulus decomposition is
\texttt{thm:mpv\_normalize} in
\path{blueprint/src/chapter/ch08_canonical.tex}, lines 125--139.}

The omission is mathematically meaningful, because the source definition is
used in two ways downstream: it fixes the dominant block on the unit circle,
which is what the spectral and irreducibility arguments need, and it forces
every sub-dominant block strictly inside the unit disk. The latter is the
conventional content used by the proportional MPV comparison below.

\section{The formal proportional comparison}

The formal route to the proportional Fundamental Theorem replaces the source
identity \texttt{eq:II\_cor2\_power\_sum} by a different mathematical object:
two coefficient sequences indexed by chain length, together with the assumption
that they converge to nonzero limits. Concretely, the structure
\texttt{ProportionalDecompositionData}\footnote{%
\path{TNLean/MPS/BNT/Construction.lean}, lines 512--537.}
packages, for each block index $j$, a function
$N\mapsto a_{N,j}\in\mathbb C$ and a designated limit $a_j^\infty\in\mathbb C$
satisfying
\[
  a_{N,j}\to a_j^\infty\text{ as }N\to\infty,
  \qquad a_j^\infty\ne 0,
\]
together with the analogous data on the $B$-side and a sequence of
proportionality scalars $c_N$ converging to a nonzero limit $c$. The
identification of the scalars is then derived from these convergence
hypotheses by a comparison of limits, rather than from the source
power-sum identity.

For the natural realization in which the $j$th source sector has one selected
representative, or one copy in the coefficient multiplicity notation
($r_j=1$), and the coefficient is the source weight itself, the formal choice
of array is
\[
  a_{N,j}=(\mu_j^A)^N.
\]
\footnote{The corresponding blueprint passage is the remark
\texttt{rem:cf\_coeff\_arrays} following \texttt{lem:ft\_proportional} in
\path{blueprint/src/chapter/ch11_fundamental_theorem_proof.tex},
lines 172--202.} With this
choice, the convergence requirement $a_{N,j}\to a_j^\infty\ne 0$ becomes a
condition on the powers $(\mu_j^A)^N$.

\section{The mathematical incompatibility}

Throughout this section, the comparison is restricted to the strict-weight,
one-representative-per-block specialization just described, namely
$a_{N,j}=(\mu_j^A)^N$ and $b_{N,k}=(\mu_k^B)^N$. It is not a claim about the
more general sector coefficient arrays
$c_{S,j}^{(N)}=\sum_q \mu_{j,q}^N$ that retain repeated copies within a BNT
sector.

Under the source normalization $|\mu_1^A|=1$ and $|\mu_j^A|<1$ for $j\ge 2$,
the powers $(\mu_j^A)^N$ behave as follows. For each sub-dominant block,
\[
  |(\mu_j^A)^N|=|\mu_j^A|^N\to 0\quad\text{as }N\to\infty,
\]
so the only candidate limit is $a_j^\infty=0$, contradicting the formal
hypothesis $a_j^\infty\ne 0$. For the dominant block, the modulus stays on
the unit circle, and the sequence $(\mu_1^A)^N$ converges only in the special
case $\mu_1^A=1$; nontrivial roots of unity are periodic rather than
convergent, and a generic phase has no limit.

The conclusion is a structural mismatch. With the source normalization in
force, the formal data $\texttt{ProportionalDecompositionData}$ cannot be
instantiated for a multi-block ($r>1$) family using $a_{N,j}=(\mu_j^A)^N$,
because the limit hypothesis fails on every sub-dominant block. Without the
source normalization, the formal hypothesis is sometimes satisfiable, but the
resulting object is no longer the canonical-form datum from
\cite{Cirac2016MPDO_arXiv}.

The paper does not face this difficulty, because it never asks for the
$N\to\infty$ limit of $(\mu_j^A)^N$. The proof reads off the multiset of
weights from the polynomial identity \texttt{eq:II\_cor2\_power\_sum} viewed
as an identity in $N$. With the BNT linear independence supplied by the
direct-sum span theorem, the same identity, restricted to a single
sufficiently large witness $N=N_0$, is enough to recover the multiset
equality and the $g_a=g_b$ statement.

\section{Cross-check with the blueprint}

The blueprint statements of \texttt{lem:ft\_proportional} and
\texttt{lem:ft\_proportional\_nt} carry the same convergence-and-nonzero-limit
data as the formal structure, with the symbols $a_j^\infty,b_k^\infty,c^\infty$
in place of \texttt{aLim}, \texttt{bLim}, \texttt{cLim}. The accompanying
remark \texttt{rem:cf\_coeff\_arrays} in the Fundamental Theorem proof chapter
acknowledges the strict-weight specialization $a_{N,j}=(\mu_j^A)^N$ and the
more general sector arrays $c_{S,j}^{(N)}=\sum_q\mu_{j,q}^N$, but does not flag
the restricted-specialization incompatibility described above. The remark
\texttt{rem:normalcf\_bnt\_gap} in the BNT chapter discusses a different point
(spectral versus algebraic normality of a single block), so it does not fill in
the missing comparison.

The blueprint BNT definitions are symmetric with the formal definition: the
strict-modulus ordering is recorded; the normalization $|\mu_1|=1$ is not.

\section{What a corrected formalization would say}

Two changes are needed to bring the formal development into agreement with the
source.

The first change is to record the source convention $|\mu_1|=1$. Once this
normalization is present, the implicit bounds $|\mu_k|\le 1$ on the remaining
blocks follow from the strict modulus ordering.\footnote{%
The formal field carrying this ordering is \texttt{mu\_strict\_anti} in
\texttt{IsNormalCanonicalFormBNT}; see
\path{TNLean/MPS/BNT/Construction.lean}, lines 184--185.} This should be a
field of \texttt{IsNormalCanonicalFormBNT} and of the corresponding blueprint
definition. It is a transcription of the source definition, not an additional
assumption.

The second change is to reformulate the proportional comparison so that it
extracts coefficients from a single sufficiently large witness length $N_0$.
The mathematical input is the BNT linear independence at length $N_0$, which
is supplied by the direct-sum span theorem on a basis of normal tensors. The
output is, blockwise,
\[
  (\mu_j^A)^{N_0}=c^{N_0}\,(\mu_j^B)^{N_0},\qquad j=1,\ldots,g,
\]
which fixes $\mu_j^A$ in terms of $\mu_j^B$ and $c$ up to an $N_0$th root of
unity. The phase ambiguity is then absorbed into the per-block phase $\phi_j$
already present in the source statement
$B_j=e^{i\phi_j}Y_j A_j Y_j^{-1}$. No convergence statement is needed; in
particular no nonzero limit hypothesis on $(\mu_j^A)^N$ is required.

The reformulated comparison is what the source actually proves
\cite[proof of Corollary~\texttt{II\_cor2}]{Cirac2016MPDO_arXiv}, transposed
into the BNT-direct-sum-span input that the formal route already isolates
elsewhere in the development.

\section{Conclusion}

The verdict is that the canonical-form definition and the proportional MPV
comparison currently differ from
\cite{Cirac2016MPDO_arXiv} in two coupled ways. The canonical-form definition
omits the source convention $|\mu_1|=1$. The proportional comparison replaces
the source's fixed-length power-sum identity by a limit hypothesis on
$(\mu_j^A)^N$ that fails on the sub-dominant blocks under the very
normalization the source imposes.

Each change can be made independently of the other, but they are most
naturally addressed together: adding the missing normalization makes the
mismatch between the two proof models visible at the level of types, since the
limit model becomes uninstantiable on the source's intended class of inputs.
Once both changes are in place, the formal proportional Fundamental Theorem
can be aligned with the source argument without weakening the statement and
without introducing convergence hypotheses absent from
\cite{Cirac2016MPDO_arXiv}.

The corresponding tracked items are
\href{https://github.com/LionSR/TNLean/issues/1554}{issue~\#1554} for the
canonical-form normalization and
\href{https://github.com/LionSR/TNLean/issues/1555}{issue~\#1555} for the
fixed-length reformulation of the proportional comparison.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
