\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Per-block RFP Isometry Form and the Source Isometry Condition in CPSV16}
\author{}
\date{2026-06-14}

\begin{document}
\maketitle

\section{Source statement}

The pure-state RFP characterization in Section~3 of
\cite{Cirac2016MPDO_arXiv} writes a canonical-form RFP tensor as
\begin{align}
  A^i
    =
  \bigoplus_{j=1}^g\bigoplus_{q=1}^{r_j}
  \mu_{j,q}X_{j,q}\Lambda_jU_j^iX_{j,q}^{-1},
  \tag{\path{III_CFI_RFP}}
\end{align}
where the matrices $\Lambda_j$ are diagonal, positive, and trace-normalized.
The same theorem requires a joint isometry condition
\begin{align}
  \sum_{i=1}^s
  (U_j^i)_{\alpha,\beta}\,
  \overline{(U_{j'}^i)_{\alpha',\beta'}}
  =
  \delta_{j,j'}\delta_{\alpha,\alpha'}\delta_{\beta,\beta'}.
  \tag{\path{eq:III_isometry}}
\end{align}
The diagonal case $j=j'$ is a full pair-index orthonormality condition in the
two virtual indices $(\alpha,\beta)$.  The factor $\delta_{j,j'}$ adds a
cross-block orthogonality condition: for $j\ne j'$, the physical-index
isometries associated with the two normal-tensor blocks are mutually
orthogonal.  Corollary~III.cor3 invokes this same isometry condition for the
BNT elements.\footnote{Local source:
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:543--554} and
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:583--589}.}

\section{Formal boundary}

The current formal theorem applies the single-block isometry form separately to
each block of an indexed family.\footnote{Formal declaration:
\leanid{MPSTensor.rfp_nt_structural_full_blocks} in
\doclink{TNLean/MPS/RFP/StructuralFull}.  Introduced in \ghpr{2785}.}  For each
block $k$ it gives
\begin{align}
  A_k^i
    &=
  X_k\Lambda_kU_k^iX_k^{-1},\\
  \sum_i (U_k^i)^\dagger U_k^i
    &= I,\\
  \sum_i
  \overline{(U_k^i)_{\alpha,\beta}}\,
  (U_k^i)_{\alpha',\beta'}
    &=
  D_k^{-1}\delta_{\alpha,\alpha'}\delta_{\beta,\beta'}.
\end{align}
The second equation is the contracted identity
\begin{align}
  \sum_{i,\alpha}
  \overline{(U_k^i)_{\alpha,\beta}}\,
  (U_k^i)_{\alpha,\beta'}
  =
  \delta_{\beta,\beta'}.
\end{align}
The third equation records the diagonal pair-index orthonormality in the
normalization used by the formal theorem.  Compared with the source convention,
the entries of $U_k^i$ carry the normalized matrix-unit factor $D_k^{-1/2}$;
therefore the compatible pair-index equation is
\begin{align}
  \sum_i
  \overline{(U_k^i)_{\alpha,\beta}}\,
  (U_k^i)_{\alpha',\beta'}
  =
  D_k^{-1}\delta_{\alpha,\alpha'}\delta_{\beta,\beta'}.
\end{align}
The remaining normalization comparison with the source equation
\eqref{eq:III_isometry} is therefore explicit rather than hidden.

A subsequent comparison theorem rescales the same single-block decomposition by
$U_k^i\mapsto D_k^{1/2}U_k^i$ and
$\Lambda_k\mapsto D_k^{-1/2}\Lambda_k$.\footnote{Formal declarations:
\leanid{MPSTensor.rfp_nt_structural_full_unit_pair} and
\leanid{MPSTensor.rfp_nt_structural_full_blocks_unit_pair} in
\doclink{TNLean/MPS/RFP/StructuralFull}.}  This leaves the product
$\Lambda_k U_k^i$ unchanged and gives the diagonal pair-index equation in the
source unit convention
\begin{align}
  \sum_i
  \overline{(U_k^i)_{\alpha,\beta}}\,
  (U_k^i)_{\alpha',\beta'}
  =
  \delta_{\alpha,\alpha'}\delta_{\beta,\beta'}.
\end{align}
Thus the scalar normalization mismatch is no longer hidden in the formal
statement: both conventions are available, and their comparison is explicit.
The Appendix~B structural datum used in the conditional parent-Hamiltonian
theorem chooses the unit pair-index convention, so the remaining coefficient
problem starts from the source normalization.
Appendix~B first gives the coefficient form of the basic-vector expression
\begin{align}
  C_L(\sigma)
  =
  \sum_{\alpha_0,\ldots,\alpha_{L-1}}
  \prod_{t=0}^{L-1}
  \Lambda_{\alpha_t}
  (U^{\sigma_t})_{\alpha_t,\alpha_{t+1}},
  \qquad \alpha_L=\alpha_0,
\end{align}
which is the coefficient form of
\(|V^{(L)}(A_j)\rangle=U^{\otimes L}|\varphi_j\rangle^{\otimes L}\), with
\(\varphi_j\) shared between \(b_t\) and \(a_{t+1}\).  The disjoint adjacent
physical-pair condition used by the current conditional parent-Hamiltonian
theorem is therefore a separate statement, not the source expression itself.
The formal development proves this cyclic coefficient formula for the original
tensor after the Appendix~B change of basis, for every nonempty chain length
and a chosen structural datum:
\[
  V^{(L)}(A)_\sigma=C_L(\sigma).
\]
This is the declaration
\leanid{MPSTensor.AppendixBStructuralData.mpv_eq_cyclicVirtualPairState}.  It
does not prove the disjoint adjacent physical-pair condition.
The pair-index isometry above is an orthogonality equation after summing over
the physical letter \(i\).  It is therefore not, by itself, a coefficient
identity for a fixed physical word \(\sigma\).  Any proof of a fixed-word
adjacent-pair factorization must use the source basic-vector expression
\(U^{\otimes L}\varphi_j^{\otimes L}\), or an equivalent construction of the
corresponding local projectors, rather than applying the summed isometry as if
the physical labels were also summed in \(C_L(\sigma)\).
This obstruction is genuine, not merely a missing formal proof.  For example,
take two virtual indices and physical labels \(i=(a,b)\), with
\[
  (U^{(a,b)})_{x,y}=\delta_{x,a}\delta_{y,b},
  \qquad
  A^{(a,b)}=\lambda_a E_{a,b},
\]
where \(\lambda_0,\lambda_1>0\).  The unit pair-index isometry holds.  For the
length-four word
\[
  \sigma=((0,1),(1,0),(1,0),(0,1)),
\]
the cyclic coefficient \(C_4(\sigma)\) vanishes, because the middle virtual
matching condition fails.  The disjoint adjacent-pair product is instead
\[
  (\lambda_0\lambda_1)(\lambda_1\lambda_0)
  =
  \lambda_0^2\lambda_1^2\ne 0.
\]
Thus the adjacent physical-pair coefficient identity is not a consequence of
the unit pair-index isometry alone.

There is still no conclusion relating $U_k$ and $U_\ell$ for $k\ne \ell$.  Thus
the formal theorems record only a separate isometry condition for each block;
they do not record the cross-block equations
\begin{align}
  \sum_i
  (U_k^i)_{\alpha,\beta}\,
  \overline{(U_\ell^i)_{\alpha',\beta'}}
  =
  0
  \qquad (k\ne \ell),
\end{align}
which are part of the source condition \path{eq:III_isometry}.

\section{Square-root diagonal (local correction)}

The source's displayed normal-tensor form (\path{eq_III_NT_RFP}) is written
\begin{align}
  A^i = X\,\Lambda\,U^i\,X^{-1},
  \tag{\path{eq_III_NT_RFP}}
\end{align}
with the diagonal $\Lambda$ itself, together with the unit isometry
condition\footnote{Local source:
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:1278} for the displayed equation and
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:1281-1283} for the unit isometry.}
\begin{align}
  \sum_i (U^i)_{\alpha,\beta}\,\overline{(U^i)_{\alpha',\beta'}}
    = \delta_{\alpha,\alpha'}\delta_{\beta,\beta'}.
\end{align}
This literal pairing is internally inconsistent.  The same proof constructs the
reference tensor\footnote{Local source:
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:1300}.}
\begin{align}
  \widehat A^{(\alpha',\beta')}_{\alpha,\beta}
    = \delta_{\alpha,\alpha'}\delta_{\beta,\beta'}\sqrt{\Lambda_\alpha},
\end{align}
which is left-canonical and has fixed point $\diag(\Lambda)$ with
$\tr(\Lambda)=1$.  Writing $\widehat A = \Lambda\,U$ would force
$U^{(\alpha',\beta')}_{\alpha,\beta}
  = \delta_{\alpha,\alpha'}\delta_{\beta,\beta'}/\Lambda_\alpha$, whose pair-index
sum is
\begin{align}
  \sum_i (U^i)_{\alpha,\beta}\,\overline{(U^i)_{\alpha',\beta'}}
    = \frac{\delta_{\alpha,\alpha'}\delta_{\beta,\beta'}}{\Lambda_\alpha}
    \ne \delta_{\alpha,\alpha'}\delta_{\beta,\beta'}
    \qquad(\Lambda_\alpha\ne 1),
\end{align}
so the residual after factoring $\Lambda$ is \emph{not} a unit isometry.  Writing
instead $\widehat A = \sqrt\Lambda\,U$ gives
$U^{(\alpha',\beta')}_{\alpha,\beta}=\delta_{\alpha,\alpha'}\delta_{\beta,\beta'}$,
whose pair-index sum is
$\delta_{\alpha,\alpha'}\delta_{\beta,\beta'}$, a genuine unit isometry.

The formalization therefore states the decomposition with the square-root
diagonal,
\begin{align}
  A^i = X\,\sqrt\Lambda\,U^i\,X^{-1},
\end{align}
keeping both the source trace-normalization $\tr(\Lambda)=1$ on $\Lambda$ and the
source unit isometry condition on $U$ simultaneously.  This $\Lambda\to\sqrt\Lambda$
change is a \emph{local correction} of the source display: it is mathematically
correct as stated, and the source itself uses $\sqrt{\Lambda_\alpha}$ for the
reference tensor at line~1300, so the displayed $\Lambda$ at line~1278 is a
typographic simplification of the dressing.

\section{Trace normalization (discharged)}

The source trace-normalization $\tr(\Lambda)=1$ is now formalized for a single
normal-tensor block.  A named predicate records the isometry canonical
form,\footnote{Formal declaration: \leanid{MPSTensor.IsIsometryCanonicalForm}
in \doclink{TNLean/MPS/RFP/StructuralFull}.} and a theorem proves it for a
normal canonical-form-II RFP tensor.\footnote{Formal declarations:
\leanid{MPSTensor.isIsometryCanonicalForm_of_rfp_nt} and the per-block
\leanid{MPSTensor.isIsometryCanonicalForm_of_rfp_nt_blocks}.}  The predicate
asserts the existence of an invertible $X$, a positive diagonal $\Lambda$ with
$\sum_\alpha\Lambda_\alpha=1$, and a unit pair-index isometry $U$ with
\begin{align}
  A^i = X\,\sqrt\Lambda\,U^i\,X^{-1},\qquad
  \sum_i \overline{(U^i)_{\alpha,\beta}}\,(U^i)_{\alpha',\beta'}
    = \delta_{\alpha,\alpha'}\delta_{\beta,\beta'}.
\end{align}
The square root on the diagonal matches the source reference tensor (the
square-root correction is documented separately in the previous section); the
trace
condition $\tr(\Lambda)=1$ is imposed on $\Lambda$ itself, equivalently on the
normalized diagonal fixed point $\rho=\diag(\Lambda)$,
$\Lambda_\alpha=\rho_{\alpha,\alpha}/\tr\rho$.  The identity
$\sum_\alpha\Lambda_\alpha=1$ is the trace identity
$\sum_\alpha\rho_{\alpha,\alpha}=\tr\rho$ for the diagonal fixed point.

This discharges the trace-normalization scope item.  The remaining open part of
the source isometry condition is the cross-block $\delta_{j,j'}$ orthogonality
between distinct normal-tensor blocks, treated below.

\section{Elimination plan}

A source-faithful version of Corollary~III.cor3 should be stated for a
BNT-family or whole-tensor canonical-form datum, not merely for independent
blocks.  Its conclusion should first expose the full pair-index isometry for a
single block, with the normalization compared to the source convention, and
then include a joint family $U$ satisfying
\begin{align}
  \sum_i
  (U_k^i)_{\alpha,\beta}\,
  \overline{(U_\ell^i)_{\alpha',\beta'}}
  =
  \delta_{k,\ell}\delta_{\alpha,\alpha'}\delta_{\beta,\beta'}.
\end{align}
Such a statement must account for the common physical space into which all
blocks embed; in particular, the joint isometry condition has the dimensional
content that the block-pair spaces fit orthogonally inside the physical space.

\section{Conclusion}

The current theorem is a scope restriction.  It is a useful per-block
consequence of the single-block RFP isometry form, but it is not the full
isometry assertion of Corollary~III.cor3.  The formal development now exposes
two normalization conventions.  In the $D_k^{-1/2}$ convention the residual
tensor satisfies the contracted per-block identity
\[
  \sum_i (U_k^i)^\dagger U_k^i=I
\]
and its pair-index equation has right-hand side
$D_k^{-1}\delta_{\alpha,\alpha'}\delta_{\beta,\beta'}$.  In the rescaled
unit pair-index convention the residual tensor satisfies
\[
  \sum_i
  \overline{(U_k^i)_{\alpha,\beta}}\,
  (U_k^i)_{\alpha',\beta'}
  =
  \delta_{\alpha,\alpha'}\delta_{\beta,\beta'},
\]
while the contracted identity is instead
$\sum_i (U_k^i)^\dagger U_k^i=D_k I$.  Thus the scalar normalization comparison
is explicit.  The source trace-normalization $\tr(\Lambda)=1$ is now formalized
for a single block by the named predicate
\leanid{MPSTensor.IsIsometryCanonicalForm} and the theorem
\leanid{MPSTensor.isIsometryCanonicalForm_of_rfp_nt}, with the diagonal weight
$\Lambda_\alpha=\rho_{\alpha,\alpha}/\tr\rho$ and the square-root dressing
$\sqrt\Lambda$ of the unit isometry.  The remaining open part of the source
isometry condition is the cross-block $\delta_{j,j'}$ orthogonality equations
between distinct normal-tensor blocks.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
