\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Purification RFP in Cirac--Perez-Garcia--Schuch--Verstraete 2017}
\author{}
\date{2026-06-05}

\begin{document}
\maketitle

\section{Source statement}

Cirac--P\'erez-Garcia--Schuch--Verstraete define a purification
renormalization fixed point for an MPDO tensor $M$ through a purifying MPS
tensor $A$:
\[
  \rho^{(N)}_p(M)
  =
  \tr_a\bigl(
    |\Psi^{(N)}(A)\rangle\langle\Psi^{(N)}(A)|
  \bigr).
\]
The source definition says that $M$ is a PRFP when it admits such a
representation with $A$ an RFP.\footnote{Source location:
\path{Papers/1606.00608/MPDO-22-12-17-2.tex}:750--763, Definition
\path{def:Puri-RFP}.}

The following theorem in the source states that, for tensors $M$ satisfying the
purification equation above, the three conditions
\[
  M\text{ is a PRFP},\qquad
  M\text{ has ZCL},\qquad
  \rho^{(N)}(M)\text{ has the displayed PRFP decomposition}
\]
are equivalent.\footnote{Source location:
\path{Papers/1606.00608/MPDO-22-12-17-2.tex}:775--786.}
The later warning that this definition is too weak concerns saturation of the
area law, not the implication from PRFP to zero correlation length.

\section{Deleted formalization}

The deleted predicate \leanid{MPOTensor.IsPRFP} did not formalize the source
definition. It did require a local LPDO witness and the condition
\[
  \leanid{MPSTensor.IsRFP}(\leanid{purifyingMPSTensor}(A)),
\]
so the pure-state RFP requirement itself was present. The missing source data
was the stronger purification-RFP structure around
\path{eq:MPDO-Puri-1}: after tracing the ancillary index, the source obtains a
trace-preserving completely positive map on the spin degrees of freedom and the
decomposition displayed in \path{PuriRFP}--\path{PuriRFP-BNT}. The purifying
tensor is also described by the canonical RFP form \path{III_CFI_RFP}, but this
is not the condition violated by the deleted scalar witness; in bond
dimension~$1$ the witness already has that form. The deleted predicate kept
only the local purifying family together with generic transfer-map idempotence
of the doubled MPS tensor. It did not encode the global purification equation or
the trace-preserving post-ancilla map.

With this weaker hypothesis the development proved a counterexample to PRFP
implying ZCL. That conclusion contradicted the source theorem above, so it was
a counterexample to the local Lean predicate, not to Definition
\path{def:Puri-RFP}.

For this reason the file \path{TNLean/MPS/MPDO/PRFP.lean}, the corresponding
blueprint section, and the counterexample registry entry were removed. The
coverage audit now records the source PRFP definition as open, not as
formalized.\footnote{This note records the divergence removed in
\href{https://github.com/LionSR/TNLean/pull/2386}{PR \#2386}. The faithful
replacement is tracked by
\href{https://github.com/LionSR/TNLean/issues/2383}{issue \#2383} and the
MPDO coverage tracker
\href{https://github.com/LionSR/TNLean/issues/2380}{issue \#2380}.}

\section{Faithful replacement}

A faithful formalization should state the purification RFP definition with the
positive-length global purification equation \path{eq:MPDO-Puri-1} and require
the purifying spin--ancilla tensor to be a pure-state RFP. The trace over the
ancillary index, discussed after Definition \path{def:Puri-RFP}, is a separate
one-site tpCPM; it is not the two-map mixed-state RFP condition
\path{RFPMixedTS}. The definitions are now present in
\path{TNLean/MPS/MPDO/PRFP.lean}: the global equation is
\leanid{MPOTensor.HasGlobalPurificationEquation}, the pure-state witness is
\leanid{MPOTensor.HasPurificationRFPWitness}, and the source predicate is
\leanid{MPOTensor.IsPRFP}. The ancillary-trace map is
\leanid{MPOTensor.ancillaryTraceMap}; its Kraus-form trace-preserving
complete positivity is
\leanid{MPOTensor.ancillaryTraceMap_isKrausCPTP}, and the corresponding
spin-reduction condition is
\leanid{MPOTensor.HasTracePreservingSpinReduction}. Only after the remaining
theorem is proved should the theorem
\[
  \operatorname{PRFP}(M) \Longleftrightarrow \operatorname{ZCL}(M)
\]
be attached to the source theorem.

No theorem should assert that source PRFP fails to imply ZCL. Any future
counterexample must instead be phrased for a separate local-purification
predicate and must not be cited as Definition \path{def:Puri-RFP}.

\section{Status}

A separate local predicate
\leanid{MPOTensor.IsLocalPurificationRFP} is now in
\path{TNLean/MPS/MPDO/LocalPurificationRFP.lean}. It records the tensor-level
local purification structure
\[
  M^{ij}
  =
  \left(\sum_k A^{(i,k)} \otimes \overline{A^{(j,k)}}\right)_{e}
\]
together with the requirement that the purifying tensor $A$, viewed as an MPS
tensor on the combined spin--ancilla index, is a pure-state renormalization
fixed point (\leanid{MPSTensor.IsRFP}). This predicate pins the local tensor of
$M$ and proves the useful consequence
\[
  \leanid{MPOTensor.IsLocalPurificationRFP.isMPDO}:
  \leanid{MPOTensor.IsLocalPurificationRFP}(M)
  \Longrightarrow \leanid{MPOTensor.IsMPDO}(M),
\]
via \leanid{MPOTensor.IsLPDO.isMPDO}.

This local predicate is not Definition \path{def:Puri-RFP}. The source-level
positive-length global purification equation and the trace-preserving structure
appearing after the ancillary trace are recorded separately in
\path{TNLean/MPS/MPDO/PRFP.lean}. The equivalence
\[
  \operatorname{PRFP}(M) \Longleftrightarrow \operatorname{ZCL}(M)
\]
therefore remains open.

\section{Forward implication on the physical-trace transfer}

The forward implication of the source theorem (lines 775--786), from a
purification renormalization fixed point to zero correlation length, is
formalized on the source physical-trace transfer object
$\mathcal T_M=\sum_i M^{ii}$ (Definition~4.2, lines 735--739). The theorem
\leanid{MPOTensor.physTraceTransfer_sq_of_isLocalPurificationRFP} proves that
the local purification renormalization fixed point condition makes
$\mathcal T_M$ idempotent, $\mathcal T_M\,\mathcal T_M=\mathcal T_M$, with no
extra hypothesis: the transfer matrix of the pure-state purifying tensor is
idempotent, and closing the physical legs of $M$ transports this idempotence
through the bond identification and the Kronecker factor swap.

The source-faithful conclusion that $M$ has source zero correlation length,
\leanid{MPOTensor.isSourceZCL_of_isLocalPurificationRFP}, carries the extra
hypothesis $\mathcal T_M\neq 0$. This records a normalization restriction: the
unconditional implication is false. The zero tensor at $D=D'=d_K=1$ with
$A^{(i,0)}=0$ satisfies the local purification condition, since its purifying
tensor is identically zero and $0\circ 0=0$ makes that tensor a pure-state
renormalization fixed point, yet $\mathcal T_M=0$, so the conclusion fails. The
hypothesis $\mathcal T_M\neq 0$ reflects the source purification renormalization
fixed point being a normalized pure-state fixed point with leading eigenvalue
$1$, a normalization that \leanid{MPOTensor.IsLocalPurificationRFP} drops.
Eliminating the hypothesis requires the normalized purification condition, which
remains open together with the full equivalence.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
