\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\input{./command}

\title{Global and Per-Sector Unit-Weight Witnesses in the CPSV16 BNT Argument}
\author{}
\date{2026-05-30}

\begin{document}
\maketitle

\section{The source normalization}

Cirac--Perez-Garcia--Schuch--Verstraete 2016 normalizes the raw coefficients in
the basis-of-normal-tensors (BNT) expansion by a single global condition.
Section II.C, line 246 states the condition for the one-index canonical-form
coefficients \(\mu_k\).  The BNT display at lines 287--289 then refines this
index into a sector index and a copy index.  In the resulting two-layer
notation
\[
  A^i =
  \bigoplus_{j=1}^{g}\bigoplus_{q=1}^{r_j}
  \mu_{j,q} X_{j,q} A_j^i X_{j,q}^{-1},
\]
the source assumes
\[
  |\mu_{j,q}| \le 1 \quad \text{for all } (j,q),
  \qquad
  |\mu_{j_\ast,q_\ast}| = 1 \quad \text{for some } (j_\ast,q_\ast).
\]
This is the same normalization, expressed after the two-layer BNT refinement.

Thus the paper-level canonical form does not state that every BNT sector
\(j\) contains a copy \(q\) of modulus one.  A sector may contain only strictly
subunit copy weights, while some other sector carries the global unit witness.

\section{The formal boundary}

The formal predicate \leanid{MPSTensor.IsBNTCanonicalForm} records the source
normalization faithfully: it has a bound on all copy weights and a single
global unit-weight witness.  It does not contain the per-sector condition
\[
  \forall j,\ \exists q,\ |\mu_{j,q}| = 1.
\]

Several full-basis matching and global-gauge theorems formerly assumed this
per-sector condition explicitly.  The condition was the price of the
\emph{asymptotic} matching route, which applied the Cesaro non-decay lemma
sector by sector in the CPSV16 Appendix MPV line-1182 matching argument: a
strictly subunit sector coefficient \(c^{(j)}_N=\sum_q \mu_{j,q}^N\) tends to
zero, so it is invisible to a projection onto a single sector unless that sector
carries a unit-modulus copy.

This restriction has since been removed.  The exact linear-independence matcher
replaces the asymptotic projection: at a fixed sufficiently large length, BNT
linear independence detects whether a sector coefficient vanishes, and
\leanid{MPSTensor.IsBNTCanonicalForm.coeff_not_eventually_zero} shows that
\(c^{(j)}_N\) is nonzero infinitely often for \emph{any} nonzero copy weights,
with no modulus assumption.  The matching and global-gauge declarations
therefore carry exactly the source hypothesis set (BNT canonical form, with
proportional or equal MPVs) and no longer impose the per-sector unit-weight
condition.

\section{Equal-modulus declaration path}

The current formal assembly is organized through
\path{TNLean/MPS/FundamentalTheorem/SectorBNT/}.  There is no separate
\path{TNLean/MPS/FundamentalTheorem/Multi/} assembly directory on the current
main branch.

The equal-modulus question is logically separate from the per-sector
unit-witness restriction above.  The current SectorBNT coefficient comparison
works with raw sector coefficients
\[
  c_N^{(j)} = \sum_q \mu_{j,q}^N
\]
and the copy weights are recovered from these coefficients by the finite
power-sum argument.  Equal-modulus copies such as the two copies in
\(C \oplus (-C)\), and unequal-modulus copies such as \(C \oplus (1/2)C\),
are both represented by the raw two-layer SectorBNT formulation.

The declaration path is as follows.
\begin{itemize}
  \item \leanid{MPSTensor.IsBNTCanonicalForm} records the raw copy weights
        \(\mu_{j,q}\), their coefficient sums, the bound
        \(|\mu_{j,q}| \le 1\), and the single global unit witness.  It imposes
        neither a strict ordering of the moduli nor an equal-modulus layer.
  \item Sectorwise non-decay is supplied, with no modulus assumption, by
        \begin{center}
        {\small\leanid{MPSTensor.IsBNTCanonicalForm.coeff_not_eventually_zero}},
        \end{center}
        which shows the raw coefficient sum is nonzero infinitely often for any
        nonzero copy weights.
  \item Eventual equality of the sector coefficient sequences is converted into
        equality of copy weights by
        \leanid{MPSTensor.matched_sector_weight_multiset_eq} and
        \leanid{MPSTensor.matched_sector_weight_equiv}, using the finite
        power-sum recovery theorem
        \begin{center}
        {\small\leanid{Matrix.sum_pow_eq_implies_card_eq_and_multiset_eq_of_le_max_card}}.
        \end{center}
  \item \leanid{MPSTensor.SectorBNTCopyWeightMatching.of_coeff_identity}
        packages this recovery into the copy-weight matching used by the
        proportional global-gauge theorem.
  \item In the equal-MPV path, the declarations
        \begin{center}
        {\small\leanid{MPSTensor.ft_sector_bnt_equal_sector_data}}\\
        {\small\leanid{MPSTensor.ft_sector_bnt_equal_global_gauge}}\\
        {\small\leanid{MPSTensor.ft_sector_bnt_equal_mps_gaugeEquiv_literal}}
        \end{center}
        use the same coefficient-to-copy-weight step.
\end{itemize}

Thus equal-modulus examples such as \(C \oplus (-C)\) are handled by the raw
coefficient comparison and finite power-sum recovery after sector matching.  No
theorem in this SectorBNT declaration path assumes a strict ordering of the
copy-weight moduli.

The extra hypothesis that once distinguished these statements from the source
was not strictness of moduli but the per-sector existence of a unit-modulus copy
weight, which is stronger than the single global unit witness stated in CPSV16
line 246.  As described in the next section, that hypothesis has been eliminated
by the exact linear-independence matcher, so only the single global unit witness
remains.

\section{How the restriction was eliminated}

The restriction is removed by the second of the two strategies above, in its
exact form.  Rather than projecting asymptotically onto a single sector, the
matching works at a fixed sufficiently large length.  The \(P\)-sectors are
split into those whose overlap decays against every \(Q\)-sector and the rest;
each non-decaying sector is rewritten as an exact scalar multiple of its matched
\(Q\)-sector, and the residual is a single linear combination over the combined
BNT family, which is eventually linearly independent.  Linear independence at a
fixed length forces every coefficient to vanish, so any putatively unmatched
sector would have \(c^{(j_0)}_N=0\) for all large \(N\), contradicting
\leanid{MPSTensor.IsBNTCanonicalForm.coeff_not_eventually_zero}.  No power of a
copy weight is required to have modulus one, and no Cesaro projection or
coefficient limit is taken.  The proportional case is identical except that the
proportionality scalar \(c_N\) rescales only the aggregate \(Q\)-side
coefficients, never the isolated target \(P\)-coefficient.

The matcher is \leanid{MPSTensor.exists_block_match_exact} for the equal case
and \leanid{MPSTensor.exists_block_match_exact_of_eventuallyProportional} for the
proportional case.  Consequently the BNT sector-matching declarations are no
longer scope-restricted: they carry exactly the source hypothesis set and may be
cited as the CPSV16 theorem and corollary on the BNT canonical-form surface.

\section{Affected declarations}

The following declarations \emph{formerly} carried the additional per-sector
unit-modulus hypothesis and no longer do:
\begin{itemize}
  \item \leanid{MPSTensor.forall_k_exists_j_nondecaying_overlap_of_sameMPV},
  \item \leanid{MPSTensor.bijective_match_of_sameMPV},
  \item \leanid{MPSTensor.ft_sector_bnt_proportional_sector_match_witnesses},
  \item \leanid{MPSTensor.ft_sector_bnt_proportional_global_gauge_of_copy_weight_matching},
  \item \leanid{MPSTensor.ft_sector_bnt_proportional_global_gauge_of_coeff_identity},
  \item \leanid{MPSTensor.ft_sector_bnt_equal_global_gauge},
  \item \leanid{MPSTensor.ft_sector_bnt_equal_mps_gaugeEquiv_literal}.
\end{itemize}

The faithful structural normalization remains the global unit-weight field
\path{MPSTensor.IsBNTCanonicalForm.weight_unit_exists} in
\path{TNLean/MPS/FundamentalTheorem/SectorBNT/Basic.lean}; the source assumes
only this single global witness, which the exact matcher is now content with.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
