\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Rate-Quantified BNT Cross-Overlap Decay\\
       for the Paper-Faithful BNT Surface}
\author{}
\date{2026-05-13}

\begin{document}
\maketitle

\section{Notation}

Fix a sector decomposition $P$ with basis blocks
$\{P.\mathtt{basis}\,j\}_{j=0}^{g-1}$ and a spectral level
$\lambda_j\in\mathbb C\setminus\{0\}$ with
$\|\lambda_0\|>\|\lambda_1\|>\cdots>\|\lambda_{g-1}\|$ and
$\|\lambda_0\|=1$. The bilinear cross-overlap between basis blocks
$j$ and $k$ at length $N$ is
\begin{align}
    O_{j,k}(N) := 
    \mathtt{mpvOverlap}\bigl(P.\mathtt{basis}\,j,
                             P.\mathtt{basis}\,k\bigr)(N)
    = \sum_{\sigma}\mathtt{mpv}(P.\mathtt{basis}\,j)(\sigma)
                     \overline{\mathtt{mpv}(P.\mathtt{basis}\,k)(\sigma)},
\end{align}
written \leanid{mpvOverlap} in the formalization.\footnote{%
    Declaration \leanid{MPSTensor.mpvOverlap} in
    \doclink{TNLean/MPS/Overlap/Basic}.}

For every ordered pair $j\ne k$ we are interested in a geometric
bound
\begin{align}
    \|O_{j,k}(N)\| \le C_{j,k} \eta_{j,k}^N
    \qquad(\forall N\in\mathbb N),
    \label{eq:rate}
\end{align}
with $C_{j,k}\ge 0$ and a rate $\eta_{j,k}\in[0,1)$.

\section{Cited assertion}

The Step~1 argument at
\cite[lines~1170--1192]{Cirac2016MPDO_arXiv} relies on
$O_{j,k_0}(N)\to 0$ for $j\ne k_0$ to discard
cross-overlap terms when projecting an assembled BNT identity onto a
distinguished basis block $V^{(N)}(P.\mathtt{basis}\,k_0)$. The
qualitative limit is implicit in the BNT structure: distinct normal
blocks satisfy
\begin{align}
    O_{j,k}(N)\xrightarrow[N\to\infty]{} 0
    \qquad(j\ne k).
    \label{eq:qual}
\end{align}
The argument as written does not quantify the rate of decay in
\eqref{eq:qual}.

Definition~4.3 of~\cite{Cirac2021Matrix} packages
\eqref{eq:qual} indirectly under \emph{eventual linear independence} of
the basis MPVs and does not state a rate. The formal counterpart is
the predicate
\leanid{MPSTensor.HasBNTSectorData}\footnote{%
    \doclink{TNLean/MPS/FundamentalTheorem/SectorDecomposition} at
    \leanid{HasBNTSectorData}.}
in the formalization, and \eqref{eq:qual} is provided pointwise by
\leanid{MPSTensor.mpvOverlap\_tendsto\_zero\_of\_not\_gaugePhaseEquiv\_cast\_left\_of\_irreducible\_TP}.\footnote{%
    \doclink{TNLean/MPS/Overlap/CastDecay}.}
Neither source equips $\eta_{j,k}$ with a quantitative relation
to the spectral weight ratios.

\section{The rate condition}

The rate condition that would discharge the non-dominant projection argument
takes a sector decomposition $P$ and a spectral level
$\lambda:\mathrm{Fin}\,P.\mathtt{basisCount}\to\mathbb C$. It asserts the
existence of non-negative families $C,\eta$ such that for every $j\ne k$,
\begin{align}
    0\le C_{j,k},\quad 0\le\eta_{j,k}<1,
        \qquad
    \eta_{j,k}\cdot\|\lambda_j/\lambda_k\|<1,
    \label{eq:hyp}
\end{align}
together with the geometric bound~\eqref{eq:rate} for all $N$.

This condition is not currently a formal declaration.  It arose during the
retired two-layer sector-refinement route, where a common spectral level
$\lambda_j$ was separated from within-sector phases.  The present
paper-faithful route keeps the raw CPSV weights $\mu_{j,q}$ in the core BNT
predicate and does not impose such a factorization.

The surviving formal statements used in the current route are the qualitative
and coefficient estimates in
\doclink{TNLean/MPS/FundamentalTheorem/PaperBNT/Api}.  In particular,
\leanid{IsBNTCanonicalForm.cross\_overlap\_basis\_tendsto\_zero} gives
the limit~\eqref{eq:qual},
\leanid{IsBNTCanonicalForm.combined\_family\_eventually\_li} gives the
combined-family linear-independence conclusion from vanishing cross-overlaps,
\leanid{IsBNTCanonicalForm.norm\_coeff\_le\_copies} bounds the raw sector
coefficient under the CPSV line-246 normalization, and
\leanid{IsBNTCanonicalForm.coeff\_not\_tendsto\_zero\_at\_block} gives the
non-vanishing conclusion for a block carrying a unit-modulus copy.

\section{Where the rate condition would be used}

The retired two-layer per-block-projection lemmas carried an abstract
non-cancellation hypothesis ruling out asymptotic cancellation of the assembled
overlap against the chosen projection block~$k_0$.  The obstruction analysis in
\path{audits/2026-05-13_cpsv16_ft_path_beta_pr2_5_design.md} shows that
for a non-dominant $k_0$ the discharge requires, after isolating
$V^{(N)}(P.\mathtt{basis}\,k_0)$ on the right and renormalising by
$\lambda_{k_0}^N$, that every $j\ne k_0$ contribute a term whose modulus
is bounded by
\begin{align}
    \|O_{j,k_0}(N)\|\cdot\|\lambda_j/\lambda_{k_0}\|^N
    \le C_{j,k_0}(\eta_{j,k_0}\cdot\|\lambda_j/\lambda_{k_0}\|)^N.
    \label{eq:product}
\end{align}
The right-hand side of~\eqref{eq:product} converges to $0$ exactly when
\eqref{eq:hyp} holds. The qualitative limit~\eqref{eq:qual} alone
gives $\eta_{j,k_0}<1$ but not the comparison
$\eta_{j,k_0}\cdot\|\lambda_j/\lambda_{k_0}\|<1$, which is the
content of the rate condition.

\section{Classification}

\emph{Open gap.}  The rate condition above is a structural input beyond what
\cite[\S II]{Cirac2016MPDO_arXiv} and
\cite[Definition~4.3]{Cirac2021Matrix} make explicit. Its rate is in
principle derivable from the spectral gap of the mixed transfer operator that
powers
\leanid{mpvOverlap\_tendsto\_zero\_of\_not\_gaugePhaseEquiv\_cast\_left\_of\_irreducible\_TP},
but the compatibility inequality
$\eta_{j,k}\cdot\|\lambda_j/\lambda_k\|<1$ is not implied by the
BNT structure and must be either imposed or established analytically.

The analytic derivation from the transfer matrix spectral gap is the
subject of \emph{Option~2} in the design memo
\path{audits/2026-05-13_cpsv16_ft_path_beta_pr2_5_design.md} and is
not addressed in this note.  The current PaperBNT route instead records the
source-faithful qualitative overlap decay, coefficient boundedness, and
coefficient non-vanishing separately in
\doclink{TNLean/MPS/FundamentalTheorem/PaperBNT/Api}; any future use of the
rate condition should be added as a theorem-level hypothesis or proved from
the mixed-transfer spectral gap, not by reviving the deleted two-layer
projection surface.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
