\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Strong Subadditivity as a Standalone Axiom and Its Lieb-Concavity
Elimination Route}
\author{}
\date{2026-06-22}

\begin{document}
\maketitle

\section{The assertion and its current formal status}

Strong subadditivity of the von Neumann entropy is the inequality
\begin{align}
  S(\rho_{ABC}) + S(\rho_{B}) \le S(\rho_{AB}) + S(\rho_{BC})
  \tag{SSA}
  \label{eq:ssa}
\end{align}
for every density operator $\rho_{ABC}$ on a tripartite finite-dimensional space
$\mathcal H_A \otimes \mathcal H_B \otimes \mathcal H_C$, where
$S(\rho) = -\tr(\rho \log \rho)$ and the reduced operators are the corresponding
partial traces \cite{LiebRuskai1973}.

In the present development the inequality~\eqref{eq:ssa} is asserted as a
standalone axiom.\footnote{The axiom is
\leanid{strong_subadditivity} in \doclink{TNLean/Axioms/Entropy}
(\path{TNLean/Axioms/Entropy.lean}). It is stated with the
eigenvalue-based entropy $S(\rho) = \sum_i \eta(\lambda_i)$, $\eta(x) = -x\log x$,
of \doclink{TNLean/Analysis/Entropy}.}
It is \emph{not} derived from the operator-concavity input that the same
development already isolates for the Ando--Lieb theory, namely Lieb's joint
concavity theorem.\footnote{Lieb concavity is the axiom
\leanid{lieb_concavity_axiom} in \doclink{TNLean/Axioms/OperatorConvexity}
(\path{TNLean/Axioms/OperatorConvexity.lean}), consumed by
\leanid{lieb_concavity} in \doclink{TNLean/Channel/Schwarz/AndoLieb}.}
The two axioms are therefore logically independent in the current
formalization: discharging~\eqref{eq:ssa} into a theorem built on Lieb
concavity is open work, not a completed reduction. This note records the
mathematical route from the one axiom to the other and names the intermediate
layers that are missing.

\section{The point at issue}

The standard derivation of~\eqref{eq:ssa} factors through the quantum relative
entropy and its behaviour under coarse-graining. Several of the intermediate
objects are now formalized and several are still missing, so the gap is not a
single missing lemma but a chain of them; the per-layer status is recorded below
and summarized under \emph{Current formal status}. In source order the chain is
the following.

\begin{enumerate}
  \item \emph{Entropy bridge.} The identity
    $S(\rho) = -\tr(\rho \log \rho)$ relating the trace-of-logarithm form to the
    eigenvalue form $\sum_i \eta(\lambda_i)$ used in the formal definition. This
    bridge is the prerequisite that makes the trace-functional manipulations of
    the relative entropy meaningful for the formal entropy, and it is dispatched
    separately.\footnote{Tracked on \ghissue{784}.}
  \item \emph{Relative entropy.} The definition
    \begin{align}
      D(\rho \,\|\, \sigma) = \tr\bigl(\rho(\log \rho - \log \sigma)\bigr),
    \end{align}
    for density operators $\rho,\sigma$ with $\ker\sigma \subseteq \ker\rho$.
  \item \emph{Klein's inequality.} Nonnegativity $D(\rho\,\|\,\sigma) \ge 0$,
    with equality exactly when $\rho = \sigma$. This is the order property that
    makes $D$ a divergence. The nonnegativity half is now formalized:
    \leanid{quantumRelativeEntropy_nonneg} in
    \doclink{TNLean/Analysis/KleinInequality}
    (\path{TNLean/Analysis/KleinInequality.lean}) proves
    $D(\rho\,\|\,\sigma)\ge 0$ for density operators $\rho,\sigma$ with $\sigma$
    of full rank, by the eigenbasis reduction to the doubly stochastic overlap
    matrix and the scalar Gibbs inequality $\log x\le x-1$; it consumes neither
    Lieb concavity nor operator Jensen. The equality case and the
    singular-$\sigma$ extension under $\ker\sigma\subseteq\ker\rho$ remain open.
  \item \emph{Joint convexity.} Convexity of
    $(\rho,\sigma) \mapsto D(\rho\,\|\,\sigma)$ in the pair of arguments. This is
    the step that consumes Lieb's joint concavity theorem: the concavity of
    $(A,B)\mapsto \tr\!\bigl(K^{\dagger} A^{s} K B^{1-s}\bigr)$ for
    $s\in[0,1]$ yields, in the limit governing the derivative at $s=1$, the
    convexity of the relative entropy. The present formal Lieb axiom is stated
    for positive definite inputs, whereas the relative-entropy domain in layer
    (2) allows singular reference states provided
    $\ker\sigma\subseteq\ker\rho$. Therefore this layer also needs the standard
    extension from the full-rank case to the support-condition case: replace the
    pair by strictly positive regularizations, prove convexity there, and pass to
    the limit using continuity on the common support and lower semicontinuity at
    the boundary. Without this regularization step, the later support lemma would
    place the singular pairs in the domain of $D$ but would not yet connect them
    to the positive-definite Lieb input. The positive-definite case of this layer
    is now formalized (\leanid{convexOn_quantumRelativeEntropy}); only the
    singular-$\sigma$ extension remains.
  \item \emph{Data-processing inequality (DPI).} Monotonicity
    $D(\Lambda\rho \,\|\, \Lambda\sigma) \le D(\rho\,\|\,\sigma)$ under every
    trace-preserving completely positive map $\Lambda$, in particular under the
    partial trace. Monotonicity under the partial trace is the form of DPI used
    below; it follows from joint convexity together with unitary invariance of
    $D$ and the ancilla-additivity identity
    $D(\rho \otimes \tau \,\|\, \sigma \otimes \tau) = D(\rho\,\|\,\sigma)$ for a
    fixed ancilla state $\tau$. The standard twirling argument writes the partial
    trace $\tr_C$ as an average of unitary conjugations on $\mathcal H_C$ applied
    after appending the maximally mixed ancilla $\tau = \mathbf 1_C / d_C$, so all
    three ingredients --- joint convexity, unitary invariance, and
    ancilla-additivity --- are needed; joint convexity and unitary invariance
    alone do not suffice.
  \item \emph{Final assembly.} The inequality~\eqref{eq:ssa} read as a single
    instance of DPI under the partial trace over $C$. Take the reference state
    $\sigma_{ABC} = (\mathbf 1_A / d_A) \otimes \rho_{BC}$, the tensor product of
    the maximally mixed state $\mathbf 1_A / d_A$ on $\mathcal H_A$
    ($d_A = \dim \mathcal H_A$) with the reduced state $\rho_{BC}$ on
    $\mathcal H_B \otimes \mathcal H_C$. Being a density operator is \emph{not}
    enough to place the pair $(\rho_{ABC},\sigma_{ABC})$ in the domain fixed in
    layer (2): that domain is the kernel-containment condition
    $\ker\sigma_{ABC} \subseteq \ker\rho_{ABC}$, and as soon as $\rho_{BC}$ fails
    to be full rank the reference state $\sigma_{ABC}$ is singular, so the
    condition is a genuine hypothesis to be discharged and is also what licenses
    the displayed logarithm. The containment is supplied by the following
    \emph{marginal support lemma}: for any density operator $\rho_{ABC}\ge 0$ with
    marginal $\rho_{BC} = \tr_A \rho_{ABC}$,
    \begin{align}
      \mathcal H_A \otimes \ker\rho_{BC} \subseteq \ker\rho_{ABC},
      \qquad\text{equivalently}\qquad
      \operatorname{supp}\rho_{ABC} \subseteq
        \mathcal H_A \otimes \operatorname{supp}\rho_{BC}.
      \label{eq:marginal-support}
    \end{align}
    To see~\eqref{eq:marginal-support}, let $P_{BC}$ be the projection onto
    $\operatorname{supp}\rho_{BC}$ and $Q = \mathbf 1_A \otimes (\mathbf 1_{BC} -
    P_{BC})$ the projection onto $\mathcal H_A \otimes \ker\rho_{BC}$. Using
    $\tr_A \rho_{ABC} = \rho_{BC}$ and $(\mathbf 1_{BC} - P_{BC})\,\rho_{BC} = 0$,
    \begin{align}
      \tr\bigl(Q\,\rho_{ABC}\bigr)
        = \tr\bigl((\mathbf 1_{BC} - P_{BC})\,\tr_A\rho_{ABC}\bigr)
        = \tr\bigl((\mathbf 1_{BC} - P_{BC})\,\rho_{BC}\bigr) = 0 .
    \end{align}
    Since $\rho_{ABC}\ge 0$ and $Q$ is a projection,
    $\tr(Q\,\rho_{ABC}) = \lVert Q\,\rho_{ABC}^{1/2}\rVert_2^2 = 0$ forces
    $Q\,\rho_{ABC}^{1/2} = 0$, hence $Q\,\rho_{ABC} = 0$, so the range of $Q$ ---
    namely $\mathcal H_A \otimes \ker\rho_{BC}$ --- lies in $\ker\rho_{ABC}$.
    Because $\mathbf 1_A / d_A$ is full rank we have $\ker\sigma_{ABC} =
    \mathcal H_A \otimes \ker\rho_{BC}$, so~\eqref{eq:marginal-support} gives
    exactly $\ker\sigma_{ABC} \subseteq \ker\rho_{ABC}$ and places the pair in the
    domain of layer (2). On $\operatorname{supp}\sigma_{ABC}$, where the relative
    entropy is evaluated, the logarithm splits as
    $\log\bigl((\mathbf 1_A / d_A) \otimes \rho_{BC}\bigr)
      = -(\log d_A)\,\mathbf 1_{ABC} + \mathbf 1_A \otimes \log\rho_{BC}$
    (with $\log\rho_{BC}$ taken on $\operatorname{supp}\rho_{BC}$), and the
    support condition guarantees $\rho_{ABC}$ carries no weight off this support,
    so the relative entropy evaluates to
    \begin{align}
      D\bigl(\rho_{ABC} \,\|\, (\mathbf 1_A / d_A) \otimes \rho_{BC}\bigr)
        = \log d_A + S(\rho_{BC}) - S(\rho_{ABC}).
    \end{align}
    Applying the partial trace $\Lambda = \tr_C$ sends
    $\rho_{ABC} \mapsto \rho_{AB}$ and
    $(\mathbf 1_A / d_A) \otimes \rho_{BC} \mapsto
    (\mathbf 1_A / d_A) \otimes \rho_{B}$. The image pair inherits its own support
    condition from~\eqref{eq:marginal-support} applied to $\rho_{AB}$ with
    marginal $\rho_{B} = \tr_A\rho_{AB}$, so it again lies in the domain of layer
    (2) and the image relative entropy is
    \begin{align}
      D\bigl(\rho_{AB} \,\|\, (\mathbf 1_A / d_A) \otimes \rho_{B}\bigr)
        = \log d_A + S(\rho_{B}) - S(\rho_{AB}).
    \end{align}
    Monotonicity under $\Lambda$, namely
    $D(\Lambda\rho \,\|\, \Lambda\sigma) \le D(\rho\,\|\,\sigma)$, then reads
    $\log d_A + S(\rho_{B}) - S(\rho_{AB}) \le
     \log d_A + S(\rho_{BC}) - S(\rho_{ABC})$; the common $\log d_A$ cancels,
    leaving $S(\rho_{B}) - S(\rho_{AB}) \le S(\rho_{BC}) - S(\rho_{ABC})$, which on
    rearrangement is exactly~\eqref{eq:ssa}; equivalently it is the
    nonnegativity of the conditional mutual information $I(A:C\mid B)\ge 0$. The
    discarded subsystem is $C$ and the reference state carries the maximally mixed
    state on the \emph{retained} factor $A$; discarding $A$ instead against
    $\rho_A \otimes \rho_{BC}$ would only yield subadditivity
    $I(A:BC)\ge 0$, not strong subadditivity.
\end{enumerate}

The mathematical question is therefore not whether~\eqref{eq:ssa} is true --- it
is a theorem of Lieb and Ruskai --- but which of the layers (2)--(6) must be
built so that the standalone axiom can be deleted and replaced by a proof whose
only entropy-side input is the Lieb-concavity axiom already present.

\section{The elimination route}

The route deletes the axiom \leanid{strong_subadditivity} once the chain above is
in place. Concretely:

\begin{itemize}
  \item Layer (1) is the entropy bridge and is independent of Lieb concavity;
    it is the lemma that lets the eigenvalue-based formal entropy be treated as a
    trace of a matrix logarithm.
  \item Layers (2) and (3) define the divergence and establish its
    nonnegativity; they need only the matrix logarithm and its operator
    monotonicity, not Lieb concavity. Layer (2) is in place
    (\leanid{quantumRelativeEntropy}) and layer (3), Klein's inequality, is now
    formalized for full-rank $\sigma$ (\leanid{quantumRelativeEntropy_nonneg}),
    via the eigenbasis reduction and the scalar Gibbs inequality rather than
    operator monotonicity of the logarithm.
  \item Layer (4) is the single consumer of \leanid{lieb_concavity_axiom}: joint
    convexity of $D$ is the entropy-theoretic content of Lieb's theorem. This
    layer is now formalized on the positive-definite domain:
    \leanid{convexOn_quantumRelativeEntropy} in
    \doclink{TNLean/Channel/Schwarz/RelativeEntropyConvexity}
    (\path{TNLean/Channel/Schwarz/RelativeEntropyConvexity.lean}) proves that
    $(\rho,\sigma)\mapsto D(\rho\,\|\,\sigma)$ is jointly convex on pairs of
    positive definite matrices, via the Lindblad/Uhlmann approximant
    $g_s(\rho,\sigma) = (\tr\rho
    - \operatorname{Re}\tr(\rho^s\sigma^{1-s}))/(1-s)$: each $g_s$ is jointly
    convex for $s\in[0,1)$ (\leanid{convexOn_relativeEntropyApprox}, the
    consumer of \leanid{lieb_concavity_axiom}) and converges pointwise to $D$ as
    $s\to1^-$ (\leanid{tendsto_relativeEntropyApprox}), and the pointwise limit
    of convex functions is convex. The two-calculus trace double-sum identity
    \leanid{re_trace_cfc_mul_cfc_eq_double_sum} underlying the limit consumes no
    Lieb input. The extension from the positive-definite case to singular
    reference states under $\ker\sigma\subseteq\ker\rho$ (the regularization
    step described above) remains open.
  \item Layer (5) combines joint convexity, unitary invariance, and
    ancilla-additivity $D(\rho\otimes\tau\,\|\,\sigma\otimes\tau)
    = D(\rho\,\|\,\sigma)$ into the partial-trace monotonicity statement via the
    twirling representation of $\tr_C$. The unitary-invariance ingredient is now
    formalized for arbitrary Hermitian arguments:
    \leanid{quantumRelativeEntropy_conj_unitary} in
    \doclink{TNLean/Channel/Schwarz/RelativeEntropyUnitaryInvariance}
    (\path{TNLean/Channel/Schwarz/RelativeEntropyUnitaryInvariance.lean}) proves
    $D(U\rho U^\dagger\,\|\,U\sigma U^\dagger) = D(\rho\,\|\,\sigma)$ for every
    unitary $U$, by the covariance of the matrix logarithm under unitary
    conjugation (\leanid{Matrix.log_conj_unitary}, itself the $f=\log$ case of the
    functional-calculus covariance \leanid{Matrix.cfc_conj_unitary}) followed by
    trace cyclicity. This ingredient needs no Lieb input. The
    ancilla-additivity ingredient is now formalized as well, on the
    positive-definite domain:
    \leanid{quantumRelativeEntropy_kronecker} in
    \doclink{TNLean/Channel/Schwarz/RelativeEntropyAncillaAdditivity}
    (\path{TNLean/Channel/Schwarz/RelativeEntropyAncillaAdditivity.lean}) proves
    $D(\rho\otimes\tau\,\|\,\sigma\otimes\tau) = D(\rho\,\|\,\sigma)$ for positive
    definite $\rho,\sigma$ and a positive definite $\tau$ of unit trace. It
    reduces to the logarithm of a tensor product
    $\log(\rho\otimes\tau) = \log\rho\otimes\mathbf 1 + \mathbf 1\otimes\log\tau$
    (\leanid{Matrix.log_kronecker}): the one-sided splits
    $\log(A\otimes\mathbf 1) = \log A\otimes\mathbf 1$
    (\leanid{Matrix.cfc_kronecker_one}) and
    $\log(\mathbf 1\otimes B) = \mathbf 1\otimes\log B$
    (\leanid{Matrix.cfc_one_kronecker}) follow from the functional calculus
    through the unital tensor embeddings, and recombining them uses the logarithm
    of a commuting product $\log(XY) = \log X + \log Y$ for commuting positive
    definite $X,Y$ (\leanid{Matrix.PosDef.cfc_log_mul} in
    \doclink{TNLean/Analysis/CfcLogAdditive},
    \path{TNLean/Analysis/CfcLogAdditive.lean}). The difference of tensor
    logarithms then cancels the $\mathbf 1\otimes\log\tau$ terms, the trace of the
    Kronecker product factors as a product of traces, and the unit trace of $\tau$
    leaves $D(\rho\,\|\,\sigma)$. These results depend only on the standard Lean
    axioms and need no Lieb input. The third ingredient of layer (5) is the
    twirling identity, which writes
    $\tr_C\rho\otimes(\mathbf 1_C/d_C)$ as a uniform average of conjugations by a
    unitary $1$-design on $C$. Because the current dependency carries no Haar
    measure on the matrix unitary group, the average is taken over the finite
    Heisenberg--Weyl $1$-design. Its design property is now formalized:
    \leanid{Matrix.sum_weyl_conj} in
    \doclink{TNLean/Channel/Schwarz/WeylTwirl}
    (\path{TNLean/Channel/Schwarz/WeylTwirl.lean}) proves that for a primitive
    $d$-th root of unity the uniform average of the conjugations by the $d^2$ Weyl
    operators $W(a,b) = X^a Z^b$ equals the completely depolarizing channel
    $M \mapsto (\tr M / d)\,\mathbf 1$, via the root-of-unity character-sum
    orthogonality (\leanid{Matrix.sum_rootOfUnity_pow_eq}). This result depends
    only on the standard Lean axioms and needs no Lieb input. Lifting that
    design property to the ancilla factor presents $\tr_C$ through the average:
    \leanid{Matrix.sum_kronecker_one_weyl_conj} in
    \doclink{TNLean/Channel/Schwarz/RelativeEntropyDataProcessing}
    (\path{TNLean/Channel/Schwarz/RelativeEntropyDataProcessing.lean}) proves that
    the uniform average of the conjugations by the $d_C^2$ unitaries
    $\mathbf 1_S \otimes W(a,b)$ equals $(\tr_C M)\otimes(\mathbf 1_C/d_C)$. With
    all three ingredients in place, layer (5) is now formalized on the
    positive-definite domain: \leanid{quantumRelativeEntropy_partialTraceRight_le}
    in the same file proves
    $D(\tr_C\rho\,\|\,\tr_C\sigma)\le D(\rho\,\|\,\sigma)$ for positive definite
    $\rho,\sigma$, via the ancilla-additivity expansion of the reduced-state
    relative entropy, the twirling identity, joint convexity over the $d_C^2$
    equally weighted conjugated pairs, and unitary invariance of each term. The
    transport of joint convexity to the product index uses the reindexing
    invariance \leanid{quantumRelativeEntropy_submatrix_equiv}.
  \item Layer (6) instantiates layer (5) at the tripartite splitting to
    recover~\eqref{eq:ssa}; it additionally needs the marginal support
    lemma~\eqref{eq:marginal-support}, which puts the singular reference state
    $(\mathbf 1_A/d_A)\otimes\rho_{BC}$ into the domain of layer (2) and so is a
    prerequisite of the instantiation rather than a consequence of normalization.
    The marginal support lemma is now formalized
    (\leanid{Matrix.marginal_support} in \doclink{TNLean/Analysis/MarginalSupport},
    \path{TNLean/Analysis/MarginalSupport.lean}): for a positive semidefinite
    $\rho_{ABC}$ a Hermitian idempotent on $B\otimes C$ annihilating the marginal
    $\tr_A\rho_{ABC}$ has its identity-on-$A$ lift annihilate $\rho_{ABC}$. Its
    operator core --- a projection of zero expectation against a positive
    semidefinite state annihilates it
    (\leanid{Matrix.PosSemidef.proj_mul_eq_zero_of_trace_eq_zero}) --- and the
    partial-trace adjoint identity it uses
    (\leanid{Matrix.traceA_ABC_lift_trace}) are independent of Lieb concavity.
    These three results depend only on the standard Lean axioms; the remaining
    layer-(6) work is the instantiation, which is gated on layer (5).
\end{itemize}

\paragraph{Current formal status.} Layers (1)--(5) and the layer-(6) marginal
support prerequisite are formalized; layers (4) and (5) on the positive-definite
domain. Layer (5), data-processing (monotonicity of $D$ under the partial trace),
combines the layer-(4) joint convexity with unitary invariance and
ancilla-additivity $D(\rho\otimes\tau\,\|\,\sigma\otimes\tau) =
D(\rho\,\|\,\sigma)$ via the twirling representation of $\tr_C$. Joint convexity
(layer 4, on the positive-definite domain), unitary invariance
(\leanid{quantumRelativeEntropy_conj_unitary}, for arbitrary Hermitian
arguments), ancilla-additivity
(\leanid{quantumRelativeEntropy_kronecker}, on the positive-definite domain), the
finite Heisenberg--Weyl $1$-design twirl (\leanid{Matrix.sum_weyl_conj}), and its
ancilla-factor lift presenting $\tr_C$ through the twirl
(\leanid{Matrix.sum_kronecker_one_weyl_conj}) are formalized, and combine into the
partial-trace monotonicity
\leanid{quantumRelativeEntropy_partialTraceRight_le} on the positive-definite
domain. Only the layer-(6) instantiation remains; until it is in place the
standalone axiom \leanid{strong_subadditivity} stands.

After layer (6) the inequality is a theorem, the standalone axiom is removed, and
the only remaining entropy-side axiom on this route is the Lieb-concavity axiom.
The downstream consumers in the stable entropy interface
\footnote{The stable interface re-exports the inequality from
\doclink{TNLean/Entropy/StrongSubadditivity}; consumers import it from there and
not from the axiom module.}
require no change because the statement they import is unchanged; only its
justification changes from axiom to theorem.

\section{Verdict}

The strong-subadditivity inequality is presently an axiom that is logically
disconnected from the Lieb-concavity axiom available in the same repository. The
elimination route is the classical relative-entropy chain: entropy bridge,
definition of relative entropy, Klein's inequality, joint convexity (the step
consuming Lieb concavity), data-processing under the partial trace, and final
assembly at the tripartite splitting. Layers (2) and (3) are formalized (the
latter for full-rank $\sigma$), and layer (4) is formalized on the
positive-definite domain (\leanid{convexOn_quantumRelativeEntropy}), so the
Lieb-concavity axiom is now connected to the joint convexity of the relative
entropy. Layer (5), data-processing, is formalized on the positive-definite
domain (\leanid{quantumRelativeEntropy_partialTraceRight_le}): its unitary
invariance (\leanid{quantumRelativeEntropy_conj_unitary}), ancilla additivity
(\leanid{quantumRelativeEntropy_kronecker}), finite Heisenberg--Weyl $1$-design
twirl (\leanid{Matrix.sum_weyl_conj}), and the ancilla-factor lift presenting
$\tr_C$ through the twirl (\leanid{Matrix.sum_kronecker_one_weyl_conj}) combine
through joint convexity into the partial-trace monotonicity
$D(\tr_C\rho\,\|\,\tr_C\sigma)\le D(\rho\,\|\,\sigma)$. The remaining work is the
singular-$\sigma$ extension of layers (3)--(5) and the instantiation layer (6);
until those are formalized, the standalone axiom records a result the development
assumes but does not yet derive from its operator-convexity input.

\begin{thebibliography}{9}

\bibitem{LiebRuskai1973}
  E.~H.~Lieb and M.~B.~Ruskai,
  \emph{Proof of the strong subadditivity of quantum-mechanical entropy},
  Journal of Mathematical Physics \textbf{14} (1973), 1938--1941.

\bibitem{Lieb1973Convex}
  E.~H.~Lieb,
  \emph{Convex trace functions and the Wigner--Yanase--Dyson conjecture},
  Advances in Mathematics \textbf{11} (1973), 267--288.

\end{thebibliography}

\end{document}
