\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}

\input{./command}

\title{The Fixed-Block Cancellation Step in CPSV16 Theorem II.1}
\author{}
\date{2026-05-11}

\begin{document}
\maketitle

\section{Source passage}

This note concerns the proof of CPSV16
\cite[Theorem~\texttt{thm1}, lines~1181--1185]{Cirac2016MPDO_arXiv}.
In the proof one fixes a block \(B_k\).  The source then asserts that the
asymptotic orthogonality alternative
\[
  \langle V^{(N)}(A_j),V^{(N)}(B_k)\rangle \longrightarrow 0
  \qquad (1\le j\le g_A)
  \tag{$D_k$}
\]
is impossible, because it would contradict proportionality of the full MPV
families by
\cite[Corollary~\texttt{Lem1}, lines~1130--1133]{Cirac2016MPDO_arXiv}.
The same sentence, with \(A\) and \(B\) interchanged, gives the opposite
block-count inequality.

\section{The identities}

Write
\[
  a_j(N)=V^{(N)}(A_j),
  \qquad
  b_\ell(N)=V^{(N)}(B_\ell).
\]
The proportionality of the total MPV families is an equality of the weighted
sums
\[
  \sum_{j=1}^{g_A} \mu_j^N a_j(N)
  =
  c_N\sum_{\ell=1}^{g_B}\nu_\ell^N b_\ell(N),
  \qquad c_N\ne 0 .
  \tag{$P_N$}
\]
The BNT normalization gives, for large \(N\),
\[
  \langle a_i(N),a_j(N)\rangle\longrightarrow \delta_{ij},
  \qquad
  \langle b_k(N),b_k(N)\rangle\longrightarrow 1 .
  \tag{BNT}
\]
Combining \((D_k)\) with \((BNT)\), Corollary~\texttt{Lem1} gives exactly
\[
  \operatorname{LinIndep}\{a_1(N),\ldots,a_{g_A}(N),b_k(N)\}
  \qquad(N\gg0).
  \tag{$LI_k$}
\]
This is the local conclusion supplied by the cited source passage.  It is the
faithful statement currently formalized for the fixed-right and fixed-left
applications of Lemma~\texttt{Lem1}.

\section{What \((LI_k)\) proves immediately}

If \(g_B=1\), then \((P_N)\) has no \(B\)-tail.  Separating the selected block
gives
\[
  \sum_{j=1}^{g_A}\mu_j^N a_j(N)-c_N\nu_k^N b_k(N)=0.
  \tag{1}
\]
By \((LI_k)\), the coefficient of \(b_k(N)\) in \((1)\) must vanish:
\[
  c_N\nu_k^N=0.
\]
This contradicts \(c_N\ne0\) and the nonzero BNT weight \(\nu_k\).  Thus the
no-tail base case is a direct coefficient-extraction argument.

If \(k\) is the leading \(B\)-block, there is also a dominant-projection
argument.  After normalizing by the leading weights, testing \((P_N)\) against
\(b_k(N)\) gives a term of the form
\[
  c_N\left(\frac{\nu_k}{\mu_1}\right)^N
  \langle b_k(N),b_k(N)\rangle
\]
on the \(B\)-side.  Under \((D_k)\), the \(A\)-side tends to \(0\), while
\(\langle b_k(N),b_k(N)\rangle\to1\).  For the leading block the remaining
normalized \(B\)-tail is strictly smaller, so the selected term cannot be
cancelled by larger weights.  This proves the leading-block contradiction
without assuming linear independence of the whole residual family.

\section{Where the general fixed-block proof is missing}

For an arbitrary \(k\), separating the selected summand in \((P_N)\) gives
\[
  \sum_{j=1}^{g_A}\mu_j^N a_j(N)
  -
  c_N\nu_k^N b_k(N)
  =
  c_N\sum_{\ell\ne k}\nu_\ell^N b_\ell(N).
  \tag{2}
\]
The family in \((LI_k)\) controls only linear relations of the form
\[
  \sum_{j=1}^{g_A}\alpha_j(N)a_j(N)+\beta(N)b_k(N)=0.
  \tag{3}
\]
It does not apply directly to \((2)\), because the right-hand side of \((2)\)
contains vectors outside the family \((LI_k)\).  To extract the coefficient of
\(b_k(N)\) from \((2)\), one would need an additional isolation statement, for
example
\[
  c_N\sum_{\ell\ne k}\nu_\ell^N b_\ell(N)
  \in
  \operatorname{span}\{a_1(N),\ldots,a_{g_A}(N)\},
  \tag{$I_k$}
\]
or equivalently a functional \(f_N\) such that
\[
  f_N(a_j(N))=0,\qquad
  f_N(b_\ell(N))=0\ (\ell\ne k),\qquad
  f_N(b_k(N))=1.
  \tag{$F_k$}
\]
Neither \((I_k)\) nor \((F_k)\) follows from \((LI_k)\).  In purely algebraic
terms, \((LI_k)\) rules out
\[
  b_k(N)\in \operatorname{span}\{a_j(N)\}_j,
\]
but the needed contradiction rules out the stronger possibility
\[
  c_N\nu_k^N b_k(N)
  \in
  \operatorname{span}\{a_j(N)\}_j
  +
  \operatorname{span}\{b_\ell(N):\ell\ne k\}.
  \tag{4}
\]
Statement \((4)\) is precisely the fixed-block cancellation difficulty.

The dominant-projection proof also does not isolate an arbitrary fixed block.
Testing \((P_N)\) against \(b_k(N)\), after normalization by a leading weight,
leaves
\[
  c_N\sum_{\ell=1}^{g_B}
  \left(\frac{\nu_\ell}{\mu_1}\right)^N
  \langle b_\ell(N),b_k(N)\rangle .
  \tag{5}
\]
When \(k\) is not leading, terms with \(|\nu_\ell|>|\nu_k|\) may dominate
\((5)\).  If instead one normalizes by \(\nu_k^N\), then terms with
\(|\nu_\ell|>|\nu_k|\) grow.  Thus the leading-block argument is not a proof of
the arbitrary fixed-block assertion at line~1182 of the source.

\section{Symmetric fixed-left statement}

The proof with \(A\) and \(B\) interchanged has the same form.  Fixing an
\(A_i\), the decay alternative is
\[
  \langle a_i(N),b_\ell(N)\rangle\longrightarrow 0
  \qquad(1\le \ell\le g_B),
  \tag{$D'_i$}
\]
and Lemma~\texttt{Lem1} supplies
\[
  \operatorname{LinIndep}\{a_i(N),b_1(N),\ldots,b_{g_B}(N)\}
  \qquad(N\gg0).
  \tag{$LI'_i$}
\]
The proportionality identity still contains the remaining \(A\)-summands:
\[
  c_N\sum_{\ell=1}^{g_B}\nu_\ell^N b_\ell(N)
  -
  \mu_i^N a_i(N)
  =
  \sum_{j\ne i}\mu_j^N a_j(N).
  \tag{6}
\]
Thus the missing left-handed step is the symmetric isolation of \(a_i(N)\) from
the residual \(A\)-tail.

\section{Formal boundary}

The existing conditional tail lemmas assume a larger independence statement,
schematically
\[
  \operatorname{LinIndep}
  \bigl(\{a_j(N)\}_j\cup\{b_\ell(N):\ell\ne k\}\cup\{b_k(N)\}\bigr)
  \qquad(N\gg0).
  \tag{$G_k$}
\]
From \((G_k)\), coefficient extraction from \((2)\) is immediate.  However
\((G_k)\) is stronger than Corollary~\texttt{Lem1} as used at
\cite[Theorem~\texttt{thm1}, line~1182]{Cirac2016MPDO_arXiv}.  The source
sentence supplies \((LI_k)\), not global residual-family independence
\((G_k)\).  Therefore the conditional tail lemmas are valid local algebraic
tools, but they are not yet the faithful formalization of the fixed-block
cancellation sentence in the source.

\section{Formal declarations}

The two named proof obligations are
\[
\begin{aligned}
&\leanid{MPSTensor.fixed\_right\_all\_overlaps\_decay\_false\_of\_eventuallyNonzeroProportionalMPV₂\_CFBNT},\\
&\leanid{MPSTensor.fixed\_left\_all\_overlaps\_decay\_false\_of\_eventuallyNonzeroProportionalMPV₂\_CFBNT}.
\end{aligned}
\]
They are statements of the cancellation step implicit in the source passage,
not strengthened variants of the source theorem.  The formal proof obligation
is tracked by issue~\#1607.

The no-tail base cases, where the selected side has no other summands, are
already formalized.  In those cases ordinary coefficient extraction from the
eventual linear independence supplied by Lemma~\texttt{Lem1} forces a nonzero
selected coefficient to vanish.  The corresponding declarations are
\[
\begin{aligned}
&\leanid{MPSTensor.fixed\_right\_all\_overlaps\_decay\_false\_of\_eventuallyNonzeroProportionalMPV₂\_CFBNT\_finOne},\\
&\leanid{MPSTensor.fixed\_left\_all\_overlaps\_decay\_false\_of\_eventuallyNonzeroProportionalMPV₂\_CFBNT\_finOne}.
\end{aligned}
\]

The leading-block cases are also formalized.  They are discharged by the
dominant projection contradiction already available for eventually nonzero
proportional BNT families:
\[
\begin{aligned}
&\leanid{MPSTensor.fixed\_right\_leading\_all\_overlaps\_decay\_false\_of\_eventuallyNonzeroProportionalMPV₂\_CFBNT},\\
&\leanid{MPSTensor.fixed\_left\_leading\_all\_overlaps\_decay\_false\_of\_eventuallyNonzeroProportionalMPV₂\_CFBNT}.
\end{aligned}
\]

The first partner-identification step for peeling is now also formalized:
\begin{flushleft}\small
\leanid{MPSTensor.leading\_right\_nondecaying\_partner\_eq\_leading\_left\_of\_eventuallyNonzeroProportionalMPV₂\_CFBNT}.
\end{flushleft}
It proves that, under eventual nonzero proportionality, a non-decaying partner
of the leading \(B\)-block must be the leading \(A\)-block.  This is still only
a preparatory step: it identifies the pair eligible for dominant tail
subtraction, but it does not by itself prove the arbitrary fixed-block
cancellation statements.

The resulting leading phase relation and asymptotic leading-tail estimate are
recorded in:
\begin{flushleft}\small
\leanid{MPSTensor.exists\_leading\_phase\_tail\_diff\_tendsto\_zero\_of\_eventuallyNonzeroProportionalMPV₂\_CFBNT}.
\end{flushleft}
This lemma supplies the phase relation for the leading blocks, written
\(A_0,B_0\) in the zero-indexed formal declarations and \(A_1,B_1\) in the
formulas above.  It proves that the normalized leading-erased tail difference
tends to zero.  It is not the missing exact selected-summand cancellation: the
remaining proof must still extract the leading coefficient identity, or an
equivalent tail proportionality statement, without assuming residual-family
linear independence beyond the fixed-block Lemma~\texttt{Lem1} input.

There are also conditional tail-peeling auxiliary lemmas:
\[
\begin{aligned}
&\leanid{MPSTensor.eventually\_selected\_weighted\_mpvState\_eq\_smul\_of\_phase\_sum\_and\_li},\\
&\leanid{MPSTensor.eventuallyNonzeroProportionalMPV₂\_tail\_succAbove\_of\_phase\_sum\_li},\\
&\leanid{MPSTensor.eventuallyNonzeroProportionalMPV₂\_tail\_succAbove\_of\_phase\_sum\_li\_left}.
\end{aligned}
\]
These are correct algebraic statements, but their linear-independence
hypotheses are stronger than the fixed-block application of
Lemma~\texttt{Lem1}.  The right-handed auxiliary lemma assumes eventual linear
independence of all \(A\)-blocks together with the whole remaining \(B\)-tail;
the left-handed auxiliary lemma assumes the symmetric condition.  In contrast,
the source sentence at line 1182 supplies only the family consisting of all
blocks on one side together with the single fixed block on the other side.
These auxiliary lemmas are therefore conditional local tools, not
formalizations of the fixed-block cancellation step.

\section{Elimination plan}

The intended elimination is to prove the two fixed-block contradiction
declarations from the BNT hypotheses, eventual nonzero proportionality, and the
already formalized fixed-block Lemma~\texttt{Lem1} inputs.  A proof may use an
induction through matched block removal, using the leading-block
contradictions above as the case in which the selected block has become
dominant, or a direct argument isolating the selected summand in the quotient
by the span of the opposite family.  It must not assume global linear
independence of all remaining blocks, since that is not a consequence of the
BNT hypotheses once a block on one side may be phase-equivalent to a block on
the other.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
