\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{The Strong-Subadditivity Equality Characterization as an External
Stand-In, and Its One-Directional Hygiene Split}
\author{}
\date{2026-06-22}

\begin{document}
\maketitle

\section{The assertion and its current formal status}

Equality in strong subadditivity,
\begin{align}
  S(\rho_{ABC}) + S(\rho_{B}) = S(\rho_{AB}) + S(\rho_{BC}),
  \tag{SSA$_{=}$}
  \label{eq:ssa-eq}
\end{align}
is characterized by a structural decomposition of the middle subsystem. The
characterization states that~\eqref{eq:ssa-eq} holds if and only if, after a
unitary change of basis on $\mathcal H_B$, the middle space decomposes as a
finite orthogonal direct sum
\begin{align}
  \mathcal H_B \;\cong\; \bigoplus_{j} \mathcal H_{B_j^{L}} \otimes
  \mathcal H_{B_j^{R}},
\end{align}
and in this basis the state is block diagonal,
\begin{align}
  \rho_{ABC} \;\cong\; \bigoplus_{j} p_j \,
  \rho_{A B_j^{L}} \otimes \rho_{B_j^{R} C},
  \tag{QMC}
  \label{eq:qmc}
\end{align}
with probabilities $p_j \ge 0$, $\sum_j p_j = 1$, and density operators
$\rho_{A B_j^{L}}$, $\rho_{B_j^{R} C}$ on the indicated factors. This is the
quantum-Markov-chain structure of the tripartite state
\cite{Ruskai2002, HaydenJozsaPetzWinter2004, Hayashi2006}.

In the present development the biconditional
\eqref{eq:ssa-eq} $\Leftrightarrow$~\eqref{eq:qmc} is asserted as a standalone
axiom.\footnote{The axiom is \leanid{hayashi_ssa_equality_characterization} in
\doclink{TNLean/Axioms/Entropy} (\path{TNLean/Axioms/Entropy.lean}, line~238).
The equality predicate is \leanid{IsSSAEquality} in
\doclink{TNLean/Analysis/Entropy} (\path{TNLean/Analysis/Entropy.lean},
line~559), and the right-hand structure is recorded by
\leanid{HayashiMarkovDecomposition} in \doclink{TNLean/Axioms/Entropy}
(\path{TNLean/Axioms/Entropy.lean}, line~142), whose fields store the
finite decomposition, the unitary basis change, the weights $p_j$, the component
density operators, and the block-diagonal equality.} It is an external stand-in:
the characterization is a theorem of Ruskai, of Hayden--Jozsa--Petz--Winter, and
of Hayashi, but its proof is not formalized here.

\section{The point at issue: only the forward direction is consumed}

The biconditional has two directions of very different difficulty and very
different usage.

\paragraph{Forward direction (consumed).}
The implication \eqref{eq:ssa-eq} $\Rightarrow$~\eqref{eq:qmc} --- equality
forces the quantum-Markov-chain structure --- is the only direction used
downstream. Its proof is the deep part of the characterization: it requires
recovery-map theory and the Petz transpose channel, that is, the analysis of
when the data-processing inequality for relative entropy is saturated and the
reconstruction of the discarded subsystem from a recovery map. This recovery-map
and Petz machinery is absent from the underlying library, which is why the
forward direction is axiomatized rather than proved.

\paragraph{Reverse direction (tractable, currently unused).}
The implication \eqref{eq:qmc} $\Rightarrow$~\eqref{eq:ssa-eq} --- a state of the
block-diagonal product form attains equality --- is elementary by comparison and
is computed directly from additivity properties of the entropy. For a state of
the form~\eqref{eq:qmc} the von Neumann entropy decomposes as
\begin{align}
  S\!\left(\bigoplus_j p_j\, \omega_j\right)
  = H(\{p_j\}) + \sum_j p_j\, S(\omega_j),
  \qquad
  S(\omega \otimes \tau) = S(\omega) + S(\tau),
\end{align}
where $H(\{p_j\}) = -\sum_j p_j \log p_j$ is the Shannon entropy of the weights.
Substituting~\eqref{eq:qmc} and its three marginals into both sides
of~\eqref{eq:ssa-eq} and cancelling the matching terms gives the equality with no
recovery-map input. The reverse direction is therefore provable from
entropy-additivity sub-lemmas alone, but it is not at present consumed by any
downstream result.

\section{The one-directional hygiene split}

Because only the forward direction is both deep and used, an optional cleanup
keeps the axiom boundary minimal: state the axiom as the single forward
implication \eqref{eq:ssa-eq} $\Rightarrow$~\eqref{eq:qmc}, and prove the reverse
implication \eqref{eq:qmc} $\Rightarrow$~\eqref{eq:ssa-eq} as a theorem. The full
biconditional is then recovered by combining the two, but the axiomatic content
is reduced to exactly the part that genuinely needs the missing theory.

The reverse half needs the following entropy-additivity sub-lemmas, all of which
are within reach of the present entropy interface:

\begin{itemize}
  \item additivity over a finite weighted orthogonal direct sum:
    $S\!\left(\bigoplus_j p_j\,\omega_j\right) = H(\{p_j\}) + \sum_j p_j
    S(\omega_j)$, i.e.\ the entropy of a block-diagonal state with weights;
  \item additivity over a tensor factorization:
    $S(\omega \otimes \tau) = S(\omega) + S(\tau)$;
  \item invariance of the entropy under a unitary basis change and under
    reindexing of the underlying space, so that the basis change on $\mathcal
    H_B$ and the direct-sum reindexing in~\eqref{eq:qmc} leave every entropy term
    unchanged.
\end{itemize}

With these three facts the reverse implication is a finite computation; the
forward implication remains an axiom until recovery-map and Petz theory are
available.

\section{Verdict}

The strong-subadditivity equality characterization is an external stand-in for a
theorem of Ruskai, of Hayden--Jozsa--Petz--Winter, and of Hayashi. Only the
forward direction, equality implies quantum-Markov-chain structure, is consumed
downstream, and it is precisely the direction that needs recovery-map and Petz
theory not present in the library. The reverse direction is an elementary
consequence of entropy additivity over weighted direct sums and tensor factors
together with unitary and reindexing invariance, and could be proved now; doing
so would let the axiom be narrowed to the single load-bearing forward
implication. Until the recovery-map theory is formalized, the forward direction
remains the irreducible axiomatic content of this characterization.

\begin{thebibliography}{9}

\bibitem{Ruskai2002}
  M.~B.~Ruskai,
  \emph{Inequalities for quantum entropy: a review with conditions for
  equality},
  Journal of Mathematical Physics \textbf{43} (2002), 4358--4375.

\bibitem{HaydenJozsaPetzWinter2004}
  P.~Hayden, R.~Jozsa, D.~Petz, and A.~Winter,
  \emph{Structure of states which satisfy strong subadditivity of quantum
  entropy with equality},
  Communications in Mathematical Physics \textbf{246} (2004), 359--374.

\bibitem{Hayashi2006}
  M.~Hayashi,
  \emph{Quantum Information: An Introduction},
  Springer, 2006; Theorem 5.24.

\end{thebibliography}

\end{document}
