\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Pure zero correlation length and local orthogonality in
Cirac--Perez-Garcia--Schuch--Verstraete 2017}
\author{}
\date{2026-06-10}

\begin{document}
\maketitle

\section{Source statement}

In Section~3.2 of \cite{Cirac2016MPDO_arXiv}, correlations independent of
distance are defined by requiring, for separated observables, that translating
one observable through any allowed separation does not change the expectation
value.\footnote{Local source:
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:437--445}, label
\path{ZCL}.}  The same subsection defines local orthogonality for a
basis-of-normal-tensors family by the mixed-sector equations
\begin{align}
  \mathbb E_{j,j'}=\sum_i A_j^i\otimes \overline{A_{j'}^i}=0,
  \qquad j\ne j',
  \tag{LO}
\end{align}
and then defines zero correlation length as the conjunction of CID and local
orthogonality.\footnote{Local source:
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:467--479}.}

For a single tensor $A$, write
\begin{align}
  \mathbb E_A=\sum_i A^i\otimes \overline{A^i}
\end{align}
for the transfer matrix.

The pure-state theorem states that, for a tensor $A$ in canonical form,
\begin{align}
  \mathrm{ZCL}(A)
  \quad\Longleftrightarrow\quad
  \mathbb E_A^2=\mathbb E_A.
  \tag{\path{TheoremZCLPure}}
\end{align}
In the proof, the source first identifies local orthogonality with the equations
$\mathbb E_{j,j'}=0$ for $j\ne j'$.  After assuming those equations, it proves
that $\mathbb E_A^2=\mathbb E_A$ is equivalent to CID.\footnote{Local source:
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:1248--1268}.}

\section{Formal boundary}

The current formal predicate
\leanid{MPSTensor.IsLocallyOrthogonal} is a one-tensor convention:
\begin{align}
  \leanid{MPSTensor.IsLocallyOrthogonal}(A)
  \quad:\!\!\Longleftrightarrow\quad
  \mathbb E_A^2=\mathbb E_A.
\end{align}
Consequently
\begin{align}
  \leanid{MPSTensor.IsZCL}(A)
  \quad:\!\!\Longleftrightarrow\quad
  \bigl(\mathbb E_A^2=\mathbb E_A\bigr)\wedge \mathrm{CID}(A).
\end{align}
The theorem \leanid{MPSTensor.zcl_iff_idempotent_transfer} therefore proves
the single-block equivalence
\begin{align}
  \bigl(\mathbb E_A^2=\mathbb E_A\bigr)\wedge \mathrm{CID}(A)
  \quad\Longleftrightarrow\quad
  \mathbb E_A^2=\mathbb E_A.
\end{align}
Its nontrivial direction is the implication
\[
  \mathbb E_A^2=\mathbb E_A \Longrightarrow \mathrm{CID}(A),
\]
obtained from $\mathbb E_A^n=\mathbb E_A$ for all $n\ge 1$.  The reverse
direction of that displayed equivalence is immediate because the left-hand side
already contains $\mathbb E_A^2=\mathbb E_A$.

This is a useful local statement, but it is not the source theorem
\path{TheoremZCLPure}: the source theorem uses the BNT-family predicate
\leanid{MPSTensor.IsBNTZCL}, whose local-orthogonality part is the
mixed-sector equations (LO).

\section{Elimination plan}

The formalization now has a BNT-family version of local orthogonality carrying
the equations
\[
  \mathbb E_{j,j'}=0 \qquad (j\ne j'),
\]
and a BNT-family ZCL predicate
\[
  \mathrm{ZCL}_{\mathrm{BNT}}(A)
  :=
  \mathrm{CID}(A)\wedge
  \bigl(\forall j\ne j',\, \mathbb E_{j,j'}=0\bigr).
\]
The remaining target theorem for tensors in canonical form is then
\[
  \mathrm{ZCL}_{\mathrm{BNT}}(A)
  \Longleftrightarrow
  \mathbb E_A^2=\mathbb E_A.
\]
The existing single-block theorem supplies only the CID consequence of
idempotence and should remain marked as a scope restriction until the
canonical-form theorem is stated with the BNT predicate.\footnote{Tracked by
\ghissue{2597}.}

The present theorem is therefore a scope restriction: it records the one-block
idempotence/CID consequence, while the BNT-level equivalence remains open.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
