\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\newcommand{\MPV}[2]{V^{(#1)}\!\left(#2\right)}

\title{The Length-Zero (``Zero-Tail'') Bookkeeping in the\\
  Blocked Canonical-Form Decomposition}
\author{}
\date{2026-05-30}

\begin{document}
\maketitle

\noindent
This note explains why carrying the blocked canonical-form decomposition as an
\emph{all-length} matrix-product-vector identity is the wrong formal shape, and
records the clean replacement together with the one place where the formal
development still keeps the awkward form.\footnote{%
  Context: PRs \ghpr{2153}, \ghpr{2156}, and \ghpr{2167} removed roughly a
  thousand lines of ``zero-tail'' plumbing built on the all-length form
  discussed here. This note records why that form is a non-source-faithful
  overcomplication and what remains.}

\section{The assertion}

Let $A$ be a translation-invariant matrix-product tensor of bond
dimension~$D$, with one-site matrices $A^i\in M_D(\mathbb C)$. For a length
$N$ and a word $\sigma=(i_1,\dots,i_N)$ write the matrix-product-vector
coefficient
\begin{align}
  \MPV{N}{A}_\sigma = \tr\!\left(A^{i_1}\cdots A^{i_N}\right).
\end{align}
The blocked canonical-form reduction of
\cite[Section~2.3]{Cirac2016MPDO_arXiv}, following
\cite[Theorem~3]{PerezGarcia2007MPSReps}, states that after blocking by some
period $p\ge 1$ the tensor is, up to a gauge, the direct sum of an all-zero
block and a weighted family of normal blocks,
\begin{align}
  A^{[p]} \cong  \mathbf 0_{z}\ \oplus\ \bigoplus_{k} \mu_k\,B_k,\label{eq:zero-block}
\end{align}
where $\mathbf 0_{z}$ is the zero tensor on a bond space of dimension~$z$,
the $B_k$ are normal (primitive, trace-preserving, irreducible) blocks of
bond dimension $D_k$, the weights satisfy $\mu_k\neq 0$, and
$z+\sum_k D_k = D$ is the bond dimension of $A^{[p]}$ in the formal blocking
convention.

\section{The point at issue}

The direct sum~\eqref{eq:zero-block} gives, at every length $N$ and word
$\sigma$,
\begin{align}
  \MPV{N}{A^{[p]}}_\sigma
  = \MPV{N}{\mathbf 0_{z}}_\sigma
  + \MPV{N}{\textstyle\bigoplus_k \mu_k B_k}_\sigma.\label{eq:all-length}
\end{align}
The zero block contributes nothing except at the empty word: for $N\ge 1$ a
product of $A^i$-blocks that passes through the zero summand vanishes, so
\begin{align}
  \MPV{N}{\mathbf 0_{z}}_\sigma = 0 \qquad (N\ge 1),
\end{align}
whereas at $N=0$ the only coefficient is the trace of the identity on the
bond space,
\begin{align}
  \MPV{0}{C}_{\varnothing} = \tr(\mathbf 1_{D_C}) = D_C.
\end{align}
Hence the \emph{entire} content of the length-zero instance
of~\eqref{eq:all-length} is the dimension count
\begin{align}
  z + \sum_k D_k = D,\label{eq:dim}
\end{align}
which is already part of~\eqref{eq:zero-block}. Every length $N\ge 1$ carries
the genuine mathematical content: the weighted normal-block comparison
$\MPV{N}{A^{[p]}}_\sigma = \MPV{N}{\bigoplus_k \mu_k B_k}_\sigma$.

\noindent\emph{Why the all-length form is the wrong shape.}
Stating the reduction as the all-length identity~\eqref{eq:all-length}
attaches the trivial dimension fact~\eqref{eq:dim} to the substantive
positive-length comparison and forces both to travel together through blocking,
reindexing, and sector grouping. In the formal development this produced a
family of transport lemmas whose only purpose was to move the empty-word
instance of~\eqref{eq:all-length} across each such step.\footnote{%
  These were the \leanid{zeroTail_*} lemmas in
  \path{TNLean/MPS/CanonicalForm/SectorComparison/}; they are removed in
  PRs \ghpr{2156} and \ghpr{2167} once the comparison is stated at positive
  length.}
This is also not source-faithful: the article introduces the zero block once,
to account for the bond dimension, and then discards it, comparing only the
nonzero normal blocks in the thermodynamic limit. The zero summand
$\mathbf 0_z$ is an artifact of keeping a common bond space in the formal
statement; it carries no state and no normal-block structure.

\section{The clean statement}

Separate the two facts. The positive-length comparison
\begin{align}
  \MPV{N}{A^{[p]}}_\sigma
  = \MPV{N}{\textstyle\bigoplus_k \mu_k B_k}_\sigma
  \qquad (N\ge 1)\label{eq:pos}
\end{align}
carries all of the mathematics, and the length-zero coefficient is recovered
\emph{once}, where an all-length statement is genuinely needed, from the single
dimension identity~\eqref{eq:dim}: if two tensors agree at all positive lengths
and have equal bond dimension, they agree at length zero as well, because the
empty-word coefficient is the bond dimension.\footnote{%
  Formally, positive-length equality is \leanid{SameMPV₂Pos} and the one-step
  upgrade is
  \leanid{SameMPV₂Pos.toSameMPV₂_of_bondDim_eq} in
  \doclink{TNLean/MPS/Defs}. The canonical-form predicate now stores
  \leanid{SameMPV₂Pos} together with a single bond-dimension field rather than
  the all-length decomposition; see \ghpr{2153}.}
This is the source-faithful shape: the nonzero normal blocks are compared at
positive length, and the zero block appears only as the scalar~$z$
in~\eqref{eq:dim}.

\section{Formal status}

\leanid{exists_tp_primitive_blockDecomp_after_blocking}, blueprint
\path{thm:tp_primitive_blockdecomp} in
\path{blueprint/src/chapter/ch11b_after_blocking.tex}, now uses the
source-faithful shape: its MPV clause is the positive-length comparison
in~\eqref{eq:pos}, and the empty-word contribution is the separate identity
\(z+\sum_k D_k=D\).  The former support lemma
\leanid{zeroTail_toTensorFromBlocks_blockPower} has therefore been removed.

\section{Conclusion}

The blocked canonical form~\eqref{eq:zero-block} is correct, and the dimension
count~\eqref{eq:dim} is part of it. The former formal defect was one of shape:
stating the reduction as the all-length matrix-product-vector
identity~\eqref{eq:all-length} fused a trivial empty-word dimension count with
the substantive positive-length comparison. The faithful statement is the
positive-length comparison~\eqref{eq:pos} with the length-zero coefficient
restored once from~\eqref{eq:dim}. The development now uses this shape in the
after-blocking primitive decomposition and its downstream consumers.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
