\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Per-block RFP Isometry Form and Cross-Block Orthogonality in CPSV16}
\author{}
\date{2026-06-14}

\begin{document}
\maketitle

\section{Source statement}

The pure-state RFP characterization in Section~3 of
\cite{Cirac2016MPDO_arXiv} writes a canonical-form RFP tensor as
\begin{align}
  A^i
    =
  \bigoplus_{j=1}^g\bigoplus_{q=1}^{r_j}
  \mu_{j,q}X_{j,q}\Lambda_jU_j^iX_{j,q}^{-1},
  \tag{\path{III_CFI_RFP}}
\end{align}
where the matrices $\Lambda_j$ are diagonal, positive, and trace-normalized.
The same theorem requires a joint isometry condition
\begin{align}
  \sum_{i=1}^s
  (U_j^i)_{\alpha,\beta}\,
  \overline{(U_{j'}^i)_{\alpha',\beta'}}
  =
  \delta_{j,j'}\delta_{\alpha,\alpha'}\delta_{\beta,\beta'}.
  \tag{\path{eq:III_isometry}}
\end{align}
The factor $\delta_{j,j'}$ is a cross-block orthogonality condition: for
$j\ne j'$, the physical-index isometries associated with the two normal-tensor
blocks are mutually orthogonal.  Corollary~III.cor3 invokes this same isometry
condition for the BNT elements.\footnote{Local source:
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:543--554} and
\path{Papers/1606.00608/MPDO-22-12-17-2.tex:583--589}.}

\section{Formal boundary}

The current formal theorem applies the single-block isometry form separately to
each block of an indexed family.\footnote{Formal declaration:
\leanid{MPSTensor.rfp_nt_structural_full_blocks} in
\doclink{TNLean/MPS/RFP/StructuralFull}.  Introduced in \ghpr{2785}.}  For each
block $k$ it gives
\begin{align}
  A_k^i
    &=
  X_k\Lambda_kU_k^iX_k^{-1},\\
  \sum_i (U_k^i)^\dagger U_k^i
    &= I.
\end{align}
There is no conclusion relating $U_k$ and $U_\ell$ for $k\ne \ell$.  Thus the
formal theorem records a separate isometry condition for each block, but it does
not record the cross-block equations
\begin{align}
  \sum_i
  (U_k^i)_{\alpha,\beta}\,
  \overline{(U_\ell^i)_{\alpha',\beta'}}
  =
  0
  \qquad (k\ne \ell),
\end{align}
which are part of the source condition \path{eq:III_isometry}.

\section{Elimination plan}

A source-faithful version of Corollary~III.cor3 should be stated for a
BNT-family or whole-tensor canonical-form datum, not merely for independent
blocks.  Its conclusion should include a joint family $U$ satisfying
\begin{align}
  \sum_i
  (U_k^i)_{\alpha,\beta}\,
  \overline{(U_\ell^i)_{\alpha',\beta'}}
  =
  \delta_{k,\ell}\delta_{\alpha,\alpha'}\delta_{\beta,\beta'}.
\end{align}
Such a statement must account for the common physical space into which all
blocks embed; in particular, the joint isometry condition has the dimensional
content that the block-pair spaces fit orthogonally inside the physical space.

\section{Conclusion}

The current theorem is a scope restriction.  It is a useful per-block
consequence of the single-block RFP isometry form, but it is not the full
isometry assertion of Corollary~III.cor3 because it omits the cross-block
$\delta_{j,j'}$ orthogonality equations.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
