\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{The Direct-Sum Input from the Older MPS Representation Paper}
\author{}
\date{2026-05-05}

\begin{document}
\maketitle

\section{Source statement}

The finite-length direct-sum input used behind the block-injective
canonical-form route is not proved in the 2016 MPDO source.  The 2016 paper
cites the older representation paper for the block-injective-after-blocking
step \cite[lines~340--345]{Cirac2016MPDO_arXiv}.  The local source for the
older argument is \path{Papers/quant-ph_0608197/MPSarchive.tex}.

That source first introduces the injectivity condition \(C1\): for some
\(L_0\), the map
\[
  \Gamma_{L_0}(X)
    =
    \sum_{i_1,\ldots,i_{L_0}}
      \operatorname{tr}(X A_{i_1}\cdots A_{i_{L_0}})
      |i_1,\ldots,i_{L_0}\rangle
\]
is injective.  It immediately identifies this with the exact-length word
products of length \(L_0\) spanning the full matrix algebra
\cite[lines~888--908]{PerezGarcia2007MPSReps}.

The technical algebraic lemma then says that, for nonzero \(D\times D\)
matrices \(C\), every matrix \(X\) is a finite sum
\[
  X = \sum_i R_i C S_i .
\]
The proof reduces \(C\) to a matrix unit by elementary matrix operations and
then reconstructs \(X\) from permuted copies of that matrix unit
\cite[lines~1333--1344]{PerezGarcia2007MPSReps}.

The next lemma is the direct-sum statement.  Its local hypotheses are important:
the tensor has \(b \ge 2\) canonical blocks \(A^1,\ldots,A^b\), each block
satisfies \(C1\), the common length is
\(L_0=\max_j L_0^{A^j}\), the block sizes are ordered, and the block states are
assumed pairwise different
\cite[lines~1321--1330]{PerezGarcia2007MPSReps}.  Under these hypotheses, if
\(L \ge 3(b-1)(L_0+1)\), then
\[
  \sum_{j=1}^b \mathcal G_L^{A^j}
\]
is a direct sum \cite[lines~1346--1408]{PerezGarcia2007MPSReps}.  The proof is
by induction on the number of blocks.  The base case uses \(C1\) and the
matrix-factorization lemma to show that a nontrivial linear dependence would
force inclusion, then equality, of the two length-\(L_0+1\) MPS subspaces,
contradicting uniqueness of the associated injective parent-space
characterization.  The induction step isolates one block with a local
projector, applies the induction hypothesis to the remaining blocks, and then
uses \(C1\) plus the factorization lemma to force the remaining coefficient to
vanish.

\section{MPS translation}

For the non-periodic Fundamental Theorem route, the mathematical consequence is
homogeneous and positive-length.  After the canonical-form and
basis-of-normal-tensors reduction, the selected representatives should admit one
length at which their simultaneous word evaluations separate the direct-sum
blocks.  In trace-dual form, if
\[
  \sum_{j=1}^g \operatorname{tr}(\Delta_j A_j^w)=0
  \qquad\text{for every word } w \text{ of length } L,
\]
then each \(\Delta_j\) is zero.  This is the finite-length direct-sum span input
needed by the BNT pair-separation route.

This statement is not a consequence of the empty word and is not a cumulative
span statement.  The exact length \(L\) is part of the content.  Therefore the
formal proof should not replace the source input by a hypothesis that only says
the cumulative pair algebra is full, nor by a theorem that uses the length-zero
identity as padding.

\section{Formal boundary}

The current formal development now formalizes the matrix-factorization input
from Lemma~\texttt{lem1}: \leanid{Matrix.span_range_mul_nonzero_mul_eq_top}
states that \(\operatorname{span}\{RCS\}=M_D(\mathbb C)\) whenever
\(C\ne0\), and
\leanid{MPSTensor.span_range_evalWord_mul_nonzero_mul_evalWord_eq_top}
translates this into the exact-length word-product form
\(\operatorname{span}\{A_u X A_v: |u|=|v|=L\}=M_D(\mathbb C)\) under
\(L\)-block injectivity and \(X\ne0\).

It also formalizes the next finite-dimensional part of the two-block
direct-sum proof.  The trace-dual three-block reduction gives inclusion of the
length-\(L\) image spaces
\(\mathcal G_L^A\subseteq\mathcal G_L^B\).  Since block injectivity makes
\(\dim\mathcal G_L^A=D_A^2\) and \(\dim\mathcal G_L^B=D_B^2\), this inclusion
and the source-paper size ordering \(D_B\le D_A\) force \(D_A=D_B\) and
\(\mathcal G_L^A=\mathcal G_L^B\).  This is recorded by
\leanid{MPSTensor.bondDim_eq_and_groundSpace_eq_of_three_block_trace_relation_left_of_dim_ge}.
The strict-size branch is also formalized:
\leanid{MPSTensor.not_three_block_trace_relation_left_of_isNBlkInjective_of_dim_gt}
states that, if \(D_B<D_A\), a nonzero coefficient on the larger block cannot
satisfy the homogeneous three-block trace relation.
\leanid{MPSTensor.pairTraceSeparatingAt_threeBlock_of_isNBlkInjective_of_dim_gt}
is the corresponding trace-separation form: once the smaller block is also
block-injective at the three-block length, the strict-size branch gives
homogeneous pair trace separation at that exact length.
\leanid{MPSTensor.pairTraceSeparatingAt_threeBlock_of_isNBlkInjective_of_dim_ne}
removes the ordering convention, using the symmetry of pair trace separation.
The conditional directness package
\leanid{MPSTensor.groundSpace_inf_eq_bot_of_not_bondDim_eq_and_groundSpace_eq_of_dim_ge}
turns any external proof of the source contradiction
\(\neg(D_A=D_B \wedge \mathcal G_L^A=\mathcal G_L^B)\) into
\(\mathcal G_{3L}^A\cap\mathcal G_{3L}^B=\{0\}\).
The equal-size uniqueness branch is now isolated as well:
\leanid{MPSTensor.not_bondDim_eq_and_groundSpace_eq_of_mpvSubmodule_ne}
shows that, for injective tensors, equal local image spaces would force equality
of the periodic MPV lines, so a separately supplied distinct-MPV-line hypothesis
rules out the equal-size collapse.
\leanid{MPSTensor.groundSpace_inf_eq_bot_of_exists_not_forall_mpv_eq_mul_of_dim_ge}
combines this uniqueness input with the finite-dimensional step in the form
used downstream: the caller supplies some chain length \(N\ge \max\{2,L\}\) at
which the two MPV states are not pointwise proportional.
\leanid{MPSTensor.exists_ge_not_forall_mpv_eq_mul_of_not_gaugePhaseEquiv_of_irreducible_TP}
proves the analytic source of such witnesses for irreducible trace-preserving
blocks with normalized self-overlaps: if two blocks are not gauge-phase
equivalent, then arbitrarily far out there is a chain length at which their MPV
states are not proportional.
\leanid{MPSTensor.exists_ge_not_forall_mpv_eq_mul_of_blocksNotGaugePhaseEquiv_of_irreducible_TP}
is the corresponding already-blocked BNT-family wrapper for same-dimensional
distinct blocks.  The direct-sum file records the pointwise conversion by
\leanid{MPSTensor.groundSpace_inf_eq_bot_of_exists_not_forall_mpv_eq_mul_of_dim_ge}.
\leanid{MPSTensor.groundSpace_inf_eq_bot_of_blocksNotGaugePhaseEquiv_same_dim_of_dim_ge}
then packages these two ingredients for the same-dimensional equal-size
direct-sum branch: BNT separation supplies the long-chain non-proportional MPV
witness, while injectivity and length-\(L\) block injectivity are still stated
as explicit direct-sum hypotheses.
\leanid{MPSTensor.pairTraceSeparatingAt_of_groundSpace_inf_eq_bot_of_isNBlkInjective}
turns zero intersection of the two image spaces at a fixed length into
homogeneous pair trace separation, provided the two boundary-to-chain maps are
injective at that fixed length.  Hence
\leanid{MPSTensor.pairTraceSeparatingAt_threeBlock_of_exists_not_forall_mpv_eq_mul_of_dim_ge}
and
\leanid{MPSTensor.pairTraceSeparatingAt_threeBlock_of_blocksNotGaugePhaseEquiv_same_dim_of_dim_ge}
record the pair-trace versions consumed by selector and word-span
constructions.
\leanid{MPSTensor.forall_pairTraceSeparatingAt_threeBlock_of_blocksNotGaugePhaseEquiv}
combines the same-dimension BNT branch with the strict-size branch for all
ordered pairs of distinct blocks at the common length \(3L\), and
\leanid{MPSTensor.exists_pos_productWordSpan_of_isCanonicalFormBNT_of_directSum_threeBlock}
feeds that common pair separation into the selector/product-span construction.
The already-injective BNT surface is now closed by the later direct-sum wrappers
\leanid{MPSTensor.forall_pairTraceSeparatingAt_of_isCanonicalFormBNT_of_directSum_injectiveBlocks},
\leanid{MPSTensor.exists_pos_productWordSpan_of_isCanonicalFormBNT_of_directSum_injectiveBlocks}, and
\leanid{MPSTensor.exists_wordTupleSpanTop_of_isCanonicalFormBNT_of_directSum_injectiveBlocks}:
when the selected BNT blocks are already injective at the required common
length, the exact-length direct-sum span is a theorem rather than a remaining
identity-padding assumption.

The current formal development also includes downstream lemmas that use
the direct-sum input.
\begin{itemize}
  \item \leanid{WordTupleSpanTop} expresses exact-length span of a direct-sum
        block family.
  \item \leanid{PairWordTupleSpanTop} and \leanid{PairTraceSeparatingAt}
        express the two-block trace-dual form used in the pair route.
  \item The period-window theorems in
        \path{TNLean/MPS/MPDO/BiCFDerivation/PairHomogenization.lean} turn a
        positive homogeneous window of \leanid{PairWordTupleSpanTop} data into
        homogeneous trace separation.
  \item The BNT-level reformulations in
        \path{TNLean/MPS/CanonicalForm/BlockDiagonalCommutant.lean} turn
        pairwise period-window data into a common \leanid{WordTupleSpanTop}
        length for a canonical-form BNT family.
\end{itemize}

\section{Conclusion}

The exact-length direct-sum input from the older representation paper has now
been formalized for the already-injective BNT theorem surface in the
one-site-injective specialization used by the current canonical-form API.  In
that setting, the BNT representatives, their pairwise not-gauge-phase-equivalence
hypotheses, and the fixed lengths supplied by one-site injectivity imply the
homogeneous pair-trace separation and product-word-span conclusions needed
downstream.  This does not claim the full paper statement with an arbitrary
pre-existing injectivity length \(L_0\), nor does it reproduce the paper's
ordered block-size hypotheses verbatim; those features are replaced here by the
\leanid{IsCanonicalFormBNT} and irreducible-block assumptions of the formal
surface.  The old warning against replacing exact-length span by cumulative span
therefore remains mathematically important, but it is no longer an open gap for
this already-injective \leanid{IsCanonicalFormBNT} specialization.

For the full after-blocking normal-CF-BNT pipeline, the conclusion is still
conditional.  The bounded and common injective-blocking results now supply the
Wielandt part of the input, but the final CPSV theorem is not yet unconditional:
the proof still has to transport the blocked BNT separation hypotheses through
the chosen reblocking/reindexing maps and feed the blocked not-GPE facts into
\leanid{CommonPrimitiveBNTCoverHypotheses}.  That transport is the remaining
bridge between the already-injective direct-sum theorem surface and the
canonical-form after-blocking comparison.

Accordingly, the blueprint should present the David/P\'erez-Garc\'ia
exact-length direct-sum input as formalized for the already-injective BNT
specialization, while keeping the after-blocking CPSV statement visibly
conditional until the reblocking transport and blocked BNT-separation handling
have landed.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
