\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Common Blocking and Span Equality in the Non-periodic MPS Fundamental Theorem}
\author{}
\date{2026-04-30}

\begin{document}
\maketitle

\section{The compressed source argument}

The proof of the non-periodic matrix product state fundamental theorem passes
from arbitrary tensors with the same matrix product vectors to
basis-of-normal-tensors comparison data. The source argument is the MPDO
canonical-form paper \cite{Cirac2016MPDO_arXiv,Cirac2017MPDO_AnnPhys}, together
with the later review \cite{Cirac2021Matrix}.\footnote{For traceability, this
note corresponds to \href{https://github.com/LionSR/TNLean/issues/1047}{issue \#1047}.}
The cited papers are not being questioned. The point is that a formal proof has
to separate several canonical-form steps which are compressed in the source
prose.

For a tensor $A=(A^i)_{i=1}^d$, the source paper defines the matrix product
vectors by
\[
  |V^{(N)}(A)\rangle
  = \sum_{i_1,\ldots,i_N=1}^d
      \operatorname{tr}(A^{i_1}\cdots A^{i_N})
      |i_1\rangle\otimes\cdots\otimes |i_N\rangle,
\]
for $N=1,2,3,\ldots$ \cite[Definition~\texttt{MPV}, lines~130--138]{Cirac2016MPDO_arXiv}.
After removing invariant subspaces, the same paper writes the tensor in block form
\[
  A^i = \bigoplus_{k=1}^r \mu_k A_k^i,
\]
allowing zero blocks and scaling the nonzero blocks so that their transfer maps
have spectral radius one. Each nonzero irreducible block may still have peripheral
period. Blocking by a common multiple of the periods removes that periodicity
\cite[lines~214--231]{Cirac2016MPDO_arXiv}. The review follows the same
canonical-form argument \cite[lines~1798--1820]{Cirac2021Matrix}.

The later comparison is made through bases of normal tensors. A basis of normal
tensors is a minimal family $(A_j)_{j=1}^g$ of normal tensors such that every
matrix product vector of $A$ is a linear combination of the matrix product vectors
of the $A_j$, and such that those vectors are eventually linearly independent.
Equivalently, every normal block in the canonical form is a phase times a gauge
transform of exactly one basis tensor, with the family chosen minimally
\cite[Proposition~\texttt{prop:char-BNT}, lines~271--280]{Cirac2016MPDO_arXiv}.
The review states the same characterization
\cite[Proposition~\texttt{prop:char-BNT}, lines~1846--1859]{Cirac2021Matrix}.
For a tensor in canonical form one then has an expansion
\[
  |V^{(N)}(A)\rangle
    = \sum_{j=1}^g \left(\sum_{q=1}^{r_j} \mu_{j,q}^N\right)
        |V^{(N)}(A_j)\rangle.
\]
This is the basis-of-normal-tensors expansion used in the equal-case proof
\cite[lines~283--302]{Cirac2016MPDO_arXiv}. The review uses the same expansion
\cite[lines~1864--1885]{Cirac2021Matrix}.

Injectivity enters at a different point. The source first defines injective and
block-injective canonical form, and then invokes the quantum Wielandt theorem to
say that, after a bounded further blocking, every normal tensor becomes injective
\cite[lines~317--332]{Cirac2016MPDO_arXiv}. The review records the corresponding
$3D^5$ bound for block-injective canonical form
\cite[lines~2120--2124]{Cirac2021Matrix}.
This step is not the same as period removal: period removal gives primitive normal
blocks, while the overlap-rigidity arguments used later require one-site
injectivity, or an explicit replacement after further blocking.

Finally, the source comparison for the equal case is compact. Equality or
proportionality of the matrix product vector families first matches the two bases
of normal tensors by phase and gauge. After relabeling, the coefficient
multiplicities are compared through the power sums
\[
  \sum_{q=1}^{r_{a,j}} \mu_{j,q}^N
  = \sum_{q=1}^{r_{b,j}} (\nu_{j,q}e^{i\phi_j})^N,
  \qquad \text{for all } j \text{ and } N.
\]
The elementary power-sum lemma then recovers the multiplicities and the weights,
and the blockwise gauges are assembled into the global conjugating matrix
\cite[lines~1181--1192]{Cirac2016MPDO_arXiv}. The review states the resulting
proportional and equal MPV fundamental theorems in the same form
\cite[Theorem~\texttt{thm1} and Corollary~\texttt{II\_cor2}, lines~1891--1900]{Cirac2021Matrix}.

\section{The formal argument}

The formal equality relation \(\mathrm{SameMPV}_2\) compares matrix product
vector coefficients for all finite words, including the empty word.
This convention is stronger than the positive-length family in the source
definition. An all-zero block of bond dimension $D_0$ contributes nothing at
positive length, but it contributes $D_0$ at length zero, because the empty product
has trace $D_0$. Consequently the formal proof separates the zero-tail
contribution from the nonzero part, proves positive-length equality for the
nonzero parts, and records the exact $N=0$ identity separately.\footnote{See
\path{blueprint/src/chapter/ch08_canonical.tex:967--1022} and
\path{blueprint/src/chapter/ch08_canonical.tex:3036--3164}, with labels
\texttt{def:zero\_mps\_tensor}, \texttt{thm:mpv\_zero\_tensor},
\texttt{thm:zero\_block\_separation},
\texttt{thm:nonzero\_block\_block\_power\_zero\_tail\_identity}, and
\texttt{thm:nonzero\_block\_same\_mpv\_zero\_tail\_eq}. The corresponding
zero-block cancellation step in the final comparison chapter is
\path{blueprint/src/chapter/ch11_fundamental_theorem_proof.tex:1202--1217}.}

The common blocking step is also more explicit. For a single family of blocks,
the source phrase ``block by the least common multiple of the periods'' hides no
mathematical difficulty. In the two-tensor comparison, however, the formal proof
must choose one positive physical blocking length for both sides. For each
original block with period-removal length $m_k$, the common length is written as
$p=m_k e_k$. The tensor obtained by first blocking by $m_k$ and then by $e_k$
has physical labels given by nested words, while the tensor obtained by
blocking once by $p$ has physical labels in the alphabet of words of length $p$.
The formal statement therefore keeps the relabeling between these two alphabets as
part of the statement.\footnote{The common-sector data and the relabeling results
are recorded in \path{blueprint/src/chapter/ch08_canonical.tex:1645--1824} and
\path{blueprint/src/chapter/ch08_canonical.tex:3230--3284}. The main formal
statement is
\texttt{MPSTensor.fundamentalTheorem\_after\_blocking\_commonLength\_commonSector},
with the one-sided relabeling-compatible zero-tail statement
\texttt{MPSTensor.zeroTail\_commonFlat\_of\_reindexed}; see
\path{TNLean/MPS/CanonicalForm/SectorComparison/CommonSectorData.lean} and
\path{TNLean/MPS/CanonicalForm/SectorComparison/CommonSectorTransport.lean}.}
This is why the canonical-form theorem supplies relabeled common-sector
families and a conditional label-compatibility statement, rather than immediately
replacing every canonical nonzero part after blocking by its flattened sector tensor.

The formal comparison theorems downstream use primitive overlap-rigidity data for
sector bases. Those data include equality of the finite-length spans of the basis
matrix product vectors and one-site injectivity of the basis blocks. Thus the
additional Wielandt blocking needed for injectivity remains a separate refinement
after period removal. Treating it separately prevents a common period-removal
length from being used as though it already supplied the injective hypotheses
needed by the overlap-span theorem.\footnote{The span hypotheses are stated in
\path{blueprint/src/chapter/ch11_fundamental_theorem_proof.tex:590--602}, label
\texttt{def:sector\_basis\_overlap\_span\_hyps}. The remaining canonical
proof obligations are summarized in
\path{blueprint/src/chapter/ch08_canonical.tex:3289--3341}.}

The paper's compact BNT matching and power-sum comparison is represented in the
formal statement by common phase-cover and finite-length span-equality adapters. A
common phase cover consists of a common family of nonzero blocks, surjective maps
from the two block families onto that common family, and phase equivalences between
each block and its common representative. Such data imply that, for every system
size, the spans of the two nonzero-block MPV families are equal. When a BNT
proportional-decomposition comparison has already produced a permutation, matched
bond dimensions, and gauge phases, the formal construction gives the common phase
cover from that comparison and then obtains the span equality needed by the sector
comparison theorem.\footnote{See
\path{blueprint/src/chapter/ch11_fundamental_theorem_proof.tex:755--923}. The corresponding formal
objects include the common phase-cover structure and span lemmas, the block-span
and common-phase-cover overlap-span adapters, and the proportional-decomposition
specialization. They are located at
\path{TNLean/MPS/CanonicalForm/PhaseCover.lean},
\path{TNLean/MPS/CanonicalForm/PhaseClassSectorData.lean}, and
\path{TNLean/MPS/CanonicalForm/SectorComparison/CommonPrimitiveProportionalData.lean};
the main adapter names have prefix
\texttt{MPSTensor.exists\_bnt\_sectorDecomp\_pair\_with\_overlapSpan\_of\_}
and suffixes \texttt{block\_span\_eq}, \texttt{commonPhaseCover}, and
\texttt{proportionalDecompositionConclusion}.}
These lemmas state precisely what comparison data are needed; they do not by
themselves derive all of that data from an arbitrary equality of total MPV
families.

\section{Why these distinctions belong together}

These distinctions are coupled in the single conversion from an arbitrary equality
of MPV families to the hypotheses of the BNT sector-comparison theorem. The
zero-tail convention determines what is known at length zero after each blocking.
The common blocking construction determines the physical alphabet and the relabeled
sector families to which the later comparison must apply. The injectivity
refinement may further change the final block length before the overlap-rigidity
hypotheses are available. The common phase-cover or BNT proportional-decomposition
data then have to be constructed for those final nonzero-sector families, not for
an earlier family before relabeling or before injectivity blocking.

For this reason common blocking and finite-length span equality should be kept
in one account. The formal proof needs the whole chain: remove the zero tail,
choose one common physical blocking length, account for the word-relabeling, keep
the later injectivity blocking separate, and then turn BNT matching data into the
finite-length nonzero-sector span equality used by the sector comparison theorem.
The blueprint states this chain as the remaining connection between the
canonical-form chapter and the final comparison chapter.\footnote{See the
``Remaining hypotheses'' discussion in
\path{blueprint/src/chapter/ch11_fundamental_theorem_proof.tex:1277--1341}. Audit notes recording
the same remaining obligations include
\path{audits/2026-04-28_issue969_direct_common_blocking.md:139--153},
\path{audits/2026-04-28_issue970_common_phase_cover.md:81--95},
\path{audits/2026-04-29_issue942_common_exact_nonzero_sector.md:66--97}, and
\path{audits/2026-04-29_issue944_overlap_span_from_common_live.md:83--94}.}

This note should also not be read as a claim that the same-side equal-norm scalar
obstruction discussed in the broader fundamental-theorem work is a flaw in the
CPSV or CPGSV arguments. That obstruction corrected an over-strong formal statement
which tried to identify an equal-norm class with a single basis tensor. The BNT
expansion in the papers already allows multiple basis tensors and multiplicities.
The formal argument therefore moved from naive norm-class grouping to phase-class
sector decompositions and common-cover data.

\section{Conclusion}

The conclusion is that the cited papers compress several standard canonical-form
operations; no counterexample is claimed. Zero or nilpotent contributions are
invisible to positive-length MPVs, periods are removed by blocking, injectivity is
obtained by a later Wielandt-type blocking, BNT bases are matched by phase and
gauge, and coefficient multiplicities are recovered from power sums. The formal
development exposes these operations as separate statements because the formal MPV
relation includes the empty word, the two sides must be placed over one explicitly
relabeled alphabet after blocking, and the downstream comparison theorem requires
finite-length span equality and injective final blocks.

Thus the comparison gives neither a counterexample nor a theorem-level source
error. It explains why the formal proof cannot compress the argument into a
single invocation of the BNT uniqueness theorem. The relevant comparison points are
zero-tail cancellation, common primitive block construction, common blocking,
finite-length span equality, zero-tail transport, overlap-span hypotheses, and
iterated-blocking relabeling.\footnote{For traceability, these are recorded in
\href{https://github.com/LionSR/TNLean/issues/652}{issue \#652},
\href{https://github.com/LionSR/TNLean/issues/942}{issue \#942},
\href{https://github.com/LionSR/TNLean/issues/969}{issue \#969},
\href{https://github.com/LionSR/TNLean/issues/970}{issue \#970},
\href{https://github.com/LionSR/TNLean/issues/971}{issue \#971},
\href{https://github.com/LionSR/TNLean/issues/944}{issue \#944}, and
\href{https://github.com/LionSR/TNLean/issues/990}{issue \#990}.}

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
