\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{docs/paper-gaps/command}

\title{Paper Comparison for BNT Inputs in the Non-periodic MPS Fundamental Theorem}
\author{}
\date{2026-05-06}

\begin{document}
\maketitle

\section{Source passages and mathematical content}

The non-periodic matrix product vector Fundamental Theorem in
\cite{Cirac2016MPDO_arXiv,Cirac2017MPDO_AnnPhys,Cirac2021Matrix} is a theorem
about tensors already put in canonical form. The comparison is not stated as a
theorem about arbitrary block-diagonal data plus unrelated finite-length
hypotheses. The finite-length hypotheses now isolated in the formalization are
not separate assumptions in the paper theorem. They are names for the
mathematical ingredients supplied by the source argument: canonical-form
reduction, BNT selection, fixed-length block-injective span after blocking,
phase-gauge matching, and power-sum comparison.

This note uses the paper source, not issue numbers, as the comparison standard.
The formal issue trackers are useful for scheduling work, but the mathematical
boundary is the following one.  A conditional formal theorem matches the source
only when its extra hypotheses are exactly one of the source ingredients listed
below, or when the text states that the source ingredient has not yet been
proved.

\begin{description}
  \item[Positive-length MPVs.]
  The paper works with
  \[
    \langle i_1,\ldots,i_N|V^{(N)}(A)\rangle
      = \operatorname{tr}(A^{i_1}\cdots A^{i_N}),
    \qquad N=1,2,\ldots .
  \]
  A formal theorem should not use the empty word as a paper-level hypothesis or
  conclusion.

  \item[Canonical form.]
  The source first reaches
  \[
    A^i=\bigoplus_k \mu_k A_k^i,\qquad \sum_k D_k\le D,
  \]
  and then removes periods by blocking.  The formal route should therefore keep
  zero blocks and period-removal data separate from later injectivity blocking.

  \item[BNT decomposition.]
  The source writes
  \[
    A^i=X\left[\bigoplus_j(M_j\otimes A_j^i)\right]X^{-1},
  \]
  and
  \[
    |V^{(N)}(A)\rangle
      =\sum_j\left(\sum_q\mu_{j,q}^N\right)|V^{(N)}(A_j)\rangle .
  \]
  The comparison is between the selected BNT representatives, not the raw
  cyclic-sector or repeated-copy lists produced on the way to them.

  \item[Block-injective span.]
  After the cited blocking step, every matrix in
  \[
    \bigoplus_jM_{D_j}(\mathbb C)
  \]
  is a linear combination of the block-evaluations \(\tilde A^i\).  The formal
  proof must prove this homogeneous fixed-length direct-sum span, or keep it
  explicitly as a hypothesis.

  \item[Equal-case weights.]
  Once the BNT representatives have been matched, the source compares
  \[
    \sum_q\mu_{j,q}^N
      =
    \sum_q(\nu_{j,q}e^{i\phi_j})^N
  \]
  for all matched \(j\) and \(N\), and only then applies the finite power-sum
  lemma.
\end{description}

\subsection{Verdict}

There is no mathematical discrepancy with the source theorem if the conditional
formal hypotheses are read as names for source inputs that still have to be
produced from the canonical-form proof.  The source theorem assumes two tensors
already in canonical form and then compares their bases of normal tensors.  It
does not assume arbitrary phase covers, raw cyclic-sector lists, cumulative
pair-span fullness, or length-zero matrix product vector equality.

The present formal boundary is therefore a proof gap, not a different theorem.
The remaining inputs are the representative-family BNT cover, the exact-length
direct-sum span supplied by block-injective canonical form, and the transport of
those data through the common blocking and word-reindexing steps.  Until these
inputs are proved, the downstream after-blocking theorem should stay visibly
conditional.

The phrase ``zero tail'' is formalization terminology.  The 2016 source says
only that, after extracting the nonzero invariant blocks,
\(\sum_k D_k \le D\), and explains this as the possible presence of zero blocks
\cite[lines~214--219]{Cirac2016MPDO_arXiv}.  Thus a zero tail in the formal
text means the unused all-zero block summand of dimension
\(D - \sum_k D_k\).  It is not a paper term and should not be used without this
definition.

\subsection{Source label map}

The comparison below uses the labels in the local 2016 source wherever the
paper gives one:
\[
  A^i = \sum_{k=1}^r P_k B^i P_k
      = \bigoplus_{k=1}^r \mu_k A_k^i
  \tag{\texttt{eq:II\_Aiplusk1}}
\]
is the canonical-form reduction; the following source sentence adds
\(\sum_k D_k \le D\), explaining the possible all-zero leftover block.  The
canonical-form display is the equation labelled \texttt{eq:II\_CF1}:
\[
  A^i = \bigoplus_{k=1}^r \mu_k A_k^i
  \tag{\texttt{eq:II\_CF1}}
\]
The phase-gauge relation used in the BNT characterization is labelled
\texttt{eq:II:A=XAX}:
\[
  \tilde A_k = e^{i\phi_k} X_k A_j X_k^{-1}
  \tag{\texttt{eq:II:A=XAX}}
\]
The BNT characterization is \texttt{prop:char-BNT}, and the BNT decomposition is
the subequation labelled \texttt{eq:II\_ABasicTensors}:
\[
  A^i =
  X\left[\bigoplus_{j=1}^g(M_j\otimes A_j^i)\right]X^{-1}
  =
  \bigoplus_{j=1}^g\bigoplus_{q=1}^{r_j}
    \mu_{j,q}X_{j,q}A_j^iX_{j,q}^{-1}
  \tag{\texttt{eq:II\_ABasicTensors}}
\]
Its matrix product vector expansion is the displayed formula labelled
\texttt{decBSV}.  The block-injective definition and bound are labelled
\texttt{defnbi} and \texttt{propblockinj}.  The proportional Fundamental
Theorem is \texttt{thm1}; the equal-MPV corollary is \texttt{II\_cor2}.

In the proof appendix, the normal-block overlap and phase-gauge criterion is
\texttt{equalMPS}, the phase equality corollary is \texttt{eqV}, and the finite
power-sum lemma is \texttt{Lem:app\_simple}.  These labels are the source
objects that the formal BNT comparison should match.  Statements about raw
flattened cyclic sectors, common phase covers, or cumulative pair algebras are
formal intermediates; they are source-faithful only insofar as they are proved
to imply the labelled BNT comparison objects above.

\subsection{Formula-by-formula translation}

Reading the labelled source formulas literally gives the following translation
rules for the formal theorem surface.

\begin{enumerate}
  \item The displays \texttt{eq:II\_Aiplusk1} and \texttt{eq:II\_CF1} give a
  block diagonal tensor
  \[
    A^i=\bigoplus_k \mu_k A_k^i
  \]
  whose displayed blocks have dimensions summing to at most the original bond
  dimension.  The formal zero-tail parameter is therefore only the dimension of
  the omitted all-zero summand.  It is not an additional normal tensor, a BNT
  representative, or a separate scalar weight in the paper decomposition.

  \item The phase-gauge relation \texttt{eq:II:A=XAX} and the characterization
  \texttt{prop:char-BNT} quotient the canonical-form blocks by phase and gauge.
  Thus the paper's BNT is a list of representatives of equivalence classes.
  A formal common phase cover on raw cyclic sectors is not yet the paper's BNT
  matching statement until it has been pushed down to these representatives.

  \item The BNT expansion \texttt{eq:II\_ABasicTensors} and \texttt{decBSV}
  separate the representative index \(j\) from the copy index \(q\):
  \[
    |V^{(N)}(A)\rangle
      =
    \sum_j\left(\sum_q\mu_{j,q}^N\right)|V^{(N)}(A_j)\rangle .
  \]
  Repeated blocks therefore appear in the paper through the power sums of the
  weights attached to one representative.  They should not be exposed as a new
  family of theorem-level sector-matching statements unless the statement is
  exactly about recovering those multiplicities or weights.

  \item The source definitions \texttt{defnbi} and \texttt{propblockinj} give
  exact physical-length span after blocking:
  \[
    \operatorname{span}\{\tilde A^i\}_i
      =
    \bigoplus_j M_{D_j}(\mathbb C)
  \]
  for the blocked physical alphabet.  Cumulative algebra fullness, or
  membership of the identity at length zero, is not this assertion.

  \item The Fundamental Theorem \texttt{thm1} compares the BNT representatives
  of two tensors in canonical form.  The equal-case corollary \texttt{II\_cor2}
  then uses \texttt{Lem:app\_simple} on the already matched coefficient power
  sums.  Hence the formal equal-case route should first establish the
  representative matching and only then recover multiplicities and weights.
\end{enumerate}

These translation rules are also the naming rule for this part of the
formalization.  Results that merely package raw cyclic-sector data into one of
these hypotheses are supporting lemmas.  Theorems should be reserved for the
source-level milestones above, or for final statements that explicitly include
the remaining paper-facing hypotheses.

\subsection{Positive-length matrix product vectors}

The 2016 paper defines the matrix product vector family only for positive chain
lengths \(N=1,2,\ldots\)
\cite[Definition~\texttt{MPV}, source lines~130--151]{Cirac2016MPDO_arXiv}.
Thus the source comparison is
\[
  \operatorname{tr}(A^{i_1}\cdots A^{i_N})
  =
  \operatorname{tr}(B^{i_1}\cdots B^{i_N})
  \qquad (N \ge 1),
\]
or its proportional analogue. A formal relation that also includes the empty
word is a stronger auxiliary convention. It is useful for algebraic closure, but
the source theorem does not get a positive-length conclusion from the
length-zero trace.

\subsection{Canonical form and period removal}

The first source step removes invariant subspaces. Starting from a tensor
\(B\), the 2016 paper replaces it by block-diagonal matrices
\[
  A^i=\bigoplus_{k=1}^r \mu_k A_k^i,\qquad \sum_kD_k\le D
\]
which generate the same positive-length matrix product vectors
\cite[lines~214--225]{Cirac2016MPDO_arXiv}. Each block is then treated through
its completely positive map. Periodic peripheral eigenvalues are removed by
blocking a common multiple of their periods
\cite[lines~227--231]{Cirac2016MPDO_arXiv}. After that blocking, the blocks are
normal in the sense used in the paper, and the tensor is in canonical form
\cite[lines~233--255]{Cirac2016MPDO_arXiv}. The review gives the same
description: block diagonalization, period-removing blocking, primitive
transfer maps, and the resulting canonical form
\cite[lines~1793--1839]{Cirac2021Matrix}.

This is the first important separation of tasks. Period removal is a blocking
step used to make the peripheral spectrum primitive. It is not the later
finite-length injectivity or block-injectivity step.

\subsection{Basis of normal tensors}

The next source step chooses a basis of normal tensors. In the 2016 paper, such
a basis \(A_1,\ldots,A_g\) is a family of normal tensors whose matrix product
vectors span the matrix product vectors of \(A\) at each length and are
eventually linearly independent
\cite[lines~271--274]{Cirac2016MPDO_arXiv}. The subsequent characterization
identifies the basis tensors with representatives of the phase-gauge
equivalence classes of normal blocks: every normal block in the canonical form
is a phase times an invertible conjugate of one selected basis tensor, and no
two selected basis tensors are related in that way
\cite[lines~278--280]{Cirac2016MPDO_arXiv}. The review repeats the same
definition and characterization
\cite[lines~1846--1859]{Cirac2021Matrix}.

For a tensor in canonical form, the source then writes the matrix product vector
family through the basis of normal tensors as
\[
  |V^{(N)}(A)\rangle
    = \sum_{j=1}^g
        \left(\sum_{q=1}^{r_j}\mu_{j,q}^N\right)
        |V^{(N)}(A_j)\rangle.
\]
These are the data compared between two tensors in the equal or proportional
case
\cite[lines~283--302]{Cirac2016MPDO_arXiv};
\cite[lines~1864--1885]{Cirac2021Matrix}.

Consequently, the comparison data belong to the selected BNT representatives,
not to every raw block that appears before quotienting by phase-gauge
equivalence. Repeated copies and cyclic representatives contribute through the
power sums of their weights.

\subsection{The Fundamental Theorem and the equal case}

The Fundamental Theorem itself says that, for two tensors \(A\) and \(B\) in
canonical form, proportionality of all matrix product vector families matches
their bases of normal tensors up to a permutation, phases, and invertible gauge
matrices
\cite[lines~347--352]{Cirac2016MPDO_arXiv};
\cite[lines~1890--1900]{Cirac2021Matrix}. In the equal case, the source then
compares the coefficients in the BNT decompositions by the power-sum identity
\[
  \sum_{q=1}^{r_{a,j}}\mu_{j,q}^N
  =
  \sum_{q=1}^{r_{b,j}}(\nu_{j,q}e^{i\phi_j})^N
\]
for each matched basis tensor and all lengths \(N\). The elementary power-sum
lemma recovers the multiplicities and weights, and the blockwise gauges assemble
into the global conjugating matrix
\cite[lines~1155--1192]{Cirac2016MPDO_arXiv}.

\section{Fixed-length span after blocking}

There are two different blocking operations in the source account. The
period-removing blocking makes the transfer maps primitive. The Wielandt
blocking supplies finite-length injectivity after normality is available. These
steps should not be merged.

For one normal tensor, the quantum Wielandt theorem is cited as saying that
after at most \(D^4\) blockings the tensor becomes injective
\cite[line~332]{Cirac2016MPDO_arXiv}. The older canonical-form paper defines
injectivity by the map
\[
  \Gamma_L(X)
    =
    \sum_{i_1,\ldots,i_L}
      \operatorname{tr}(X A_{i_1}\cdots A_{i_L})
      |i_1,\ldots,i_L\rangle,
\]
and observes that injectivity of \(\Gamma_L\) is equivalent to the length-\(L\)
word products spanning the whole matrix algebra
\cite[lines~888--908]{PerezGarcia2007MPSReps}. Thus the relevant conclusion is
a homogeneous fixed-length span statement after a positive blocking length, not
mere membership in a cumulative span. The 2021 review records a sharper
MPS-specific Wielandt bound for normal tensors
\cite[lines~1942--1945]{Cirac2021Matrix}; the point here is the nature of the
conclusion, not which numerical bound is ultimately used.

For several basis blocks, the paper needs block-injective canonical form. In the
2016 source this is the assertion that one common physical tensor spans every
block matrix simultaneously:
\[
  \bigl(A_1^i,\ldots,A_g^i\bigr)_i
  \quad\text{spans}\quad
  \bigoplus_{j=1}^g M_{D_j}(\mathbb C)
\]
after the appropriate blocking
\cite[lines~317--332]{Cirac2016MPDO_arXiv}. The source then cites the
P\'erez-Garc\'ia--Verstraete--Wolf--Cirac bound: if \(L\) is a length at which
each normal tensor is injective, then after \(3(L+1)(D-1)\) blockings the
canonical-form tensor is block-injective; together with \(L\le D^4\), this gives
the \(3D^5\) bound
\cite[lines~340--345]{Cirac2016MPDO_arXiv}. The review states the same
block-injective conclusion and bound
\cite[lines~2114--2124]{Cirac2021Matrix}.

The older representation paper contains the direct-sum input behind this cited
step.  Its condition \(C1\) is exact-length injectivity of \(\Gamma_{L_0}\), or
equivalently length-\(L_0\) word products spanning the full matrix algebra
\cite[lines~888--908]{PerezGarcia2007MPSReps}.  It then fixes canonical blocks
which each satisfy \(C1\), sets \(L_0=\max_j L_0^{A^j}\), orders their bond
dimensions, and assumes the block states are pairwise different
\cite[lines~1321--1330]{PerezGarcia2007MPSReps}.  Under those hypotheses, a
matrix factorization lemma and a direct-sum lemma show that the sum of the
corresponding length-\(L\) MPS subspaces is direct once
\(L \ge 3(b-1)(L_0+1)\)
\cite[lines~1333--1408]{PerezGarcia2007MPSReps}.  The formal proof boundary is
therefore a homogeneous fixed-length direct-sum span or its trace-dual
separation form.  The note
\path{docs/paper-gaps/david2006_direct_sum_input.tex} records this source input
separately because it is a cited proof component, not a named theorem in the
2016 source.

This block-injective conclusion is stronger than separate injectivity of the
individual blocks. It includes the cross-block separation needed to prescribe
different matrices in different basis components at one common length. In dual
form, one fixed word length \(L\), after the prescribed blocking, gives the
following implication. If, for every word \(w\) with \(|w|=L\),
\[
  \sum_{j=1}^g \operatorname{tr}(\Delta_j A_j^w)=0
\]
then every test matrix \(\Delta_j\) is zero. Equivalently, the homogeneous
length-\(L\) word evaluations span the direct sum
\[
  \bigoplus_{j=1}^g M_{D_j}(\mathbb C).
\]
This fixed-length trace-dual statement is the mathematical content behind the
pair trace-separation predicates used in the non-periodic proof. It is not a
separate theorem in the papers; it is the dual form of block-injective
direct-sum span. In the later MPDO argument the same source explicitly writes
\(C_L(V)=\operatorname{span}(V^L)\) and invokes the block-injective proposition
to obtain \(C_L(V)=\mathcal A_1\) for some fixed \(L\)
\cite[lines~1984--1988]{Cirac2016MPDO_arXiv}. This is again an exact-length
span statement.

\section{The homogeneous padding point}

The paper-level injectivity and block-injectivity conclusions are homogeneous.
They concern words of a specified positive length after blocking. They are not
consequences of the empty word lying in a cumulative word span.

This distinction matters for the pair representation used in the comparison
proof. A cumulative span includes the empty word. Therefore the identity pair
belongs to a cumulative span at length zero for every pair of tensors. That fact
gives no positive-length padding. For example, over a one-letter alphabet with
\(A_0=B_0=0\) in bond dimension one, the length-one pair span is generated by
the zero pair, while the identity pair appears at length zero. Hence the
implication from cumulative identity membership to identity membership in all
later homogeneous spans is false.

Cumulative fullness alone is still too weak. In bond dimension one with one
letter and pair word values \((2^n,3^n)\), the cumulative span through length
one is all of \(\mathbb C\oplus\mathbb C\), since \((1,1)\) and \((2,3)\) are
linearly independent. At each positive homogeneous length \(n\), however, the
span is only the line through \((2^n,3^n)\), so it does not contain \((1,1)\).
Thus the source hypotheses must not be discarded down to cumulative pair-span
fullness.

The source target is fixed-length block-injective direct-sum span after the
prescribed blocking. A Burnside--Jacobson homogeneous-padding argument would be
an alternative proof of the same kind of exact-length trace separation; it must
retain the BNT, non-gauge-equivalence, and normalization data that make the
fixed-length conclusion true. It is not a standalone consequence of full
cumulative pair-span, and it is not a restatement of the empty-word observation.

\section{Comparison after common blocking}

The equal-case source proof compares BNT decompositions after the canonical form
has been chosen. The after-blocking theorem reaches this situation through a
longer route: discard the formal zero-tail summand just defined, choose one
common physical blocking length for both tensors, reindex nested blocked words
against flat words, separate period removal from later injectivity blocking, and
then compare the resulting nonzero sector families.

The common-blocking question is mathematical rather than clerical. If a block has
period-removal length \(m_k\) and the final common length is \(p=m_ke_k\), the
tensor obtained by blocking first by \(m_k\) and then by \(e_k\) has nested
physical labels, whereas the tensor obtained by blocking once by \(p\) has flat
word labels. Paper notation treats these identifications silently; the
theorem must prove the reindexing statements.

The BNT comparison data should therefore be produced for the final nonzero
sector families after all chosen blockings and reindexings. In source terms,
this means the proof must construct the matched basis tensors, the phase-gauge
equivalences, and the power-sum comparison for exactly those families. The
common primitive sectors should first be grouped into representative BNT
classes, with repeated cyclic representatives and transported weights accounted
for by the power-sum comparison. Common phase covers and proportional
decomposition data match the source theorem only for those representative or
grouped BNT families, not for the raw flattened cyclic-sector list. The raw list
can contain several cyclic representatives of the same original block with the
same transported weight, so it is not the basis of normal tensors used in the
paper comparison.

\section{Formal translation}

The source comparison above translates into the following formal targets.

\begin{itemize}
  \item The comparison relation should distinguish positive-length MPV equality
  from any auxiliary length-zero convention. If a theorem uses the empty word,
  the statement must say how that convention is removed before applying the
  paper theorem.

  \item The canonical-form reduction has two different blocking roles:
  period-removal blocking produces primitive normal blocks, while later
  Wielandt or block-injectivity blocking produces exact finite-length span.

  \item The BNT trace-separation theorem is the dual form of the fixed-length
  direct-sum span supplied by block-injective canonical form. The formal proof
  should therefore prove direct-sum span for the selected BNT representatives,
  or explicitly keep that fixed-length span as a hypothesis.

  \item Phase-cover and proportional-decomposition data must be obtained from
  BNT matching and the equal-case power-sum comparison. They should not be
  passed as unrelated assumptions on the raw flattened cyclic-sector list.

  \item The after-blocking comparison needs the paper-facing inputs for the
  representative BNT families: positive dimensions, injectivity, normalization,
  self-overlap limits, off-overlap decay for inequivalent representatives, and
  equality of finite-length matrix product vector spans. The auxiliary structure
  used to collect these inputs is only a formal device; the mathematical
  obligation is to derive the listed assertions for the representatives selected
  by the canonical-form and BNT construction.
\end{itemize}

\section{Current formal anchors}

The present formalization already separates the source theorem into the same
mathematical layers.

\begin{itemize}
  \item \path{TNLean/MPS/BNT/Construction.lean} defines
  \leanid{IsCanonicalFormBNT} as canonical form plus strict weight ordering and
  non-equivalence of distinct representatives. This is a restricted
  already-grouped surface: it suppresses the repeated-copy data
  \(r_j,\mu_{j,q}\) that appear in the BNT expansion and Appendix BNT
  characterization of \cite{Cirac2016MPDO_arXiv}.
  In the same file, \leanid{ProportionalDecompositionData} records explicit
  decomposition coefficients used by a conditional proportional comparison.
  The split comparison theorems
  \path{fundamentalTheorem_of_separated_CFBNT_data} and
  \path{fundamentalTheorem_of_separated_normalCFBNT_data} are therefore
  strict-representative special cases, not the general BNT theorem stated in
  the source paper.

  \item \path{TNLean/MPS/CanonicalForm/SectorComparison.lean} records the
  after-blocking theorem in conditional form. The structure
  \leanid{BlockedNormalFormHypotheses} supplies the remaining
  comparison data for the common primitive nonzero-sector families produced from
  the two original tensors. The conclusion
  \leanid{BlockedNormalFormSectorMatching} states the BNT sector
  matching after one positive common blocking length.

  \item \path{TNLean/MPS/CanonicalForm/SectorComparison/ProportionalComparison.lean}
  gives the current route from produced primitive sectors to the BNT-level
  comparison theorem. Its \leanid{CommonPrimitiveBNTCoverHypotheses} boundary is
  paper-facing only if those hypotheses are supplied for the selected
  representative BNT families.
\end{itemize}

\section{Named formal gaps}

The current non-periodic proof route still has three paper-facing obligations
which should not be hidden behind similarly shaped auxiliary hypotheses.

\begin{enumerate}
  \item Produce the common representative BNT cover data for the collapsed
  nonzero-sector families selected by canonical-form reduction and the BNT
  characterization.  This is the representative-family version of the paper's
  phase-gauge matching theorem, not a condition on every raw block produced by
  an intermediate cyclic decomposition.

  \item Prove the homogeneous trace-separation statement for distinct ordered
  BNT block pairs.  In paper notation this is the trace-dual form of the
  block-injective span of \(\oplus_jM_{D_j}(\mathbb C)\) at one common positive
  length.
  It is not the weaker statement that the cumulative algebra generated by a
  pair representation is full.

  \item Thread that fixed-length separation through the after-blocking
  comparison.  The blocking used to remove periods, the blocking used to obtain
  injectivity, and the reindexing from nested words to flat words must be
  accounted for before applying the BNT-level Fundamental Theorem.
\end{enumerate}

\section{Source files checked}

The comparison above is based on the LaTeX sources in the repository, not only
on published theorem names.

\begin{itemize}
  \item \path{Papers/1606.00608/MPDO-22-12-17-2.tex}: lines~130--151 define
  positive-length MPVs; lines~214--255 give canonical form and period removal;
  lines~271--302 define and use BNTs; lines~317--345 define injectivity,
  block-injective canonical form, and the \(3D^5\) bound; lines~347--360 state
  the Fundamental Theorem and equal-case corollary; lines~1155--1192 contain
  the power-sum comparison in the equal case; lines~1984--1988 use
  \(C_L(V)=\operatorname{span}(V^L)\) as an exact-length span.

  \item \path{Papers/quant-ph_0608197/MPSarchive.tex}: lines~742--763 state the
  translationally invariant canonical form; lines~888--908 define
  \(\Gamma_L\) and identify its injectivity with length-\(L\) word products
  spanning the matrix algebra.

  \item \path{Papers/1210.6613/UncleHMPS.tex}: lines~247--278 record the
  standard form after blocking, including the block-diagonal span condition and
  its injective/block-injective special cases.

  \item \path{Papers/2011.12127/TN-Review-main.tex}: lines~1793--1900 review
  canonical form, BNTs, and the Fundamental Theorem; lines~1942--1945 record the
  normal-tensor Wielandt bound; lines~2114--2124 state the block-injective
  canonical-form bound.
\end{itemize}

\section{Closing criterion}

A formal proof obligation in this part of the theorem is discharged only when the
following three assertions agree.

First, the source assertion has been identified at the level of the mathematical
objects above: canonical form, basis of normal tensors, block-injective
fixed-length span, phase-gauge BNT matching, or power-sum comparison. Second,
the formal theorem states the same assertion for the same blocked and reindexed
families, or explicitly records the extra hypothesis that remains. Third, the
final comparison theorem uses that assertion rather than an unrelated assumption
with a similar shape.

Conditional theorems are useful because they expose the missing mathematical
inputs precisely. They should not be described as closing the source theorem
until the corresponding assertion has been proved or the final theorem is
deliberately restated with that hypothesis.

\bibliographystyle{alpha}
\bibliography{docs/paper-gaps/references}

\end{document}
