\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{PEPS Gauge Uniqueness and Balanced Edge Scalars}
\author{}
\date{2026-04-30}

\begin{document}
\maketitle

\section{The assertion}

The uniqueness clause in the Fundamental Theorem for injective PEPS on a simple
graph appears in \cite[Theorem~2 and Section~3]{Molnar2018NormalPEPS}.
The argument first blocks the graph around a chosen edge to a three-site
injective MPS. The one-dimensional fundamental theorem then assigns a gauge to
that edge. Repeating this construction over all edges gives a family of edge
gauges, and a local two-tensor comparison gives scalar proportionalities at the
vertices.\footnote{In the repository source, the edge-blocking step is at
\path{Papers/1804.04964/paper_normal.tex:981--1009}; the assignment of edge
gauges is at \path{Papers/1804.04964/paper_normal.tex:1037--1065}; the local
scalar proportionality argument is at
\path{Papers/1804.04964/paper_normal.tex:1066--1092} and
\path{Papers/1804.04964/paper_normal.tex:1204--1207}.}

The theorem is then stated for PEPS on graphs with no double edges and no
self-loops. It says that two injective PEPS generate the same state if and only
if their tensors are related by local gauges. In short, the source says that the
local gauges are ``unique up to a multiplicative constant.''\footnote{The
quoted uniqueness phrase occurs in
\path{Papers/1804.04964/paper_normal.tex:1207--1210}. For traceability, this
note corresponds to \ghissue{1045}.}
Interpreted as uniqueness up to one common scalar multiplying all edge gauges,
this statement is too small for a general graph.

\section{The corrected quotient}

Let $G=(V,E)$ be a finite simple graph with a fixed orientation of each edge.
For an edge $e$, a local gauge family assigns an invertible matrix $X_e$ on the
bond space of $e$. At the tail of $e$ the endpoint action is $X_e$, and at the
head of $e$ it is $X_e^{-1}$.

A family of nonzero edge scalars $c=(c_e)_{e\in E}$ determines endpoint scalars
\[
  c_{v,e}=\begin{cases}
    c_e, & \text{if } v \text{ is the tail of } e,\\
    c_e^{-1}, & \text{if } v \text{ is the head of } e.
  \end{cases}
\]
The family is vertex-balanced if
\[
  \prod_{e\ni v} c_{v,e}=1
  \qquad\text{for every } v\in V.
\]
Two gauge families $X$ and $Y$ are equivalent modulo balanced edge scalars if
there is such a vertex-balanced family $c$ with
\[
  X_{v,e}=c_{v,e}Y_{v,e}
  \qquad\text{for every incident pair } (v,e),
\]
where $X_{v,e}$ and $Y_{v,e}$ denote the oriented endpoint actions of the two
families.

This is the graph-correct quotient. Rescaling the incident endpoint gauges at a
vertex $v$ by the factors $c_{v,e}$ multiplies the gauged local tensor at $v$ by
$\prod_{e\ni v} c_{v,e}$. If the scalar family is vertex-balanced, this product
is $1$ at every vertex, so all local gauged tensors are unchanged. Thus local
gauge uniqueness can at best be uniqueness modulo these balanced edge scalars.
The formal statement uses this quotient as the corrected uniqueness statement;
the quotient relation is an equivalence relation on gauge
families.\footnote{The formal definitions are
\leanid{TNLean.PEPS.edgeScalarAt}, \leanid{TNLean.PEPS.IsVertexBalanced}, and
\leanid{TNLean.PEPS.GaugeEquivModEdgeScalars}; the quotient relation is
\leanid{TNLean.PEPS.GaugeEquivModEdgeScalars.setoid} in
\path{TNLean/PEPS/FundamentalTheorem.lean}. The corresponding blueprint
entries are in \path{blueprint/src/chapter/ch13a_peps_ft.tex}.}

\section{The triangle witness}

The obstruction already appears on the connected triangle. Let the vertices be
$1<2<3$, orient every edge from the smaller endpoint to the larger endpoint, and
let every bond dimension be $1$. Then an edge gauge is just a nonzero complex
number. Take any nonzero injective bond-dimension-one PEPS tensor $A$, and
compare the identity gauge family with the scalar gauge family
\[
  c_{12}=t,\qquad c_{23}=t,\qquad c_{13}=t^{-1},
  \qquad t\in\mathbb C^\times,\quad t^2\ne 1.
\]
At the three vertices the oriented products are
\[
  c_{12}c_{13}=1,\qquad
  c_{12}^{-1}c_{23}=1,\qquad
  c_{13}^{-1}c_{23}^{-1}=1.
\]
Hence the family is vertex-balanced. Applying this gauge family to $A$ gives
back the same local tensor at every vertex, so it is a second gauge family
relating $A$ to itself.

There is no single global scalar $s$ with $c_{12}=c_{13}=c_{23}=s$. Indeed,
$c_{12}=c_{13}$ would give $t=t^{-1}$, contradicting $t^2\ne 1$. Therefore the
global-scalar uniqueness interpretation is false even for a connected simple
graph with one-dimensional bonds. The same example fits the balanced
edge-scalar quotient exactly.\footnote{This is the counterexample recorded in
\path{docs/counterexamples.md:127--135}. It originated in
\ghissue{762}; the formal
statement was corrected in \ghpr{790}.}

\section{Formal statement}

The discrepancy is a statement-level divergence, not only an unfinished proof.
The triangle example satisfies the graph-theoretic setting of the uniqueness
clause and produces two genuine gauge families which differ by a
balanced edge-scalar family but not by one global scalar. Consequently the
source phrase must either be read with this quotient understood, or corrected to
say uniqueness modulo balanced edge scalars.

The corrected quotient is now proved under the standing normal-PEPS convention
that every bond space is nonzero. From equality of the gauged local tensors at
each vertex, linear independence promotes the statement to equality of the
products of incident edge-gauge entries over every configuration pair. Since
each oriented endpoint gauge is invertible and, by the nonzero-bond convention,
indexed by a nonempty type, the many-leg scalar extraction yields, at each
vertex, nonzero proportionality constants with product $1$. The lower-endpoint
constant of each edge defines a single edge scalar; the inverse relation
transports it to the upper endpoint, and invertibility makes each constant
unique, so the per-vertex products identify with the oriented products of the
edge scalars and are therefore $1$.\footnote{For traceability, the declaration
is \texttt{TNLean.PEPS.gauge\_unique\_mod\_edge\_scalars}; the many-leg
extraction is \texttt{TNLean.PEPS.piProduct\_forms\_scalar} in
\path{TNLean/PEPS/TensorFactorScalar.lean}. This point is recorded in
\ghissue{842}; \ghpr{879} reduced the formal statement to the scalar-ratio and
global-reconciliation step, since completed.}

\paragraph{Scope restriction: nonzero bond spaces.}
The hypothesis $\dim_e>0$ on every edge $e$ is genuinely required, not a
convenience. If a bond space is empty, the local virtual configurations at its
two endpoints vanish, so the linear-independence hypothesis at those vertices is
vacuous. A positive-dimensional bond adjacent to such an empty-bond endpoint is
then left entirely unconstrained by the hypotheses: one can choose two gauge
families that agree everywhere except on that bond, where they differ by no
scalar, yet both satisfy all hypotheses vacuously. The conclusion fails for such
a pair. The nonzero-bond convention is exactly the normal-PEPS assumption used
in the existence direction (\texttt{hposA} in
\texttt{TNLean.PEPS.fundamentalTheorem\_PEPS}), so the uniqueness clause carries
the same convention.

Thus the corrected mathematical statement is uniqueness modulo balanced edge
scalars under nonzero bond spaces, and the triangle example refutes the smaller
global-scalar quotient.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
