\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}
\providecommand{\leanid}[1]{\path{#1}}

\title{Injective PEPS Fundamental Theorem: Section 3 Route}
\author{}
\date{2026-05-06}

\begin{document}
\maketitle

\section{The source argument}

The injective PEPS part of \cite{Molnar2018NormalPEPS} proves the
Fundamental Theorem for tensor networks on finite graphs without double edges
and without self-loops. The proof is local. It does not slice a two-dimensional
network into a long one-dimensional chain. Instead, it fixes an edge, blocks
the graph around that edge into a three-site injective MPS, applies the
one-dimensional injective theorem, and repeats this construction over all
edges.\footnote{The edge-blocking construction is
\path{Papers/1804.04964/paper_normal.tex:981--1009}. The equality of the two
PEPS before blocking is at
\path{Papers/1804.04964/paper_normal.tex:1010--1036}. The application of the
three-site injective MPS theorem is at
\path{Papers/1804.04964/paper_normal.tex:1037--1065}.}

For a chosen edge $e=(u,v)$, all vertices other than $u$ and $v$ are
grouped into the middle tensor. The two endpoint tensors and the middle tensor
form a three-site MPS. Injectivity of the PEPS tensors implies injectivity of
this blocked MPS. Since the two original PEPS represent the same state, the two
blocked MPS represent the same state. The one-dimensional injective theorem
therefore gives an algebra isomorphism between matrix insertions on the chosen
bond. Equivalently, an arbitrary matrix $X$ inserted on that bond in one
network is matched by a matrix on the corresponding bond in the other network.

After choosing the resulting edge gauges for every edge and absorbing them into
one tensor family, the paper obtains the following stronger comparison: for
every edge and every matrix $X$, inserting $X$ on that edge in the first
PEPS gives the same state as inserting $X$ on the same edge in the modified
second PEPS.\footnote{This is the displayed equation labelled
\path{eq:inj_equal_edge} in
\path{Papers/1804.04964/paper_normal.tex:1037--1065}.}

\section{Current formal representation}

The finite graph, edge, incident-edge, tensor, virtual configuration, state
coefficient, equality of states, and vertex injectivity are represented in
\path{TNLean/PEPS/Defs.lean}. Vertex injectivity is the linear independence of
the local physical vectors indexed by local virtual configurations, matching
the paper's statement that each tensor is injective as a map from the virtual
space to the physical space.

The local linear-algebraic realization of virtual operations at an injective
vertex is represented in \path{TNLean/PEPS/VirtualInsertion.lean}. The chosen
left inverse of the local tensor map and the resulting physical realization of
local virtual endomorphisms correspond to the same observation used in the
three-site MPS proof: a virtual operation can be represented by a physical
operation on a neighboring site. The basis computation
\leanid{TNLean.PEPS.localIncidentMatrixOp_single} and its tensor-map form
\leanid{TNLean.PEPS.localTensorMap_localIncidentMatrixOp_single} identify the
one-edge matrix action on a single local virtual boundary with the expected
sum over the new distinguished edge index.

The edge-centered blocking data are represented in
\path{TNLean/PEPS/Blocking.lean}. The existing declarations split local virtual
configurations at a distinguished incident edge and split the vertex product
into the left endpoint, the middle region, and the right endpoint. The
edge-blocked coefficient is proved equal to the original PEPS coefficient.

The next formal layer is the matrix-insertion comparison. The declarations
\leanid{TNLean.PEPS.EdgeInsertedBoundaryConfig},
\leanid{TNLean.PEPS.edgeBoundaryToInsertedBoundaryConfig},
\leanid{TNLean.PEPS.EdgeDiagonalInsertedBoundaryConfig},
\leanid{TNLean.PEPS.edgeBoundaryConfigEquivDiagonalInsertedBoundaryConfig},
\leanid{TNLean.PEPS.edgeMiddleConfigToOpenMiddleConfig},
\leanid{TNLean.PEPS.edgeOpenMiddleConfigToMiddleConfig},
\leanid{TNLean.PEPS.edgeMiddleConfigEquivOpenMiddleConfig},
\leanid{TNLean.PEPS.edgeOpenMiddleWeight},
\leanid{TNLean.PEPS.edgeBoundaryLeftIndexEquivInsertedBoundaryConfig},
\leanid{TNLean.PEPS.edgeBoundaryRightIndexEquivInsertedBoundaryConfig}, and
\leanid{TNLean.PEPS.edgeInsertedCoeff} represent the open edge contraction used
when an arbitrary matrix is inserted on the distinguished edge. These
definitions are the formal counterpart of the matrix insertion appearing before
and after the equation labelled \path{eq:inj_equal_edge}.
The two maps
\leanid{TNLean.PEPS.edgeMiddleConfigToOpenMiddleConfig} and
\leanid{TNLean.PEPS.edgeOpenMiddleConfigToMiddleConfig} have been assembled into
the finite equivalence
\leanid{TNLean.PEPS.edgeMiddleConfigEquivOpenMiddleConfig}. This isolates the
reindexing needed for the identity-insertion specialization. The current Lean
layer includes the corresponding middle-tensor reindexing theorem
\leanid{TNLean.PEPS.edgeMiddleWeight_eq_edgeOpenMiddleWeight}. The diagonal
summand has also been isolated in
\leanid{TNLean.PEPS.edgeInsertedCoeff_identity_diagonal_summand}, and
\leanid{TNLean.PEPS.edgeInsertedCoeff_identity_offDiagonal_summand} records the
pointwise off-diagonal vanishing. The full identity-insertion specialization is
\leanid{TNLean.PEPS.edgeInsertedCoeff_identity}; its proof uses
\leanid{TNLean.PEPS.edgeBoundaryConfigEquivDiagonalInsertedBoundaryConfig} to
reindex the finite inserted-boundary sum along the diagonal.

\section{Formalized Section 3 route}

The following statements record how the formal proof follows the Section~3
route.\footnote{This list and the ledger below are the status record referenced
by the Section~3 docstrings and blueprint comments.} They form a status account:
each item names the formal statement that carries the corresponding mathematical
step.

\begin{enumerate}
\item Vertex injectivity passes to the edge-blocked three-site
  object. This is the formal version of the paper's assertion that the
  regrouped MPS is injective. The middle physical index for this regrouped
  object is now explicit as
  \leanid{TNLean.PEPS.EdgeMiddlePhysicalConfig}, and the middle contractions
  have the middle-indexed forms \leanid{TNLean.PEPS.edgeMiddleWeightOn} and
  \leanid{TNLean.PEPS.edgeOpenMiddleWeightOn}. In the blueprint the
  injectivity statement is \leanid{TNLean.PEPS.EdgeBlockedThreeSiteInjective}.
  The singleton base case is now formalized by
  \leanid{TNLean.PEPS.singletonRegionTensorFamily} and
  \leanid{TNLean.PEPS.IsVertexInjective.singletonRegionTensorInjective}: for a
  one-vertex block $\{v\}$,
  $T_{A,\{v\}}(\eta,\sigma)=A_v(\eta,\sigma)$, hence vertex injectivity gives
  $\operatorname{inj}(T_{A,\{v\}})$. The passage to the abstract
  region-injectivity predicate is recorded by
  \leanid{TNLean.PEPS.SingletonRegionInjectivityComparison} and
  \leanid{TNLean.PEPS.SingletonRegionInjectivityFromVertexInjectivity.ofSingletonComparison}.
  With the abstract union-closure lemma, this gives every nonempty finite
  region by
  \leanid{TNLean.PEPS.RegionInjectivityUnionClosure.finset_injective_of_singletons}:
  $R=\bigcup_{v\in R}\{v\}$ and $\kappa(\{v\})$ for every $v\in R$ imply
  $\kappa(R)$.
  The corresponding nonempty-middle edge-middle tensor consequence is recorded
  in \path{thm:peps_edgeMiddleTensorInjective_of_singletons}:
  for $M_e=V\setminus\{u,v\}$, the assumptions
  $M_e\ne\varnothing$ and
  $M_e=\bigcup_{w\in M_e}\{w\}$ give $\kappa(M_e)$; the edge-middle
  comparison then gives injectivity of the middle tensor.
  The endpoint reduction then gives the one-edge three-site statement
  \path{thm:peps_edgeBlockedThreeSiteInjective_of_singletons}, and
  applying this to all edges satisfying $M_e\ne\varnothing$ gives
  \path{thm:peps_edgeBlockedThreeSiteInjective_all_of_singletons}.
  The graph-cardinality variant
  \leanid{TNLean.PEPS.EdgeMiddleRegionInjectivityComparison.%
edgeBlockedThreeSiteInjective_all_two_lt_card}
  covers the common case $2<|V|$: since $|M_e|=|V|-2>0$ for every edge,
  the nonempty-middle hypothesis is automatic.
  Vertex injectivity supplies the two endpoint assertions by
  \leanid{TNLean.PEPS.IsVertexInjective.edgeBlockedEndpointTensorMaps_injective}
  and, for all edges at once, by
  \leanid{TNLean.PEPS.IsVertexInjective.edgeBlockedEndpointTensorMaps_injective_all},
  and
  \leanid{TNLean.PEPS.IsVertexInjective.edgeBlockedThreeSiteInjective_of_middle}
  reduces the proof to the middle-block injectivity statement. Its
  all-edge form is
  \leanid{TNLean.PEPS.IsVertexInjective.edgeBlockedThreeSiteInjective_all_of_middle}.
  The
  comparison
  \leanid{TNLean.PEPS.EdgeMiddleRegionInjectivityComparison}
  records the conditional passage from finite-region injectivity of
  $V\setminus\{u,v\}$ to the edge-middle tensor family. Its theorem
  \leanid{TNLean.PEPS.EdgeMiddleRegionInjectivityComparison.edgeBlockedThreeSiteInjective}
  then combines this comparison with vertex injectivity to obtain the
  edge-blocked three-site assertion at one edge. The all-edge middle-tensor
  comparison is
  \leanid{TNLean.PEPS.EdgeMiddleRegionInjectivityComparison.edgeMiddleTensorInjective_all}.
  The all-edge conditional form
  is
  \leanid{TNLean.PEPS.EdgeMiddleRegionInjectivityComparison.edgeBlockedThreeSiteInjective_all}.
  The finite induction on exposed middle vertices is separated as
  \leanid{TNLean.PEPS.FiniteRegionKernelDescent}: for kernel conditions
  $K_c(S)$ attached to a coefficient family $c=(c_\rho)_\rho$, the implications
  $K_c(S)\Rightarrow K_c(S\setminus\{j\})$ imply
  $K_c(M_e)\Rightarrow K_c(\varnothing)$. The edge-middle instance of this
  coefficient argument is recorded by
  \leanid{TNLean.PEPS.EdgeMiddleKernelDescentData}: the zero relation
  $\sum_\rho c_\rho T_{A,M_e}^\rho(\tau)=0$ for every middle physical index
  $\tau$ gives $K_c(M_e)$, and $K_c(\varnothing)$ gives $c_\rho=0$ for every
  $\rho$. Consequently
  \leanid{TNLean.PEPS.EdgeMiddleKernelDescentData.edgeMiddleTensorInjective}
  proves $\operatorname{inj}(T_{A,M_e})$ from such a datum, and
  \leanid{TNLean.PEPS.IsVertexInjective.edgeBlockedThreeSiteInjective_of_kernelDescent}
  combines it with the endpoint injectivity assertions.
  The implication from vertex injectivity to middle-block injectivity is false
  without a positivity hypothesis on the virtual bond dimensions. If some
  complement bond makes
  \leanid{TNLean.PEPS.EdgeComplementConfig} empty, then the middle tensor
  family is identically zero
  (\leanid{TNLean.PEPS.edgeMiddleTensorFamily_eq_zero_of_isEmpty}), and the
  family is not linearly independent when the boundary-label index is nonempty
  (\leanid{TNLean.PEPS.not_edgeMiddleTensorInjective_of_isEmpty}). A four-cycle
  with a zero-dimensional middle-to-middle edge gives the corresponding
  mathematical example in which vertex injectivity can survive vacuously at
  vertices incident to that bond. The source avoids this because injective PEPS
  have nonzero-dimensional virtual bond spaces. The hypothesis
  $\forall f,\ 0<\dim(f)$ is therefore part of
  \leanid{TNLean.PEPS.IsVertexInjective.edgeBlockedThreeSiteInjective}; the two
  lemmas above record the formally checked non-injectivity obstruction.

  The mathematical step represented by the restored theorem is the contraction
  theorem used in the paper: contracting vertex-injective tensors
  over that finite middle region gives an injective blocked tensor. The source
  proves this in one step from the one-sided inverse, where the inverse of a
  contraction of injective tensors is the contraction of the inverses up to the
  connecting bond dimension (\path{Papers/1804.04964/paper_normal.tex}, lines
  205--250). The kernel descent recorded above is the finite-induction route
  towards that one-step fact, not a separate argument in the source. The full
  transfer is stated as
  \leanid{TNLean.PEPS.IsVertexInjective.edgeBlockedThreeSiteInjective} with
  the positive-bond hypothesis; its proof consists of the three facts below.

  The per-vertex one-sided-inverse step is expressed by
  \leanid{TNLean.PEPS.IsVertexInjective.localCoeff_eq_zero_of_contract_zero}:
  a coefficient family on the local virtual configurations at a vertex $v$ that
  contracts to the zero physical vector vanishes. The middle tensor is a genuine
  per-vertex product,
  $T_{A,M_e}(\rho,\tau)=\sum_{\zeta}\prod_{v\in M_e}A_v(\zeta,\tau_v)$, so a
  single vertex $j$ is removed by
  $\prod_{v\in M_e}=A_j\cdot\prod_{v\in M_e\setminus\{j\}}$. Three facts
  instantiate the kernel descent from vertex
  injectivity:
  \begin{enumerate}
  \item the star-bond split equivalence
    \leanid{TNLean.PEPS.edgeComplementConfigSplitAt}, decomposing the complement
    bond configuration into the star bonds of $j$ (its local virtual
    configuration) and the remaining bonds, with a first-component law matching
    $\textsf{edgeComplementValue}$; the shared bond between two middle vertices
    is read back from the star factor, which is why the peeled coefficient
    genuinely depends on $j$'s local virtual configuration;
  \item the partial-contraction kernel condition over a finite open region $S$
    together with its one-vertex deletion $K_c(S)\Rightarrow K_c(S\setminus
    \{j\})$, obtained by fixing the exposed bonds, ranging the open physical
    leg of $j$, and applying
    \leanid{TNLean.PEPS.IsVertexInjective.localCoeff_eq_zero_of_contract_zero};
  \item the terminal isolation $K_c(\varnothing)\Rightarrow c=0$ over the
    residual endpoint boundary labels, by surjectivity of the boundary
    labelling, which is where the positive bond dimensions are used.
  \end{enumerate}
  These assemble into
  \leanid{TNLean.PEPS.edgeMiddleKernelDescentData} and close the target by the
  existing reductions.
\item The two internal steps in Lemma \path{inj_isomorph} are formalized for
  the edge-blocked three-site MPS. The local endpoint part of the
  virtual-to-physical realization $X\mapsto O_1,O_2$ is formalized by
  \path{thm:peps_virtualInsertionPhysicalRealization}: vertex injectivity turns
  the one-leg virtual matrix action at either endpoint into a physical operator
  on that endpoint. The endpoint coefficient-level direction is formalized by
  \path{thm:peps_edgeInsertedCoeff_endpointPhysicalRealization}: after fixing
  ordinary edge-boundary data, the left and right physical operators supply the
  independent inserted-edge index at the corresponding endpoint. On the left
  endpoint the formal statement uses the transpose of the inserted matrix,
  because the ordinary boundary supplies the right bond index. The source
  recovery $O_1,O_2\mapsto W$ is proved as
  \leanid{TNLean.PEPS.physical_to_virtual_insertion}. The
  middle block is contracted out by
  \leanid{TNLean.PEPS.resonate_middle_inverted}, leaving the bond-contracted
  endpoint identity. Inverting the right endpoint
  (\leanid{TNLean.PEPS.resonate_invert_right_endpoint}) reads the common matrix
  $M$ off the action of $O_1$; inverting the left endpoint
  (\leanid{TNLean.PEPS.resonate_invert_left_endpoint}) reads the action of $O_2$
  off as a bond matrix, and
  \leanid{TNLean.PEPS.resonate_endpoint_coeff_reconcile} forces the two to
  agree, which is the source step $V=W$. The positivity hypothesis on the
  virtual bond dimensions supplies the reference residual configurations used to
  fix $M$.
  Its local endpoint recovery component is also formalized by
  \leanid{TNLean.PEPS.edgeEndpointLocalVirtualOpOfPhysicalOp_eq_of_projected_realization_eq}:
  once the compressed endpoint actions are known to realize the same bond
  matrix, the local virtual pullbacks recover that matrix on the shared bond.
  The image-preserving conditional form
  \leanid{TNLean.PEPS.edgePhysicalToVirtualInsertion_of_projected_realization_eq}
  records the endpoint conclusion under explicit projected-realization and
  image-preservation hypotheses.
\item Equality of PEPS states is upgraded to equality of edge-inserted
  coefficients after the three-site reduction. This is the step that produces
  the matrix-algebra isomorphism on the chosen bond. The inserted-coefficient
  correspondence $X\mapsto Y$ is now formalized by
  \leanid{TNLean.PEPS.exists_edgeInsertedCoeff_eq}: for every inserted matrix
  $X$ on the chosen bond of the first blocked PEPS there is a matrix $Y$ on the
  corresponding bond of the second blocked PEPS with equal edge-inserted
  coefficients at every physical configuration. Its proof rests on the
  state and endpoint-operator comparison
  \leanid{TNLean.PEPS.edgeRealizationSum_left_eq_sum_edgeBlockedCoeff} and
  \leanid{TNLean.PEPS.edgeRealizationSum_right_eq_sum_edgeBlockedCoeff}, which
  rewrite the endpoint realization sums of
  \leanid{TNLean.PEPS.edgeInsertedCoeff_eq_sum_left_physicalRealization} and
  \leanid{TNLean.PEPS.edgeInsertedCoeff_eq_sum_right_physicalRealization} as
  weighted sums of the edge-blocked coefficient (using the endpoint
  update-invariance of the open middle weight,
  \leanid{TNLean.PEPS.edgeOpenMiddleWeight_update_left} and
  \leanid{TNLean.PEPS.edgeOpenMiddleWeight_update_right}); equality of PEPS
  states then transfers them between the two families
  (\leanid{TNLean.PEPS.edgeRealizationSum_left_sameState},
  \leanid{TNLean.PEPS.edgeRealizationSum_right_sameState}), and
  \leanid{TNLean.PEPS.physical_to_virtual_insertion} supplies $Y$ on the second
  family. The correspondence $X\mapsto Y$ is now realized as an \emph{explicit}
  matrix-to-matrix map \leanid{TNLean.PEPS.edgeTransferMatrix}, built as a
  composition of algebra maps: insert $X$ on the right endpoint of the first
  family and realize it (\leanid{TNLean.PEPS.edgeRightInsertionOp}), transfer the
  resulting physical operator to the second family across equal states, and read
  off the matrix it realizes there. The load-bearing identity that the explicit
  map agrees with the matrix recovered by
  \leanid{TNLean.PEPS.physical_to_virtual_insertion} is
  \leanid{TNLean.PEPS.edgeRightInsertionOp_realizes_edgeTransferMatrix}. Because
  each step preserves composition, the explicit map is multiplicative
  (\leanid{TNLean.PEPS.edgeTransferMatrix_mul}); it is also additive and
  homogeneous (\leanid{TNLean.PEPS.edgeTransferMatrix_add},
  \leanid{TNLean.PEPS.edgeTransferMatrix_smul}) and satisfies the coefficient
  identity (\leanid{TNLean.PEPS.edgeTransferMatrix_edgeInsertedCoeff}). The
  multiplicativity that was \emph{not} visible from the coefficient identity is
  thereby resolved.

  The full positive-bond theorem
  \leanid{TNLean.PEPS.isEdgeBlockedInsertionAlgebraIsomorphism} now proves that
  this correspondence is a $\mathbb C$-algebra isomorphism, under the explicit
  hypotheses that every virtual bond of both tensor families has positive
  dimension. These hypotheses make explicit the nonzero virtual spaces used in
  Section~3, for example the reference residual configurations and the full
  matrix algebras on the virtual bonds. They are not consequences of
  injectivity. With a zero-dimensional bond, injectivity and state equality can
  hold vacuously while the physical-to-virtual recovery and gauge conclusion
  fail. The theorem-level obstruction is checked in
  \leanid{TNLean.PEPS.FundamentalTheoremCounterexample.fundamentalTheoremPEPSStatement_false};
  the corresponding intermediate recovery failure is checked in
  \leanid{TNLean.PEPS.PhysicalToVirtualCounterexample}, and the edge-middle
  obstruction is
  \leanid{TNLean.PEPS.not_edgeMiddleTensorInjective_of_isEmpty}. Thus the
  positive-bond assumptions are the formal expression of the source's nonzero
  virtual bond spaces, not an unproved consequence. The last step in
  the positive-bond theorem is the injectivity of the edge-inserted coefficient
  in the inserted matrix: equal coefficients at every physical configuration
  determine the matrix. This gives the unit law $\Phi(1)=1$ (the identity
  insertion has the same coefficient as $\Phi(1)$ by equal states), injectivity
  of the transfer map, and surjectivity by the symmetric $A\leftrightarrow B$
  construction. The injectivity is the inverse-direction content of the
  resonate inversion used inside
  \leanid{TNLean.PEPS.physical_to_virtual_insertion}.
\item The algebra isomorphism is converted into an invertible edge gauge by the
  same finite-dimensional matrix-algebra argument as the one-dimensional
  injective theorem. This is formalized by
  \leanid{TNLean.PEPS.edgeGaugeFromInsertionAlgebraIsomorphism}: an algebra
  isomorphism between two full matrix algebras first identifies their sizes and
  is then conjugation by an invertible matrix. The per-edge gauges used in the
  global proof are collected by
  \leanid{TNLean.PEPS.exists_edgeGaugeFamily}.
\item After all edge gauges have been absorbed into the second tensor family,
  producing the modified tensors $\widetilde B_v$, the analog of
  \path{eq:inj_equal_edge} is formalized by
  \leanid{TNLean.PEPS.post_absorption_edge_insertion_equality}. Its conclusion
  constructs
  \leanid{TNLean.PEPS.PostAbsorptionEdgeInsertionEquality}: arbitrary insertions
  of the same matrix on the same edge give equal PEPS coefficients for the
  first tensor family and the absorbed second tensor family.
\item The generalized two-injective-tensor comparison from
  \path{lem:inj_equal_tensors_2} is formalized by
  \leanid{TNLean.PEPS.two_injective_tensor_insertion_comparison}. It says that
  equality of all virtual-bond insertions for two pairs of injective tensors
  forces $A_1=\lambda B_1$ and $A_2=\lambda^{-1}B_2$. The formal
  statement is an abstract version with nonempty spectator external boundary
  spaces; the source lemma is recovered by taking these spaces to be one-point
  spaces. The rank-one cancellation after coefficientwise product equality is
  now formalized by
  \leanid{TNLean.PEPS.twoBlockReciprocalScalarProportional_of_pointwise_mul_eq}:
  if
  $A_1(\eta_1,\mu,\sigma_1)A_2(\eta_2,\nu,\sigma_2)
  =B_1(\eta_1,\mu,\sigma_1)B_2(\eta_2,\nu,\sigma_2)$ for all indices, then
  injectivity of $A_1$ and $A_2$ gives the reciprocal scalar
  proportionality. The one-shared-bond coefficient-separation case is now
  formalized by
  \leanid{TNLean.PEPS.two_injective_tensor_insertion_comparison_singletonBond}:
  inserting the matrix unit $E_{\mu,\nu}$ extracts the single product
  $A_1(\eta_1,\mu,\sigma_1)A_2(\eta_2,\nu,\sigma_2)$, and the pointwise
  product cancellation theorem then gives reciprocal scalar proportionality.
  The coefficientwise product separation above is on the source path only in the
  single-shared-bond case
  (\leanid{TNLean.PEPS.two_injective_tensor_insertion_comparison_singletonBond}),
  where a matrix-unit insertion extracts the product directly. In the many-bond
  case the source does not separate the product: it inverts $A_2$, reads off the
  gauges $Z,U,W$ on three leg pairs, and uses their identity-form compatibility
  to force each gauge to be a scalar
  (\leanid{TNLean.PEPS.threeLeg_residual_forms_scalar};
  \path{Papers/1804.04964/paper_normal.tex}, lines 1157--1204). The inversion
  and leg-regrouping are contained in the proved core comparison
  \leanid{TNLean.PEPS.two_injective_tensor_insertion_comparison_core}.
\item Both injective blocks of the one-vertex-versus-complement comparison are
  now formalized. The selected vertex block is
  \leanid{TNLean.PEPS.vertexTwoBlock} with injectivity
  \leanid{TNLean.PEPS.isTwoBlockInjective_vertexTwoBlock}. The complement block
  is \leanid{TNLean.PEPS.complementTwoBlock}, the contraction of $A$ over
  $V\setminus\{v\}$ opened along the star $\textsf{IncidentEdge}(G,v)$ as the
  shared boundary; its injectivity
  \leanid{TNLean.PEPS.isTwoBlockInjective_complementTwoBlock} follows from vertex
  injectivity and positive bond dimensions through the star-boundary kernel
  descent \leanid{TNLean.PEPS.vertexComplementTensorInjective_of_isVertexInjective}.
  This is the same contraction-of-injective-tensors fact used for the edge-middle
  block, with the single distinguished vertex $v$ replacing the two endpoints of
  a distinguished edge: a global virtual configuration is split at each
  complement vertex $j\ne v$
  (\leanid{TNLean.PEPS.vertexConfigSplitAt}), the kernel condition descends by
  the one-sided inverse at $j$, and the terminal empty-region step uses positive
  bond dimensions to fill the interior virtual indices.
\item The final two-tensor comparison is applied by blocking one vertex against
  its complement. The paper uses this to prove
  $A_i=\lambda_i\widetilde B_i$ at each vertex, and then incorporates the
  scalars into the local gauges. This specialization is represented by
  \leanid{TNLean.PEPS.one_vertex_complement_comparison}. The formal statement
  uses the abstract two-block notation with nonempty external boundary spaces;
  the source comparison is obtained from the one-vertex and complement blocks
  after \path{eq:inj_equal_edge}.
\item The boundary-insertion argument removes the separate
  bond-dimension-identification hypothesis in
  \leanid{TNLean.PEPS.fundamentalTheorem_PEPS_of_bondDim}. For each edge, the
  algebra equivalence
  \leanid{TNLean.PEPS.edgeTransferAlgEquiv} identifies the two full matrix
  algebras of virtual insertions. The dimension count
  \leanid{TNLean.PEPS.bondDim_eq_of_matrixAlgEquiv} forces equality of the
  corresponding bond dimensions, and
  \leanid{TNLean.PEPS.fundamentalTheorem_PEPS} then gives the source
  gauge-equivalence conclusion under the standing positive-bond and
  connectivity assumptions.
\end{enumerate}

\section{Source-to-formalization ledger}

The following ledger records the present status of the Section~3 proof route.
It is meant to prevent the proof from drifting away from the source argument.

\begin{itemize}
\item \path{eq:block_to_mps} edge blocking:
  \path{edgeBlockedCoeff_eq_stateCoeff} is locally proved; the transfer of
  injectivity is stated as the positive-bond theorem
  \leanid{TNLean.PEPS.IsVertexInjective.edgeBlockedThreeSiteInjective}. The
  middle physical index of this three-site object is formalized by
  \leanid{TNLean.PEPS.EdgeMiddlePhysicalConfig}. The named
  injectivity statement is \leanid{TNLean.PEPS.EdgeBlockedThreeSiteInjective},
  and the endpoint assertions are proved from vertex injectivity, both for a
  single chosen edge and uniformly over all edges. The comparison from
  edge-middle region injectivity to edge-middle tensor injectivity is also
  available uniformly over all edges. The all-edge
  conditional implication from middle-region injectivity is
  \leanid{TNLean.PEPS.EdgeMiddleRegionInjectivityComparison.edgeBlockedThreeSiteInjective_all}.
\item Identity insertion on the chosen bond:
  \path{thm:peps_edgeInsertedCoeff_identity} is formalized by finite diagonal
  reindexing.
\item Lemma \path{inj_isomorph}, $X\mapsto O_1,O_2$:
  \path{thm:peps_virtualInsertionPhysicalRealization} formalizes the endpoint
  realization, and
  \path{thm:peps_edgeInsertedCoeff_endpointPhysicalRealization} formalizes the
  endpoint coefficient expansion.
\item Lemma \path{inj_isomorph}, $O_1,O_2\mapsto W$:
  \leanid{TNLean.PEPS.physical_to_virtual_insertion} proves the source recovery
  step. The middle block is contracted out by
  \leanid{TNLean.PEPS.resonate_middle_inverted}; the two endpoints are inverted
  by \leanid{TNLean.PEPS.resonate_invert_right_endpoint} and
  \leanid{TNLean.PEPS.resonate_invert_left_endpoint}; the two extracted matrices
  are reconciled by
  \leanid{TNLean.PEPS.resonate_endpoint_coeff_reconcile}. The positivity
  hypothesis on the virtual bond dimensions supplies the reference residual
  configurations used to fix the recovered matrix.
  The endpoint projected-recovery component for a common bond matrix is also
  formalized by
  \leanid{TNLean.PEPS.edgeEndpointLocalVirtualOpOfPhysicalOp_eq_of_projected_realization_eq},
  and the conditional theorem
  \leanid{TNLean.PEPS.edgePhysicalToVirtualInsertion_of_projected_realization_eq}
  records the endpoint conclusion under explicit image-preservation hypotheses.
\item Matrix-insertion algebra isomorphism, positive-bond variant:
  \leanid{TNLean.PEPS.isEdgeBlockedInsertionAlgebraIsomorphism} states the
  edge-blocked matrix-algebra correspondence and its inserted-coefficient
  equality under positive bond dimensions for both tensor families. The
  inserted-coefficient correspondence $X\mapsto Y$ is formalized by
  \leanid{TNLean.PEPS.exists_edgeInsertedCoeff_eq}; its witness is the explicit
  transfer matrix and its coefficient identity is
  \leanid{TNLean.PEPS.edgeTransferMatrix_edgeInsertedCoeff}. The
  algebra-equivalence theorem
  \leanid{TNLean.PEPS.isEdgeBlockedInsertionAlgebraIsomorphism} is now proved in
  this positive-bond scope: well-definedness, injectivity, surjectivity, and the
  homomorphism laws including multiplicativity are part of the formalized
  statement. The positive-bond hypotheses express the source's nonzero virtual
  bond spaces; they are not consequences of injectivity.
\item Edge gauge from the algebra isomorphism:
  \leanid{TNLean.PEPS.edgeGaugeFromInsertionAlgebraIsomorphism} proves the
  Skolem--Noether step from an algebra equivalence of full matrix algebras to
  conjugation by an invertible edge matrix. The edgewise family used later is
  \leanid{TNLean.PEPS.exists_edgeGaugeFamily}.
\item Absorption of edge gauges into $B$:
  \leanid{TNLean.PEPS.absorbEdgeGauges} defines the absorbed tensor family, and
  \leanid{TNLean.PEPS.post_absorption_edge_insertion_equality} applies the
  edge-gauge family to it.
\item Equation \path{eq:inj_equal_edge}:
  \leanid{TNLean.PEPS.post_absorption_edge_insertion_equality} constructs
  \leanid{TNLean.PEPS.PostAbsorptionEdgeInsertionEquality}.
\item Lemma \path{inj_equal_tensors_2}, scalar step:
  the rank-one product cancellation is
  \leanid{TNLean.PEPS.twoBlockReciprocalScalarProportional_of_pointwise_mul_eq},
  and the many-bond scalar forcing is
  \leanid{TNLean.PEPS.threeLeg_residual_forms_scalar}.
\item Lemma \path{inj_equal_tensors_2}, comparison:
  \leanid{TNLean.PEPS.two_injective_tensor_insertion_comparison} states the
  abstract two-block insertion comparison with nonempty external boundary
  spaces and reciprocal scalar proportionality. The source lemma is the
  one-point external-boundary specialization.
\item One vertex versus its complement:
  \leanid{TNLean.PEPS.one_vertex_complement_comparison} records the
  application of the generalized two-block comparison to a selected vertex and
  its complement, under the abstract nonempty external-boundary formulation.
\item Final gauge equivalence and uniqueness:
  \leanid{TNLean.PEPS.fundamentalTheorem_PEPS_of_bondDim} proves the
  gauge-equivalence theorem after the bond dimensions are identified;
  \leanid{TNLean.PEPS.fundamentalTheorem_PEPS} removes that separate hypothesis
  by the matrix-algebra dimension count, and
  \leanid{TNLean.PEPS.gauge_unique_mod_edge_scalars} proves uniqueness modulo
  edge scalars.
\end{itemize}

\section{Blueprint diagrams}

The blueprint now uses semantic tensor-network commands for the Section 3 moves
that should remain visually attached to the proof.

\begin{itemize}
\item $\backslash$\texttt{TNPEPSEdgeBlockingReduction} depicts the chosen edge
  and the passage to the three-site chain. Its endpoint labels follow
  \path{eq:block_to_mps}: the two circled endpoints are $A'_1$ and $A'_2$,
  while the boxed complement is $A'_3$.
\item $\backslash$\texttt{TNPEPSEdgeInsertedCoeff} depicts the single
  edge-blocked contraction with one virtual matrix $X$ inserted on the chosen
  bond. This is the diagram attached to the definition of the inserted-edge
  coefficient, before comparing two tensor networks.
\item $\backslash$\texttt{TNPEPSThreeSiteInsertionComparison} depicts the
  comparison of a virtual insertion in the two three-site chains, as in the
  statement of \path{lem:inj_isomorph}.
\item $\backslash$\texttt{TNPEPSInsertionPhysicalRealization} depicts the
  realization of a virtual insertion as a physical operation on either
  neighboring site.
\item $\backslash$\texttt{TNPEPSPhysicalToVirtualInsertion} depicts the
  converse step in which two neighboring physical operations determine a
  virtual operation on the shared bond.
\item $\backslash$\texttt{TNPEPSEdgeInsertionEquality} depicts the full
  PEPS comparison after the edge gauges have been absorbed into the second
  tensor family, matching the five-vertex graph and the distinguished
  $(2,5)$-edge insertion in \path{eq:inj_equal_edge}.
\item $\backslash$\texttt{TNPEPSEdgeGaugeAbsorption} depicts the local
  absorption of incident oriented edge gauges into a $B$-tensor, producing
  $\widetilde B_v$.
\item $\backslash$\texttt{TNPEPSTwoInjectiveTensorInsertionComparison}
  depicts the generalized two-injective-tensor comparison with an arbitrary
  matrix inserted on a shared virtual bond.
\item $\backslash$\texttt{TNPEPSTwoInjectiveGaugeScalarReduction} depicts the
  residual virtual operators $Z,U,W$ that appear after applying an inverse
  for one injective block in the proof of
  \path{lem:inj_equal_tensors_2}.
\item $\backslash$\texttt{TNPEPSOneVertexComplementComparison} depicts the
  final one-vertex-versus-complement comparison used to prove
  $A_v\propto\widetilde B_v$. It follows the source paper's displayed
  comparison by drawing the one-vertex enlargement and the smaller injective
  region as tensor-network tensors, rather than as an auxiliary formal block.
\end{itemize}

These diagrams deliberately keep the local dot convention used in the blueprint,
but they preserve the distinctions that matter in the paper: tensor sites,
blocked regions, virtual matrix insertions, and physical operations are not
drawn as the same object.

\section{Conclusion}

The injective PEPS Fundamental Theorem is now formalized, faithful to
\cite[Section~3]{Molnar2018NormalPEPS}. Under the source's standing
positive-bond-dimension and connectivity assumptions, the existence clause
\leanid{TNLean.PEPS.fundamentalTheorem\_PEPS} and the uniqueness clause
\leanid{TNLean.PEPS.gauge\_unique\_mod\_edge\_scalars} are proved and
axiom-clean. The three-site injectivity transfer, the matrix-insertion
comparison as a \(\mathbb C\)-algebra isomorphism, the edge gauge construction,
and the one-vertex-versus-complement comparison are all complete. The
positive-bond-dimension and connectivity hypotheses are not removable: dropping
positivity is refuted by
\leanid{TNLean.PEPS.FundamentalTheoremCounterexample.fundamentalTheoremPEPSStatement_false},
and dropping connectivity is refuted by
\leanid{TNLean.PEPS.GaugeConsistencyConnectivityCounterexample}. Thus they are
genuine standing assumptions of the source rather than steps to be discharged.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
