\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}
\providecommand{\ghissue}[1]{GitHub issue~\##1}
\providecommand{\leanid}[1]{\path{#1}}

\title{Normal PEPS Fundamental Theorem: Section 3 Route}
\author{}
\date{2026-05-06}

\begin{document}
\maketitle

\section{The source argument}

The normal-PEPS part of \cite{Molnar2018NormalPEPS} begins after the injective
PEPS theorem. Its purpose is not to replace the injective proof, but to make
the injective proof applicable after a controlled blocking. A normal PEPS is
one for which sufficiently large regions can be blocked into injective tensors.
Once such injective regions are available, the proof again applies the
three-site injective-chain argument around a chosen edge.

The source proof has three layers. First, Lemma \path{injective_union} proves
that the union of two injective regions is injective. The proof blocks the
network into the four regions $A\setminus B$, $A\cap B$,
$B\setminus A$, and $(A\cup B)^c$. If a boundary tensor kills the
contraction over $A\cup B$, injectivity of $A$ first removes the
$A\setminus B$ and $A\cap B$ part, and injectivity of $B$ then removes
the remaining $A\cap B$ and $B\setminus A$ part.
\footnote{The source is
    \path{Papers/1804.04964/paper_normal.tex:1324--1400}.}

Second, for the translationally invariant square-lattice theorem, the paper
assumes that every $2\times 3$ and $3\times 2$ region is injective. By
repeatedly applying the union lemma, the regions denoted $R$, $S$, and $T$
in the paper are injective. The finite-lattice block denoted $A_3$ in the proof
of Theorem~\path{3} has the same complementary shape as $T$ but lives in the
full finite lattice around the chosen edge; the present coordinate definition
records only the displayed local-window complement. The regions $R$ and $S$
differ by one site, and their complements are injective when the system is
large enough. These are the regions used in the last comparison step.
\footnote{The regions $R$, $S$, and $T$ are drawn at
    \path{Papers/1804.04964/paper_normal.tex:1407--1443}.}

Third, the proof of Theorem \path{3} blocks a neighborhood of each edge into
three injective pieces. The red region is $A_1$, the blue region is $A_2$,
and the complement is $A_3$. This produces the same three-site injective MPS
configuration used in the injective PEPS proof. Lemma \path{inj_isomorph}
therefore gives a gauge on every edge. Translation invariance reduces these
edge gauges to one horizontal matrix $X$ and one vertical matrix $Y$. After
absorbing them into the second tensor, the final comparison of the $R$ and
$S$ regions gives $A\propto \widetilde B$.
\footnote{The proof of Theorem \path{3} is
    \path{Papers/1804.04964/paper_normal.tex:1449--1572}.}

The general theorem \path{normal} removes the square-lattice specialization.
It assumes precisely the blockings needed above: every edge can be blocked into
a three-site injective chain, and every site admits two injective regions with
injective complements that differ exactly in that site.
\footnote{The general theorem is
    \path{Papers/1804.04964/paper_normal.tex:1574--1585}.}

\section{Current formal representation}

The injective PEPS objects used by the normal theorem are represented in
\path{TNLean/PEPS/Defs.lean}, \path{TNLean/PEPS/VirtualInsertion.lean}, and
\path{TNLean/PEPS/Blocking.lean}. Chapter 13a of the blueprint has a
separate normal-PEPS section after the injective theorem. This separation is
important: the normal theorem needs extra blocking hypotheses before the
injective proof route can be invoked.

The present blueprint nodes are the following.

\begin{itemize}
    \item \path{lem:peps_injectiveRegion_union} states the union-of-injective
          regions lemma. It is tracked by \ghissue{1374}. It records the
          four-region decomposition and the two successive inverse applications
          used in the source proof.
    \item \path{lem:peps_regionInjectivityUnionClosure_finite} records the
          finite nonempty iteration of the binary union-closure assertion.
    \item \path{def:peps_normalSquareBlockingRegions} records the square-lattice
          regions $R$, $S$, and $T$. It is part of \ghissue{1375}. The
          coordinate rectangular hypotheses, union closure, and a supplied
          rectangular cover of the displayed $T$-region imply this
          abstract structure.
    \item \path{thm:peps_normalEdgeBlockingToInjectiveChain} states that the
          normal square-lattice hypotheses allow a blocking around every edge into
          three injective regions. It is part of \ghissue{1375}.
    \item \path{thm:peps_tiNormalGaugeAbsorption} records the translationally
          invariant gauge absorption with horizontal gauge $X$ and vertical gauge
          $Y$. It is part of \ghissue{1375}.
    \item \path{thm:fundamentalTheorem_normalSquarePEPS} records the square-lattice
          theorem corresponding to Theorem \path{3}. It is part of \ghissue{1375}.
    \item \path{thm:fundamentalTheorem_normalPEPS} records the general normal PEPS
          theorem following Theorem \path{3}. It is part of \ghissue{1375}.
\end{itemize}

The finite rectangular square lattice is represented by pairs of horizontal
and vertical coordinates. Its graph has horizontal and vertical
nearest-neighbor edges, named coordinate right/up edges, and orientation
predicates distinguishing the two possible directions. Every edge has exactly
one of these orientations. With the ordered endpoints convention for edges,
horizontal edges are coordinate right edges from their left endpoints, and
vertical edges are coordinate upward edges from their lower endpoints.

The four-region decomposition from Lemma \path{injective_union} is formalized
both by named regions and by an indexed family in the source-paper order:
$A\setminus B$, $A\cap B$, $B\setminus A$, and $(A\cup B)^c$. The indexed
family reconstructs $A$, $B$, $A\cup B$, $A^c$, $B^c$, and the full vertex set.
The same reconstructions are available as finite unions over the selected
indexed subfamilies $\{0,1\}$, $\{1,2\}$, $\{0,1,2\}$, $\{2,3\}$, and
$\{0,3\}$. The singleton subfamily $\{3\}$ gives the outside block and is the
complement of the selected inside subfamily $\{0,1,2\}$. Its selected
complementary pairs match the two places where the source proof applies
injectivity of $A$ and then injectivity of $B$. The
selected subfamilies $\{0,1\}$ and $\{2,3\}$ are disjoint, as are
$\{1,2\}$ and $\{0,3\}$; these are the disjointness forms needed for the two
inverse applications. The abstract union-closure hypothesis also has the
corresponding indexed inside-union form: injectivity of the selected subfamilies
$\{0,1\}$ and $\{1,2\}$ implies injectivity of the inside subfamily
$\{0,1,2\}$. Every vertex belongs to exactly one indexed piece. These facts
prepare the tensor-level boundary-contraction proof, but do not prove that
theorem.

The rectangular vocabulary separates coordinate products from source-paper
contiguous blocks. Coordinate rectangles are products of finite coordinate
sets, whereas contiguous coordinate rectangles are products of coordinate
intervals. The $2\times3$ and $3\times2$ coordinate-product predicates assert
only the two cardinalities; the contiguous-rectangle predicates express the
source-paper blocks. The square-lattice rectangular injectivity hypotheses use
these two contiguous predicates and map to the abstract rectangular-injectivity
bundle. Bounded contiguous rectangles of sizes $2\times3$ and $3\times2$
satisfy the corresponding source-rectangle predicates.

The displayed regions $R$ and $S$ are explicit unions of contiguous
rectangles. Region $R$ is the union of one contiguous $2\times3$ rectangle and
one contiguous $3\times2$ rectangle. Region $S$ is the union of two contiguous
$2\times3$ rectangles. Their coordinate definitions agree with the source
claim that they differ in one site: $R\subseteq S$, and $S\setminus R$ is the
lower-right site of the displayed $3\times3$ square. Assuming the
union-closure assertion supplied by the source lemma on unions of injective
regions, these decompositions imply injectivity of $R$ and $S$. The formal
development also contains the finite nonempty iteration of this union-closure
assertion, matching the repeated unions used after Lemma
\path{injective_union}.

The displayed local model for $T$ is the complement, inside the source
$5\times6$ window, of the two edge blocks in the source picture. These removed
blocks are one contiguous $2\times3$ rectangle and one contiguous $3\times2$
rectangle, hence they are injective under the coordinate rectangular
hypotheses. They are disjoint, and together with the local-window $T$ region
they reconstruct the $5\times6$ window. This local-window region is distinct
from the finite-lattice complementary block $A_3$ used later in the proof of
Theorem~\path{3}. The finite-lattice block is the complement of the two edge
blocks in the full square lattice; its elementary disjointness and covering
identities are formalized separately.

A target-parametric rectangular-cover criterion supplies a sufficient
injectivity route: if a square-lattice region has a nonempty cover by
contiguous $2\times3$ and $3\times2$ rectangles, then rectangular injectivity
and finite union closure imply injectivity of that region. The formal special
cases are the local-window $T$ region and the finite-lattice complementary
block. This criterion is only a sufficient local model, not the source
argument itself.

The origin-based local-window model for $T$ has no such rectangular cover
whenever the ambient lattice contains the corresponding $5\times6$ window.
The obstruction is a point-forcing argument. For the local-window target
$T_{\mathrm{loc}}$ at the origin, let $R^{a\times b}_{x,y}$ denote the
contiguous $a\times b$ rectangle with lower-left corner $(x,y)$. Then
\begin{align}
    p &:= (0,0) \in T_{\mathrm{loc}},                                    \\
    p \in R^{2\times3}_{x,y}
        &\Longrightarrow q_{23}:=(0,2)\in R^{2\times3}_{x,y},
        \qquad q_{23}\notin T_{\mathrm{loc}},                             \\
    p \in R^{3\times2}_{x,y}
        &\Longrightarrow q_{32}:=(2,1)\in R^{3\times2}_{x,y},
        \qquad q_{32}\notin T_{\mathrm{loc}}.
\end{align}
Thus any exact cover of $T_{\mathrm{loc}}$ by $2\times3$ and $3\times2$
rectangles would cover $p$ by one of its pieces and would therefore also cover
a point outside $T_{\mathrm{loc}}$. The same calculation applies to the
rotated local-window model for vertical edges in a $6\times5$ window, to the
origin-based horizontal edge-complement model in the $5\times7$ frame, and to
the origin-based vertical edge-complement model in the $7\times5$ frame. The
forced points are
\[
\begin{array}{c|c|c|c}
    \text{target} & p & q_{23} & q_{32} \\
    \hline
    T_{\mathrm{loc}} & (0,0) & (0,2) & (2,1) \\
    T_{\mathrm{loc}}^{\mathrm{rot}} & (0,0) & (1,2) & (2,1) \\
    A_{3,\mathrm{hor}} & (0,0) & (0,2) & (2,1) \\
    A_{3,\mathrm{ver}} & (0,0) & (1,2) & (2,1)
\end{array}
\]
The formal point-forcing lemma is
\leanid{TNLean.PEPS.not_squareLatticeRectangleCover_of_forced_points}. Its
four present applications are the corresponding no-cover theorems for the
local $T$ models and the edge-complement models. These no-cover results do not
contradict the paper. They show that the source's injectivity argument for $T$
and $A_3$ must be aligned with the finite PEPS geometry, rather than read
directly from the present exact-cover criterion.

The normalized horizontal edge frame has a collar identity: in the
$5\times7$ frame, $A_3$ is the local-window $T$ region together with two
top-collar contiguous $3\times2$ rectangles. Thus local $T$-injectivity implies
injectivity of $A_3$ by union closure and rectangular injectivity. The
vertical counterpart is a $7\times5$ frame in which the edge complement is the
rotated local $T$ region together with two right-collar contiguous
$2\times3$ rectangles. A rectangular cover of the rotated local $T$ region
therefore implies injectivity of the vertical complementary block.

The injectivity of $A_3$ is now established directly, with no cover hypothesis
and no detour through the local-window $T$ region. The local $T$ region is a
fixed small window whose own corner is forced, so it has no rectangular cover by
contained source rectangles; but $A_3$ is the complement of the removed L-shape
in the \emph{full} finite lattice, and when the removed shape sits in the
interior the corner has room. The realignment is a finite-geometry cover of
$A_3$ by four full-length bands around the removed shape — below, above, left,
and right — together with two filler rectangles inside the shape's bounding box.
Each band is a contiguous rectangle whose two side lengths are at least the
source $2\times3$ or $3\times2$ shapes, hence injective as a finite union of
those shapes; each filler is one source rectangle. The union of these six
rectangles equals the complement of the removed shape, so union closure and
rectangular injectivity give injectivity of $A_3$ unconditionally for an
interior edge.\footnote{Formal declarations:
\leanid{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.wideRectangle_injective},
\leanid{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.shortRectangle_injective},
\leanid{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.interiorEdgeComplement_injective},
and
\leanid{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.interiorVerticalEdgeComplement_injective}.}
The interior placement asks for one row and column of margin below and to the
left of the removed shape and enough total room above and to the right, so the
shape is surrounded on every side.

The normalized horizontal and vertical edge blockings give one-edge data with
red, blue, and complementary regions. The distinguished edges have the
expected orientations; their endpoints lie in the red and blue regions; the
three regions are pairwise disjoint and cover the finite lattice. This one-edge
datum is the local form of the later every-edge blocking hypotheses. A family
of such one-edge data determines those hypotheses, and each datum supplies
the three injective regions used in the associated edge-blocked chain.

The translated horizontal and vertical pictures carry the same data at an
arbitrary coordinate origin. When the translated frame fits in the finite
lattice, the translated edge is identified with the corresponding coordinate
right or upward edge. A rectangular cover of the translated complementary
region supplies the one-edge blocking datum and hence the three injective
regions. The normalized pictures are the origin specializations of these
translated pictures, while the normalized and translated data remain in
separate Lean modules.

The conditional every-edge construction has two equivalent forms. A translated
edge-window assignment gives, for every edge, a horizontal or vertical window
and a rectangular cover of its complement. Such an assignment implies the
normal edge-blocking hypotheses. An oriented margin-and-cover datum records the
same input directly on each square-lattice edge: the edge orientation, the
coordinate inequalities placing the translated frame, and the complementary
cover. A family of these data also determines the normal edge-blocking
hypotheses, and the resulting hypotheses recover the supplied one-edge datum
at each edge.

With the interior $A_3$ injectivity in hand, the every-edge construction is now
cover-free at interior edges. For an edge with interior margins — horizontal
with enough room to place the surrounded horizontal removed shape, or vertical
with enough room for the rotated one — the translated edge-blocking datum is
built with no rectangular cover: its complementary-block injectivity is the
interior $A_3$ injectivity above. A family of these cover-free interior data over
all edges assembles the normal edge-blocking hypotheses, with the injective
chain, endpoint membership, pairwise disjointness, and coverage available at
every edge.\footnote{Formal declarations:
\leanid{TNLean.PEPS.horizontalSquareLatticeEdge_blockingDatum_interior},
\leanid{TNLean.PEPS.verticalSquareLatticeEdge_blockingDatum_interior},
\leanid{TNLean.PEPS.NormalSquareInteriorEdgeDatum}, and
\leanid{TNLean.PEPS.normalSquareEdgeBlockingHypotheses_of_interiorData}.}

For the present open rectangular coordinate graph, the margin predicates
describe only interior translated frames. A coordinate right edge must have
enough left and right margin; a coordinate upward edge must have enough bottom
and top margin. Hence the cover-free interior construction supplies data only on
the interior edges of the open $7\times7$ graph; the earlier translated-window
and oriented margin-and-cover criteria have the same interior restriction.
Explicit boundary obstructions are recorded on all four sides. The remaining
every-edge construction must therefore include the boundary geometry: the
interior $A_3$ injectivity removes the local and rotated $T$-cover obligation for
interior edges, leaving the boundary edges of the open rectangular model.

Verdict: this note records an open gap with a scope restriction. The earlier
vocabulary gap for contiguous rectangular blocks is closed. The earlier
coordinate rectangular-injectivity structure is also present, and the
indexed finite-region geometry of Lemma \path{injective_union} is formalized
down to the selected complement and closure forms used by the two inverse
applications. The rectangular decompositions of $R$ and $S$ are formalized, and
the conditional derivations of injectivity for $R$ and $S$ from union closure
are also formalized. The two edge blocks removed from the displayed $T$-window
are injective under the coordinate rectangular hypotheses. The finite-lattice
complementary block around the chosen edge has a set-theoretic definition.
The exact-cover criterion by $2\times3$ and $3\times2$ rectangles is only a
restricted sufficient condition, and the origin-based instances above show
that it is not the source's injectivity argument. Injectivity of the
edge-complementary block $A_3$ is now derived unconditionally for an interior
edge, by the finite-geometry band-and-filler cover of the full-lattice
complement; the local-window $T$ region is no longer on the critical path,
since the source object is $A_3$ and the local window is a fixed small frame
whose corner is forced. The every-edge blocking construction is cover-free at
interior edges and assembles the normal edge-blocking hypotheses there. The
tensor-level union theorem is proved for positive bond dimensions, including the
overlapping case (the overlap re-insertion). The remaining open gap is the
boundary geometry of the open rectangular model: edges near the lattice
boundary lack interior margins, so the cover-free construction does not yet
reach them; on a translation-invariant lattice every edge is interior and the
construction is complete. Injectivity of $R$ and $S$ is already derived from
union closure and rectangular injectivity.

The normal route itself is tracked by \ghissue{1373}. The normal-section
diagram work is tracked by \ghissue{1376}. The general PEPS status tracker is
\ghissue{1252}.

\section{Remaining mathematical obligations}

The normal PEPS theorem cannot be obtained by citing the injective theorem
alone. The lists below are the authoritative record for the square-lattice
normal blocking route. Lean docstrings and blueprint entries should point here
rather than repeat this status in full.

The following part of the formalization is already present.
\begin{itemize}
    \item Region injectivity is represented by
          \leanid{TNLean.PEPS.RegionInjectivityData}; the same infrastructure
          contains the region-complement definitions
          \leanid{TNLean.PEPS.regionComplement} and
          \leanid{TNLean.PEPS.regionOutsideUnion}, and the finite-region
          union-closure hypothesis
          \leanid{TNLean.PEPS.RegionInjectivityUnionClosure}.
    \item The four-region decomposition for Lemma \path{injective_union} is
          formalized both by named regions and by the indexed family
          \(A\setminus B\), \(A\cap B\), \(B\setminus A\), and
          \((A\cup B)^c\), including the reconstruction, disjointness, and
          selected-complement identities used by the two inverse applications.
    \item The three contraction maps for Lemma \path{injective_union} are
          separated as
          \leanid{TNLean.PEPS.InjectiveUnionContractionMaps}.
    \item The abstract contraction-chain criterion for Lemma
          \path{injective_union} is formalized by
          \leanid{TNLean.PEPS.InjectiveUnionContractionChain}: if the three
          implications
          \[
              \mathcal C_{\{0,1,2\}}(X)=0
              \Longrightarrow \mathcal C_{\{2\}}(X)=0
              \Longrightarrow \mathcal C_{\{1,2\}}(X)=0
              \Longrightarrow X=0
          \]
          are supplied, then the union contraction has trivial kernel.
    \item The square-lattice coordinate model, contiguous \(2\times3\) and
          \(3\times2\) rectangles, and the coordinate rectangular-injectivity
          hypotheses are present.
    \item The source regions \(R\) and \(S\) are explicit finite coordinate
          regions, their rectangular decompositions are formalized, and their
          injectivity follows conditionally from rectangular injectivity and
          union closure.
    \item The displayed local \(T\)-region, the finite-lattice
          edge-complement regions, their sufficient rectangular-cover criteria,
          the no-cover diagnostics for the present origin-based local models,
          and the normalized collar decompositions are formalized.
    \item The one-edge, translated-window, and margin-cover hypotheses
          for normal edge blocking are present, together with their endpoint,
          disjointness, coverage, and injective-chain consequences.
    \item The general normal PEPS blocking hypotheses
          \leanid{TNLean.PEPS.NormalPEPSBlockingHypotheses} provide their
          edge and one-site consequences directly, so downstream square-lattice
          arguments can use these hypotheses rather than restating the
          edge-blocking and one-site hypotheses separately.
\end{itemize}

The remaining obligations are the following.

\begin{enumerate}
    \item The tensor-level union-of-injective-regions lemma is now proved, with
          the explicit positive-bond hypothesis the formalization carries. The
          proof follows the source by applying the left inverses for $A$ and $B$
          in sequence. The disjoint case is formalized over a bare three-block
          geometry (no red-block injectivity and no distinguished edge): for
          disjoint injective regions $S$ and $T$ with positive bond dimensions,
          $S\cup T$ is injective, and its nonempty finite iteration over
          pairwise-disjoint families is also available.\footnote{Formal
          declarations:
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_union_disjoint} and
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_biUnion_disjoint}.}
          The general overlapping statement
          $\operatorname{inj}(A)\wedge\operatorname{inj}(B)\Rightarrow
          \operatorname{inj}(A\cup B)$ is now closed for positive bond
          dimensions.\footnote{Formal declaration:
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_union_overlap}, with
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_union_disjoint} the
          disjoint specialization.} The displayed regions $R$ and $S$ are unions
          of \emph{overlapping} rectangles, so they require this four-region
          two-inverse descent rather than the disjoint engine.

          \emph{The closure (the overlap re-insertion).} The remaining
          obstruction recorded below was the $P_0$-outer freedom of the right
          rebuild: the right-rebuild row is indexed by $R_2$ boundary
          configurations, and a row over $R_2$ alone cannot separate the
          $P_0$-outer host indices. The closure carries those indices through the
          re-insertion as the source's open legs of the final tensor. The union
          boundary edges partition into the $R_2$ boundary edges and the
          $P_0$-outer edges (the union boundary edges running from
          $R_1\setminus R_2$ to $(R_1\cup R_2)^c$), and the $P_0$-outer edges are
          disjoint from the $R_2$, overlap, and difference boundary edges while
          lying on the boundary of $R_1$. Fixing the $P_0$-outer label to a
          reference $\delta$ and restricting the coefficient family to the
          $P_0$-fiber of $\delta$ gives a row whose right coupling, read through
          the overlap-crossing collapse and the bridge coefficient identity, is a
          sum of the left first strips restricted to the same fiber. Each
          restricted strip is either zero (the $R_1$ boundary label determines the
          $P_0$-outer label, so a mismatched fiber kills it) or the full strip
          (which already vanishes). The right rebuild then forces the fiber row to
          vanish after inverting $R_2$, and the partition determinacy selects a
          single host term: fixing $\delta$ to the $P_0$-outer part and $b_2$ to
          the $R_2$ part of a configuration realizing a host label forces the
          coefficient to vanish there. Union host-boundary surjectivity covers
          every label.\footnote{Formal declarations:
          \leanid{TNLean.PEPS.p0OuterLabel},
          \leanid{TNLean.PEPS.regionBoundaryLabel_union_eq_of_R₂_p0Outer},
          \leanid{TNLean.PEPS.overlapLeft_firstStrip_fiber_weightCombination_eq_zero},
          \leanid{TNLean.PEPS.overlapFiber_bridge_rightCoupling_eq_zero},
          \leanid{TNLean.PEPS.overlapFiber_bridgeRow_eq_zero}, and
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_union_overlap}.} The
          analysis below records how this closure was reached and is retained as
          the development history.

          The two host three-block geometries of the overlapping descent are now
          formalized over a bare geometry, together with the first inverse
          application. With the four parts $P_0=A\setminus B$, $P_1=A\cap B$,
          $P_2=B\setminus A$, and $P_3=(A\cup B)^c$, the host of both geometries
          is $A\cup B=P_0\cup P_1\cup P_2$ and their shared inner block is the
          residual coupling block $P_2=B\setminus A$. The left geometry places
          $A$ as the blue block and $P_2$ as the complement block; the right
          geometry places the overlap $P_1$ as the blue block and $P_2$ as the
          complement block over the smaller host $B$. The first inverse
          application then strips $A$ from the host annihilation and leaves the
          residual coupling through $P_2$ vanishing for every $P_2$ physical leg
          and every $A$-boundary configuration.\footnote{Formal declarations:
          \leanid{TNLean.PEPS.overlapLeftGeometry},
          \leanid{TNLean.PEPS.overlapRightGeometry}, and
          \leanid{TNLean.PEPS.overlap_firstStrip}, all built on the bare
          three-block factorization machinery of
          \leanid{TNLean.PEPS.ThreeBlockGeometry}.}

          The remaining sub-step is the re-insertion bridge of the source's
          ``plug back the tensor over $A\cap B$'' move, which closes through the
          right geometry over the smaller host $B$. The two residual couplings
          share the $P_2$ block but constrain different edge sets. The left
          geometry's coupling leaves the internal $P_1$--$P_0$ edges free, since
          both endpoints of such an edge lie in $A$ and the edge is neither an
          $A\cup B$-boundary edge nor an $A$-boundary edge. The right geometry's
          $B$-host constraint pins exactly these edges, because a $P_1$--$P_0$
          edge crosses the boundary of $B$ ($P_1\subseteq B$, $P_0\cap B=
          \emptyset$). The bridge is therefore the multiplicity collapse over the
          $P_1$--$P_0$ crossing edges: contracting the overlap tensor along its
          boundary with $A\setminus B$ rebuilds the $B$-block weight from the left
          geometry's $P_2$ coupling, summed over the free $P_1$--$P_0$ indices.
          This is the larger-host analogue of the landed complement collapse
          \leanid{TNLean.PEPS.ThreeBlockGeometry.regionBlockedWeight_complement_eq_smul_constrained}
          and
          \leanid{TNLean.PEPS.ThreeBlockGeometry.blueRedCrossingBondProd_smul_threeBlockComplCoeff_eq},
          with the crossing-edge family taken to be the $P_1$--$P_0$ edges and the
          produced block taken to be $B$. After the collapse, the right
          geometry's blue-side factorization
          \leanid{TNLean.PEPS.ThreeBlockGeometry.regionInteriorBondProd_smul_threeBlockBlueWeight_eq}
          reads the rebuilt coupling as a combination of $B$ blocked weights, the
          left inverse for $B$ forces the reconstructed host residual coefficients
          to vanish, and host-boundary surjectivity
          (\leanid{TNLean.PEPS.ThreeBlockGeometry.exists_regionBoundaryLabel_host_eq})
          makes every host residual realized, closing $c=0$. The new content is
          the $P_1$--$P_0$ crossing collapse; the rest reuses the landed
          three-block factorizations and the first inverse application
          \leanid{TNLean.PEPS.overlap_firstStrip}.

          The right-geometry blue-side rebuild step is now landed:
          \leanid{TNLean.PEPS.overlapRight_bondProd_smul_hostWeight_combination_eq_zero}
          takes a coefficient row over the $B$-host boundary configurations whose
          right-coupling combination
          $\sum_{b_2} \mathrm{row}(b_2)\cdot\mathrm{cc}_R(b_2,\sigma_{P_2},\beta_{P_1})$
          vanishes for every difference leg $\sigma_{P_2}$ and overlap boundary
          $\beta_{P_1}$, and produces a vanishing $B$-host-weight combination read
          through the fused overlap/difference leg, by the blue-side factorization
          \leanid{TNLean.PEPS.regionInteriorBondProd_smul_threeBlockBlueWeight_eq}.
          The $P_1$--$P_0$ crossing edges are also characterized
          (\leanid{TNLean.PEPS.IsOverlapCrossingEdge}): both endpoints in $R_1$,
          a $B=R_2$ boundary edge but neither an $A=R_1$ nor an $A\cup B$ boundary
          edge, and a blue/red crossing edge of the right geometry
          (\leanid{TNLean.PEPS.isBlueRedCrossingEdge_right_of_overlapCrossing},
          using $P_0=R_1\setminus R_2\subseteq R_2^c$).

          The $P_0$-outer bridge is now landed. The $P_0$-outer edges (the
          $A\cup B$ boundary edges that are not $B$ boundary edges, running
          $P_0\to(A\cup B)^c$) are formalized
          (\leanid{TNLean.PEPS.IsP0OuterEdge}) together with the gluing along
          them (\leanid{TNLean.PEPS.p0OuterGlue}) and along the $P_1$--$P_0$
          crossing edges (\leanid{TNLean.PEPS.interCrossGlue}). The two host
          residuals are reconciled by two gathering identities
          (\leanid{TNLean.PEPS.sum_hostGlue_mul_rightIndicator_eq} and
          \leanid{TNLean.PEPS.sum_interGlue_mul_leftIndicator_eq}), each proved by
          uniqueness of the realizing $B$- respectively $A$-boundary label (the
          determinacy lemmas
          \leanid{TNLean.PEPS.regionBoundaryLabel_R₂_eq_of_inter_sdiff} and
          \leanid{TNLean.PEPS.regionBoundaryLabel_R₁_eq_of_union_sdiff}) plus the
          gluing; both couplings reduce, by the landed crossing collapse, to the
          same $P_2$ blocked-region weights with an existence-indicator
          coefficient (\leanid{TNLean.PEPS.ThreeBlockGeometry.crossingBond_smul_complCoeff_combination_eq}).
          The scalar heart of the bridge is
          \leanid{TNLean.PEPS.overlapBridge_coeff_eq}. With the
          boundary-configuration transport between the geometry-native host
          $\mathrm{univ}\setminus g.\mathrm{red}$ and the literal union regions
          (\leanid{TNLean.PEPS.regionBoundaryConfigCongr}), the first strip read
          through the crossing collapse vanishes
          (\leanid{TNLean.PEPS.overlapLeft_firstStrip_weightCombination_eq_zero}),
          and the bridge produces the rebuild hypothesis:
          \leanid{TNLean.PEPS.overlap_bridge_rightCoupling_eq_zero} discharges
          $\sum_{b_2}\mathrm{row}(b_2)\cdot\mathrm{cc}_R(b_2,\sigma_{P_2},
          \beta_{P_1})=0$ for the bridge row
          \leanid{TNLean.PEPS.overlapBridgeRow}.

          The single remaining sub-step is the final $c=0$ extraction. The
          summed bridge row, fed to the rebuild and inverted by injectivity of
          $B$, yields the vanishing of $\sum_{\mathrm{bdry}}c(\mathrm{bdry})\cdot
          \mathbb{1}[\exists q:\mathrm{host}\,q=\mathrm{bdry}\wedge\mathrm{lab}_B
          \,q=b_2]$ for every $B$-boundary $b_2$. This pins $c$ only up to the
          $P_0$-outer freedom (the host $A\cup B$ residual is determined by the
          pair ($A$-boundary, $B$-boundary), so a row over $B$ alone cannot
          separate the $P_0$-outer indices). The source pins those indices
          through the $A$-boundary parameter of the first strip, which is the
          open-leg parameter of its ``$X$ with open $A$-legs''. The closure
          therefore needs the $A$-boundary-parametrized rebuild row (route (b)
          of the prior analysis): inverting $B$ at each fixed $A$-boundary
          configuration $\beta_1$, then reading the unique host residual
          reconstructed from the pair ($\beta_1$, $B$-boundary), forces each $X$
          coefficient to vanish.

          Three pieces of this closure are now formalized, and the precise
          remaining obstruction is isolated. The host-reconstruction
          determinacy is in place: the host $A\cup B$ boundary label is
          determined by the pair ($A$-boundary, $B$-boundary), so once a
          $\beta_1$-parametrized row over $B$ vanishes, the unique host label
          reconstructed from ($\beta_1$, $b_2$) makes the final extraction a
          single-term selection forcing $c=0$ at every realizable host label,
          and host-boundary surjectivity covers every label. The difference
          $B\setminus A$ label is likewise determined by the host and
          $A$-boundary labels. The overlap-crossing multiplicity collapse --- the
          bond-level content of the source's ``plug back the tensor over
          $A\cap B$'' move --- is also formalized: the joint existence indicator
          over (host, $A$-boundary $\beta_1$, overlap, difference) factors as the
          product of the overlap-glue indicator over ($A$-boundary, overlap) and
          the left first-strip indicator over (host, $A$-boundary, difference),
          the glue re-contracting the overlap along the $P_1$--$P_0$ crossing
          edges using the shared $A$-boundary label $\beta_1$.\footnote{Formal
          declarations:
          \leanid{TNLean.PEPS.regionBoundaryLabel_union_eq_of_R₁_R₂},
          \leanid{TNLean.PEPS.regionBoundaryLabel_sdiff_eq_of_union_R₁}, and
          \leanid{TNLean.PEPS.overlapJointIndicator_eq_interGlue_mul_leftIndicator}.}

          What remains is not a port of the landed summed-row machinery but a
          $\beta_1$-parametrized \emph{rebuild}. The landed rebuild
          \leanid{TNLean.PEPS.overlapRight_bondProd_smul_hostWeight_combination_eq_zero}
          consumes a row over the $B$-boundary configurations. Such a row cannot
          carry the $\beta_1$-constraint on the overlap-to-difference edges (the
          edges from $A\cap B$ to $B\setminus A$): these are $A$-boundary,
          overlap-boundary, and difference-boundary edges, but they are
          \emph{not} $B$-boundary edges (both endpoints lie in $B$), so they are
          invisible to a row indexed by $B$-boundary configurations. Folding
          $\beta_1$ into a $B$-row therefore overcounts: a $B$-row witness
          carrying $A$-boundary $=\beta_1$ and a separate $B$-row witness
          carrying the overlap and difference labels need not agree on the
          overlap-to-difference edges, so the $B$-row gathering does not reduce
          to the $\beta_1$-restricted first strip. The closure consequently
          requires a rebuild whose row spans the full host $A\cup B$ boundary
          (equivalently, a rebuild carrying the open-$A$-legs $\beta_1$ alongside
          the $B$-boundary, with a $\beta_1$-aware blue-side factorization),
          after which the overlap-crossing collapse above and the left first
          strip discharge the right coupling for each $\beta_1$ and the
          host-reconstruction determinacy closes $c=0$. This $\beta_1$-spanning
          rebuild is the genuine remaining mathematical step.

          The host-spanning rebuild is now landed, and the remaining obstruction
          is isolated to a single inversion. The host three-block geometry over
          $A\cup B$ that places $B$ as its blue block, the difference $A\setminus
          B$ as its complement block, and $(A\cup B)^c$ as its red block is
          formalized (\leanid{TNLean.PEPS.overlapHostGeometry}). Its blue/red
          crossings run between $B$ and $A\setminus B$, so the landed generic blue
          strip inverts $B$ on it: from a coefficient family annihilating the
          $A\cup B$ host weights, the host complement coupling vanishes for every
          difference leg over $A\setminus B$ and every $B$-boundary configuration
          (\leanid{TNLean.PEPS.overlapHost_complCoeff_combination_eq_zero}). The
          landed crossing collapse then reads this as a vanishing combination of
          the difference $A\setminus B$ blocked weights, the coefficient at the
          difference boundary $b_0$ being the host coefficient sum over residuals
          jointly realizable with $B$-boundary $b_2$ and difference boundary $b_0$
          (\leanid{TNLean.PEPS.overlapHost_differenceWeight_combination_eq_zero}).
          This coefficient is a single host coefficient: the union host boundary
          label is determined by the pair ($B$-boundary, $A\setminus B$-boundary)
          (\leanid{TNLean.PEPS.regionBoundaryLabel_union_eq_of_R₂_sdiff}, every
          union boundary edge being a $B$-boundary edge or a difference boundary
          edge), so the joint-indicator sum selects a single term equal to the
          coefficient at the unique realizing host label
          (\leanid{TNLean.PEPS.overlapHost_coeff_select}). The supporting
          region-equality transports for the rebuild's read-off and the
          surjectivity of the fused complement leg are landed
          (\leanid{TNLean.PEPS.regionBlockedWeight_congr},
          \leanid{TNLean.PEPS.ThreeBlockGeometry.complPhysical_surjective}).

          What remains is exactly the inversion of the difference $A\setminus B$
          weights. The host rebuild leaves a vanishing combination of the
          $A\setminus B$ blocked weights whose coefficient is a single host
          coefficient; were the $A\setminus B$ block injective, host-boundary
          surjectivity would close $c=0$ immediately. The hypotheses supply
          injectivity of the full regions $A$ and $B$, not of the strict sub-block
          $A\setminus B$. The blue-strip mechanism inverts a region only when that
          region is the blue block of a three-block geometry, that is, when its
          complement is disjoint from it; the difference $A\setminus B$ lies inside
          $A$, so injectivity of $A$ cannot be turned into an inversion of
          $A\setminus B$ by this mechanism. The source closes this instead through
          the ``plug back the tensor over $A\cap B$'' move, which re-couples the
          difference weights with the overlap $A\cap B$ into $A$ weights and
          inverts $A$ as a map of its physical legs, leaving the open-$A$-legs
          configuration $\beta_1$ uncontracted through the re-insertion. The two
          strips run independently --- inverting $A$ leaves a vanishing combination
          of $B\setminus A$ weights, inverting $B$ leaves a vanishing combination of
          $A\setminus B$ weights --- and neither difference block is injective on
          its own, so the genuine remaining step is the overlap re-insertion
          combining the two strips through the shared overlap block, which the
          present boundary-weight-combination formulation does not yet express.
          A simple $\beta_1$-parametrized row over $B$-boundary configurations does
          not suffice: a row indexed by $B$-boundary cannot see the
          overlap-to-difference edges, and the corresponding gathering identity
          fails (a configuration realizing the $B$-boundary half and a
          configuration realizing the difference half need not agree on the
          overlap-to-difference edges, so their two halves do not force the joint
          four-label witness).

          \emph{Resolution.} The obstruction above is that a row indexed by the
          full $A$-boundary over-determines the $R_1$ label on the
          overlap-to-difference edges, which lie inside $B$ and are invisible to a
          $B$-boundary row. The closure instead parametrizes by the $P_0$-outer
          label alone, the part of the $A$-boundary on the union boundary edges
          running into $(A\cup B)^c$. These edges are disjoint from the $R_2$,
          overlap, and difference boundary edges, so the $P_0$-outer label is an
          independent parameter that the right rebuild does not touch; the
          gathering glue along the $P_0$-outer edges then combines the two halves
          cleanly. This is the route described in the closure paragraph at the
          head of this obligation.
    \item Injectivity of the complementary block \(A_3\) is now realigned with
          the finite PEPS geometry and proved unconditionally for an interior
          edge. The local-window \(T\) region is a fixed small window whose own
          corner is forced and so has no rectangular cover; the source object is
          instead \(A_3\), the complement of the removed L-shape in the full
          finite lattice, which for an interior edge is the union of four
          full-length bands around the removed shape and two filler rectangles.
          Each band is a large contiguous rectangle, injective as a finite union
          of source \(2\times3\) (respectively \(3\times2\)) shapes; the union of
          the six pieces equals \(A_3\), so union closure proves it injective
          with no cover hypothesis. The same realignment holds for the rotated
          vertical complementary block. The earlier exact-cover criterion remains
          a useful sufficient condition, but it is no longer needed for interior
          edges.\footnote{Formal declarations:
          \leanid{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.wideRectangle_injective},
          \leanid{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.shortRectangle_injective},
          \leanid{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.interiorEdgeComplement_injective},
          and
          \leanid{TNLean.PEPS.NormalSquareLatticeRectangleInjectivityHypotheses.interiorVerticalEdgeComplement_injective}.}
    \item The every-edge finite-geometry construction is now cover-free at
          interior edges. Each interior horizontal or vertical edge gets a
          translated edge-blocking datum whose complementary-block injectivity is
          the interior \(A_3\) injectivity, with no rectangular cover; a family of
          these cover-free interior data over all edges assembles the normal
          edge-blocking hypotheses, with the injective chain, endpoint
          membership, pairwise disjointness, and coverage at every
          edge.\footnote{Formal declarations:
          \leanid{TNLean.PEPS.horizontalSquareLatticeEdge_blockingDatum_interior},
          \leanid{TNLean.PEPS.verticalSquareLatticeEdge_blockingDatum_interior},
          \leanid{TNLean.PEPS.NormalSquareInteriorEdgeDatum}, and
          \leanid{TNLean.PEPS.normalSquareEdgeBlockingHypotheses_of_interiorData}.}
          The remaining task is the boundary geometry of the open rectangular
          model, where some edges lack interior margins; on a translation-invariant
          (periodic) lattice every edge is interior and the construction is
          complete.
    \item Reuse the injective-chain insertion route after this blocking:
          \ghissue{1367}, \ghissue{1368}, \ghissue{1363}, \ghissue{1364},
          \ghissue{1361}, and \ghissue{1360}. The formal material for this route
          separates into unconditional region-insertion algebra, the
          identity-insertion bridge to the closed state, and a conditional
          matrix-transfer datum.

          The region-inserted coefficient is linear in the inserted matrix and,
          under the same positive-bond restriction as in the edge case, determines
          that matrix from all region and complement physical configurations.
          These facts use linear independence of the blocked weight family of
          \(R\) and of the complement \(V\setminus R\), and give the region
          analogue of the injectivity of the edge-inserted coefficient.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.regionInsertedCoeff_add},
          \leanid{TNLean.PEPS.regionInsertedCoeff_smul},
          \leanid{TNLean.PEPS.regionInsertedCoeff_injective}, and
          \leanid{TNLean.PEPS.edgeInsertedCoeff_injective}.}
          Once an explicit edge matrix transfer has matched region-inserted
          coefficients, is multiplicative, and preserves the unit, it gives a
          region insertion-algebra isomorphism and hence, by Skolem--Noether, a
          per-edge gauge matrix.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.RegionInsertionTransfer},
          \leanid{TNLean.PEPS.isRegionBlockedInsertionAlgebraIsomorphism_of_transfer}, and
          \leanid{TNLean.PEPS.exists_regionEdgeGauge_of_transfer}.}

          The identity-insertion bridge is the following equation. Here
          \(C_A(R,f,X;\sigma,\tau)\) denotes the region-inserted coefficient for
          the tensor \(A\), the region \(R\), the boundary edge \(f\), the boundary
          matrix \(X\), the region physical configuration \(\sigma\), and the
          complement physical configuration \(\tau\). The symbol \(D_A(e)\)
          denotes the bond dimension of the edge \(e\), \(\partial R\) denotes the
          set of boundary edges of \(R\), and
          \(\langle\sigma,\tau\mid\Psi_A\rangle\) denotes the closed-state
          coefficient of the assembled physical configuration:
          \[
              C_A(R,f,1;\sigma,\tau)
              =
              \Bigl(\prod_{e\notin \partial R} D_A(e)\Bigr)
              \langle \sigma,\tau \mid \Psi_A\rangle
          \tag{4.1}
          \]
          The proof decomposes the closed coefficient into region and complement
          contractions, rewrites identity insertion as a double sum over
          boundary-agreeing virtual configurations, and collapses that double sum
          by merging the two virtual configurations. The factor
          \(\prod_{e\notin\partial R}D_A(e)\) is therefore part of the formula,
          not a normalization that can be discarded.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.prod_assembleRegionσ_split},
          \leanid{TNLean.PEPS.regionInsertedCoeff_identity_eq_doubleSum},
          \leanid{TNLean.PEPS.regionMerge},
          \leanid{TNLean.PEPS.regionProd_eq_merge},
          \leanid{TNLean.PEPS.complementProd_eq_merge},
          \leanid{TNLean.PEPS.regionInteriorBondProd},
          \leanid{TNLean.PEPS.stateCoeff_eq_regionComplement}, and
          \leanid{TNLean.PEPS.regionInsertedCoeff_one_eq_stateCoeff}.}
          Under matched bond dimensions the corresponding interior bond products
          for \(A\) and \(B\) agree, so the same factor is available in the
          matched-coefficient transfer.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.regionInteriorBondProd_congr} and
          \leanid{TNLean.PEPS.regionInsertedCoeff_transfer_of_realizes}.}

          For a second tensor \(B\), let \(T_f(X)\) denote the boundary matrix
          assigned to \(X\) on the edge \(f\) by the desired region insertion
          transfer. The formal candidate pulls the in-region endpoint operator of
          \(A\) back through \(B\)'s vertex and reads it off through a fixed
          residual reference frame, in analogy with the edge transfer matrix.
          The available recovery theorem is conditional: it reduces the matched
          coefficient identity to two hypotheses, namely that the transferred
          endpoint operator preserves the image of \(B\)'s local tensor map at the
          in-region vertex and that its virtual pullback through \(B\) has the
          single-edge insertion form on \(f\).\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.regionTransferMatrix},
          \leanid{TNLean.PEPS.edgeTransferMatrix}, and
          \leanid{TNLean.PEPS.regionTransferMatrix_realizes_of_image}.}
          These are the region analogues of the two consequences extracted by the
          edge physical-to-virtual recovery.\footnote{
          Formal declaration:
          \leanid{TNLean.PEPS.physical_to_virtual_insertion}.}

          A useful image-preservation argument is available only in the
          vertex-injective regime. If \(B\) is vertex injective, then the
          state-vector coefficients at the in-region vertex span the local
          virtual coefficient space by row--column rank duality for the
          vertex-complement tensor family of \(B\). Under
          \(\operatorname{SameState}(A,B)\) this identifies the local tensor images
          of \(A\) and \(B\), and the transferred endpoint operator preserves that
          image.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.span_stateOpenCoeff_eq_top},
          \leanid{TNLean.PEPS.vertexComplementTensorInjective_of_isVertexInjective},
          \leanid{TNLean.PEPS.range_localTensorMap_eq_span_stateVec},
          \leanid{TNLean.PEPS.range_localTensorMap_eq_of_sameState}, and
          \leanid{TNLean.PEPS.regionInsertionOp_localProjectorAt_eq}.}
          This does not discharge the normal-PEPS route of
          \cite[Section~3]{Molnar2018NormalPEPS}: the source theorem assumes
          injectivity after blocking, not single-vertex injectivity. The
          region-injective proof still requires the same spanning and
          image-preservation statement derived from blocked-region injectivity.

          Even in the vertex-injective setting, the full realization is
          conditional on the remaining matrix-structure transfer: the virtual
          pullback of the transferred endpoint operator must be the matrix
          insertion of \(T_f(X)^{\mathsf T}\) on \(f\).\footnote{
          Formal declaration:
          \leanid{TNLean.PEPS.regionTransferRealizesAt_of_hform}.}
          Equivalently, the change of frame between \(A\)'s and \(B\)'s coefficient
          spaces at the in-region vertex must intertwine the single-bond
          insertions on \(f\). This is the region analogue of the incident-matrix
          form read from the edge resonate identity.

          This matrix-structure transfer is now reduced to a single structural
          fact. Because the region transfer matrix is, by definition, the read-off
          of the virtual pullback, the matrix-structure hypothesis holds exactly
          when that pullback is of single-bond insertion form on \(f\); the
          read-off then inverts the insertion and the round-trip closes the
          hypothesis. The image-preservation half is unconditional in the
          vertex-injective setting, so the pullback already realizes the
          transferred endpoint operator on every local tensor image of \(B\) at the
          in-region vertex. What remains is to show the pullback is of single-bond
          insertion form, which is the residual-independence the edge resonate
          identity supplies.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.hform_of_isIncidentMatrixForm} and
          \leanid{TNLean.PEPS.localTensorMap_localVirtualOpOfPhysicalOpAt_regionInsertionOp}.}

          The pin available from the closed-state transfer alone is non-circular but
          insufficient on its own. Evaluating the transferred endpoint operator of
          \(A\) on \(B\)'s state vector at the endpoint physical leg recovers
          \(A\)'s region-inserted coefficient; varying the
          region configuration at the endpoint vertex makes the leg range over the
          whole physical space while the state vector stays fixed, so the pin determines
          the endpoint operator's output on every state vector. But this only equates
          the operator with the read-off matrix \emph{at the single reference residual
          configuration} used to define \(T_f\); extending it to all residual
          configurations is the residual-independence the read-off must satisfy, and
          a single pin per region configuration cannot supply it. The closed-state
          transfer reads only one bond contraction (through \(R\)'s endpoint vertex),
          so it cannot reconcile the two residual frames the way the edge proof does.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.regionInsertionOp_regionStateVec_pin} and
          \leanid{TNLean.PEPS.regionStateVec_update_vmem}.}

          The faithful route mirrors the edge proof's two-endpoint structure with
          \(R\) and its set complement \(\mathrm{univ}\setminus R\) as the two contracted
          blocks of the single boundary edge \(f\): the in-region endpoint vertex
          plays the role of one edge endpoint, the out-of-region endpoint vertex of
          \(f\) (the in-region endpoint of \(f\) for the complement region) plays the
          other, and the empty interior of the single edge \(f\) makes the middle
          block trivial. The region-inserted coefficient already carries this doubled
          boundary-configuration sum (one configuration read by \(R\), one by the
          complement, coupled by the inserted matrix on \(f\)). The missing steps are:
          (i) a complement-side realization, the analogue of
          the realization of the region-inserted coefficient applied to the complement
          region with \(f\) as its boundary edge; (ii) a region resonate identity
          equating the \(v\)-side and complement-side
          realizations of \(A\)'s region-inserted coefficient, transferred to \(B\) by
          \(\operatorname{SameState}\); (iii) inverting both endpoints through the
          blocked-region left inverses (using the blocked-region
          injectivity of \(R\) and of the complement); and (iv) a region reconcile
          forcing
          the two read-off matrices to coincide. This is the region-granularity port
          of the edge insertion-realization argument and remains the open
          obligation. Once it is supplied, the transfer datum and the per-edge gauge
          follow mechanically.

          Step (i) now has a cast-free reading. The complement-side reading is the
          region-inserted coefficient read with the region and its set complement
          exchanged and the inserted matrix transposed on \(f\). The exchange is the
          block-swap symmetry of the abstract two-block inserted coefficient, which
          stays over \(R\)'s boundary-edge index set and so avoids the dependent-type
          reindexing \(\mathrm{univ}\setminus(\mathrm{univ}\setminus R)=R\). What still
          requires that reindexing is factoring the complement block through the
          out-of-region endpoint vertex's local tensor map for the inversion of step
          (iii).\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.twoBlockInsertedCoeff_swap} and
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_complementTwoBlock}.}

          Steps (ii) and the inversion infrastructure of step (iii) are now
          formalized. The out-of-region-endpoint pin reads the region-inserted
          coefficient through the out-of-region endpoint vertex by instancing the
          in-region pin at the set complement and casting back through the
          complement-side cast identity. Equating it with the in-region-endpoint pin
          gives the region resonate identity: the in-region and out-of-region endpoint
          operators of the first tensor, applied to the second tensor's closed state
          vectors, agree at the two endpoint legs. The region-inserted coefficient
          factors, as a function of the complement (resp. region) physical
          configuration, through the blocked-region tensor map of the set complement
          (resp. of the region); each blocked-region left inverse reads off the
          corresponding row function. What remains is the reconcile of step (iv):
          exhibiting the first tensor's region-inserted coefficient, as a function of
          the region physical configuration, in the image of the second tensor's
          region blocked tensor map (the membership that the resonate transfer across
          \(\operatorname{SameState}\) supplies), then matching the two read-off row
          functions and concluding incident-matrix form of the virtual pullback
          through \leanid{TNLean.PEPS.hform_of_isIncidentMatrixForm}.

          The remaining content of step (iv) is now isolated to a single predicate.
          Incident-matrix form of the virtual pullback, for every inserted matrix,
          assembles the realization bundle and hence the region insertion transfer
          datum and the per-edge gauge, all unconditionally on top of that predicate.
          The predicate names exactly the reconcile: the virtual pullback of the
          transferred in-region endpoint operator is of single-bond insertion form on
          the boundary edge. With it in both directions the per-edge gauge follows
          mechanically, so the open obligation is reduced to establishing this
          predicate from the region resonate identity and the two blocked-endpoint
          inversions.

          The reconcile predicate is in turn reduced to a single coefficient
          identity. The leg-wise pin reads the in-region endpoint operator of the
          first tensor, applied to the second tensor's closed state vector, leg by
          leg as the first tensor's region-inserted coefficient scanned over the
          endpoint physical configuration. The closed state-vector coefficients span
          the full local virtual coefficient space at the in-region endpoint vertex,
          so a realization on the state vectors extends to all coefficients. Together
          these show: if for every inserted matrix on the first tensor's bond there
          is a matrix on the second tensor's bond whose region-inserted coefficient
          equals the first's at every physical configuration, then the reconcile
          predicate holds, with the read-off matrix the transpose of the matching
          matrix. The whole remaining obligation is therefore exactly this
          coefficient transfer: producing, for each inserted matrix, a matching
          matrix on the second tensor. This is the content the region resonate
          identity and the two blocked-endpoint inversions must supply; it is
          equivalent to the insertion-algebra isomorphism of the source lemma.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.regionInsertionOp_regionStateVec_pin_leg},
          \leanid{TNLean.PEPS.regionInsertionOp_realizes_of_realizes_on_stateVec},
          \leanid{TNLean.PEPS.regionInsertionOp_regionStateVec_eq_of_coeff_eq},
          \leanid{TNLean.PEPS.isIncidentMatrixForm_of_coeff_eq}, and
          \leanid{TNLean.PEPS.regionResonateReconcile_of_coeff_transfer}.}

          The coefficient transfer is itself reduced to a single structural
          statement about a doubly-indexed transfer coefficient. The first tensor's
          region-inserted coefficient factors through both blocked endpoints of the
          second tensor at once: the v-side complement factorization gives the
          coefficient as a function of the complement physical configuration, and
          the symmetric region factorization (read off the complement-side cast
          identity and transported across the double complement) gives it as a
          function of the region physical configuration. Combining the two, the
          coefficient is the double sum of a transfer coefficient against the second
          tensor's region and complement blocked weights, with the transfer
          coefficient read off by the two blocked-endpoint left inverses; the
          required membership in the second tensor's complement block image holds
          because the region row is a finite linear combination of the v-side
          complement-image coefficients. Once the transfer coefficient is shown to
          have the incident-matrix coupling form of a single matrix on the boundary
          bond --- coupling only the boundary-edge legs and contracting the residual
          legs by the identity --- that matrix is the matching matrix the
          coefficient transfer asks for. This incident-matrix form of the transfer
          coefficient is the only remaining content; it is the region analogue of
          the source step \(V=W\), forced by the resonate identity together with the
          full injectivity of both blocked endpoints.

          The incident-matrix form of the transfer coefficient is now reduced, by a
          sorry-free chain, to the incident-matrix form of the virtual pullback
          itself. The transfer-coefficient column at a fixed complement boundary
          configuration is the second tensor's region blocked left inverse of the
          v-side row. The v-side row is the first tensor's in-region endpoint
          operator, from the inserted matrix, applied to the second tensor's region
          weight vector; since that weight vector is the second tensor's local tensor
          map of the region open coefficient, the unconditional image-preservation
          realization rewrites the v-side row as the second tensor's local tensor map
          of the virtual pullback applied to the region open coefficient. If the
          virtual pullback is of incident-matrix form, the second tensor's own
          in-region endpoint operator of that matrix realizes it, and the inner-sum
          realization expands the v-side row to the incident-matrix sum against the
          region blocked weights; the region blocked left inverse then collapses each
          weight to the standard basis configuration, reading off the incident-matrix
          coupling of the transfer coefficient. The reconcile predicate
          \leanid{TNLean.PEPS.RegionResonateReconcile} is exactly this
          incident-matrix form of the virtual pullback for every inserted matrix, so
          the whole open obligation is now this single predicate. Establishing it
          requires the out-of-region (complement-vertex) inversion of the region
          resonate identity, the region analogue of the right-endpoint inversion that
          supplies the residual-independence of the read-off matrix at the edge
          level; the in-region pin alone fixes the matrix only at the reference
          residual.\footnote{
          Formal declarations (the sorry-free reduction):
          \leanid{TNLean.PEPS.transferCoeff_column_eq_regionBlockedLeftInverse_vSideRow},
          \leanid{TNLean.PEPS.vSideRow_eq_localTensorMap_virtualPullback},
          \leanid{TNLean.PEPS.vSideRow_eq_incidentSum_of_virtualPullback_incidentForm},
          and
          \leanid{TNLean.PEPS.transferCoeff_eq_incidentForm_of_virtualPullback_incidentForm}.}\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_complement_blockedMap_B},
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_region_blockedMap_B},
          \leanid{TNLean.PEPS.regionBlockedTensorMap_doubleCompl},
          \leanid{TNLean.PEPS.regionRowB_mem_range},
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_doubleSum_transferCoeff},
          \leanid{TNLean.PEPS.vSideRow_eq_region_blockedMap_transferCoeff}, and
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_of_transferCoeff_form}.}\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.RegionResonateReconcile},
          \leanid{TNLean.PEPS.regionTransferRealizes_of_isIncidentMatrixForm},
          \leanid{TNLean.PEPS.regionTransferRealizes_of_reconcile},
          \leanid{TNLean.PEPS.regionInsertionTransfer_of_reconcile}, and
          \leanid{TNLean.PEPS.exists_regionEdgeGauge_of_reconcile}.}\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.regionInsertionOp_regionStateVec_pin_compl},
          \leanid{TNLean.PEPS.region_resonate_identity},
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_complement_blockedMap},
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_region_blockedMap},
          \leanid{TNLean.PEPS.regionBlockedLeftInverse_complement_regionInsertedCoeff}, and
          \leanid{TNLean.PEPS.regionBlockedLeftInverse_region_regionInsertedCoeff}.}\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.regionInsertedCoeff},
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_regionRealizationSum},
          \leanid{TNLean.PEPS.resonate_invert_right_endpoint},
          \leanid{TNLean.PEPS.resonate_invert_left_endpoint},
          \leanid{TNLean.PEPS.resonate_endpoint_coeff_reconcile},
          \leanid{TNLean.PEPS.regionInsertionTransfer_of_realizes}, and
          \leanid{TNLean.PEPS.exists_regionEdgeGauge_of_transfer}.}

          The reconcile above is also reachable in a \emph{block frame} that uses no
          single-vertex injectivity. The foundational input is the block-level image
          coincidence: under \(\operatorname{SameState}\), with the complement block
          blocked-tensor injective and positive bond dimensions, the range of the
          region blocked tensor map is the same for the two tensors, because both
          ranges equal the span of the partial states across the region cut, and the
          partial states depend only on the closed state coefficient. The same fact
          for the complement block follows by transporting blocked-tensor injectivity
          across the double complement. With these two image coincidences, the first
          tensor's region-inserted coefficient, read as a function of the region
          physical configuration, lies in the second tensor's region block image
          (region side), and the transferred row, as a function of the complement
          physical configuration, lies in the second tensor's complement block image
          (complement side); the two blocked left inverses read off the same transfer
          coefficient as in the vertex frame, and the first tensor's coefficient is
          the double sum of that transfer coefficient against the second tensor's two
          blocked weights. None of this uses single-vertex injectivity.

          In the block frame the remaining content is isolated cleanly. By double
          blocked injectivity of the second tensor, the double-sum representation of a
          coefficient by a transfer coefficient is unique, so the transfer
          coefficient of an inserted matrix has the incident-matrix coupling form of a
          single bond matrix if and only if the first tensor's region-inserted
          coefficient of that inserted matrix equals the second tensor's of that bond
          matrix. The whole open obligation is therefore equivalent to the coefficient
          transfer alone --- for every inserted matrix on the first tensor's bond, a
          matching bond matrix on the second tensor --- with the matching matrix
          carried directly by the equivalence. The identity insertion is matched
          across \(\operatorname{SameState}\), so the transfer coefficient of the
          identity is already the incident-matrix coupling of the identity. The
          residual open content is the incident-matrix coupling form of the transfer
          coefficient for a general inserted matrix: the transfer coefficient must be
          local on the boundary bond, depending on the two boundary configurations
          only through their bond legs. This locality is the bond-level content of the
          source step \(V=W\) and is the one fact the block-frame foundation does not
          supply, because the block image coincidence preserves only the full weight
          span and not the per-leg decomposition across the boundary bond.\footnote{
          Formal declarations (block frame, sorry-free):
          \leanid{TNLean.PEPS.range_regionBlockedTensorMap_eq_of_sameState},
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_doubleCompl},
          \leanid{TNLean.PEPS.range_regionBlockedTensorMap_compl_eq_of_sameState},
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_blockTransferRow},
          \leanid{TNLean.PEPS.blockTransferRow_mem_range},
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_doubleSum_transferCoeff_block},
          \leanid{TNLean.PEPS.doubleSum_kernel_injective},
          \leanid{TNLean.PEPS.transferCoeff_eq_incidentKernel_iff_coeff_eq},
          \leanid{TNLean.PEPS.transferCoeff_one_eq_incidentKernel_one},
          \leanid{TNLean.PEPS.coeffTransfer_iff_transferCoeff_incidentForm}, and
          \leanid{TNLean.PEPS.exists_regionEdgeGauge_of_blockCoeffTransfer}.}

          The block-frame foundation is now assembled around the edge-blocking
          triple. The union of two injective blocks is injective; applied to the blue
          and complement blocks of a one-edge blocking datum, this gives injectivity of
          the host block, the set complement of the red block, which every block-frame
          backbone lemma needs at the complement of the red region. With it, the four
          block injectivities of the two tensors at the red region and at its host are
          all read off a shared one-edge blocking datum, recording that the two tensors
          are blocked around the same edge into the same red, blue, and complement
          regions, together with positive bond dimensions. On top of these, the
          per-edge gauge is assembled from one predicate alone: bond locality of the
          transfer coefficient on the boundary edge, in both directions. Bond locality
          is proved equivalent to the coefficient transfer, so the open content of the
          block frame is now exactly this single predicate, with the per-edge gauge and
          all four injectivities derived from it and the shared blocking datum without
          single-vertex injectivity. The identity transfer of a single tensor is
          bond-local, the diagonal anchor of the cross-tensor transfer. The remaining
          open obligation is bond locality of the cross-tensor transfer for a general
          inserted matrix, the block-level content of the step \(V=W\); it is the only
          fact between the assembled foundation and the per-edge gauge.\footnote{
          Formal declarations (block frame, sorry-free):
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_union},
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_compl_red},
          \leanid{TNLean.PEPS.SharedNormalEdgeBlockingData},
          \leanid{TNLean.PEPS.IsBondLocalTransferKernel},
          \leanid{TNLean.PEPS.bondLocal_iff_coeffTransfer},
          \leanid{TNLean.PEPS.coeffTransfer_of_bondLocal},
          \leanid{TNLean.PEPS.isBondLocalTransferKernel_self}, and
          \leanid{TNLean.PEPS.SharedNormalEdgeBlockingData.exists_regionEdgeGauge}.}

          The block-frame transfer kernel \(K_M(\mu,\nu')\) is the double blocked
          left inverse of the first tensor's region-inserted coefficient of \(M\)
          through the second tensor's region and complement blocks. Its bond locality
          \leanid{TNLean.PEPS.IsBondLocalTransferKernel} is the incident-matrix coupling
          form: the kernel couples the two boundary configurations only through their
          legs on the boundary edge \(f\). The single remaining open step is the port of
          the edge resonate inversion to this block granularity, which hits one
          structural mismatch that is not a matter of length. The edge inversion
          \leanid{TNLean.PEPS.resonate_invert_right_endpoint} consumes the
          bond-contracted resonate identity
          \leanid{TNLean.PEPS.resonate_middle_inverted}, which pairs the
          left-endpoint and right-endpoint physical operators \emph{indexed by the
          shared bond} \(k\): the two endpoints are two single vertices sharing one
          bond, and the two operators act on the same physical space \(\mathbb C^d\).
          In the block frame the two endpoints are the red region and its complement
          sharing the boundary edge \(f\), and the only available factorizations of the
          first tensor's coefficient through the second tensor's complement block in the
          bond-\(f\)-paired form needed by the inversion --- the v-side and
          \(\sigma\)-side blocked maps
          (\leanid{TNLean.PEPS.regionInsertedCoeff_eq_complement_blockedMap_B},
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_region_blockedMap_B}) ---
          carry the single-vertex hypothesis \(\text{LinearIndependent}\,(A_v)\) at the
          endpoint vertex of \(f\). The block-image coincidence
          \leanid{TNLean.PEPS.range_regionBlockedTensorMap_eq_of_sameState} supplies
          the full-weight-span identity through the second tensor's whole region block
          as a single basis change, but it does not supply the per-\(f\)-leg bond
          pairing the inversion consumes; the only block-only complement-side reading
          of the coefficient is through the double left inverse \(K_M\) itself, which is
          the very kernel whose bond locality is the goal. The block-frame port
          therefore requires a vertex-injectivity-free bond-\(f\)-paired complement
          factorization of the coefficient before the edge inversion structure
          transfers, and constructing that factorization is the remaining open step.

          The two reduction tracks now meet at a single named bridge. The
          block-frame predicate
          \leanid{TNLean.PEPS.IsBondLocalTransferKernel} follows from the
          virtual-pullback predicate
          \leanid{TNLean.PEPS.RegionResonateReconcile} by reading the incident-matrix
          form of the virtual pullback off as the incident-matrix coupling of the
          transfer kernel
          (\leanid{TNLean.PEPS.isBondLocalTransferKernel_of_regionResonateReconcile},
          with its coefficient-transfer corollary
          \leanid{TNLean.PEPS.coeffTransfer_of_regionResonateReconcile}). This bridge
          is unconditional on top of the reconcile predicate, but the reconcile and
          the read-off it feeds carry the single-vertex hypotheses
          \(\text{LinearIndependent}\,(A_v)\), \(\text{LinearIndependent}\,(B_v)\) at
          the endpoint vertex of \(f\), together with vertex injectivity of both
          tensors, because the image-preservation root
          \leanid{TNLean.PEPS.span_stateOpenCoeff_eq_top} --- the spanning of the
          closed state-vector coefficients at the endpoint vertex --- is derived only
          from \leanid{TNLean.PEPS.vertexComplementTensorInjective_of_isVertexInjective}.
          Thus the bridge closes the bond locality precisely in the vertex-injective
          regime, isolating the residual normal-setting content to one fact: the same
          endpoint-vertex spanning, derived from blocked-region injectivity of the red
          region and its complement rather than from single-vertex injectivity. The
          endpoint vertex of \(f\) is a single vertex of a normal (region-injective,
          not vertex-injective) tensor and need not itself be injective, so no
          double-complement transport of the present
          \leanid{TNLean.PEPS.range_regionBlockedTensorMap_eq_of_sameState} reaches it;
          this endpoint-vertex spanning from blocked-region injectivity is the
          precise remaining open step.

          \emph{The coarse three-site route avoids the endpoint-vertex spanning
          entirely.} The endpoint-vertex obstruction above is an artifact of running
          the comparison on the original single vertex at the end of the boundary
          edge. The source never uses that vertex. After blocking, the three sites of
          the chain are the blocks themselves, and the component family of each
          super-site is the blocked-region weight family, injective by the blocking
          hypothesis. Realizing this literally, the three blocks are assembled into a
          fresh tensor on a three-vertex complete graph: the red block at one vertex,
          the blue block at a second, the complement block at the third, with one
          super-bond for the distinguished edge, one for the red-to-complement
          crossings, and one for the blue-to-complement crossings, and a uniform padded
          physical dimension carrying every region's physical configuration through a
          fixed surjection. Each super-site's component family is the blocked-region
          weight family reindexed by the super-site's legs and read through the padded
          physical surjection; reindexing by a bijection and pulling back along a
          surjection both preserve linear independence, so each super-site is injective
          directly from the blocked-region injectivity of its block, with no
          single-vertex hypothesis. The single-vertex three-site comparison then
          applies to this coarse tensor verbatim: the two endpoint super-sites are
          injective by the red and blue blocking injectivities, the middle super-site
          is the single complement super-site, injective by the complement blocking
          injectivity, and the proven contraction-of-injective-tensors argument
          supplies the middle-block injectivity of the coarse chain. The coarse chain
          is therefore edge-blocked three-site injective at its distinguished super-bond
          with no single-vertex injectivity of the original tensor, and the proven
          injective inserted-coefficient correspondence runs on the coarse super-bond,
          producing, for every matrix inserted on the coarse super-bond of the first
          coarse tensor, a matrix on the second giving the same coarse inserted
          coefficient at every coarse physical configuration.\footnote{Formal
          declarations:
          \leanid{TNLean.PEPS.coarseGraph},
          \leanid{TNLean.PEPS.CoarseBlockingFrame},
          \leanid{TNLean.PEPS.CoarseBlockingFrame.coarseTensor},
          \leanid{TNLean.PEPS.CoarseBlockingFrame.coarseTensor_isVertexInjective},
          \leanid{TNLean.PEPS.CoarseBlockingFrame.coarseTensor_edgeBlockedThreeSiteInjective},
          and \leanid{TNLean.PEPS.coarse_exists_edgeInsertedCoeff_eq}. The coarse
          construction and the injectivity transport are unconditional on top of the
          blocked-region injectivities, with no single-vertex spanning.}

          Two obligations remain to descend the coarse super-bond correspondence to the
          original bond locality. The first is the state gluing: the coarse state
          coefficient must equal a positive constant times the original state
          coefficient of the assembled physical configuration, so that the original
          same-state hypothesis transports to the coarse tensors. The coarse state
          coefficient is the sum over the three super-bonds of the product of the three
          blocked-region weights, the three-block analogue of the two-block region
          gluing already formalized for one region against its complement. This gluing
          carries a compatibility requirement the present coarse frame does not yet
          record: each super-bond is incident to two super-sites, so its value is read
          by both incident per-site leg identifications, and those two identifications
          must send the shared super-bond value to the same original crossing-bond
          value. The distinguished super-bond is incident to the red and blue
          super-sites, the red-to-complement super-bond to the red and complement
          super-sites, the blue-to-complement super-bond to the blue and complement
          super-sites; for each, the two incident leg identifications must agree on the
          shared crossing bonds. Without this agreement the three blocked-region weights
          are not glued along their shared crossing edges and the coarse state
          coefficient does not factor through the original state coefficient. The second
          obligation is the descent identity: the coarse inserted coefficient on the
          distinguished super-bond must equal a positive constant times the original
          region-inserted coefficient of the same matrix, the three-block analogue of
          the landed identity reading the region-inserted coefficient of the identity as
          a constant times the state coefficient. With the super-bond identified with
          the original distinguished bond, this descent turns the coarse super-bond
          correspondence into the bond locality of the original transfer kernel and
          instantiates the per-edge gauge from the source hypotheses. Both obligations
          are state-gluing combinatorics over the crossing bonds, the three-block
          extension of the existing one-region-against-complement gluing; neither
          reintroduces single-vertex spanning.

          \emph{The compatibility requirement of the first obligation is now recorded.}
          The independent per-super-site leg identifications of the plain coarse frame
          are refined by a coherent coarse frame that fixes one bond model per coarse
          super-edge --- an equivalence between the coarse bond and the original
          configurations on the edges crossing between the two incident regions --- and
          requires each super-site's leg identification to factor through these shared
          bond models on its incident super-edges. The geometric content is that every
          boundary edge of a region crosses to exactly one partner region, so the two
          super-edges incident to a super-site carry exactly the two crossing bundles of
          its region's boundary. The payoff is the shared-super-bond agreement: two
          super-sites incident to a super-edge, fed matching super-bond values, read
          every shared crossing edge to the same original virtual leg. This is the
          well-posedness the state gluing needs --- the three blocked-region weights are
          read at a single consistent assignment of the original crossing bonds.

          \emph{The well-posedness layer is now fully reduced to the per-edge bond
          models.} A partition hypothesis records that the three regions are pairwise
          disjoint and cover the vertex set, the geometry of the source edge blocking,
          detached from the tensor construction; from it the three regions form a
          three-block geometry, unlocking the landed three-block factorizations (the
          fused complement physical leg and the host vertex-product split over the
          complement of the red region). The partition gives the crossing
          classification: every boundary edge of a region crosses to exactly one
          partner region, so a red boundary edge is a red-to-blue or a red-to-complement
          crossing edge, and likewise at the blue and complement super-sites. Combined
          with the factoring fields, this expresses each super-site's leg identification
          of a coarse configuration entirely through the two bond models on its incident
          super-edges: the red boundary configuration reads each red boundary edge off
          the red-to-blue bond model (if it crosses to blue) or the red-to-complement
          bond model (if it crosses to the complement), and symmetrically at the blue
          and complement super-sites. For a single coarse configuration the three
          induced boundary configurations therefore agree on every shared crossing
          edge, the single-configuration form the merge collapse contracts against the
          original crossing edges.\footnote{Formal declarations (the well-posedness
          layer):
          \leanid{TNLean.PEPS.IsCrossingEdge},
          \leanid{TNLean.PEPS.CrossingConfig},
          \leanid{TNLean.PEPS.CoherentCoarseBlockingFrame},
          \leanid{TNLean.PEPS.CoarseBlockingFrame.IsPartition},
          \leanid{TNLean.PEPS.CoarseBlockingFrame.toThreeBlockGeometry},
          \leanid{TNLean.PEPS.CoarseBlockingFrame.isCrossingEdge_red_blue_or_red_complement},
          \leanid{TNLean.PEPS.CoarseBlockingFrame.isCrossingEdge_red_blue_or_blue_complement},
          \leanid{TNLean.PEPS.CoherentCoarseBlockingFrame.legEquivRed_apply_eq},
          \leanid{TNLean.PEPS.CoherentCoarseBlockingFrame.legEquivBlue_apply_eq},
          \leanid{TNLean.PEPS.CoherentCoarseBlockingFrame.legEquivComplement_apply_eq},
          and
          \leanid{TNLean.PEPS.CoherentCoarseBlockingFrame.legEquivRed_eq_legEquivBlue_on_rb_single}.}

          With this layer in place, the residual content of both obligations is the
          purely combinatorial three-region merge collapse: the triple sum over coarse
          virtual configurations of the boundary-coupled product of the three region
          weights refactors, through the partition of the original edges into the three
          regions' interiors and the three crossing bundles, as the product of the three
          regions' interior bond dimensions times the original closed-state coefficient
          of the assembled physical configuration. The single remaining mathematical
          step is the \emph{global fiber-collapse bijection}: the bond models give a
          bijection between the coarse virtual configurations and the original
          inter-region crossing configurations, under which the triple sum's
          pairwise-agreement constraints become the boundary-agreement of an original
          configuration triple, and merging that triple into one global configuration
          (region-incident edges read from the region's own configuration, the rest
          free) collapses the sum to the closed-state coefficient with the three
          non-incident bond products as multiplicity. This is the three-region extension
          of the two-block merge/fiber collapse
          \leanid{TNLean.PEPS.stateCoeff_eq_regionComplement} of the landed
          one-region-against-complement gluing --- a fiber count over the three regions'
          non-incident edges threading the per-edge bond models.

          This collapse is now formalized. The bond models assemble into one product
          equivalence between coarse virtual configurations and triples of original
          inter-region crossing configurations, under which the coarse state sum reindexes
          as a sum over triples of original virtual configurations agreeing on the crossing
          edges; merging each agreeing triple into one global configuration (red-incident
          edges read the red configuration, the remaining blue-incident edges the blue, the
          rest the complement) collapses the sum to the closed-state coefficient, the fiber
          over a merged configuration being parameterised by the virtual indices each
          configuration is free to choose away from its own region. The constant is the
          product of the three regions' non-incident bond products. The same-state
          hypothesis transports through the collapse: the assembled physical configuration
          depends only on the shared regions and the merge-collapse constants coincide under
          matched bond dimensions. None of this reintroduces single-vertex spanning.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.crossingTripleEquiv},
          \leanid{TNLean.PEPS.threeRegionSum_eq_agreeingTripleSum},
          \leanid{TNLean.PEPS.triFiber_card},
          \leanid{TNLean.PEPS.agreeingTripleSum_collapse},
          \leanid{TNLean.PEPS.stateCoeff_coarseTensor_collapse}, and
          \leanid{TNLean.PEPS.coarseTensor_sameState_of_sameState}.}

          What remained was the inserted-coefficient descent (the same collapse with a matrix
          carried on the coarse red-to-blue super-bond by its bond model) and the kernel
          closure (dividing the positive non-incident bond products and applying
          \leanid{TNLean.PEPS.bondLocal_iff_coeffTransfer} in both directions). Both are now
          landed; the per-edge gauge is delivered through the coherent-frame interface
          \leanid{TNLean.PEPS.exists_regionEdgeGauge_of_coherentFrames}, which discharges the
          three hypotheses \leanid{hbondAB}, \leanid{hbondBA}, and \leanid{hmul} of the
          abstract \leanid{TNLean.PEPS.SharedNormalEdgeBlockingData.exists_regionEdgeGauge}
          from the two coherent frames and the single-crossing hypothesis, returning the same
          gauge conclusion. The final assembly is recorded below.

          The descent is not a single-edge assembly, because of a granularity gap recorded
          here. The coarse red-to-blue super-bond carries the \emph{whole} bundle of
          red-to-blue crossing edges, not a single boundary edge: its bond model is an
          equivalence with the red-to-blue crossing configurations
          (\leanid{TNLean.PEPS.CrossingConfig}), and the coarse edge-inserted coefficient
          inserts a matrix on that whole bundle. The collapse with such a matrix therefore
          descends to a whole-bundle inserted coefficient
          (\leanid{TNLean.PEPS.redBundleInsertedCoeff}): a doubled red-boundary sum coupling
          the two configurations through their red-to-blue crossing labels
          (\leanid{TNLean.PEPS.redBoundaryRBCrossing}) while the red-to-complement crossings
          are contracted diagonally (\leanid{TNLean.PEPS.SameAwayFromRBBundle}). This
          coefficient is additive and homogeneous in the inserted matrix
          (\leanid{TNLean.PEPS.redBundleInsertedCoeff_add},
          \leanid{TNLean.PEPS.redBundleInsertedCoeff_smul}). The relaxed crossing agreement
          of a triple under the inserted matrix --- agreeing on the red-to-complement and
          blue-to-complement crossings, with the red-to-blue crossings left free for the
          matrix to couple --- is recorded by
          \leanid{TNLean.PEPS.CrossTripleAgreesAwayRB}.

          The open-bond reindexing of the descent is now landed. The coarse edge-inserted
          coefficient at the red-to-blue super-bond expands, through the open-bond form of the
          matrix insertion, as a sum over a coarse virtual configuration and a free right
          super-bond index of the matrix coupling their two super-bond values times the three
          region weights, the blue weight read through the override of the coarse
          configuration on the super-bond (\leanid{TNLean.PEPS.edgeInsertedCoeff_eq_pairSum},
          \leanid{TNLean.PEPS.edgeInsertedCoeff_coarseTensor_eq_threeRegionSum}). Carrying the
          matrix through the coherent bond models reindexes this onto the relaxed crossing
          triples: the matrix becomes the bond-model-conjugated matrix
          (\leanid{TNLean.PEPS.bondModelMatrix}) at the two configurations' red-to-blue
          crossing labels, summed over triples agreeing away from the red-to-blue crossings
          (\leanid{TNLean.PEPS.mCoupledThreeRegionSum_eq_relaxedTripleSum}, built on the
          constraint-set collapse \leanid{TNLean.PEPS.pairConfig_constraint_set} and the
          per-pair triple expansion \leanid{TNLean.PEPS.perPair_threeRegionProduct_eq}). This
          is the matrix-carrying analogue of the closed-state reindexing
          \leanid{TNLean.PEPS.threeRegionSum_eq_agreeingTripleSum}.

          The bundle-to-single-edge bridge is also landed. The single-edge predicate
          \leanid{TNLean.PEPS.IsBondLocalTransferKernel} that the per-edge gauge consumes, and
          the equivalence \leanid{TNLean.PEPS.bondLocal_iff_coeffTransfer} that reduces it to a
          single-edge coefficient transfer, are stated for one boundary edge $f$. When the red
          region crosses to the blue region across the single distinguished edge $e$ --- the
          source's geometry, where the red sites surround one endpoint and the blue sites the
          other --- the whole red-to-blue crossing bundle is one virtual leg on $e$
          (\leanid{TNLean.PEPS.singletonCrossingEquiv}), the bundle agreement is the
          single-edge agreement \leanid{TNLean.PEPS.SameAwayFromBond} on $e$
          (\leanid{TNLean.PEPS.sameAwayFromRBBundle_iff_sameAwayFromBond}), and the
          whole-bundle red inserted coefficient of the carried bond matrix is the single-edge
          \leanid{TNLean.PEPS.regionInsertedCoeff} of that matrix on $e$
          (\leanid{TNLean.PEPS.redBundleInsertedCoeff_singleton}). The single-crossing
          hypothesis is the explicit assumption that the red-to-blue crossings are exactly the
          singleton $\{e\}$, which the source's coordinate blocking discharges
          (\path{TNLean/PEPS/NormalBlocking.lean:251--264,367--379}).

          With these two landed pieces, the descent on both tensors (sharing the non-incident
          bond products under matched bond dimensions, by
          \leanid{TNLean.PEPS.coarseTensor_sameState_of_sameState}), the coarse coefficient
          transfer \leanid{TNLean.PEPS.coarse_exists_edgeInsertedCoeff_eq}, the surjectivity
          \leanid{TNLean.PEPS.coarseProj_surjective}, and the positivity of the non-incident
          bond products would give the single-edge coefficient transfer, hence
          \leanid{TNLean.PEPS.IsBondLocalTransferKernel} in both directions and the per-edge
          gauge.

          The \emph{relaxed-triple merge collapse} is now landed
          (\leanid{TNLean.PEPS.relaxedTripleSum_collapse}), and with it the whole
          inserted-coefficient descent (\leanid{TNLean.PEPS.edgeInsertedCoeff_coarseTensor_descent}):
          the coarse edge-inserted coefficient at the red-to-blue super-bond equals a constant
          times the whole-bundle red inserted coefficient
          \leanid{TNLean.PEPS.redBundleInsertedCoeff} of the bond-model-conjugated matrix, with
          the host physical leg the fused blue/complement leg. The collapse is the
          matrix-carrying analogue of \leanid{TNLean.PEPS.agreeingTripleSum_collapse}, a
          \emph{two-sided} merge: the red configuration $\zeta_r$ merges into the red-boundary
          index $\mu$ (collapsing to \leanid{TNLean.PEPS.regionBlockedWeight} of $\mu$ with no
          fiber multiplicity, the blocked-region weight already summing the red product over
          every configuration carrying the red boundary label), while the blue and complement
          configurations $(\zeta_b,\zeta_c)$, agreeing on the blue-to-complement crossings,
          merge into the host configuration \leanid{TNLean.PEPS.hostMerge} carrying the host
          boundary index $\nu$ over $\mathrm{univ}\setminus\mathrm{red}$. The
          bond-model-conjugated matrix couples $\mu$ and $\nu$ through their red-to-blue
          crossing labels (\leanid{TNLean.PEPS.redBoundaryRBCrossing}), and the
          red-to-complement crossing agreement is the bundle agreement
          \leanid{TNLean.PEPS.SameAwayFromRBBundle}.

          The merged summand $\mathrm{matrix}(\mu_\mathrm{rb},\nu_\mathrm{rb})\cdot
          \mathrm{redProd}\,\zeta_r\cdot\mathrm{hostProd}(\mathrm{hostMerge})$ is
          \leanid{TNLean.PEPS.relaxedTriple_summand_eq}: the complement and blue vertex products
          read the host merge unchanged (\leanid{TNLean.PEPS.complProd_hostMerge},
          \leanid{TNLean.PEPS.blueProd_hostMerge}), so the host vertex product of the merge,
          read with the fused blue/complement physical leg, is the complement product against
          the blue product (\leanid{TNLean.PEPS.hostProd_hostMerge_eq}); the two boundary labels
          and the matrix coupling are identified by
          \leanid{TNLean.PEPS.sameAwayFromRBBundle_hostMerge},
          \leanid{TNLean.PEPS.redBoundaryRBCrossing_hostMerge}, and
          \leanid{TNLean.PEPS.crossingLabel_eq_redBoundaryRBCrossing}. The host side collapses
          through the host-merge fiber count
          \leanid{TNLean.PEPS.hostMergeFiber_card}: the relaxed pairs agreeing on the
          blue-to-complement crossings whose merge is a fixed configuration biject with the free
          virtual indices, the fiber multiplicity being
          \leanid{TNLean.PEPS.hostMergeFiberProd} (the complement non-incident bond product times
          the free blue bond product), a function of the bond dimensions alone and hence shared
          across the two tensors under matched bond dimensions. The fiberwise collapse
          \leanid{TNLean.PEPS.hostMerge_fiberwise_collapse} and the relaxed agreement split
          \leanid{TNLean.PEPS.crossTripleAgreesAwayRB_iff} wire the two-stage grouping against the
          configuration expansion
          \leanid{TNLean.PEPS.redBundleInsertedCoeff_bondModelMatrix_eq_configSum} of the target
          coefficient. The descent uses no single-vertex spanning.

          The final assembly is now landed. Under the single-crossing hypothesis the
          whole inserted-coefficient descent reads, through the bond-model and the
          single-leg crossing equivalence, as a constant times the single-edge
          region-inserted coefficient on $e$
          (\leanid{TNLean.PEPS.edgeInsertedCoeff_coarseTensor_descent_single}), the
          constant the shared positive host-merge fiber product
          (\leanid{TNLean.PEPS.hostMergeFiberProd_pos},
          \leanid{TNLean.PEPS.hostMergeFiberProd_eq}). Cancelling that constant after the
          descent on both coherent frames, and feeding the coarse coefficient transfer and
          the surjectivity of the descent's physical legs
          (\leanid{TNLean.PEPS.exists_descent_legs}), gives the single-edge coefficient
          transfer in both directions
          (\leanid{TNLean.PEPS.exists_regionInsertedCoeff_eq_sharedRegion},
          \leanid{TNLean.PEPS.exists_regionInsertedCoeff_eq_sharedRegion_symm}), hence the
          two bond-local transfer kernels
          (\leanid{TNLean.PEPS.isBondLocalTransferKernel_of_coherentFrames}). The forward
          multiplicativity is the concrete single-edge transfer's multiplicativity, the
          coarse edge transfer matrix conjugated by the two bond models
          (\leanid{TNLean.PEPS.regionEdgeTransfer},
          \leanid{TNLean.PEPS.regionEdgeTransfer_mul}), carried through
          \leanid{TNLean.PEPS.regionInsertedCoeff_injective}
          (\leanid{TNLean.PEPS.coeffTransferMap_eq_regionEdgeTransfer}). With the two
          kernels and the forward multiplicativity, the per-edge gauge follows
          (\leanid{TNLean.PEPS.exists_regionEdgeGauge_of_coherentFrames}): the forward
          per-edge matrix transfer on $e$ is conjugation by an invertible gauge matrix and
          the bond dimensions on $e$ coincide. The whole assembly uses no single-vertex
          injectivity of the original tensor; the only injectivity input is the
          blocked-region injectivity of the three regions, carried by the coherent coarse
          blocking frame. This closes, modulo review, the block-level step $V=W$ and the
          forward multiplicativity --- the last residual content of this obligation ---
          for two coherent frames sharing the three regions, the bond dimensions, and the
          single distinguished crossing edge.

          The coherent-frame interface is now fed from a one-edge blocking datum. The
          coherent coarse blocking frame of a one-edge blocking datum over the concrete
          region-injectivity predicate of a tensor is assembled
          (\leanid{TNLean.PEPS.coherentFrameOfBlockingData}), and two such data over the
          two tensors, sharing the three regions and the single distinguished crossing
          edge, deliver the per-edge gauge: the bond dimensions on $e$ coincide and the
          forward per-edge matrix transfer on $e$ is conjugation by an invertible matrix
          (\leanid{TNLean.PEPS.exists_regionEdgeGauge_of_blockingData}), with no
          single-vertex injectivity.

          The single-crossing hypothesis is discharged from the coordinate geometry of
          the square lattice. For the translated horizontal edge blocking the red block
          occupies the columns $x_0,x_0{+}1$ and the rows $y_0{+}2,\dots,y_0{+}4$ and the
          blue block the columns $x_0{+}2,\dots,x_0{+}4$ and the rows $y_0{+}1,y_0{+}2$,
          two adjacent rectangles meeting at the single nearest-neighbour pair
          $(x_0{+}1,y_0{+}2)$--$(x_0{+}2,y_0{+}2)$, the distinguished edge; the rotated
          picture holds for the vertical blocking. Hence an edge crosses between the red
          and blue blocks if and only if it is the distinguished edge
          (\leanid{TNLean.PEPS.isCrossingEdge_normalSquareHorizontalTranslatedEdge},
          \leanid{TNLean.PEPS.isCrossingEdge_normalSquareVerticalTranslatedEdge}), with
          interior-data wrappers tying the single-crossing hypothesis to the cover-free
          interior blocking data. The region projections of the interior data are exposed
          syntactically as the coordinate blocks
          (\leanid{TNLean.PEPS.normalSquareHorizontalTranslatedEdge_blockingDatum_interior_red}
          and companions).

          What remains for the per-edge gauge family is a decidable-equality coherence
          bridge. The square-lattice geometry layer builds its blocking data with the
          canonical product decidable equality on $\mathrm{Fin}\,w\times\mathrm{Fin}\,h$,
          while the gauge interface lives over the lattice's linear order and uses the
          order's decidable equality. The two instances are equal as a subsingleton, so
          they carry the same mathematics, but they are not definitionally equal, and the
          two layers meet exactly when an interior datum is fed to the gauge interface.
          Feeding the gauge per interior edge therefore requires aligning the two
          decidable-equality instances --- by transporting the datum across the
          subsingleton equality, or by a uniform instance choice across the geometry and
          gauge files. This is a finite engineering step with no further mathematical
          content: the gauge of a one-edge datum
          (\leanid{TNLean.PEPS.exists_regionEdgeGauge_of_blockingData}) and the
          single-crossing discharge are already in place. The orientation-uniform
          reduction (obligation~6) and the boundary geometry (obligation~4) remain the
          open mathematical pieces above this bridge.
    \item In the translationally invariant case, show that the edge gauges obtained
          after blocking reduce to one horizontal gauge $X$ and one vertical gauge
          $Y$, up to scalar.

          The carrier of this reduction is now formalized. A square-lattice
          bond-dimension function is orientation uniform when every horizontal edge
          has one dimension and every vertical edge has another; given that, a single
          horizontal matrix and a single vertical matrix determine a gauge family by
          transport across the bond-dimension equalities, and a gauge family of this
          shape is called orientation uniform. The absorption step incorporates such
          an orientation-uniform family into the second tensor, preserving the
          orientation-uniform bond dimensions and vertex injectivity; the
          post-absorption edge-insertion equality is taken as the conditional kernel
          input, in the shape produced unconditionally in the vertex-injective case.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.SquareLatticeUniformBondDim},
          \leanid{TNLean.PEPS.orientationUniformGauge},
          \leanid{TNLean.PEPS.IsOrientationUniformGaugeFamily},
          \leanid{TNLean.PEPS.exists_orientationUniformGaugeFamily},
          \leanid{TNLean.PEPS.tiAbsorb},
          \leanid{TNLean.PEPS.isVertexInjective_tiAbsorb},
          \leanid{TNLean.PEPS.tiNormalGaugeAbsorption}, and
          \leanid{TNLean.PEPS.exists_postAbsorption_of_vertexInjective}.}

          The uniqueness-up-to-scalar step is now formalized. The conjugating
          matrix realizing a forward per-edge transfer is determined up to a
          multiplicative constant: if two invertible matrices induce the same
          conjugation on every bond matrix then their quotient commutes with the
          whole bond matrix algebra, hence lies in its center, the scalars, so the
          two differ by a nonzero constant. Reading this through the bond-dimension
          reindexing gives the same statement for the per-edge transfer map of an
          edge: two gauges with the same transfer map differ by a constant. This is
          the source's statement that the horizontal and vertical matrices are unique
          up to a multiplicative constant, at the level of a single edge.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.gl_conj_unique_scalar} and
          \leanid{TNLean.PEPS.edgeGauge_unique_scalar}.}

          Because each per-edge gauge is fixed only up to a constant, the reduction
          to one horizontal and one vertical matrix holds only up to a per-edge
          scalar: a gauge family is orientation uniform up to scalars when each edge
          gauge is the transported reference matrix rescaled by a nonzero per-edge
          constant. The assembly across the two orientation classes is formalized:
          collecting, on each class, the per-edge ``agree with the reference up to a
          constant'' facts produces one orientation-uniform-up-to-scalar family, with
          the exact orientation-uniform family as the all-constants-one
          specialization.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.IsOrientationUniformGaugeFamilyModScalar},
          \leanid{TNLean.PEPS.isOrientationUniformGaugeFamilyModScalar_of_isOrientationUniform},
          and
          \leanid{TNLean.PEPS.isOrientationUniformGaugeFamilyModScalar_of_classAgreement}.}

          The per-edge constants are absorbed at the absorption step when they are
          vertex balanced. A per-edge scalar rescaling whose oriented product around
          every vertex is one is a balanced edge-scalar reweighting, and such a
          reweighting leaves the absorbed tensor unchanged. Hence the
          scalar-carrying orientation-uniform gauge and the underlying exact
          orientation-uniform gauge incorporate into the same modified second tensor,
          so the conditional square-lattice theorem may consume an exact
          orientation-uniform gauge even though the blocking delivers one only up to
          balanced constants.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.gaugeEquivModEdgeScalars_of_matrixScalar} and
          \leanid{TNLean.PEPS.absorbEdgeGauges_eq_of_matrixScalar}.}

          The periodic lattice on which this transport lives is now formalized,
          together with the covariance of the blocking machinery under it. The
          discrete torus carries cyclic nearest-neighbour edges on which the
          coordinate translations are graph automorphisms, a tensor is
          translation invariant when fixed by transport along every translation, and
          such a tensor has orientation-uniform bond dimensions on each
          class.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.torusGraph},
          \leanid{TNLean.PEPS.translate},
          \leanid{TNLean.PEPS.IsTorusTranslationInvariant},
          \leanid{TNLean.PEPS.TorusUniformBondDim}, and
          \leanid{TNLean.PEPS.torusUniformBondDim_of_translationInvariant}.}
          The blocking data themselves transport: the blocked-region weight of a
          tensor transported along a graph isomorphism is the original weight
          reindexed by the edge action and the vertex bijection, so blocked-region
          injectivity transports, and a one-edge blocking datum transports to a datum
          for the transported tensor at the image edge with the single red-to-blue
          crossing carried across. Specialized to a translation-invariant tensor,
          this carries one reference blocking datum at the reference edge of an
          orientation class to a datum at every translate of that edge, for the same
          tensor.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.regionBlockedWeight_transport},
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_transport},
          \leanid{TNLean.PEPS.transportBlockingDataAlong},
          \leanid{TNLean.PEPS.isCrossingEdge_transportBlockingDataAlong},
          \leanid{TNLean.PEPS.translateBlockingData}, and
          \leanid{TNLean.PEPS.isCrossingEdge_translateBlockingData}.}
          The reference blocking datum at each class's reference edge is now constructed
          concretely.  For a vertex-injective tensor with positive bond dimensions every
          finite vertex region is injective, so the singleton endpoint blocks and their
          complement form a one-edge blocking datum with no rectangular geometry: the
          left endpoint is the red block, the right endpoint the blue block, the rest of
          the torus the complement, and the only red-to-blue crossing is the edge itself.
          This supplies the reference datum at the reference horizontal and vertical edges
          directly, with no wraparound-avoiding rectangle bounds, since singleton blocks
          need no room to be injective once the tensor is vertex injective.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.singletonEdgeBlockingData},
          \leanid{TNLean.PEPS.isCrossingEdge_singletonEdgeBlockingData},
          \leanid{TNLean.PEPS.torusHorizontalReferenceBlockingDatum}, and
          \leanid{TNLean.PEPS.torusVerticalReferenceBlockingDatum}.}
          \textbf{Scope restriction (NormalTorusRectangleInjectivityHypotheses):}
          this singleton datum is a vertex-injective specialization of
          the normality hypothesis in Theorem~3 of arXiv:1804.04964, lines
          1449--1452.  The source-faithful
          torus reference datum uses the rectangular injectivity hypotheses
          directly: the reference red and blue blocks are the two source
          rectangles adjacent to the distinguished edge, and the complement is
          the union of the surrounding bands and fillers.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.torusHorizontalRectangleBlockingDatum},
          \leanid{TNLean.PEPS.torusVerticalRectangleBlockingDatum},
          \leanid{TNLean.PEPS.isCrossingEdge_torusHorizontalRectangleBlockingDatum},
          and
          \leanid{TNLean.PEPS.isCrossingEdge_torusVerticalRectangleBlockingDatum}.}
          The per-edge gauge interface applies verbatim on the torus (it is stated
          over a general finite ordered vertex set), and the orientation-uniform
          reduction is ported to the torus, so the assembly above a class-agreement
          input is in place there too.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.exists_regionEdgeGauge_torus},
          \leanid{TNLean.PEPS.torusOrientationUniformGauge},
          \leanid{TNLean.PEPS.IsTorusOrientationUniformGaugeFamilyModScalar},
          \leanid{TNLean.PEPS.isTorusOrientationUniformGaugeFamilyModScalar_of_classAgreement},
          and
          \leanid{TNLean.PEPS.isTorusOrientationUniformGaugeFamilyModScalar_of_translationInvariant}.}

          The covariance of the region-inserted coefficient is now established, and
          with it the determinacy that closes the class-agreement input above the
          per-edge gauge of obligation~5. The region-inserted coefficient of a tensor
          transported along a graph isomorphism, over the image region with the
          inserted matrix carried to the image bond and the boundary edge to its image,
          equals the original coefficient over the original region; the boundary
          configurations reindex by the edge action, the matrix coupling and the
          away-from-the-bond agreement are preserved, and the region and complement
          weights both transport. No endpoint orientation enters, since the coefficient
          inserts the matrix between the two single boundary values on the edge.
          \footnote{Formal declarations:
          \leanid{TNLean.PEPS.regionComplementWeight_transport},
          \leanid{TNLean.PEPS.sameAwayFromBond_regionBoundaryConfigMap}, and
          \leanid{TNLean.PEPS.regionInsertedCoeff_transport}.}
          Specialized to a translation-invariant pair, this carries the coefficient
          identity at a reference edge to the same identity at every translate of that
          edge, for the same two tensors, since the tensors are fixed by transport.
          \footnote{Formal declarations:
          \leanid{TNLean.PEPS.bondDim_boundaryEdgeMap_translate},
          \leanid{TNLean.PEPS.regionInsertedCoeff_congrTensor}, and
          \leanid{TNLean.PEPS.regionInsertedCoeff_translate_coeffIdentity}.}

          The non-constructive transfer map of the gauge interface is determined by the
          coefficient identity it satisfies: with the second tensor's region and
          complement injective and positive bond dimensions, two matrix maps realizing
          the same coefficient identity coincide. Hence the per-edge transfer map at a
          translate equals the reference transfer map carried across the translation,
          two independently chosen per-edge gauges realize the same conjugation map, and
          the per-edge uniqueness up to scalar supplies the constants; gathering the two
          orientation classes yields the orientation-uniform-up-to-scalar family.
          \footnote{Formal declarations:
          \leanid{TNLean.PEPS.regionInsertedCoeff_transferMap_unique},
          \leanid{TNLean.PEPS.regionTransferMap_eq_of_coeffIdentities}, and
          \leanid{TNLean.PEPS.isTorusOrientationUniformGaugeFamilyModScalar_of_conjCovariance}.}
          The algebraic reduction from matched coefficient identities to the conjugation
          covariance is now in place. The per-edge gauge at a torus edge carries its
          defining coefficient identity, so the gauge conjugation itself satisfies the
          identity. Reindexing across a bond-dimension equality commutes with
          conjugation, identifying the transported reference matrix with the conjugation
          by the reindexed reference. Two gauges realizing the same coefficient identity
          by conjugation induce the same conjugation map on every bond matrix, by the
          transfer-map determinacy. These supply, edge by edge, the conjugation
          coincidence the orientation-uniform selection consumes.\footnote{Formal
          declarations: \leanid{TNLean.PEPS.exists_regionEdgeGauge_torus_coeff},
          \leanid{TNLean.PEPS.reindexAlgEquiv_gaugeConj},
          \leanid{TNLean.PEPS.gaugeConj_eq_of_coeffIdentities}, and
          \leanid{TNLean.PEPS.gaugeConj_eq_of_coeffIdentities_torus}.}
          What remains is the geometric production of the two matched coefficient
          identities at every edge: reading the per-edge gauge off the reference blocking
          datum transported along each class translation, and matching its region and
          boundary edge with those of the transported reference identity. The transported
          blocking datum reads its three blocks from the translation, which may reorder
          the edge endpoints and so place the reference red region in the blue slot; the
          orientation-class matching must account for this branch. This construction
          still depends on the open per-edge gauge of obligation~5 for the transfer maps
          themselves.

          The endpoint-reordering branch is now handled at the level of the gauge-interface
          inputs. The transported datum's complement block is always the image of the
          reference complement, and its red block is one of the two image blocks; whichever
          image block plays the red role, that block is injective and the union of the
          remaining two covers the host region of the red block and is injective under
          union closure. So both injectivity inputs the per-edge gauge consumes are
          available on both branches, with no case split, and the region-inserted
          coefficient is endpoint-symmetric. What is left is to route the orientation-class
          matching through the image of the reference red region rather than the
          possibly-swapped red slot, which the region-polymorphic coefficient covariance
          already supplies.\footnote{Formal declarations:
          \leanid{TNLean.PEPS.transportBlockingDataAlong_complement},
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_transportBlockingDataAlong_red},
          and
          \leanid{TNLean.PEPS.regionBlockedTensorInjective_host_transportBlockingDataAlong}.}

          The reference blocking datum is now produced from the source's own
          rectangular-injectivity hypotheses on the torus, with no single-vertex
          injectivity. Contiguous torus rectangles read the half-open coordinate
          intervals through the value embedding, so wraparound-avoiding offsets recover
          the contiguous block and full-coordinate bands cover the whole range; the
          $2\times3$ and $3\times2$ shape predicates, the native rectangular-injectivity
          hypotheses, and the sliding-window tiling are the torus ports of the open-lattice
          development. The horizontal and vertical reference edges carry the red (removed
          source rectangle), blue (the other removed source rectangle), and complementary
          (the band-and-filler cover) blocks; all three are injective by rectangular
          injectivity and union closure, they partition the torus, and the cyclic-step
          value lemmas read a wraparound-free step as the ordinary step, so the single
          red-to-blue crossing is the reference edge. This is the source-faithful
          replacement for the vertex-injective singleton scaffolding.\footnote{Formal
          declarations:
          \leanid{TNLean.PEPS.NormalTorusRectangleInjectivityHypotheses},
          \leanid{TNLean.PEPS.NormalTorusRectangleInjectivityHypotheses.horizontalEdgeComplement_injective},
          \leanid{TNLean.PEPS.torusHorizontalRectangleBlockingDatum},
          \leanid{TNLean.PEPS.torusVerticalRectangleBlockingDatum},
          \leanid{TNLean.PEPS.isCrossingEdge_torusHorizontalEdge}, and
          \leanid{TNLean.PEPS.isCrossingEdge_torusVerticalEdge}.}

          Feeding this faithful datum to the per-edge gauge engine yields, at the reference
          edge, the per-edge conjugation gauge realizing the region-insertion coefficient
          identity, with the second tensor's red and host injectivities supplied by its own
          rectangle hypotheses and union closure. The conjugation-covariance input the
          orientation-uniform selection consumes is packaged as a per-edge coefficient-identity
          witness — a region with single boundary edge $e$, the second tensor's
          region and host injectivity, and the two coefficient identities the per-edge gauge
          and the transported reference gauge realize. Each witness yields the conjugation
          coincidence by the transfer-map determinacy, and gathering the two orientation
          classes assembles the orientation-uniform-up-to-scalar family. The reference edge
          inhabits this witness interface directly.

          The translation production is now carried out. The reference coefficient identity,
          realized by conjugation with the reference gauge, transports along each class
          translation to the same identity at every edge of the class, realized by conjugation
          with the reference gauge carried across the bond-dimension reindexing. The translated
          region is the image of the reference region; the second tensor's image region and its
          host are injective there by transport of the reference injectivities, since the
          second tensor is fixed by the translation; and an arbitrary region or complement
          configuration on the image region is the transport of a unique configuration on the
          reference region. Composing the two reindexings of the transported gauge collapses to
          the single transported orientation matrix the selection expects, so the per-edge
          witness against that matrix holds at every edge. Gathering the horizontal and vertical
          witness families through the orientation-uniform selection produces, from the source's
          rectangle hypotheses on a translation-invariant pair, the
          orientation-uniform-up-to-scalar family unconditionally: the per-edge gauge family is
          the single reference matrix per class transported to every edge, the source's
          ``$X$ and $Y$, the same matrix on all horizontal (vertical) edges'', with the per-edge
          scalars all one.\footnote{Formal declarations:
          \leanid{TNLean.PEPS.regionInsertedCoeff_translate_coeffIdentity_conj},
          \leanid{TNLean.PEPS.edgeCoeffIdentityWitness_translate},
          \leanid{TNLean.PEPS.edgeCoeffIdentityWitness_translateUniform},
          \leanid{TNLean.PEPS.exists_edgeCoeffIdentityWitness_horizontalFamily},
          \leanid{TNLean.PEPS.exists_edgeCoeffIdentityWitness_verticalFamily}, and
          \leanid{TNLean.PEPS.isTorusOrientationUniformGaugeFamilyModScalar_torus_rectangle}.}

          One strengthening remains. The family produced above takes the per-edge gauge to be the
          transported reference matrix itself, so the witness's two coefficient identities are the
          same single identity and the per-edge scalars are all one. The source's stronger claim is
          that \emph{any} per-edge gauge independently produced by the edge-blocking engine at each
          edge agrees with the transported reference up to a scalar. Reading that independent engine
          gauge over the image of the reference region requires a blocking datum at the translated
          edge whose red block is the image of the reference red block; but the lexicographic edge
          convention reorders the endpoints under a wraparound translation, so the transported
          datum's red block is the image of the reference \emph{blue} block on that branch, and the
          reference coefficient identity — tied to the reference red region — does not transport to
          it. Closing the independent-gauge strengthening therefore needs a branch-free reference
          identity (one stated over the reference region's host orientation as well), which the
          present reference engine does not supply. The unconditional family above does not depend
          on this strengthening.

          The final region comparison setup is also formalized: a region and its set
          complement are wrapped as the two blocks consumed by the two-injective
          comparison, and from matched region-inserted coefficients of the two
          tensors the region blocks are scalar proportional, the region analogue of
          the one-vertex-versus-complement comparison. For vertex-injective tensors
          the four block-injectivity hypotheses are discharged uniformly, leaving the
          region-inserted-coefficient equality as the only conditional input.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.regionComplement_comparison} and
          \leanid{TNLean.PEPS.regionComplement_comparison_of_vertexInjective}.}

          The region-inserted-coefficient equality consumed by the region comparison is the
          \emph{absorbed} form: \(A\) inserting \(M\) on \(f\) matches \(\widetilde B=
          \operatorname{applyGauge} B\,X\) inserting the same \(M\) on \(f\). The per-edge gauge
          delivered by the blocking is, however, the \emph{conjugation} form: \(A\) inserting
          \(M\) matches \(B\) inserting \(X^{\mathsf T} M (X^{-1})^{\mathsf T}\) on \(f\). The
          bridge between the two is the region-granularity port of the open-edge gauge
          cancellation \leanid{TNLean.PEPS.edgeInsertedCoeff_applyGauge}: inserting \(M\) on the
          gauged tensor cancels every interior gauge of the region and of its complement and
          conjugates \(M\) by the gauge on the single boundary edge \(f\). This bridge is now
          formalized as \leanid{TNLean.PEPS.regionInsertedCoeff_applyGauge}.

          The proof factors through a region-to-edge identity. First, the
          double-global-configuration form of the region-inserted coefficient
          (\leanid{TNLean.PEPS.regionInsertedCoeff_eq_doubleSum}) collects its region and complement
          vertex weights into one global vertex product over all vertices, the region side reading
          the first configuration of the agreeing pair and the complement side the second
          (\leanid{TNLean.PEPS.pairOuter}), with the physical legs assembled; this holds both for the
          gauged tensor (\leanid{TNLean.PEPS.regionComplProd_gauge_eq}) and gauge-free
          (\leanid{TNLean.PEPS.regionComplProd_eq}). The outer reading
          \leanid{TNLean.PEPS.pairOuter} of an agreeing pair is consistent off \(f\)
          (\leanid{TNLean.PEPS.pairOuter_isConsistentOff}): only the single boundary edge \(f\) keeps
          two distinct endpoint indices. The agreeing pairs over a fixed consistent-off-\(f\)
          configuration are parameterised by the free non-boundary legs, so each fiber has the
          interior bond multiplicity (\leanid{TNLean.PEPS.pairOuterFiber_card}), and summing a
          function of \leanid{TNLean.PEPS.pairOuter} over the agreeing pairs is that multiplicity
          times the sum over consistent-off-\(f\) configurations
          (\leanid{TNLean.PEPS.sum_pairOuter_fiber_collapse}). Identifying the latter with the
          edge-inserted coefficient at \(f\) of the assembled physical configuration gives the
          region-to-edge identity \leanid{TNLean.PEPS.regionInsertedCoeff_eq_smul_edgeInsertedCoeff}:
          the region/complement contraction is the single-bond cut at \(f\) up to the
          gauge-invariant interior bond multiplicity, with the inserted matrix oriented to the edge
          convention by \leanid{TNLean.PEPS.regionEdgeOrient} (identity when \(f\)'s left endpoint
          lies in \(R\), transpose otherwise). The edge gauge cancellation
          \leanid{TNLean.PEPS.edgeInsertedCoeff_applyGauge} then transports across the identity,
          giving the bridge with the conjugation by \(X_f^{\mathsf T}\) on the orientation-corrected
          matrix.

          The conjugation-to-absorbed conversion is now formalized, both at a single comparison
          region and edge-level (so region-independent). At a region \(R\) with boundary edge
          \(f\), the engine's conjugation-form coefficient identity \(A\) inserting \(M\) equals
          \(B\) inserting \(Z\,M\,Z^{-1}\) becomes the absorbed equality \(A\) inserting \(M\)
          equals \(\operatorname{applyGauge} B\,X\) inserting the same \(M\), with the absorbing
          matrix at \(f\) the orientation-adapted transpose of the engine gauge
          (\leanid{TNLean.PEPS.absorbedBoundaryGauge}: \(Z^{\mathsf T}\) when \(f\)'s left
          endpoint lies in \(R\), \(Z^{-1}\) otherwise, so the bridge's
          \(X^{\mathsf T}\!\cdot\!\operatorname{orient}\!\cdot\!(X^{-1})^{\mathsf T}\) conjugation
          reproduces the engine's \(Z\,\cdot\,Z^{-1}\) conjugation). The matrix reconciliation is
          \leanid{TNLean.PEPS.regionEdgeOrient_absorbedBoundaryGauge}; the per-region conversion is
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_applyGauge_of_conjIdentity}. Cancelling the
          shared positive interior multiplicity in the region-to-edge identity reduces the region
          absorbed equality to the edge-level absorbed equality at \(f\) over the bare edge
          (\leanid{TNLean.PEPS.edgeInsertedCoeff_eq_applyGauge_of_region}), and the converse
          multiplies it back up to the region absorbed equality at \emph{every} region with \(f\)
          on its boundary (\leanid{TNLean.PEPS.regionInsertedCoeff_eq_applyGauge_of_edge}). The two
          directions are the route note's region-independence: the absorbed plain equality over
          any comparison region follows from the single edge-level identity at the boundary edge,
          which is what the region comparison \leanid{TNLean.PEPS.regionComplement_comparison}
          consumes.

          On the torus the per-edge coefficient identity is produced at a single boundary edge of
          each \emph{red blocking region} of the orientation-uniform family
          (\leanid{TNLean.PEPS.exists_regionEdgeGauge_torus_coeff},
          \leanid{TNLean.PEPS.isTorusOrientationUniformGaugeFamilyModScalar_torus_rectangle}). The
          per-vertex relation \(A_v=\lambda\cdot\operatorname{gaugeVertex} B\,X\,v\) that the
          torus Fundamental Theorem
          (\leanid{TNLean.PEPS.fundamentalTheorem_normalTorusPEPS}) takes as the hypothesis
          \path{hPV} still needs two pieces above the region-independence just recorded. First, the
          absorbing matrix \leanid{TNLean.PEPS.absorbedBoundaryGauge} is the orientation-adapted
          transpose of each \emph{per-edge} engine gauge \(Z_e\), so the absorbed equality holds
          against a per-edge absorbing family rather than against the single
          orientation-uniform \(X=\operatorname{torusOrientationUniformGauge}\) that
          \leanid{TNLean.PEPS.isTorusOrientationUniformGaugeFamilyModScalar_torus_rectangle}
          returns and that \path{hPV} is stated against; reconciling the per-edge absorbing family
          with that orientation-uniform \(X\) (the transpose/orientation bookkeeping across the two
          orientation classes) is the first open piece. Second, feeding the absorbed equality to
          \leanid{TNLean.PEPS.regionComplement_comparison} at the source comparison regions \(R\)
          and \(S\) (which differ in one site, both blocked-tensor injective with injective
          complements) gives \(A_R\propto\widetilde B_R\) and \(A_S\propto\widetilde B_S\);
          quotienting the two region proportionalities to isolate the single differing vertex is
          the second open piece. The block-granularity one-site quotient that performs this
          isolation without single-vertex injectivity --- the region analogue of
          \leanid{TNLean.PEPS.perVertexScalar}, whose present form rests on
          \leanid{TNLean.PEPS.one_vertex_complement_comparison} and edge-level vertex injectivity
          --- needs a factorization of the blocked-region weight of \(\operatorname{withSite} v\)
          through the blocked-region weight of \(\operatorname{withoutSite} v\) and the single
          vertex tensor at \(v\) (the one-vertex analogue of
          \leanid{TNLean.PEPS.regionBlockedWeight_eq_localTensorMap}).

          The foundational layer of this quotient is now landed. Writing the differing regions as
          \(S=\operatorname{insert} v\,R\) with \(v\notin R\), the vertex product over the sites of
          \(S\) factors as the inserted-site tensor times the vertex product over \(R\)
          (\leanid{TNLean.PEPS.prod_region_insert_split}), and this lifts to the constrained
          blocked-weight sum (\leanid{TNLean.PEPS.regionBlockedWeight_insert_eq_sum_split}): the
          blocked weight of \(S\) is the \(\mu\)-constrained sum of \(A.\mathrm{component}\,v\)
          against the residual vertex product over \(R\). The boundary-configuration half is also
          landed: an \(R\)-boundary edge not incident to \(v\) is an \(S\)-boundary edge
          (\leanid{TNLean.PEPS.isRegionBoundaryEdge_insert_of_not_incident}), one incident to \(v\)
          becomes internal to \(S\), and the boundary label of \(R\) is folded from the \(S\)-boundary
          label away from \(v\) and the local virtual configuration at \(v\) on the \(v\)-incident
          edges (\leanid{TNLean.PEPS.boundaryLabelOfInsert},
          \leanid{TNLean.PEPS.regionBoundaryLabel_eq_boundaryLabelOfInsert}).

          The residual-multiplicity factorization, the inserted-site analogue of
          \leanid{TNLean.PEPS.regionInsertedCoeff_eq_smul_edgeInsertedCoeff}, is now landed. The
          residual sum at a fixed local configuration \(\eta\) at \(v\)
          (\leanid{TNLean.PEPS.insertResidual}) is not the bare blocked weight of \(R\): it carries
          (i) a consistency delta forcing \(\mu\) and \(\eta\) to agree on the \(v\)-incident edges
          running to \((\operatorname{insert} v\,R)^c\) (these are \(S\)-boundary but not
          \(R\)-boundary), formalized as the vanishing of the residual on a mismatched outer
          \(v\)-incident edge (\leanid{TNLean.PEPS.insertResidual_eq_zero_of_inconsistent}); and
          (ii) the interior bond multiplicity of the \(S\)-boundary edges away from \(R\), which are
          free in the residual sum but pinned by \(\mu\), the inserted-site overcounting
          multiplicity \leanid{TNLean.PEPS.insertOuterBondProd}. The residual refines the
          bridge-label blocked weight of \(R\)
          (\leanid{TNLean.PEPS.insertResidual_eq_filter_regionBlockedWeight}), and grouping the
          bridge-label blocked-weight sum by overwriting the free outer edges with \(\mu\) collapses
          it to the multiplicity times the residual when \(\mu\) and \(\eta\) are consistent
          (\leanid{TNLean.PEPS.regionBlockedWeight_bridge_eq_smul_insertResidual}, with the fiber
          cardinality \leanid{TNLean.PEPS.insertFiber_card}). Assembling the local-configuration
          grouping (\leanid{TNLean.PEPS.regionBlockedWeight_insert_eq_sum_localConfig}) with this
          collapse gives the inserted-site factorization
          (\leanid{TNLean.PEPS.insertOuterBondProd_smul_regionBlockedWeight_insert} and
          \leanid{TNLean.PEPS.insertOuterBondProd_smul_insertResidual_eq}): the inserted-site
          multiplicity times the blocked weight of \(S\) is \(\sum_\eta
          A.\mathrm{component}\,v\,\eta\,(\sigma_v)\) against the bridge-label blocked weight of
          \(R\) on the consistent configurations. The multiplicity is bond-dimension data, hence
          identical for \(A\) and for the reindexed \(\widetilde B\); it therefore cancels in the
          quotient of the two region proportionalities, leaving \(\sum_\eta
          A.\mathrm{component}\,v\,\eta \cdot (c_R\,\widetilde B\,\text{-residual})\).

          The scalar extraction on top of this factorization is now landed. Assembling the
          inserted-site factorization with the residual collapse reads the inserted-site
          multiplicity times the blocked weight of \(\operatorname{insert} v\,R\) as the sum over the
          inserted-site local configurations of the inserted-site tensor against the bridge-label
          blocked weight of \(R\) on the configurations consistent with \(\mu\), and zero otherwise
          (\leanid{TNLean.PEPS.insertOuterBondProd_smul_regionBlockedWeight_insert_eq_bridge}). A
          consistent local configuration at \(v\) is determined by \(\mu\) and its bridge label --- it
          reads the bridge label on the \(v\)-incident edges bounding \(R\) and \(\mu\) on the
          remaining \(v\)-incident edges --- so two consistent configurations with the same bridge
          label coincide (\leanid{TNLean.PEPS.localConfig_eq_of_insertConsistent}). Feeding the two
          region proportionalities \(A_R=c_R\,\widetilde B_R\) and \(A_S=c_S\,\widetilde B_S\) (the
          outputs of \leanid{TNLean.PEPS.regionComplement_comparison} at \(R\) and
          \(S=\operatorname{insert} v\,R\)) through this factorization, cancelling the bond-only
          inserted-site multiplicity, and substituting the \(R\)-proportionality leaves a vanishing
          combination of \(\widetilde B\)'s \(R\)-blocked family weighted by
          \(c_R\,A_v(\eta)-c_S\,\widetilde B_v(\eta)\) over the consistent \(\eta\). The consistent
          configurations inject into their bridge labels, so linear independence of \(\widetilde
          B\)'s \(R\)-blocked family separates the combination to a single term each, forcing the
          per-vertex relation \(A_v=(c_S/c_R)\cdot\widetilde B_v\) at every local configuration
          (\leanid{TNLean.PEPS.component_eq_of_regionProportional}). When the comparison tensor is the
          gauge-absorbed second tensor \(\operatorname{applyGauge} B\,X\), the reindexed inserted-site
          tensor is the gauge action \(\operatorname{gaugeVertex} B\,X\,v\), so the extraction reads
          as the per-vertex gauge relation
          \(A_v=(c_S/c_R)\cdot\operatorname{gaugeVertex} B\,X\,v\)
          (\leanid{TNLean.PEPS.component_eq_gaugeVertex_of_twoBlockProportional}), exactly the shape
          of \path{hPV}. Consequently the torus scalar condition follows directly from the two
          two-block proportionalities at every vertex with a common ratio
          (\leanid{TNLean.PEPS.lambda_pow_card_torus_eq_one_of_twoBlockProportional}).

          The production of a \emph{single} per-edge family carrying the absorbed equality at every
          edge is now landed. The conjugation-to-absorbed conversion is packaged so the bare-edge
          absorbed equality at a single edge follows from one coefficient-identity witness, depending
          only on the gauge at that edge (the open-edge gauge cancellation isolates the boundary
          edge), so the per-edge absorbing choices are independent
          (\leanid{TNLean.PEPS.edgeAbsorbed_of_conjIdentity},
          \leanid{TNLean.PEPS.edgeAbsorbed_of_edgeCoeffIdentityWitness}). Feeding the two
          orientation-class coefficient-identity witness families to that conversion produces one
          per-edge gauge family \(X\) --- the orientation-adapted absorbing gauge read off each chosen
          witness --- for which the bare-edge absorbed equality against \(\operatorname{applyGauge}
          B\,X\) holds at every torus edge
          (\leanid{TNLean.PEPS.exists_edgeAbsorbed_torus_rectangle}). The region-independence step
          then multiplies this back up to the region absorbed equality, supplying the comparison's
          \path{hregion} and producing the region-block proportionality \(A_R\propto\widetilde B_R\)
          at \emph{every} comparison region from the bare-edge equality
          (\leanid{TNLean.PEPS.regionAbsorbed_of_edgeAbsorbed},
          \leanid{TNLean.PEPS.twoBlockProportional_of_edgeAbsorbed}); the proportionality scalar is
          unique once the comparison block is injective
          (\leanid{TNLean.PEPS.twoBlockScalarProportional_unique}). The torus scalar condition is then
          assembled directly from the bare-edge absorbed equality plus the per-vertex two-block
          proportionalities with a common ratio
          (\leanid{TNLean.PEPS.lambda_pow_card_torus_eq_one_of_edgeAbsorbed}).

          What remains for an unconditional \path{hPV}, and hence for an unconditional torus
          Fundamental Theorem, are three pieces, all consuming the landed \(X\) above. First, the
          \emph{orientation-uniformity} of that absorbing family --- that it is one horizontal matrix
          and one vertical matrix up to per-edge scalars, the shape the Fundamental Theorem's gauge
          conclusion \leanid{TNLean.PEPS.IsTorusOrientationUniformGaugeFamilyModScalar} asserts. The
          absorbing matrix \leanid{TNLean.PEPS.absorbedBoundaryGauge} is \(Z^{\mathsf T}\) or
          \(Z^{-1}\) according to whether the boundary edge's left endpoint lies in the red blocking
          region, and the engine gauge \(Z_e\) is the transported reference; both are
          orientation-determined, but the endpoint-membership branch varies across the translates of
          a class with wraparound, which is the transpose/orientation bookkeeping that remains.
          Second, the blocked-tensor injectivity of the comparison regions of the reindexed
          \(\operatorname{applyGauge} B\,X\) (the gauge-absorbed analogue of the rectangle injectivity
          of \(B\)), and of \(A\)'s one-site-different comparison regions and their complements.
          Third, the translation-invariant common ratio \(c_S/c_R\) of the
          (now uniquely-named, \leanid{TNLean.PEPS.twoBlockScalarProportional_unique}) proportionality
          scalars across the vertices. The scalar extraction itself is no longer on the critical
          path.
    \item Prove the scalar condition $\lambda^{nm}=1$ for the square-lattice
          theorem, and state separately the scalar freedom in the non-translation
          invariant normal theorem.

          The square-lattice scalar condition is formalized: when all per-vertex
          scalars equal a single $\lambda$, their product over the
          $\text{width}\times\text{height}$ lattice is $\lambda$ raised to the site
          count, which the injective theorem forces to be one, giving
          $\lambda^{\,\text{width}\cdot\text{height}}=1$. The square-lattice theorem
          skeleton records the top-down assembly: from the orientation-uniform gauge
          absorption datum and the per-vertex single-scalar relation (the region
          comparison output) the scalar condition follows. A second entry to the same
          assembly consumes the gauge in the shape the edge blocking actually
          delivers: an orientation-uniform gauge fixed only up to vertex-balanced
          per-edge constants, which absorb identically, so the exact
          orientation-uniform reference carries the post-absorption equality and the
          per-vertex relation against it yields the same scalar condition. Both
          skeletons remain conditional on the open per-edge gauge of
          obligation~5.\footnote{
          Formal declarations:
          \leanid{TNLean.PEPS.card_squareLatticeVertex},
          \leanid{TNLean.PEPS.lambda_pow_card_eq_one},
          \leanid{TNLean.PEPS.fundamentalTheorem_normalSquarePEPS},
          \leanid{TNLean.PEPS.tiNormalGaugeAbsorptionData_of_modScalar}, and
          \leanid{TNLean.PEPS.fundamentalTheorem_normalSquarePEPS_of_modScalarGauge}.}
\end{enumerate}

\section{Source-to-formalization ledger}

\begin{itemize}
    \item Lemma \path{injective_union}: blueprint
          \path{lem:peps_injectiveRegion_union}, issue \#1374. The
          positive-bond formalization is the sibling node
          \path{lem:peps_injectiveRegion_union_pos}
          (\leanid{TNLean.PEPS.regionBlockedTensorInjective_union_overlap}); the
          unrestricted node remains the source statement without the
          positive-bond hypothesis.
    \item Finite iteration of Lemma \path{injective_union}: blueprint
          \path{lem:peps_regionInjectivityUnionClosure_finite}, issue \#1374.
    \item Regions $R$, $S$, and $T$: blueprint
          \path{def:peps_normalSquareBlockingRegions}, issue \#1375. This
          node also records the conditional implication from coordinate
          rectangular injectivity, union closure, and a rectangular cover of
          the displayed $T$-region to the abstract $R,S,T$ blocking-region
          structure.\footnote{The formal declaration is
          \leanid{TNLean.PEPS.normalSquareBlockingRegions_of_TCover}.}
    \item Injectivity of the two edge blocks removed from the displayed
          $T$-window, and of their union under union closure: blueprint
          \path{lem:peps_normalSquareRegionT_edgeBlocks_injective}, issue
          \#2004.
    \item General rectangular-cover criterion for square-lattice regions:
          blueprint \path{lem:peps_normalSquareRectangleCover_injective}, issue
          \#2004.
    \item Rectangular-cover criterion for injectivity of the displayed
          $T$-region: blueprint
          \path{lem:peps_normalSquareRegionT_injective_of_cover}, issue
          \#2004. This node also contains the rotated vertical local
          $T$-region cover criterion.
    \item Obstruction to applying that criterion directly to the present
          origin-based local-window $T$ models and edge-complement models:
          blueprint
          \path{lem:peps_normalSquareRegionT_no_five_by_six_cover}, issue
          \#2004.
    \item Finite-lattice complementary block around an edge: blueprint
          \path{lem:peps_normalSquareEdgeComplementRegion}, issue \#2004.
    \item Rectangular-cover criterion for injectivity of that block: blueprint
          \path{lem:peps_normalSquareEdgeComplementRegion_injective_of_cover},
          using the cover definition
          \path{def:peps_normalSquareRectangleCover}, issue \#2004.
    \item Extension of a local $T$ rectangular cover to a rectangular cover of
          the normalized horizontal-edge complementary block by adjoining the
          two top-collar rectangles: blueprint
          \path{lem:peps_normalSquareEdgeComplementRegion_coverOfT}, issue
          \#2004.
    \item Normalized $5\times7$ top-collar decomposition of that block:
          blueprint
          \path{lem:peps_normalSquareEdgeComplementRegion_topCollar}, issue
          \#2004. This node also contains a normalized $7\times5$
          collar identity for the analogous rectangular frame and the rotated
          vertical edge-complement collar identity.
    \item Normalized horizontal and vertical red/blue/complement blocking data:
          blueprint
          \path{lem:peps_normalSquareHorizontalEdge_blockingData}, issue
          \#2004. These use the common one-edge blocking-data structure.
          They also expose the corresponding injective-chain conclusions. The
          horizontal complement injectivity is reduced to a local $T$-cover,
          and the vertical complement injectivity is reduced to a rotated
          local $T$-cover.
    \item Translated horizontal red/blue/complement blocking data:
          blueprint
          \path{lem:peps_normalSquareHorizontalTranslatedEdge_blockingData},
          issue \#2004. This is the coordinate-origin-parametric horizontal
          step toward the every-edge construction. Its coordinate origin
          specializes to the normalized horizontal picture.
    \item Translated vertical red/blue/complement blocking data:
          blueprint
          \path{lem:peps_normalSquareVerticalTranslatedEdge_blockingData},
          issue \#2004. This is the coordinate-origin-parametric vertical
          step toward the every-edge construction. Its coordinate origin
          specializes to the normalized vertical picture.
    \item Conditional construction from translated edge windows:
          blueprint
          \path{thm:peps_normalSquareTranslatedEdgeBlockingHypotheses_of_windows},
          issue \#2004. This records the formal boundary before proving that
          every edge in the finite square lattice admits such a translated
          window. Coordinate right and upward edges are tied to the
          corresponding translated windows, and the margin hypotheses for
          actual coordinate right/up edges are explicit. Arbitrary horizontal
          and vertical edges inherit these criteria from their coordinate
          representatives. The oriented margin-and-cover data structure records
          the remaining per-edge input in the open rectangular coordinate
          model. One such datum directly supplies one-edge blocking data
          together with the corresponding endpoint, injectivity, disjointness,
          and coverage conclusions, and a family of such data determines the
          normal edge-blocking hypotheses. The resulting hypotheses recover the
          supplied datum and the same combined edge-blocking conclusions at
          each edge. The horizontal and vertical margin
          predicates are named, the coordinate right/up and oriented-edge
          window constructors accept those predicates directly, translated
          windows on coordinate right/up edges force those predicates, and the
          open $7\times7$ graph does not admit a family of these
          translated windows over all edges; four explicit boundary examples
          are recorded as a single conjunction. The per-edge margin-cover structure
          stores the same predicates, and the resulting open-boundary obstruction for
          horizontal edges without enough left or right margin and vertical
          edges without enough bottom or top margin is also formalized,
          including explicit $7\times7$ examples and their four-sided
          conjunction. Thus the present margin criterion is a sufficient interior
          criterion rather than the final every-edge construction.
          Consequently, the nonexistence of a family of this margin-cover data
          over all edges of the current open $7\times7$ graph is also
          formalized.
    \item A family of one-edge blocking data determining the uniform
          every-edge hypotheses, including the combined endpoint, injectivity,
          disjointness, and coverage assertion: blueprint
          \path{thm:peps_normalEdgeBlockingHypotheses_injectiveChain}, issue
          \#2004.
    \item Edge blocking into red, blue, and complementary injective regions:
          blueprint \path{thm:peps_normalEdgeBlockingToInjectiveChain}, issue \#1375.
    \item Application of Lemma \path{inj_isomorph} after normal blocking:
          downstream issues \#1367, \#1368, \#1363, and \#1364.
    \item Translationally invariant gauge absorption:
          blueprint \path{thm:peps_tiNormalGaugeAbsorption}, issue \#1375.
    \item Final $R$ and $S$ comparison:
          blueprint \path{thm:peps_oneVertexComplementComparison}, issue \#1360.
    \item Theorem \path{3}: blueprint
          \path{thm:fundamentalTheorem_normalSquarePEPS}, issue \#1375.
    \item General theorem \path{normal}: blueprint
          \path{thm:fundamentalTheorem_normalPEPS}, issue \#1375.  A
          scope-restricted connected-graph form is now proved; see the next
          section.
\end{itemize}

\section{The general-graph assembly and its scope restrictions}

The general theorem \path{normal} is now assembled on an arbitrary connected
finite simple graph, in a scope-restricted form.\footnote{Formal declarations:
    \leanid{TNLean.PEPS.fundamentalTheorem_normalPEPS} (existence, blueprint
    \path{thm:fundamentalTheorem_normalPEPS_connected}) and
    \leanid{TNLean.PEPS.fundamentalTheorem_normalPEPS_gauge_unique}
    (uniqueness, blueprint
    \path{thm:fundamentalTheorem_normalPEPS_gauge_unique}).}
The route is the source's: the blocking hypotheses produce a per-edge gauge
from the coherent three-block frames at every edge, the gauges are absorbed
into the second tensor, the one-site comparisons give a nonzero scalar
$\lambda_v$ at every site, the closed state forces $\prod_v\lambda_v=1$, and
the spanning-tree edge-scalar solve of the injective theorem absorbs the
scalars into the gauges, leaving the scalar-free local gauge relation.  The
uniqueness clause is the per-edge conjugator determinacy: two families
realizing the scalar-free relation are proportional at every edge.

The deltas against the source statement (lines 1576--1583 of
\path{Papers/1804.04964/paper_normal.tex}) are the following.

\begin{enumerate}
    \item \emph{Shared blockings, injective for both tensors.}  The formal
          hypotheses instantiate the blocking bundle at the conjunction
          predicate: each blocking region is injective for both tensors at
          once.  This is the reading under which the source's final comparison
          contracts both tensors over the same regions; the source phrase
          ``\emph{they} can be blocked'' is about both states.
    \item \emph{Single crossing edge.}  The per-edge gauge lives on the
          distinguished edge only when that edge is the entire bond between
          the red and blue blocks; with several red-to-blue crossings the
          isomorphism lemma would produce a gauge on the collected bond, not
          one matrix per edge.  The formalization takes the single-crossing
          condition as an explicit hypothesis, reading ``blocked \emph{around}
          every edge'' as in the source's Theorem \path{3} picture, where the
          two blocks adjacent to the edge touch only along it.  The recorded
          blocking bundle does not carry this condition as a field.
    \item \emph{Matched bond dimensions.}  The equality of the two bond
          dimension assignments is a hypothesis, as in the torus and
          open-lattice assemblies.  The elimination plan is the region-level
          analogue of the injective theorem's edgewise derivation: the blocked
          insertion-algebra equivalence on each bond forces equal sizes.
    \item \emph{Positive bonds and connectivity.}  These are the
          counterexample-backed corrections carried over from the injective
          theorem (\path{docs/paper-gaps/peps_injective_ft_section3_route.tex}
          and \path{docs/paper-gaps/peps_gaugeConsistency_connectivity_gap.tex});
          the scalar absorption is exactly the step that requires
          connectivity, matching the source's remark that the proportionality
          constants can be incorporated into the gauges.
\end{enumerate}

The third delta is now eliminated: the equality of the two bond dimension
assignments is no longer a hypothesis of the existence
theorem.\footnote{Formal declaration:
    \leanid{TNLean.PEPS.bondDim_eq_of_normalBlocking} (blueprint
    \path{thm:peps_bondDim_eq_of_normalBlocking}), consumed by
    \leanid{TNLean.PEPS.fundamentalTheorem_normalPEPS}.}  The derivation is
the recorded elimination plan, the region-level analogue of the injective
theorem's edgewise derivation, run on the coarse three-site chains: around
each edge the two blocked chains have state coefficients equal to positive
merge-collapse constants times the original state coefficients, hence
proportional states; multiplying one super-site of each chain by the other
chain's constant restores an exact state equality while preserving the
super-site injectivities and the bond dimensions, the blocked
insertion-algebra equivalence on the distinguished coarse bond then forces
equal sizes (lines 560--583 of \path{Papers/1804.04964/paper_normal.tex}, the
remark that the algebra isomorphism makes the bond dimensions on the two
sides agree), and under the single-crossing hypothesis the coarse bond
carries the original bond dimension of the distinguished edge for each
tensor.  The per-edge uniqueness clause keeps the equality among its
hypotheses, since the families of gauges it compares are typed by the
identification of the two bond spaces that the existence theorem produces.

Two degenerate one-site comparisons allowed by the hypothesis bundle are
handled directly rather than through the boundary-edge engine: when the
smaller comparison region is empty its block is the count of global virtual
configurations, and when the one-site completion is the full vertex set its
block is the closed state coefficient; in both cases the comparison
proportionality holds with scalar one.  Every other comparison region has a
boundary edge by connectivity.

\section{Blueprint diagrams}

The normal route has its own tensor-network diagrams. These are not substitutes
for the injective theorem diagrams; they attach the additional normal blocking
layer to named proof nodes.\footnote{Chapter~13a calls the public TikZ commands
    in \path{blueprint/src/macros/tn_print.tex}. The web blueprint renders
    cached SVGs from the same commands through
    \path{blueprint/src/Packages/tn_diagrams.py}.}

\begin{itemize}
    \item Lemma \path{injective_union} shows the four regions
          $A\setminus B$, $A\cap B$, $B\setminus A$, and $(A\cup B)^c$.
    \item Definition \path{peps_normalSquareBlockingRegions} shows the regions
          $R$, $S$, and $T$ from the square-lattice proof.
    \item Lemma \path{peps_normalSquareEdgeComplementRegion_topCollar} shows the
          finite-lattice complementary block $A_3$ as the local $T$-region
          together with the two top-collar rectangles in the normalized
          $5\times7$ frame.
    \item Theorem \path{peps_normalEdgeBlockingToInjectiveChain} shows the
          red/blue/complementary blocking around a distinguished edge and the
          resulting three-site chain.
    \item Theorem \path{peps_tiNormalGaugeAbsorption} shows the final
          translationally invariant local gauge formula with horizontal gauge
          $X$ and vertical gauge $Y$.
\end{itemize}

\section{Conclusion}

The normal PEPS route is separated from the injective PEPS route in the
blueprint and in the GitHub issue tree. The tensor-level proof of Lemma
\path{injective_union} is now complete for positive bond dimensions, in both the
disjoint and the overlapping case; the region-injectivity vocabulary, the
four-region decomposition, and the abstract contraction-chain criterion that
supported it are present. The next genuine mathematical steps are the
source-faithful local and rotated $T$-injectivity argument and the every-edge
finite-geometry construction. After these normal blocking hypotheses are proved,
the downstream proof should reuse the injective edge-insertion route rather than
duplicate it. The translationally invariant gauge absorption and the scalar
condition remain the final steps of the normal theorem.

\section{The vertex-frame obstruction and the block-frame route}

The region insertion machinery formalized so far inserts the comparison matrix
at the \emph{in-region endpoint vertex} $v$ of the boundary edge $f$ and
realizes it through $v$'s tensor (\path{regionInsertionOp}). With this frame the
coefficient transfer reduces cleanly to a single residual-frame statement: the
$v$-side row of the first tensor is of single-bond insertion form. Every step of
that reduction uses only blocked-region injectivity of the region and its
complement.

The remaining $v$-side row statement, however, does not close under blocked-region
injectivity alone. Discharging it evaluates the first tensor's endpoint operator,
which factors as
$\Phi^A_v \circ (\text{incident insertion}) \circ \Psi^A_v$ with $\Psi^A_v$ a left
inverse of $\Phi^A_v=\path{localTensorMap}\ A\ v$, on the \emph{second} tensor's
region weight vectors. This requires the cross-tensor endpoint image inclusion
$\operatorname{range}(\Phi^B_v)\subseteq\operatorname{range}(\Phi^A_v)$, i.e.\ the
two single-vertex tensor images at $v$ coincide. That coincidence is available only
from single-vertex spanning (\path{span_stateOpenCoeff_eq_top}, derived from
\path{vertexComplementTensorInjective_of_isVertexInjective}), which is genuine
\emph{vertex} injectivity. Blocked-region injectivity is linear independence of the
blocked weight family indexed by boundary configurations; for a region with more
than one site it does not force the single-vertex marginal at $v$ to be injective,
and two tensors with the same global state can have distinct $v$-column spaces. The
disanalogy with the edge argument is structural: the edge resonate identity is
single-tensor (the abstract operators act on one family's components), whereas the
region resonate identity reads the first tensor's operator on the second tensor's
vectors, introducing a cross-tensor image mismatch absent at the edge level.

Hence the vertex-endpoint frame closes only the \emph{vertex-injective}
specialization, where the injective Fundamental Theorem already supplies the gauge
(\path{fundamentalTheorem_normalPEPS_of_vertexInjective}). The source's actual
argument is block-framed: the injective blocks play the role of super-sites, and the
insertion is realized on the \emph{block's} open boundary leg through the blocked
tensor, so the image-preservation step becomes the block-level range coincidence
$\operatorname{range}(\text{blocked map of }A)=\operatorname{range}(\text{blocked map of }B)$,
which is provable from blocked-region injectivity together with the same-state
hypothesis (the blocked-weight analogue of
\path{range_localTensorMap_eq_of_sameState}). The conditional chain from the
coefficient transfer to the per-edge gauge is already block-injective-clean
(\path{exists_regionEdgeGauge_of_coeffTransfer}, no single-vertex injectivity), so a
block-framed proof of the coefficient transfer feeds it directly. The block-framed
realization---a block-leg insertion operator with its own block-level resonate
inversion---is the genuine remaining obligation for the general region-injective
theorem and the recommended next development.

This block-framed realization is now delivered through the coarse three-site route.
The three injective blocks are assembled into a coarse tensor on a three-vertex
complete graph whose super-sites are injective directly from the blocked-region
injectivities, with no single-vertex injectivity of the original tensor; the proven
injective inserted-coefficient correspondence runs on its distinguished super-bond,
the bond-model collapse descends the coarse coefficient to the single-edge
region-inserted coefficient, and the shared positive fiber multiplicity cancels. The
result is the block-framed coefficient transfer in both directions
(\path{exists_regionInsertedCoeff_eq_sharedRegion},
\path{exists_regionInsertedCoeff_eq_sharedRegion_symm}), the two bond-local transfer
kernels (\path{isBondLocalTransferKernel_of_coherentFrames}), the forward
multiplicativity (\path{regionEdgeTransfer_mul}), and hence the per-edge gauge
(\path{exists_regionEdgeGauge_of_coherentFrames}), for two coherent frames sharing
the three regions, the bond dimensions, and the single distinguished crossing edge.
The block-level step $V=W$ and the forward multiplicativity --- the genuine remaining
obligation above --- are thereby closed modulo review, conditional only on the
coherent coarse blocking frames the source's edge geometry supplies. What remains for
the general theorem is constructing those frames at every edge from the source's
finite-lattice blocking, the geometry obligation recorded earlier in this note.

\section*{Closure on the torus: the unconditional theorem}

The geometry obligation is discharged on the discrete torus. The coherent frames are
instantiated at every edge from the contiguous-rectangle injectivity hypotheses, and
the per-edge gauges are absorbed into the second tensor, giving the bare-edge absorbed
inserted-coefficient equality at every torus edge. Three further steps complete the
assembly. First, a gauge preserves blocked-region linear independence
(\path{regionBlockedTensorInjective_applyGauge}): the gauged blocked weights are the
original weights mixed by an invertible boundary coupling, the product over the
boundary edges of the surviving endpoint gauges. Second, the absorbing family is
constructed covariantly: each edge carries the transported reference gauge of its
orientation class, so the family commutes with the torus translations
(\path{exists_torusCovariantAbsorbedGaugeFamily}). Third, translation covariance
transports the corner-region comparison to every site with the same proportionality
scalars (\path{twoBlockScalarProportional_translate}), so the inserted-site extraction
yields one ratio $\lambda = c_S/c_R$ at every vertex. The result is the unconditional
torus theorem (\path{fundamentalTheorem_normalTorusPEPS_unconditional}): from
translation invariance, matched positive bond dimensions, the same state, and the
rectangle injectivity hypotheses alone, there are a translation-covariant gauge family
and a single scalar $\lambda$ with $A_v = \lambda\cdot(\text{gauge action of } B)_v$ at
every site and $\lambda^{nm}=1$. The union closure of injective regions is derived
from the positive bond dimensions, not assumed.

One convention finding accompanies the closure. The ordered-edge normal form stores a
wraparound edge with its endpoints swapped, so the absorbing family carries the
transposed inverse of the class matrix on the seam edges. A family that is one matrix
per orientation class up to scalars in the stored convention cannot satisfy the
absorbed equality at the seam (it would force $Z^{\mathsf T}$ proportional to
$Z^{-1}$), so the lexicographic uniform-up-to-scalar predicate misstates the source's
``same matrix on all horizontal (vertical) edges'' on the torus; translation
covariance is the orientation-faithful form. The earlier conditional torus statement,
whose per-vertex hypothesis is phrased against a lexicographically uniform family, is
unsatisfiable at seam edges in general and is superseded.

The superseded apparatus has been removed. Gone are: the conditional torus
statement itself (\path{fundamentalTheorem_normalTorusPEPS}); the lexicographic
uniform-up-to-scalar predicate and its selection lemmas, which gathered per-edge
agreement with one matrix per orientation class into a uniform family
(\path{IsTorusOrientationUniformGaugeFamilyModScalar},
\path{TorusOrientationGauge.lean}, \path{TorusTINormalGauge.lean},
\path{TorusClassAgreement.lean}); the production of such a family from the
rectangle hypotheses and its per-edge feeders
(\path{isTorusOrientationUniformGaugeFamilyModScalar_torus_rectangle},
\path{isTorusOrientationUniformGaugeFamilyModScalar_of_edgeWitnesses},
\path{edgeCoeffIdentityWitness_translateUniform},
\path{TorusRectangleWitness.lean}); and the per-edge-chosen absorbed family
superseded by the covariant construction
(\path{exists_edgeAbsorbed_torus_rectangle}). The witness interface, the witness
transport along translations, the per-class witness families, the single-edge
absorbed-equality conversion, and the determinacy bridges all remain: they carry
the covariant route of the unconditional theorem.

The uniqueness clause of Theorem~3 --- ``$X$ and $Y$ are unique up to a
multiplicative constant'' --- is closed. Two families realizing the bare-edge
absorbed equality at an edge are proportional there
(\path{torusAbsorbedGauge_unique_scalar}): each family's absorbed equality lifts to
the region inserted-coefficient identity in conjugation form
(\path{regionConjIdentity_of_edgeAbsorbed}), the transfer map is determined by the
identity, and the centralizer of the bond matrix algebra is the scalars. For
translation-covariant families the per-edge scalar is one constant per orientation
class, transported along the translations
(\path{torusCovariantAbsorbedGauge_unique_classScalar}); the honest class statement
carries the constant $c$ on the edges stored in the natural cyclic order and
$c^{-1}$ on the seam edges, because the covariance places the transposed inverse of
the class matrix on a seam edge and the transposed inverse of a proportionality
inverts its constant. The clause also holds against the per-vertex relation itself
(\path{torusGauge_unique_scalar_of_perVertex}): the inserted-edge coefficient is
multilinear in the site tensors, so the relation $A_v = \lambda\cdot(\text{gauge
action of }B)_v$ with $\lambda^{nm}=1$ already implies the bare-edge absorbed
equality at every edge (\path{edgeAbsorbed_of_perVertex}).

What remains against the source statement of Theorem~3: the torus sizes are
$n,m\geq 9$ (large enough to host the two reference rectangles) where the source
states $n,m\geq 7$; and the source's square-lattice region form of the theorem is
covered separately by the conditional square-lattice statement recorded earlier in
this note.

\section*{Removal of the conditional square-lattice skeleton}

The conditional square-lattice skeleton recorded earlier in this note has been
removed. Its assembly assumed single-site injectivity of $A$
(\path{IsVertexInjective}), the hypothesis the normal theorem exists to avoid: its
scalar condition rested on the vertex-injective product identity rather than the
region-injective one (\path{prod_perVertexScalar_eq_one_of_regionInjective}), and
the interior-window open-lattice theorem
(\path{fundamentalTheorem_normalSquarePEPS_unconditional}) and the general-graph
gauge-equivalence theorem (\path{fundamentalTheorem_normalPEPS}) now cover the
square-lattice content from region injectivity alone. Gone are: the conditional
statements themselves (\path{fundamentalTheorem_normalSquarePEPS},
\path{fundamentalTheorem_normalSquarePEPS_of_modScalarGauge},
\path{lambda_pow_card_eq_one}, \path{card_squareLatticeVertex}); the
orientation-uniform carrier on the open lattice
(\path{SquareLatticeUniformBondDim}, \path{orientationUniformGauge},
\path{IsOrientationUniformGaugeFamily},
\path{IsOrientationUniformGaugeFamilyModScalar}, their selection lemmas, and
\path{NormalSquareTI.lean}); and the translation-invariant absorption datum and its
feeders (\path{tiAbsorb}, \path{TINormalGaugeAbsorptionData},
\path{tiNormalGaugeAbsorption}, \path{tiNormalGaugeAbsorptionData_of_modScalar},
\path{exists_postAbsorption_of_vertexInjective},
\path{gaugeEquivModEdgeScalars_of_matrixScalar},
\path{absorbEdgeGauges_eq_of_matrixScalar}, and
\path{NormalSquareTIAbsorption.lean}). The vertex-injective product identity
\path{prod_perVertexScalar_eq_one} itself remains: it serves the injective theorem
(Theorem~2), where single-site injectivity is the source hypothesis. A
source-exact translationally invariant square-lattice statement awaits a
translation-invariance predicate on the open lattice, as the earlier sections
record.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
