\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}
\providecommand{\leanid}[1]{\path{#1}}

\title{Injective PEPS Gauge Consistency Needs a Connectivity Hypothesis}
\author{}
\date{2026-06-06}

\begin{document}
\maketitle

\section{The assertion}

The gauge-consistency step of the Fundamental Theorem for injective PEPS on a
simple graph asserts that two vertex-injective tensors generating the same state
are related by one global edge-gauge family. In the source this is the passage
after the equation labelled \path{eq:inj_equal_edge}: blocking one vertex against
its complement gives, for each vertex, a scalar proportionality
$A_v=\lambda_v\,\widetilde B_v$, and the source then says ``the constants
$\lambda_v$ can be incorporated into the gauge
transformations.''\footnote{\path{Papers/1804.04964/paper_normal.tex:1207}.}

The formal statement is \leanid{TNLean.PEPS.gaugeConsistency}: for all
$v,\eta,\sigma$,
\[
  B_v(\eta)_\sigma = \operatorname{gaugeVertex}(A,X)_v(\eta)_\sigma ,
\]
for a single global gauge family $X$ on the edges. The headline
\leanid{TNLean.PEPS.fundamentalTheorem_PEPS} reduces to this step.

\section{The statement is false without connectivity}

The reduction to scalars is correct, but the absorption of the vertex scalars
$\lambda_v$ into the edge gauges is \emph{not} possible on an arbitrary graph.
The reason is structural. Each $\lambda_v$ multiplies the whole vertex tensor;
absorbing it into the edge gauges requires nonzero edge scalars $s_e$ whose
oriented product around each vertex equals $\lambda_v^{-1}$. The oriented
incidence product of edge scalars at a vertex contributes $s_e$ at the tail and
$s_e^{-1}$ at the head, so for every assignment of edge scalars the product of
all vertex values is $1$ on each connected component. The vertex scalars
$\lambda_v$ are therefore absorbable only when their reciprocal family has
product $1$ on each component, which is forced by the state equality only for a
connected graph (where there is a single component).

For a disconnected graph the per-component products need not be $1$, and the
exact single-gauge conclusion fails. The simplest witness is the empty graph.

\paragraph{Witness (empty graph on two vertices).}
Take $G=(\{0,1\},\varnothing)$, physical dimension $1$, and tensors
\[
  A_0=2,\quad A_1=3,\qquad B_0=6,\quad B_1=1 .
\]
Both are vertex injective (each vertex tensor is a single nonzero number). The
states agree:
\[
  \langle A\rangle = A_0\,A_1 = 6 = B_0\,B_1 = \langle B\rangle .
\]
All bond dimensions are vacuously positive (there are no edges). On the empty
graph the only edge-gauge family is the empty one, and
$\operatorname{gaugeVertex}(A,X)_v=A_v$ for every $v$, so the conclusion forces
$A=B$. But $A_0=2\ne 6=B_0$. Hence the conclusion of
\leanid{TNLean.PEPS.gaugeConsistency} fails on this input.

This construction has been machine-checked and is axiom-clean (only
\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}, with no
\texttt{sorryAx}): the negation of the conclusion of
\leanid{TNLean.PEPS.gaugeConsistency} for this input is a theorem. The same
input refutes \leanid{TNLean.PEPS.fundamentalTheorem_PEPS}, since the empty edge
gauge gives $\operatorname{GaugeEquiv}A\,B\iff A=B$.

\section{The correct hypothesis}

The source theorem is intended for connected PEPS; a PEPS on a disconnected
graph is a product of independent states and is outside the scope of the local
blocking argument. The faithful repair is to add a connectivity hypothesis to
\leanid{TNLean.PEPS.gaugeConsistency} and \leanid{TNLean.PEPS.fundamentalTheorem_PEPS}
(for example \texttt{G.Connected}, or \texttt{G.Preconnected} together with
nonemptiness of $V$). Under connectivity the vertex scalars satisfy
$\prod_v\lambda_v=1$ (from the nonvanishing state equality), and a spanning-tree
construction produces edge scalars realizing $\lambda_v^{-1}$ as the oriented
incidence product at every vertex.

\section{What is established}

\begin{itemize}
\item \emph{Per-vertex scalar.} The reduction to per-vertex scalars is
  derivable sorry-free: combining the post-absorption insertion equality
  \leanid{TNLean.PEPS.post_absorption_edge_insertion_equality} with the four
  two-block injectivity facts
  (\leanid{TNLean.PEPS.isTwoBlockInjective_vertexTwoBlock},
  \leanid{TNLean.PEPS.isTwoBlockInjective_complementTwoBlock} for $A$ and the
  reindexed $\widetilde B$), the bridge
  \leanid{TNLean.PEPS.sameTwoBlockInsertions_of_edgeInsertedCoeff_eq}, and
  \leanid{TNLean.PEPS.one_vertex_complement_comparison} yields, at every
  non-isolated vertex $v$, a nonzero $c_v$ with $A_v=c_v\,\widetilde B_v$.
\item \emph{Gauge inversion.} Inverting the absorbed gauge recovers
  $B_v=c_v^{-1}\operatorname{gaugeVertex}(A,Z^{-1})_v$, reducing the goal to the
  edge-scalar solve.
\item \emph{Remaining step.} The edge-scalar solve
  $\prod_{e\ni v}s_{v,e}=c_v^{-1}$ is solvable exactly when $G$ is connected and
  $\prod_v c_v=1$. This is the spanning-tree coboundary step and is the only
  remaining obligation once the connectivity hypothesis is added.
\end{itemize}

This connectivity gap is distinct from the balanced-edge-scalar gap of
\path{docs/paper-gaps/peps_gauge_edge_scalars.tex}: that note concerns the
\emph{uniqueness} clause (\leanid{TNLean.PEPS.gauge_unique_mod_edge_scalars}),
where even a connected graph admits a non-global edge-scalar freedom; this note
concerns the \emph{existence} clause, where disconnectedness breaks the single
global gauge outright.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
