\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Scope of the Formal Canonical-Form Reductions Relative to PGVWC07}
\author{}
\date{2026-05-12}

\begin{document}
\maketitle

\section{The source theorem}

The translation-invariant canonical-form theorem of
P\'erez-Garc\'ia, Verstraete, Wolf, and Cirac starts from an arbitrary
translation-invariant MPS representation on a finite ring
\cite[Theorem~\texttt{Th:TIcanonical}, lines~742--763]{PerezGarcia2007MPSReps}.
It states that the matrices can always be decomposed as
\[
    A_i =
    \begin{pmatrix}
        \lambda_1 A_i^1 &0               &0\\
        0               &\lambda_2 A_i^2 &0\\
        0               &0               &\cdots
    \end{pmatrix},
\]
where $1 \geq \lambda_j > 0$ and each block satisfies:
\[
    \sum_i A_i^j A_i^{j\dagger} = \mathbf{1},\qquad
    \sum_i A_i^{j\dagger}\Lambda^j A_i^j = \Lambda^j,
\]
for a diagonal positive full-rank matrix $\Lambda^j$, and $\mathbf{1}$ is the
only fixed point of
\[
    \mathcal{E}_j(X) = \sum_i A_i^j X A_i^{j\dagger}.
\]
If the initial representation has bond dimension $D$, the canonical
representation has bond dimension at most $D$.

The statement does not assume injectivity, a condition C1, blockwise
primitivity, a prior block decomposition, strict separation of the weights,
or that only one block occurs in each weight class. These properties enter
elsewhere in the paper as later hypotheses, consequences, or genericity
conditions. For example, the injectivity statement appears only after the
canonical form has been obtained
\cite[Proposition~\texttt{prop-inj}, lines~911--950]{PerezGarcia2007MPSReps},
and the generic uniqueness discussion is a separate theorem
\cite[Theorem~\texttt{thm-uniq}, lines~1002--1015]{PerezGarcia2007MPSReps}.

\section{Faithful proof order}

The proof of Theorem~\texttt{Th:TIcanonical} is short but highly ordered, and
the formalization should preserve that order rather than replace it by later
canonical-form theorems. The source proof proceeds as follows.

\begin{enumerate}
    \item Lines~765--770 normalize the spectral radius and, when the positive
          fixed point is full rank, gauge the tensor by \(X^{-1/2}\) and
          \(X^{1/2}\) to obtain the unital condition
          \(\sum_i B_iB_i^\dagger=\mathbf{1}\).

    \item Lines~771--783 treat a non-full-rank positive fixed point. The
          support projection \(P_R\) is not an independent hypothesis: it is
          derived from the spectral decomposition of the singular positive
          fixed point, using the positivity contradiction in lines~775--783.
          This is the point at which the invariant-subspace hypothesis is
          proved; it is not an assumption imported from a prepared block
          decomposition.

    \item Lines~785--815 use the derived relation
          \(A_iP_R=P_RA_iP_R\) to split the trace over the finite ring into the
          \(R\) and \(R^\perp\) restrictions. The cyclic trace calculation is a
          finite-ring argument: the \(P_R\)-term passes through the product and
          cancels against \(P_{R^\perp}\). This is the source of the recursive
          block decomposition and of the bond-dimension non-increase.

    \item Lines~816--826 iterate the same splitting on each block and then use
          a non-scalar fixed point \(X\neq\mathbf{1}\) to produce the positive
          non-full-rank fixed point
          \(\mathbf{1}-\lambda_1^{-1}X\). This is the step that enforces
          uniqueness of the identity fixed point in each final block.

    \item Lines~827--832 apply the analogous argument to the dual map and use a
          unitary conjugation to diagonalize the resulting full-rank positive
          fixed point, producing the matrices \(\Lambda^j\) in the theorem.
\end{enumerate}

Thus the source theorem is not a primitive-block theorem, an injective-block
theorem, or a prepared-family theorem. Nor is the invariant-projection lemma by
itself enough: the projection must be obtained from the positive fixed-point
argument in the cited lines, and the finite-ring trace split must then be
carried out before the recursive decomposition is asserted. A formal proof of
Theorem~\texttt{Th:TIcanonical} should reconstruct these steps from an
arbitrary translation-invariant representation and use later notions only as
lemmas after their hypotheses have been obtained in this proof order.

\section{Normalization orientation}

There is also a convention difference in the direction of the canonical
normalization. In the source theorem, the block transfer map is written as
\[
    \mathcal{E}_j(X) = \sum_i A_i^j X A_i^{j\dagger}
\]
and imposes the unital condition
\[
    \sum_i A_i^j A_i^{j\dagger} = \mathbf{1}.
\]
The local canonical-form chapter usually works in the dual
trace-preserving, or left-canonical, orientation
\[
    \sum_i A_i^{j\dagger} A_i^j = \mathbf{1}.
\]
This is a convention difference when it is accompanied by the corresponding
adjoint transfer map or conjugate-transposed Kraus family. It is nevertheless
important to state the convention explicitly: a theorem written only in the
left-canonical orientation should not be read literally as PGVWC07
Theorem~\texttt{Th:TIcanonical} unless the dualization or gauge conversion
has also been supplied.

\section{The formal reductions}

The currently checked statements in the canonical-form reduction chapter
prove conditional reductions from prior block hypotheses. In particular,
the primitive fixed-point reduction assumes
that the blocks are injective, trace-preserving in the chosen gauge, strictly
ordered by weight modulus, nonzero, positive-dimensional, and primitive. The
peripheral-spectrum variant assumes the same injectivity, normalization,
strict weight ordering, nonzero weights, and positive block dimensions,
together with the statement that each block transfer map has peripheral
spectrum $\{1\}$. The normal-canonical-form reduction starts from a weighted
block decomposition already equivalent to the original state and assumes
irreducibility, trace-preserving normalization, primitivity, pairwise distinct
weight moduli, nonzero weights, and positive bond dimensions.

These are correct conditional consequences, but they are not formalizations
of the full PGVWC07 translation-invariant canonical-form theorem. Their
hypotheses are stronger and more structured than the source theorem's
hypotheses. Treating them as the full source theorem would import extra
assumptions into any later argument that cites canonical-form existence.

\section{Current proof-path audit}
\label{sec:current-proof-path-audit}

The earlier existence path now has the following formal pieces.

\begin{enumerate}
    \item
          \path|MPSTensor.spectralUnitalGauge_isUnital_of_transferMap_eigenvector|
          formalizes the spectral-radius normalization part of
          Theorem~\texttt{Th:TIcanonical}, lines 765--769.  Given a positive
          definite eigenvector \(\rho\) with
          \(\mathcal{E}_A(\rho)=r\rho\) and \(r>0\), it proves that
          \(B_i=r^{-1/2}\rho^{-1/2}A_i\rho^{1/2}\) satisfies
          \(\sum_i B_iB_i^\dagger=\mathbf{1}\).

    \item \path|MPSTensor.unitalGauge_isUnital_of_transferMap_fixedPoint|
          formalizes the full-rank fixed-point gauge in
          Theorem~\texttt{Th:TIcanonical}, lines 767--769.  Given a positive
          definite fixed point \(\rho\) of \(\mathcal{E}_A\), it constructs the
          right-canonical gauge \(B_i=\rho^{-1/2}A_i\rho^{1/2}\) and proves
          \(\sum_i B_iB_i^\dagger=\mathbf{1}\).  The companion theorem
          \path|MPSTensor.sameMPV_unitalGauge| records preservation of the
          finite-ring coefficients under this similarity.

    \item
          \path|MPSTensor.exists_singular_posSemidef_fixedPoint_of_unital_nonScalar_fixedPoint|
          formalizes the fixed-point shift in
          Theorem~\texttt{Th:TIcanonical}, lines 819--826.  For a unital block,
          a Hermitian fixed point which is not a scalar multiple of the identity
          gives the singular positive fixed point
          \(\lambda_{\max}\mathbf{1}-X\).  The companion theorem
          \path|MPSTensor.exists_twoBlock_decomp_of_unital_nonScalar_fixedPoint|
          composes this shift with the support-projection split, giving a
          strict two-block MPV-equivalent decomposition.  The formal hypothesis
          uses ``non-scalar'' rather than merely \(X\ne\mathbf{1}\), since every
          scalar multiple of the identity is fixed by a unital linear map and
          cannot force a nontrivial split.

    \item
          \path|MPSTensor.fixed_eq_scalar_of_isIrreducibleTensor_unital|
          formalizes the fixed-point endpoint in
          Theorem~\texttt{Th:TIcanonical}, lines 816--826.  Once the block is
          irreducible and unital, every fixed point of its transfer map is a
          scalar multiple of \(\mathbf{1}\).  The proof uses the MPS
          irreducible-tensor/irreducible-map bridge and the already formalized
          channel theorem \path|MPSTensor.fixed_eq_scalar_of_irreducible_unital|
          (Wolf Theorem 6.6).  Thus the identity is the unique fixed point
          after the usual scalar normalization of the fixed-point line.

    \item
          \path|MPSTensor.exists_unitary_diag_posDef_adjointFixedPoint_of_unital_of_isIrreducibleTensor|
          formalizes the dual-map diagonalization in
          Theorem~\texttt{Th:TIcanonical}, lines 827--832, conditional on having
          already reached an irreducible unital block.  The proof applies the
          existing trace-preserving diagonalization theorem to the
          conjugate-transposed Kraus family \(i\mapsto A_i^\dagger\).  This is
          not a duplicate spectral proof: unitality of \(A\) is trace
          preservation for \(A^\dagger\), and translating back gives a unitary
          gauge \(B_i=U^\dagger A_iU\) with
          \[
              \sum_i B_iB_i^\dagger=\mathbf{1},
              \qquad
              \sum_i B_i^\dagger\Lambda B_i=\Lambda,
          \]
          for diagonal positive-definite \(\Lambda\).

    \item
          \path|MPSTensor.exists_unital_data_of_irreducible| formalizes the
          Perron--Frobenius unital-gauge step in
          Theorem~\texttt{Th:TIcanonical}, lines 765--770, for a single
          irreducible nonzero block.  It obtains a positive-definite
          eigenvector \(\rho\) for the transfer map by applying the adjoint
          Perron--Frobenius theorem to the conjugate-transposed Kraus family,
          and then forms
          \(B_i=r^{-1/2}\rho^{-1/2}A_i\rho^{1/2}\) with
          \(\sum_i B_iB_i^\dagger=\mathbf{1}\).  This is a scoped
          single-block input, not the full recursive canonical-form theorem.

    \item
          \path|MPSTensor.exists_pgvwc07_unital_dualDiag_data_of_irreducible|
          composes the single-block source steps in
          Theorem~\texttt{Th:TIcanonical}, lines 765--770 and 816--832.  Starting
          from one irreducible nonzero block, it constructs the
          Perron--Frobenius unital representative
          \(B_i=r^{-1/2}\rho^{-1/2}A_i\rho^{1/2}\), proves that the fixed
          points of \(\mathcal{E}_B\) are scalar multiples of the identity, and
          then applies the dual fixed-point diagonalization to a unitary
          conjugate of \(B\).  This removes a previous composition gap for a
          single block.  It is still not the full source theorem, because the
          construction has not yet been threaded through the recursive
          finite-ring block decomposition, the positive weights, and the
          total bond-dimension bound.

    \item
          \path|MPSTensor.exists_pgvwc07_unital_dualDiag_blockwise|
          applies the preceding single-block theorem to a prepared family of
          nonzero irreducible summands with unit weights.  It constructs the
          positive spectral-radius weights, unital block representatives,
          scalar transfer-map fixed points, and diagonal positive-definite dual
          fixed points block by block, while preserving the weighted MPV
          family.  This closes the blockwise assembly step after the
          irreducible summands have already been obtained.  It is still not
          the full source theorem, because the statement assumes the prepared
          nonzero irreducible decomposition and does not yet incorporate the
          arbitrary-input zero-block separation or the final total
          bond-dimension bound.

    \item
          \path|MPSTensor.exists_pgvwc07_unital_dualDiag_from_arbitrary_with_zeroTail|
          composes the zero-block separation with the preceding blockwise
          PGVWC07 theorem.  Starting from an arbitrary tensor, it retains the
          all-zero summands as one zero block and puts the nonzero part into
          the unital orientation, with positive spectral-radius weights,
          scalar transfer-map fixed points, and diagonal positive-definite dual
          fixed points.  This removes the prepared-family hypothesis from the
          blockwise statement while still keeping the zero-block contribution
          explicit in the finite-ring MPV identity.  It is still not the full
          source theorem; this theorem itself does not include the final
          total bond-dimension bound.

    \item
          \path|MPSTensor.exists_pgvwc07_unital_dualDiag_from_arbitrary_with_zeroTail_bondDimBound|
          adds the length-zero trace calculation to the preceding arbitrary
          tensor theorem.  It proves
          \(D_0+\sum_k D_k=D\), where \(D_0\) is the retained zero-block bond
          dimension and the \(D_k\) are the nonzero block dimensions.  Hence
          \(\sum_k D_k\leq D\), matching the total bond-dimension estimate in
          Theorem~\texttt{Th:TIcanonical}, lines 761--762, while still keeping
          the zero block explicit.

    \item
          \path|MPSTensor.exists_pgvwc07_unital_dualDiag_from_arbitrary_posMPV_bondDimBound|
          removes the explicit zero-block summand from the displayed MPV
          identity by restricting the equality to positive chain lengths.
          The zero block has no positive-length coefficients, so the nonzero
          weighted direct sum has the same positive-length MPV family as the
          original tensor, and the estimate \(\sum_k D_k\leq D\) is retained.
          This matches the source theorem on nonempty rings.  A statement that
          also quantifies over the empty word must either keep the zero block or
          separately account for its bond dimension.

    \item
          \path|MPSTensor.PGVWC07PositiveLengthWitness| and
          \path|MPSTensor.exists_pgvwc07_positiveLengthWitness| record the
          preceding positive-length theorem as a single package of data.  The
          fields contain the nonzero unital blocks, positive real weights,
          diagonal positive-definite dual fixed points, scalar transfer-map
          fixed-point conclusions, positive block dimensions, positive-length MPV equality,
          and the total bond-dimension bound.  Nonvanishing of the weights is
          a direct consequence of their positive-real form, so nonvanishing is
          not an additional field of the witness.  This is not a new proof
          route; it is an equivalent formulation of the proved positive-length
          theorem, intended to prevent later arguments from repeating the long
          existential conclusion.

    \item
          \path|MPSTensor.PGVWC07PositiveLengthWitness.exists_weight_normalization|
          formalizes the finite-family scalar normalization corresponding to
          Theorem~\texttt{Th:TIcanonical}, lines 765--766, for a witness with
          at least one nonzero block.  Dividing the positive real weights by
          their maximum gives normalized positive weights with
          \(|\nu_k|\leq 1\) and equality for one block.  The theorem also
          records the resulting global factor \(s^N\) in length-\(N\) MPV
          coefficients.

    \item
          \path|MPSTensor.PGVWC07PositiveLengthWitness.block_count_pos_of_exists_ne_zero_mpv|
          proves that a positive-length witness is nonempty whenever the
          original tensor has at least one nonzero positive-length MPV
          coefficient.  If the block family were empty, the block-diagonal MPV
          expansion would be an empty sum at every positive length, contradicting
          the witness equality with the chosen nonzero coefficient.

    \item
          \path|MPSTensor.exists_pgvwc07_normalized_exact_form_after_rescaling_of_exists_ne_zero_mpv|
          composes the arbitrary-input witness theorem with the preceding
          nonemptiness lemma and the exact weight-normalization theorem.
          Under the nonzero positive-length MPV
          hypothesis, it produces a positive scalar \(s\) and normalized
          weights \(|\nu_k|\leq 1\), with equality for one block, such that the
          globally rescaled tensor \(s^{-1}A\) has exactly the same
          positive-length MPV coefficients as the normalized weighted block
          tensor.  This formalizes the exact finite-ring effect of the
          ``without loss of generality'' spectral-radius normalization in
          Theorem~\texttt{Th:TIcanonical}, lines 765--766.

    \item
          \path|MPSTensor.exists_pgvwc07_normalized_exact_form_after_rescaling_or_forall_pos_mpv_eq_zero|
          records the arbitrary-input zero/nonzero dichotomy for the exact
          rescaled formulation.  Either all positive-length MPV coefficients of
          the original tensor vanish, or the preceding exact rescaled normalized
          PGVWC07 form exists.  This still keeps the zero positive-length branch
          explicit and therefore does not by itself settle the final length-zero
          convention for Theorem~\texttt{Th:TIcanonical}.

    \item
          \path|MPSTensor.exists_pgvwc07_normalized_exact_form_after_rescaling_allow_empty|
          removes the disjunction in the preceding theorem by adopting the
          positive-length convention that the zero branch has an empty
          nonzero-block family.  Thus the positive-weight, unital,
          scalar-fixed-point, and dual-diagonal block conditions are vacuous in
          the branch where every positive-length MPV coefficient vanishes.  In
          the nonzero branch the family is nonempty and still has a normalized
          weight of norm one.  This is an exact positive-length formulation:
          after a positive global rescaling of the original tensor, the
          normalized weighted block tensor has the same coefficients on every
          nonempty ring, and \(\sum_k D_k\leq D\).

    \item
          \path|MPSTensor.exists_pgvwc07_unital_dualDiag_from_arbitrary_with_zeroTail_bondDimBound|
          records the complementary explicit all-zero direct-summand convention before the
          final finite-family weight normalization.  It keeps the all-zero summand
          dimension from the arbitrary-input decomposition, proves
          positive-length agreement with the weighted nonzero blocks, and
          records the length-zero dimension identity
          \(D_0+\sum_kD_k=D\), together with the bond-dimension bound.

    \item \path|MPSTensor.exists_irreducible_blockDecomp| formalizes the
          recursive invariant-projection splitting used in the proof of
          Theorem~\texttt{Th:TIcanonical}, lines 765--833. It starts from an
          arbitrary tensor and gives a finite direct sum of irreducible blocks
          with the same matrix-product-vector coefficients. This is a faithful
          partial proof step. It is not the full source theorem, because it does
          not construct the weights $\lambda_j$, the unital block orientation,
          the diagonal full-rank matrices $\Lambda^j$, the uniqueness of the
          identity fixed point, or the final bond-dimension bound.

    \item \path|MPSTensor.exists_irreducible_blockDecomp_nonzeroBlocks|
          separates the all-zero blocks from the irreducible summands and keeps
          the length-zero trace contribution. This matches the paper's statement
          that zero blocks may occur, but it is still only an intermediate
          decomposition before Perron--Frobenius normalization.

    \item \path|MPSTensor.exists_tp_gauge_from_arbitrary_with_zeroTail|
          applies Perron--Frobenius normalization block by block after the
          all-zero summands have been separated. It produces nonzero
          irreducible blocks in the local left-canonical orientation
          $\sum_i (B_k^i)^\dagger B_k^i = \mathbf{1}$. This is the dual
          orientation of the unital condition in
          Theorem~\texttt{Th:TIcanonical}; the conversion is a convention issue
          only when the adjoint transfer map is carried along explicitly.

    \item \path|MPSTensor.exists_normalCanonicalForm_of_primitive_blockDecomp|
          is a prepared-block theorem. It assumes a weighted block decomposition,
          irreducibility, trace-preserving normalization, primitivity, pairwise
          distinct weight moduli, nonzero weights, and positive bond dimensions.
          It should remain a conditional corollary, not a citation target for
          PGVWC07 Theorem~\texttt{Th:TIcanonical}.
\end{enumerate}

Thus the primitive, peripheral-primitive, and normal-canonical-form reductions
listed above should not be read as the PGVWC07 translation-invariant
canonical-form theorem.  The blueprint instead uses the checked positive-length
statement as the theorem statement for
Theorem~\texttt{Th:TIcanonical}.  The finite-family part of the paper's
convention \(1\geq \lambda_j>0\) is formalized by rescaling all positive
weights by their maximum.  In exact coefficients this introduces the global
factor \(s^N\) at length \(N\), so the primary exact statement rescales the
original tensor by \(s^{-1}\).

\section{Formal boundary}

The currently checked primitive, peripheral-primitive, and
normal-canonical-form reductions should therefore be read as conditional
canonical-form theorems rather than as
formalizations of the arbitrary-input source theorem.\footnote{The formal
declarations are listed in Section~\ref{sec:current-proof-path-audit}.}
They may be used after their hypotheses have been proved from earlier
source-faithful reductions, but they should not be cited as giving canonical
form directly from an arbitrary translation-invariant MPS representation.

\section{Current conclusion}

The formalization now has a theorem cited for PGVWC07
Theorem~\texttt{Th:TIcanonical}, in the positive-length convention used by the
blueprint:
\path|MPSTensor.exists_pgvwc07_normalized_exact_form_after_rescaling_allow_empty|.
It starts from an arbitrary translation-invariant MPS representation.  After a
positive global rescaling, it constructs a finite family of positive weights,
unital blocks, diagonal full-rank dual fixed points, scalar transfer-map fixed
points, and the bond-dimension bound.  The nonzero-block family is allowed to
be empty exactly in the case where every positive-length MPV coefficient
vanishes.

The explicit zero-block/length-zero bookkeeping is recorded earlier by
\path|MPSTensor.exists_pgvwc07_unital_dualDiag_from_arbitrary_with_zeroTail_bondDimBound|.
It keeps the all-zero summand dimension, gives positive-length agreement with the
weighted nonzero blocks, and proves \(D_0+\sum_kD_k=D\), together with the
bond-dimension bound.  This is the companion length-zero convention before the
final finite-family weight normalization, while the blueprint cites the
positive-length theorem as the primary statement.

The remaining conditional primitive, peripheral-primitive, and
normal-canonical-form reductions are later consequences used only after their
extra hypotheses have been established.  They are not alternate citations for
Theorem~\texttt{Th:TIcanonical}.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
