\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Scope of the Formal Block-Separation Lemmas Relative to PGVWC07 Uniqueness}
\author{}
\date{2026-05-12}

\begin{document}
\maketitle

\section{Source theorem}

The translation-invariant canonical-form uniqueness theorem of
P\'erez-Garc\'ia, Verstraete, Wolf, and Cirac starts with two TI canonical
\(D\)-MPS representations of the same finite-ring state.  We write their
matrices as \(B_i\) and \(C_i\), following the source theorem
\cite[Theorem~\texttt{thm-uniq}, lines~1002--1015]{PerezGarcia2007MPSReps}.
For a TI representation \(A_i\), PGVWC07 defines
\[
  \Gamma_L(X)
  =
  \sum_{i_1,\ldots,i_L}
    \operatorname{tr}\!\left[X A_{i_1}\cdots A_{i_L}\right]
    |i_1\cdots i_L\rangle .
\]
The finite injectivity hypothesis called C1 in the source is the existence of
a finite length \(L_0\) for which \(\Gamma_{L_0}\) is injective
\cite[lines~888--908]{PerezGarcia2007MPSReps}.  The theorem assumes:
\begin{enumerate}
  \item this finite injectivity hypothesis holds;
  \item the open-boundary-condition (OBC) canonical representation of the MPS
    state is unique;
  \item the length satisfies \(N > 2L_0 + D^4\).
\end{enumerate}
The conclusion is the existence of a unitary matrix \(U\) such that
\[
  B_i = U C_i U^\dagger
\]
for every physical index \(i\), with the usual permutation-and-phase
interpretation when the Schmidt weights are non-degenerate.

\section{Formal boundary}

The checked block-separation lemmas in Chapter~9 are not formalizations of
PGVWC07 Theorem~\texttt{thm-uniq}. They are conditional same-structure
comparison results. Their hypotheses include blockwise normalization, injective
or irreducible block hypotheses, a common block index set and dimension
function, and strict ordering of the weight moduli.\footnote{The corresponding
formal declarations are
\leanid{MPSTensor.per\_block\_sameMPV\_of\_\allowbreak canonical\_form},
\leanid{MPSTensor.per\_block\_sameMPV\_of\_\allowbreak normal\_canonical\_form},
\leanid{MPSTensor.fundamentalTheorem\_\allowbreak canonicalForm}, and
\leanid{MPSTensor.fundamentalTheorem\_\allowbreak canonicalForm\_explicit}.}

These assumptions are useful local inputs for later MPS arguments, but they are
not the hypothesis set of the PGVWC07 uniqueness theorem. In particular, the
source theorem does not assume a same-structure block decomposition of the two
representations, nor does it assume the Chapter~9 strict-modulus separation
lemma as an input. Its proof uses the OBC canonical representation and the
Horn--Johnson matrix lemma route described after
\cite[Theorem~\texttt{thm-uniq}, lines~1002--1015]{PerezGarcia2007MPSReps}.

Thus this is a theorem-level gap. PGVWC07 Theorem~\texttt{thm-uniq} is not
currently formalized. The available formal lemmas are conditional
same-structure consequences with stronger hypotheses, and therefore must not be
cited as formalizations of that source theorem.

\section{Required formalization}

A source-faithful formalization of PGVWC07 Theorem~\texttt{thm-uniq} should
state the finite-ring injectivity hypothesis above, uniqueness of the OBC
canonical representation, and \(N > 2L_0 + D^4\), and should conclude a single unitary
intertwiner between the two TI canonical \(D\)-MPS representations. The current
Chapter~9 block-separation lemmas may be used only as local consequences after
their stronger hypotheses have been proved in the argument at hand.

\section{Tracked consequences}

Blueprint statements that point to Chapter~9 block separation must explicitly
state the strict weight-modulus hypothesis when their Lean declarations require
it. Downstream CPSV16 uses should not cite these statements as PGVWC07
canonical-form uniqueness unless the finite-injectivity, OBC uniqueness,
finite-size-bound, and unitary-conclusion theorem have been formalized separately.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
