\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\input{command}

\title{PGWSVC08 String Order: Physical Endpoints versus Virtual Boundary
  Witnesses}
\author{}
\date{2026-06-11}

\begin{document}
\maketitle

\section{The source statement}

The string-order paper~\cite{PerezGarcia2008StringOrder} defines string
order for a translationally invariant state $\Psi$ by the existence of a
local unitary $u \ne \mathbb 1$ and local operators $x, y$ (hermitian
without loss of generality) acting on the physical sites flanking the
string, such that
\begin{align}
  \lim_{N\to\infty} |S_N(x,y,u,\Psi)| > 0,
  \qquad
  S_N(x,y,u,\Psi) := \langle\Psi|\, x \otimes u^{\otimes N} \otimes y\,
  |\Psi\rangle.
  \tag{\path{SOP}}
\end{align}%
\footnote{Source: \cite[lines~112--122]{PerezGarcia2008StringOrder},
  local file \path{Papers/0802.0447/StringOrder-v10.tex}.  A footnote of
  the source (lines~124--127) asserts, by citation to
  Verstraete--Mart{\'i}n-Delgado--Cirac, Phys.\ Rev.\ Lett.\ 92, 087201,
  that an ``alternative but equivalent'' definition gives the endpoints
  access to the virtual levels; the source does not prove this
  equivalence.}
Both the unitary and the endpoint operators are existentially
quantified; the fixed-endpoint string order parameter of the
Haldane-chain literature is the historical special case.  For a pure
finitely correlated state the correlator takes the transfer form
\begin{align}
  S_N(x,y,u,\Psi_\infty)
  = \tr\bigl[\Lambda\, \mathcal E_x\, \mathcal E_u^N\, \mathcal E_y
  (\mathbb 1)\bigr],
  \tag{\path{SOPMP}}
\end{align}%
\footnote{Source: \cite[lines~176--181]{PerezGarcia2008StringOrder}.}
where $\mathcal E_x, \mathcal E_y$ are the transfer maps twisted by the
endpoint operators.  The source's Theorem~1 then states: a pure
finitely correlated state has string order if and only if there exist a
unitary $\tilde u \ne \mathbb 1$, a matrix $V$, and indices $n, m$ with
\begin{align}
  \mathcal E_{\tilde u}(V) = V,
  \qquad
  \tr(V \Lambda A_n A_m^\dagger) \ne 0.
  \tag{\path{Th1}}
\end{align}%
\footnote{Source: \cite[lines~270--276]{PerezGarcia2008StringOrder}.
  The second condition is the realizability of the endpoint by the
  single-site operator $x = |n\rangle\langle m|$ with
  $y = x^\dagger \tilde u$, per the reduction at lines~257--268.}
Immediately afterwards (lines~278--296) the source shows that the second
condition can be dropped when $x$ and $y$ are products of observables on
$D^2$ spins: the spans
$S_m := \operatorname{span}\{A_{n_1}\cdots A_{n_m} A_{m_m}^\dagger
\cdots A_{m_1}^\dagger\}$ grow strictly until they exhaust the full
matrix algebra $M_D(\mathbb C)$, so endpoint operators on sufficiently
many physical sites reach every $D \times D$ matrix.

\section{The formalized physical and virtual forms}

The transfer-form physical correlator is now a formal definition:
\[
  S_N(x,y,u)
  = \tr\bigl[\Lambda\, \mathcal E_x\, \mathcal E_u^N\, \mathcal E_y
  (\mathbb 1)\bigr].
\]
\footnote{\leanid{MPSTensor.physicalStringOrderParam},
  \path{TNLean/MPS/Symmetry/StringOrderDefs.lean}.}
The source-style existential predicate is also formalized: there exist a
unitary $u\ne \mathbb 1$, physical endpoint operators $x,y$, and a
number $s>0$ such that $|S_N(x,y,u)|\to s$.\footnote{\leanid{MPSTensor.HasPhysicalStringOrder},
  same file.}

The main theorems presently use a virtual-boundary form.  The boundary
string-order expression is
$\tr(\Lambda X\, \mathcal E_u^L(Y))$ with arbitrary
$X, Y \in M_D(\mathbb C)$,\footnote{\leanid{MPSTensor.stringOrderBoundaryParam},
  \path{TNLean/MPS/Symmetry/StringOrderDefs.lean}.} and string order is
the existence of such $X, Y$ together with a constant $c > 0$ bounding
$\lVert \tr(\Lambda X\, \mathcal E_u^L(Y)) \rVert$ from below at every
length $L$.\footnote{\leanid{MPSTensor.HasStringOrder}, same file.}  The
equivalence of string order with local symmetry, and the universality of
string order for injective on-site-symmetric tensors with canonical
normalisations, are proved for this virtual
form.\footnote{\leanid{MPSTensor.stringOrder_iff_localSymmetry},
  \leanid{MPSTensor.hasStringOrder_of_symmetric_injective},
  \leanid{MPSTensor.hasStringOrder_iff_of_symmetric_injective},
  \path{TNLean/MPS/Symmetry/StringOrder.lean}.}

The virtual statement differs from the physical source statement in four ways.

\begin{enumerate}
  \item \emph{Physical versus virtual endpoints.}  The source
        quantifies over physical operators $x, y$ of bounded site
        support; the formal definition quantifies over arbitrary virtual
        matrices $X, Y$.  A physical endpoint expands to a particular
        virtual matrix in the transfer picture, and the source's span
        argument (lines~278--296) provides the converse for injective
        tensors, but no formal statement connects the two notions.
  \item \emph{Limit versus uniform bound.}  The source requires a
        positive limit $\lim_{N\to\infty} |S_N| > 0$; the formal
        definition requires a lower bound uniform in the length.  A
        sequence with positive limit can vanish at finitely many small
        lengths and so fail the uniform bound for that witness pair.
  \item \emph{Left boundary form.}  The source's left endpoint acts
        through the adjoint channel, $\mathcal E_x^\dagger(\Lambda)$;
        the formal expression multiplies, $\Lambda X$.  These have the
        same expressive range when $\Lambda$ is positive definite, as it
        is in all theorems concerned, but the formal expression is not
        the literal transfer form of the source.
  \item \emph{Fixed versus existential twist.}  The source's ``has
        string order'' quantifies existentially over the unitary $u$;
        the formal predicate fixes $u$ as a parameter, in the style of
        the source's Theorem~2.  The formal version is the finer
        statement, and the existential form is recoverable from it.
\end{enumerate}

\section{Verdict}

The virtual-boundary reformulation is mathematically sound, and for the
injective tensors that all concerned theorems treat, the existential
physical-endpoint and existential virtual-boundary criteria are
equivalent by the source's own span argument.  The physical correlator
and the positive-limit source predicate are now formal definitions, but
that equivalence is still the part with no formal counterpart: the
development proves its string-order theorems for the virtual form only.
Elimination: prove the realizability of virtual boundary matrices by
physical endpoint operators for injective tensors (a starting point is the
single-site realization
\leanid{MPSTensor.physRealize}\footnote{\path{TNLean/MPS/Chain/VirtualInsertion.lean},
  currently not imported by the string-order files.}), and derive the
source-faithful Theorem~1; tracked in \ghissue{2660}.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
