\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Example Parent Hamiltonians in Cirac--Perez-Garcia--Schuch--Verstraete 2021}
\author{}
\date{2026-06-04}

\begin{document}
\maketitle

\section{Purpose}

This note records the source scope for the example-level parent-Hamiltonian
targets in \href{https://github.com/LionSR/TNLean/issues/2315}{issue \#2315}.
The issue concerns the GHZ, cluster, and AKLT tensors already present in the
formal development.  Its mathematical aim is natural: connect these tensors to
their parent Hamiltonians and to the corresponding ground-space statements.
The point of this note is only to separate what is stated in the local review
source from standard example formulas which need an additional citation or an
explicit derivation.

\section{Statements present in the review}

The review gives the GHZ MPS tensor by
\[
  A^i_{\alpha\beta}=\delta_{i=\alpha=\beta},
\]
and writes the corresponding state as
\[
  |\mathrm{GHZ}\rangle=\sum_{i=0}^{d-1}|i,i,\ldots,i\rangle
\]
\cite[lines~2341--2345]{Cirac2021Matrix}.  Earlier, in the discussion of
symmetry breaking, it identifies the ferromagnetic Ising Hamiltonian
\[
  H=-\sum_i Z_iZ_{i+1}
\]
as the paradigmatic model whose symmetric ground state is the superposition of
all spins up and all spins down, and writes the same MPS tensor as
$A^\uparrow=|0\rangle\langle 0|$ and
$A^\downarrow=|1\rangle\langle 1|$
\cite[lines~1194--1196]{Cirac2021Matrix}.  In the stability discussion it also
says that the parent Hamiltonian of a GHZ state is a ferromagnetic Ising
Hamiltonian with a two-fold degenerate ground space
\cite[line~2205]{Cirac2021Matrix}.  The two-dimensional GHZ PEPS tensor is
recorded separately in the examples appendix
\cite[lines~2417--2421]{Cirac2021Matrix}.
Another local source writes the two-site parent interaction explicitly as the
projector complement of
\[
  \operatorname{span}\{|00\rangle,|11\rangle\},
\]
namely, up to an additive constant,
\[
  \tfrac12 I-\big(|00\rangle\langle 00|+|11\rangle\langle 11|\big).
\]
\footnote{See \path{Papers/1210.6613/UncleHMPS.tex}:432--455.}

For the one-dimensional cluster state, the review gives the MPS tensor
\[
  A^0=|0\rangle\langle +|,\qquad A^1=|1\rangle\langle -|,
\]
and explains its controlled-$Z$ preparation from $|+\rangle^{\otimes N}$
\cite[lines~2364--2369]{Cirac2021Matrix}.  The two-dimensional cluster PEPS is
also displayed in the examples appendix
\cite[lines~2424--2429]{Cirac2021Matrix}.  In the local review source, however,
the one-dimensional stabilizer Hamiltonian
\[
  -\sum_i Z_{i-1}X_iZ_{i+1}
\]
is not stated in the one-dimensional cluster example passage.  The older MPS
review says instead that cluster states are unique ground states of the
three-body interactions
\[
  \sum_i \sigma^z_i\sigma^x_{i+1}\sigma^z_{i+2},
\]
with an explicit two-matrix MPS representation
\cite[local TeX lines~374--387]{PerezGarcia2007MPSReps}.  Thus the issue should
use the $ZXZ$ convention for the controlled-$Z$ cluster state unless a
local-basis variant is explicitly introduced and proved equivalent.  The
general parent-Hamiltonian theorem for normal MPS does imply a unique ground
state for a suitable parent Hamiltonian once the normal-tensor hypotheses and
the range-reduction theorem have been formalized
\cite[lines~2087--2090]{Cirac2021Matrix}; it is not, by itself, the explicit
cluster stabilizer formula.

For the one-dimensional AKLT state, the review constructs the tensor from
spin-$\tfrac12$ singlets and projection onto the joint spin-$1$ subspace, and
then gives the matrices in two bases
\cite[lines~2372--2393]{Cirac2021Matrix}.  In the parent-Hamiltonian section it
states that the AKLT model has injectivity length $L_0=2$, while a two-site
parent Hamiltonian is already sufficient for a unique ground state; the cited
check is
\[
  \mathcal G_{1,2}\cap\mathcal G_{2,3}=\mathcal G_{1,2,3}
\]
\cite[line~2095]{Cirac2021Matrix}.  In the SPT discussion it also records the
dangling spin-$\tfrac12$ edge modes: for open boundary conditions the parent
Hamiltonian has at least the square of the virtual irrep dimension as ground
state degeneracy, and for AKLT this gives a four-dimensional ground space with
only one spin-$0$ state
\cite[lines~1169--1174]{Cirac2021Matrix}.  The standard description of the
AKLT Hamiltonian as a sum of spin-$2$ projectors is not displayed in these local
review passages; it should be cited to the AKLT source or derived from the
spin-addition calculation before it is made a source-faithful target here.
What is displayed in the local sources is the polynomial spin Hamiltonian
\[
  H=\sum_i \vec S_i\cdot\vec S_{i+1}
      +\frac13\big(\vec S_i\cdot\vec S_{i+1}\big)^2,
\]
together with the unnormalised matrices
$\{\sigma^z,\sqrt2\sigma^+,-\sqrt2\sigma^-\}$
\cite[local TeX lines~340--349]{PerezGarcia2007MPSReps}.  The same polynomial
form, with an additive constant, appears in the local Schuch--P\'erez-Garc\'ia--Cirac
source.\footnote{See \path{Papers/1010.3732/paper_v3.tex}:2126--2130.}

\section{Current formal boundary}

The formal example files already contain the tensors and their elementary
structure.  The GHZ tensor is defined as
$A^0=|0\rangle\langle 0|$ and $A^1=|1\rangle\langle 1|$, and the file proves
that it is an RFP, is not injective, and has the expected $\mathbb Z_2$
symmetry.\footnote{See \path{TNLean/MPS/Examples/GHZ.lean}:54--86 and the
module summary at \path{TNLean/MPS/Examples/GHZ.lean}:21--29.}
The cluster file defines the reflected one-dimensional convention
$A^0=|+\rangle\langle 0|$ and $A^1=|-\rangle\langle 1|$, proves
non-injectivity at one site, and proves two-block injectivity.\footnote{See
\path{TNLean/MPS/Examples/Cluster.lean}:62--78 and
\path{TNLean/MPS/Examples/Cluster.lean}:200--215.}
The AKLT file defines the spin-$1$ tensor and proves two-block injectivity, hence
normality.\footnote{See \path{TNLean/MPS/Examples/AKLT.lean}:49--58 and
\path{TNLean/MPS/Examples/AKLT.lean}:222--237.}

The general parent-Hamiltonian API is also present.  The local interaction is the
orthogonal projector onto $\mathcal G_L(A)^\perp$, and the finite-chain parent
Hamiltonian is the sum of translated local terms.\footnote{See
\path{TNLean/MPS/ParentHamiltonian/Defs.lean}:66--75 and
\path{TNLean/MPS/ParentHamiltonian/Defs.lean}:23--27.}
The theorem that the parent Hamiltonian annihilates the MPS vector is already
available.\footnote{See
\path{TNLean/MPS/ParentHamiltonian/Basic.lean}:118--128.}
The normal-MPS range-reduction theorem still depends on the boundary-closing
comparison, which is the remaining formal gap for the general unique-ground-state
statement.\footnote{Formal declaration:
\leanid{wrapped_mirror_witness_agree_of_chainGroundSpace}. See
\path{TNLean/MPS/ParentHamiltonian/UniqueGroundState.lean}:821--852 and
\path{TNLean/MPS/ParentHamiltonian/UniqueGroundState.lean}:854--870.}

\section{Source-faithful order of work}

A source-faithful formalization of issue \#2315 should proceed in the following
order.

\begin{enumerate}
  \item For GHZ, first state the Ising parent-Hamiltonian comparison using the
        review terminology: the ferromagnetic Ising Hamiltonian and its
        two-dimensional ground space.  The two-site projector description can be
        taken from the local uncle-Hamiltonian source.
  \item For cluster, first connect the existing normality proof to the general
        parent-Hamiltonian construction, or prove the three-body $ZXZ$
        interaction stated in the older MPS review.  Any other convention should
        be stated only after recording the local-basis or index-shift equivalence.
  \item For AKLT, first use the source's stated two-site criterion
        $\mathcal G_{1,2}\cap\mathcal G_{2,3}=\mathcal G_{1,2,3}$ and the
        already formalized two-block injectivity.  The polynomial AKLT
        Hamiltonian is locally sourced; the spin-$2$ projector wording should be
        introduced only after citing or proving its equivalence with that
        polynomial form.  The open-boundary fourfold degeneracy should be treated
        separately from the periodic unique-ground-state statement.
\end{enumerate}

\section{Verdict}

This is a scope restriction.  Issue \#2315 is source-faithful only when split
into separate targets: the review-stated parent-Hamiltonian consequences, the
explicit stabilizer or spin-projector formulas, and the open-boundary edge-mode
ground spaces are distinct mathematical statements with distinct source
obligations.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
