\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Source-Notation Gap: The Majumdar--Ghosh Example Tensor}
\author{}
\date{2026-06-16}

\begin{document}
\maketitle

\section{The assertion}

The review \cite{Cirac2021Matrix} describes the Majumdar--Ghosh model as the
one-dimensional resonating-valence-bond state, the superposition of the two
nearest-neighbour singlet coverings of a periodic spin-$\tfrac12$ chain.  It
then writes the MPS tensor in the shorthand
\begin{align}
    A
    =
    \bigl[
      |0\rangle\bigl((02|+(20|\bigr)
      +
      |1\rangle\bigl((12|+(21|\bigr)
    \bigr]\otimes Y,
    \qquad
    Y=
    \begin{pmatrix}0&-1\\1&0\end{pmatrix},
    \tag{\cite[lines~2397--2405]{Cirac2021Matrix}}
\end{align}
where $Y$ is the singlet tensor used earlier in the examples appendix.

The formal example file introduces instead the following bond-dimension-three
matrices:
\begin{align}
    A^0 =
    \begin{pmatrix}
        0&1&0\\
        0&0&0\\
        1/\sqrt2&0&0
    \end{pmatrix},
    \qquad
    A^1 =
    \begin{pmatrix}
        0&0&1\\
        -1/\sqrt2&0&0\\
        0&0&0
    \end{pmatrix}.
    \tag{formal example}
\end{align}
The formalized results about this tensor are its left-canonical equation, its
transfer map on the identity, non-injectivity, and non-normality.\footnote{Formal
declarations:
\leanid{MPSTensor.majumdarGhoshTensor},
\leanid{MPSTensor.majumdarGhosh\_left\_canonical},
\leanid{MPSTensor.majumdarGhosh\_transferMap\_one},
\leanid{MPSTensor.majumdarGhosh\_not\_isInjective}, and
\leanid{MPSTensor.majumdarGhosh\_not\_isNormal} in
\doclink{TNLean/MPS/Examples/MajumdarGhosh}.}

\section{The point at issue}

The review formula is not a literal bond-dimension-three formula.  If
$(ab|$ is read as $|a\rangle\langle b|$ on a three-dimensional auxiliary
factor and the displayed tensor product with $Y$ is taken literally, then the
bond space is
\begin{align}
    \mathbb C^3\otimes\mathbb C^2,
\end{align}
so the bond dimension is $6$.  Writing
\begin{align}
    B^0=E_{02}+E_{20},\qquad B^1=E_{12}+E_{21},
\end{align}
the literal matrices would be $\widetilde A^i=B^i\otimes Y$.  Their two-site
periodic contractions factor:
\begin{align}
    \tr(\widetilde A^i\widetilde A^j)
    =
    \tr(B^iB^j)\,\tr(Y^2).
\end{align}
Since $Y^2=-I_2$ and
\begin{align}
    \tr\bigl((B^0)^2\bigr)=2,\qquad
    \tr\bigl((B^1)^2\bigr)=2,\qquad
    \tr(B^0B^1)=\tr(B^1B^0)=0,
\end{align}
the literal two-letter amplitudes have nonzero same-letter contractions and
zero opposite-letter contractions.  This is not the support pattern of a
two-site singlet, whose coefficients are carried by the opposite-letter words
$01$ and $10$ with opposite signs.

Thus the review formula must be interpreted as valence-bond shorthand rather
than as a literal $6\times6$ MPS representation.  The bond-dimension-three
matrices above are the contracted representative used in the formal example:
the two spin-$\tfrac12$ values of an open singlet partner and the closed
inter-dimer link are kept as the three bond states, and the antisymmetry of
$Y$ supplies the relative sign in $A^1$.

\section{Current formal boundary}

The repository does not yet contain a theorem identifying the matrix-product
vectors of the displayed bond-dimension-three tensor with the Majumdar--Ghosh
dimer superposition on even rings.  Such a theorem would have to state the
finite-size vector identity, with the parity, normalization, and cyclic
convention fixed, for example by comparing the coefficient
\begin{align}
    \tr(A^{i_1}\cdots A^{i_N})
\end{align}
with the sum of the two products of singlet coefficients on the coverings
$(1,2)(3,4)\cdots$ and $(2,3)(4,5)\cdots(N,1)$.

Until that theorem is proved, the formalized object should be described as the
bond-dimension-three contracted representative used for the algebraic example,
not as a formally established presentation of the Majumdar--Ghosh state itself.
The elementary statements already proved about it remain independent of this
identification.

\section{Verdict}

This is an open gap in the example's physical identification.  The literal
reading of the review shorthand does not give the intended singlet-pair support
pattern, while the contracted $D=3$ tensor has not yet been proved in Lean to
produce the Majumdar--Ghosh dimer superposition.  The present formalization
therefore records only the concrete tensor and its algebraic properties.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
