\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\input{command}

\title{Scope Note: The AKLT On-Site Symmetry Is Formalized for the
$\mathbb{Z}_2\times\mathbb{Z}_2$ Subgroup, Not the Full $\mathrm{SO}(3)$}
\author{}
\date{\today}

\begin{document}
\maketitle

\section{The assertion}

The review \cite{Cirac2021Matrix} records that the AKLT matrix product state lies
in a non-trivial symmetry-protected topological phase, and that this phase ``can
be protected by multiple distinct physical symmetries: on-site $\mathrm{SO}(3)$
(with virtual symmetry given by a spin-$\tfrac12$ representation), on-site
$\mathbb{Z}_2\times\mathbb{Z}_2$ (the smallest subgroup of $\mathrm{SO}(3)$ still
exhibiting a non-trivial 2-cocycle), time-reversal, or reflection
symmetry.''\footnote{\cite{Cirac2021Matrix}, around line 1159 of the source.}
In each case the physical symmetry acts on the spin-$1$ site, while the
entanglement (virtual) degrees of freedom carry a projective representation whose
cohomology class is the order-two generator of $H^2$ of the symmetry group with
$\mathrm{U}(1)$ coefficients.

The strongest such statement is on-site $\mathrm{SO}(3)$: the spin-$1$
representation acts on the physical index, and the virtual index transforms in
the projective spin-$\tfrac12$ representation, whose class generates
$H^2(\mathrm{SO}(3),\mathrm{U}(1))\cong\mathbb{Z}_2$.

\section{The point at issue}

Two different statements must be separated: the on-site $\mathrm{SO}(3)$
\emph{symmetry}, and the cohomological \emph{classification} of the phase it
protects. They have very different formalization costs.

\paragraph{The symmetry.}
The symmetry statement says that for every rotation $R\in\mathrm{SO}(3)$ the
twisted tensor is gauge equivalent to the original; for each $R$ this asks only
for the \emph{existence} of a virtual gauge. Most of the required scaffolding is
already available. The groups $\mathrm{SO}(3)$ and $\mathrm{SU}(2)$ exist as
matrix groups, with their group structure, as
$\mathrm{specialOrthogonalGroup}\,(\mathrm{Fin}\,3)\,\mathbb{R}$ and
$\mathrm{specialUnitaryGroup}\,(\mathrm{Fin}\,2)\,\mathbb{C}$. The spin-$1$
representation is not an extra construction at all: it is the defining action of
$\mathrm{SO}(3)$ on the three physical states (followed by a fixed change of
basis), so no exponentiation of angular-momentum generators is needed. The one
genuinely missing input is the surjectivity of the double cover
$\mathrm{SU}(2)\to\mathrm{SO}(3)$: that every rotation is realized by
conjugating the Pauli vector $(\sigma_x,\sigma_y,\sigma_z)$ by an
$\mathrm{SU}(2)$ element. The library does not yet prove this (the Clifford spin
group leaves cover-surjectivity onto the orthogonal group as an explicit open
task, and the quaternion development has no rotation map), but it is a classical
finite-dimensional computation, a single self-contained development rather than a
research project. With that lemma in hand, the on-site $\mathrm{SO}(3)$ symmetry
of the AKLT tensor is within reach of the existing on-site-symmetry predicate:
for each $R$ one takes the gauge to be an $\mathrm{SU}(2)$ preimage of $R$.

\paragraph{The classification.}
The claim that the resulting projective phase is the order-two generator of
$H^2(\mathrm{SO}(3),\mathrm{U}(1))\cong\mathbb{Z}_2$ is a different matter
entirely. It requires the continuous (Borel, or Lie-group) cohomology of
$\mathrm{SO}(3)$, equivalently its Schur multiplier. The group cohomology in the
library is purely algebraic and treats $\mathrm{SO}(3)$ as a discrete group, so
it computes the wrong object; there is no Schur multiplier and no
projective-representation API, and the nascent continuous-cohomology scaffold has
no computations and is not yet reconciled with the discrete theory. This piece is
a substantial development on its own.

\section{The formalized statement}

The formalization proves the $\mathbb{Z}_2\times\mathbb{Z}_2$ case, which the
review singles out as the smallest subgroup of $\mathrm{SO}(3)$ that still carries
a non-trivial $2$-cocycle. Concretely, the two generators act on the spin-$1$
physical index by two commuting involutions, the images of the
$\pi$-rotations about two orthogonal axes, and each twisted tensor is gauge
equivalent to the original. The virtual gauges are
\[
  G_1=\begin{pmatrix}0&1\\-1&0\end{pmatrix}=i\sigma_y,
  \qquad
  G_2=\sigma_z=\begin{pmatrix}1&0\\0&-1\end{pmatrix},
\]
and they anticommute, $G_1G_2=-G_2G_1$. The group commutator therefore acts on
the bond space by $-I$, so the assignment of gauges to group elements is a
projective representation with the non-trivial class in
$H^2(\mathbb{Z}_2\times\mathbb{Z}_2,\mathrm{U}(1))\cong\mathbb{Z}_2$. This is the
mathematical content of the symmetry-protected phase: the phase $-1$ cannot be
removed by rephasing the gauges, so it distinguishes the AKLT chain from a
trivial product state.\footnote{Formalized as \leanid{aklt\_isOnSiteSymmetric\_Z2Z2}
in \doclink{TNLean/MPS/Examples/AKLT}.}

The $\mathbb{Z}_2\times\mathbb{Z}_2$ statement is a genuine restriction of the
$\mathrm{SO}(3)$ statement, not a derivation of it: it exhibits the same
projective class on a finite subgroup, but it does not establish symmetry under
the full continuous group.

\section{Conclusion}

The AKLT example is formalized at the level of the finite
$\mathbb{Z}_2\times\mathbb{Z}_2\subset\mathrm{SO}(3)$ symmetry, which captures the
projective-representation obstruction that witnesses the phase. The full on-site
$\mathrm{SO}(3)$ symmetry asserted by \cite{Cirac2021Matrix} is not yet
formalized, but it is not blocked by any deep obstruction: it is a moderate
development gated almost entirely on the surjectivity of
$\mathrm{SU}(2)\to\mathrm{SO}(3)$, which would be a worthwhile addition to the
library in its own right. The accompanying classification of the phase by
$H^2(\mathrm{SO}(3),\mathrm{U}(1))\cong\mathbb{Z}_2$ is the genuinely large piece,
requiring continuous group cohomology. The blueprint entry for the
$\mathrm{SO}(3)$ statement is therefore kept without a formal tag until the
symmetry is proved, while the $\mathbb{Z}_2\times\mathbb{Z}_2$ statement is tagged
against the formal declaration.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
