\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{The Trace Sign in Wolf's Lorentz-Cone Converse}
\author{}
\date{2026-06-23}

\begin{document}
\maketitle

\section{The assertion}

Wolf's Proposition~3.9 states two elementary trace inequalities for Hermitian
matrices.\footnote{\cite[Proposition~3.9]{Wolf2012Quantum};
local source \path{Notes/WolfNoteTexSource/ch03_positive_not_completely.tex},
lines 693--722.}  The forward implication says that a positive semidefinite
matrix satisfies
\begin{align}
  \tr(A^2) \le \tr(A)^2.
\end{align}
The converse is printed as
\begin{align}
  \tr(A)^2 \ge (d-1)\tr(A^2) \Longrightarrow A \ge 0.
  \tag{\path{Wolf Prop. 3.9 converse}}
\end{align}

\section{The point at issue}

The printed converse cannot be true without a sign condition on the trace.  For
example, if \(A=-I_d\) and \(d\geq 1\), then \(A\) is Hermitian and not positive
semidefinite, while
\begin{align}
  \tr(A)^2=d^2 \ge d(d-1)=(d-1)\tr(A^2).
\end{align}
Thus the squared inequality only controls the Lorentz cone up to time
orientation; it does not distinguish the positive and negative trace sheets.

This is also visible in the proof printed after the proposition.  The argument
decomposes \(A=A_+-A_-\) and uses the estimate
\begin{align}
  \tr(A) \le \tr(A_+).
\end{align}
To contradict the squared inequality one needs the left hand side to have the
same orientation as the nonnegative quantity being bounded.  A nonnegative
trace hypothesis supplies exactly this orientation.

\section{Corrected statement}

The formal theorem proves the trace-nonnegative version:
\begin{align}
  A=A^\dagger,\qquad
  \tr(A)\ge 0,\qquad
  (d-1)\tr(A^2)\le \tr(A)^2
  \Longrightarrow A\ge 0.
\end{align}
In the complex matrix formalization the traces are written with real parts,
which agrees with the displayed real quantities for Hermitian \(A\).\footnote{
Formal declaration
\leanid{Matrix.IsHermitian.posSemidef_of_trace_re_nonneg_of_card_sub_one_mul_trace_sq_re_le}
in \doclink{TNLean/Analysis/MatrixTraceInequalities}; tracked by
\ghissue{3402}.}

The proof is a finite eigenvalue argument.  If some eigenvalue \(\lambda_0\) is
negative, write
\begin{align}
  T=\sum_i\lambda_i,\qquad
  R=\sum_{i\ne 0}\lambda_i .
\end{align}
Since \(T\ge0\) and \(T=\lambda_0+R\), one has \(T<R\), and hence
\(T^2<R^2\).  Cauchy's inequality for the remaining \(d-1\) eigenvalues gives
\begin{align}
  R^2\le (d-1)\sum_{i\ne 0}\lambda_i^2
       \le (d-1)\sum_i\lambda_i^2,
\end{align}
contradicting the assumed Lorentz inequality.  Therefore all eigenvalues are
nonnegative.

\section{Conclusion}

The unrestricted printed converse should not be marked as formalized.  The
formal development records the future-cone, trace-nonnegative converse as the
mathematically valid statement used in the positive-map discussion.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
