\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Wolf Chapter~5 Operator Jensen and Lieb Concavity Boundary}
\author{}
\date{2026-06-23}

\begin{document}
\maketitle

\section{Scope}

This note records the mathematical content of the Chapter~5
operator-convexity boundary\footnote{%
    The declarations are
    \leanid{posMap_rpow_concave_jensen}, \leanid{posMap_rpow_convex_jensen},
    \leanid{posMap_log_concave_jensen}, and \leanid{lieb_concavity_axiom} in
    \doclink{TNLean/Axioms/OperatorConvexity}; the first, third, and fourth are
    now theorems (the Lieb declaration in its positive-definite, boundary-exponent
    scope, see Section~\ref{sec:lieb}); none remains an axiom.}
together with the cited sources, the precise gaps, and the first concrete step
toward eliminating each remaining gap.  The positive-map Jensen declarations
underlie \cite[Corollary~5.2]{Wolf2012Quantum} (the second corollary of Wolf's
\S5.4) and rest on the operator-convexity and
operator-monotonicity-versus-positive-maps results
\cite[Theorems~5.10--5.13]{Wolf2012Quantum}; the Lieb declaration is the
Ando--Lieb trace functional \cite[Theorem~5.15]{Wolf2012Quantum}.  (Wolf's Theorem~5.1 is
Douglas' theorem and is unrelated to this development.)

The notation is fixed throughout.  Operators act on a finite-dimensional Hilbert
space, identified with $D\times D$ complex matrices.  The order $A\le B$ is the
Loewner order ($B-A$ positive semidefinite).  A linear map $T$ on matrices is
\emph{positive} if it sends positive semidefinite operators to positive
semidefinite operators, \emph{subunital} if $T(\mathbf 1)\le\mathbf 1$, and
\emph{unital} if $T(\mathbf 1)=\mathbf 1$.  The power $A^p$ for $A\ge 0$ and the
logarithm $\log A$ for $A>0$ are defined by the continuous functional calculus.

\section{The positive-map Jensen declarations}

\subsection{The assertions}\label{ssec:assertions}

For a positive subunital map $T$ the development asserts the operator Jensen
inequalities
\begin{align}
  T(A^p) &\le (T A)^p, & p&\in[0,1], \ A\ge 0, \tag{\path{posMap_rpow_concave_jensen}}\\
  (T A)^p &\le T(A^p), & p&\in[1,2], \ A\ge 0, \tag{\path{posMap_rpow_convex_jensen}}
\end{align}
and, for a positive \emph{unital} map $T$ and $A>0$,
\begin{align}
  T(\log A) &\le \log(T A). \tag{\path{posMap_log_concave_jensen}}
\end{align}
These are the operator-valued Jensen inequalities behind Wolf
Corollary~5.2.\footnote{%
    They are consumed by
    \leanid{cor52_item1_rpow_of_subunital},
    \leanid{cor52_item2_rpow_of_subunital}, and
    \leanid{cor52_item3_log_of_subunital} in
    \doclink{TNLean/Channel/Schwarz/OperatorMonotone}.}
The first and third are concavity inequalities (for the operator-concave
functions $x\mapsto x^p$ on $[0,1]$ and $x\mapsto\log x$); the second is a
convexity inequality (for the operator-convex function $x\mapsto x^p$ on
$[1,2]$).  The logarithmic case requires $T$ unital, not merely subunital: with
$T$ only subunital the inequality fails because $\log$ is unbounded below near
$0$.

\subsection{The point at issue: the genuine gap is positive-map Jensen}

Mathlib~4.31 supplies the integral
representation\footnote{%
    \leanid{CFC.exists_measure_nnrpow_eq_integral_cfcₙ_rpowIntegrand₀₁}
    and its $[1,2]$ companion in
    \path{Mathlib/Analysis/SpecialFunctions/ContinuousFunctionalCalculus/Rpow/IntegralRepresentation.lean},
    together with the Bochner-integral convexity transfer for C\textsuperscript{*}-algebras.}
and, through it, the operator concavity of $x\mapsto x^p$ on $[0,1]$
(\leanid{CFC.concaveOn_rpow}) and of $\log$ (\leanid{CFC.concaveOn_log}).  The
scalar input of the convex case, operator convexity of $x\mapsto x^p$ on
$[1,2]$, is also now present as \leanid{CFC.convexOn_rpow} in
\doclink{TNLean/Analysis/RpowConvexity}.  Thus the scalar functional-calculus
inputs for the three Jensen declarations are available in the formal
development.

The remaining gap is the step \emph{from} operator concavity (a statement about a
single operator) \emph{to} the Jensen inequality $T(fA)\le f(TA)$ for a positive
map.  This is the Choi--Davis--Jensen inequality (the positive-map form of the
Hansen--Pedersen operator Jensen inequality): for an operator-concave $f$ and a
positive unital map $T$ one has $T(fA)\le f(TA)$, with the subunital version
obtained by dilating $T$ to a unital map.  Mathlib~4.31 contains no such
general theorem for positive maps.  The concave real-power case has now been
discharged by the special Löwner-integrand route recorded in
\leanid{TNLean.OperatorJensen.positiveMap_rpowIntegrand₀₁_jensen}, together
with ordered Bochner integral monotonicity.  The logarithmic case has also been
discharged, not by a general Choi--Davis--Jensen theorem, but by taking the
right limit $p\to 0^+$ in the concave real-power inequality and using
\leanid{CFC.tendsto_cfc_rpow_sub_one_log}.  The convex real-power case is now
likewise proved, by the same positive-map Jensen step applied to the convex
integrand of $x\mapsto x^p$ on $[1,2]$.

\subsection{The compression scaffold is faithful to bare
positivity}\label{ssec:caveat}

A faithfulness worry would arise if the only available proof route needed a
\emph{completely positive} $T$, since the statements assume only that $T$ is
positive (the hypothesis recorded in the formal statements is
\leanid{IsPositiveMap}) and a complete-positivity hypothesis would be the
stronger-hypothesis variant forbidden by the faithfulness rule.  That worry does
not apply here.

Wolf's own proof of the positive-map operator Jensen results
(\cite[Theorems~5.10--5.13]{Wolf2012Quantum}) proves them first for completely
positive maps, then \emph{reduces} the bare positive-map case to the completely
positive case: $A$ and $f(A)$ lie in the abelian $C^*$-algebra generated by $A$
and $\mathbf 1$, on which every positive map is automatically completely
positive (Wolf Prop.~1.6) and admits a completely positive extension to all of
$\mathcal M_d$ (Wolf Prop.~1.7).  Thus no obstruction specific to maps that are
positive but not completely positive survives, and the source corollary itself
is stated for
positive subunital maps, not for channels: \cite[Corollary~5.2]{Wolf2012Quantum}
hypothesises a positive map with $T(\mathbf 1)\le\mathbf 1$, which is exactly the
\leanid{IsPositiveMap} plus subunitality recorded in the two power companions
\leanid{cor52_item1_rpow_of_subunital} and
\leanid{cor52_item2_rpow_of_subunital}.

\paragraph{Local fix (log unitality).}  The logarithmic companion
\leanid{cor52_item3_log_of_subunital} departs from this subunital hypothesis: it
assumes $T$ \emph{unital}, $T(\mathbf 1)=\mathbf 1$, rather than $T(\mathbf
1)\le\mathbf 1$.  This is a deliberate strengthening, not an exact match to the
subunital form of the source corollary: as recorded in \S\ref{ssec:assertions},
the subunital logarithmic inequality $T(\log A)\le\log(TA)$ fails because $\log$
is unbounded below near $0$, so the unital restriction is the mathematically
correct domain for the logarithmic case.  This is a local fix (a hypothesis
sharpened to the regime where the inequality holds), documented here so the
deviation from the verbatim subunital reading of
\cite[Corollary~5.2]{Wolf2012Quantum} is auditable; the two power companions
above retain the bare subunital hypothesis.

The partial scaffold already present in the repository for the concave power
case\footnote{%
    \doclink{TNLean/Channel/Schwarz/OperatorJensenAux} (finite-POVM compression,
    Schur-complement inverse bound, and the resolvent inequality).}
realises this reduction at the matrix level.  It is a finite-POVM compression
argument in terms of a positive family $\{C_i\}$ and a defect block, not a global
Kraus or Stinespring representation of $T$.  Its POVM elements are supplied by
the positive matrices $T(P_i)$, where the $P_i$ are the spectral projections of
$A$; these exist for any positive map (indeed $\sum_i T(P_i)=T(\mathbf 1)$, the
identity used in Wolf's block-positivity step in the proof of
  \cite[Theorem~5.12]{Wolf2012Quantum}).  The scaffold
  therefore applies to a bare positive map in the finite-POVM resolvent reduction
  under the \leanid{IsPositiveMap} hypothesis --- it does not require, and must not
  be relabelled with, complete positivity.  It now proves the concave
  real-power Jensen inequality after integration.  The logarithmic Jensen theorem
  follows from this real-power result by the right-limit formula for
  $(A^p-\mathbf 1)/p$ as $p\to 0^+$.  The convex real-power Jensen declaration
  is now proved the same way, with the convex integrand on $[1,2]$.

For the concave real-power case, the analytic step has now been supplied: the
Löwner-integral argument carries the pointwise finite-POVM resolvent inequality
(\leanid{povm_resolvent_inv_le}) through the integral representation of
\texttt{rpow} to the operator inequality
\leanid{posMap_rpow_concave_jensen}.  The logarithmic inequality
\leanid{posMap_log_concave_jensen} is then obtained by applying
\leanid{CFC.tendsto_cfc_rpow_sub_one_log} to both $A$ and $T(A)$ and closing the
Loewner inequality by the closedness of the finite-dimensional order graph.

\subsection{Current formal inputs}\label{ssec:formal-inputs}

The auxiliary ingredients needed before the remaining positive-map Jensen
assemblies are now present.  The theorem \leanid{CFC.convexOn_rpow} supplies operator
convexity of $x\mapsto x^p$ for $p\in[1,2]$, completing the scalar
functional-calculus input for \leanid{posMap_rpow_convex_jensen}.

The closedness of the finite-dimensional Loewner cone also gives the ordered
Bochner integral step needed for the integral route.  The positive-cone
specialization is recorded as
\leanid{TNLean.OperatorJensen.integral_nonneg_matrix_of_ae}: a matrix-valued
Bochner integral of an almost-everywhere positive semidefinite function is
positive semidefinite.  Almost-everywhere Loewner inequalities are integrated
directly by the general ordered Bochner theorem \leanid{integral_mono_ae}.

\subsection{Verdict}

The three positive-map Jensen statements are mathematically standard
\cite[Theorems~5.10--5.13 and Corollary~5.2]{Wolf2012Quantum}.  The concave
real-power statement \leanid{posMap_rpow_concave_jensen} has been discharged by
the finite-POVM resolvent scaffold, the pointwise integrand inequality, and
ordered Bochner integral monotonicity.  The logarithmic statement
\leanid{posMap_log_concave_jensen} has been discharged as the right limit of the
concave real-power result.  The convex real-power statement
\leanid{posMap_rpow_convex_jensen} has likewise been discharged, by the same
positive-map operator Jensen step applied to the convex integrand.
The integral representation of the power integrand, previously named as the
obstruction, is present.

\section{The Lieb concavity theorem (boundary exponent)}\label{sec:lieb}

\subsection{The assertion}

For $s\in[0,1]$, an arbitrary matrix $K$, and positive-definite $A,B$, the map
\begin{align}
  (A,B)\ \longmapsto\ \tr\!\bigl(K^\dagger A^s K B^{1-s}\bigr)
  \tag{\path{lieb_concavity_axiom}}
\end{align}
is jointly concave on pairs of positive-definite matrices.  This is the
boundary-exponent specialization of the Lieb concavity theorem
(\cite[Theorem~5.15, the Ando--Lieb functional]{Wolf2012Quantum}), the trace
functional underlying the Wigner--Yanase--Dyson conjecture.  It is now
\emph{proved} --- \leanid{lieb_concavity_axiom} is a theorem,
derived from the commuting-superoperator integral representation
(\S\ref{ssec:lieb-gap}) --- and \leanid{lieb_concavity_psd} extends it to
positive-semidefinite inputs by regularization.  The theorem
\leanid{lieb_concavity_subboundary} now derives the two-exponent region
$x,y\ge 0$, $x+y\le 1$, by reducing to the boundary theorem after the change of
variables $r=x+y$, $s=x/r$ when $r>0$.  Thus the finite-dimensional matrix
formalization now covers the exponent range of Wolf's Theorem~5.15.

\paragraph{Sub-boundary exponents (resolved).}  Wolf's Theorem~5.15 is stated
for two exponents $x,y\ge 0$ with $x+y\le 1$.  The first formal theorem
\leanid{lieb_concavity_axiom} records only the boundary line $x=s$, $y=1-s$,
where $x+y=1$, and \leanid{lieb_concavity_psd} extends that boundary theorem to
positive-semidefinite inputs.  The sub-boundary theorem
\leanid{lieb_concavity_subboundary} removes the exponent restriction: for
$r=x+y=0$ the functional is constant, and for $0<r\le 1$ it rewrites
$A^x,B^y$ as $(A^r)^s,(B^r)^{1-s}$ and combines operator concavity of
$C\mapsto C^r$, monotonicity of the boundary trace functional, and
\leanid{lieb_concavity_psd}.

\paragraph{Positive-definite versus positive semidefinite (resolved).}  Wolf's
Theorem~5.15 states the Ando--Lieb concavity for positive (semidefinite)
operators, including singular ones.  The first formalization
\leanid{lieb_concavity_axiom} was stated only for positive-\emph{definite}
$A,B$ (its hypotheses are \leanid{Matrix.PosDef}).  The positive-semidefinite
case is now also proved, as \leanid{lieb_concavity_psd}, with hypotheses
\leanid{Matrix.PosSemidef} on $A_1,A_2,B_1,B_2$.  The proof is by
regularization: each input is replaced by $A_i+\varepsilon\Id$ (positive
definite for $\varepsilon>0$), the positive-definite theorem is applied for
every $\varepsilon>0$, and the limit $\varepsilon\to 0^+$ recovers the
semidefinite inequality.  The regularized powers converge because $x\mapsto x^s$
is continuous on the nonnegative reals for $0<s<1$, including at the origin ---
the logarithmic singularity that obstructs the entropy-continuity route does not
arise here --- and the trace functional is continuous.  This scope restriction
is therefore eliminated; the exponent restriction is eliminated by
\leanid{lieb_concavity_subboundary}.

\subsection{The point at issue}\label{ssec:lieb-gap}

The matrix identity
$A^s B^{1-s}=\tfrac{\sin(\pi s)}{\pi}\int_0^\infty t^{s-1}A\,(A+tB)^{-1}B\,\mathrm dt$
holds only when $A$ and $B$ commute; for noncommuting $A,B$ the product
$A^sB^{1-s}$ is not even self-adjoint, and this resolvent formula is \emph{not}
the representation behind the Lieb functional.  The standard route works instead
with the \emph{left and right multiplication superoperators} on $\mathcal M_d$,
\begin{align}
  L_A : X \longmapsto AX, \qquad R_B : X \longmapsto XB,
\end{align}
which always commute ($L_A R_B = R_B L_A$) and are positive on the
Hilbert--Schmidt space $(\mathcal M_d,\langle X,Y\rangle = \tr(X^\dagger Y))$
when $A,B\ge 0$.  The Lieb functional is the Hilbert--Schmidt pairing
\begin{align}
  \tr\!\bigl(K^\dagger A^s K B^{1-s}\bigr)
    = \bigl\langle K,\ L_A^{\,s}\,R_B^{\,1-s}\,K\bigr\rangle,
\end{align}
and the Ando/Lieb integral representation is the commuting-superoperator identity
\begin{align}
  L_A^{\,s}\,R_B^{\,1-s}
    = \frac{\sin(\pi s)}{\pi}\int_0^\infty t^{s-1}\,
      L_A\,(L_A + t\,R_B)^{-1}\,R_B\ \mathrm dt .
\end{align}
Joint concavity of the superoperator integrand in $(A,B)$ for each fixed $t$ ---
via operator monotonicity of the resolvent $(L_A+tR_B)^{-1}$ in $A,B$ --- followed
by pairing against $K$ in the Hilbert--Schmidt inner product, integration, and
linearity, yields the result.

Mathlib~4.31 contains neither this superoperator integral representation nor a
matrix-mean / resolvent-monotonicity API of the required form for the commuting
pair $(L_A,R_B)$.  Both were therefore developed within the formalization
itself rather than drawn from the library.

\subsection{Verdict}

The Lieb concavity statement is correct
\cite[Theorem~5.15]{Wolf2012Quantum}.  Its elimination required the Ando
integral representation of the commuting left/right multiplication superoperators
$L_A^{\,s}R_B^{\,1-s}$ (not the matrix product $A^sB^{1-s}$, whose resolvent
formula holds only when $A,B$ commute) together with resolvent monotonicity of
$(L_A+tR_B)^{-1}$.  Both were formalized over a sequence of steps --- the scalar
boundary integral, the operator integral representation on the Kronecker model,
the concavity of the resolvent integrand, and the resulting operator concavity ---
and \leanid{lieb_concavity_axiom} is now a theorem in its positive-definite,
boundary-exponent scope.  The positive-semidefinite case is also proved, as
\leanid{lieb_concavity_psd}, by regularizing each input to positive definite and
passing to the limit, and \leanid{lieb_concavity_subboundary} derives the full
two-exponent range $x,y\ge 0$, $x+y\le 1$.  Three downstream corollaries build on
the positive-definite version (the $K=\mathbf 1$ case and separate concavity in
each argument); these are now axiom-free.  The former positive-semidefinite and
sub-boundary exponent restrictions for the finite-dimensional matrix
Ando--Lieb theorem have both been eliminated.

\section*{References}

Beyond \cite{Wolf2012Quantum}, the underlying matrix-analysis facts are:
E.~H.\ Lieb, \emph{Convex trace functions and the Wigner--Yanase--Dyson
conjecture}, Adv.\ Math.\ \textbf{11} (1973) 267--288;
T.~Ando, \emph{Concavity of certain maps on positive definite matrices and
applications to Hadamard products}, Linear Algebra Appl.\ \textbf{26} (1979)
203--241;
F.~Hansen and G.~K.\ Pedersen, \emph{Jensen's operator inequality}, Bull.\
London Math.\ Soc.\ \textbf{35} (2003) 553--564;
R.~Bhatia, \emph{Matrix Analysis}, Springer GTM~169 (1997), Chapter~V.
The relevant formal-library inputs now available are
\leanid{CFC.exists_measure_nnrpow_eq_integral_cfcₙ_rpowIntegrand₀₁},
\leanid{CFC.concaveOn_rpow}, \leanid{CFC.convexOn_rpow}, and
\leanid{CFC.concaveOn_log}.  Thus the scalar functional-calculus inputs are
available, and all four positive-map Jensen and Lieb declarations of this
boundary are now proved theorems.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
