\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Top-index scope of the spectral n-positivity criterion (Wolf Prop.~3.2)}
\author{}
\date{2026-06-24}

\begin{document}
\maketitle

\section{Source statement}

Proposition~3.2 of Wolf's \emph{Quantum Channels \& Operations}
(\path{Notes/WolfNoteTexSource/ch03_positive_not_completely.tex}, lines
153--179) bounds the expectation of a Hermitian operator
$\tau\in\mathcal{M}_{D'}\otimes\mathcal{M}_{D}$ in vectors of fixed Schmidt
rank. Writing $\tau=\sum_i\nu_i|\phi_i\rangle\langle\phi_i|$ with normalized
eigenvectors, $\rho_i=\tr_2|\phi_i\rangle\langle\phi_i|$ the reduced density on
the first factor $\mathbb{C}^{D'}$, $\nu_0$ the smallest positive eigenvalue and
$\nu$ the largest one, the infimum over normalized vectors $\psi$ of Schmidt
rank $n$ obeys
\begin{align}
  \inf_{\psi}\langle\psi|\tau|\psi\rangle
  &\ge \nu_0 + \sum_{i:\nu_i\le 0}(\nu_i-\nu_0)\,\|\rho_i\|_{(n)},
  \tag{3.7}
\end{align}
and, when a single eigenvalue $\nu_-$ is non-positive,
\begin{align}
  \inf_{\psi}\langle\psi|\tau|\psi\rangle
  &\le \nu + (\nu_- - \nu)\,\|\rho_-\|_{(n)}.
  \tag{3.8}
\end{align}
Each $\rho_i$ acts on the first factor $\mathbb{C}^{D'}$, so its Ky-Fan
$n$-norm $\|\rho_i\|_{(n)}$ is meaningful for $1\le n\le D'$. The Schmidt rank
$n$ ranges over $1\le n\le D'$, so the top value $n=D'$ is within the source's
scope.

\section{Formalized statement}

For Schmidt ranks $1\le n<D'$ the formal theorems\footnote{Formal declarations:
    \leanid{Matrix.IsHermitian.spectral_lower_bound} and
    \leanid{Matrix.IsHermitian.exists_le_spectral_upper_bound} in
    \doclink{TNLean/Channel/NPositivitySpectralCriterion}.}
establish these bounds in the per-vector form: every normalized vector $\psi$ of
Schmidt rank at most $n$ satisfies the lower bound~(3.7), and some such $\psi$
realizes the upper bound~(3.8).

\section{Status}

The top index $n=D'$ is resolved by a route that bypasses the $n<D'$ restriction
of the maximal-overlap and Ky-Fan principles. At $n=D'$ the Schmidt-rank
constraint is vacuous: every normalized vector qualifies, $\|\rho_i\|_{(D')}
=\tr\rho_i=1$, and both bounds collapse to the elementary spectral estimate
$\nu_{\min}\le\langle\psi|\tau|\psi\rangle\le\nu_{\max}$. This estimate is the
Rayleigh characterization of the extreme eigenvalues, which follows directly from
the Rayleigh expansion $\langle\psi|\tau|\psi\rangle=\sum_i\nu_i
|\langle\phi_i|\psi\rangle|^2$ and Parseval's identity $\sum_i
|\langle\phi_i|\psi\rangle|^2=1$ by comparing each $\nu_i$ against $\nu_{\min}$
respectively $\nu_{\max}$; both extremes are attained at the corresponding
eigenvector. The Ky-Fan machinery of the $n<D'$ proof is not invoked.

The top-index bounds are recorded as
\leanid{Matrix.IsHermitian.spectral_lower_bound_top} (with attainment
\leanid{Matrix.IsHermitian.exists_spectral_lower_bound_top}) and
\leanid{Matrix.IsHermitian.spectral_upper_bound_top} (with attainment
\leanid{Matrix.IsHermitian.exists_spectral_upper_bound_top}), giving
$\min_\psi\langle\psi|\tau|\psi\rangle=\nu_{\min}$ and the dual identity for the
maximum, with no Schmidt-rank restriction on $\psi$.

\section{Classification}

\emph{Resolved.} The full source range $1\le n\le D'$ of bounds~(3.7) and~(3.8)
is now formalized: the per-vector bounds
\leanid{Matrix.IsHermitian.spectral_lower_bound} and
\leanid{Matrix.IsHermitian.exists_le_spectral_upper_bound} for $1\le n<D'$, and
the Rayleigh characterization \leanid{Matrix.IsHermitian.spectral_lower_bound_top}
and \leanid{Matrix.IsHermitian.spectral_upper_bound_top} for the
completely-positive endpoint $n=D'$, by the route detailed in the Status above.
No hypothesis foreign to the source is used.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
