\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{The Square-Dimension Restriction in Wolf's Generic Normal Form}
\author{}
\date{2026-06-30}

\begin{document}
\maketitle

\section{The assertion}

Wolf's Proposition~2.9 is a normal-form statement for completely positive maps
with full Kraus rank.\footnote{\cite[Proposition~2.9]{Wolf2012Quantum};
local source \path{Notes/WolfNoteTexSource/ch02_representations.tex},
lines 923--929.}  In the source it reads: let
\begin{align}
  T : \mathcal{M}_{d_1} \to \mathcal{M}_{d_2}
\end{align}
be a completely positive map with full Kraus rank.  Then there exist invertible
completely positive maps \(\Phi_i \in \mathcal B(\mathcal{M}_{d_i})\) of Kraus
rank one such that \(T' := \Phi_2 T \Phi_1\) is doubly-stochastic, i.e.\ both
\(T'(\one)\) and \(T'^{*}(\one)\) are proportional to the identity matrix.

The two filterings act on \emph{different} matrix algebras: \(\Phi_1\) on the
input algebra \(\mathcal{M}_{d_1}\) and \(\Phi_2\) on the output algebra
\(\mathcal{M}_{d_2}\), with no constraint relating \(d_1\) and \(d_2\).

\section{The point at issue}

The formal development states the proposition only in the equal-dimension case
\(d_1 = d_2 = D\):
\begin{align}
  T : \MN{D} \to \MN{D},
\end{align}
with both filterings acting on the same algebra \(\MN{D}\).\footnote{Formal
declaration \leanid{Wolf.exists\_normal\_form\_generic} in
\doclink{TNLean/Channel/LorentzNormalForm}.}  The Choi matrix
\(\tau = (T \otimes \mathrm{id})(\ketbra{\Omega}{\Omega})\) is then a square
matrix on \(\C^{D} \otimes \C^{D}\), and the minimisation runs over
\(\mathrm{SL}(D,\C) \times \mathrm{SL}(D,\C)\) filterings of a single dimension.
The compactness lemma \leanid{Wolf.infimum\_is\_attained} and the
trace--determinant AM--GM equality
\leanid{Wolf.posSemidef\_pow\_det\_eq\_trace\_pow\_iff} that drive the proof are
likewise written for a square Choi matrix on \(\C^{D} \otimes \C^{D}\).

The square case is the one needed downstream: the qubit Lorentz normal form
(\leanid{Wolf.exists\_lorentz\_normal\_form\_qubit}) is the \(D = 2\)
specialisation, and there \(d_1 = d_2 = 2\).  The general rectangular statement
of the source is therefore not yet formalised; the equal-dimension instance is.

\section{The restricted statement}

The formal theorem proves the square instance:
\begin{align}
  T : \MN{D} \to \MN{D}, \qquad
  \tau = \mathrm{choi}(T) \succ 0
  \ \Longrightarrow\
  \exists\, \Phi_1, \Phi_2,\
  \Phi_2 \circ T \circ \Phi_1\ \text{doubly-stochastic},
\end{align}
with both \(\Phi_1, \Phi_2\) SL-filterings on \(\MN{D}\).  This is a sub-case of
the source proposition, obtained by setting \(d_1 = d_2\); it is mathematically
correct as stated, just narrower than Wolf's rectangular formulation.

\section{Elimination plan}

A faithful formalization of the source proposition requires generalising the
two supporting lemmas to independent input and output dimensions:
\begin{itemize}
  \item \leanid{Wolf.infimum\_is\_attained} should range over a Choi matrix on
    \(\C^{d_2} \otimes \C^{d_1}\) with filterings in
    \(\mathrm{SL}(d_2,\C) \times \mathrm{SL}(d_1,\C)\);
  \item \leanid{Wolf.exists\_normal\_form\_generic} should then be restated for
    \(T : \mathcal{M}_{d_1} \to \mathcal{M}_{d_2}\) with separate filterings on
    each factor.
\end{itemize}
Until that work is done, the square version is kept as an explicitly-labelled
specialisation, and the source-faithful rectangular proposition remains
unformalised.  The blueprint entry \texttt{thm:exists\_normal\_form\_generic}
and the Lean docstring carry the matching scope marker.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
