\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Top-index scope of the maximal-overlap lemma (Wolf Lemma 3.1)}
\author{}
\date{2026-06-24}

\begin{document}
\maketitle

\section{Source statement}

Lemma~3.1 of Wolf's \emph{Quantum Channels \& Operations}
(\path{Notes/WolfNoteTexSource/ch03_positive_not_completely.tex}, lines
119--128) states that, for a normalized vector
$\phi\in\mathbb{C}^{D'}\otimes\mathbb{C}^{D}$ with reduced density matrix
$\rho=\tr_2|\phi\rangle\langle\phi|$ on the first factor $\mathbb{C}^{D'}$,
\begin{align}
  \sup_{\psi}|\langle\phi|\psi\rangle|^2 = \|\rho\|_{(n)},
\end{align}
where the supremum runs over normalized vectors $\psi$ of Schmidt rank $n$ and
$\|\cdot\|_{(n)}$ is the Ky-Fan $n$-norm. Since $\rho$ acts on the first factor
$\mathbb{C}^{D'}$, its Ky-Fan $n$-norm is meaningful for $1\le n\le D'$. In the
application (n-positivity, Wolf Prop.~3.2) the Schmidt rank $n$ ranges over
$1\le n\le D'$, so the top value $n=D'$ is within the source's scope.

\section{Formalized statement}

The formal theorem\footnote{Formal declaration:
    \leanid{Matrix.maximalSchmidtOverlap_eq_kyFanNorm} in
    \doclink{TNLean/Channel/MaximalOverlap}.} establishes the same identity,
strengthened to attainment (the \Lean{} \texttt{IsGreatest} predicate), under
the hypotheses $1\le n$ and $n<D'$, where $D'$ is the dimension of the first
tensor factor on which $\rho$ acts.

\section{The gap}

The formalization omits the top index $n=D'$. At $n=D'$ the Schmidt-rank
constraint is vacuous and $\|\rho\|_{(D')}=\tr\rho=1$; the supremum equals $1$
and is attained at $\psi=\phi$.

The restriction is inherited from the underlying maximum principle,\footnote{Formal
    declaration: \leanid{Matrix.IsHermitian.kyFanNorm_eq_sup_trace}.} which
characterizes the Ky-Fan $k$-norm as a maximum over rank-exactly-$k$ orthogonal
projections only for $k<D'$. At $k=D'$ the only rank-$D'$ projection is the
identity, and the eigenvalue sum already equals the trace; the variational
statement degenerates rather than fails.

\section{Elimination plan}

Add a top-index case to the Ky-Fan maximum principle (or a companion lemma)
covering $k=D'$, where the maximizing projection is the identity and the value
is $\operatorname{Re}\tr(A)$. The overlap lemma then extends to $1\le n\le D'$
by feeding that case into the same transfer argument.

\section{Classification}

\emph{Resolved.} The top index $n=D'$ is now formalized as the companion theorem
\leanid{Matrix.maximalSchmidtOverlap_eq_kyFanNorm_top}: there the Schmidt-rank
constraint is vacuous, the Ky-Fan $D'$-norm collapses to the trace
$\|\rho\|_{(D')}=\tr\rho=1$, and the value is attained at $\psi=\phi$ with the
upper bound supplied by Cauchy--Schwarz. Together with
\leanid{Matrix.maximalSchmidtOverlap_eq_kyFanNorm} (the range $1\le n<D'$) this
covers the full source range $1\le n\le D'$ of Wolf's Lemma~3.1, exactly as the
elimination plan above anticipated.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
