\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{./command}
\providecommand{\leanid}[1]{\path{#1}}

\title{The nonempty-family condition in Wolf's Radon--Nikodym theorem}
\author{}
\date{2026-06-16}

\begin{document}
\maketitle

\section{Source statement}

Wolf's Radon--Nikodym theorem for quantum instruments states that, if
\(\{T_i\}\) is a family of completely positive maps satisfying
\(\sum_i T_i=T\) and
\[
  T(A)=V^\dagger(A\otimes \mathbf 1_r)V,
\]
then there are positive operators \(P_i\in\mathcal M_r\) such that
\[
  \sum_i P_i=\mathbf 1_r,
  \qquad
  T_i(A)=V^\dagger(A\otimes P_i)V.
\]
The local source is
\path{Notes/WolfNoteTexSource/ch02_representations.tex}:402--407, titled
``Radon--Nikodym for quantum instruments'' in
\cite{Wolf2012Quantum}.

\section{Local boundary}

The Lean theorem \leanid{IsCPMap.radon_nikodym_of_stinespring} makes the index
type nonempty.  This is not a strengthening of a mathematical hypothesis in the
substantive case; it is the finite-type boundary needed for the identity
\[
  \sum_i P_i=\mathbf 1_r
\]
to be meaningful as a resolution of the identity by instrument components.  If
the index type were empty, then the hypotheses \(\sum_i T_i=T\) force
\(T=0\).  One may still supply the zero Stinespring matrix
\(V=0\), but the conclusion would ask
\[
  0=\mathbf 1_r
\]
in \(\mathcal M_r\), which is false for a nonzero dilation dimension.

Thus the formal theorem states the source theorem in the nonempty instrument
case.  The blueprint entry records the same boundary explicitly by saying
``nonempty finite family.''

\section{Elimination plan}

No mathematical proof gap is expected for the nonempty case.  An empty-index
variant would need a different conclusion, for example a vacuous statement with
zero dilation space or no identity resolution.  Such a statement is not the
instrument theorem used in the development.

Verdict: local correction of a well-posedness boundary in the printed finite-family
statement.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
