\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Top Schmidt-rank index in the $n$-positivity threshold of $T_\eta$
(Wolf Chapter 3)}
\author{}
\date{2026-06-24}

\begin{document}
\maketitle

\section{Source statement}

In Chapter~3 of Wolf's lecture notes \cite{Wolf2012Quantum}, the one-parameter
family
\begin{align}
  T_\eta(\rho)=\tr(\rho)\,\mathbf 1-\frac{\rho}{\eta},
  \qquad \eta\in\mathbb R_+,
  \tag{3.11}
\end{align}
on $M_D(\mathbb C)$ (the formal map is written $\tr(\rho)\,\mathbf 1-\eta^{-1}\rho$,
which agrees on $\eta>0$ and is moreover defined at $\eta=0$) is analyzed
through its Choi--Jamiolkowski operator
$\tau_\eta=D^{-1}\mathbf 1-\eta^{-1}\ket\Omega\bra\Omega$.  All positive
eigenvalues of $\tau_\eta$ equal $1/D$, and the single non-positive eigenvalue
is $\nu_-=1/D-1/\eta$ with reduced density $\rho_-=\mathbf 1/D$, so its
Ky--Fan $n$-norm is $\|\rho_-\|_{(n)}=n/D$.  The source concludes:
\begin{align}
  T_\eta\ \text{is $n$-positive}\iff \eta\ge n,
  \qquad n=1,\dots,D,
\end{align}
where the range of $n$ runs up to the dimension $D$, in which case
$n$-positivity is complete positivity.  This proves that every inclusion in
Wolf's chain
\begin{align}
  \mathcal T_{\mathrm{cp}}=\mathcal T_D\subseteq
  \mathcal T_{D-1}\subseteq\cdots\subseteq\mathcal T_2\subseteq\mathcal T_1
  \tag{3.3}
\end{align}
is strict.\footnote{Local source:
\path{Notes/WolfNoteTexSource/ch03_positive_not_completely.tex:181--193},
equation label \path{eq:ch3:3.11}.}

\section{Point at issue}

The formal statement establishes the equivalence
\begin{align}
  T_\eta\ \text{is $k$-positive}\iff \eta\ge k
\end{align}
for Schmidt-rank index $1\le k<D$, omitting the top index $k=D$.  The
restriction is inherited, not introduced here: the equivalence is derived from
the maximal-overlap principle (Wolf Lemma~3.1), whose formal version
$\sup_\psi|\braket{\Omega}{\psi}|^2=\|\rho\|_{(k)}$ over Schmidt-rank-$k$
normalized vectors is available only for $1\le k<D$.  That bound in turn rests
on the Ky--Fan maximum principle, stated for orthogonal projections of rank
exactly $k$ with $k<D$.  At the top index $k=D$ the Ky--Fan $D$-norm is the
full trace and the maximizing projection is the identity; the present
formalization of the maximum principle does not cover that boundary case.

Consequently the strictness witness produced here, a $k$-positive map that fails
to be $(k+1)$-positive, is available for $1\le k$ with $k+1<D$.  This covers
every inclusion in the chain~(3.3) except the topmost one,
$\mathcal T_D\subseteq\mathcal T_{D-1}$, which compares complete positivity with
$(D-1)$-positivity and requires the $k=D$ index.

\section{Status}

The boundary case $k=D$ is resolved by a route that bypasses the $k<D$
restriction of the maximal-overlap and Ky--Fan principles entirely.  On
$M_D(\mathbb C)$, $D$-positivity is complete positivity
(\leanid{isCPMap_iff_isNPositiveMap_card}), which is positive semidefiniteness of
the Choi operator $\tau_\eta$.  Its quadratic form on a vector $\psi$ is
$D^{-1}\|\psi\|^2-\eta^{-1}|\braket{\Omega}{\psi}|^2$, and since $\ket\Omega$ is
normalized the Cauchy--Schwarz bound $|\braket{\Omega}{\psi}|^2\le\|\psi\|^2$
(saturated at $\psi=\Omega$) makes that form nonnegative for every $\psi$ exactly
when $\eta\ge D$.  Hence $T_\eta$ is $D$-positive iff $\eta\ge D$, the equivalence
at the top index.

The formal declarations are \leanid{Matrix.isNPositiveMap_tEta_iff} for the
range $1\le k<D$ and \leanid{Matrix.isNPositiveMap_tEta_card_iff} for the top
index $k=D$, together covering the full equivalence~(3.11) over $1\le k\le D$, in
\doclink{TNLean/Channel/NPositivityChainStrict}.  The strictness witness
\leanid{Matrix.exists_isNPositiveMap_not_succ} remains stated for $k+1<D$; the
strictness of the top inclusion $\mathcal T_D\subsetneq\mathcal T_{D-1}$ follows
from the $k=D-1$ and $k=D$ thresholds but is not recorded as a separate witness.

\section{Classification}

\emph{Resolved.} The full source range $1\le k\le D$ of equivalence~(3.11) is now
formalized: \leanid{Matrix.isNPositiveMap_tEta_iff} for $1\le k<D$ and
\leanid{Matrix.isNPositiveMap_tEta_card_iff} for the completely-positive endpoint
$k=D$, the latter by the Choi-operator route detailed in the Status above.  No
hypothesis foreign to the source is used.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
