\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}

\input{command}

\title{Schmidt-number premise of Wolf's reduction criterion (Wolf eq.~(3.18))}
\author{}
\date{2026-06-24}

\begin{document}
\maketitle

\section{Source statement}

The reduction-criterion paragraph of Wolf's \emph{Quantum Channels \&
Operations} (\path{Notes/WolfNoteTexSource/ch03_positive_not_completely.tex},
lines 329--338) introduces the maps $T_n(X)=\tr(X)\,\mathbf 1-X/n$, notes that
each is $n$-positive and self-dual, and concludes that
\begin{align}
  n\,\rho_1\otimes\mathbf 1\ge\rho
  \qquad\text{and}\qquad
  n\,\mathbf 1\otimes\rho_2\ge\rho
  \tag{3.18}
\end{align}
is \emph{necessary for a bipartite state $\rho$ to have Schmidt number $n$},
with $\rho_1$ and $\rho_2$ the two reduced density matrices. The source assertion
is the implication
\begin{align}
  \operatorname{sn}(\rho)\le n
  \;\Longrightarrow\;
  \bigl(n\,\rho_1\otimes\mathbf 1\ge\rho\ \text{and}\
  n\,\mathbf 1\otimes\rho_2\ge\rho\bigr),
  \tag{\path{source}}
\end{align}
where $\operatorname{sn}(\rho)$ is the Schmidt number of $\rho$. Its proof has
two steps. A state of Schmidt number at most $n$ is a convex combination of
pure states of Schmidt rank at most $n$, so $n$-positivity of $T_n$ gives
\begin{align}
  \operatorname{sn}(\rho)\le n
  \;\Longrightarrow\;
  (T_n\otimes\operatorname{id})(\rho)\ge 0,
  \tag{step 1}
\end{align}
and the algebraic identity $(T_n\otimes\operatorname{id})(\rho)=\mathbf 1\otimes\rho_2-n^{-1}\rho$
then yields
\begin{align}
  (T_n\otimes\operatorname{id})(\rho)\ge 0
  \;\Longrightarrow\;
  n\,\mathbf 1\otimes\rho_2\ge\rho,
  \tag{step 2}
\end{align}
with the companion inequality obtained from the swapped state.

\section{Formalized statement}

The formal theorems\footnote{Formal declarations:
    \leanid{Matrix.reductionCriterion_left} and
    \leanid{Matrix.reductionCriterion_right} in
    \doclink{TNLean/Channel/ReductionCriterion}.} establish step~2 only. For a
bipartite matrix $\rho$ and $n\ge 1$,
\begin{align}
  (T_n\otimes\operatorname{id})(\rho)\ge 0
  &\;\Longrightarrow\;
  n\,\mathbf 1\otimes\rho_2\ge\rho,\\
  (T_n\otimes\operatorname{id})(\rho^{F})\ge 0
  &\;\Longrightarrow\;
  n\,\rho_1\otimes\mathbf 1\ge\rho,
\end{align}
where $\rho^{F}$ is the image of $\rho$ under the factor swap and the second
hypothesis equals Wolf's symmetric condition $(\operatorname{id}\otimes T_n)(\rho)\ge 0$. The
underlying identity $(T_n\otimes\operatorname{id})(\rho)=\mathbf 1\otimes\rho_2-n^{-1}\rho$ is
also formalized.\footnote{Formal declaration:
    \leanid{Matrix.tensorMapId_tEta_eq}.}

\section{The gap}

The formalized statement carries the hypothesis $(T_n\otimes\operatorname{id})(\rho)\ge 0$,
which is the intermediate produced by step~1 rather than the source premise
$\operatorname{sn}(\rho)\le n$. Step~1 is not yet formalized because the
repository has no Schmidt-number predicate for mixed states, equivalently no
separable-state predicate from which the Schmidt-number filtration is built. The
same missing foundation is the obstruction to the separability extension of the
reduction criterion as an entanglement criterion.

The $n$-positivity content needed for step~1 is itself available: the map $T_n$
is $n$-positive for all $1\le n<D$.\footnote{Formal declaration:
    \leanid{Matrix.isNPositiveMap_tEta_iff} in
    \doclink{TNLean/Channel/NPositivityChainStrict}, the reduction map being
    $T_\eta$ at $\eta=n$ by \leanid{Matrix.reductionMap_eq_tEta}.} What is
absent is the predicate $\operatorname{sn}(\rho)\le n$ and the lemma that it
implies $\rho$ lies in the cone on which the $n$-positive ampliation acts
positively.

\section{Elimination plan}

Introduce a Schmidt-number predicate for bipartite matrices, built from a
separable-state predicate, and prove step~1: a state of Schmidt number at most
$n$ has $(T_n\otimes\operatorname{id})(\rho)\ge 0$, by writing it as a convex combination of
Schmidt-rank-$\le n$ pure states and applying the $n$-positivity already
formalized. Composing step~1 with the formalized step~2 recovers Wolf's
eq.~(3.18) with its Schmidt-number premise. The separability foundation is
shared with the entanglement-criterion extension and is tracked alongside the
PPT and Schmidt-number formalization tasks.

\section{Classification}

\emph{Scope restriction.} The formal theorems prove the second step of Wolf's
two-step argument: the operator implication from $(T_n\otimes\operatorname{id})(\rho)\ge 0$ to
the inequalities of eq.~(3.18). The conclusions match the source exactly; the
hypothesis is the derived intermediate of step~1, so the formalized statement is
strictly weaker than, and is a sub-step of, the source implication. No
hypothesis foreign to the source is introduced, so the deviation is a scope
restriction rather than an unfaithful theorem.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
